### 2013-08-02 exact_learning/paper.tex
init

\section{Introduction}
Many classical computer vision applications such as stereo, optical flow, semantic
segmentation and visual grouping can be naturally formulated as image labeling tasks.

Arguably the most popular way to approach such labeling problems is via graphical
models, such as Markov random fields (MRFs) and conditional random fields (CRFs).
MRFs and CRFs provide a principled way of integrating local evidence and
modeling spacial dependencies, which are strong in most image-based tasks.

While in earlier approaches, model parameters were set by hand or using
cross-validation, more recently parameters are often learned using a max-margin
approach.

Most models employ linear energy functions of unary and pairwise interactions,
trained using structural support vector machines (SSVMs). While linear energy
functions lead to learning problems that are convex in the parameters, complex
constraints complicate their optimization. Additionally, inference (or more precisely
loss-augmented prediction) is a crucial part in learning, and can often not be
performed exactly, due to loops in the neighborhood graphs.  % TODO clean up!
Approximations in the inference then lead to approximate learning.

We look at semantic image segmentation, learning a model of pairwise
interactions on the popular MSRC-21 dataset.
The contribution of this work is threefold:
\begin{itemize}
    \item We analyze the simultaneous use of multiple approximate inference
        methods for learning SSVMs using the cutting plane method, relating
        approximate learning to the exact optimum.
    \item We introduce an efficient caching scheme to accelerate cutting plane
        training.
    \item We demonstrate that using a combination of under-generating and exact
        inference methods, we can learn an SSVM exactly in a practical
        application, even in the presence of loopy graphs.
\end{itemize}

While empirically exact learning yields results comparable to those using
approximate inference alone, certification of optimality allows treating
learning as a black-box, allowing the researcher to focus attention on
designing the model for the application at hand.

\section{Related Work}

Max margin learning for structured prediction was introduced by
\citet{taskar2003max}, in the form of maximum-margin Markov models. Later, this
framework was generalized to structural support vector machines by
\citet{tsochantaridis2006large}. Both works assume tractable loss-augmented
inference.

Currently the most widely used method is the one-slack cutting plane formulation
introduced by \citet{joachims2009cutting}.
This work also introduced the caching of constraints,
which serves as a basis for our work. We improve upon their caching scheme, and in
particular consider how it interacts with approximate inference algorithms.

% TODO learning for inference?
Recently, there has been much research in learning structured prediction models
where standard exact inference techniques are not applicable.
The influence of approximate inference on structural support vector machine
learning was first analyzed by  \citet{finley2008training}.
\citet{finley2008training} show convergence results for under-generating and
over-generating inference procedures, meaning methods that find suboptimal, but
feasible solutions, and optimal solutions from a larger (unfeasible) set,
respectively.
\citet{finley2008training} demonstrate that over-generating approaches---in
particular linear programming (LP)---perform best on the considered model.
They also give a bound on the empirical risk for this case.
In contrast, we aim at optimizing the non-relaxed objective directly, yielding
the original, tighter bound.

As using LP relaxations was considered too costly for typical computer vision
approaches, later work employed graph-cuts~\citep{szummer2008learning} or Loopy
Believe Propagation (LBP)~\citep{lucchi2011spatial}. These works use a single
inference algorithm for during the whole learning process, and can not provide any
bounds on the true objective or the empirical risk. In contrast, we combine
different inference methods that are more appropriate for different stages of
learning.

Recently, \citet{meshi2010learning}, \citet{hazan2010primal} and
\citet{komodakis2011efficient} introduced formulations for joint inference and
learning using duality.
In particular, \citet{hazan2010primal} demonstrate the performance of their
model on an image denoising task, where it is possible to learn a large number
of parameters efficiently.
While these approaches show great promise, in particular for pixel-level or
large-scale problems, they perform approximate inference and learning, and do
not relate their results back to the original SSVM objective they approximate.

\begin{algorithm*}[t]
    \caption{Cutting Plane Training of Structural SVMs \label{alg_cutting_plane}}

\begin{algorithmic}[1]
    \Require Training samples $\{ (x^{1}, y^{1}), \dots, (x^{n}, y^{n})\}$, regularization parameter $C$, stopping tolerance $\epsilon$, separation oracle $I$.
    \Ensure Parameters $\theta$, slack $\xi$
    \State $\W \leftarrow \emptyset$
    \Repeat
    \State $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi$\label{restricted}\\
    \begin{equation*}
        \text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \W:
        \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i,
            \hat{y}^i)] \right \rangle \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
    \end{equation*}
    \For {i=1, \dots, n}
    \State
    $\hat{y}^i \leftarrow I(x^i, y^i, \theta) \approx \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y})] \right \rangle$ \label{get_cutting_plane}
    \EndFor
    \State $\W \leftarrow \W \cup \{ (\hat{y}^1, \dots, \hat{y}^n) \} $
    \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \Delta(y^i, \hat{y}^i) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y}^i)] \right \rangle$
    \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
\end{algorithmic}
\end{algorithm*}

\section{Efficient Cutting Plane Training of SSVMs}\label{learning}

\subsection{The Cutting Plane Method}\label{cutting_plane}

When learning for structured prediction in the max-margin framework of
\citet{tsochantaridis2006large}, predictions are made as
\[ \argmax_{y \in \mathcal{Y}} f(x, y, \theta),\]
where $x \in \mathcal{X}$ is the input, $y \in \mathcal{Y}$ the prediction, and $\theta$ are the parameters to be learned.
We will assume $y$ to be multivariate, $y=(y_1, \dots, y_k)$ with possibly varying $k$.

The function $f$ measures compatibility of $x$ and $y$ and is a linear function of the parameters $\theta$:
\[f(x, y, \theta) = \langle \theta, \psi(x, y) \rangle.\]
Here $\psi(x, y)$ is a joint feature vector of $x$ and $y$. Specifying a particular SSVM model
amounts to specifying $\psi$.

For a given loss $\Delta$, the parameters $\theta$ are learned by minimizing
the loss-based soft-margin objective
\begin{equation}\label{mainequation}
\min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  r_i(\theta)
\end{equation}
with regularization parameter $C$, where $r_i$ is a hinge-loss-like upper
bound on the empirical $\Delta$-risk:
\[r_i(x^i, y^i, y) = \left [\max_{y \in \mathcal{Y}} \Delta(y^i, y) + \left<\theta, \psi(x^i, y) - \psi(x^i, y^i) \right > \right]_+ \]

We solve the following reformulation of Equation~\ref{mainequation}, known as one-slack QP formulation:
\begin{align}\label{oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
        &\left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i,
            \hat{y}^i)] \right \rangle \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
\end{align}
using the cutting plane method described in
Algorithm~\ref{alg_cutting_plane}~\citep{joachims2009cutting}.

The cutting plane method alternates between solving Equation~\eqref{oneslack}
with a working set $\W$ of constraints, and expanding the working set using the
current $\theta$ by finding $\mathbf{y}$ corresponding to the most violated constraint,
using a separation oracle $I$.
We investigate the construction of $\W$ and the influence of $I$
on learning.

%TODO unstructured!!

Intuitively, the one-slack formulation corresponds to joining all training
samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
interactions between variables corresponding to different data points.
Consequently, only a single constraint is added in each iteration of
Algorithm~\ref{alg_cutting_plane}, leading to very small $\W$. We further use
pruning of unused constraints, as suggested by \citet{joachims2009cutting},
resulting in the size of $\W$ being in the order of hundreds for all experiments.

We also use another enhancement of the cutting plane algorithm introduced by
\citet{joachims2009cutting}, the caching oracle. For each training example $(x^i, y^i)$,
we maintain a set $C^i$ of the last $r$ results of loss-augmented inference
(line~\ref{get_cutting_plane} in Algorithm~\ref{alg_cutting_plane}).
For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we find
\[ 
    \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y})] \right \rangle
\]
by enumeration of $C^i$ and continue until line~\ref{convergence_check}.
Only if $\xi' - \xi < \epsilon$, i.e.\ the produced constraint is not violated, we
return to line~\ref{get_cutting_plane} and actually invoke the separation
oracle $I$.

In computer vision applications, or more generally in graph labeling problems,
$\psi$ is often given as a factor graph over $y$, typically using only unary and pairwise functions:
\[ \langle \theta, \psi(x, y) \rangle = \sum_{i=0}^k \langle \theta_u,  \psi_u(x, y_k) \rangle + \sum_{(i, j) \in E} \langle \theta_p, \psi_p(x, y_k, y_l) \rangle, \]
where $E$ a set of pairwise relations. In this form, parameters $\theta_u$ and $\theta_p$ for unary and
pairwise terms are shared across all entries and pairs.
The decomposition of $\psi$ into only unary and pairwise interactions allows
for efficient inference schemes, based on message passing or graph cuts.

There are two groups of inference procedures, as identified in
\citet{finley2008training}: under-generating and over-generating approaches.
An under-generating approach satisfies $I(x^i, y^i, \theta) \in
\mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane}
of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
hand, will solve the loss-augmented inference in line~\ref{get_cutting_plane}
exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.  We will elaborate on
the properties of under-generating and over-generating inference procedures in
Section~\ref{bounds}.

\subsection{Bounding the Objective}\label{bounds}
Even using approximate inference procedures, several statements
about the original exact objective (Equation~\ref{mainequation}) can be
obtained.

Let $o_{\W}(\theta)$ denote objective \eqref{oneslack} with
given parameters $\theta$ restricted to a working set $\W$, as computed in
line~\ref{restricted} of Algorithm~\ref{alg_cutting_plane} and  let
\[
    o^I(\theta) = C\xi' + \frac{||\theta||}{2}^2
\]
when using inference algorithm $I$, i.e. $o^I(\theta)$ is the approximation of the primal
objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.

Depending on the properties of the inference procedure $I$ used, it is easy to see:
\begin{enumerate}
    \item If all constraints $\hat{y}$ in  $\W$ are feasible, i.e.\ generated
        by an under-generating or exact inference mechanism, then $o_{\W}$ is
        an lower bound on the true optimum $o(\theta^*)$.

    \item If $I$ is an over-generating or exact algorithm, $o^I$ is an upper
        bound on $o(\theta^*)$.
\end{enumerate}

These two observations can be used in a number of ways.
\citet{finley2008training} used 2. to show that learning with an
over-generating approach minimizes an upper bound on the empirical risk.

We can also use these observations to judge the suboptimality of a given parameter $\theta$,
i.e.\ see how far the current objective is from the true optimum.
Learning with any under-generating approach, we can use 1. to maintain a lower bound
on the objective. At any point during learning, in particular if no more constraints
can be found, we can then use 2., to also find an upper bound.
This way, we can empirically bound the estimation error, using only approximate
inference.
We will now describe how we can further use 1. to both speed up and improve learning.

\subsection{Combining Inference Procedures}\label{combining}
The cutting plane method described in Section~\ref{cutting_plane} relies only
on some separation oracle $I$ that produces violated constraints.

Using any under-generating oracle $I$, learning can proceed as long as a
constraint is found that is violated by more than the stopping tolerance
$\epsilon$.  Which constraint is used next has an impact on the speed of
convergence, but not on correctness. Therefore, as long as an under-generating
method does generate constraints, optimization makes progress on the objective.

Instead of choosing a single oracle, we propose to use a succession of
algorithms, moving from fast methods to more exact methods as training
proceeds. This strategy not only accelerates training, it even makes it
possible to train with exact inference methods, which is infeasible otherwise.

In particular, we employ three strategies for producing constraints,
moving from one to the next if no more constraints can be found:
\begin{enumerate*}
    \item Produce a constraint using previous, cached inference results.
    \item Use a fast under-generating algorithm.
    \item Use exact inference.
\end{enumerate*}

While using more different oracles is certainly possible, we found
that using just these three methods performed very well in practice.  This
combination allows us to make fast progress initially and guarantee optimality
in the end.
Notably, it is not necessary for an algorithm used as 3) to always produce
exact results. For guaranteeing optimality of the model, it is sufficient that
we obtain a certificate of optimality when learning stops.

\subsection{Dynamic Constraint Selection}\label{caching}
Combining inference algorithm as described in Section~\ref{combining}
accelerates calls to the separation oracle by using faster, less accurate
methods. On the down-side, this can lead to the inclusion of many constraints
that make little progress in the overall optimization, resulting in much more
iterations of the cutting plane algorithm. We found this particularly problematic
with constraints produced by the cached oracle.

We can overcome this problem by defining a more elaborate schedule to switch
between oracles, instead of switching only if no violated constraint can be
found any more. Our proposed schedule is based on the intuition that we only
trust a separation oracle as long as the current primal objective did not move
far from the primal objective as computed with the stronger inference
procedure.

%More precisely, let $o_{\W}(\theta)$ denote objective \eqref{oneslack} with
%given parameters $\theta$ restricted to a working set $\W$, as computed in
%line~\ref{restricted} of Algorithm~\ref{alg_cutting_plane} and  let
%\[
    %o^I(\theta) = C\xi' + \frac{||\theta||}{2}^2
    %\]
%when using inference algorithm $I$, i.e. $o^I(\theta)$ is the approximation of the primal
%objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.
In the following, we will use the notation of Section~\ref{bounds} and indicate
the choices of oracle $I$ with $Q$ for a chosen inference algorithm and $C$ for
using cached constraints.

To determine whether to produce inference results from the cache or to run the inference algorithm,
we solve the QP once with a constraint from the cache. If the resulting $o^C$ verifies
\begin{equation}\label{cache_test}
    o^C - o^Q < \frac{1}{2} (o^Q - o_{\W})
\end{equation}
we continue using the caching oracle. Otherwise we run the inference algorithm again.
For testing \eqref{cache_test}, the last known value of $o^Q$ is used, as recomputing it would defy
the purpose of the cache.

It is easy to see that our heuristic runs inference only $O(\log(o^Q -
o_{\W}))$ times more often than the strategy from \citet{joachims2009cutting} in the
worst case.

\section{Experiments}
\subsection{Inference Algorithms}
As a fast under-generating inference algorithm, we used $\alpha$-expansion
moves based on non-submodular graph-cuts using Quadratid Pseudo-Boolean
Optimization (QPBO)~\citep{rother2007optimizing}.  This move-making strategy
can be seen as a simple instantiation of the more general framework of fusion
moves, as introduced by \citet{lempitsky2010fusion}

For exact inference, we use the recently developed Alternating Direction Dual
Decomposition (AD$^3$) method of \citet{martins2011augmented}. AD$^3$ produces
a solution to the linear programming relaxation, which we use as the basis for
branch-and-bound.

\begin{table}
    \begin{center}
    \begin{tabular}{lcc}
        \toprule
                    & Average & Global \\
        \cmidrule{1-3}
    Unary terms only & 77.7& 83.2 \\
    Pairwise model (move making)& 79.6&84.6\\
    Pairwise model (exact)& 79.0 & 84.3\\
        \cmidrule{1-3}
    \citet{ladicky2009associative} & 75.8& 85.0\\
    \citet{gonfaus2010harmony} & 77&  75\\
    \citet{lucchi2013learning} & 78.9& 83.7\\
    \bottomrule
    \end{tabular}
    \end{center}
    \caption{Accuracies on the MSRC-21 Dataset.  We compare a baseline model,
    our exact and approximately learned models and state-of-the-art
    approaches.\label{msrcacc}}

\end{table}

%\subsection{Multi-Label Classification}
%TODO

%\subsection{Snake}
%TODO

\subsection{Image Segmentation}
We choose superpixel-based semantic image segmentation as sample application
for this work.  CRF based models have a history of success in this application,
with much current work investigating models and
learning~\citep{gonfaus2010harmony, lucchi2013learning, ladicky2009associative,
kohli2009robust, krahenbuhl2012efficient}.  We use the MSRC-21 and Pascal VOC 20120 datasets, two
widely used benchmark in the field.

We use the same model and pairwise features for the two datasets.
Each image is represented as a neighborhood graph of superpixels.
For each image, we extract approximately 100 superpixels using 
the SLIC algorithm~\citep{achanta2012slic}.

We introduce pairwise interactions between neighboring superpixels, as is
standard in the literature. Pairwise potentials are founded on two
image-based features: color contrast between superpixels, and relative location
(coded as angle), in addition to a bias term.

We set the stopping criterion $\epsilon=10^{-4}$, though using only
the under-generating method, training always stopped prematurely as no violated
constraints could be found any more.

\subsection{Caching}
First, we compare our caching scheme, as described in Section~\ref{combining}, with the
scheme of \citet{joachims2009cutting}, which produces constrains from the cache
as long as possible, and with not using caching of constraints at all.  For this experiment,
we only use the under-generating move-making inference on the MSRC-21 dataset. Times until convergence
are 397s for our heuristic, 1453s for the heuristic of
\citet{joachims2009cutting}, and 2661s for using no cache, with all strategies
reaching essentially the same objective.

Figure~\ref{caching} shows a visual comparison that highlights the differences
between the methods. Note that neither $o^Q$ nor $o^C$ provide valid upper bounds on the objective,
which is particularly visible for $o^C$ using the method of \cite{joachims2009cutting}.
Using no cache leads to a relatively smooth, but slow convergence, as inference is run often.
Using the method of \citet{joachims2009cutting}, each run of the separation oracle is followed by
quick progress of the dual objective $o_\W$, which flattens out quickly. Much time is then spent adding
constraints that do not improve the dual solution.
Our heuristic instead probes the cache, and only proceeds using cached constraints if the resulting
$o^C$ is not to far from $o^Q$.

\begin{figure*}
\centering
\includegraphics[width=.7\linewidth]{caching}
\caption{%
Training time comparison using different caching heuristics.
Large dots correspond to $o^Q$, small dots correspond to $o^C$,
and the line shows $o_\W$. See the text for details.\label{caching}}
\end{figure*}
% much simpler model?

\subsubsection{MSRC-21 Dataset}
For the MSRC-21 Dataset, we use unary potentials based on bag-of-words of SIFT
features and color features.  Following \citet{lucchi2011spatial} and
\citet{fulkerson2009class}, we augment the description of each superpixel by a
bag-of-word descriptor of the whole image. To obtain the unary potentials for
our CRF model, we train a linear SVM using the additive $\chi^2$ transform
introduced by \citet{vedaldi2010efficient}. Additionally, we use the unary
potentials provided by \citet{krahenbuhl2012efficient}, which are based on
TextonBoost~\citep{shotton2006textonboost}. This leads to $42 = 2 \cdot 21$
unary features for each node.

The resulting model has around 100 output variables per image, each taking one of 21
labels. The model is trained on 335 images from the standard training and
validation split.

\begin{table}
    \begin{center}
    \begin{tabular}{lc}
        \toprule
                    & Jaccard \\
        \cmidrule{1-2}
    Unary terms only &  27.5 \\
    Pairwise model (move making)& 30.2\\
    Pairwise model (exact) & 30.4\\
        \cmidrule{1-2}
    \citet{dann2012pottics} & 27.4\\
    \citet{krahenbuhl2012efficient} & 30.2\\
    \citet{krahenbuhlparameter} & 30.8\\
    \bottomrule
    \end{tabular}
    \end{center}
    \caption{Accuracies on the Pascal VOC Dataset. We compare our approach
    against approaches using the same unary potentials.\label{pascalacc}}

\end{table}

\subsubsection{Pascal VOC 2010}
For the Pascal VOC 2010 dataset, we follow the procedure from \citet{krahenbuhl2012efficient}
in using the official ``validation'' set as our evaluation set, and splitting the training set again.
We use the unary potentials provided by the same work, and compare only against methods
using the same setup and potentials, \citet{krahenbuhlparameter} and \citet{dann2012pottics}.
Note that state-of-the-art approaches, some not build on the CRF framework, obtain
around 40\% Jaccard score (also call VOC score), notably \cite{xia2012segmentation}, who 
evaluate on the Pascal VOC 2010 ``test'' set.

\begin{table}
    \begin{center}
    \begin{tabular}{lcc}
    \toprule
                    & Move-making & Exact \\
    \cmidrule{1-3}
    Dual Objective $o_\W$ & 65.10 & 67.66  \\
    Estimated Objective $o^I$ &  67.62& 67.66\\
    True Primal Objective $o^E$& 69.92& 67.66\\
    \bottomrule
    \end{tabular}
    \end{center}
    \caption{Objective function values on the MSRC-21 Dataset}
    \label{msrc_objective}
\end{table}

\subsubsection{Results}
We compare classification results using different inference schemes in
with results from the literature. As a sanity check, we
also provide results without pairwise interactions.

Results on the MSRC-21 dataset are shown in Table~\ref{msrcacc}.
We find that our model is on par with state-of-the-art approaches, implying
that our model is realistic for this task. In particular, our results are comparable to those of
\citet{lucchi2013learning}, who use a stochastic subgradient method with working sets.
Their best model takes 583s for training, while training our model exactly takes 1814s.
We find it remarkable that it is possible to guarantee optimality in time of
the same order of magnitude that a stochastic subgradient procedure with
approximate inference takes. Using exact learning and inference does not increase accuracy
on this dataset.
Learning the structured prediction model using move-making inference alone
takes 4 minutes, while guaranteeing optimality up to  $\epsilon=10^{-4}$
takes only 18 minutes.

Results on the Pascal-VOC dataset are shown in Table~\ref{pascalacc}.
We compare against several approaches using the same unary potentials.
For completeness, we also list state-of-the-art approaches not based on CRF models.
Notably, out model matches or exceeds the performance of the much more involved approaches of
\citet{krahenbuhl2012efficient} and \citet{dann2012pottics} which use the same
unary potentials.
Using exact learning and inference slightly increased performance on this dataset.
Learning took 25 minutes using move-making alone and 100 minutes to guarantee optimality
up to $\epsilon=10^{-4}$.
A visual comparison of selected cases is shown in Figure~\ref{visual}.

The objective function values using only the under-generating move-making and
the exact inference are detailed in Table~\ref{msrc_objective} and Table~\ref{pascal_objective}.
We see that a significant gap between the cutting plane objective and the primal objective
remains when using only under-generating inference.
Additionally, the estimated primal objective $o^I$ using under-generating inference is
too optimistic, as can be expected. This underlines the fact that
under-generating approaches can not be used to upper-bound the primal
objective.

\begin{table}
    \begin{center}
    \begin{tabular}{lcc}
    \toprule
                    & Move-making & Exact \\
    \cmidrule{1-3}
    Dual Objective $o_\W$ &92.06& 92.24\\
    Estimated Objective $o^I$ & 92.07 &92.24\\
    True Primal Objective $o^E$&92.35& 92.24  \\
    \bottomrule
    \end{tabular}
    \end{center}
    \caption{Objective function values on the Pascal Dataset.}
    \label{pascal_objective}
\end{table}

\subsection{Implementation Details}
We implemented the cutting plane Algorithm~\ref{alg_cutting_plane} and our pairwise
model in Python.
Our cutting plane solver for Equation~\eqref{mainequation} uses
\texttt{cvxopt}~\citep{dahl2006cvxopt} for solving the QP in the inner loop. Code for
our structured prediction framework will be released under an open source
license upon acceptance.

We used the SLIC implementation provided by \citet{achanta2012slic} to extract superpixels and
the SIFT implementation in the \texttt{vlfeat} package~\citep{vedaldi08vlfeat}.
For clustering visual words, piecewise training of unary potentials and the
approximation to the $\chi^2$-kernel, we made use of the \texttt{scikit-learn}
machine learning package~\citep{pedregosa2011scikit}.
The move-making algorithm using QPBO is implemented with the help of the QPBO-I
method provided by \citet{rother2007optimizing}.
We use the excellent implementation of AD$^3$ provided by the authors of
\citet{martins2011augmented}.

Thanks to using these high-quality implementations, running the whole pipeline
for the pairwise model takes less than an hour on a 12 core CPU\@. Solving the
QP is done in a single thread, while inference is parallelized over all cores.

\section{Discussion}
In this work we demonstrated that it is possible to learn state-of-the-art CRF models
exactly using structural support vector machines, despite the model containing many loops.
The key to efficient learning is the combination of different inference mechanisms and
a novel caching scheme for the one-slack cutting plane method, in combination
with state-of-the-art inference methods.

We show that guaranteeing exact results is feasible in a practical setting, and
hope that this result provides a new perspective onto learning loopy models for
computer vision applications.

Even though exact learning does not necessarily lead to a large improvement in
accuracy, it frees the practitioner from worrying about optimization and
approximation issues, leaving more room for improving the model, instead of the
optimization.

While our model is trained on a relatively small dataset, we expect that scaling to
datasets of several thousand images, such as Pascal VOC, is unproblematic.
The time for a single iteration of the cutting plane solver will only grow
linearly for the inference. The cost of solving the QP is actually independent
of the number of samples, as only a single constraint will be added in each
iteration.

We do not expect learning of pixel-level models, which typically have tens or
hundreds of thousands of variables, to be efficient using exact inference. However we
believe our results will carry over to other super-pixel based approaches.
Using other over-generating techniques, such as duality-based message passing
algorithms, it might be possible to obtain meaningful bounds on the true
objective, even in the pixel-level domain.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figure}
\caption{%
    Visual comparison of the result of exact and approximate learning on
    selected images from the test set.  From left to right: the input image,
    prediction using approximate learning, prediction using exact learning, and
    ground truth.
\label{visual}}
\end{figure*}

### 2013-08-02 main.tex
init

\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{scrreprt}

\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{algorithm}
\usepackage{mdwlist}
\usepackage{url}
\usepackage[noend]{algpseudocode}
\usepackage[small]{caption}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\W}{\mathcal{W}}

\begin{document}

%\title{Learning a Loopy Model For Semantic Segmentation Exactly}

\begin{titlepage}
\begin{center}
\vspace*{1in}
\textbf{{\LARGE PhD Thesis}}
\par
\vspace{1.5in} {\large Andreas Mueller}
 \par \vfill A Thesis submitted for the degree of Doctor of Philosophy
\par \vspace{0.5in}
Department IV, Institute for Computer Science
\par \vspace{0.5in}
Friedrich-Willhelm Universitaet Bonn \par
\vspace{0.5in} September 2013 \end{center}

\end{titlepage}

\pagenumbering{Roman}
\tableofcontents
\listoffigures
\listoftables
 \chapter*{Acknowledgements}
 \begin{abstract} abstract
\end{abstract}
\pagenumbering{arabic}

\include{exact_learning/paper}

\end{document}

### 2013-08-02 itm/appendix.tex
copy *& paste everything

\documentclass[runningheads]{llncs}
\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%\usepackage[numbers]{natbib}
\usepackage{csquotes}
\usepackage[natbib, style=numeric-comp, firstinits=true, sorting=none, maxnames=2]{biblatex}

\renewcommand{\eqref}[1]{Equation~(\ref{#1})}
\mainmatter

\begin{document}

\begin{theorem}
Consider the two cluster case with $n$ data points.
Let $n_0, n_1$ be the cluster sizes and $L_0, L_1$ be the length of the minimum spanning trees in the obtimum solution.
The non-optimality of a solution obtained by ITM the search as above is bounded as
\begin{align}
    \hat{I}^* - \hat{I}_{\text{ITM}} \leq 
         d \left [ \log \Bigl(\frac{L_0 + e}{L_0}\Bigr) + \frac{n_1}{n} \log \Bigl(\frac{L_0}{L_1} \Bigr) \right] 
\end{align}
\begin{proof}
Remove an edge of the MST with such that the resulting clustering with
MST lengths $L'_0, L'_1$ and $n'_0, n'_1$ the corresponding cluster sizes fulfill $L'_0 \geq L'_1$ and $n'_1 \geq n'_0$.

Wlog. $L_0 \geq L_1$.
Let $2e$ be the length of the longest edge in the MST of the whole dataset minus its shortest edge.
Then, by construction $2 (L_0 + e) \geq L_0 + L_1 + 2e \geq L'_0 + L'_1$.
By monotonicity of the logarithm, $\log(L_0 + e) \geq \log(\frac{L'_0 + L'_1}{2})$.
Using the concavity of the logarithm, we obtain
\begin{align}\label{concavity}
    \log(L_0 + e) \geq \frac{1}{2} \log(L'_0) + \frac{1}{2} \log(L'_1) \geq \frac{n_0}{n}\log(L'_0) + \frac{n_1}{n}\log(L'_1)
\end{align}

Any clustering obtained in this way is certain to have an objective not higher than the solution obtained by ITM\@.
For ease of analysis we will compare the optimum solution against any clustering that fulfills this restriction.
Then 
\begin{align}
    \hat{I}^* - \hat{I}_{\text{ITM}}
    & = -\frac{d}{n} \left [n_0 \log(L_0) + n_1 \log(L_1) - n_0 \log(L'_0) - n_1 \log(L'_1) \right ] + Z\\
    %& = -\frac{d}{n} \left [n_0 \log(L_0) + n_1 \log(L_0) + n_1 \log \Bigl (\frac{L_1}{L_0} \Bigr ) - n_0 \log(L'_0) - n_1 \log(L'_1) \right ] \\
    & = -\frac{d}{n} \left [n \log(L_0) + n_1 \log\Bigl (\frac{L_1}{L_0} \Bigr) - n_0 \log(L'_0) - n_1 \log(L'_1) \right] +Z\\
    & \leq -\frac{d}{n} \left [n \log(L_0) + n_1 \log\Bigl (\frac{L_1}{L_0} \Bigr) - n\log(L_0 + e) \right] +Z\\
    & = -\frac{d}{n} \left [n \log(\frac{L_0}{L_0 + e}) + n_1 \log\Bigl (\frac{L_1}{L_0} \Bigr) \right] +Z\\
    & = d \left [ \log(\frac{L_0 + e}{L_0}) + \frac{n_1}{n} \log \Bigl(\frac{L_0}{L_1} \Bigr) \right] + Z, 
\end{align}
which concludes the proof.
\end{proof}

\end{theorem}
\end{document}

### 2013-08-02 itm/dagm2012submission.tex
copy *& paste everything

% last updated in April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010
% Modified for DAGM 2010 by SR
% Modified for DAGM 2011 by HF&CB, updated lncs.cls and splncs03.bst
% Modified for DAGM-OAGM 2012 by TP, updated lncs.cls and splncs03.bst

\documentclass[runningheads]{llncs}
\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.

%
% Definition of \DAGMreviewversion which 
% defines the use of line numbers etc.
\newcommand\DAGMreviewversion{
  \usepackage{lineno}
  \usepackage{color}
  \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
  \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}} 
  \linenumbers
}
%
% Please activate this command (\DAGMreviewversion) for a draft
% version of your manuscript which will be used during the review
% process.
% DO NOT USE this command for the camera ready version of your paper.
%
\DAGMreviewversion

\begin{document}
\pagestyle{headings}
\mainmatter

%
% Insert your submission number here
%
\def\DAGM12SubNumber{***}

%
% Replace with your title
%
\title{Author Guidelines for DAGM-OAGM Submission}

%
% DO NOT MODIFY these for the draft version that is used for the
% review process.
% 
%
\titlerunning{DAGM-OAGM 2012 Submission \#\DAGM12SubNumber{}. CONFIDENTIAL
  REVIEW COPY.}
\authorrunning{DAGM-OAGM 2012 Submission \#\DAGM12SubNumber{}. CONFIDENTIAL
  REVIEW COPY.}
\author{Anonymous DAGM-OAGM submission}
\institute{Paper ID \DAGM12SubNumber}

\maketitle

\begin{abstract}
The abstract should summarize the contents of the paper and should
contain at least 70 and at most 300 words. It should be set in 9-point
font size and should be inset 1.0~cm from the right and left margins.
\dots
\end{abstract}

\section{Introduction}

Please follow the steps outlined below when submitting your
manuscript\footnote{These instructions have been adapted from DAGM 2011.}.

%-------------------------------------------------------------------------

\subsection{Language}

All manuscripts must be in English.

\subsection{Paper length}
The maximum allowed paper length is 10 pages. Overlength papers will
simply not be reviewed. This includes papers where the margins and
formatting are deemed to have been significantly altered from those
laid down by this style guide.  The reason such papers will not be
reviewed is that there is no provision for supervised revisions of
manuscripts. The reviewing process cannot determine the suitability of
the paper for presentation in 10 pages if it is reviewed in 12.

\subsection{Dual submission}

By submitting a manuscript to DAGM-OAGM, the author(s) assert(s) that it has
not been previously published in substantially similar
form. Furthermore, no paper which contains significant overlap with
the contributions of this paper either has been or will be submitted
during the DAGM-OAGM 2012 review period to either a journal or a
conference.

If there are any papers that may appear to the reviewers to violate
this condition, then it is your responsibility to (1) cite these
papers -- preserving anonymity as described in section~\ref{sec:blind}
of this example paper, (2) argue in the body of your paper why your
DAGM-OAGM paper is nontrivially different from these concurrent
submissions, and (3) include anonymized versions of those papers in
the supplemental material.

\subsection{Supplemental Material}

Authors may optionally upload supplemental material. Typically, this
material might include result videos that cannot be included in
the main paper, anonymized related submissions to other conferences
and journals, and appendices or technical reports containing extended
proofs and mathematical derivations that are not essential for
understanding of the paper. Note that the contents of the supplemental
material should be referred to appropriately in the paper, and that
reviewers are not obliged to look at the submitted material.

All supplemental material must be zipped or tarred into a single
file. There is a 30MB limit on the size of this file. The deadline for
supplemental material is three days after the main paper deadline. To
limit the load on the servers, we ask authors to either submit the
supplemental material well before the main paper deadline, or after
the main paper deadline.

%------------------------------------------------------------------------- 
\subsection{Line numbering}

All lines should be numbered, as in this example document.  This makes
reviewing more efficient, because reviewers can refer to a line on a
page.  If you are preparing a document using a non-\LaTeX\
document preparation system, please arrange for an equivalent line numbering.

\subsection{Mathematics}

Please number all of your sections and displayed equations.  Again,
this makes reviewing more efficient, because reviewers can refer to a
line on a page.  Also, it is important for readers to be able to refer
to any particular equation.  Just because you didn't refer to it in
the text doesn't mean some future reader might not need to refer to
it.  It is cumbersome to have to use circumlocutions like ``the
equation second from the top of page 3 column 1''.  (Note that the
line numbering will not be present in the final copy, so is not an
alternative to equation numbers).  Some authors might benefit from
reading Mermin's description of how to write mathematics:
\url{http://www.cvpr.org/doc/mermin.pdf}.

\section{Blind review}
\label{sec:blind}

Many authors misunderstand the concept of anonymizing for blind
review.  Blind review does not mean that one must remove
citations to one's own work---in fact it is often impossible to
review a paper unless the previous citations are known and
available.

Blind review means that you do not use the words ``my'' or ``our''
when citing previous work.  That is all.  (But see below for 
techreports).

Saying ``this builds on the work of Lucy Smith [1]'' does not say
that you are Lucy Smith, it says that you are building on her
work.  If you are Smith and Jones, do not say ``as we show in
[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
paper, include reference 7 as you would any other cited work.

An example of a bad paper:
\begin{quote}
\begin{center}
    An analysis of the frobnicatable foo filter. 
\end{center}
\vspace{0.5em}

   In this paper we present a performance analysis of our
   previous paper [1], and show it to be inferior to all
   previously known methods.  Why the previous paper was
   accepted without this analysis is beyond me.

   [1] Removed for blind review
\end{quote}

An example of an excellent paper:

\begin{quote}
\begin{center}
     An analysis of the frobnicatable foo filter. 
\end{center}
\vspace{0.5em}

   In this paper we present a performance analysis of the
   paper of Smith [1], and show it to be inferior to
   all previously known methods.  Why the previous paper
   was accepted without this analysis is beyond me.

   [1] Smith, L and Jones, C. ``The frobnicatable foo
   filter, a fundamental contribution to human knowledge''.
   Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time,
which covers similar or overlapping material, you may need to refer to that
submission in order to explain the differences, just as you would if you
had previously published related work.  In such cases, include the
anonymized parallel submission~\cite{Authors10} as additional material and
cite it as
\begin{quote}
1. Authors. ``The frobnicatable foo filter'', BMVC 2010 Submission ID 324,
Supplied as additional material {\tt bmvc10.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be
found elsewhere, and refer them to a technical report.  For conference
submissions, the paper must stand on its own, and not {\em require} the
reviewer to go to a techreport for further details.  Thus, you may say in
the body of the paper ``further details may be found
in~\cite{Authors10b}''.  Then submit the techreport as additional material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool which
is widely known to be restricted to a single institution.  For example,
let's say it's 1969, you have solved a key problem on the Apollo lander,
and you believe that the DAGM-OAGM audience would like to hear about your
solution.  The work is a development of your celebrated 1968 paper entitled
``Zero-g frobnication: How being the only people in the world with access to
the Apollo lander source code makes us a wow at parties'', by Zeus.

You can handle this paper like any other.  Don't write ``We show how to
improve our previous work [Anonymous, 1968].  This time we tested the
algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors. Instead
write the following:
\begin{quotation}
\noindent
   We describe a system for zero-g frobnication.  This
   system is new because it handles the following cases:
   A, B.  Previous systems [Zeus et al. 1968] didn't
   handle case B properly.  Ours handles it by including
   a foo term in the bar integral.

   ...

   The proposed system was integrated with the Apollo
   lunar lander, and went all the way to the moon, don't
   you know.  It displayed the following behaviours
   which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention,
reads better than the first version, and does not explicitly name you as
the authors.  A reviewer might think it likely that the new paper was
written by Zeus, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been
contracted to solve problem B. \\

\noindent FAQ: Are acknowledgements OK?  No. Please {\bf omit
acknowledgements} in your review copy; they can go in the final copy.

\section{Manuscript Preparation}
This is an edited version of Springer LNCS instructions adapted for
DAGM-OAGM 2012 first paper submission.

You are strongly encouraged to use \LaTeX2$_\varepsilon$ for the
preparation of your
camera-ready manuscript together with the corresponding Springer
class file \verb+llncs.cls+.

We would like to stress that the class/style files and the template
should not be manipulated and that the guidelines regarding font sizes
and format should be adhered to. This is to ensure that the end product
is as homogeneous as possible.

\subsection{Printing Area}
The printing area is $122  \; \mbox{mm} \times 193 \;
\mbox{mm}$.
The text should be justified to occupy the full line width,
so that the right margin is not ragged, with words hyphenated as
appropriate. Please fill pages so that the length of the text
is no less than 180~mm.

\subsection{Layout, Typeface, Font Sizes, and Numbering}
Use 10-point type for the name(s) of the author(s) and 9-point type for
the address(es) and the abstract. For the main text, please use 10-point
type and single-line spacing.
We recommend using Computer Modern Roman (CM) fonts, Times, or one
of the similar typefaces widely used in photo-typesetting.
(In these typefaces the letters have serifs, i.e., short endstrokes at
the head and the foot of letters.)
Italic type may be used to emphasize words in running text. Bold
type and underlining should be avoided.
With these sizes, the interline distance should be set so that some 45
lines occur on a full-text page.

\subsubsection{Headings.}

Headings should be capitalized
(i.e., nouns, verbs, and all other words
except articles, prepositions, and conjunctions should be set with an
initial capital) and should,
with the exception of the title, be aligned to the left.
Words joined by a hyphen are subject to a special rule. If the first
word can stand alone, the second word should be capitalized.
The font sizes
are given in Table~\ref{table:headings}.
\setlength{\tabcolsep}{4pt}
\begin{table}
\begin{center}
\caption{Font sizes of headings. Table captions should always be
positioned {\it above} the tables. The final sentence of a table
caption should end without a full stop}
\label{table:headings}
\begin{tabular}{lll}
\hline\noalign{\smallskip}
Heading level & Example & Font size and style\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
Title (centered)  & {\Large \bf Lecture Notes \dots} & 14 point, bold\\
1st-level heading & {\large \bf 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bf 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bf Headings.} Text follows \dots & 10 point, bold
\\
4th-level heading & {\it Remark.} Text follows \dots & 10 point,
italic\\
\hline
\end{tabular}
\end{center}
\end{table}
\setlength{\tabcolsep}{1.4pt}

Here are
some examples of headings: ``Criteria to Disprove Context-Freeness of
Collage Languages'', ``On Correcting the Intrusion of Tracing
Non-deterministic Programs by Software'', ``A User-Friendly and
Extendable Data Distribution System'', ``Multi-flip Networks:
Parallelizing GenSAT'', ``Self-determinations of Man''.

\subsubsection{Lemmas, Propositions, and Theorems.}

The numbers accorded to lemmas, propositions, and theorems etc. should
appear in consecutive order, starting with the number 1, and not, for
example, with the number 11.

\subsection{Figures and Photographs}
\label{sect:figures}

Please produce your figures electronically and integrate
them into your text file. For \LaTeX\ users we recommend using package
\verb+graphicx+ or the style files \verb+psfig+ or \verb+epsf+.

Check that in line drawings, lines are not
interrupted and have constant width. Grids and details within the
figures must be clearly readable and may not be written one on top of
the other. Line drawings should have a resolution of at least 800 dpi
(preferably 1200 dpi).
For digital halftones 300 dpi is usually sufficient.
The lettering in figures should have a height of 2~mm (10-point type).
Figures should be scaled up or down accordingly.
Please do not use any absolute coordinates in figures.

Figures should be numbered and should have a caption which should
always be positioned {\it under} the figures, in contrast to the caption
belonging to a table, which should always appear {\it above} the table.
Please center the captions between the margins and set them in
9-point type
(Fig.~\ref{fig:example} shows an example).
The distance between text and figure should be about 8~mm, the
distance between figure and caption about 5~mm.
\begin{figure}
\centering
\includegraphics[height=7.2cm]{eijkel2}
\caption{One kernel at $x_s$ ({\it dotted kernel}) or two kernels at
$x_i$ and $x_j$ ({\it left and right}) lead to the same summed estimate
at $x_s$. This shows a figure consisting of different types of
lines. Elements of the figure described in the caption should be set in
italics,
in parentheses, as shown in this sample caption. The last
sentence of a figure caption should generally end without a full stop}
\label{fig:example}
\end{figure}

If possible (e.g. if you use \LaTeX) please define figures as floating
objects. \LaTeX\ users, please avoid using the location
parameter ``h'' for ``here''. If you have to insert a pagebreak before a
figure, please ensure that the previous page is completely filled.

\subsection{Formulas}

Displayed equations or formulas are centered and set on a separate
line (with an extra line or halfline space above and below). Displayed
expressions should be numbered for reference. The numbers should be
consecutive within each section or within the contribution,
with numbers enclosed in parentheses and set on the right margin.
For example,
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt \;  .
\end{equation}

Please punctuate a displayed equation in the same way as ordinary
text but with a small space before the end punctuation.

\subsection{Program Code}

Program listings or program commands in the text are normally set in
typewriter font, e.g., CMTT10 or Courier.

\medskip

\noindent
{\it Example of a Computer Program}
\begin{verbatim}
program Inflation (Output)
  {Assuming annual inflation rates of 7%, 8%, and 10%,...
   years};
   const
     MaxYears = 10;
   var
     Year: 0..MaxYears;
     Factor1, Factor2, Factor3: Real;
   begin
     Year := 0;
     Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
     WriteLn('Year  7% 8% 10%'); WriteLn;
     repeat
       Year := Year + 1;
       Factor1 := Factor1 * 1.07;
       Factor2 := Factor2 * 1.08;
       Factor3 := Factor3 * 1.10;
       WriteLn(Year:5,Factor1:7:3,Factor2:7:3,Factor3:7:3)
     until Year = MaxYears
end.
\end{verbatim}
%
\noindent
{\small (Example from Jensen K., Wirth N. (1991) Pascal user manual and
report. Springer, New York)}

\subsection{Footnotes}

The superscript numeral used to refer to a footnote appears in the text
either directly after the word to be discussed or -- in relation to a
phrase or a sentence -- following the punctuation sign (comma,
semicolon, or full stop). Footnotes should appear at the bottom of
the
normal text area, with a line of about 2~cm in \TeX\ and about 5~cm in
Word set
immediately above them.\footnote{The footnote numeral is set flush left
and the text follows with the usual word spacing. Second and subsequent
lines are indented. Footnotes should end with a full stop.}

\subsection{Citations}

The list of references is headed ``References" and is not assigned a
number
in the decimal system of headings. The list should be set in small print
and placed at the end of your contribution, in front of the appendix,
if one exists.
Please do not insert a pagebreak before the list of references if the
page is not completely filled.
An example is given at the
end of this information sheet. For citations in the text please use
square brackets and consecutive numbers: \cite{Alpher02},
\cite{Alpher03}, \cite{Alpher04} \dots

\bibliographystyle{splncs03}
\bibliography{egbib}

\clearpage\mbox{}Page \thepage\ of the manuscript.
\clearpage\mbox{}Page \thepage\ of the manuscript.
This is the last page of the manuscript. 
\par\vfill\par
Now we have reached the maximum size of the DAGM-OAGM 2012 submission.
\end{document}

### 2013-08-02 itm/paper.tex
copy *& paste everything

\section{Introduction}
Clustering data is one of the fundamental problems in machine learning.  In
clustering, the goal is to divide data points into homogeneous subsets, called
clusters.
Many different formulations of the clustering problem are given in the literature.
Most algorithms are based on ad-hoc criteria such as intra-cluster similarity
and inter-cluster dissimilarity.
%
An alternative approach is to formalize clustering using an information
theoretic framework, where one considers inputs as well as cluster assignments
as random variables.  The goal is then to find an assignment of data points to
clusters that maximizes the mutual information between the assignments and the
observations.

In this work, we rely on a non-parametric estimator of the data entropy to find 
clusterings of maximum mutual information.
%
The use of non-parametric estimates allows a data-driven approach, without 
making strong assumptions on the form of the data distribution.
%
As a consequence, we obtain a very flexible model that, e.g{.}, allows 
non-convex clusters.
%
The resulting objective is easy to evaluate, but difficult to optimize
over. We overcome this by proposing an efficient approximate optimization
based on the Euclidean minimum spanning tree algorithm. 
%
Because the estimator and the optimization are both parameter-free, 
the only free parameter of the algorithm is the number of clusters, 
which makes it very easy to use in practice.
%
The contributions of this work are:
\begin{itemize}
    \item Proposing the use of a MST-based entropy estimator in information theoretic clustering.
    \item Give a fast algorithm for a relaxed version of the resulting problem.
    \item Show the practicality on a number of synthetic and real datasets.
\end{itemize}

%CHL: Maybe use this somewhere else, here it sounds redundant.
%While other algorithms have to rely on continuous, gradient-based techniques,
%we can rely on fast, combinatorial optimization.
%
%The run time of the proposed algorithm is dominated by the computation of the Euclidean spanning tree of the
%data, which is possible essentially in $O(n \log(n))$~\cite{dtb2010, eppstein1999spanning}.

\section{Related Work}
The most commonly used clustering algorithm is the $k$-Means algorithm, also
known as Lloyd's algorithm \citep{macqueen1967some,lloyd1982least}. While
$k$-Means often works well in practice, one of its main drawbacks is the
restriction in cluster shape. 
%
They are given by the Voronoi tessellation of the cluster means and therefore always convex.
%Not only are clusters as found by $k$-Means convex,
%they are given by the Voronoi tessellation of the cluster means.

Another widely used method is spectral clustering~\citep{shi2000normalized,ng2002spectral}, 
which solves a graph partitioning problem on a similarity
graph constructed from the data. While spectral clustering is much more
flexible than $k$-Means it is quite sensitive to the particular choice of graph
construction and similarity measure. It is also computationally expensive to
compute, because clustering $n$ points requires computing the eigenvalues and
-vectors of an $n\times n$ matrix.

Information theoretic approaches to clustering were first investigated in the
context of document classification. In this setting, training examples are
described by a discrete distribution over words, leading to the task of
\emph{distributional clustering}, which was later related to the Information
Bottleneck method by \citep{slonim1999agglomerative}.
This setting was described in detail by \citep{dhillon2003divisive}. In
distributional clustering, it is assumed that the distribution of the data is
known explicitly (for example as word counts), which is not the case in our
setting.

Later, \citet{banerjee2005clustering} introduced the concept of Bregman
Information, generalizing mutual information of distributions, and showed how
this leads to a natural formulation of several clustering algorithms.
%
\citet{barber2006kernelized} construct a soft clustering by using a parametric model of $p(Y\given X)$.
%
The framework of mutual information based clustering was extended to
non-parametric entropy estimates by \citet{faivishevsky2010nonparametric}.
They use a nearest neighbor based estimator of the mutual information, called
MeanNN, that takes into account all possible neighborhoods, therefore combining
global and local influences. The approximate mutual information is maximized
using local search over labels.
%discuss this later in detail
%While optimizing MeanNN allows for non-convex
%clusters, in practice the result is strongly biased towards compact, convex
%regions. This can be explained by a close connections of MeanNN to $k$-Means,
%see Section~\ref{discussion} for an in-depth discussion.

Clustering algorithms based on minimum spanning trees have been studied early
on in the statistics community, due to their efficiency.  One of the earliest
methods is single-link agglomerative clustering~\citep{gower1969minimum}.
Single-link agglomerative clustering can be understood as a minimum spanning
tree-based approach in which the largest edge is removed until the desired
number of components is reached.  \citet{zahn1971graph} refines this criterion
by cutting edges that are longer than other edges in the vicinity. This
approach requires tuning several constants by hand. More recently,
\citet{grygorash2006minimum} proposed a hierarchical MST-based clustering
approach that iteratively cuts edges, merges points in the resulting
components, and rebuilds the spanning tree.
We will limit our discussion to the most widely used algorithm from~\citep{gower1969minimum}.
%
\section{Information Theoretic Clustering Using Nonparametric Entropy-Estimates}
In general, the goal of clustering can be formulated as follows: 
given a finite collection of samples $\mathbf{x} = (x_1, \dotsc, x_n)$, we want to 
assign cluster-memberships $\mathbf{y} = (y_1, \dotsc, y_n ), y_i \in \{1, \dotsc k\}$ to these samples.
%
We adopt the viewpoint of information theoretic clustering of \citet{gokcay2002information}, 
where the $x_i$ are considered i.i.d{.} samples from a distribution $p(X)$, and the 
$y_i$ are found such that the mutual information $I(X, Y)$ between the distribution $p(X)$ and the
assigned labels $p(Y)$ is maximized.
We can rewrite this objective as 
\begin{align}
         I(X, Y)
        &= D_{\text{KL}}(p(X, Y) \parallel p(X) p(Y)) = H(X) - \sum_{y=1}^k p(Y{=}y) H(X \given Y {=} y)
\end{align}
where
\begin{itemize}
\item $D_\text{KL} = \int_\mathcal{X} p(X) \ln(\frac{p(X)}{q(X)})dX$ is the Kullback-Leibler divergence,
\item $H(X) = \int_\mathcal{X} p(X) \ln(p(X))dX$ is the differential entropy, and
\item $H(X \given Y{=}y) = \int_\mathcal{X} p(X \given Y {=}y) \ln(p(X \given Y {=} y))dX$ is the conditional differential entropy.
    \end{itemize}
Expressing the mutual information in terms of the entropy is convenient, since
the objective then decomposes over the values of $Y$. 
%
Additionally, $H(X)$ is independent of the distribution of $Y$ and therefore
does not influence the search over $\mathbf{y}$.

Because we are given only a finite sample from $p(X)$, there is no way to
exactly compute $I(X, Y)$, and this is still true if we fix a set of cluster
indicators $y_i$ 
%
Possible ways to overcome this are:
\begin{enumerate}
    \item Fit a parametric model $\hat{p}(X, Y \given \theta)$ to the observations.
    \item Use a non-parametric model $\hat{x}$ to approximate $p(X, Y)$. 
    \item Estimate $H(X \given Y)$ directly using a non-parametric estimate.
\end{enumerate}
We choose the third option, as it is the most flexible while avoiding
the curse of dimensionality that comes with using non-parametric density estimates.

Let $\mathbf{x}_y$ be the set of $x_i$ with label $y$. 
Given a non-parametric density estimator $H_\text{est}$
we have $H_{est}(\mathbf{x}_y) \approx H(X \given Y {=} y)$, leading to the clustering problem
\begin{align}\label{general_objective}
     \max_{\mathbf{y}}\quad - \sum_{y=1}^k p(Y{=}y) H_\text{est}(\mathbf{x}_y),
\end{align}
where the probability $p(Y{=}y)$ is given by the empirical frequency of $y$, 
\mbox{$p(Y\,{=}\,y)= \frac{n_y}{n}$} for $n_y = \frac{|\{i \mid y_i = y \}|}{n}$.

\subsection{Minimum Spanning Tree Based Entropy Estimation}
From now on, we assume that $\mathcal{X}=\mathbb{R}^d$ and $p(X)$ is
absolute continuous.
This setting allows the use of the non-parametric entropy estimate of
\citet{hero1999asymptotic}, that constructs a minimum spanning tree of the data
and obtains an estimate of the data entropy from the logarithm of the length of
the spanning tree.
%
More precisely, the entropy estimate of a dataset $\mathbf{x} = (x_1, \dotsc, x_n)$ is given by
\begin{align}\label{hmst}
    H_{mst}(\mathbf{x}) = d \log(L) - (d-1) \log(n) + \log(\beta_d).
\end{align}
where $L$ is the length of a minimum spanning tree $T(\mathbf{x})$ of $\mathbf{x}$ and $\beta_d$ is an
unknown, but data-independent constant.
The estimator $H_{mst}$ is consistent in the sense that $H_{mst}(\mathbf{x}) \rightarrow H(X)$
for $n \rightarrow \infty$~\citep{hero1999asymptotic}.
Using \eqref{hmst} as a non-parametric entropy estimate in \eqref{general_objective} yields
the problem to maximize $\hat{I}(\mathbf{x}, \mathbf{y})$ with
\begin{align}\label{objective}
     \hat{I}(\mathbf{x}, \mathbf{y}) := & -\sum_{y=0}^k p(y) \Bigl[ d \log(L_y) - (d-1) \log{n_y} \Bigr] + C,\\
    = & - \sum_{y=0}^k p(y) \Bigl[ d \log(\bar{L}_y  ) + \log{n_y} \Bigr] + C'\\
    = & - d \sum_{y=0}^k p(y) \log(\bar{L}_y  ) - \sum_{y=0}^k p(y) \log{p(y) + C''} \label{objective_simplified}
\end{align}
where $n_y$ is the cardinality of $\mathbf{x}_y$, $L_y$ is the length of the
minimum spanning tree $T(\mathbf{x}_y)$ and $C$, $C'$ and $C''$ are constants independent of
$\mathbf{y}$.  We defined $\bar{L}_y := \frac{L_y}{n_y}$, the mean edge length
per node in $T(\mathbf{x}_y)$.

\eqref{objective_simplified} has a natural interpretation: The first term
penalizes long spanning trees, weighted by the size of the cluster. The second
term favors a high entropy of $p(y)$, leading to balanced clusters. Note that
there is a natural trade-off between enforcing intra-cluster similarity,
expressed through $L$ and the balancing of cluster sizes. This trade-off is
similar to formulating an objective in terms of a loss and a regularizer. In
contrast to the ``loss+regularizer'' setup, where the trade-off needs to be
specified by the user, the trade-off in \eqref{objective_simplified}, given by
the factor $d$, is a direct consequence of the entropy estimator.

The reliance on the dimensionality of the ambient space $\mathbb{R}^d$ can be
seen as the requirement that $d$ is actually the intrinsic dimensionality of
the data.
This requirement is made explicit in our assumptions of an absolute continuous
data density: If the support of $p(X)$ was a lower-dimensional sub-manifold of
$\mathbb{R}^d$, $p(X)$ could not be absolute continuous.

\begin{algorithm}[t]
\caption{Information Theoretic MST-based Clustering}

\begin{algorithmic}\label{overall}
    \Require Points $\mathbf{x}$, desired number of clusters $k$.
    \Ensure Clustering $\mathbf{y}$ of $\mathbf{x}$
    \State $G \leftarrow T(\mathbf{x})$
    \For{$i=0,\dotsc,k-1$}
        \For{$G_j, j=0,\dotsc, i$ connected components of $G$}
            \State $e_j \leftarrow \text{SplitCluster}(G_j)$
        \EndFor
        \State $l \leftarrow \displaystyle \argmax_j\  \hat{I}(G_j \setminus e_j)$
        \State $G \leftarrow G \setminus e_l$
    \EndFor
    \Statex
    \Function{SplitCluster}{$G$}
        %\Require Euclidean minimum spanning tree $T$, dimension $d$
        %\Ensure $e^* = \displaystyle \argmax_{e\in E(T)} \ \hat{I}(G \setminus e )$
        \State Pick arbitrary root $x_0$ of $G$.
        \For{node $x$ starting from leaves}
            \State $w_x \leftarrow \displaystyle \sum_{c \in \text{children}(x)} w_c + d(x, c)$
            \State $n_x \leftarrow \displaystyle 1 + \sum_{c \in \text{children}(x)} n_c$
        \EndFor
        \For{node $x$ starting from root}
        \State $w'_x \leftarrow w'_{\text{par}(x)} + w_{\text{par}(x)} - w_x + d(x, \text{par}(x))$
        \EndFor
        \For{$e \in E(G), e=(c, p)$, $p$ parent of $c$}
            \State $v_c \leftarrow w'_p + w_p - w_c - d(p,c)$
            \State $m_c \leftarrow n - n_c$
            \State objective$(e) \leftarrow d m_c \ln(m_c) - (d-1) m_c \ln(v_c) + d n_c \ln(n_c) - (d-1) n_c \ln (w_c)$
        \EndFor
        \State $e^* \leftarrow \displaystyle \argmax_{e \in E(G)} \ \text{objective}(e)$

    \EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Finding Euclidean Minimum Spanning Tree Clusterings}
The objective given by \eqref{objective} is a non-linear combinatorial
optimization problem. It has two properties that make it hard to optimize:
\begin{enumerate}
    \item The objective depends in a non-linear way on $L_y$.
        This makes linear programming techniques, that proved successful for
        other combinatorial task, not directly applicable.
    \item $L_y$ is defined in terms of minimum spanning trees. This set is hard
        to characterize, as changing the cluster membership of a single node may change
        the two minimum spanning trees involved completely.
\end{enumerate}

\begin{figure}
\centering
\includegraphics[width=.3\textwidth]{illustration_left}
\includegraphics[width=.3\textwidth]{illustration_center}
\includegraphics[width=.3\textwidth]{illustration_right}
\caption{Illustration of the optimization algorithm for $k=3$ on synthetic
dataset. \emph{Left}: Euclidean minimum spanning tree of the data.
\emph{Center}: The edge that yields the best two-cluster partition in terms of
\eqref{objective} was removed, yielding two connected components. \emph{Right}:
Another edge from the forest was removed, resulting in the desired number of
three components. Note that the edge that are removed are not the longest edges
but form a trade-off between edge length and cluster size.}
\label{illustration}
\end{figure}

For the above reasons, we propose a simple procedure to approximately solve
\eqref{objective}.
%filler removed Let us first fix some notation.
Consider a graph $G$ with nodes $\mathbf{x}$ and edge weights given by the
Euclidean distances between points.  The connected components of $G$ induce a
clustering $\mathbf{y}(G)$ of $\mathbf{x}$, by assigning $x_i$ and $x_j$ the
same cluster if and only if they are in the same connected component of $G$.
Define
\setlength{\abovedisplayskip}{1mm}
\setlength{\belowdisplayskip}{1mm}
\begin{align}\label{graphobjective}
\hat{I}(G) := - \sum_{y=0}^k p(y) \Bigl[ d \log(L_{G,y}) - (d-1) \log{n_y} \Bigr],
\end{align}
where $y$ enumerates the connected components $G_0, \dotsc, G_k$ of $G$, $n_y =
|V(G_y)|$ is the number of nodes in $G_y$ and $L_{G,y} = \sum_{e \in E(G_y)}
w(e)$ is the sum of the weights of all edges in the connected component $G_y$.
%
Then $\hat{I}(G) \geq \hat{I}(\mathbf{x}, \mathbf{y}(G))$, by the definition of
the minimum spanning tree, and equality holds if and only if $G_y$ is the
minimum spanning tree of its nodes for all $y$.
%
We try to find a graph $G$ with $k$ components, such that $\hat{I}(G)$ is maximal. We can restrict
ourself to optimizing over the set $\mathcal{F}$ of forests over $\mathbf{x}$ with $k$ components,
as adding edges inside connected components will only decrease the objective.
Thus we can formulate the clustering problem equivalently as
$\displaystyle \max_{G \in \mathcal{F}} \hat{I}(G)$.

Optimization over forests remains hard, and we further restrict ourself to
solutions from $\mathcal{G} := \{ F \in \mathcal{F} \mid F \text{\ subgraph of\
} T(\mathbf{x}) \}$ for a given minimum spanning tree $T(\mathbf{x})$, leading to
the problem $\max_{G \in \mathcal{G}} \hat{I}(G)$.  This restriction allows for
a very fast, combinatorial optimization procedure.

For the two class case, optimization of the above objective can be solved
exactly and efficiently by searching over all of $\mathcal{G}$.
This amounts to searching for the edge $e$ that maximizes $\hat{I}(T(\mathbf{x})
\setminus e)$. The naive algorithm that computes the objective for each edge
separately has run time that is quadratic in the number of data points.  To
improve upon this, we use a dynamic programming approach as described in
Algorithm~1, in function SplitCluster, which has only linear
complexity. Using this algorithm, run time in the two cluster case is dominated
by computing $T(\mathbf{x})$.
%
We extend this algorithm to the case of more than two clusters in a greedy way:
Starting with the full spanning tree of $\mathbf{x}$, we remove the edge
yielding the lowest value of \eqref{graphobjective} until the number of
components equals the number of desired clusters. The overall procedure is
summarized in Algorithm~1, an illustration can be found in
\figref{illustration}. We refer to Algorithm~1 as \emph{Information
Theoretic MST-based (ITM) clustering}.

We use the dual-tree Boruvka algorithm~\citep{dtb2010} to compute the minimum
spanning tree, which has runtime close to $O(n \log(n) \alpha(n))$. Here
$\alpha$ is the inverse of the Ackerman function, which grows so slowly as to
be considered constant in practice. The dynamic programming solution of
Algorithm~1 has a run time of $O(n)$ per removed edge, leading to an overall
run time of $O(n \log(n) \alpha(n)+ nk)$. The $O(nk)$ comes from a worst case
scenario, in which each step in the hierarchical clustering procedure only
splits off a constant number of points. In a more realistic setting, we expect
that the individual clusters are much smaller than the original dataset. In
this case, the $O(nk)$ factor would improve to $O(n\log(k))$.

\begin{figure}[t]
\centering
\includegraphics[width=.2\textwidth]{qualitative_blobs_km}
\includegraphics[width=.2\textwidth]{qualitative_blobs_mean_nn}
\includegraphics[width=.2\textwidth]{qualitative_blobs_aggl}
\includegraphics[width=.2\textwidth]{qualitative_blobs_mst} \\

\includegraphics[width=.2\textwidth]{qualitative_circles_km}
\includegraphics[width=.2\textwidth]{qualitative_circles_mean_nn}
\includegraphics[width=.2\textwidth]{qualitative_circles_aggl}
\includegraphics[width=.2\textwidth]{qualitative_circles_mst} \\

\includegraphics[width=.2\textwidth]{qualitative_circles2_km}
\includegraphics[width=.2\textwidth]{qualitative_circles2_mean_nn}
\includegraphics[width=.2\textwidth]{qualitative_circles2_aggl}
\includegraphics[width=.2\textwidth]{qualitative_circles2_mst} \\

\includegraphics[width=.2\textwidth]{qualitative_twomoons_km}
\includegraphics[width=.2\textwidth]{qualitative_twomoons_mean_nn}
\includegraphics[width=.2\textwidth]{qualitative_twomoons_aggl}
\includegraphics[width=.2\textwidth]{qualitative_twomoons_mst} \\ 
\caption{Comparison of $k$-Means (left), MeanNN (center left), single link
(center right) and ITM (right) on four synthetic datasets.  Without the need to
tune parameters, ITM can adjust to different cluster shapes.  MeanNN is able to
recover non-convex clusters (third row) but often produces similar results to
$k$-Means (second and last row). Single link clustering is very sensitive
to noise, as it does not take cluster size into account.}
\label{qualitative}
\end{figure}

\section{Experiments}
We compared ITM to the popular $k$-Means algorithm 
\citep{macqueen1967some,lloyd1982least}, to the MeanNN algorithm of
\citet{faivishevsky2010nonparametric} and to single-link agglomerative
clustering~\citep{gower1969minimum}. The similarities between single-link
agglomerative clustering and the proposed MST-based optimization make it a 
good baseline for tree-based clustering approaches.

A comparison of ITM, MeanNN and the baseline methods, $k$-Means and 
single link agglomerative clustering, in terms of their objective, 
optimization and complexity can be found in Table~\ref{nowotab}. 
We implemented the ITM clustering procedure as well as MeanNN in Python.  We
used the $k$-Means implementation available in the scikit-learn
library~\citep{pedregosa2011scikit}.  We use the dual tree Boruvka algorithm
implemented in the mlpack machine learning library~\citep{mlpack2011}. The source
code is available online\footnote{\url{https://github.com/amueller/information-theoretic-mst}}.
\enlargethispage{5mm}

\subsection{Experimental Setup}
For both $k$-Means and MeanNN, we restart the algorithm ten times using
different random initializations, keeping the result with the best objective
value. As ITM is deterministic there is no need for random restarts.  All of
the algorithms we compare work with a fixed number of clusters, which we set to
the number of classes in the dataset for all experiments.

As single link agglomerative clustering is sensitive to outliers, we set
a hard limit on the minimum number of samples per cluster of five for
the quantitative analysis.

\begin{table}[t]
\caption{Comparing properties of related algorithms.}\label{nowotab}
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lccc}
\toprule
Algorithm &     Objective &     Deterministic&      Complexity \\
\cmidrule{1-4}
$k$-Means &     $\displaystyle \sum_y \sum_{i, y_i = y} \| x_i - \mu_y \|^2$ & No  & $O(nk)$ per iteration
%\footnote{The worst-time complexity of the Lloyd-Algorithm is exponential~\cite{arthur2006slow}, though online algorithm are observed to converge in linear time~\cite{bottou1995convergence}.}
\\
MeanNN &    $\displaystyle \sum_y \log\left(\frac{1}{|\mathbf{x}_y|}\sum_{i,j, y_i=y_j=y} \| x_i - x_j \|^2 \right)$ & No & $O(n^2)$ per iteration\\
Single Link &   -- &    Yes    &    $O(n\log n)$\\
ITM & $\displaystyle \sum_{y=0}^k d p(y) \log(\bar{L}_y  ) + p(y) \log{p(y)}$ & Yes & $O(\alpha(n) n \log n + nk)$\\
\bottomrule
\end{tabularx}
\end{table}

%As our objective is motivated by limit properties of the entropy estimate of \eqref{hmst},
%we expect the objective to work best in a setting with many samples. To avoid tautological
%solutions, we set a hard lower limit to the number of points in each cluster.
%This limit was set to $10$ for all experiments, independent of the number of samples and clusters.

\subsection{Qualitative Results}
\figref{qualitative} shows qualitative results on three synthetic datasets.
For well separated, convex clusters, all four algorithms produce the same
clustering (see top row).  If the structure of the data is more complex, the
advantage of the proposed method is apparent.  Note that there was no need to
specify any other parameters than the number of clusters to produce these
results.  It is also noteworthy that the results of MeanNN are very close to
those produces by $k$-Means in most cases. This similarity can be explained
by the close relation of the objective functions, listed in Table~\ref{nowotab}.

\subsection{Quantitative Results}
We present results on several standard datasets from the UCI repository, 
selecting datasets that span a wide range of combinations of number
of samples, features and clusters.
To satisfy the assumption of absolute continuity of the data distribution, 
we restrict ourself to data with continuous features.

We evaluated the experiments using the \emph{adjusted Rand index 
(ARI}~\citep{hubert1985comparing} and \emph{normalized mutual
information (NMI)}~\citep{strehl2003cluster}, two popular measures 
of cluster quality~\citep{gomes2010discriminative,kamvar2003spectral}.
The Rand index~\citep{rand1971objective} between two clusterings counts on how
many pairs of points two clusterings agree. The adjusted Rand index contains a
calibration against chance performance.

Table~\ref{results} summarizes the results. The two entropy-based methods 
(MeanNN, ITM) have a clear advantage of the other methods, with ITM finding 
better clusterings than MeanNN in the majority of cases. 
%
The single link agglomerative clustering procedure produces reasonable 
results on datasets with little noise and well-separated clusters, but 
fails otherwise.
%
When inspecting the results, we observed that ITM produced several very small
clusters on the \emph{faces} dataset. Indeed, increasing the minimum cluster 
size to 6 or more improved the results to 0.59/0.84. A possible explanation 
for this is that very small clusters make the entropy estimate less reliable. 
%
The single-link method also benefited from this, improving its results to 
0.42/0.82.
The run time of computing the ITM clustering was dominated by the computation
of the MST of the data. The implementation in mlpack took
60 seconds on a desktop computer for \emph{usps}, the largest dataset in our 
experiments.
%
The other methods had run times in the order of seconds, but given the 
different implementations we used, this should not be interpreted as a 
general statement about the speed of the individual methods.

\begin{table}[t]
\caption{Scores (ARI/NMI) of $k$-Means, MeanNN, single link agglomerative clustering and
ITM on several benchmark datasets (higher is better). The best score for
each dataset is printed in bold.}\label{results}
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lrrrcccc}
\toprule
\multicolumn{4}{c}{Dataset} &\multicolumn{4}{c}{Results}\\
\cmidrule{1-4}
\cmidrule{5-8}
Description &     n &    d &   k &  $k$-Means        & MeanNN                & SL                       &   ITM\\
\cmidrule{1-4}                   \cmidrule{5-5}      \cmidrule{6-6}          \cmidrule{7-7}             \cmidrule{8-8}
digits      &  1797 &   64 &  10 &     0.62\,/\,0.71 &    0.67\,/\,0.76      &0.10\,/\,0.50             &\B{0.85}\,/\,\B{0.89}\\ 
faces       &   400 & 4096 &  40 &     0.41\,/\,0.76 & \B{0.49}\,/\,\B{0.80} &0.08\,/\,0.69             &0.02\,/\,0.49\\ %\footnotemark[4]
iris        &   150 &    4 &   3 &     0.72\,/\,0.76 &    0.75\,/\,0.78      &0.55\,/\,0.72             &\B{0.88}\,/\,\B{0.87}\\
usps        &  9298 &  256 &  10 &     0.52\,/\,0.61 & \B{0.54}\,/\,\B{0.65} &0.00\,/\,0.04             &   0.44\,/\,0.58\\ 
vehicle     &   846 &   18 &   4 &     0.10\,/\,0.15 &    0.09\,/\,0.11      &0.00\,/\,0.04             &\B{0.10}\,/\,\B{0.14}\\
vowel       &   990 &   10 &  11 &     0.17\,/\,0.37 &    0.19\,/\,\B{0.40}  &0.00\,/\,0.16             &\B{0.20}\,/\,0.39\\
waveform    &  5000 &   21 &   2 & \B{0.37}\,/\,0.35 &    0.30\,/\,\B{0.38}  &0.00\,/\,0.00             &    0.23\,/\,0.22\\
\bottomrule
\end{tabularx}
\end{table}

\section{Conclusions}
In this work we proposed the use of a minimum spanning tree based,
non-parametric entropy estimator in information theoretic clustering, ITM\@.
Thereby we extended the work of \citet{faivishevsky2010nonparametric} to a more
flexible and efficient entropy estimate.  We proposed an approximate optimization
method by formulating the clustering problem as a search over graphs.
The resulting algorithm is deterministic has sub-quadratic run time.
%
Empirical comparisons showed that the proposed method outperforms standard
algorithms and the non-parametric entropy based clustering of
\citep{faivishevsky2010nonparametric} on multiple benchmark datasets. We
demonstrated that ITM is able to detect non-convex clusters,
even in the presence of noise.
%
In contrast to other algorithms that can handle non-convex clusters, ITM has no
tuning parameters, as the objective presents a natural trade-off between
balancing cluster sizes and enforcing intra-cluster similarity.

A limitation of the proposed algorithm is that it is based on the assumption 
of an absolute continuous data distribution. This assumption eliminates the 
possibility of using categorical variables and data that lies on a submanifold 
of the input space. 
%
In future work we plan to investigate a way to overcome this limitation, 
for example by estimating the intrinsic dimensionality of the 
data~\citep{pettis1979intrinsic}.
%
We will also investigate optimizations of the objective 
\eqref{graphobjective} that go beyond the proposed method. 
Move-making algorithms seem a promising way to refine solutions 
found by Algorithm~1. Branch and bound techniques could provide 
an alternative approach.

\bibliography{paper.bib}

### 2013-08-02 main.tex
use chapters
copy *& paste everything

@@
 \usepackage[small]{caption}
+\usepackage{setspace}
 
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\newcommand{\figref}[1]{Figure~\ref{#1}}
+\renewcommand{\eqref}[1]{Equation~(\ref{#1})}
+\algrenewcommand{\algorithmicrequire}{\textbf{Input:}}
+\algrenewcommand{\algorithmicensure}{\textbf{Output:}}
+
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
 \newcommand{\W}{\mathcal{W}}
+
+\newcommand{\given}{\, | \,}
+\newcommand{\B}[1]{\textbf{#1}}
 
@@
 \pagenumbering{Roman}
+
 \tableofcontents
-\listoffigures
-\listoftables
+
+%\listoffigures
+%\listoftables
 \chapter*{Acknowledgements}
+
 \begin{abstract} abstract
 \end{abstract}
+
 \pagenumbering{arabic}
+
+\chapter{Information Theoretic Clustering for Object Class Discovey}
+\include{itm/paper}
+
+\chapter{Weakly Supervised Object segmentation via Multiple Instance Learning}
+\include{partial_supervision/paper}
+\include{mil/paper}
+
+\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}
+
+\include{pystruct/paper}
+
+\chapter{Efficient Exact Learning of Loopy CRFs}
 
@@
 
+\chapter{Hierarchical Latent Variable Models for Semantic Segmentation}
+
+\bibliography{main}
 \end{document}

### 2013-08-02 mil/paper.tex
copy *& paste everything

%\newtheorem{corollary}{Corollary}
%\newtheorem{corollarya}{Corollary}
%\newtheorem{lemma}{Lemma}
%\newtheorem{theorem}{Theorem}
%\newtheorem{theorema}{Theorem}

%\newcolumntype{R}{>{\raggedleft\arraybackslash}X}
%\newcolumntype{C}{>{\centered\arraybackslash}X}

%\newcommand{\fix}{\marginpar{FIX}} \newcommand{\new}{\marginpar{NEW}}
%\newcommand\etal{~et\,al.~}

%\aistatsauthor{Andreas M\"uller\thanks{This work was founded by the B-IT research
%school.}  \and Sven Behnke\\ {amueller@ais.uni-bonn.de},
%{behnke@cs.uni-bonn.de}} {Department of Computer Science\\
%University of Bonn\\ 53113 Bonn, Germany \\ }

\section{Introduction}\seclabel{intro}
Multiple instance learning is a supervised learning scenario first proposed by \citet{dietterich1997solving}.
In multiple instance learning, examples are multi sets of instances, called bags, and each bag is assigned a class label.
It is assumed that each instance has a true label, defined by an underlying concept. The label of a bag is positive, if and only if one of
the instances it contains is labeled positive. The true label of instances is assumed to be unobserved by the learner who only has access to bag labels.

In the original formulation of multiple instance learning, the goal of the learner
was to predict the label of bags. Some recent works also aim at predicting
the label of individual instances, a much harder problem.
This can be viewed as learning an instance-level classifier using a form of weak supervision.

Learning to predict instance labels has been attempted in some recent work~\citep{liconvex2010,zhang2002dd}
but there has been very little theoretical study when and to which extend this is possible.
In this work, we explore which assumptions about the distributions of instances and labels
are necessary to learn instance-level classifiers from bag labels only.
We provide bounds on the instance label prediction error based on the
bag label prediction error. The bounds can be used to judge a learner when no
instance labeling exists for training or testing.
%If one is interested in classifying instances, such a bound is very important.
There are two applications that we target:
\begin{description}
\item[Setting 1:] Bag labels for a validation set are known and instance predictions need to be judged.
\item[Setting 2:] Bag labels for a test set can be bound using standard learning theoretic bounds and instance
predictions need to be judged.
\end{description}
The first case is very common in practical settings, where
cross validation is often used to set parameters of a predictor. This is
particularly important in support vector machines (SVMs) which are a popular
learning method for multiple instance problems~\citep{andrews2003support}.
If no instance labels are present for training, it is only possible to
choose parameters to minimize bag prediction error, which might be suboptimal if one
wants to minimize instance label misclassification. 
The bounds can also be used to find the amount of positive instances in positive bags,
known as the witness rate, which is an integral part of established multiple
instance algorithms~\citep{zhang2002dd,liconvex2010}.
The second case can be used to convert established bounds on the bag prediction error
to instance prediction error.

We show that it is possible to bound instance prediction error without
ever observing any instance labels even in fairly unrestricted settings.
To the authors knowledge, this is the first result of this kind.

Detailed proofs of all propositions can be found in the appendix.

\section{Related Work}
The theoretical analysis of multiple instance learning started with \citet{auer1997approximating} and later \citet{blum1998note}.
These works assume that instances are drawn i.i.d.\ from a single distribution over instances. Blum and Kalai showed that in this
setting, if the underlying concept is PAC-learnable in polynomial time, then the multiple instance learning problem is PAC-learnable
in polynomial time. Auer et\,al.\ showed that if there is an algorithm to PAC-learn multiple instance learning with axis-parallel
rectangles that is polynomial in size of the bag and dimension of the instance features, then it is possible to polynomially
PAC-learn DNF-formulars, a problem that is solvable only if $\text{RP}=\text{NP}$.
Recently, \citet{sabato2009homogeneous,DBLP:journals/corr/abs-1107-2021} proved bounds on the VC dimension and
fat shattering dimension of multiple instance learners by the corresponding dimension for a base learner.

All of the results above only consider predicting bag labels. There has been very little theoretical work
on predicting instance labels. In spite of the large amount of research on multiple instance learning algorithms 
(for example \citet{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles})
to the authors knowledge, the only quantitative evaluations of predicting
instance labels are by \citet{gehler2007deterministic} and by \citet{liconvex2010}. While these works show promising results,
no analysis of instance label prediction is given.

Sabato et\,al.\ also studied multiple instance learning as a way
to reduce labeling complexity~\citep{sabato2010reducing}.
In the setting of reducing label complexity, the learner creates
the bags itself. The instances can therefore be assumed to be globally i.i.d.\ and
the distribution of labels inside the bags is known a priori.
Under these assumptions, the instance label prediction error is bound using the bag label error.
This result is used to evaluate the sample complexity of learning from bags using empirical risk minimization.
In this work, we generalize some results from \citet{sabato2010reducing} to the setting where the distribution
of instance labels is unknown and to cases where instances inside a bag may be dependent.

\section{Preliminaries}
Throughout this work, $\mathcal{X}$ denotes the space of all instances. Bags are multi sets of instance $X=\{x_1,\dotsc, x_r\}$.
As usual in supervised learning tasks, we assume our training examples ($X_i,Y_i)$ to be i.i.d.\ from an unknown distribution.
We denote the label of an instance $x_i$ by $y_i$ and the label of the bag $X_i$ as $Y_i = \max\{y_j | x_j \in  X_i\}$.
We assume the bag size $r\in \mathbb{N}$ to be constant over all bags in a given problem. This restriction simplifies the treatment
significantly while our results can in great parts be carried over to settings with varying bag sizes.
We consider arbitrary classifiers $h_I\colon \mathcal{X} \rightarrow \{0, 1\}$ with associated bag classifier
$h_B(X) = \max \{h(x_1),\dotsc,h(x_r)\}$.
The goal is to learn a classifier $h$ which minimizes expected instance misclassification, given training bags and bag labels
$X_i$ and $Y_i$.
We do not consider specific hypotheses classes but instead try to bound the expected error of $h$ using the expected error of $h_B$.
During all of this work, we assume that the true instance labels are never observed.
We use the following notation: $p(h_I(x)=0):= h_0$, the negative rate of the classifier,
$p(h_I(x) \neq y) := e_I$, the instance error, and $p(h_B(X) \neq Y) := e_B$, the bag error.

\section{Necessary Assumptions}\seclabel{assumptions}
If we allow instances inside bags and their labels to have arbitrary dependencies, we are trying to find
$y_1,\dotsc,y_r$ maximizing $p(y_1,\dotsc, y_r | x_1, \dotsc, x_r)$. This is a structured 
prediction problem for which there is no hope solving it, given only bag labels.
Therefore, as usual in multiple instance learning, we assume there is an underlying concept for instances,
i.\,e.\ the label of each instance depends only on this instance.
In other words, we assume $p(y_1,\dotsc, y_r | x_1, \dotsc, x_r)=p(y_1|x_1) \cdot p(y_r|x_r)$.

If we allow arbitrary dependencies between the instances $x_i$ in a given bag, there is a simple example 
\citep{sabato2009homogeneous} that shows that learning instance classification is not necessarily possible.
Let $\mathcal{X}$ consist of two
distinct points $x_1$ and $x_2$ with corresponding labels $y_1 = 1$ $y_2 = -1$. If the distribution
of instances inside bags is such that each bag either contains two copies of $x_1$ or both of $x_1$ and $x_2$, then
by predicting all bags as positive, one can obtain a perfect bag classifier, while it is impossible to say anything
about the label of $x_2$.
Therefore, we need to make additional assumptions about the distribution of instances inside bags to enable
predicting instance labels.

\section{Independent Case}\seclabel{iid}
The simplest case in which learning instance prediction is possible is the case when instances are globally i.i.d.,
i.\,e.\ when instances inside a bag are independent of each other. This is a very strong assumption that might be
violated in practical applications. On the other hand, strong results can be proved in this setting.

\begin{theorem}\label{basicthm}
Let all instances be independent. With bags $X$, instances $x$, true labels $y,Y$, $h_I$ any classifier
on instances and $h_B$ the resulting bag classifier, we have
\begin{align*}
    %\eqlabel{iid}
   e_I = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}.
\end{align*}
\end{theorem}
This is a (slight) generalization of Theorem 1 in \citet{sabato2010reducing} and serves as a basis for all further bounds.

In particular:
\begin{corollary}\label{perfect}
If all instances are independent and bag classification is perfect, meaning $e_B=0$, then instance classification is also perfect ($e_I =0$).
\end{corollary}
This is an very interesting result in that if we do not assume independence, this corollary is wrong,
as shown by the example in \Secref{assumptions}.

In practice, $p(y=0)$ is unknown, as we assume a setting where no instance labels are available.
Thus finding a bound independent of $p(y=0)$ is desirable. The following theorem accomplishes this:

\begin{theorem}\label{iidbound}
If $h_0^r  \geq e_B$ and all instances are independent, we have
\begin{align}
    p(h_I(x)\neq y) \leq h_0 - \left ( h_0^r - e_B \right ) ^ \frac{1}{r}
\end{align}
\end{theorem}

This result, which applies to the Setting $2$ from \Secref{intro},
can be used in several ways. First, it can give a general idea about the quality of instance-level
predictions, without assuming any knowledge of the distribution of instance labels. In this inequality,
$h_0$ can simply be read off on training and/or test data. The other quantity, $e_B$ can be
estimated using standard supervised generalization bounds.

If we assume Setting $1$ from \Secref{intro}, where bag labels are known, we can say even more:
\begin{theorem}\label{equality}
Let all instances be independent. Given $p(Y)$, the instance error $e_I$ can be calculated by
\begin{align}
    %\eqlabel{iid}
e_I = h_0 + p(Y=0)^\frac{1}{r} - 2 \left (\frac{1}{2} ( h_0^r + p(Y=0) - e_B) \right)^\frac{1}{r}.
\end{align}
\end{theorem}
If bag labels are given, $p(Y=0)$ and $ e_B$ can simply be estimated from validation data.
This means, in the i.i.d.\ case, instance-level error can simply be calculated from bag label and classifier statistics.

\section{Restricted Dependent Case}
As remarked in \Secref{assumptions}, it is not possible to bound the instance classification error
when allowing arbitrary dependencies between instances. On the other hand, the i.i.d.\ assumption of \Secref{iid}
might be to strong in a practical setting.
Therefore, we propose an assumption that is both very general and allows learning of instance classification:
We assume that each bag has a ``hidden cause'' $Z$, conditioned on which all instances are independent.
More formally, we assume
\begin{align}
    p(x_1,\dotsc,x_r)=\int_Z p(x_1)\cdot \dotsc \cdot p(x_r) dZ.
\end{align}
The intuition behind this assumption is as follows. In the classical case of drug activity prediction~\citep{dietterich1997solving}, a bag corresponds
to a molecule and each instance corresponds to a different geometric configurations. These are certainly not independent.
If one knew the ``true'' geometric characteristics of the molecule, each configuration could be derived from that.

Another popular example of multiple instance learning is image or scene classification~\citep{zhou2007multi,zha2008joint,zhou2009multi}.
In this case, each image
is represented as a collection of interest points or segments.
For the sake of the argument let us assume that interest points or segments correspond to objects.
Again, objects in a scene are not independent. For example, if we know an image contains a chair and a TV set,
it is very unlikely to contain a bus.
On the other hand, if we knew something about the scene, such as ``this is a living room scene'', then the appearance
of certain objects does not tell us much about other objects.

The idea of introducing a hidden cause is similar to the idea of viewing a bag as a manifold and instances as points
sampled on this manifold, as in \citet{ICML2011Babenko_74}.

Conditioned on the hidden cause $Z$ for a given bag, the proof of Theorem~\ref{basicthm} still holds, yielding
\begin{align}\eqlabel{conditioned}
&p(h_I(x) \neq y | Z) =p(h_I(x)=0|Z) + p(y=0|Z)\\
&\textstyle- 2  [\frac{1}{2} ( p(h_I(x)=0|Z)^r + p(y=0|Z)^r\\
& - p(h_B(X) \neq Y|Z)) ]^\frac{1}{r}
\end{align}
Similarly, the idea of Corollary~\ref{perfect} yields:

\begin{corollary}\label{perfect}
If the bag classification is perfect, meaning $e_B=0$, then instance classification is also perfect.
\end{corollary}

While this is a very simple result, is has strong implications. In particular, albeit we generalized our
assumptions significantly, learning instance label prediction is still possible.
As we assume no knowledge of the distribution of the hidden bag variable $Z$ or even about its domain, it is significantly harder
to bound the instance error than in the i.i.d.\ case.
Using Theorem~\ref{iidbound} and taking expected values with respect to $Z$, we obtain
\begin{align}
    & p(h_I(x)\neq y) \\\
    & \leq h_0 - \mathbb{E}_Z \left [\left ( p(h_I(x)=0 | Z)^r - p(h_B(X)\neq Y|Z) \right ) ^ \frac{1}{r} \right ].
\end{align}

By bounding the right hand side we obtain:
\begin{theorem}\label{bound}
    Assume all instances within a bag are independent given an unobserved bag variable. If $h_0^r \geq e_B$ the instance error $e_I$ can
    be bound by
\begin{align}
    e_I \leq h_0 -  \left ( h_0^r - e_B) \right ).
\end{align}
\end{theorem}
While this bound might not be very tight -- for example if the variables are i.i.d.\ and $r$ is big, it is clear
that it is very loose compared to Theorem~\ref{iidbound} -- it is to the authors knowledge the first theorem that shows that instance-level
predictions can not be arbitrarily bad, even in a complicated dependent setting.

If we turn to the setting where bag labels are known, the bound of Theorem~\ref{bound} can be extended outside $( h_0^r > e_B)$
using the same statement for $p(y=0)$ instead of $h_0$. If there are places where both $h_0^r < e_B$ and
$p(y=0)^r < e_B$, we can use another bound, finally yielding:

\begin{theorem}\label{globalbound}
    If all instances within a bag are independent given an unobserved bag variable and the distribution of
    bag labeld $P(Y)$ is known, the instance error $e_I$ can
    be bound by
\begin{align}
    p(h_I(x)\neq y) \leq \text{min}\{ b_h, b_Y, b_g \} 
\end{align}
with
\begin{align*}
b_g &= h_0 + p(Y=0)^\frac{1}{r} - h_0^r - p(Y=0) + e_B,  \\
b_h &= h_0 - h_0^r + e_B \text{\ if } h_0^r > e_B \text{, else }\infty, \\
b_y &= p(Y=0)^\frac{1}{r} - p(Y=0) + e_B \\
&\text{\ if } p(Y=0) > e_B \text{, else }\infty. \\
\end{align*}
\end{theorem}

\begin{figure}[tbp]
	\begin{center}
		\includegraphics[height=40mm]{images/gehler_decision_boundary.png}\hspace{5px}
		\includegraphics[height=40mm]{images/gehler_instances.png}
	\end{center}
	\caption{Visualization of the dataset from \citet{gehler2007deterministic}. Left: Regions in $\mathbb{R}^2$
    corresponding to negative and positive instances. Right: Training points for 40\% positive rate. Red denotes negative
    bags, blue denotes positive bags. Crosses denote positive instances, circles negative instances.}
	\figlabel{gehler}

\end{figure}
To get a better handle on the behavior of instance errors, it may also be useful to lower-bound $e_I$.
This is possible using Theorem~\ref{basicthm}:

\begin{theorem}\label{lowerbound}
If the instance labels are independent inside a bag given a hidden cause, the instance
error $e_I$ has the following lower bound:
\begin{multline*}
e_I \geq h_0 + p(y=0)\\ 
- 2 \left (\frac{1}{2} (p(h_B(X)=0) + p(Y=0) - e_B) \right)^\frac{1}{r}.
\end{multline*}
\end{theorem}

\section{Experiments}
We evaluate the tightness of our bounds and their utility for finding parameters on the sythetic dataset of \citet{gehler2007deterministic}.
This dataset allows us to vary $p(y=0)$, which is half of the positive rate in positive bags.
A visualization of the dataset, which we refer to as Gehler-org is shown in \Figref{gehler}. In Gehler-org, there are as many positive
as negative bags, 30 of which are commonly used for training and 100 of which are used for testing.
We use 1000 bags for testing, resulting in more stable estimates of errors and bounds.
When generating the dataset, one fixes a rate of positive instances, also known as witness rate, which determines the ratio
of positive instances in a positive bag.

To evaluate our bounds in the i.i.d.\ setting, we have to modify the dataset slightly. We fix bag size at ten instances and generate instances
with a fraction of positive instances of half the witness rate. Then we randomly assign these instances to bags.
This way, the distribution over instances stays the same as in Gehler-org while instances inside bags become independent.
We refer to this dataset as Gehler-iid.
As a multiple instance learning algorithm, we choose SVM-SVR from \citet{liconvex2010}, as a state-of-the-art method that
provides instance predictions. Support vector regression is learned to estimate likelihood ratios of label instances.

\begin{figure}[tbp]
	\begin{center}
		\includegraphics[height=63mm]{images/iid_01.png}
		\includegraphics[height=63mm]{images/iid_02.png}
		\includegraphics[height=63mm]{images/iid_03.png}
	\end{center}
	\caption{Error rates and bounds for dataset Gehler-iid. Top, center and bottom correspond to positive rates of 0.1, 0.2 and 0.3 respectively.
    We only look at low positive rates as in the i.i.d.\ case, high positive rates result in all bags being positive.
    See the text for details.}
	\figlabel{gehler_iid}
\end{figure}

\Figref{gehler_iid} compares the true instance error $e_I$, the bag error $e_B$, the estimate of $e_I$ using Theorem~\ref{equality}
and the bound of Theorem~\ref{iidbound} for datasets with different values of the positive instance rate.
The $x-axis$ shows different thresholds on estimated likelihood ratios, resulting in varying $h_0$.
It is apparent that the estimate of Theorem~\ref{equality} reflects the true error closely, with only little noise from the restricted
sample size. Note that the upper bound of Theorem~\ref{iidbound} is not defined everywhere. While it is quite tight when
$h_I^r \gg e_B$, it is not when $h_I^r \approx e_B$.

We also performed experiments using a dependent version of Gehler-org
with probabilistic constraints. Ten examples are drawn per bag and each example is positive with probability
of the fixed positive rate. We denote this dataset by Gehler-soft. In the original dataset, the number of positives and negatives in a positive bag
was deterministic, making the instances strongly dependent.

The new dataset Gehler-soft violates the i.i.d.\ assumption but satisfies the
restricted dependent assumption. \Figref{gehler_soft}
shows the values of the bound from Theorem~\ref{globalbound} and Theorem~\ref{lowerbound} compared to the
true instance error $e_I$ for different settings of the positive instance rate.
Inspecting the lower bound, which uses the same computation as the estimate in the i.i.d.\ case, we see that due to the limited
dataset size, the estimated lower bound actually lies above the actual error in some cases. On the other hand, for low positive
rates, the lower bound follows the true error quite closely.
The upper bound behaves quite different, being quite loose for low instance rates and thighter for high instance rates.
While the bounds on the true error might not always be very thight, it should be observed that they are clearly
non-trivial. This is surprising since we made only very general assumptions on the distribution of instances.

\begin{figure}[tbp]
	\begin{center}
		\includegraphics[height=63mm]{images/non_iid_02.png}
		\includegraphics[height=63mm]{images/non_iid_05.png}
		\includegraphics[height=63mm]{images/non_iid_09.png}
	\end{center}
	\caption{Error rates and bounds for dataset Gehler-soft. Top, center and bottom correspond to positive rates of 0.2, 0.5 and 0.9 respectively.
    See the text for details.}
	\figlabel{gehler_soft}
\end{figure}

\section{Conclusion}
We provided several bounds on instance prediction error in a multiple instance learning setting.
For the case that instances inside a bag are independent, we proved a strong upper bound on the error.
We showed that when bag labels are present, it is even possible to directly estimate the true instance error.

We introduced a restricted dependent setting for multiple instance learning, in which instances inside a bag
are only independent given an unobserved variable. We showed that without any further assumptions
on the distribution of this variable, it is possible to learn instance label prediction from bag labels
in this setting. We also provided upper and lower bounds on the instance error, given known bag labels
or an estimage of the bag error.
Empirical evaluation on a synthetic dataset with known dependen structure and known instance labels
showed that the bound are non-trivial and yield insight into the behavior of instance label errors.
We hope that this work establishes a basis for further research on instance label prediction
in multiple instance learning.
\bibliographystyle{plainnat} \bibliography{paper}

\section{Appendix}
\begin{lemma}\label{binarylemma}
Given two (dependent) binary random variables $x,y$:
\begin{equation}
p(x \neq y) = p(x=0) + p(y=0) - 2 p(x=0, y=0)
\end{equation}

\begin{proof}
\begin{align}
p(x \neq y) & = p(x=0, y=1) + p(x=1, y=0) \\
                &= p(x=0) - p(x=0, y=0)\\
                &  + p(y=0) - p(x=0, y=0) \\
                &= p(x=0) + p(y=0) -2 p(x=0, y=0)
\end{align}
\end{proof}
\end{lemma}

\begin{theorema}
Let all instances be independent. With bags $X$, instances $x$, true labels $y,Y$, $h_I$ any classifier
on instances and $h_B$ the resulting bag classifier, we have
\begin{align*}
   e_I = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}
\end{align*}
\begin{proof}

\begin{align*}
e_B & = p(h_B(X) = 0) + p(Y=0) -2 p(h_B(X)=0, Y=0) \\\
&\text{using the lemma}\\
& = h_0^r + p(y=0)^r -2 p(h_I(x)=0, y=0)^r\\\
&\text{by independence}\\
& = h_0^r + p(y=0)^r -2 \left ( \frac{1}{2} (h_0 + p(y=0) - e_I) \right ) ^r \\
&\text{using Lemma~\ref{binarylemma} on }p(h_I(x)=0, y=0.
\end{align*}
Solving for $p(h_I(x)\neq y)$:
\begin{align*}
    & \left ( \frac{1}{2} (h_0 + p(y=0) - e_I )\right ) ^r = \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right )\\
    & \Rightarrow \frac{1}{2} (h_0 + p(y=0) - e_I)   = \left ( \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right ) \right )^\frac{1}{r}\\
    & \Rightarrow e_I   = h_0 + p(y=0) - 2 \left ( \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right ) \right )^\frac{1}{r}
\end{align*}
\end{proof}
\end{theorema}

\begin{corollarya}
If all instances are independent and bag classification is perfect, meaning $e_B=0$, then instance classification is also perfect ($e_I =0$).

\begin{proof}
Consider Theorem~\ref{basicthm}. If $e_B=0$ we have
\begin{align*}
p(h_I(x) \neq y) = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r)) \right)^\frac{1}{r}
\end{align*}
By the convexity of the power function the right hand side of the equation is always greater or equal zero.
Since $e_I \geq 0$, both sides are equal zero.
\end{proof}
\end{corollarya}

\begin{theorema}
    If $h_0^r  \geq e_B$ and all instances are independent, we have
\begin{align}
    p(h_I(x)\neq y) \leq h_0 - \left ( h_0^r - e_B) \right ) ^ \frac{1}{r}
\end{align}
\begin{proof}
    Assume $h_0^r  \geq e_B$.
    From Theorem~\ref{basicthm} we have
\begin{align*}
   e_I &= h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}\\
   &\leq h_0 + p(y=0) - \left [( h_0^r + p(y=0)^r - e_B) \right]^\frac{1}{r}\\
   &- \left ( p(y=0)^r \right )^\frac{1}{r}\\
   &= h_0 - \left (( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}\\
\end{align*}
\end{proof}
\end{theorema}

\begin{theorema}
Let all instances be independent. Given $p(Y)$, the instance error $e_I$ can be calculated by
\begin{align}
    %\eqlabel{iid}
e_I = h_0 + p(Y=0)^\frac{1}{r} - 2 \left (\frac{1}{2} ( h_0^r + p(Y=0) - e_B) \right)^\frac{1}{r}.
\end{align}
\begin{proof}
    This follows from Theorem~\ref{basicthm} by using $p(Y=0)=p(y=0)^r$.
\end{proof}
\end{theorema}

\begin{theorema}
    Assume all instances within a bag are independent given an unobserved bag variable. If $h_0^r \geq e_B$ the instance error $e_I$ can
    be bound by
\begin{align}
    e_I \leq h_0 -  \left ( h_0^r - e_B) \right ).
\end{align}
\begin{proof}
    Using the proof of Theorem~\ref{iidbound}, we can obtain
\begin{multline}
    p(h_I(x)\neq y| Z) \leq p(h_I(x)=0|Z) \\
    - \left ( p(h_B(x)=0|Z) - p(h_B(X)\neq Y| Z) \right ) ^ \frac{1}{r}
\end{multline}
Taking expected values with respect to $Z$ on both sides yields
\begin{align}
    e_I \leq h_0 - \mathbb{E}_Z \left ( p(h_I(x)=0|Z)^r - p(h_B(X)\neq Y| Z) \right ) ^ \frac{1}{r}
\end{align}
Since $0 \leq p(h_I(x)=0|Z)^r - p(h_B(X)\neq Y| Z) \leq 1$, the $r$-th root of this term can
be lower bounded by the identity function, resulting in
\begin{multline}
    e_I \leq h_0 - \mathbb{E}_Z \left[  p(h_I(x)=0|Z)^r \right] - e_B 
\end{multline}
The theorem follows with Jensen's inequality.
\end{proof}
\end{theorema}

\begin{theorema}
    If all instances within a bag are independent given an unobserved bag variable and the distribution of
    bag labeld $P(Y)$ is known, the instance error $e_I$ can
    be bound by
\begin{align}
    p(h_I(x)\neq y) \leq \text{min}\{ b_h, b_Y, b_g \} 
\end{align}
with
\begin{align}
b_g &= h_0 + p(Y=0)^\frac{1}{r} - h_0^r - p(Y=0) + e_B  \\
b_h &= h_0 - h_0^r + e_B \text{\ if } h_0^r > e_B \text{, else }\infty \\
b_y &= p(Y=0)^\frac{1}{r} - p(Y=0) + e_B \\
&\text{\ if } p(Y=0) > e_B \text{, else }\infty \\
\end{align}
\begin{proof}
    The bounds given by $b_h$ follows from Theorem~\ref{bound}, and $b_y$ as described in the text.
    If both  $h_0^r < e_B$ and $p(y=0)^r < e_B$, then $0 \leq ( h_0^r + p(y=0)^r - e_B) \leq 1$ and
    we can lower bound the $r$-th root as in the proof of Theorem~\ref{bound}. This yields the bound given
    by $b_g$.
\end{proof}
\end{theorema}

\begin{theorema}
If the instance labels are independent inside a bag given a hidden cause, the instance
error $e_I$ has the following lower bound:
\begin{multline*}
e_I \geq h_0 + p(y=0) - \\
2 \left (\frac{1}{2} (p(h_B(X)=0) + p(Y=0) - e_B) \right)^\frac{1}{r}
\end{multline*}
\begin{proof}
    This follows from taking the expected value with respect to $Z$ of \eqref{conditioned} and using Jensen's
    inequality.
\end{proof}
\end{theorema}

### 2013-08-02 partial_supervision/fullpaper.tex
copy *& paste everything

\documentclass{llncs} % For LaTeX2e \usepackage{amssymb,amsmath,array}
\usepackage[draft]{fixme} \usepackage{booktabs} \usepackage{graphicx}
\usepackage{refstyle} \usepackage[utf8]{inputenc} \usepackage{xspace}

\usepackage{amssymb,amsmath,array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{refstyle}
\usepackage[utf8]{inputenc}
\usepackage{xspace}
\usepackage{tabularx}
\usepackage{array}
\usepackage{microtype}
\usepackage[numbers]{natbib}
\makeatletter
\renewcommand\bibsection%
{
\section*{\refname
\@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}
}
\makeatother

\newcolumntype{R}{>{\raggedleft\arraybackslash}X}
\newcolumntype{C}{>{\centered\arraybackslash}X}

\newcommand{\fix}{\marginpar{FIX}} \newcommand{\new}{\marginpar{NEW}}
\newcommand\etal{~et.\,al.~}

\begin{document}
\mainmatter              % start of the contributions

\title{Multi-Instance Methods for Partially Supervised Image
Segmentation}

\author{Andreas M\"uller\thanks{This work was founded by the B-IT research
school.}  \and Sven Behnke\\ \email{amueller@ais.uni-bonn.de},
\email{behnke@cs.uni-bonn.de}} \institute{Autonomous Intelligent Systems\\
Department of Computer Science\\
University of Bonn\\ 53113 Bonn, Germany \\ }

\maketitle

\begin{abstract}
In this paper, we propose a new partially supervised multi-class image segmentation algorithm.
We focus on the multi-class, single-label setup, where each image is assigned
one of multiple classes. 
We formulate the problem of image segmentation as a multi-instance task on a given
set of overlapping candidate segments. Using these candidate segments, we solve the multi-instance, multi-class problem
using multi-instance kernels with an SVM\@. This computationally advantageous approach, which requires only convex optimization,
yields encouraging results on the challenging problem of partially supervised image segmentation.
\end{abstract}

\section{Introduction}
The task of multi-class image segmentation is to create a pixel-wise labeling of an input image
into regions belonging to one of several semantic classes.
Most algorithms for this setting work with strong supervision: a pixel-wise labeling
of training images. Methods that are used in this setting include random forests~\citep{schroff2008object} and support vector machines (SVM). Usually
the output of such algorithms is further processed by a conditional random field (CRF~\citep{ladicky2009associative,gonfaus2010harmony,jiang2009efficient}).
While these methods reach high accuracy, it is very time consuming to create
pixel-level ground truth for real-world applications.
This problem can be addressed in several ways:
the LabelMe effort~\citep{russell2008labelme} tries to use the ``wisdom of crowds'' to obtain human labelings.
Another possibility is to use only weak supervision, which is the approach we follow here.

In the weakly supervised setting, the ground
 truth for a given image is a list of semantic classes that occur in this image, instead of a pixel-level
labeling as in the strongly supervised setting.
Image-level labels are much easier to obtain, e.~g.\  through online image libraries such as flickr and facebook.

The task of multi-class segmentation is often split up in a segmentation
and a recognition part. Random forest methods often classify each pixel
separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
computational burden of classifying each pixel separately, but have
two drawbacks:
\begin{enumerate}
    \item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
    \item Segment boundaries are decided on the lowest level by generating
        superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
\end{enumerate}
In our approach, we work with a set of candidate segments, generated using constrained parametric min-cuts~\citep{carreira2010constrained}.
For each image, these segments are a set of overlapping, object-like regions, which serve
as candidates for object locations.

We formulate weakly supervised multi-class image segmentation as a multi-instance problem, based upon candidate segments.
In multi-instance learning~\citep{dietterich1997solving}, each training example is given as a multi-set of instances, called a bag.
Each instance is represented as a feature vector $x$ and a label $y$.
A bag is labeled positive if it contains at least one positive example, and negative otherwise.
During training, only the labels of the training bags, not of the instances inside the bags, are known.
The goal is to learn a classifier for unseen bags. 
Formally, let $\mathcal{X}$ be the set of instances. To simply notation, we assume that bags are simply sets, not multi-sets.
Then a bag is an element of the power set $2^\mathcal{X}$ and the task is
to learn a function
\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}  \end{equation}
from a set of training examples of the form $(X_i,y_i)$ with bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.
The $f_{MI}$ function stems from the so-called underlying concept, given by an (unknown) function
$f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
\begin{equation}\eqlabel{multi_instance}
f_{MI}(X)= \max_{x \in X} f_{I}(x).
\end{equation}

Sometimes, the goal of finding $f_{MI}$ is extended to finding labels not only
on bag-level but also for all the instances within a bag
\citep{liconvex2010,zha2008joint}, i.\,e.\ finding $f_{I}$.\\ 
Even though finding $f_{I}$ is sometimes included in the task statement, there
has been very little work that actually reported accuracy on instance label prediction. Part
of the reason for this might be that for many of the datasets used in
multi-instance learning no ground truth exists.

We look explicitly at accuracy on instance-level since we are interested in
actually segmenting images, not just classifying them. For multi-class image
segmentation, there are some hand-labeled datasets that provide ground truth
on pixel level. We use this ground truth to evaluate the performance of our
method. This approach does not exactly correspond to instance-level ground truth --
since the instances are segments, not pixels -- but relates to it closely.

In this work, we explore the 
application of multi-instance learning algorithms to the task of partially supervised image segmentation.
Multi-instance learning is a natural formulation for image classification and has been
successfully applied in this task~\citep{zhou2007multi}. We propose to go a step further and apply
multi-instance learning to the task of object-class segmentation in natural
images. To our knowledge, all previous
methods in the field use strong supervision, meaning manual pixel-wise annotation of
training images. This approach does not
scale to larger datasets, especially if one expects consistency and quality
in the segmentations.

\section{Related Work}
\subsection{Proposal Object Segments}\label{related_segments}
Most work on multi-class segmentation focuses on strong supervision on superpixel level. There is still little work on using candidate segments.
The method we use for generating candidate segments is Constraint Parametric Min-Cuts (CPMC) from \citet{carreira2010constrained}.
This method creates a wide variety of overlapping segments. Support vector regression (SVR) is trained on these segments
to estimate the overlap of segments with ground truth object-class labeling
from the Pascal VOC dataset~\citep{pascal}. This provides a ranking of candidate segments, according to how ``object-like'' they are, which
allows for selecting only a limited number of very object-like segments.
The method performed well on a variety of datasets.
A similar approach was investigated by \citet{endres2010category}.
%Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
%algorithm.

\subsection{Multi-Instance Methods}
Multi-instance learning was formally introduced in \citet{dietterich1997solving}.
Since then, many algorithms were proposed to solve the multi-instance learning problem
using many different approaches \citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
We will discuss only those that are relevant to this work.

\citet{gaertner2002multi} introduced the concept of a multi-instance
kernel on bags, defined in terms of a kernel on instances. 
The basic principle of multi-instance kernel is similar to a soft-max over instances in
each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. They show that the multi-instance kernel
is able to separate bags if and only if the original kernel on instances is able to separate the underlying concepts.
The method of \citet{gaertner2002multi} has a particular appeal in that it neatly transforms a multi-instance problem into a standard
classification problem by changing the kernel. The downside of this approach is that it does
not directly label instances, only bags.
 %and that it assumes that instances inside a bag have i.i.d.
%labels.

\citet{zhou2009multi} explicitly address non-i.i.d.\  labels, leading to an algorithm that can take
advantage of correlations inside bags. Computational costs of their algorithm does not scale well with
the number of instances, although a heuristic algorithm is proposed to overcome this restriction.
\citet{zhou2009multi} demonstrated only a slight advantage of their algorithm over the MI-kernel of \citet{gaertner2002multi}, so
we use the MI-kernel for better scalability.

\citet{liconvex2010} compute likelihood ratios for instances, giving a new convex formulation of the multi-instance problem.
Using these likelihood ratios, classification can be performed directly on the instances, provided an appropriate
threshold for classifying instances as positive is known.
We circumvent this problem by applying the same classifier to instances and bags, thereby obtaining hard class
decisions for each instance.

%\subsection{Multi Instance Multi Label (MIML) Methods}
%Zhou and Zhang~\citet{zhou2007multi} were some of the first to tackle the problem of multi-instance multi-label tasks.
%They proposed two methods, MIMLBoost and MIMLSvm. MIMLBoost transforms a MIML task into
%a multi-instance learning task by reformulating the problem of \eqref{miml} as
%\begin{equation} f_{MIL} \colon 2^\mathcal{X}\times \mathcal{Y} \rightarrow \{ -1, +1\}.\end{equation}
%This corresponds to learning a single multi-instance classifier per class.
%MIMLSvm, on the other hand, transforms a MIML task into a multi-label task by constructing a mapping
%\begin{equation} \phi \colon 2^\mathcal{X} \rightarrow \mathcal{Z} \end{equation}
%into a new instance space $\mathcal{Z}$ by medioid clustering.
%Our approach fits well into the category of MIMLBoost: we also transform the MIML task
%into a multi-instance task that we then solve using MI-kernels.
%While our problem formulation is similar to Zhou and Zhang, we aim not only at image classification but
%also at object segmentation.

%Zha\etal\citet{zha2008joint} use a Hidden Conditional Random Field (HCRF), directly capturing the semantics of each instance.
%The HCRF model used is a complex probabilistic model that in practice can only be trained using contrastive divergence.
%For inference, Gibbs sampling is used, making recall very time consuming.
%Nguyen~\citet{nguyen2010new} proposed an iterative algorithm for MIML that alternates between phases of label propagation, in which
%labels are assigned to instances, and margin-maximization, in which feature weights are adjusted.

%The algorithms of Zha and Nguyen have the benefit of explicitly exploiting correlations between labels.
%It is noteworthy, however, that both rely heavily on approximate optimizations. We concentrate our efforts
%on methods that are faster and scale better with the number of instances and bags.

%In our work, we demonstrate how a computationally efficient algorithm combined with strong object segmentation hypotheses can efficiently solve MIML
%tasks as they occur in weakly supervised multi-class image segmentation, using only a single convex optimization.

\subsection{Semantic Scene Segmentation via Multi-Instance Learning}
Recently, several methods have been proposed to obtain semantic segmentations of images using only
image-level supervision~\citep{vezhnevets2011weakly,verbeek2007region,vezhnevets2010towards}. \citet{vezhnevets2011weakly},
for example, report impressive results on the MSRC dataset.

While semantic segmentation is closely related to multi-class image segmentation, there are important
distinctions: In semantic segmentation, each pixel has a semantic annotation, also containing ``background''
classes like ``sky'', ``grass'' and ``water''. In multi-class image segmentation, the focus is on
objects, with possibly large parts of the image being labeled as unspecific ``background''.
The unspecific background class contains much more clutter than for example ``grass'' and is therefore
much harder to model.
This makes disseminating the interesting part in multi-class object recognition challenging, since it
is not necessary possible to identify non-object regions easily.

\section{Multi-Instance Kernels for Image Segmentation}
\subsection{Constraint Parametric Min Cuts (CPMC)}
To generate proposal segments, we use the CPCM framework from \citet{carreira2010constrained}. We construct initial
segments using graph cuts, on the image graph. The energy function for these cuts uses pixel
color and the response of the global probability of boundary (gPb) detector~\citep{maire2008using}. As much as ten thousand initial segments are generated from foreground
and background seeds. A fast rejection based on segment size and ratio cut~\citep{wang2003image} reduced these to about 2000 overlapping
segments per image. Then, the segments are ranked according to a measure of object-likeness that is based on region and Gestalt properties.
This ranking is computed using an SVR model~\citep{carreira2010constrained}, which is available online.
For computing the global probability of boundary (gPb), we used the CUDA implementation
of \citet{catanzaro2009efficient}, instead of the original one, for speed.

\subsection{Multi-Instance Learning using MI-Kernels}
Since scalability is very important in real-world computer vision
applications, and natural images might need hundreds of segments to
account for all possible object boundaries, we use the efficient
multi-instance kernels~\citep{gaertner2002multi}.
Multi-instance kernels are a form of set kernels that transform a kernel
on instance level to a kernel on bag level.
We reduce the multi-instance multi-class problem to a multi-instance problem by using the one-vs-all approach.

With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
as
\begin{equation}\eqlabel{mi_kernel}
k_{MI}(X,X') := \sum_{x \in X, x' \in X'} k^p(x,x'),
\end{equation}
where $p \in \mathbb{N}$ is a parameter~\citep{gaertner2002multi}. As we use the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
are again RBF-kernels, we will not consider $p$ in the following.

We normalize the kernel $k_{MI}$ \citep{gaertner2002multi} using
\begin{equation}
k(X,X') := \frac{k_{MI}(X,X')}{\sqrt{k_{MI}(X,X)k_{MI}(X',X')}}.
\end{equation}
Training an SVM with this kernel produces a bag-level classifier for each class, which we will refer to as MIK.
This procedure is very efficient since the resulting kernel matrix is of size
number of bags, which is much smaller than a kernel matrix of size number of
instances, as is commonly used in the
literature~\citep{andrews2003support,nguyen2010new,zhang2008m3miml}.  Another
advantage over other methods is, that it uses a single convex optimization,
whereas other approaches often use iterative algorithms~\citep{andrews2003support} or need to fit complex
probabilistic models~\citep{zha2008joint}.

While using MIK has many advantages, it produces only an instance-level
classifier. We propose to transform a bag-level classifier $f_{MI}$ as given by
the SVM and \eqref{mi_kernel} into an instance-level classifier by setting
$f_{I}(x):=f_{MI}(\{x\})$, in other words, by considering each instance as its own
bag.

\subsection{Segment Features}
To describe single segments, we make use of densely computed SIFT~\citep{lowe2004distinctive} and ColorSIFT~\citep{van2009evaluating} features, from which
we compute bag of visual word histograms. Additionally, we use histograms of oriented gradients~\citep{dalal2005histograms} on the segments.
We use RBF-kernels for all of the features,
constructing one MI-kernel per feature. These are then combined using multiple kernel
learning to produce a single kernel matrix. This kernel matrix can then be used for all classes,
making classification particularly efficient.

\subsection{Combining Segments}\seclabel{combiningsegments}
The framework described above yields an image-level and a segment-level classification.
In our setup, each segment might be given multiple labels. To obtain a pixel-level object-class segmentation, we have to combine these. 
When building the segmentation for a given image, we only consider classes whose presence was predicted on image level.
Since we do not make use of the ground truth segmentation during training, we cannot learn an optimal combination as 
in \citet{li2010object} but perform a simple majority vote instead.
We merge segments into pixel-level class labels by setting
the label $y_x$ of a pixel $x$ according to
\begin{equation}
    y_x = \text{argmax}_{y \in Y} \#\{S_i | p \in S_i \land y_{S_i}=y \},
\end{equation}
where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an
image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
assigned the class with the highest number of class segments containing it.
This simple heuristic yields good results in practice.

\begin{figure}[tbp]
	\begin{center}
        \includegraphics[height=35mm]{images/scheme-crop.pdf}
	\end{center}
        \caption{Overview of our method. See text for details.}
	\figlabel{results}
\end{figure}
\section{Experiments}
\subsection{Instance-Level Predictions using Multi-Instance Kernels}
To assess the validity of instance-level predictions using multi-instance kernels, we transform $f_{I}$
back to an instance-level classifier, using the multi-instance learning
assumption (\eqref{multi_instance}). We refer to these instance-based MIK predictions
as MIK-instance. In all experiments, the parameters of the MI-Kernel
and SVM are adjusted using MIK and then used with both MIK and MIK-instance.
This facilitates very fast parameter scans since MIK is very efficient to
compute. Note that we cannot adjust parameters using instance prediction error, as we assume
no instance labels to be known.

\begin{table}
    \centering
    \caption{Bag level performance of various MIL algorithms on the standard Musk
    datasets. All but MIK provide instance-level labeling. }
    \begin{tabularx}{0.9\linewidth}{@{\extracolsep{\fill}}lccccccc}
    \toprule
        & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-instance \\
    \cmidrule(rl){2-8}
    Musk1 & 87.9 &84.9 &  87.4 &  77.9     & 84.3 & 88.0& 88.0 \\
    Musk2 & 85.4 &84.8 &  83.6 &  84.3     & 90.5 & 89.3& 85.2 \\
    \bottomrule
    \end{tabularx}
    \tablabel{musk-acc}
\end{table}
We compared the performance of MIK, MIK-instance and state-of-the-art MI
methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
\tabref{musk-acc}. Results were obtained using 10-fold
cross-validation. While the computational complexity of MIK-instance is
very low compared to the other methods, it achieves competitive results.
Using instance-level labels results in a slight loss of accuracy of
MIK-instances, compared to MIK\@. This small degradation of performance is quite surprising, since the model was
not trained to provide any instance-level labels.

For multi-class image segmentation, it is beneficial to have a low witness
rate, i.\,e.\ only a few instances are assumed to be positive in a positive
bag. Since an object might not be very prominent in an image, only a fraction
of segments might correspond to the object.
\tabref{musk-witness} compares the witness rates of MIK-instance,
miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
datasets. MIK-instance is able to achieve similar accuracy
with much less witnesses than the other methods.  Note that Musk1 consists of
very small bags while Musk2 contains significantly larger bags, more similar to the
image/segment setup.
\begin{table}
    \centering
    \caption{MIL algorithms and the empirical witness rates of the
    classifiers.}
    \begin{tabularx}{.8\linewidth}{@{\extracolsep{\fill}}lcccc}
    \toprule
    & \multicolumn2c{Musk1}  & \multicolumn2c{Musk2}  \\
                &accuracy & witness-rate & accuracy & witness-rate  \\
    \cmidrule(rl){2-3}
    \cmidrule(rl){4-5}
    mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
    SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
    MIK-instance& 88.0          & 99\%                &  85.2          & 62.3\%\\
    \bottomrule
    \end{tabularx}
    \tablabel{musk-witness}
\end{table}

\subsection{Partially Supervised Image Segmentation on Graz 02}

We evaluate the performance of the proposed algorithm for object-class
segmentation on the challenging Graz-02 dataset.
This dataset contains 1096 images of three object classes, bike, car and person.
Each image may contain multiple instances of the same class.

%From each image, we extract overlapping object-like segments using constraint
%parametric min cuts (CPMC,\citet{carreira2010constrained}). We use the top 200 ranked segments and discard the
%rest.
%Each segment is described using SIFT~\citet{lowe2004distinctive} and
%ColorSIFT~\citet{van2009evaluating} features, from which we compute bag of
%visual word histograms as well as histograms of oriented
%gradients~\citet{dalal2005histograms}. We construct one MI-Kernel per
%feature which are combined using multiple kernel learning.
We adjusted parameters on a hold-out validation set using bag-level information and used the
training and test sets as given by the dataset.
It is straight-forward to extend the binary MIK method to the multi-class
setting using a one-vs-all strategy.
We train one MKL-SVM per class using MIK and predict class labels on segment level
using MIK-instance. If at least one SVM classifies a segment as positive,
it is associated with the most confident class. Otherwise, it is assigned
``background'' or no class.
This yields a classification of each segment into one of four classes: car, bike, person,
or background. We merge segments into pixel-level class labels as described
in \secref{combiningsegments}.

\begin{figure}[tbp]
	\begin{center}
        \includegraphics[width=28mm]{images/car1_img.png}\hspace*{0.7ex}
        \includegraphics[width=28mm]{images/car1_gt.png}\hspace*{0.7ex}
        \includegraphics[width=28mm]{images/car1_pos.png}\hspace*{0.7ex}
        \includegraphics[width=28mm]{images/car1_neg.png}\hspace*{0.7ex}\\
        \vspace{1mm}
        \includegraphics[width=28mm]{images/person1_img.png}\hspace*{0.7ex}
        \includegraphics[width=28mm]{images/person1_gt.png}\hspace*{0.7ex}
        \includegraphics[width=28mm]{images/person1_pos.png}\hspace*{0.7ex}
        \includegraphics[width=28mm]{images/person1_neg.png}\hspace*{0.7ex}
	\end{center}
        \vspace{-2mm}
        \caption{Qualitative results on on the Graz-02 dataset. Top: Results on
        category ``car''. Bottom: Results on category ``person''. From left to
        right: original image, ground truth segmentation, segment votes for
        correct class, segment votes against correct class (red many, blue few votes).}
	\figlabel{results}
        \vspace{-0mm}
\end{figure}

\begin{table}
    \centering
    \caption{Pixel-level accuracy on the Graz-02 dataset.}
    \begin{tabularx}{0.8\textwidth}{lRRR}
    \toprule
                & car & bike & person \\
    \cmidrule(l){2-2}
    \cmidrule(l){3-3}
    \cmidrule(l){4-4}
        This work&   0.30&  0.45&  0.43 \\
        Best strongly supervised approaches~\citep{fulkerson2009class,schulz2011}&   0.72&  0.72&  0.66 \\
    \bottomrule
    \end{tabularx}
    \tablabel{graz}
\end{table}
Per-class pixel accuracies are reported in \tabref{graz}; some qualitative results are shown in \figref{results}. The overall
accuracy on images labels, which is the task that was actually trained, is $90\%$.
While the performance of our multiple-instance based approach is far
from current methods that use pixel-level annotations, whose pixel-level accuracy is around $70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level,
it can serve as a baseline for research on weakly supervised methods for image segmentation.

\section{Conclusions}
We proposed an algorithm for object-class segmentation using only weak supervision based on
multiple-instance learning. In our approach each image is represented as a bag of object-like
proposal segments.

We described a way to extent bag level predictions made by the multi-instance
kernel method to instance level while remaining competitive with the state-of-the-art
in bag label prediction.

Finally we evaluated the proposed object-class segmentation method on the challenging
Graz02 dataset. While not reaching the performance of methods using strong supervision,
our result can work as a baseline for further research into weakly supervised object class
segmentation.

In future work, we plan to scale our approach to much larger image datasets. As much
more images with weak annotations are available than with pixel level segmentation,
we hope that we can improve upon the state-of-the-art in object-class segmentation
by making use of larger bodies of training images.

%\bibliographystyle{splncs03} 
\bibliographystyle{splncsnat} 
\bibliography{paper} \end{document}

### 2013-08-02 partial_supervision/paper.tex
copy *& paste everything

--- 2013-08-02 partial_supervision/fullpaper.tex
@@
-\documentclass{llncs} % For LaTeX2e \usepackage{amssymb,amsmath,array}
-\usepackage[draft]{fixme} \usepackage{booktabs} \usepackage{graphicx}
-\usepackage{refstyle} \usepackage[utf8]{inputenc} \usepackage{xspace}
+In this work, we explore the 
+application of multi-instance learning algorithms to the task of partially supervised image segmentation.
+Multi-Instance learning is a natural formulation for image classification and has been
+successfully applied in this task~\cite{zhou2007multi}. We propose to go a step further and apply
+multi-instance learning to the task of object class segmentation in natural
+images.  In object class segmentation, the goal is to create a pixel-wise
+labeling of an input image into one of several semantic
+classes.  Multi-class image segmentation receives much attention in the
+computer vision community at the moment. To our knowledge, all previous
+methods in the field use strong supervision, meaning manual pixel-wise annotation of
+training images. This approach does not
+scale to larger datasets, especially if one expects consistency and quality
+in the segmentations.
 
-\usepackage{amssymb,amsmath,array}
-\usepackage{booktabs}
-\usepackage{graphicx}
-\usepackage{refstyle}
-\usepackage[utf8]{inputenc}
-\usepackage{xspace}
-\usepackage{tabularx}
-\usepackage{array}
-\usepackage{microtype}
-\usepackage[numbers]{natbib}
-\makeatletter
-\renewcommand\bibsection%
-{
-\section*{\refname
-\@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}
-}
-\makeatother
+In this work, we investigate the use of multi-instance
+learning to obtain multi-class image segmentations using ground truth labeling
+only on the image level.
+We focus on the multi-class, single label setup, where each image is assigned
+one of multiple classes. We formulate multi-class image segmentation as
+a multi-instance learning problem by considering each image as a bag of
+overlapping candidate segments.
 
-\newcolumntype{R}{>{\raggedleft\arraybackslash}X}
-\newcolumntype{C}{>{\centered\arraybackslash}X}
-
-\newcommand{\fix}{\marginpar{FIX}} \newcommand{\new}{\marginpar{NEW}}
-\newcommand\etal{~et.\,al.~}
-
-\begin{document}
-\mainmatter              % start of the contributions
-
-\title{Multi-Instance Methods for Partially Supervised Image
-Segmentation}
-
-\author{Andreas M\"uller\thanks{This work was founded by the B-IT research
-school.}  \and Sven Behnke\\ \email{amueller@ais.uni-bonn.de},
-\email{behnke@cs.uni-bonn.de}} \institute{Autonomous Intelligent Systems\\
-Department of Computer Science\\
-University of Bonn\\ 53113 Bonn, Germany \\ }
-
-\maketitle
-
-\begin{abstract}
-In this paper, we propose a new partially supervised multi-class image segmentation algorithm.
-We focus on the multi-class, single-label setup, where each image is assigned
-one of multiple classes. 
-We formulate the problem of image segmentation as a multi-instance task on a given
-set of overlapping candidate segments. Using these candidate segments, we solve the multi-instance, multi-class problem
-using multi-instance kernels with an SVM\@. This computationally advantageous approach, which requires only convex optimization,
-yields encouraging results on the challenging problem of partially supervised image segmentation.
-\end{abstract}
-
-\section{Introduction}
-The task of multi-class image segmentation is to create a pixel-wise labeling of an input image
-into regions belonging to one of several semantic classes.
-Most algorithms for this setting work with strong supervision: a pixel-wise labeling
-of training images. Methods that are used in this setting include random forests~\citep{schroff2008object} and support vector machines (SVM). Usually
-the output of such algorithms is further processed by a conditional random field (CRF~\citep{ladicky2009associative,gonfaus2010harmony,jiang2009efficient}).
-While these methods reach high accuracy, it is very time consuming to create
-pixel-level ground truth for real-world applications.
-This problem can be addressed in several ways:
-the LabelMe effort~\citep{russell2008labelme} tries to use the ``wisdom of crowds'' to obtain human labelings.
-Another possibility is to use only weak supervision, which is the approach we follow here.
-
-In the weakly supervised setting, the ground
- truth for a given image is a list of semantic classes that occur in this image, instead of a pixel-level
-labeling as in the strongly supervised setting.
-Image-level labels are much easier to obtain, e.~g.\  through online image libraries such as flickr and facebook.
-
-The task of multi-class segmentation is often split up in a segmentation
-and a recognition part. Random forest methods often classify each pixel
-separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
-over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
-computational burden of classifying each pixel separately, but have
-two drawbacks:
-\begin{enumerate}
-    \item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
-    \item Segment boundaries are decided on the lowest level by generating
-        superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
-\end{enumerate}
-In our approach, we work with a set of candidate segments, generated using constrained parametric min-cuts~\citep{carreira2010constrained}.
-For each image, these segments are a set of overlapping, object-like regions, which serve
-as candidates for object locations.
-
-We formulate weakly supervised multi-class image segmentation as a multi-instance problem, based upon candidate segments.
-In multi-instance learning~\citep{dietterich1997solving}, each training example is given as a multi-set of instances, called a bag.
-Each instance is represented as a feature vector $x$ and a label $y$.
-A bag is labeled positive if it contains at least one positive example, and negative otherwise.
-During training, only the labels of the training bags, not of the instances inside the bags, are known.
-The goal is to learn a classifier for unseen bags. 
-Formally, let $\mathcal{X}$ be the set of instances. To simply notation, we assume that bags are simply sets, not multi-sets.
-Then a bag is an element of the power set $2^\mathcal{X}$ and the task is
-to learn a function
-\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}  \end{equation}
-from a set of training examples of the form $(X_i,y_i)$ with bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.
-The $f_{MI}$ function stems from the so-called underlying concept, given by an (unknown) function
+The standard multi-instance setup is as follows. Let $\mathcal{X}$ be a set of
+instances. Then a bag is an element of the power set $2^\mathcal{X}$ and the
+task is to learn a function
+%\begin{equation}
+$f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}$
+%\end{equation}
+from a set of training examples of the form $(X_i,y_i)$ with
+bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.  This function
+stems from the so-called underlying concept, given by an (unknown) function
 $f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
+\vspace{-3mm}
 \begin{equation}\eqlabel{multi_instance}
 f_{MI}(X)= \max_{x \in X} f_{I}(x).
+\vspace{-3mm}
 \end{equation} 
-
+%In our case, instances are overlapping segments, obtained using constraint
+%parametric min-cuts~\cite{carreira2010constrained}.  Each image is a bag
+%consisting of hundreds of these segments.
 Sometimes, the goal of finding $f_{MI}$ is extended to finding labels not only
 on bag-level but also for all the instances within a bag
-\citep{liconvex2010,zha2008joint}, i.\,e.\ finding $f_{I}$.\\ 
+\cite{liconvex2010,zha2008joint}, i.\ e.\ finding $f_{I}$.\\ 
 Even though finding $f_{I}$ is sometimes included in the task statement, there
-has been very little work that actually reported accuracy on instance label prediction. Part
+has been very little work that actually reported accuracy on instance-level. Part
 of the reason for this might be that for many of the datasets used in
@@
 segmentation, there are some hand-labeled datasets that provide ground truth
-on pixel level. We use this ground truth to evaluate the performance of our
-method. This approach does not exactly correspond to instance-level ground truth --
+on pixel-level. We use this ground truth to evaluate the performance of our
+method. This does not exactly correspond to instance-level ground truth --
 since the instances are segments, not pixels -- but relates to it closely.
 
-In this work, we explore the 
-application of multi-instance learning algorithms to the task of partially supervised image segmentation.
-Multi-instance learning is a natural formulation for image classification and has been
-successfully applied in this task~\citep{zhou2007multi}. We propose to go a step further and apply
-multi-instance learning to the task of object-class segmentation in natural
-images. To our knowledge, all previous
-methods in the field use strong supervision, meaning manual pixel-wise annotation of
-training images. This approach does not
-scale to larger datasets, especially if one expects consistency and quality
-in the segmentations.
-
-\section{Related Work}
-\subsection{Proposal Object Segments}\label{related_segments}
-Most work on multi-class segmentation focuses on strong supervision on superpixel level. There is still little work on using candidate segments.
-The method we use for generating candidate segments is Constraint Parametric Min-Cuts (CPMC) from \citet{carreira2010constrained}.
-This method creates a wide variety of overlapping segments. Support vector regression (SVR) is trained on these segments
-to estimate the overlap of segments with ground truth object-class labeling
-from the Pascal VOC dataset~\citep{pascal}. This provides a ranking of candidate segments, according to how ``object-like'' they are, which
-allows for selecting only a limited number of very object-like segments.
-The method performed well on a variety of datasets.
-A similar approach was investigated by \citet{endres2010category}.
-%Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
-%algorithm.
-
-\subsection{Multi-Instance Methods}
-Multi-instance learning was formally introduced in \citet{dietterich1997solving}.
-Since then, many algorithms were proposed to solve the multi-instance learning problem
-using many different approaches \citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
-We will discuss only those that are relevant to this work.
-
-\citet{gaertner2002multi} introduced the concept of a multi-instance
-kernel on bags, defined in terms of a kernel on instances. 
-The basic principle of multi-instance kernel is similar to a soft-max over instances in
-each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. They show that the multi-instance kernel
-is able to separate bags if and only if the original kernel on instances is able to separate the underlying concepts.
-The method of \citet{gaertner2002multi} has a particular appeal in that it neatly transforms a multi-instance problem into a standard
-classification problem by changing the kernel. The downside of this approach is that it does
-not directly label instances, only bags.
- %and that it assumes that instances inside a bag have i.i.d.
-%labels.
-
-\citet{zhou2009multi} explicitly address non-i.i.d.\  labels, leading to an algorithm that can take
-advantage of correlations inside bags. Computational costs of their algorithm does not scale well with
-the number of instances, although a heuristic algorithm is proposed to overcome this restriction.
-\citet{zhou2009multi} demonstrated only a slight advantage of their algorithm over the MI-kernel of \citet{gaertner2002multi}, so
-we use the MI-kernel for better scalability.
-
-\citet{liconvex2010} compute likelihood ratios for instances, giving a new convex formulation of the multi-instance problem.
-Using these likelihood ratios, classification can be performed directly on the instances, provided an appropriate
-threshold for classifying instances as positive is known.
-We circumvent this problem by applying the same classifier to instances and bags, thereby obtaining hard class
-decisions for each instance.
-
-%\subsection{Multi Instance Multi Label (MIML) Methods}
-%Zhou and Zhang~\citet{zhou2007multi} were some of the first to tackle the problem of multi-instance multi-label tasks.
-%They proposed two methods, MIMLBoost and MIMLSvm. MIMLBoost transforms a MIML task into
-%a multi-instance learning task by reformulating the problem of \eqref{miml} as
-%\begin{equation} f_{MIL} \colon 2^\mathcal{X}\times \mathcal{Y} \rightarrow \{ -1, +1\}.\end{equation}
-%This corresponds to learning a single multi-instance classifier per class.
-%MIMLSvm, on the other hand, transforms a MIML task into a multi-label task by constructing a mapping
-%\begin{equation} \phi \colon 2^\mathcal{X} \rightarrow \mathcal{Z} \end{equation}
-%into a new instance space $\mathcal{Z}$ by medioid clustering.
-%Our approach fits well into the category of MIMLBoost: we also transform the MIML task
-%into a multi-instance task that we then solve using MI-kernels.
-%While our problem formulation is similar to Zhou and Zhang, we aim not only at image classification but
-%also at object segmentation.
-
-%Zha\etal\citet{zha2008joint} use a Hidden Conditional Random Field (HCRF), directly capturing the semantics of each instance.
-%The HCRF model used is a complex probabilistic model that in practice can only be trained using contrastive divergence.
-%For inference, Gibbs sampling is used, making recall very time consuming.
-%Nguyen~\citet{nguyen2010new} proposed an iterative algorithm for MIML that alternates between phases of label propagation, in which
-%labels are assigned to instances, and margin-maximization, in which feature weights are adjusted.
-
-%The algorithms of Zha and Nguyen have the benefit of explicitly exploiting correlations between labels.
-%It is noteworthy, however, that both rely heavily on approximate optimizations. We concentrate our efforts
-%on methods that are faster and scale better with the number of instances and bags.
-
-%In our work, we demonstrate how a computationally efficient algorithm combined with strong object segmentation hypotheses can efficiently solve MIML
-%tasks as they occur in weakly supervised multi-class image segmentation, using only a single convex optimization.
-
-\subsection{Semantic Scene Segmentation via Multi-Instance Learning}
-Recently, several methods have been proposed to obtain semantic segmentations of images using only
-image-level supervision~\citep{vezhnevets2011weakly,verbeek2007region,vezhnevets2010towards}. \citet{vezhnevets2011weakly},
-for example, report impressive results on the MSRC dataset.
-
-While semantic segmentation is closely related to multi-class image segmentation, there are important
-distinctions: In semantic segmentation, each pixel has a semantic annotation, also containing ``background''
-classes like ``sky'', ``grass'' and ``water''. In multi-class image segmentation, the focus is on
-objects, with possibly large parts of the image being labeled as unspecific ``background''.
-The unspecific background class contains much more clutter than for example ``grass'' and is therefore
-much harder to model.
-This makes disseminating the interesting part in multi-class object recognition challenging, since it
-is not necessary possible to identify non-object regions easily.
-
-\section{Multi-Instance Kernels for Image Segmentation}
-\subsection{Constraint Parametric Min Cuts (CPMC)}
-To generate proposal segments, we use the CPCM framework from \citet{carreira2010constrained}. We construct initial
-segments using graph cuts, on the image graph. The energy function for these cuts uses pixel
-color and the response of the global probability of boundary (gPb) detector~\citep{maire2008using}. As much as ten thousand initial segments are generated from foreground
-and background seeds. A fast rejection based on segment size and ratio cut~\citep{wang2003image} reduced these to about 2000 overlapping
-segments per image. Then, the segments are ranked according to a measure of object-likeness that is based on region and Gestalt properties.
-This ranking is computed using an SVR model~\citep{carreira2010constrained}, which is available online.
-For computing the global probability of boundary (gPb), we used the CUDA implementation
-of \citet{catanzaro2009efficient}, instead of the original one, for speed.
-
-\subsection{Multi-Instance Learning using MI-Kernels}
 Since scalability is very important in real-world computer vision
@@
 account for all possible object boundaries, we use the efficient
-multi-instance kernels~\citep{gaertner2002multi}.
+multi-instance kernels~\cite{graetner2002multi}.
 Multi-instance kernels are a form of set kernels that transform a kernel
-on instance level to a kernel on bag level.
-We reduce the multi-instance multi-class problem to a multi-instance problem by using the one-vs-all approach.
+on instance-level to a kernel on bag level.
+With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the
+corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
+as 
+\vspace{-2mm}
+\begin{equation}
+    \eqlabel{mi_kernel}
+    k_{MI}(X,X') := \sum_{x \in X, x' \in X'} k^p(x,x'),
+\vspace{-2.5mm}
+\end{equation}
+where $p \in \mathbb{N}$ is a parameter.  As we use the RBF-kernel
+$k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels are again
+RBF-kernels, we will not consider $p$ in the following.
+We normalize the kernel $k_{MI}$ in feature space.
 
-With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
-as
-\begin{equation}\eqlabel{mi_kernel}
-k_{MI}(X,X') := \sum_{x \in X, x' \in X'} k^p(x,x'),
-\end{equation}
-where $p \in \mathbb{N}$ is a parameter~\citep{gaertner2002multi}. As we use the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
-are again RBF-kernels, we will not consider $p$ in the following.
-
-We normalize the kernel $k_{MI}$ \citep{gaertner2002multi} using
-\begin{equation}
-k(X,X') := \frac{k_{MI}(X,X')}{\sqrt{k_{MI}(X,X)k_{MI}(X',X')}}.
-\end{equation}
+%\begin{equation}
+    %k(X,X') := \frac{k_{MI}(X,X')}{\sqrt{k_{MI}(X,X)k_{MI}(X',X')}}.
+%\end{equation}
 Training an SVM with this kernel produces a bag-level classifier for each class, which we will refer to as MIK.
-This procedure is very efficient since the resulting kernel matrix is of size
-number of bags, which is much smaller than a kernel matrix of size number of
+This procedure is very efficient since the resulting Gram matrix is of size
+number of bags, which is much smaller than a Gram matrix of size number of
 instances, as is commonly used in the
-literature~\citep{andrews2003support,nguyen2010new,zhang2008m3miml}.  Another
+literature~\cite{andrews2003support,nguyen2010new,zhang2008m3miml}.  Another
 advantage over other methods is, that it uses a single convex optimization,
-whereas other approaches often use iterative algorithms~\citep{andrews2003support} or need to fit complex
-probabilistic models~\citep{zha2008joint}.
+whereas other approaches often use iterative algorithms~\cite{andrews2003support} or need to fit complex
+probabilistic models~\cite{zha2008joint}.
 
 While using MIK has many advantages, it produces only an instance-level
-classifier. We propose to transform a bag-level classifier $f_{MI}$ as given by
+classifier. We propose to transform a bag lavel classifier $f_{MI}$ as given by
 the SVM and \eqref{mi_kernel} into an instance-level classifier by setting
@@
 
-\subsection{Segment Features}
-To describe single segments, we make use of densely computed SIFT~\citep{lowe2004distinctive} and ColorSIFT~\citep{van2009evaluating} features, from which
-we compute bag of visual word histograms. Additionally, we use histograms of oriented gradients~\citep{dalal2005histograms} on the segments.
-We use RBF-kernels for all of the features,
-constructing one MI-kernel per feature. These are then combined using multiple kernel
-learning to produce a single kernel matrix. This kernel matrix can then be used for all classes,
-making classification particularly efficient.
-
-\subsection{Combining Segments}\seclabel{combiningsegments}
-The framework described above yields an image-level and a segment-level classification.
-In our setup, each segment might be given multiple labels. To obtain a pixel-level object-class segmentation, we have to combine these. 
-When building the segmentation for a given image, we only consider classes whose presence was predicted on image level.
-Since we do not make use of the ground truth segmentation during training, we cannot learn an optimal combination as 
-in \citet{li2010object} but perform a simple majority vote instead.
-We merge segments into pixel-level class labels by setting
-the label $y_x$ of a pixel $x$ according to
-\begin{equation}
-    y_x = \text{argmax}_{y \in Y} \#\{S_i | p \in S_i \land y_{S_i}=y \},
-\end{equation}
-where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an
-image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
-assigned the class with the highest number of class segments containing it.
-This simple heuristic yields good results in practice.
-
-\begin{figure}[tbp]
-	\begin{center}
-        \includegraphics[height=35mm]{images/scheme-crop.pdf}
-	\end{center}
-        \caption{Overview of our method. See text for details.}
-	\figlabel{results}
-\end{figure}
-\section{Experiments}
-\subsection{Instance-Level Predictions using Multi-Instance Kernels}
-To assess the validity of instance-level predictions using multi-instance kernels, we transform $f_{I}$
+To assess the validity of predictions made by $f_{I}$, we transform it
 back to an instance-level classifier, using the multi-instance learning
-assumption (\eqref{multi_instance}). We refer to these instance-based MIK predictions
+assumption \eqref{multi_instance}. We refer to these instance-based MIK predictions
 as MIK-instance. In all experiments, the parameters of the MI-Kernel
@@
 This facilitates very fast parameter scans since MIK is very efficient to
-compute. Note that we cannot adjust parameters using instance prediction error, as we assume
-no instance labels to be known.
+compute.
 
@@
     \centering
-    \caption{Bag level performance of various MIL algorithms on the standard Musk
-    datasets. All but MIK provide instance-level labeling. }
+    \vspace{-4mm}
     \begin{tabularx}{0.9\linewidth}{@{\extracolsep{\fill}}lccccccc}
@@
     \end{tabularx}
+    \vspace{1mm}
+    \caption{Bag level performance of various MIL algorithms on the standard Musk
+    datasets. All but MIK provide instance-level labeling. }
+    \vspace{-8mm}
     \tablabel{musk-acc}
@@
 We compared the performance of MIK, MIK-instance and state-of-the-art MI
-methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
+methods on the Musk benchmark datasets~\cite{dietterich1997solving}, see
 \tabref{musk-acc}. Results were obtained using 10-fold
@@
 For multi-class image segmentation, it is beneficial to have a low witness
-rate, i.\,e.\ only a few instances are assumed to be positive in a positive
+rate, i.\ e.\ only a few instances are assumed to be positive in a positive
 bag. Since an object might not be very prominent in an image, only a fraction
@@
 \tabref{musk-witness} compares the witness rates of MIK-instance,
-miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
+miSVM~\cite{andrews2003support} and SVR-SVM~\cite{liconvex2010} on the Musk
 datasets. MIK-instance is able to achieve similar accuracy
@@
     \centering
-    \caption{MIL algorithms and the empirical witness rates of the
-    classifiers.}
+    \vspace{-6mm}
     \begin{tabularx}{.8\linewidth}{@{\extracolsep{\fill}}lcccc}
@@
     \end{tabularx}
+    \vspace{1mm}
+    \caption{MIL algorithms and the empirical witness rates of the
+    classifiers.}
+    \vspace{-10mm}
     \tablabel{musk-witness}
 \end{table}
-
-\subsection{Partially Supervised Image Segmentation on Graz 02}
-
-We evaluate the performance of the proposed algorithm for object-class
+%The MIK-instance method can be extended to the multi-class setup using a
+%straight-forward 1-vs-all strategy.
+We evaluate the performance of the proposed algorithm for image class
 segmentation on the challenging Graz-02 dataset.
@@
 Each image may contain multiple instances of the same class.
-
-%From each image, we extract overlapping object-like segments using constraint
-%parametric min cuts (CPMC,\citet{carreira2010constrained}). We use the top 200 ranked segments and discard the
-%rest.
-%Each segment is described using SIFT~\citet{lowe2004distinctive} and
-%ColorSIFT~\citet{van2009evaluating} features, from which we compute bag of
-%visual word histograms as well as histograms of oriented
-%gradients~\citet{dalal2005histograms}. We construct one MI-Kernel per
-%feature which are combined using multiple kernel learning.
-We adjusted parameters on a hold-out validation set using bag-level information and used the
+From each image, we extract overlapping object-like segments using constraint
+parametric min cuts (CPMC,\cite{carreira2010constrained}). We use the top 200 ranked segments and discard the
+rest.
+Each segment is described using SIFT~\cite{lowe2004distinctive} and
+ColorSIFT~\cite{van2009evaluating} features, from which we compute bag of
+visual word histograms as well as histograms of oriented
+gradients~\cite{dalal2005histograms}. We construct one MI-Kernel per
+feature which are combined using multiple kernel learning.
+We adjusted parameters on a hold-out validation set and used the
 training and test sets as given by the dataset.
@@
 using MIK-instance. If at least one SVM classifies a segment as positive,
-it is associated with the most confident class. Otherwise, it is assigned
+it is associated with the most confident class. Otherwise it is assigned
 ``background'' or no class.
-This yields a classification of each segment into one of four classes: car, bike, person,
-or background. We merge segments into pixel-level class labels as described
-in \secref{combiningsegments}.
+This yields a classification of each segment into one of car, bike, person
+or background. We merge segments into pixel-level class labels by setting
+the label $y_x$ of a pixel $x$ according to
+\vspace{-3mm}
+\begin{equation}
+    y_x = \text{argmax}_{y \in Y} \frac{\#\{S_i | p \in S_i \land y_{S_i}=y \}}{
+    \#\{S_i | p \in S_i \}},
+\vspace{-3mm}
+\end{equation}
+where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an
+image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
+assigned the class with the highest ratio of class segments vs non-class
+segments containing it.
 
+\begin{table}
+    \centering
+    \vspace{-6mm}
+    \begin{tabularx}{0.8\textwidth}{lRRR}
+    \toprule
+                & car & bike & person \\
+    \cmidrule(l){2-2}
+    \cmidrule(l){3-3}
+    \cmidrule(l){4-4}
+        Pixel-level accuracy&   19.1\%&  27.7\%&  24.6\% \\
+    \bottomrule
+    \end{tabularx}
+    \vspace{1mm}
+    \caption{Pixel-level accuracy on the Graz-02 dataset.}
+    \vspace{-8mm}
+    \tablabel{graz}
+\end{table}
+If the ratio $\frac{\#\{S_i | p \in S_i \land y(S_i)=y \}}{\#\{S_i | p \in S_i \}}$ 
+is smaller than $0.5$ for all labels, i.\ e.\ for all labels there are more
+non-class segments than class segments at a given pixel, the pixel
+is assigned to the background.
+Per-class pixel accuracies are reported in \tabref{graz}, some qualitiative results are shown in \figref{results}. The overall
+accuracy on images labels, which is the task that was actually trained, is $87\%$.
+While the performance of our multiple-instance based approach is far
+from current methods that use pixel-level annotations, whose pixel-level accuracy is around $70\%$~\cite{fulkerson2009class} on pixel-level,
+it can serve as a baseline for research on weakly supervised methods for image segmentation..
 \begin{figure}[tbp]
@@
 	\end{center}
-        \vspace{-2mm}
-        \caption{Qualitative results on on the Graz-02 dataset. Top: Results on
+        \vspace{-7mm}
+        \caption{Qualitative resultson on the Graz-02 dataset. Top: Results on
         category ``car''. Bottom: Results on category ``person''. From left to
@@
 	\figlabel{results}
-        \vspace{-0mm}
+        \vspace{-7mm}
 \end{figure}
-
-\begin{table}
-    \centering
-    \caption{Pixel-level accuracy on the Graz-02 dataset.}
-    \begin{tabularx}{0.8\textwidth}{lRRR}
-    \toprule
-                & car & bike & person \\
-    \cmidrule(l){2-2}
-    \cmidrule(l){3-3}
-    \cmidrule(l){4-4}
-        This work&   0.30&  0.45&  0.43 \\
-        Best strongly supervised approaches~\citep{fulkerson2009class,schulz2011}&   0.72&  0.72&  0.66 \\
-    \bottomrule
-    \end{tabularx}
-    \tablabel{graz}
-\end{table}
-Per-class pixel accuracies are reported in \tabref{graz}; some qualitative results are shown in \figref{results}. The overall
-accuracy on images labels, which is the task that was actually trained, is $90\%$.
-While the performance of our multiple-instance based approach is far
-from current methods that use pixel-level annotations, whose pixel-level accuracy is around $70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level,
-it can serve as a baseline for research on weakly supervised methods for image segmentation.
-
-\section{Conclusions}
-We proposed an algorithm for object-class segmentation using only weak supervision based on
-multiple-instance learning. In our approach each image is represented as a bag of object-like
-proposal segments.
-
-We described a way to extent bag level predictions made by the multi-instance
-kernel method to instance level while remaining competitive with the state-of-the-art
-in bag label prediction.
-
-Finally we evaluated the proposed object-class segmentation method on the challenging
-Graz02 dataset. While not reaching the performance of methods using strong supervision,
-our result can work as a baseline for further research into weakly supervised object class
-segmentation.
-
-In future work, we plan to scale our approach to much larger image datasets. As much
-more images with weak annotations are available than with pixel level segmentation,
-we hope that we can improve upon the state-of-the-art in object-class segmentation
-by making use of larger bodies of training images.
-
-%\bibliographystyle{splncs03} 
-\bibliographystyle{splncsnat} 
-\bibliography{paper} \end{document}

### 2013-08-02 pystruct/paper.tex
copy *& paste everything

% Definitions of handy macros can go here

%\DeclareMathOperator*{\argmax}{arg\,max}
%\DeclareMathOperator*{\argmin}{arg\,min}

%\newcommand{\dataset}{{\cal D}}
%\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}

%\newcommand{\x}{\checkmark}
%\renewcommand{\o}{$\times$}
%\newcommand{\pystruct}{{\sc PyStruct} }

\section{Introduction}
In recent years there has been a wealth of research in methods for learning structured prediction,
as well as in applications such as natural language processing and computer vision.
Unfortunately only few implementations are publicly available, and many applications build upon
the non-free implementation of~\citet{joachims2009cutting}.

\pystruct on the other hand aims at providing a simple-to-use interface, in the high-level Python language.
This allows practitioners to efficiently test a range of models, as well as allowing researchers
to compare to baseline methods much more easily.
\pystruct is BSD-Licensed, allowing modification and redistribution of the code,
as well as use in commercial applications.  By embracing paradimes established
in the scientific python community, and reusing the interface of the
widely-used {\sc scikit-learn} library~\citep{pedregosa2011scikit}, \pystruct
can be used in existing projects, replacing standard classifiers.
The online documentation and examples help new
users understand the somewhat abstract ideas behind structured pediction.

%The organization of this paper is as follows. Section~\ref{sec:api} revisits
%the basic concepts of structured learning algorithms, and how these map to the
%\pystruct interface. Section~\ref{sec:examples} exemplifies the use of \pystruct
%on some standard structured prediction benchmarks and showcasts how \pystruct
%can interact with existing libraries, foremost scikit-learn and scikit-image.
%Section~\ref{sec:benchmarks} finally compares \pystruct with SVM$^\text{struct}$,
%arguably the most widely used implementation of structural SVMs, and
%SVM$^\text{python}$, a Python interface or the same.

\section{Structured Prediction and Casting it into Software}\label{sec:api}

Structured prediction can be defined as making a prediction $f(x)$ by maximizing an
compatibility function $g$ between an input $x$ and the possible labels
$y$~\cite{nowozin2011structured}. Most current approaches use linear functions,
leading to:
\begin{equation}\label{eq:main_equation}
    f(x) = \argmax_{y \in \mathcal{Y}} \langle \theta, \Psi(x, y) \rangle 
\end{equation}
Here, $y$ is a \emph{structured} label in the sense that it is more complicated than a single
output class. Examples include predicting a label for each word in a sentence, or predicting
a label for each pixel in an image.
Learning structured prediction means learning the parameters $\theta$ from training data
$(x_1, y_1), \dotsc, (x_n, y_n)$.

Using the above formulation learning can be broken down into three sub-problems:
\begin{itemize}
    \item Optimizing the objective with respect to $\theta$.
    \item Encoding the structure of the problem in a joint features function $\Psi$.
    \item Solving the maximization problem in Equation~\ref{eq:main_equation}.
\end{itemize}

The later two problems are usually tightly coupled, as the maximization in
Equation~\ref{eq:main_equation} is usually only feasible by exploiting the
structure of $\Psi$, while the first one is usually treated as independent.
\pystruct takes an object-oriented approach to decouple the task-dependent
implementation of 2) and 3) from the general algorithms used to solve 1).

Parameter estimation is done in \texttt{learner} classes, which currently support
cutting plane algorithms for structural SVMs (SSVMs), subgradient methods for SSVMs,
the structured perceptron and latent variable SSVMs.
Encoding the structure of the problem is done using \texttt{model} classes, that
compute $\Psi$, and encode the structure of the problem.
\pystruct implements models for many common cases, such as multi-class and
multi-label classification, conditional random fields with constant or
data-dependent pairwise potentials, and several latent variable models.
The maximization for finding $y$ in Equation~\ref{eq:main_equation} is carried out
using highly optimized implementations from external libraries. \pystruct
includes direct support for using {\sc OpenGM}, {\sc LibDAI}~\citep{Mooij_libDAI_10} and {\sc
AD$^3$}~\citep{martins2011augmented}. It also includes an interface to a
general purpose linear programming solver.

Table~\ref{table:comparision} lists implemented algorithms and models, and
compares them to other public structured prediction libraries, together with
the programming language and the project license.

\begin{table}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccccccc}
\toprule
Package &     Language &     License&\multicolumn{4}{c}{Algorithms}&\multicolumn{3}{c}{Models} \\
\cmidrule(r){1-1} \cmidrule(r){2-2} \cmidrule(r){3-3} \cmidrule(r){4-7} \cmidrule{8-10}
&             &&                     \footnotesize{CP}& \footnotesize{SG}& \footnotesize{LV}& \footnotesize{ML}& \footnotesize{Chain} & \footnotesize{Graph} & \footnotesize{LDCRF}\\
\sc{\pystruct}&      Python &       BSD            & \x    & \x      & \x   & \o & \x     & \x     & \x \\
SVM$^\text{\sc{struct}}$ & C++ & non-free         & \x    & \o      & \x   & \o & \o     & \o     & \o \\
\sc{Dlib}         & C++        & boost            & \x    & \x      & \o   & \o & \x     & \x     &\o\\
\sc{CRFsuite}     & C++        & BSD              & \o    & \o      & \o   & \x & \x     & \o     &\o\\

\bottomrule
\end{tabularx}
\vspace{-3mm}
    \caption{Comparison of structured prediction software packages. CP stands
for cutting plane optimization of SSVMs~\cite{joachims2009cutting}, SG for
online subgradient optimization of SSVMs~\cite{ratliff2007online}, LV for
latent variable SSVMs~\cite{yu2009learning}, ML for maximum likelihood
training, Chain for chain-structured models
with pairwise interactionsr, Graph for arbitrary graphs with pairwise
interactions, and LDCRF for latent dynamic CRF~\citep{morency2007latent}.\label{table:comparision}}
\end{table}

\section{Project Goals}\label{sec:goals}

\paragraph{Completeness}\pystruct aims at providing full pipelines that can be
    used in applications. It contains model formulation for many typical
    applications.  This is in contrast to SVM$^\text{struct}$ that provides no
    models at all, requiring the user to develop siginificant amounts of code, even
    for simple tasks.

\paragraph{Modularity} \pystruct separates the algorithms of parameter estimation and
     inference from the task-dependent formulation of $\Psi$. This allows
     practictioners, for example in computer vision or natural language
     processing, to improve their model without changing any optimization
     code. On the other hand, researchers working on better inference or
     parameter learning can easily benchmark their improvements on a wide
     array of applications.

\paragraph{Efficiency}
     While \pystruct focuses on usability, providing efficient and competative
     implementations is important to allow fast prototyping and scaling to
     large datasets.

\paragraph{Documentation and Examples}
     \pystruct provides full documentation of all classes and functions.  It
     also provides examples for many important applications, such as
     sequence tagging, multi-label classification and image segmentation.
     Furthermore, standard benchmarks are included as examples, which allows
     easy comparision with other approaches.

\paragraph{Testing}
     \pystruct contains a testing-suite with 80\% line-coverage. It also employs continuous integration
     to ensure stability and seamless user experience.

\paragraph{Integration}
     To improve usability, \pystruct is interoperable with other numeric and scientific python projects,
     such as {\sc scikit-learn}~\citep{pedregosa2011scikit},
     {\sc mahotas}~\citep{coelho:mahotas}, {\sc gensim}~\citep{rehurek_lrec} and {\sc scikit-image}.
     This allows users to build powerful applications with little effort. In
     particular, most of the model-selection methods of scikit-learn can be used
     directly with \pystruct.

\section{Usage Example: Semantic Image Segmentation}\label{sec:examples}
%\subsection{Multi-class classification}
%Multi-class classification is a very simple case of structured prediction, that serves
%well to exemplify the usage. We load the iris dataset, and compare the usage of \pystruct
%with a call to LibLinears Crammer-Singer SVM using scikit-learn.
%\lstinputlisting[language=Python, frame=single, basicstyle=\scriptsize,
    %numbers=left, commentstyle=\color{blue}\normalfont,
%columns=flexible, firstline=11]{multi_class_example.py}

%\subsection{OCR sequence classification}
%Chain structured CRFs are commonly used in natural language processif, for
%example for part-of-speech tagging or named entity recognition.  Here we
%demonstrate the use of a chain CRF on the OCR dataset from. The ChainCRF provides
%a very simnple interface for this problem. Each example is presented as an
%array of input features of shape % (length_of_sequence, number_of_features).
%% listing, training time, accuracy, comparision with kraehenbuehl

\subsection{Semantic Image Segmentation}
Conditional random fields are an important tool for semantic image segmentation.
We show how to train an $n$-slack support vector machines on a superpixel-based CRF
on the popular Pascal dataset. We use unary potentials generated using TextonBoost
from \cite{krahenbuhl2012efficient}. The superpixels are generated using SLIC.%
\footnote{The preprocessed data can be downloaded at \dots.}
Each sample is represented as a tuple consisting of input features and a graph representation.

\lstinputlisting[language=Python, frame=single, basicstyle=\scriptsize,
    numbers=left, commentstyle=\color{blue}\normalfont,
columns=flexible,firstline=32, lastline=37]{pascal_crf.py}
Lines 1-3 declare a quite complex model pairwise model on an arbitray graph.
Here \texttt{class\_weight} reweights the hamming loss according to class
frequencies. The parametric pairwise interactions have three features: a
constant feature, color similarity, and relative vertical position. The first
are declared to be symmetric with respect to the direction of an edge, the last
is antisymmetric. The inference method used is QPBO-fustion moves.  Line 5
creates an \texttt{learner} object that will learn the parameters for the given
model using the n-slack cutting plane method, and line 6 performs the actual
learning.  Using this simple setup, we achieve an accuracy of 30.3 on the
valiation set following the setup of \citet{krahenbuhl2012efficient}, who
report 30.2 using a more complex approach. Training the structured model takes
approximately 30 minutes using a single i7 core (using more would be possible).

%TODO not here
%This makes it very easy to incoporate features or segments generated by computer vision libraries such
%as scikit-image or mahotas, and to plot results using visualization tools like matplotlib. 
% listing, training time, accuracy, comparision with kraehenbuehl

\section{Experiments}\label{sec:benchmarks}
In this section, we compare our implementation with SVM$~\text{multiclass}$ and
SVM$~\text{python}$ with respect to run-time and optimzation accuracy.
While SVM$~\text{python}$ allows integration with Python programs, it does not use native
Python data structures, such as Numpy arrays. This makes interacting with existing software
cumbersome and inefficient, which is reflected in the run-time comparision below.
We also compare against the special-purpose implementation of the Crammer-Singer multi-class SVM in
{\sc LibLinear}~\citep{fan2008liblinear}.

We do not provide comparisions on more complex problems--such as ones including
non-trivial inference--as the authors of SVM$^\text{struct}$ did not provide
any code or examples for this case.

MNIST, covertype?

% multiclass on mnist? liblinear seems to die!!!!

\section{Conclusion}
This paper introduced \pystruct, a modular structured learning and prediction library.
\pystruct is geared towards ease of use, while providing efficient implementations.
\pystruct integrates itself into the scientific Python eco-system, making it easy to use with
existing libraries and applications.
Currently \pystruct focuses on max-margin and perceptron based approaches. In the future,
we plan to integrate other paradigms, such as SampleRank, pseudo-likelihood learning,
Conditional Neural Fields and approaches that allow for a better integration of inference
and learning.

### 2013-08-02 pystruct/multi_class_example.py

@@ -1,0 +1,18 @@
+from sklearn.datasets import load_iris
+from sklearn.svm import LinearSVC
+
+from pystruct.models import CrammerSingerSVMModel
+from pystruct.learners import OneSlackSSVM
+
+# Load three class iris data.
+iris = load_iris()
+X, y = iris.data, iris.target
+
+# PyStruct interface
+model = CrammerSingerSVMModel()
+one_slack_svm = OneSlackSSVM(model)
+one_slack_svm.fit(X, y)
+
+# scikit-learn interface for liblinear
+libsvm = LinearSVC(multi_class='crammer_singer')
+libsvm.fit(X, y)

### 2013-08-02 pystruct/pascal_crf.py

@@ -1,0 +1,40 @@
+import numpy as np
+
+from pystruct import learners
+import pystruct.models as crfs
+
+from datasets.pascal import PascalSegmentation
+from pascal_helpers import load_pascal
+from latent_crf_experiments.utils import (discard_void, add_edges,
+                                          add_edge_features)
+
+
+from IPython.core.debugger import Tracer
+tracer = Tracer()
+
+
+def main():
+    ds = PascalSegmentation()
+    # load training data
+    edge_type = "pairwise"
+    which = "kTrain"
+    data_train = load_pascal(which=which, sp_type="cpmc")
+
+    data_train = add_edges(data_train, edge_type)
+    data_train = add_edge_features(ds, data_train)
+    data_train = discard_void(ds, data_train, ds.void_label)
+
+    X, Y = data_train.X, data_train.Y
+
+    class_weights = 1. / np.bincount(np.hstack(Y))
+    class_weights *= 21. / np.sum(class_weights)
+
+    model = crfs.EdgeFeatureGraphCRF(
+        class_weight=class_weights, symmetric_edge_features=[0, 1],
+        antisymmetric_edge_features=[2], inference_method='qpbo')
+
+    ssvm = learners.NSlackSSVM(model, C=0.01, n_jobs=-1)
+    ssvm.fit(X, Y)
+
+if __name__ == "__main__":
+    main()

### 2013-08-02 partial_supervision/paper.tex
minor fixes, images

--- 2013-08-02 partial_supervision/fullpaper.tex
@@
-\documentclass{llncs} % For LaTeX2e \usepackage{amssymb,amsmath,array}
-\usepackage[draft]{fixme} \usepackage{booktabs} \usepackage{graphicx}
-\usepackage{refstyle} \usepackage[utf8]{inputenc} \usepackage{xspace}
-
-\usepackage{amssymb,amsmath,array}
-\usepackage{booktabs}
-\usepackage{graphicx}
-\usepackage{refstyle}
-\usepackage[utf8]{inputenc}
-\usepackage{xspace}
-\usepackage{tabularx}
-\usepackage{array}
-\usepackage{microtype}
-\usepackage[numbers]{natbib}
-\makeatletter
-\renewcommand\bibsection%
-{
-\section*{\refname
-\@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}
-}
-\makeatother
-
-\newcolumntype{R}{>{\raggedleft\arraybackslash}X}
-\newcolumntype{C}{>{\centered\arraybackslash}X}
-
-\newcommand{\fix}{\marginpar{FIX}} \newcommand{\new}{\marginpar{NEW}}
-\newcommand\etal{~et.\,al.~}
-
-\begin{document}
-\mainmatter              % start of the contributions
-
-\title{Multi-Instance Methods for Partially Supervised Image
-Segmentation}
-
-\author{Andreas M\"uller\thanks{This work was founded by the B-IT research
-school.}  \and Sven Behnke\\ \email{amueller@ais.uni-bonn.de},
-\email{behnke@cs.uni-bonn.de}} \institute{Autonomous Intelligent Systems\\
-Department of Computer Science\\
-University of Bonn\\ 53113 Bonn, Germany \\ }
-
-\maketitle
-
-\begin{abstract}
-In this paper, we propose a new partially supervised multi-class image segmentation algorithm.
-We focus on the multi-class, single-label setup, where each image is assigned
-one of multiple classes. 
-We formulate the problem of image segmentation as a multi-instance task on a given
-set of overlapping candidate segments. Using these candidate segments, we solve the multi-instance, multi-class problem
-using multi-instance kernels with an SVM\@. This computationally advantageous approach, which requires only convex optimization,
-yields encouraging results on the challenging problem of partially supervised image segmentation.
-\end{abstract}
-
 \section{Introduction}
@@
 	\end{center}
-        \vspace{-2mm}
         \caption{Qualitative results on on the Graz-02 dataset. Top: Results on
@@
 	\figlabel{results}
-        \vspace{-0mm}
 \end{figure}
@@
 by making use of larger bodies of training images.
-
-%\bibliographystyle{splncs03} 
-\bibliographystyle{splncsnat} 
-\bibliography{paper} \end{document}

### 2013-08-02 make.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+rubber -I itm/images -I mil/images -I partial_supervision -I mil -I exact_learning -I pystruct main.tex

### 2013-08-03 exact_learning/paper.tex
fixing references to equations and figures

@@
 the loss-based soft-margin objective
-\begin{equation}\label{mainequation}
+\begin{equation}\label{eq:mainequation}
 \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  r_i(\theta)
@@
 
-We solve the following reformulation of Equation~\ref{mainequation}, known as one-slack QP formulation:
-\begin{align}\label{oneslack}
+We solve the following reformulation of Equation~\eqref{mainequation}, known as one-slack QP formulation:
+\begin{align}\label{eq:oneslack}
     \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
@@
 Even using approximate inference procedures, several statements
-about the original exact objective (Equation~\ref{mainequation}) can be
+about the original exact objective (Equation~\eqref{mainequation}) can be
 obtained.
@@
 we solve the QP once with a constraint from the cache. If the resulting $o^C$ verifies
-\begin{equation}\label{cache_test}
+\begin{equation}\label{eq:cache_test}
     o^C - o^Q < \frac{1}{2} (o^Q - o_{\W})

### 2013-08-03 pystruct/paper.tex
minor fixes, images
use natbib citation commands everywehere.

@@
-% Definitions of handy macros can go here
-
-%\DeclareMathOperator*{\argmax}{arg\,max}
-%\DeclareMathOperator*{\argmin}{arg\,min}
-
-%\newcommand{\dataset}{{\cal D}}
-%\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
-
-%\newcommand{\x}{\checkmark}
-%\renewcommand{\o}{$\times$}
-%\newcommand{\pystruct}{{\sc PyStruct} }
-
 \section{Introduction}
@@
 compatibility function $g$ between an input $x$ and the possible labels
-$y$~\cite{nowozin2011structured}. Most current approaches use linear functions,
+$y$~\citep{nowozin2011structured}. Most current approaches use linear functions,
 leading to:
@@
 \end{tabularx}
-\vspace{-3mm}
     \caption{Comparison of structured prediction software packages. CP stands
-for cutting plane optimization of SSVMs~\cite{joachims2009cutting}, SG for
-online subgradient optimization of SSVMs~\cite{ratliff2007online}, LV for
-latent variable SSVMs~\cite{yu2009learning}, ML for maximum likelihood
+for cutting plane optimization of SSVMs~\citep{joachims2009cutting}, SG for
+online subgradient optimization of SSVMs~\citep{ratliff2007online}, LV for
+latent variable SSVMs~\citep{yu2009learning}, ML for maximum likelihood
 training, Chain for chain-structured models

### 2013-08-03 main.tex
minor fixes, images
fix bibliography
use natbib citation commands everywehere.
use garamond :)

@@
-\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{scrreprt}
+\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{scrbook}
 
@@
 \usepackage{calc}
+\usepackage{refstyle}
 \usepackage{amssymb}
@@
 \usepackage{tabularx}
+\usepackage{graphicx}
 \usepackage{algorithm}
@@
 \usepackage{setspace}
+\usepackage{color}
+\usepackage{listings}
+\usepackage[urw-garamond]{mathdesign}
+\usepackage{garamondx}
+%\usepackage{layout}
 
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
-\newcommand{\figref}[1]{Figure~\ref{#1}}
-\renewcommand{\eqref}[1]{Equation~(\ref{#1})}
+
+\newcommand{\x}{\checkmark}
+\renewcommand{\o}{$\times$}
+\newcommand{\pystruct}{{\sc PyStruct} }
+
 \algrenewcommand{\algorithmicrequire}{\textbf{Input:}}
 \algrenewcommand{\algorithmicensure}{\textbf{Output:}}
+
+\newtheorem{corollary}{Corollary}
+\newtheorem{corollarya}{Corollary}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+\newtheorem{theorema}{Theorem}
+
+\newcolumntype{R}{>{\raggedleft\arraybackslash}X}
+\newcolumntype{C}{>{\centered\arraybackslash}X}
+
+\newcommand{\fix}{\marginpar{FIX}} \newcommand{\new}{\marginpar{NEW}}
+\newcommand\etal{~et\,al.~}
 
@@
 \begin{document}
-
-%\title{Learning a Loopy Model For Semantic Segmentation Exactly}
+%\layout
 
@@
 \vspace*{1in}
-\textbf{{\LARGE PhD Thesis}}
+\textbf{{\LARGE Dissertation}}
 \par
-\vspace{1.5in} {\large Andreas Mueller}
+\vspace{1.5in} {\large Andreas M\"uller}
  \par \vfill A Thesis submitted for the degree of Doctor of Philosophy
 \par \vspace{0.5in}
-Department IV, Institute for Computer Science
+Institute for Computer Science \par
+Department IV, Autonomous Intelligent Systems
 \par \vspace{0.5in}
-Friedrich-Willhelm Universitaet Bonn \par
+Friedrich-Willhelm Universit\"at Bonn \par
 \vspace{0.5in} September 2013 \end{center}
@@
 
-\begin{abstract} abstract
-\end{abstract}
+\pagenumbering{arabic}
 
-\pagenumbering{arabic}
+\chapter{introduction}
 
@@
 
+\chapter{Conclusion}
+
+\bibliographystyle{plainnat}
 \bibliography{main}

### 2013-08-14 intro.tex
ideen fuer zusaetzliches material, intro datei / gliederung

\section{Overview}
What does this theses contribute?

Why semantic segmentation?
Why is it not solved yet?

What is semantic segmentation?
Datsets? Input?

### 2013-08-14 itm/paper.tex
fix bibliography
fixing references to equations and figures
update jmlr paper in thesis

@@
-\section{Introduction}
+Before going to segmentation, we look at clustering because:
+1 we see how far we can go without supervision
+2 we need it as building block for bow
+3 we need it as initialization for latent variable models
+%TODO elaborate
+
 Clustering data is one of the fundamental problems in machine learning.  In
@@
 we have $H_{est}(\mathbf{x}_y) \approx H(X \given Y {=} y)$, leading to the clustering problem
-\begin{align}\label{general_objective}
+\begin{align}\label{eq:general_objective}
      \max_{\mathbf{y}}\quad - \sum_{y=1}^k p(Y{=}y) H_\text{est}(\mathbf{x}_y),
@@
 More precisely, the entropy estimate of a dataset $\mathbf{x} = (x_1, \dotsc, x_n)$ is given by
-\begin{align}\label{hmst}
+\begin{align}\label{eq:hmst}
     H_{mst}(\mathbf{x}) = d \log(L) - (d-1) \log(n) + \log(\beta_d).
@@
 the problem to maximize $\hat{I}(\mathbf{x}, \mathbf{y})$ with
-\begin{align}\label{objective}
+\begin{align}\label{eq:objective}
      \hat{I}(\mathbf{x}, \mathbf{y}) := & -\sum_{y=0}^k p(y) \Bigl[ d \log(L_y) - (d-1) \log{n_y} \Bigr] + C,\\
     = & - \sum_{y=0}^k p(y) \Bigl[ d \log(\bar{L}_y  ) + \log{n_y} \Bigr] + C'\\
-    = & - d \sum_{y=0}^k p(y) \log(\bar{L}_y  ) - \sum_{y=0}^k p(y) \log{p(y) + C''} \label{objective_simplified}
+    = & - d \sum_{y=0}^k p(y) \log(\bar{L}_y  ) - \sum_{y=0}^k p(y) \log{p(y) + C''} \label{eq:objective_simplified}
 \end{align}
@@
 but form a trade-off between edge length and cluster size.}
-\label{illustration}
+\label{fig:illustration}
 \end{figure}
@@
 Define
-\setlength{\abovedisplayskip}{1mm}
-\setlength{\belowdisplayskip}{1mm}
-\begin{align}\label{graphobjective}
+%\setlength{\abovedisplayskip}{1mm}
+%\setlength{\belowdisplayskip}{1mm}
+\begin{align}\label{eq:graphobjective}
 \hat{I}(G) := - \sum_{y=0}^k p(y) \Bigl[ d \log(L_{G,y}) - (d-1) \log{n_y} \Bigr],
@@
 to noise, as it does not take cluster size into account.}
-\label{qualitative}
+\label{fig:qualitative}
 \end{figure}
@@
 code is available online\footnote{\url{https://github.com/amueller/information-theoretic-mst}}.
-\enlargethispage{5mm}
 
@@
 
+% TODO more experiments: mnist / usps
+% TODO intrinsic dimensionality
+% todo segmentation unsupervised
+% TODO bag of word experiment?
+
 \section{Conclusions}
@@
 an alternative approach.
-
-\bibliography{paper.bib}

### 2013-08-14 partial_supervision/paper.tex
update jmlr paper in thesis

@@
-\section{Introduction}
 The task of multi-class image segmentation is to create a pixel-wise labeling of an input image

### 2013-08-14 pystruct/paper.tex
update jmlr paper in thesis

@@
-\section{Introduction}
 In recent years there has been a wealth of research in methods for learning structured prediction,
-as well as in applications such as natural language processing and computer vision.
-Unfortunately only few implementations are publicly available, and many applications build upon
+as well as in their application in areas such as natural language processing and computer vision.
+Unfortunately only few implementations are publicly available---many applications are based on
 the non-free implementation of~\citet{joachims2009cutting}.
 
-\pystruct on the other hand aims at providing a simple-to-use interface, in the high-level Python language.
-This allows practitioners to efficiently test a range of models, as well as allowing researchers
-to compare to baseline methods much more easily.
-\pystruct is BSD-Licensed, allowing modification and redistribution of the code,
-as well as use in commercial applications.  By embracing paradimes established
-in the scientific python community, and reusing the interface of the
-widely-used {\sc scikit-learn} library~\citep{pedregosa2011scikit}, \pystruct
-can be used in existing projects, replacing standard classifiers.
-The online documentation and examples help new
-users understand the somewhat abstract ideas behind structured pediction.
+\pystruct aims at providing a high-quality implementation with an easy-to-use
+interface, in the high-level Python language.  This allows practitioners to
+efficiently test a range of models, as well as allowing researchers to compare
+to baseline methods much more easily than this is possible with current
+implementations.  \pystruct is BSD-licensed, allowing modification and
+redistribution of the code, as well as use in commercial applications.  By
+embracing paradigms established in the scientific Python community and reusing
+the interface of the widely-used {\sc scikit-learn}
+library~\citep{pedregosa2011scikit}, \pystruct can be used in existing
+projects, replacing standard classifiers.  The online documentation and
+examples help new users understand the somewhat abstract ideas behind
+structured prediction.
 
@@
 
-Structured prediction can be defined as making a prediction $f(x)$ by maximizing an
-compatibility function $g$ between an input $x$ and the possible labels
+Structured prediction can be defined as making a prediction $f(x)$ by maximizing a
+compati\-bility function between an input $x$ and the possible labels
 $y$~\citep{nowozin2011structured}. Most current approaches use linear functions,
@@
 \begin{equation}\label{eq:main_equation}
-    f(x) = \argmax_{y \in \mathcal{Y}} \langle \theta, \Psi(x, y) \rangle 
+    f(x) = \argmax_{y \in \mathcal{Y}}\  \theta^T \Psi(x, y).
 \end{equation}
-Here, $y$ is a \emph{structured} label in the sense that it is more complicated than a single
-output class. Examples include predicting a label for each word in a sentence, or predicting
-a label for each pixel in an image.
-Learning structured prediction means learning the parameters $\theta$ from training data
-$(x_1, y_1), \dotsc, (x_n, y_n)$.
+Here, $y$ is a \emph{structured} label, $\Psi$ is a joint feature function of $x$ and $y$, and
+$\theta$ are parameters of the model. \emph{Structured} means that $y$ is more complicated than a single
+output class, for example a label for each word in a sentence or a label for
+each pixel in an image.
+Learning structured prediction means learning the parameters $\theta$ from training data.
 
-Using the above formulation learning can be broken down into three sub-problems:
-\begin{itemize}
+Using the above formulation, learning can be broken down into three sub-problems:
+\begin{enumerate*}
     \item Optimizing the objective with respect to $\theta$.
-    \item Encoding the structure of the problem in a joint features function $\Psi$.
+    \item Encoding the structure of the problem in a joint feature function $\Psi$.
     \item Solving the maximization problem in Equation~\ref{eq:main_equation}.
-\end{itemize}
-
+\end{enumerate*}
 The later two problems are usually tightly coupled, as the maximization in
@@
 \pystruct takes an object-oriented approach to decouple the task-dependent
-implementation of 2) and 3) from the general algorithms used to solve 1).
+implementation of 2. and 3. from the general algorithms used to solve 1.
 
-Parameter estimation is done in \texttt{learner} classes, which currently support
-cutting plane algorithms for structural SVMs (SSVMs), subgradient methods for SSVMs,
-the structured perceptron and latent variable SSVMs.
-Encoding the structure of the problem is done using \texttt{model} classes, that
-compute $\Psi$, and encode the structure of the problem.
+Estimating $\theta$ is done in \texttt{learner} classes, which currently support
+cutting plane algorithms for structural support vector machines~(SSVMs), subgradient methods for SSVMs,
+the structured perceptron and latent variable SSVMs. The cutting plane implementation uses the {\sc cvxopt} package \citep{dahl2006cvxopt}
+for quadratic optimization.
+Encoding the structure of the problem is done using \texttt{model} classes, which
+compute $\Psi$ and encode the structure of the problem.
 \pystruct implements models for many common cases, such as multi-class and
@@
 using highly optimized implementations from external libraries. \pystruct
-includes direct support for using {\sc OpenGM}, {\sc LibDAI}~\citep{Mooij_libDAI_10} and {\sc
-AD$^3$}~\citep{martins2011augmented}. It also includes an interface to a
-general purpose linear programming solver.
+includes support for using {\sc OpenGM}~\citep{kappes2013comparative}, {\sc
+LibDAI}~\citep{Mooij_libDAI_10}, QPBO fusion moves~\citep{rother2007optimizing},
+and {\sc AD$^3$}~\citep{martins2011augmented}. It also includes an interface to
+a general purpose linear programming solver from \textsc{cvxopt}~\cite{dahl2006cvxopt}.
 
-Table~\ref{table:comparision} lists implemented algorithms and models, and
+Table~\ref{table:comparision} lists algorithms and models that are implemented in \pystruct and
 compares them to other public structured prediction libraries, together with
@@
 &             &&                     \footnotesize{CP}& \footnotesize{SG}& \footnotesize{LV}& \footnotesize{ML}& \footnotesize{Chain} & \footnotesize{Graph} & \footnotesize{LDCRF}\\
-\sc{\pystruct}&      Python &       BSD            & \x    & \x      & \x   & \o & \x     & \x     & \x \\
-SVM$^\text{\sc{struct}}$ & C++ & non-free         & \x    & \o      & \x   & \o & \o     & \o     & \o \\
+\pystruct&      Python &       BSD$^1$  & \x$^1$    & \x      & \x   & \o & \x     & \x     & \x \\
+\svmstruct & C++ & non-free         & \x    & \o      & \x   & \o & \o     & \o     & \o \\
 \sc{Dlib}         & C++        & boost            & \x    & \x      & \o   & \o & \x     & \x     &\o\\
@@
 \end{tabularx}
-    \caption{Comparison of structured prediction software packages. CP stands
+    \caption{\label{table:comparision}Comparison of structured prediction software packages. CP stands
 for cutting plane optimization of SSVMs~\citep{joachims2009cutting}, SG for
@@
 latent variable SSVMs~\citep{yu2009learning}, ML for maximum likelihood
-training, Chain for chain-structured models
-with pairwise interactionsr, Graph for arbitrary graphs with pairwise
-interactions, and LDCRF for latent dynamic CRF~\citep{morency2007latent}.\label{table:comparision}}
+learning, Chain for chain-structured models
+with pairwise interactions, Graph for arbitrary graphs with pairwise
+interactions, and LDCRF for latent dynamic CRF~\citep{morency2007latent}.
+{\footnotesize $^1$\pystruct itself is BSD licensed, but uses the GPL-licensed package {\sc cvxopt} for cutting-plane learning.}
+}
 \end{table}
+\vfill
+\pagebreak
 
@@
     used in applications. It contains model formulation for many typical
-    applications.  This is in contrast to SVM$^\text{struct}$ that provides no
-    models at all, requiring the user to develop siginificant amounts of code, even
+    applications.  This is in contrast to \svmstruct that provides no
+    models at all, requiring the user to develop significant amounts of code, even
     for simple tasks.
@@
      inference from the task-dependent formulation of $\Psi$. This allows
-     practictioners, for example in computer vision or natural language
+     practitioners, for example in computer vision or natural language
      processing, to improve their model without changing any optimization
@@
      Furthermore, standard benchmarks are included as examples, which allows
-     easy comparision with other approaches.
+     easy comparison with the literature.
 
@@
 \paragraph{Integration}
-     To improve usability, \pystruct is interoperable with other numeric and scientific python projects,
+     To improve usability, \pystruct is interoperable with other numeric and scientific Python projects,
      such as {\sc scikit-learn}~\citep{pedregosa2011scikit},
-     {\sc mahotas}~\citep{coelho:mahotas}, {\sc gensim}~\citep{rehurek_lrec} and {\sc scikit-image}.
+     {\sc mahotas}~\citep{coelho:mahotas}, {\sc gensim}~\citep{rehurek_lrec}, and {\sc scikit-image}.
      This allows users to build powerful applications with little effort. In
-     particular, most of the model-selection methods of scikit-learn can be used
+     particular, most of the model-selection methods of {\sc scikit-learn} can be used
      directly with \pystruct.
@@
 %Multi-class classification is a very simple case of structured prediction, that serves
-%well to exemplify the usage. We load the iris dataset, and compare the usage of \pystruct
+%well to exemplify the usage. We load the iris dataset and compare the usage of \pystruct
 %with a call to LibLinears Crammer-Singer SVM using scikit-learn.
@@
 %columns=flexible, firstline=11]{multi_class_example.py}
-
 %\subsection{OCR sequence classification}
@@
 
-\subsection{Semantic Image Segmentation}
-Conditional random fields are an important tool for semantic image segmentation.
-We show how to train an $n$-slack support vector machines on a superpixel-based CRF
-on the popular Pascal dataset. We use unary potentials generated using TextonBoost
-from \cite{krahenbuhl2012efficient}. The superpixels are generated using SLIC.%
-\footnote{The preprocessed data can be downloaded at \dots.}
-Each sample is represented as a tuple consisting of input features and a graph representation.
+Conditional random fields are an important tool for semantic image
+segmentation.  We demonstrate how to learn an $n$-slack support
+vector machine on a superpixel-based CRF on the popular Pascal dataset. We use
+unary potentials generated using TextonBoost
+from \citet{krahenbuhl2012efficient}. The superpixels are generated using SLIC~\citep{achanta2012slic}.%
+\footnote{The preprocessed data can be downloaded at \url{http://www.ais.uni-bonn.de/download/datasets.html}.}
+Each sample (corresponding on one entry of the list \texttt{X}) is represented as a
+tuple consisting of input features and a graph representation.
+\begin{listing}[t]
+\begin{pythoncode}
+model = crfs.EdgeFeatureGraphCRF(
+    class_weight=inverse_frequency, symmetric_edge_features=[0, 1],
+    antisymmetric_edge_features=[2], inference_method='qpbo')
 
-\lstinputlisting[language=Python, frame=single, basicstyle=\scriptsize,
-    numbers=left, commentstyle=\color{blue}\normalfont,
-columns=flexible,firstline=32, lastline=37]{pascal_crf.py}
-Lines 1-3 declare a quite complex model pairwise model on an arbitray graph.
-Here \texttt{class\_weight} reweights the hamming loss according to class
+ssvm = learners.NSlackSSVM(model, C=0.01, n_jobs=-1)
+ssvm.fit(X, Y)
+\end{pythoncode}
+\caption{Example of defining and learning a CRF model.\label{lst:stuff}}
+\end{listing}
+
+The source code is shown in Listing 1.
+Lines 1-3 declare a model using parametric edge potentials for arbitrary graphs.
+Here \texttt{class\_weight} re-weights the hamming loss according to inverse class
 frequencies. The parametric pairwise interactions have three features: a
-constant feature, color similarity, and relative vertical position. The first
+constant feature, color similarity, and relative vertical position. The first two
 are declared to be symmetric with respect to the direction of an edge, the last
-is antisymmetric. The inference method used is QPBO-fustion moves.  Line 5
-creates an \texttt{learner} object that will learn the parameters for the given
+is antisymmetric. The inference method used is QPBO-fusion moves.  Line 5
+creates a \texttt{learner} object that will learn the parameters for the given
 model using the n-slack cutting plane method, and line 6 performs the actual
 learning.  Using this simple setup, we achieve an accuracy of 30.3 on the
-valiation set following the setup of \citet{krahenbuhl2012efficient}, who
+validation set following the protocol of \citet{krahenbuhl2012efficient}, who
 report 30.2 using a more complex approach. Training the structured model takes
-approximately 30 minutes using a single i7 core (using more would be possible).
-
-%TODO not here
-%This makes it very easy to incoporate features or segments generated by computer vision libraries such
-%as scikit-image or mahotas, and to plot results using visualization tools like matplotlib. 
-% listing, training time, accuracy, comparision with kraehenbuehl
+approximately 30 minutes using a single i7 core.
 
 \section{Experiments}\label{sec:benchmarks}
-In this section, we compare our implementation with SVM$~\text{multiclass}$ and
-SVM$~\text{python}$ with respect to run-time and optimzation accuracy.
-While SVM$~\text{python}$ allows integration with Python programs, it does not use native
-Python data structures, such as Numpy arrays. This makes interacting with existing software
-cumbersome and inefficient, which is reflected in the run-time comparision below.
-We also compare against the special-purpose implementation of the Crammer-Singer multi-class SVM in
-{\sc LibLinear}~\citep{fan2008liblinear}.
+While \pystruct focuses on usability and covers a wide range of applications, it is also
+important that the implemented learning algorithms run in acceptable time.
+In this section, we compare our implementation of the one-slack cutting plance
+algorithm with the implementation in \svmstruct.
+We compare performance of the Crammer-Singer multi-class SVM with respect to
+learning time and accuracy on the MNIST dataset of handwritten digits.
+While multi-class classification is not very interesting from a structured
+prediction point of view, this problem is well-suited to benchmark the cutting
+plane solvers with respect to accuracy and speed.
 
-We do not provide comparisions on more complex problems--such as ones including
-non-trivial inference--as the authors of SVM$^\text{struct}$ did not provide
-any code or examples for this case.
+Results are shown in \Figref{timings}. We report learning times and accuracy for
+varying regularization parameter $C$. The MNIST dataset has 60\,000 training
+examples, 784 features and 10 classes.%
+\footnote{Details about the experiment and code for the experiments can be found on the project website.}
+The figure indicates that \pystruct has competitive performance, while using
+a high-level interface in a dynamic programming language.
 
-MNIST, covertype?
-
-% multiclass on mnist? liblinear seems to die!!!!
+\begin{figure}
+\centering
+\includegraphics[width=.45\textwidth]{times_MNIST}
+\includegraphics[width=.45\textwidth]{accs_MNIST}
+\caption{Runtime comparison of \pystruct and \svmstruct for multi-class
+    classification.
+}
+\label{fig:timings}
+\end{figure}
 
 \section{Conclusion}
-This paper introduced \pystruct, a modular structured learning and prediction library.
+This paper introduced \pystruct, a modular structured learning and prediction library in Python.
 \pystruct is geared towards ease of use, while providing efficient implementations.
@@
 existing libraries and applications.
-Currently \pystruct focuses on max-margin and perceptron based approaches. In the future,
-we plan to integrate other paradigms, such as SampleRank, pseudo-likelihood learning,
-Conditional Neural Fields and approaches that allow for a better integration of inference
-and learning.
+Currently, \pystruct focuses on max-margin and perceptron-based approaches. In the future,
+we plan to integrate other paradigms, such as sampling-based learning~\citep{wick2011samplerank},
+surrogate objectives (for example pseudo-likelihood), and approaches that allow for a better integration
+of inference and learning~\citep{meshi2010learning}.

### 2013-08-14 make.sh

@@ -1,2 +1,1 @@
-#!/bin/bash
-rubber -I itm/images -I mil/images -I partial_supervision -I mil -I exact_learning -I pystruct main.tex
+rubber -c 'setlist arguments --shell-escape'  -I itm/images -I mil/images -I partial_supervision -I mil -I exact_learning -I pystruct --pdf main.tex

### 2013-08-14 mil/paper.tex
minor fixes, images
fix bibliography
remove wrong statements from MIL paper.

@@
-%\newtheorem{corollary}{Corollary}
-%\newtheorem{corollarya}{Corollary}
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{theorema}{Theorem}
-
-%\newcolumntype{R}{>{\raggedleft\arraybackslash}X}
-%\newcolumntype{C}{>{\centered\arraybackslash}X}
-
-%\newcommand{\fix}{\marginpar{FIX}} \newcommand{\new}{\marginpar{NEW}}
-%\newcommand\etal{~et\,al.~}
-
-%\aistatsauthor{Andreas M\"uller\thanks{This work was founded by the B-IT research
-%school.}  \and Sven Behnke\\ {amueller@ais.uni-bonn.de},
-%{behnke@cs.uni-bonn.de}} {Department of Computer Science\\
-%University of Bonn\\ 53113 Bonn, Germany \\ }
-
 \section{Introduction}\seclabel{intro}
@@
 when allowing arbitrary dependencies between instances. On the other hand, the i.i.d.\ assumption of \Secref{iid}
-might be to strong in a practical setting.
+might be too strong in a practical setting.
 Therefore, we propose an assumption that is both very general and allows learning of instance classification:
@@
 \end{align}
-
-By bounding the right hand side we obtain:
-\begin{theorem}\label{bound}
-    Assume all instances within a bag are independent given an unobserved bag variable. If $h_0^r \geq e_B$ the instance error $e_I$ can
-    be bound by
-\begin{align}
-    e_I \leq h_0 -  \left ( h_0^r - e_B) \right ).
-\end{align}
-\end{theorem}
-While this bound might not be very tight -- for example if the variables are i.i.d.\ and $r$ is big, it is clear
-that it is very loose compared to Theorem~\ref{iidbound} -- it is to the authors knowledge the first theorem that shows that instance-level
-predictions can not be arbitrarily bad, even in a complicated dependent setting.
-
-If we turn to the setting where bag labels are known, the bound of Theorem~\ref{bound} can be extended outside $( h_0^r > e_B)$
-using the same statement for $p(y=0)$ instead of $h_0$. If there are places where both $h_0^r < e_B$ and
-$p(y=0)^r < e_B$, we can use another bound, finally yielding:
-
-\begin{theorem}\label{globalbound}
-    If all instances within a bag are independent given an unobserved bag variable and the distribution of
-    bag labeld $P(Y)$ is known, the instance error $e_I$ can
-    be bound by
-\begin{align}
-    p(h_I(x)\neq y) \leq \text{min}\{ b_h, b_Y, b_g \} 
-\end{align}
-with
-\begin{align*}
-b_g &= h_0 + p(Y=0)^\frac{1}{r} - h_0^r - p(Y=0) + e_B,  \\
-b_h &= h_0 - h_0^r + e_B \text{\ if } h_0^r > e_B \text{, else }\infty, \\
-b_y &= p(Y=0)^\frac{1}{r} - p(Y=0) + e_B \\
-&\text{\ if } p(Y=0) > e_B \text{, else }\infty. \\
-\end{align*}
-\end{theorem}
 
@@
 in multiple instance learning.
-\bibliographystyle{plainnat} \bibliography{paper}
 
@@
 \begin{theorema}
-    Assume all instances within a bag are independent given an unobserved bag variable. If $h_0^r \geq e_B$ the instance error $e_I$ can
-    be bound by
-\begin{align}
-    e_I \leq h_0 -  \left ( h_0^r - e_B) \right ).
-\end{align}
-\begin{proof}
-    Using the proof of Theorem~\ref{iidbound}, we can obtain
-\begin{multline}
-    p(h_I(x)\neq y| Z) \leq p(h_I(x)=0|Z) \\
-    - \left ( p(h_B(x)=0|Z) - p(h_B(X)\neq Y| Z) \right ) ^ \frac{1}{r}
-\end{multline}
-Taking expected values with respect to $Z$ on both sides yields
-\begin{align}
-    e_I \leq h_0 - \mathbb{E}_Z \left ( p(h_I(x)=0|Z)^r - p(h_B(X)\neq Y| Z) \right ) ^ \frac{1}{r}
-\end{align}
-Since $0 \leq p(h_I(x)=0|Z)^r - p(h_B(X)\neq Y| Z) \leq 1$, the $r$-th root of this term can
-be lower bounded by the identity function, resulting in
-\begin{multline}
-    e_I \leq h_0 - \mathbb{E}_Z \left[  p(h_I(x)=0|Z)^r \right] - e_B 
-\end{multline}
-The theorem follows with Jensen's inequality.
-\end{proof}
-\end{theorema}
-
-\begin{theorema}
-    If all instances within a bag are independent given an unobserved bag variable and the distribution of
-    bag labeld $P(Y)$ is known, the instance error $e_I$ can
-    be bound by
-\begin{align}
-    p(h_I(x)\neq y) \leq \text{min}\{ b_h, b_Y, b_g \} 
-\end{align}
-with
-\begin{align}
-b_g &= h_0 + p(Y=0)^\frac{1}{r} - h_0^r - p(Y=0) + e_B  \\
-b_h &= h_0 - h_0^r + e_B \text{\ if } h_0^r > e_B \text{, else }\infty \\
-b_y &= p(Y=0)^\frac{1}{r} - p(Y=0) + e_B \\
-&\text{\ if } p(Y=0) > e_B \text{, else }\infty \\
-\end{align}
-\begin{proof}
-    The bounds given by $b_h$ follows from Theorem~\ref{bound}, and $b_y$ as described in the text.
-    If both  $h_0^r < e_B$ and $p(y=0)^r < e_B$, then $0 \leq ( h_0^r + p(y=0)^r - e_B) \leq 1$ and
-    we can lower bound the $r$-th root as in the proof of Theorem~\ref{bound}. This yields the bound given
-    by $b_g$.
-\end{proof}
-\end{theorema}
-
-\begin{theorema}
 If the instance labels are independent inside a bag given a hidden cause, the instance

### 2013-08-14 exact_learning/paper.tex
Moving around a lot of sections, renaming conclusion to summary, starting to merge mil and semisupervised, starting to rework structur of pystruct and exact learning.

@@
 \section{Introduction}
-Many classical computer vision applications such as stereo, optical flow, semantic
-segmentation and visual grouping can be naturally formulated as image labeling tasks.
+% Remove duplication with pystruct intro
+As discussed previously, many computer vision algorithms empoy conditional random
+field models on pixel or superpixel graphs. These graphs by nature contain loops,
+making exact prediction and loss-augmented prediction intractable.
+As loss-augmented prediction is a centreal step in the learning algorithms
+discussed in Chapter~\ref{ch:structured_pystruct}
+in the inference then lead to approximate learning.
 
-Arguably the most popular way to approach such labeling problems is via graphical
-models, such as Markov random fields (MRFs) and conditional random fields (CRFs).
-MRFs and CRFs provide a principled way of integrating local evidence and
-modeling spacial dependencies, which are strong in most image-based tasks.
+In this chapter, we investigate the impact of these approximations in
+the one-slack learning algorithms in the context of semantic 
+image segmentation. We show that despite inference being deemed intractable in
+loopy superpixel models, we are able to learn a structured support vector machine
+for this application exactly.
+learning a model of pairwise
+We will evaluate our approach on the popular MSRC-21 and Pascal VOC datasets.
 
-While in earlier approaches, model parameters were set by hand or using
-cross-validation, more recently parameters are often learned using a max-margin
-approach.
-
-Most models employ linear energy functions of unary and pairwise interactions,
-trained using structural support vector machines (SSVMs). While linear energy
-functions lead to learning problems that are convex in the parameters, complex
-constraints complicate their optimization. Additionally, inference (or more precisely
-loss-augmented prediction) is a crucial part in learning, and can often not be
-performed exactly, due to loops in the neighborhood graphs.  % TODO clean up!
-Approximations in the inference then lead to approximate learning.
-
-We look at semantic image segmentation, learning a model of pairwise
-interactions on the popular MSRC-21 dataset.
-The contribution of this work is threefold:
+The contribution of this chapter is threefold:
 \begin{itemize}
@@
 \section{Efficient Cutting Plane Training of SSVMs}\label{learning}
-
-\subsection{The Cutting Plane Method}\label{cutting_plane}
-
-When learning for structured prediction in the max-margin framework of
-\citet{tsochantaridis2006large}, predictions are made as
-\[ \argmax_{y \in \mathcal{Y}} f(x, y, \theta),\]
-where $x \in \mathcal{X}$ is the input, $y \in \mathcal{Y}$ the prediction, and $\theta$ are the parameters to be learned.
-We will assume $y$ to be multivariate, $y=(y_1, \dots, y_k)$ with possibly varying $k$.
-
-The function $f$ measures compatibility of $x$ and $y$ and is a linear function of the parameters $\theta$:
-\[f(x, y, \theta) = \langle \theta, \psi(x, y) \rangle.\]
-Here $\psi(x, y)$ is a joint feature vector of $x$ and $y$. Specifying a particular SSVM model
-amounts to specifying $\psi$.
-
-For a given loss $\Delta$, the parameters $\theta$ are learned by minimizing
-the loss-based soft-margin objective
-\begin{equation}\label{eq:mainequation}
-\min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  r_i(\theta)
-\end{equation}
-with regularization parameter $C$, where $r_i$ is a hinge-loss-like upper
-bound on the empirical $\Delta$-risk:
-\[r_i(x^i, y^i, y) = \left [\max_{y \in \mathcal{Y}} \Delta(y^i, y) + \left<\theta, \psi(x^i, y) - \psi(x^i, y^i) \right > \right]_+ \]
-
-We solve the following reformulation of Equation~\eqref{mainequation}, known as one-slack QP formulation:
-\begin{align}\label{eq:oneslack}
-    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
-    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
-        &\left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i,
-            \hat{y}^i)] \right \rangle \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
-            - \xi
-\end{align}
-using the cutting plane method described in
-Algorithm~\ref{alg_cutting_plane}~\citep{joachims2009cutting}.
-
-The cutting plane method alternates between solving Equation~\eqref{oneslack}
-with a working set $\W$ of constraints, and expanding the working set using the
-current $\theta$ by finding $\mathbf{y}$ corresponding to the most violated constraint,
-using a separation oracle $I$.
-We investigate the construction of $\W$ and the influence of $I$
-on learning.
-
-%TODO unstructured!!
-
-Intuitively, the one-slack formulation corresponds to joining all training
-samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
-interactions between variables corresponding to different data points.
-Consequently, only a single constraint is added in each iteration of
-Algorithm~\ref{alg_cutting_plane}, leading to very small $\W$. We further use
-pruning of unused constraints, as suggested by \citet{joachims2009cutting},
-resulting in the size of $\W$ being in the order of hundreds for all experiments.
-
-We also use another enhancement of the cutting plane algorithm introduced by
-\citet{joachims2009cutting}, the caching oracle. For each training example $(x^i, y^i)$,
-we maintain a set $C^i$ of the last $r$ results of loss-augmented inference
-(line~\ref{get_cutting_plane} in Algorithm~\ref{alg_cutting_plane}).
-For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we find
-\[ 
-    \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y})] \right \rangle
-\]
-by enumeration of $C^i$ and continue until line~\ref{convergence_check}.
-Only if $\xi' - \xi < \epsilon$, i.e.\ the produced constraint is not violated, we
-return to line~\ref{get_cutting_plane} and actually invoke the separation
-oracle $I$.
-
-In computer vision applications, or more generally in graph labeling problems,
-$\psi$ is often given as a factor graph over $y$, typically using only unary and pairwise functions:
-\[ \langle \theta, \psi(x, y) \rangle = \sum_{i=0}^k \langle \theta_u,  \psi_u(x, y_k) \rangle + \sum_{(i, j) \in E} \langle \theta_p, \psi_p(x, y_k, y_l) \rangle, \]
-where $E$ a set of pairwise relations. In this form, parameters $\theta_u$ and $\theta_p$ for unary and
-pairwise terms are shared across all entries and pairs.
-The decomposition of $\psi$ into only unary and pairwise interactions allows
-for efficient inference schemes, based on message passing or graph cuts.
 

### 2013-08-14 intro.tex
update jmlr paper in thesis
Moving around a lot of sections, renaming conclusion to summary, starting to merge mil and semisupervised, starting to rework structur of pystruct and exact learning.

@@
 
-Why semantic segmentation?
+Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
 Why is it not solved yet?
@@
 Datsets? Input?
+
+Structured approaches, latent variable approaches
+
+Contributions:
+- Introduces general clustering algorithm that improves pipeline in serveral places
+- Shows that weakly supervised object class segmentation (not semantic segmentation) is possible
+- Provides general efficient framework for structured prediction
+- Analyze max-margin learning algorithms in different regimes
+- shows exact learning is possible in loopy graphs and improves results (sometimes?)
+- shows importance of edge features
+- Demonstrates how latent variables benefit semantic segmentation

### 2013-08-14 itm/paper.tex
Moving around a lot of sections, renaming conclusion to summary, starting to merge mil and semisupervised, starting to rework structur of pystruct and exact learning.

@@
-Before going to segmentation, we look at clustering because:
-1 we see how far we can go without supervision
-2 we need it as building block for bow
-3 we need it as initialization for latent variable models
-%TODO elaborate
+Before we start our investigation of semantic segmentation and object class segmentation,
+we will look into a general purpose clustering algorithm.
 
-Clustering data is one of the fundamental problems in machine learning.  In
-clustering, the goal is to divide data points into homogeneous subsets, called
+In clustering, the goal is to divide data points into homogeneous subsets, called
 clusters.
 Many different formulations of the clustering problem are given in the literature.
+%
+In the context of this work, clustering plays an important role in many respects:
+\begin{itemize}
+    \item Clustering, as a purely unsupervised methods is on one end of the spectrum of algorithms we investigate.
+        When using per-image or per-pixel supervision as in the later chapters, we can use clustering to calibrate our
+        expecation of what semisupervised and supervised algorithms have at least to achieve.
+    \item As many other computer vision algorithms, our segmentation methods
+        with critically depend on bag-of-feature representations of segments or
+        images. These are build using a vocabulary of visual words, that is
+        usually created via clustering.
+    \item When discussing latent variable models in chapter~\ref{ch:latent}, we need to initialize our
+        model to overcome the inherant symmetries. Clustering methods are a
+        straight forward solution to this problem.
+\end{itemize}
+
+So not only is clustering one of the fundamental problems in machine learning, it is also
+an imortant building block for the methods in this work.
+%
 Most algorithms are based on ad-hoc criteria such as intra-cluster similarity
 and inter-cluster dissimilarity.
-%
 An alternative approach is to formalize clustering using an information
@@
 
-In this work, we rely on a non-parametric estimator of the data entropy to find 
+In the following, we rely on a non-parametric estimator of the data entropy to find 
 clusterings of maximum mutual information.
@@
 %
-The contributions of this work are:
+The contributions of this chapter are:
 \begin{itemize}
@@
 They are given by the Voronoi tessellation of the cluster means and therefore always convex.
-%Not only are clusters as found by $k$-Means convex,
-%they are given by the Voronoi tessellation of the cluster means.
+Another draw-back is the non-determinism of the procedure, caused by the
+dependence on random initialization.
 
@@
 Define
-%\setlength{\abovedisplayskip}{1mm}
-%\setlength{\belowdisplayskip}{1mm}
 \begin{align}\label{eq:graphobjective}
@@
 Thus we can formulate the clustering problem equivalently as
-$\displaystyle \max_{G \in \mathcal{F}} \hat{I}(G)$.
+\begin{equation}
+    \max_{G \in \mathcal{F}} \hat{I}(G).
+\end{equation}
 
@@
 
-\section{Conclusions}
-In this work we proposed the use of a minimum spanning tree based,
+\section{Summary}
+In this summary we proposed the use of a minimum spanning tree based,
 non-parametric entropy estimator in information theoretic clustering, ITM\@.
@@
 
-A limitation of the proposed algorithm is that it is based on the assumption 
-of an absolute continuous data distribution. This assumption eliminates the 
-possibility of using categorical variables and data that lies on a submanifold 
-of the input space. 
+%A limitation of the proposed algorithm is that it is based on the assumption 
+%of an absolute continuous data distribution. This assumption eliminates the 
+%possibility of using categorical variables and data that lies on a submanifold 
+%of the input space. 
 %
-In future work we plan to investigate a way to overcome this limitation, 
-for example by estimating the intrinsic dimensionality of the 
-data~\citep{pettis1979intrinsic}.
+%In future work we plan to investigate a way to overcome this limitation, 
+%for example by estimating the intrinsic dimensionality of the 
+%data~\citep{pettis1979intrinsic}.
 %
-We will also investigate optimizations of the objective 
-\eqref{graphobjective} that go beyond the proposed method. 
-Move-making algorithms seem a promising way to refine solutions 
-found by Algorithm~1. Branch and bound techniques could provide 
-an alternative approach.
+%We will also investigate optimizations of the objective 
+%\eqref{graphobjective} that go beyond the proposed method. 
+%Move-making algorithms seem a promising way to refine solutions 
+%found by Algorithm~1. Branch and bound techniques could provide 
+%an alternative approach.

### 2013-08-14 mil/paper.tex
Moving around a lot of sections, renaming conclusion to summary, starting to merge mil and semisupervised, starting to rework structur of pystruct and exact learning.

@@
-\section{Introduction}\seclabel{intro}
+\section{Theoretical settings}\seclabel{intro}
 Multiple instance learning is a supervised learning scenario first proposed by \citet{dietterich1997solving}.
@@
 instance labeling exists for training or testing.
+
 %If one is interested in classifying instances, such a bound is very important.
@@
 
-\section{Conclusion}
-We provided several bounds on instance prediction error in a multiple instance learning setting.
-For the case that instances inside a bag are independent, we proved a strong upper bound on the error.
-We showed that when bag labels are present, it is even possible to directly estimate the true instance error.
+\section{Summary}
+We proposed an algorithm for object-class segmentation using only weak supervision based on
+multiple-instance learning. In our approach each image is represented as a bag of object-like
+proposal segments.
 
-We introduced a restricted dependent setting for multiple instance learning, in which instances inside a bag
-are only independent given an unobserved variable. We showed that without any further assumptions
-on the distribution of this variable, it is possible to learn instance label prediction from bag labels
-in this setting. We also provided upper and lower bounds on the instance error, given known bag labels
-or an estimage of the bag error.
-Empirical evaluation on a synthetic dataset with known dependen structure and known instance labels
-showed that the bound are non-trivial and yield insight into the behavior of instance label errors.
-We hope that this work establishes a basis for further research on instance label prediction
-in multiple instance learning.
+We described a way to extent bag level predictions made by the multi-instance
+kernel method to instance level while remaining competitive with the state-of-the-art
+in bag label prediction.
 
-\section{Appendix}
-\begin{lemma}\label{binarylemma}
-Given two (dependent) binary random variables $x,y$:
-\begin{equation}
-p(x \neq y) = p(x=0) + p(y=0) - 2 p(x=0, y=0)
-\end{equation}
+We evaluated the proposed object-class segmentation method on the challenging
+Graz02 dataset. While not reaching the performance of methods using strong supervision,
+our result can work as a baseline for further research into weakly supervised object class
+segmentation.
 
-\begin{proof}
-\begin{align}
-p(x \neq y) & = p(x=0, y=1) + p(x=1, y=0) \\
-                &= p(x=0) - p(x=0, y=0)\\
-                &  + p(y=0) - p(x=0, y=0) \\
-                &= p(x=0) + p(y=0) -2 p(x=0, y=0)
-\end{align}
-\end{proof}
-\end{lemma}
+%In future work, we plan to scale our approach to much larger image datasets. As much
+%more images with weak annotations are available than with pixel level segmentation,
+%we hope that we can improve upon the state-of-the-art in object-class segmentation
+%by making use of larger bodies of training images.
+On the theoretical side, we provided several bounds on instance prediction
+error in a multiple instance learning setting.  For the case that instances
+inside a bag are independent, we proved a strong upper bound on the error.  We
+showed that when bag labels are present, it is even possible to directly
+estimate the true instance error.
 
-\begin{theorema}
-Let all instances be independent. With bags $X$, instances $x$, true labels $y,Y$, $h_I$ any classifier
-on instances and $h_B$ the resulting bag classifier, we have
-\begin{align*}
-   e_I = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}
-\end{align*}
-\begin{proof}
-
-\begin{align*}
-e_B & = p(h_B(X) = 0) + p(Y=0) -2 p(h_B(X)=0, Y=0) \\\
-&\text{using the lemma}\\
-& = h_0^r + p(y=0)^r -2 p(h_I(x)=0, y=0)^r\\\
-&\text{by independence}\\
-& = h_0^r + p(y=0)^r -2 \left ( \frac{1}{2} (h_0 + p(y=0) - e_I) \right ) ^r \\
-&\text{using Lemma~\ref{binarylemma} on }p(h_I(x)=0, y=0.
-\end{align*}
-Solving for $p(h_I(x)\neq y)$:
-\begin{align*}
-    & \left ( \frac{1}{2} (h_0 + p(y=0) - e_I )\right ) ^r = \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right )\\
-    & \Rightarrow \frac{1}{2} (h_0 + p(y=0) - e_I)   = \left ( \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right ) \right )^\frac{1}{r}\\
-    & \Rightarrow e_I   = h_0 + p(y=0) - 2 \left ( \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right ) \right )^\frac{1}{r}
-\end{align*}
-\end{proof}
-\end{theorema}
-
-\begin{corollarya}
-If all instances are independent and bag classification is perfect, meaning $e_B=0$, then instance classification is also perfect ($e_I =0$).
-
-\begin{proof}
-Consider Theorem~\ref{basicthm}. If $e_B=0$ we have
-\begin{align*}
-p(h_I(x) \neq y) = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r)) \right)^\frac{1}{r}
-\end{align*}
-By the convexity of the power function the right hand side of the equation is always greater or equal zero.
-Since $e_I \geq 0$, both sides are equal zero.
-\end{proof}
-\end{corollarya}
-
-\begin{theorema}
-    If $h_0^r  \geq e_B$ and all instances are independent, we have
-\begin{align}
-    p(h_I(x)\neq y) \leq h_0 - \left ( h_0^r - e_B) \right ) ^ \frac{1}{r}
-\end{align}
-\begin{proof}
-    Assume $h_0^r  \geq e_B$.
-    From Theorem~\ref{basicthm} we have
-\begin{align*}
-   e_I &= h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}\\
-   &\leq h_0 + p(y=0) - \left [( h_0^r + p(y=0)^r - e_B) \right]^\frac{1}{r}\\
-   &- \left ( p(y=0)^r \right )^\frac{1}{r}\\
-   &= h_0 - \left (( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}\\
-\end{align*}
-\end{proof}
-\end{theorema}
-
-\begin{theorema}
-Let all instances be independent. Given $p(Y)$, the instance error $e_I$ can be calculated by
-\begin{align}
-    %\eqlabel{iid}
-e_I = h_0 + p(Y=0)^\frac{1}{r} - 2 \left (\frac{1}{2} ( h_0^r + p(Y=0) - e_B) \right)^\frac{1}{r}.
-\end{align}
-\begin{proof}
-    This follows from Theorem~\ref{basicthm} by using $p(Y=0)=p(y=0)^r$.
-\end{proof}
-\end{theorema}
-
-\begin{theorema}
-If the instance labels are independent inside a bag given a hidden cause, the instance
-error $e_I$ has the following lower bound:
-\begin{multline*}
-e_I \geq h_0 + p(y=0) - \\
-2 \left (\frac{1}{2} (p(h_B(X)=0) + p(Y=0) - e_B) \right)^\frac{1}{r}
-\end{multline*}
-\begin{proof}
-    This follows from taking the expected value with respect to $Z$ of \eqref{conditioned} and using Jensen's
-    inequality.
-\end{proof}
-\end{theorema}
+We introduced a restricted dependent setting for multiple instance learning, in
+which instances inside a bag are only independent given an unobserved variable.
+We showed that without any further assumptions on the distribution of this
+variable, it is possible to learn instance label prediction from bag labels in
+this setting. We also provided upper and lower bounds on the instance error,
+given known bag labels or an estimage of the bag error.
+Empirical evaluation on a synthetic dataset with known dependen structure and
+known instance labels showed that the bound are non-trivial and yield insight
+into the behavior of instance label errors.  We hope that this work establishes
+a basis for further research on instance label prediction in multiple instance
+learning.

### 2013-08-14 partial_supervision/paper.tex
Moving around a lot of sections, renaming conclusion to summary, starting to merge mil and semisupervised, starting to rework structur of pystruct and exact learning.

@@
-The task of multi-class image segmentation is to create a pixel-wise labeling of an input image
-into regions belonging to one of several semantic classes.
-Most algorithms for this setting work with strong supervision: a pixel-wise labeling
-of training images. Methods that are used in this setting include random forests~\citep{schroff2008object} and support vector machines (SVM). Usually
-the output of such algorithms is further processed by a conditional random field (CRF~\citep{ladicky2009associative,gonfaus2010harmony,jiang2009efficient}).
-While these methods reach high accuracy, it is very time consuming to create
-pixel-level ground truth for real-world applications.
-This problem can be addressed in several ways:
-the LabelMe effort~\citep{russell2008labelme} tries to use the ``wisdom of crowds'' to obtain human labelings.
-Another possibility is to use only weak supervision, which is the approach we follow here.
+Most algorithms for semantic image segmentation and object-class segmentation work with strong supervision: a pixel-wise labeling
+of training images. In this chapter, we will investigate a method that works with annotation
+that is much easier to obtain: whole image labels.
 
-In the weakly supervised setting, the ground
- truth for a given image is a list of semantic classes that occur in this image, instead of a pixel-level
-labeling as in the strongly supervised setting.
-Image-level labels are much easier to obtain, e.~g.\  through online image libraries such as flickr and facebook.
+% TODO contrast with Alex
 
@@
 
-In this work, we explore the 
+This chapter explores the 
 application of multi-instance learning algorithms to the task of partially supervised image segmentation.
@@
 multi-instance learning to the task of object-class segmentation in natural
-images. To our knowledge, all previous
-methods in the field use strong supervision, meaning manual pixel-wise annotation of
-training images. This approach does not
-scale to larger datasets, especially if one expects consistency and quality
-in the segmentations.
+images.
+% FIXME relate to alex here
+To our knowledge, all previous methods in the field use strong
+supervision, meaning manual pixel-wise annotation of training images. This
+approach does not scale to larger datasets, especially if one expects
+consistency and quality in the segmentations.
+
+After our empirical evaluation on image segmentation, we will extend the existing theoretical framework
+on instance classification based on multi-instance prediction.
 
@@
 using many different approaches \citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
-We will discuss only those that are relevant to this work.
+We will discuss only those that are relevant to the present treatment.
 
@@
 \subsection{Constraint Parametric Min Cuts (CPMC)}
-To generate proposal segments, we use the CPCM framework from \citet{carreira2010constrained}. We construct initial
-segments using graph cuts, on the image graph. The energy function for these cuts uses pixel
+To generate proposal segments, we use the Constraint Parametric Min Cut (CPCM) method from \citet{carreira2010constrained}. In CPMC, initial
+segments are constructed graph cuts on the pixel grid. The energy function for these cuts uses pixel
 color and the response of the global probability of boundary (gPb) detector~\citep{maire2008using}. As much as ten thousand initial segments are generated from foreground
@@
 it can serve as a baseline for research on weakly supervised methods for image segmentation.
-
-\section{Conclusions}
-We proposed an algorithm for object-class segmentation using only weak supervision based on
-multiple-instance learning. In our approach each image is represented as a bag of object-like
-proposal segments.
-
-We described a way to extent bag level predictions made by the multi-instance
-kernel method to instance level while remaining competitive with the state-of-the-art
-in bag label prediction.
-
-Finally we evaluated the proposed object-class segmentation method on the challenging
-Graz02 dataset. While not reaching the performance of methods using strong supervision,
-our result can work as a baseline for further research into weakly supervised object class
-segmentation.
-
-In future work, we plan to scale our approach to much larger image datasets. As much
-more images with weak annotations are available than with pixel level segmentation,
-we hope that we can improve upon the state-of-the-art in object-class segmentation
-by making use of larger bodies of training images.

### 2013-08-14 pystruct/paper.tex
Moving around a lot of sections, renaming conclusion to summary, starting to merge mil and semisupervised, starting to rework structur of pystruct and exact learning.

@@
+Many classical computer vision applications such as stereo, optical flow, semantic
+segmentation and visual grouping can be naturally formulated as image labeling tasks.
+
+Arguably the most popular way to approach such labeling problems is via graphical
+models, such as Markov random fields (MRFs) and conditional random fields (CRFs).
+MRFs and CRFs provide a principled way of integrating local evidence and
+modeling spacial dependencies, which are strong in most image-based tasks.
+
+While in earlier approaches, model parameters were set by hand or using
+cross-validation, more recently parameters are often learned using a max-margin
+approach.
+
+Most models employ linear energy functions of unary and pairwise interactions,
+trained using structural support vector machines (SSVMs). While linear energy
+functions lead to learning problems that are convex in the parameters, complex
+constraints complicate their optimization.
+
 In recent years there has been a wealth of research in methods for learning structured prediction,
@@
 the non-free implementation of~\citet{joachims2009cutting}.
+
+In this chapter, we will first introduce the concepts and algorithms used in structured prediction,
+in particular in maximum margin methods. We will then accompany this by a description of our
+open source implementation of these algorithms, \pystruct.
 
@@
 
-\section{Structured Prediction and Casting it into Software}\label{sec:api}
+\section{Basic Concepts in Structured Prediction}
 
@@
 
+% TODO relate to computer vision here
+When learning for structured prediction in the max-margin framework of
+\cite{tsochantaridis2006large}, predictions are made as
+\[ \argmax_y f(x, y, \theta)\]
+where x is the input and $\theta$ are the parameters to be learned.
+Here, $f$ is a compatibility function of $x$ and $y$, that is linear in the parameters
+\[f(x, y, \theta) = \langle \theta, \Psi(x, y) \rangle\]
+where $\Psi(x, y)$ is a joint feature vector of $x$ and $y$. Specifying a specific CRF model
+amounts to specifying $\Psi$.
+
+For a given loss $\Delta$, the parameters $\theta$ are learned by minimizing
+the loss base soft-margin objective
+\begin{equation}\label{mainequation}
+    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  r_i(\theta)
+\end{equation}
+with regularization parameter $C$, where $r_i$ is an hinge-loss like upper
+bound on $\Delta$ empirical risk:
+\[r_i(x_i, y_i, y) = [\max_{y \in \mathcal{Y}} \Delta(y_i, y) + \left<\theta, \psi(x_i, y) - \psi(x_i, y_i) \right >]_+ \]
+\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
+\begin{equation}\label{subgradient}
+    g := \psi(x_i, y) - \psi(x_i, y_i) + \theta / C
+\end{equation}
+leading to a simple iterative optimization.
+
+\section{Learning Algorithms for Linear Max-Margin Structured Prediction}
+% TODO algorithms here
+\subsection{The Cutting Plane Method}\label{cutting_plane}
+
+When learning for structured prediction in the max-margin framework of
+\citet{tsochantaridis2006large}, predictions are made as
+\[ \argmax_{y \in \mathcal{Y}} f(x, y, \theta),\]
+where $x \in \mathcal{X}$ is the input, $y \in \mathcal{Y}$ the prediction, and $\theta$ are the parameters to be learned.
+We will assume $y$ to be multivariate, $y=(y_1, \dots, y_k)$ with possibly varying $k$.
+
+The function $f$ measures compatibility of $x$ and $y$ and is a linear function of the parameters $\theta$:
+\[f(x, y, \theta) = \langle \theta, \psi(x, y) \rangle.\]
+Here $\psi(x, y)$ is a joint feature vector of $x$ and $y$. Specifying a particular SSVM model
+amounts to specifying $\psi$.
+
+For a given loss $\Delta$, the parameters $\theta$ are learned by minimizing
+the loss-based soft-margin objective
+\begin{equation}\label{eq:mainequation}
+\min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  r_i(\theta)
+\end{equation}
+with regularization parameter $C$, where $r_i$ is a hinge-loss-like upper
+bound on the empirical $\Delta$-risk:
+\[r_i(x^i, y^i, y) = \left [\max_{y \in \mathcal{Y}} \Delta(y^i, y) + \left<\theta, \psi(x^i, y) - \psi(x^i, y^i) \right > \right]_+ \]
+
+We solve the following reformulation of Equation~\eqref{mainequation}, known as one-slack QP formulation:
+\begin{align}\label{eq:oneslack}
+    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
+    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
+        &\left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i,
+            \hat{y}^i)] \right \rangle \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
+            - \xi
+\end{align}
+using the cutting plane method described in
+Algorithm~\ref{alg_cutting_plane}~\citep{joachims2009cutting}.
+
+The cutting plane method alternates between solving Equation~\eqref{oneslack}
+with a working set $\W$ of constraints, and expanding the working set using the
+current $\theta$ by finding $\mathbf{y}$ corresponding to the most violated constraint,
+using a separation oracle $I$.
+We investigate the construction of $\W$ and the influence of $I$
+on learning.
+
+%TODO unstructured!!
+
+Intuitively, the one-slack formulation corresponds to joining all training
+samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
+interactions between variables corresponding to different data points.
+Consequently, only a single constraint is added in each iteration of
+Algorithm~\ref{alg_cutting_plane}, leading to very small $\W$. We further use
+pruning of unused constraints, as suggested by \citet{joachims2009cutting},
+resulting in the size of $\W$ being in the order of hundreds for all experiments.
+
+We also use another enhancement of the cutting plane algorithm introduced by
+\citet{joachims2009cutting}, the caching oracle. For each training example $(x^i, y^i)$,
+we maintain a set $C^i$ of the last $r$ results of loss-augmented inference
+(line~\ref{get_cutting_plane} in Algorithm~\ref{alg_cutting_plane}).
+For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we find
+\[ 
+    \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y})] \right \rangle
+\]
+by enumeration of $C^i$ and continue until line~\ref{convergence_check}.
+Only if $\xi' - \xi < \epsilon$, i.e.\ the produced constraint is not violated, we
+return to line~\ref{get_cutting_plane} and actually invoke the separation
+oracle $I$.
+
+In computer vision applications, or more generally in graph labeling problems,
+$\psi$ is often given as a factor graph over $y$, typically using only unary and pairwise functions:
+\[ \langle \theta, \psi(x, y) \rangle = \sum_{i=0}^k \langle \theta_u,  \psi_u(x, y_k) \rangle + \sum_{(i, j) \in E} \langle \theta_p, \psi_p(x, y_k, y_l) \rangle, \]
+where $E$ a set of pairwise relations. In this form, parameters $\theta_u$ and $\theta_p$ for unary and
+pairwise terms are shared across all entries and pairs.
+The decomposition of $\psi$ into only unary and pairwise interactions allows
+for efficient inference schemes, based on message passing or graph cuts.
+
+\section{Casting Structured Prediction into Software}\label{sec:api}
+
 Using the above formulation, learning can be broken down into three sub-problems:
@@
 
-Estimating $\theta$ is done in \texttt{learner} classes, which currently support
-cutting plane algorithms for structural support vector machines~(SSVMs), subgradient methods for SSVMs,
-the structured perceptron and latent variable SSVMs. The cutting plane implementation uses the {\sc cvxopt} package \citep{dahl2006cvxopt}
-for quadratic optimization.
+Estimating $\theta$ is done in \texttt{learner} classes, which currently
+support cutting plane algorithms for structural support vector
+machines~(SSVMs), subgradient methods for SSVMs, the structured perceptron and
+latent variable SSVMs. The cutting plane implementation uses the {\sc cvxopt}
+package \citep{dahl2006cvxopt} for quadratic optimization.
+
 Encoding the structure of the problem is done using \texttt{model} classes, which
@@
 \section{Usage Example: Semantic Image Segmentation}\label{sec:examples}
+%TODO Multi-label, chain?
+
 %\subsection{Multi-class classification}
@@
 
-\section{Conclusion}
+\section{Summary}
+%TODO summarize structured prediction
 This paper introduced \pystruct, a modular structured learning and prediction library in Python.

### 2013-08-14 mil/appendix.tex
adding appendix for mil in separate file

\section{Appendix}
\begin{lemma}\label{binarylemma}
Given two (dependent) binary random variables $x,y$:
\begin{equation}
p(x \neq y) = p(x=0) + p(y=0) - 2 p(x=0, y=0)
\end{equation}

\begin{proof}
\begin{align}
p(x \neq y) & = p(x=0, y=1) + p(x=1, y=0) \\
                &= p(x=0) - p(x=0, y=0)\\
                &  + p(y=0) - p(x=0, y=0) \\
                &= p(x=0) + p(y=0) -2 p(x=0, y=0)
\end{align}
\end{proof}
\end{lemma}

\begin{theorema}
Let all instances be independent. With bags $X$, instances $x$, true labels $y,Y$, $h_I$ any classifier
on instances and $h_B$ the resulting bag classifier, we have
\begin{align*}
   e_I = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}
\end{align*}
\begin{proof}

\begin{align*}
e_B & = p(h_B(X) = 0) + p(Y=0) -2 p(h_B(X)=0, Y=0) \\\
&\text{using the lemma}\\
& = h_0^r + p(y=0)^r -2 p(h_I(x)=0, y=0)^r\\\
&\text{by independence}\\
& = h_0^r + p(y=0)^r -2 \left ( \frac{1}{2} (h_0 + p(y=0) - e_I) \right ) ^r \\
&\text{using Lemma~\ref{binarylemma} on }p(h_I(x)=0, y=0.
\end{align*}
Solving for $p(h_I(x)\neq y)$:
\begin{align*}
    & \left ( \frac{1}{2} (h_0 + p(y=0) - e_I )\right ) ^r = \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right )\\
    & \Rightarrow \frac{1}{2} (h_0 + p(y=0) - e_I)   = \left ( \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right ) \right )^\frac{1}{r}\\
    & \Rightarrow e_I   = h_0 + p(y=0) - 2 \left ( \frac{1}{2} \left ( h_0^r + p(y=0)^r - e_B \right ) \right )^\frac{1}{r}
\end{align*}
\end{proof}
\end{theorema}

\begin{corollarya}
If all instances are independent and bag classification is perfect, meaning $e_B=0$, then instance classification is also perfect ($e_I =0$).

\begin{proof}
Consider Theorem~\ref{basicthm}. If $e_B=0$ we have
\begin{align*}
p(h_I(x) \neq y) = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r)) \right)^\frac{1}{r}
\end{align*}
By the convexity of the power function the right hand side of the equation is always greater or equal zero.
Since $e_I \geq 0$, both sides are equal zero.
\end{proof}
\end{corollarya}

\begin{theorema}
    If $h_0^r  \geq e_B$ and all instances are independent, we have
\begin{align}
    p(h_I(x)\neq y) \leq h_0 - \left ( h_0^r - e_B) \right ) ^ \frac{1}{r}
\end{align}
\begin{proof}
    Assume $h_0^r  \geq e_B$.
    From Theorem~\ref{basicthm} we have
\begin{align*}
   e_I &= h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}\\
   &\leq h_0 + p(y=0) - \left [( h_0^r + p(y=0)^r - e_B) \right]^\frac{1}{r}\\
   &- \left ( p(y=0)^r \right )^\frac{1}{r}\\
   &= h_0 - \left (( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}\\
\end{align*}
\end{proof}
\end{theorema}

\begin{theorema}
Let all instances be independent. Given $p(Y)$, the instance error $e_I$ can be calculated by
\begin{align}
    %\eqlabel{iid}
e_I = h_0 + p(Y=0)^\frac{1}{r} - 2 \left (\frac{1}{2} ( h_0^r + p(Y=0) - e_B) \right)^\frac{1}{r}.
\end{align}
\begin{proof}
    This follows from Theorem~\ref{basicthm} by using $p(Y=0)=p(y=0)^r$.
\end{proof}
\end{theorema}

\begin{theorema}
If the instance labels are independent inside a bag given a hidden cause, the instance
error $e_I$ has the following lower bound:
\begin{multline*}
e_I \geq h_0 + p(y=0) - \\
2 \left (\frac{1}{2} (p(h_B(X)=0) + p(Y=0) - e_B) \right)^\frac{1}{r}
\end{multline*}
\begin{proof}
    This follows from taking the expected value with respect to $Z$ of \eqref{conditioned} and using Jensen's
    inequality.
\end{proof}
\end{theorema}

### 2013-08-14 main.tex
ideen fuer zusaetzliches material, intro datei / gliederung
update jmlr paper in thesis
Moving around a lot of sections, renaming conclusion to summary, starting to merge mil and semisupervised, starting to rework structur of pystruct and exact learning.
add some aknowledgements

@@
 \usepackage{color}
-\usepackage{listings}
+\usepackage{minted}
 \usepackage[urw-garamond]{mathdesign}
 \usepackage{garamondx}
-%\usepackage{layout}
+\usepackage{xspace}
 
@@
 \renewcommand{\o}{$\times$}
-\newcommand{\pystruct}{{\sc PyStruct} }
+
+\definecolor{rulecolor}{rgb}{0.80,0.80,0.80}
+\newminted{python}{frame=single,rulecolor=\color{rulecolor}, linenos, xleftmargin=7mm}
+
+\newcommand{\pystruct}{\textsc{PyStruct}\xspace}
+\newcommand{\svmstruct}{SVM$^\text{struct}$\xspace}
 
@@
 \begin{document}
-%\layout
 
@@
 %\listoftables
+\chapter*{Abstract}
+
 \chapter*{Acknowledgements}
+First, I would like to thank my advisor Sven Behnke, who allowed me
+to pursue my PhD in his department and who provided funding for my studies.
+Juergen fuer review?
+
+During my work on this dissertation, my ideas where shape by many of my fellow
+students and researchers. First and foremost I am indept to Hannes Schulz,
+whose feedback proved invaluable during my studies.
+I would like to thank Christoph Lampert, for hosting me as visiting Researcher
+at the IST Austria. His guidance and advice.
+
+Carsten Rother, Sebastian Nowozin
+
+Cristian Sminchisescu?
+Phillip, Victoriia, Aurelien, Theo
+
+I have been lucky to be a part of the scikit-learn community, whos developers
+have been a constant source of encouragement and feedback to me. My understanding
+of machine learning as well as software engineering has been greatly affected
+by the constant exchange with Olivier Grisel, Lars Buitinck, Gael Varoquax,
+Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and others.
+
+Last but not least I am greatful to Anna M\uller for her support, encouragement and company
+Thank you for giving me the strength needed for this endeavour.
 
@@
 
-\chapter{introduction}
+\chapter{Overview}
+\include{intro.tex}
 
@@
 
-\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}
+\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}\label{structured_pystruct}
 
@@
 \bibliography{main}
+\include{mil/appendix}
 \end{document}

### 2013-08-14 exact_learning/paper.tex
somewhat better structure in pystruct part, misc

@@
 
-\section{Discussion}
-In this work we demonstrated that it is possible to learn state-of-the-art CRF models
+\section{Summary}
+In this chapter we demonstrated that it is possible to learn state-of-the-art CRF models
 exactly using structural support vector machines, despite the model containing many loops.

### 2013-08-14 pystruct/paper.tex
somewhat better structure in pystruct part, misc

@@
 
-%The organization of this paper is as follows. Section~\ref{sec:api} revisits
+%The organization of this chapter is as follows. Section~\ref{sec:api} revisits
 %the basic concepts of structured learning algorithms, and how these map to the
@@
 
-\section{Project Goals}\label{sec:goals}
+\subsection{Project Goals}\label{sec:goals}
 
@@
 
-\section{Usage Example: Semantic Image Segmentation}\label{sec:examples}
+\subsection{Usage Example: Semantic Image Segmentation}\label{sec:examples}
 %TODO Multi-label, chain?
@@
 
-\section{Experiments}\label{sec:benchmarks}
+\subsection{Experiments}\label{sec:benchmarks}
 While \pystruct focuses on usability and covers a wide range of applications, it is also
@@
 %TODO summarize structured prediction
-This paper introduced \pystruct, a modular structured learning and prediction library in Python.
+This chapter introduced \pystruct, a modular structured learning and prediction library in Python.
 \pystruct is geared towards ease of use, while providing efficient implementations.

### 2013-08-14 main.tex
put family in aknowledgements
make tables more consistent

@@
 
-Last but not least I am greatful to Anna M\uller for her support, encouragement and company
+I also thank my parents and my sister for their continuing support.
+
+Last but not least I am greatful to Anna M\"uller for her support, encouragement and company
 Thank you for giving me the strength needed for this endeavour.

### 2013-08-14 mil/paper.tex
make tables more consistent

@@
-\section{Theoretical settings}\seclabel{intro}
+\section{Theoretical Analysis of Instance Level Prediction}\seclabel{intro}
+%TODO clean up
 Multiple instance learning is a supervised learning scenario first proposed by \citet{dietterich1997solving}.
@@
 
-\section{Related Work}
+\subsection{Previous Analysis}
 The theoretical analysis of multiple instance learning started with \citet{auer1997approximating} and later \citet{blum1998note}.
@@
 
-\section{Preliminaries}
+\subsection{Preliminaries}
 Throughout this work, $\mathcal{X}$ denotes the space of all instances. Bags are multi sets of instance $X=\{x_1,\dotsc, x_r\}$.
@@
 
-\section{Necessary Assumptions}\seclabel{assumptions}
+\subsection{Necessary Assumptions}\seclabel{assumptions}
 If we allow instances inside bags and their labels to have arbitrary dependencies, we are trying to find
@@
 
-\section{Independent Case}\seclabel{iid}
+\subsection{Independent Case}\seclabel{iid}
 The simplest case in which learning instance prediction is possible is the case when instances are globally i.i.d.,
@@
 
-\section{Restricted Dependent Case}
+\subsection{Restricted Dependent Case}
 As remarked in \Secref{assumptions}, it is not possible to bound the instance classification error

### 2013-08-15 partial_supervision/paper.tex
make tables more consistent
making figures larger

@@
 \begin{equation}
-    y_x = \text{argmax}_{y \in Y} \#\{S_i | p \in S_i \land y_{S_i}=y \},
+    y_x = \argmax_{y \in Y} \#\{S_i | p \in S_i \land y_{S_i}=y \},
 \end{equation}
@@
 	\begin{center}
-        \includegraphics[height=35mm]{images/scheme-crop.pdf}
+        \includegraphics[width=\linewidth]{images/scheme-crop.pdf}
 	\end{center}
@@
     \centering
-    \caption{Bag level performance of various MIL algorithms on the standard Musk
-    datasets. All but MIK provide instance-level labeling. }
-    \begin{tabularx}{0.9\linewidth}{@{\extracolsep{\fill}}lccccccc}
+    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lccccccc}
     \toprule
@@
     \end{tabularx}
+    \caption{Bag level performance of various MIL algorithms on the standard Musk
+    datasets. All but MIK provide instance-level labeling. }
     \tablabel{musk-acc}
@@
     \centering
-    \caption{MIL algorithms and the empirical witness rates of the
-    classifiers.}
-    \begin{tabularx}{.8\linewidth}{@{\extracolsep{\fill}}lcccc}
+    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccc}
     \toprule
@@
     \tablabel{musk-witness}
+    \caption{MIL algorithms and the empirical witness rates of the
+    classifiers.}
 \end{table}
@@
 	\begin{center}
-        \includegraphics[width=28mm]{images/car1_img.png}\hspace*{0.7ex}
-        \includegraphics[width=28mm]{images/car1_gt.png}\hspace*{0.7ex}
-        \includegraphics[width=28mm]{images/car1_pos.png}\hspace*{0.7ex}
-        \includegraphics[width=28mm]{images/car1_neg.png}\hspace*{0.7ex}\\
+        \includegraphics[width=35mm]{images/car1_img.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/car1_gt.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/car1_pos.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/car1_neg.png}\hspace*{0.7ex}\\
         \vspace{1mm}
-        \includegraphics[width=28mm]{images/person1_img.png}\hspace*{0.7ex}
-        \includegraphics[width=28mm]{images/person1_gt.png}\hspace*{0.7ex}
-        \includegraphics[width=28mm]{images/person1_pos.png}\hspace*{0.7ex}
-        \includegraphics[width=28mm]{images/person1_neg.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/person1_img.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/person1_gt.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/person1_pos.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/person1_neg.png}\hspace*{0.7ex}
 	\end{center}
@@
     \centering
-    \caption{Pixel-level accuracy on the Graz-02 dataset.}
-    \begin{tabularx}{0.8\textwidth}{lRRR}
+    \begin{tabularx}{\textwidth}{p{208pt}RRR}
     \toprule
@@
     \end{tabularx}
+    \caption{Pixel-level accuracy on the Graz-02 dataset.}
     \tablabel{graz}

### 2013-08-15 mil/paper.tex
some reviewer comments and layouting in mil part

@@
 Learning to predict instance labels has been attempted in some recent work~\citep{liconvex2010,zhang2002dd}
-but there has been very little theoretical study when and to which extend this is possible.
+but there has been very little theoretical study when and to which extent this is possible.
 In this work, we explore which assumptions about the distributions of instances and labels
@@
 \item[Setting 1:] Bag labels for a validation set are known and instance predictions need to be judged.
-\item[Setting 2:] Bag labels for a test set can be bound using standard learning theoretic bounds and instance
+\item[Setting 2:] Bag labels for a test set can be bounded using standard learning theoretic bounds and instance
 predictions need to be judged.
@@
 These works assume that instances are drawn i.i.d.\ from a single distribution over instances. Blum and Kalai showed that in this
-setting, if the underlying concept is PAC-learnable in polynomial time, then the multiple instance learning problem is PAC-learnable
-in polynomial time. Auer et\,al.\ showed that if there is an algorithm to PAC-learn multiple instance learning with axis-parallel
+setting, if the underlying concept is PAC-learnable in polynomial time, if and only if the multiple instance learning problem is PAC-learnable
+in polynomial time with one-sided noise. Auer et\,al.\ showed that if there is an algorithm to PAC-learn multiple instance learning with axis-parallel
 rectangles that is polynomial in size of the bag and dimension of the instance features, then it is possible to polynomially
@@
 the distribution of labels inside the bags is known a priori.
-Under these assumptions, the instance label prediction error is bound using the bag label error.
+Under these assumptions, the instance label prediction error is bounded using the bag label error.
 This result is used to evaluate the sample complexity of learning from bags using empirical risk minimization.
@@
 \subsection{Preliminaries}
-Throughout this work, $\mathcal{X}$ denotes the space of all instances. Bags are multi sets of instance $X=\{x_1,\dotsc, x_r\}$.
+Throughout this work, $\mathcal{X}$ denotes the space of all instances. Bags are multi sets of instances $X=\{x_1,\dotsc, x_r\}$.
 As usual in supervised learning tasks, we assume our training examples ($X_i,Y_i)$ to be i.i.d.\ from an unknown distribution.
@@
 We assume the bag size $r\in \mathbb{N}$ to be constant over all bags in a given problem. This restriction simplifies the treatment
-significantly while our results can in great parts be carried over to settings with varying bag sizes.
+significantly while our results can in most cases be carried over to settings with varying bag sizes.
 We consider arbitrary classifiers $h_I\colon \mathcal{X} \rightarrow \{0, 1\}$ with associated bag classifier
@@
 $X_i$ and $Y_i$.
-We do not consider specific hypotheses classes but instead try to bound the expected error of $h$ using the expected error of $h_B$.
+We do not consider specific hypothesis classes but instead try to bound the expected error of $h$ using the expected error of $h_B$.
 During all of this work, we assume that the true instance labels are never observed.
@@
 \end{corollary}
-This is an very interesting result in that if we do not assume independence, this corollary is wrong,
+This is a very interesting result in that if we do not assume independence, this corollary is wrong,
 as shown by the example in \Secref{assumptions}.
@@
 \begin{align}\eqlabel{conditioned}
-&p(h_I(x) \neq y | Z) =p(h_I(x)=0|Z) + p(y=0|Z)\\
-&\textstyle- 2  [\frac{1}{2} ( p(h_I(x)=0|Z)^r + p(y=0|Z)^r\\
-& - p(h_B(X) \neq Y|Z)) ]^\frac{1}{r}
+p(h_I(x) \neq y | Z) =&p(h_I(x)=0|Z) + p(y=0|Z)\\
+&\textstyle- 2  [\frac{1}{2} ( p(h_I(x)=0|Z)^r + p(y=0|Z)^r
+ - p(h_B(X) \neq Y|Z)) ]^\frac{1}{r}
 \end{align}
@@
 Using Theorem~\ref{iidbound} and taking expected values with respect to $Z$, we obtain
-\begin{align}
-    & p(h_I(x)\neq y) \\\
-    & \leq h_0 - \mathbb{E}_Z \left [\left ( p(h_I(x)=0 | Z)^r - p(h_B(X)\neq Y|Z) \right ) ^ \frac{1}{r} \right ].
-\end{align}
+\begin{equation}
+    p(h_I(x)\neq y)
+    \leq h_0 - \mathbb{E}_Z \left [\left ( p(h_I(x)=0 | Z)^r - p(h_B(X)\neq Y|Z) \right ) ^ \frac{1}{r} \right ].
+\end{equation}
 
@@
 	\begin{center}
-		\includegraphics[height=40mm]{images/gehler_decision_boundary.png}\hspace{5px}
-		\includegraphics[height=40mm]{images/gehler_instances.png}
+		\includegraphics[width=.40 \linewidth]{images/gehler_decision_boundary.png}\hspace{5px}
+		\includegraphics[width=.40 \linewidth]{images/gehler_instances.png}
 	\end{center}
@@
 error $e_I$ has the following lower bound:
-\begin{multline*}
-e_I \geq h_0 + p(y=0)\\ 
+\begin{equation}
+e_I \geq h_0 + p(y=0)
 - 2 \left (\frac{1}{2} (p(h_B(X)=0) + p(Y=0) - e_B) \right)^\frac{1}{r}.
-\end{multline*}
+\end{equation}
 \end{theorem}
@@
 with a fraction of positive instances of half the witness rate. Then we randomly assign these instances to bags.
-This way, the distribution over instances stays the same as in Gehler-org while instances inside bags become independent.
+Using this procedure the distribution over instances stays the same as in Gehler-org while instances inside bags become independent.
 We refer to this dataset as Gehler-iid.
@@
 rates, the lower bound follows the true error quite closely.
-The upper bound behaves quite different, being quite loose for low instance rates and thighter for high instance rates.
-While the bounds on the true error might not always be very thight, it should be observed that they are clearly
+The upper bound behaves quite different, being quite loose for low instance rates and tighter for high instance rates.
+While the bounds on the true error might not always be very tight, it should be observed that they are clearly
 non-trivial. This is surprising since we made only very general assumptions on the distribution of instances.

### 2013-08-16 itm/paper.tex
make tables more consistent
making figures larger
Add inferred dimensions to the itm paper...

@@
 \begin{algorithm}[t]
-\caption{Information Theoretic MST-based Clustering}
-
 \begin{algorithmic}\label{overall}
@@
 \end{algorithmic}
+\caption{Information Theoretic MST-based Clustering}
+
 \end{algorithm}
@@
 \centering
-\includegraphics[width=.3\textwidth]{illustration_left}
-\includegraphics[width=.3\textwidth]{illustration_center}
-\includegraphics[width=.3\textwidth]{illustration_right}
+\includegraphics[width=.30\textwidth]{illustration_left}
+\includegraphics[width=.30\textwidth]{illustration_center}
+\includegraphics[width=.30\textwidth]{illustration_right}
 \caption{Illustration of the optimization algorithm for $k=3$ on synthetic
@@
 
-We use the dual-tree Boruvka algorithm~\citep{dtb2010} to compute the minimum
-spanning tree, which has runtime close to $O(n \log(n) \alpha(n))$. Here
-$\alpha$ is the inverse of the Ackerman function, which grows so slowly as to
-be considered constant in practice. The dynamic programming solution of
+We use Prim's algorithm combined with a ball tree data structure for distance
+queries to compute the minimum spanning tree of the data.
+While this procedure has no strong guarantees, we found this faster
+in practice than specialized algorithms for euclidean minimum spanning trees, 
+which achive a better theoretical runtim of $O(n \log(n) \alpha(n))$~\citep{dtb2010}. Here
+$\alpha$ is the inverse of the Ackerman function. The dynamic programming solution of
 Algorithm~1 has a run time of $O(n)$ per removed edge, leading to an overall
@@
 
+\subsection{Estimating Intrinsic Dimensionality}
+While assuming totally continuous densities is very natural from a theoretical
+point of view, it can be a hindrance in practical applications.  Often, the
+data is assumed to lie on a lower-dimensional manifold, embedded in a
+higher-dimensional space. In this case, the density is not totally continuous,
+and the dimensionality of the data can not be taken as the dimensionality of
+the embedding space.
+
+A particularly drastic example is the case of the dataset having less samples
+than features. In this case, the data clearly lies even on a linear subspace
+of the input space, and the dimensionality of the input space does not
+accurately reflect the intrinsic dimensionality of the data.
+We use the estimate analyzed in \citet{massoud2007manifold}.
+In their method, for each datapoint $x$, a local estimate $\hat{d}(x)$ of the
+dimensionality at $x$ is computed as
+\begin{equation}
+    \hat{d}(x) = \frac{\ln 2}{\ln(r_k(x) / r_{\lfloor k /2 \rfloor}(x)}.
+\end{equation}
+Here $k$ is a fixed integer and $r_k(x)$ is the distance of $x$ from its $k$th neighbor.
+We follow \citet{massoud2007manifold} and set $k =\lceil 2 \ln n\rceil$.
+The final estimate $\hat{d}$ is then computed by averaging the estimates over all $x$
+\begin{equation}
+    \hat{d} = \frac{1}{n}\sum_{x \in X} \min\left(\hat{d}(x),\ d\right).
+\end{equation}
+We found this estimate to work robustly and give sensible results in all setting.
+As we already used a ball-tree data structure to build the minimum spanning trees,
+estimating the dimensionality resulted in neglectible computational overhead.
+
 \begin{figure}[t]
 \centering
-\includegraphics[width=.2\textwidth]{qualitative_blobs_km}
-\includegraphics[width=.2\textwidth]{qualitative_blobs_mean_nn}
-\includegraphics[width=.2\textwidth]{qualitative_blobs_aggl}
-\includegraphics[width=.2\textwidth]{qualitative_blobs_mst} \\
+\includegraphics[width=.24\textwidth]{qualitative_blobs_km}
+\includegraphics[width=.24\textwidth]{qualitative_blobs_mean_nn}
+\includegraphics[width=.24\textwidth]{qualitative_blobs_aggl}
+\includegraphics[width=.24\textwidth]{qualitative_blobs_mst} \\
 
-\includegraphics[width=.2\textwidth]{qualitative_circles_km}
-\includegraphics[width=.2\textwidth]{qualitative_circles_mean_nn}
-\includegraphics[width=.2\textwidth]{qualitative_circles_aggl}
-\includegraphics[width=.2\textwidth]{qualitative_circles_mst} \\
+\includegraphics[width=.24\textwidth]{qualitative_circles_km}
+\includegraphics[width=.24\textwidth]{qualitative_circles_mean_nn}
+\includegraphics[width=.24\textwidth]{qualitative_circles_aggl}
+\includegraphics[width=.24\textwidth]{qualitative_circles_mst} \\
 
-\includegraphics[width=.2\textwidth]{qualitative_circles2_km}
-\includegraphics[width=.2\textwidth]{qualitative_circles2_mean_nn}
-\includegraphics[width=.2\textwidth]{qualitative_circles2_aggl}
-\includegraphics[width=.2\textwidth]{qualitative_circles2_mst} \\
+\includegraphics[width=.24\textwidth]{qualitative_circles2_km}
+\includegraphics[width=.24\textwidth]{qualitative_circles2_mean_nn}
+\includegraphics[width=.24\textwidth]{qualitative_circles2_aggl}
+\includegraphics[width=.24\textwidth]{qualitative_circles2_mst} \\
 
-\includegraphics[width=.2\textwidth]{qualitative_twomoons_km}
-\includegraphics[width=.2\textwidth]{qualitative_twomoons_mean_nn}
-\includegraphics[width=.2\textwidth]{qualitative_twomoons_aggl}
-\includegraphics[width=.2\textwidth]{qualitative_twomoons_mst} \\ 
+\includegraphics[width=.24\textwidth]{qualitative_twomoons_km}
+\includegraphics[width=.24\textwidth]{qualitative_twomoons_mean_nn}
+\includegraphics[width=.24\textwidth]{qualitative_twomoons_aggl}
+\includegraphics[width=.24\textwidth]{qualitative_twomoons_mst} \\ 
 \caption{Comparison of $k$-Means (left), MeanNN (center left), single link
@@
 used the $k$-Means implementation available in the scikit-learn
-library~\citep{pedregosa2011scikit}.  We use the dual tree Boruvka algorithm
-implemented in the mlpack machine learning library~\citep{mlpack2011}. The source
+library~\citep{pedregosa2011scikit}. The source
 code is available online\footnote{\url{https://github.com/amueller/information-theoretic-mst}}.
@@
 \begin{table}[t]
-\caption{Comparing properties of related algorithms.}\label{nowotab}
 \centering
@@
 \cmidrule{1-4}
-$k$-Means &     $\displaystyle \sum_y \sum_{i, y_i = y} \| x_i - \mu_y \|^2$ & No  & $O(nk)$ per iteration
-%\footnote{The worst-time complexity of the Lloyd-Algorithm is exponential~\cite{arthur2006slow}, though online algorithm are observed to converge in linear time~\cite{bottou1995convergence}.}
+$k$-Means &     $\displaystyle \sum_y \sum_{i, y_i = y} \| x_i - \mu_y \|^2$ & No  & $O(nk)$ per iteration%
+\footnote{The worst-time complexity of the Lloyd-Algorithm is exponential~\cite{arthur2006slow}, though online algorithm are observed to converge in linear time~\cite{bottou1995convergence}.}
 \\
@@
 \end{tabularx}
+\caption{Comparing properties of related algorithms.}\label{nowotab}
 \end{table}
@@
 We evaluated the experiments using the \emph{adjusted Rand index 
-(ARI}~\citep{hubert1985comparing} and \emph{normalized mutual
-information (NMI)}~\citep{strehl2003cluster}, two popular measures 
+(ARI}~\citep{hubert1985comparing}  popular measures 
 of cluster quality~\citep{gomes2010discriminative,kamvar2003spectral}.
@@
 calibration against chance performance.
+We also measured \emph{normalized mutual information
+(NMI)}~\citep{strehl2003cluster}, but do not report it here, as it resulted in
+an identical ranking of the clustering algorithms.
 
@@
 better clusterings than MeanNN in the majority of cases.
+We see that ITM does well when the intrinsic dimensionality of the data matches
+the feature dimension, but degrated otherwise (see ``faces'' and ``usps'').
+Estimating the intrinsic dimensionality of the data overcomes this weakness,
+and improves resuls in most cases. For all but one dataset, eigher ITM or ITM with
+estimated intrinsic dimensionality gives the best results of all considered
+algorithms.
 %
@@
 %
-When inspecting the results, we observed that ITM produced several very small
-clusters on the \emph{faces} dataset. Indeed, increasing the minimum cluster 
-size to 6 or more improved the results to 0.59/0.84. A possible explanation 
-for this is that very small clusters make the entropy estimate less reliable. 
+The run time of computing the ITM clustering was dominated by the computation
+of the MST of the data.
 %
-The single-link method also benefited from this, improving its results to 
-0.42/0.82.
-The run time of computing the ITM clustering was dominated by the computation
-of the MST of the data. The implementation in mlpack took
-60 seconds on a desktop computer for \emph{usps}, the largest dataset in our 
-experiments.
-%
-The other methods had run times in the order of seconds, but given the 
-different implementations we used, this should not be interpreted as a 
-general statement about the speed of the individual methods.
+% TODO runtime statements!
+% One table datasets, one table methods + runtime?
 
 \begin{table}[t]
-\caption{Scores (ARI/NMI) of $k$-Means, MeanNN, single link agglomerative clustering and
-ITM on several benchmark datasets (higher is better). The best score for
-each dataset is printed in bold.}\label{results}
 \centering
-\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lrrrcccc}
+\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lrrrccccc}
 \toprule
@@
 \cmidrule{1-4}
-\cmidrule{5-8}
-Description &     n &    d &   k &  $k$-Means        & MeanNN                & SL                       &   ITM\\
-\cmidrule{1-4}                   \cmidrule{5-5}      \cmidrule{6-6}          \cmidrule{7-7}             \cmidrule{8-8}
-digits      &  1797 &   64 &  10 &     0.62\,/\,0.71 &    0.67\,/\,0.76      &0.10\,/\,0.50             &\B{0.85}\,/\,\B{0.89}\\ 
-faces       &   400 & 4096 &  40 &     0.41\,/\,0.76 & \B{0.49}\,/\,\B{0.80} &0.08\,/\,0.69             &0.02\,/\,0.49\\ %\footnotemark[4]
-iris        &   150 &    4 &   3 &     0.72\,/\,0.76 &    0.75\,/\,0.78      &0.55\,/\,0.72             &\B{0.88}\,/\,\B{0.87}\\
-usps        &  9298 &  256 &  10 &     0.52\,/\,0.61 & \B{0.54}\,/\,\B{0.65} &0.00\,/\,0.04             &   0.44\,/\,0.58\\ 
-vehicle     &   846 &   18 &   4 &     0.10\,/\,0.15 &    0.09\,/\,0.11      &0.00\,/\,0.04             &\B{0.10}\,/\,\B{0.14}\\
-vowel       &   990 &   10 &  11 &     0.17\,/\,0.37 &    0.19\,/\,\B{0.40}  &0.00\,/\,0.16             &\B{0.20}\,/\,0.39\\
-waveform    &  5000 &   21 &   2 & \B{0.37}\,/\,0.35 &    0.30\,/\,\B{0.38}  &0.00\,/\,0.00             &    0.23\,/\,0.22\\
+\cmidrule{5-9}
+Description &     n &    d &   k &  $k$-Means      & MeanNN        & SL            &   ITM        & ITM ID \\
+\cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
+digits      &  1797 &   64 &  10 &     0.62 &      0.67            &0.10          &\B{0.85}       &0.73\\ 
+faces       &   400 & 4096 &  40 &     0.41 &      0.49            &0.08          &0.02           &\B{0.54}\\ 
+iris        &   150 &    4 &   3 &     0.72 &      0.75            &0.55          &\B{0.88}       &0.88\\
+usps        &  9298 &  256 &  10 &     0.52 &      0.54            &0.00          &   0.44        &\B{0.64}\\ 
+vehicle     &   846 &   18 &   4 &     0.10 &      0.09            &0.00          &\B{0.10}       &0.10\\
+vowel       &   990 &   10 &  11 &     0.17 &      0.19            &0.00          &\B{0.20}       &0.19\\
+waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
 \bottomrule
 \end{tabularx}
+\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative clustering and
+ITM on several benchmark datasets (higher is better). The best score for
+each dataset is printed in bold.}\label{results}
 \end{table}
 
-% TODO more experiments: mnist / usps
-% TODO intrinsic dimensionality
+% TODO more experiments: mnist
 % todo segmentation unsupervised
 % TODO bag of word experiment?
+% report times on experiments?
 
@@
 method by formulating the clustering problem as a search over graphs.
-The resulting algorithm is deterministic has sub-quadratic run time.
+The resulting algorithm is deterministic and has sub-quadratic run time.
 %
@@
 balancing cluster sizes and enforcing intra-cluster similarity.
-
-%A limitation of the proposed algorithm is that it is based on the assumption 
-%of an absolute continuous data distribution. This assumption eliminates the 
-%possibility of using categorical variables and data that lies on a submanifold 
-%of the input space. 
+%
+A limitation of the proposed algorithm is that it is based on the assumption 
+of an absolute continuous data distribution. We show that this limitation can
+be overcome in practice by estimating the intrinsic dimensionality of the data. 
 %

### 2013-08-21 exact_learning/paper.tex
make tables more consistent
Add inferred dimensions to the itm paper...
remove all occurences of i.e. and e.g. minor fixes

@@
 \]
-when using inference algorithm $I$, i.e. $o^I(\theta)$ is the approximation of the primal
+when using inference algorithm $I$, that is $o^I(\theta)$ is the approximation of the primal
 objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.
@@
 \begin{enumerate}
-    \item If all constraints $\hat{y}$ in  $\W$ are feasible, i.e.\ generated
+    \item If all constraints $\hat{y}$ in  $\W$ are feasible, that is generated
         by an under-generating or exact inference mechanism, then $o_{\W}$ is
@@
 We can also use these observations to judge the suboptimality of a given parameter $\theta$,
-i.e.\ see how far the current objective is from the true optimum.
+that is see how far the current objective is from the true optimum.
 Learning with any under-generating approach, we can use 1. to maintain a lower bound
@@
     %\]
-%when using inference algorithm $I$, i.e. $o^I(\theta)$ is the approximation of the primal
+%when using inference algorithm $I$, that is $o^I(\theta)$ is the approximation of the primal
+
 %objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.
@@
     \begin{center}
-    \begin{tabular}{lcc}
+    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcc}
         \toprule
@@
     \bottomrule
-    \end{tabular}
+    \end{tabularx}
     \end{center}
@@
 \centering
-\includegraphics[width=.7\linewidth]{caching}
+\includegraphics[width=\linewidth]{caching}
 \caption{%
@@
     \begin{center}
-    \begin{tabular}{lc}
+    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lc}
         \toprule
@@
     \begin{center}
-    \begin{tabular}{lcc}
+    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcc}
     \toprule

### 2013-08-21 itm/paper.tex
remove all occurences of i.e. and e.g. minor fixes

@@
 we will look into a general purpose clustering algorithm.
-
+%
 In clustering, the goal is to divide data points into homogeneous subsets, called
@@
         When using per-image or per-pixel supervision as in the later chapters, we can use clustering to calibrate our
-        expecation of what semisupervised and supervised algorithms have at least to achieve.
+        expecation of what semisupervised and supervised algorithms should be able to achieve.
     \item As many other computer vision algorithms, our segmentation methods
@@
 So not only is clustering one of the fundamental problems in machine learning, it is also
-an imortant building block for the methods in this work.
+an important building block for the methods in this work.
 %
@@
 %
-As a consequence, we obtain a very flexible model that, e.g{.}, allows 
+As a consequence, we obtain a very flexible model that, for example, allows 
 non-convex clusters.
@@
     \item Proposing the use of a MST-based entropy estimator in information theoretic clustering.
-    \item Give a fast algorithm for a relaxed version of the resulting problem.
-    \item Show the practicality on a number of synthetic and real datasets.
+    \item Giving a fast algorithm for a relaxed version of the resulting problem.
+    \item Showing the practicality on a number of synthetic and real datasets.
+    \item Extending the clustering to data on submanifolds by estimating implicit dimensionality.
 \end{itemize}
@@
 \section{Related Work}
-The most commonly used clustering algorithm is the $k$-Means algorithm, also
-known as Lloyd's algorithm \citep{macqueen1967some,lloyd1982least}. While
+The most commonly used clustering algorithm is the $k$-Means algorithm, originally
+investigated by \citet{macqueen1967some} and most commonly implemented using
+Lloyd's algorithm \citep{macqueen1967some,lloyd1982least}. While
 $k$-Means often works well in practice, one of its main drawbacks is the
@@
 %
-They are given by the Voronoi tessellation of the cluster means and therefore always convex.
-Another draw-back is the non-determinism of the procedure, caused by the
-dependence on random initialization.
+Clusters are given by the Voronoi tessellation of the cluster means and
+therefore are always convex.  Another draw-back is the non-determinism of the
+procedure, caused by the dependence on random initialization.
 
@@
 \begin{itemize}
-\item $D_\text{KL} = \int_\mathcal{X} p(X) \ln(\frac{p(X)}{q(X)})dX$ is the Kullback-Leibler divergence,
-\item $H(X) = \int_\mathcal{X} p(X) \ln(p(X))dX$ is the differential entropy, and
-\item $H(X \given Y{=}y) = \int_\mathcal{X} p(X \given Y {=}y) \ln(p(X \given Y {=} y))dX$ is the conditional differential entropy.
+\item $\displaystyle D_\text{KL}\, = \int_\mathcal{X} p(X) \ln\left(\frac{p(X)}{q(X)}\right)\,dX$ is the Kullback-Leibler divergence,
+\item $\displaystyle H(X)\, = \int_\mathcal{X} p(X) \ln(p(X))\,dX$ is the differential entropy, and
+\item $\displaystyle H(X \given Y{=}y)\, = \int_\mathcal{X} p(X \given Y {=}y) \ln(p(X \given Y {=} y))\,dX$ is the conditional differential entropy.
     \end{itemize}
@@
 exactly compute $I(X, Y)$, and this is still true if we fix a set of cluster
-indicators $y_i$ 
+indicators $y_i$.
 %
@@
 \end{align}
-where the probability $p(Y{=}y)$ is given by the empirical frequency of $y$, 
-\mbox{$p(Y\,{=}\,y)= \frac{n_y}{n}$} for $n_y = \frac{|\{i \mid y_i = y \}|}{n}$.
+where the probability $p(Y{=}y)$ is given by the empirical frequency of $y$: 
+\[
+    p(Y\,{=}\,y)= \frac{n_y}{n} \text{,\quad with\quad} n_y = |\{i \mid y_i = y \}|.
+\]
 
@@
 \begin{align}\label{eq:hmst}
-    H_{mst}(\mathbf{x}) = d \log(L) - (d-1) \log(n) + \log(\beta_d).
+    H_{mst}(\mathbf{x}) = d \log(L) - (d-1) \log(n) + \log(\beta_d)
 \end{align}
@@
 for $n \rightarrow \infty$~\citep{hero1999asymptotic}.
-Using \eqref{hmst} as a non-parametric entropy estimate in \eqref{general_objective} yields
+Using \Eqref{hmst} as a non-parametric entropy estimate in \Eqref{general_objective} yields
 the problem to maximize $\hat{I}(\mathbf{x}, \mathbf{y})$ with
@@
 
-\eqref{objective_simplified} has a natural interpretation: The first term
+\Eqref{objective_simplified} has a natural interpretation: The first term
 penalizes long spanning trees, weighted by the size of the cluster. The second
@@
 contrast to the ``loss+regularizer'' setup, where the trade-off needs to be
-specified by the user, the trade-off in \eqref{objective_simplified}, given by
+specified by the user, the trade-off in \Eqref{objective_simplified}, given by
 the factor $d$, is a direct consequence of the entropy estimator.
@@
         \EndFor
-        \For{node $x$ starting from root}
-        \State $w'_x \leftarrow w'_{\text{par}(x)} + w_{\text{par}(x)} - w_x + d(x, \text{par}(x))$
+        \For{node $x$}
+        \State $w'_x \leftarrow w_{\text{root}} - w_x$
         \EndFor
@@
 \subsection{Finding Euclidean Minimum Spanning Tree Clusterings}
-The objective given by \eqref{objective} is a non-linear combinatorial
+The objective given by \Eqref{objective} is a non-linear combinatorial
 optimization problem. It has two properties that make it hard to optimize:
@@
 \centering
-\includegraphics[width=.30\textwidth]{illustration_left}
-\includegraphics[width=.30\textwidth]{illustration_center}
-\includegraphics[width=.30\textwidth]{illustration_right}
+\includegraphics[width=.32\textwidth]{illustration_left}
+\includegraphics[width=.32\textwidth]{illustration_center}
+\includegraphics[width=.32\textwidth]{illustration_right}
 \caption{Illustration of the optimization algorithm for $k=3$ on synthetic
@@
 \emph{Center}: The edge that yields the best two-cluster partition in terms of
-\eqref{objective} was removed, yielding two connected components. \emph{Right}:
+\Eqref{objective} was removed, yielding two connected components. \emph{Right}:
 Another edge from the forest was removed, resulting in the desired number of
@@
 For the above reasons, we propose a simple procedure to approximately solve
-\eqref{objective}.
+\Eqref{objective}.
 %filler removed Let us first fix some notation.
@@
 %
-Then $\hat{I}(G) \geq \hat{I}(\mathbf{x}, \mathbf{y}(G))$, by the definition of
+Then $\hat{I}(G) \geq \hat{I}(\mathbf{y}(G))$, by the definition of
 the minimum spanning tree, and equality holds if and only if $G_y$ is the
@@
 Optimization over forests remains hard, and we further restrict ourself to
-solutions from $\mathcal{G} := \{ F \in \mathcal{F} \mid F \text{\ subgraph of\
+solutions of the form $\mathcal{G} := \{ F \in \mathcal{F} \mid F \text{\ subgraph of\
 } T(\mathbf{x}) \}$ for a given minimum spanning tree $T(\mathbf{x})$, leading to
-the problem $\max_{G \in \mathcal{G}} \hat{I}(G)$.  This restriction allows for
+the problem $\displaystyle \max_{G \in \mathcal{G}} \hat{I}(G)$.  This restriction allows for
 a very fast, combinatorial optimization procedure.
@@
 Starting with the full spanning tree of $\mathbf{x}$, we remove the edge
-yielding the lowest value of \eqref{graphobjective} until the number of
+yielding the lowest value of \Eqref{graphobjective} until the number of
 components equals the number of desired clusters. The overall procedure is
 summarized in Algorithm~1, an illustration can be found in
-\figref{illustration}. We refer to Algorithm~1 as \emph{Information
+\Figref{illustration}. We refer to Algorithm~1 as \emph{Information
 Theoretic MST-based (ITM) clustering}.
@@
 queries to compute the minimum spanning tree of the data.
-While this procedure has no strong guarantees, we found this faster
+While this procedure has no strong runtime guarantees, we found this faster
 in practice than specialized algorithms for euclidean minimum spanning trees, 
-which achive a better theoretical runtim of $O(n \log(n) \alpha(n))$~\citep{dtb2010}. Here
+which achieve a better theoretical runtime of $O(n \log(n) \alpha(n))$ \citep{dtb2010}. Here
 $\alpha$ is the inverse of the Ackerman function. The dynamic programming solution of
@@
 point of view, it can be a hindrance in practical applications.  Often, the
-data is assumed to lie on a lower-dimensional manifold, embedded in a
+data is assumed to lie on a submanifold, embedded in a
 higher-dimensional space. In this case, the density is not totally continuous,
@@
 We use the estimate analyzed in \citet{massoud2007manifold}.
-In their method, for each datapoint $x$, a local estimate $\hat{d}(x)$ of the
+In their method, for each data point $x$, a local estimate $\hat{d}(x)$ of the
 dimensionality at $x$ is computed as
 \begin{equation}
-    \hat{d}(x) = \frac{\ln 2}{\ln(r_k(x) / r_{\lfloor k /2 \rfloor}(x)}.
+    \hat{d}(x) = \frac{\ln 2}{\ln\left(r_k(x) / r_{\lfloor k /2 \rfloor}(x)\right)}.
 \end{equation}
@@
 \end{equation}
-We found this estimate to work robustly and give sensible results in all setting.
-As we already used a ball-tree data structure to build the minimum spanning trees,
-estimating the dimensionality resulted in neglectible computational overhead.
+We found this estimate to work robustly and give sensible results for all
+datasets we investigated.  As we already used a ball-tree data structure to
+build the minimum spanning trees, estimating the dimensionality resulted only in
+little computational overhead.
 
@@
 $k$-Means &     $\displaystyle \sum_y \sum_{i, y_i = y} \| x_i - \mu_y \|^2$ & No  & $O(nk)$ per iteration%
-\footnote{The worst-time complexity of the Lloyd-Algorithm is exponential~\cite{arthur2006slow}, though online algorithm are observed to converge in linear time~\cite{bottou1995convergence}.}
+%\footnote{The worst-time complexity of the Lloyd-Algorithm is exponential~\cite{arthur2006slow}, though online algorithm are observed to converge in linear time~\cite{bottou1995convergence}.}
 \\
@@
 
-%As our objective is motivated by limit properties of the entropy estimate of \eqref{hmst},
+%As our objective is motivated by limit properties of the entropy estimate of \Eqref{hmst},
 %we expect the objective to work best in a setting with many samples. To avoid tautological
@@
 \subsection{Qualitative Results}
-\figref{qualitative} shows qualitative results on three synthetic datasets.
+\Figref{qualitative} shows qualitative results on three synthetic datasets.
 For well separated, convex clusters, all four algorithms produce the same
@@
 We see that ITM does well when the intrinsic dimensionality of the data matches
-the feature dimension, but degrated otherwise (see ``faces'' and ``usps'').
+the feature dimension, but degraded otherwise (see ``faces'' and ``usps'').
 Estimating the intrinsic dimensionality of the data overcomes this weakness,
-and improves resuls in most cases. For all but one dataset, eigher ITM or ITM with
+and improves results in most cases. For all but one dataset, either ITM or ITM with
 estimated intrinsic dimensionality gives the best results of all considered
@@
 waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
+mnist       & 70000 &  784 &  10 &     0.37 &      xxx             &0.00          &   0.50        &0.77\\
 \bottomrule
@@
 \caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative clustering and
-ITM on several benchmark datasets (higher is better). The best score for
+ITM on several datasets (higher is better). The best score for
 each dataset is printed in bold.}\label{results}
@@
 \section{Summary}
-In this summary we proposed the use of a minimum spanning tree based,
+In this chapter we proposed the use of a minimum spanning tree based,
 non-parametric entropy estimator in information theoretic clustering, ITM\@.
@@
 %We will also investigate optimizations of the objective 
-%\eqref{graphobjective} that go beyond the proposed method. 
+%\Eqref{graphobjective} that go beyond the proposed method. 
 %Move-making algorithms seem a promising way to refine solutions 

### 2013-08-21 main.tex
remove all occurences of i.e. and e.g. minor fixes

@@
 \vspace*{1in}
-\textbf{{\LARGE Dissertation}}
+\textbf{{\LARGE On Learning Structured Prediction For Semantic Segmentation}}
 \par
-\vspace{1.5in} {\large Andreas M\"uller}
- \par \vfill A Thesis submitted for the degree of Doctor of Philosophy
+\vspace{1.5in} {\large Andreas Christian M\"uller}
 \par \vspace{0.5in}
-Institute for Computer Science \par
-Department IV, Autonomous Intelligent Systems
+Abteilung VI, Autonomie Intelligente Systeme \par
+Institut f\"ur Informatik
 \par \vspace{0.5in}
@@
 %\listoftables
+\chapter*{Zusammenfassung}
 \chapter*{Abstract}
@@
 to pursue my PhD in his department and who provided funding for my studies.
-Juergen fuer review?
+I would also like to thank J\"urgen Gall for agreeing to be my second reader.
 
-During my work on this dissertation, my ideas where shape by many of my fellow
-students and researchers. First and foremost I am indept to Hannes Schulz,
-whose feedback proved invaluable during my studies.
-I would like to thank Christoph Lampert, for hosting me as visiting Researcher
-at the IST Austria. His guidance and advice.
+During my work on this dissertation, my ideas were shaped by many of my fellow
+students and researchers. First and foremost I am in debt to Hannes Schulz,
+whose feedback proved to be invaluable during my studies.
+I would like to thank Christoph Lampert, for hosting me as a visiting researcher
+at the IST Austria. His guidance and advice helped me in directing my further
+research. My thanks are also extended to Sebastian Nowozin for his collaboration,
+many helpful discussions, and advise.
 
-Carsten Rother, Sebastian Nowozin
+I would also like to thank Carsten Rother for allowing me to work with him
+at Microsoft Research Cambridge and hosting me there.
 
-Cristian Sminchisescu?
-Phillip, Victoriia, Aurelien, Theo
-
-I have been lucky to be a part of the scikit-learn community, whos developers
+I have been lucky to be a part of the scikit-learn community, whose developers
 have been a constant source of encouragement and feedback to me. My understanding
@@
 
-Last but not least I am greatful to Anna M\"uller for her support, encouragement and company
-Thank you for giving me the strength needed for this endeavour.
+Last but not least I am greatful to Anna M\"uller for her support,
+encouragement and company. Thank you for giving me the strength needed for this
+endeavour.
 
@@
 \chapter{Overview}
-\include{intro.tex}
+\include{intro}
 
-\chapter{Information Theoretic Clustering for Object Class Discovey}
+\chapter{Information Theoretic Clustering}
 \include{itm/paper}
@@
 
-\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}\label{structured_pystruct}
+\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}\label{ch:structured_pystruct}
 
@@
 
-\chapter{Hierarchical Latent Variable Models for Semantic Segmentation}
+\chapter{Hierarchical Latent Variable Models for Semantic Segmentation}\label{ch:latent}
 

### 2013-08-21 mil/paper.tex
remove all occurences of i.e. and e.g. minor fixes

@@
 Therefore, as usual in multiple instance learning, we assume there is an underlying concept for instances,
-i.\,e.\ the label of each instance depends only on this instance.
+that is the label of each instance depends only on this instance.
 In other words, we assume $p(y_1,\dotsc, y_r | x_1, \dotsc, x_r)=p(y_1|x_1) \cdot p(y_r|x_r)$.
@@
 The simplest case in which learning instance prediction is possible is the case when instances are globally i.i.d.,
-i.\,e.\ when instances inside a bag are independent of each other. This is a very strong assumption that might be
+that is when instances inside a bag are independent of each other. This is a very strong assumption that might be
 violated in practical applications. On the other hand, strong results can be proved in this setting.
@@
 	\begin{center}
-		\includegraphics[width=.40 \linewidth]{images/gehler_decision_boundary.png}\hspace{5px}
+		\includegraphics[width=.40 \linewidth]{images/gehler_decision_boundary.png}\hspace{35px}
 		\includegraphics[width=.40 \linewidth]{images/gehler_instances.png}

### 2013-08-21 partial_supervision/paper.tex
minor layout issues in psl/mil
remove all occurences of i.e. and e.g. minor fixes

@@
 The goal is to learn a classifier for unseen bags. 
-Formally, let $\mathcal{X}$ be the set of instances. To simply notation, we assume that bags are simply sets, not multi-sets.
+Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we assume that bags are simply sets, not multi-sets.
 Then a bag is an element of the power set $2^\mathcal{X}$ and the task is
 to learn a function
-\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}  \end{equation}
-from a set of training examples of the form $(X_i,y_i)$ with bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.
-The $f_{MI}$ function stems from the so-called underlying concept, given by an (unknown) function
+\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
+Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.
+It is often assumed that the $f_{MI}$ function stems from a so-called underlying concept, given by an (unknown) function
 $f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
@@
 Sometimes, the goal of finding $f_{MI}$ is extended to finding labels not only
-on bag-level but also for all the instances within a bag
-\citep{liconvex2010,zha2008joint}, i.\,e.\ finding $f_{I}$.\\ 
+for new bags, but also for all the separate instances within a bag
+\citep{liconvex2010,zha2008joint}, that is finding $f_{I}$.\\ 
 Even though finding $f_{I}$ is sometimes included in the task statement, there
-has been very little work that actually reported accuracy on instance label prediction. Part
-of the reason for this might be that for many of the datasets used in
-multi-instance learning no ground truth exists.
+has been very little work that actually reported accuracy on instance label
+prediction. Part of the reason for this might be that for many of the datasets
+used in multi-instance learning no ground truth exists.
 
@@
 since the instances are segments, not pixels -- but relates to it closely.
+
+%TODO theoretical part reference?
+% TODO repetition below?
 
@@
 is able to separate bags if and only if the original kernel on instances is able to separate the underlying concepts.
-The method of \citet{gaertner2002multi} has a particular appeal in that it neatly transforms a multi-instance problem into a standard
+The method of \citet{gaertner2002multi} has a particular appeal in that it transforms a multi-instance problem into a standard
 classification problem by changing the kernel. The downside of this approach is that it does
@@
 For computing the global probability of boundary (gPb), we used the CUDA implementation
-of \citet{catanzaro2009efficient}, instead of the original one, for speed.
+of \citet{catanzaro2009efficient}, which provides a speedup of two orders of magnitude over the original implementation.
 
@@
 
+%TODO one classifier per class!
+
 \subsection{Segment Features}
@@
 The framework described above yields an image-level and a segment-level classification.
-In our setup, each segment might be given multiple labels. To obtain a pixel-level object-class segmentation, we have to combine these. 
+In our setup, each segment might be given multiple labels. %TODO explain!
+To obtain a pixel-level object-class segmentation, we have to combine these. 
 When building the segmentation for a given image, we only consider classes whose presence was predicted on image level.
@@
 \end{equation}
-where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an
+where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an %FIXME we didn't describe the dataset yet!
 image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
@@
     \caption{Bag level performance of various MIL algorithms on the standard Musk
-    datasets. All but MIK provide instance-level labeling. }
-    \tablabel{musk-acc}
+    datasets. All but MIK provide instance-level labeling. \tablabel{muskacc}}
+
 \end{table}
@@
 methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
-\tabref{musk-acc}. Results were obtained using 10-fold
+\Tabref{muskacc}. Results were obtained using 10-fold
 cross-validation. While the computational complexity of MIK-instance is
@@
 not trained to provide any instance-level labels.
-
+i
 For multi-class image segmentation, it is beneficial to have a low witness
-rate, i.\,e.\ only a few instances are assumed to be positive in a positive
+rate, that is only a few instances are assumed to be positive in a positive
 bag. Since an object might not be very prominent in an image, only a fraction
 of segments might correspond to the object.
-\tabref{musk-witness} compares the witness rates of MIK-instance,
+\Tabref{musk-witness} compares the witness rates of MIK-instance,
 miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
@@
     \end{tabularx}
-    \tablabel{musk-witness}
     \caption{MIL algorithms and the empirical witness rates of the
-    classifiers.}
+        classifiers.  \tablabel{musk-witness}}
 \end{table}
@@
 setting using a one-vs-all strategy.
-We train one MKL-SVM per class using MIK and predict class labels on segment level
+We train one multiple kernel learning (MKL) SVM per class using MIK and predict class labels on segment level
 using MIK-instance. If at least one SVM classifies a segment as positive,
@@
 or background. We merge segments into pixel-level class labels as described
-in \secref{combiningsegments}.
+in \Secref{combiningsegments}.
 
@@
         \includegraphics[width=35mm]{images/person1_pos.png}\hspace*{0.7ex}
-        \includegraphics[width=35mm]{images/person1_neg.png}\hspace*{0.7ex}
+        \includegraphics[width=35mm]{images/person1_neg.png}
 	\end{center}
@@
     \centering
-    \begin{tabularx}{\textwidth}{p{208pt}RRR}
+    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}p{208pt}ccc}
     \toprule
@@
     \end{tabularx}
-    \caption{Pixel-level accuracy on the Graz-02 dataset.}
-    \tablabel{graz}
+    \caption{Pixel-level accuracy on the Graz-02 dataset.
+        \tablabel{tab:graz}}
 \end{table}
-Per-class pixel accuracies are reported in \tabref{graz}; some qualitative results are shown in \figref{results}. The overall
+Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative results are shown in \figref{results}. The overall
 accuracy on images labels, which is the task that was actually trained, is $90\%$.

### 2013-08-21 pystruct/paper.tex
remove all occurences of i.e. and e.g. minor fixes

@@
 by enumeration of $C^i$ and continue until line~\ref{convergence_check}.
-Only if $\xi' - \xi < \epsilon$, i.e.\ the produced constraint is not violated, we
+Only if $\xi' - \xi < \epsilon$, that is the produced constraint is not violated, we
 return to line~\ref{get_cutting_plane} and actually invoke the separation

### 2013-08-21 make.sh
remove make.sh

@@ -1,1 +1,0 @@
-rubber -c 'setlist arguments --shell-escape'  -I itm/images -I mil/images -I partial_supervision -I mil -I exact_learning -I pystruct --pdf main.tex

### 2013-08-21 itm/paper.tex
Finish off itm :) - meannn can't run out of the box on mnist.

@@
 waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
-mnist       & 70000 &  784 &  10 &     0.37 &      xxx             &0.00          &   0.50        &0.77\\
+mnist       & 70000 &  784 &  10 &     0.37 &      N/A$^\dagger$         &0.00          &   0.50        &0.77\\
 \bottomrule
@@
 ITM on several datasets (higher is better). The best score for
-each dataset is printed in bold.}\label{results}
+each dataset is printed in bold.\\$^\dagger$\footnotesize We were unable to make MeanNN scale to 70000 data points, as storing the
+whole pairwise distance matrix seems critical.\label{results}}
 \end{table}

### 2013-08-21 mil/paper.tex
get rid of shorthands in MIL part, some refinement after first reread.

@@
 %TODO clean up
-Multiple instance learning is a supervised learning scenario first proposed by \citet{dietterich1997solving}.
-In multiple instance learning, examples are multi sets of instances, called bags, and each bag is assigned a class label.
-It is assumed that each instance has a true label, defined by an underlying concept. The label of a bag is positive, if and only if one of
-the instances it contains is labeled positive. The true label of instances is assumed to be unobserved by the learner who only has access to bag labels.
+%Multiple instance learning is a supervised learning scenario first proposed by \citet{dietterich1997solving}.
+%In multiple instance learning, examples are multi sets of instances, called bags, and each bag is assigned a class label.
+%It is assumed that each instance has a true label, defined by an underlying concept. The label of a bag is positive, if and only if one of
+%the instances it contains is labeled positive. The true label of instances is assumed to be unobserved by the learner who only has access to bag labels.
 
-In the original formulation of multiple instance learning, the goal of the learner
-was to predict the label of bags. Some recent works also aim at predicting
-the label of individual instances, a much harder problem.
-This can be viewed as learning an instance-level classifier using a form of weak supervision.
+As we mentioned above the original formulation of multiple instance learning, the goal of the learner
+it to predict only the label of bags, and only recently was this extended to predicting
+the label of individual instances, which is a much harder problem.
+This can also be viewed as learning an instance-level classifier using a form of weak supervision.
 
-Learning to predict instance labels has been attempted in some recent work~\citep{liconvex2010,zhang2002dd}
-but there has been very little theoretical study when and to which extent this is possible.
+Learning to predict instance labels has been attempted in some recent work~\citep{liconvex2010,zhang2002dd}, but
+ there has been very little theoretical study when and to which extent this is possible.
 In this work, we explore which assumptions about the distributions of instances and labels
@@
 There are two applications that we target:
+%TODO clarify
 \begin{description}
@@
 \end{description}
-The first case is very common in practical settings, where
-cross validation is often used to set parameters of a predictor. This is
-particularly important in support vector machines (SVMs) which are a popular
-learning method for multiple instance problems~\citep{andrews2003support}.
-If no instance labels are present for training, it is only possible to
-choose parameters to minimize bag prediction error, which might be suboptimal if one
-wants to minimize instance label misclassification. 
-The bounds can also be used to find the amount of positive instances in positive bags,
-known as the witness rate, which is an integral part of established multiple
-instance algorithms~\citep{zhang2002dd,liconvex2010}.
-The second case can be used to convert established bounds on the bag prediction error
-to instance prediction error.
+The first case is very common in practical settings, where cross validation is
+often used to set parameters of a predictor. This is particularly important in
+support vector machines (SVMs) which are a popular learning method for multiple
+instance problems~\citep{andrews2003support}. If no instance labels are
+present for training, it is only possible to choose parameters to minimize bag
+prediction error, which is suboptimal if one is ultimately interested in
+minimizing instance label misclassification. The bounds we provide can also be used to
+find the amount of positive instances in positive bags, known as the witness
+rate, which is an integral part of established multiple instance
+algorithms~\citep{zhang2002dd,liconvex2010}. The second case can be used to
+convert established bounds on the bag prediction error to instance prediction
+error.
 
+%FIXME what do we actually show?
 We show that it is possible to bound instance prediction error without
@@
 To the authors knowledge, this is the first result of this kind.
-
+%
 Detailed proofs of all propositions can be found in the appendix.
@@
 \subsection{Previous Analysis}
-The theoretical analysis of multiple instance learning started with \citet{auer1997approximating} and later \citet{blum1998note}.
-These works assume that instances are drawn i.i.d.\ from a single distribution over instances. Blum and Kalai showed that in this
-setting, if the underlying concept is PAC-learnable in polynomial time, if and only if the multiple instance learning problem is PAC-learnable
-in polynomial time with one-sided noise. Auer et\,al.\ showed that if there is an algorithm to PAC-learn multiple instance learning with axis-parallel
-rectangles that is polynomial in size of the bag and dimension of the instance features, then it is possible to polynomially
-PAC-learn DNF-formulars, a problem that is solvable only if $\text{RP}=\text{NP}$.
-Recently, \citet{sabato2009homogeneous,DBLP:journals/corr/abs-1107-2021} proved bounds on the VC dimension and
-fat shattering dimension of multiple instance learners by the corresponding dimension for a base learner.
+The theoretical analysis of multiple instance learning started with
+\citet{auer1997approximating} and later \citet{blum1998note}.  These works
+assume that instances are drawn i.i.d.\ from a single distribution over
+instances. Blum and Kalai showed that in this setting, the underlying concept
+is PAC-learnable in polynomial time, if and only if the multiple instance
+learning problem is PAC-learnable in polynomial time with one-sided noise.
+\citet{auer1997approximating} showed that if there is an algorithm to PAC-learn
+multiple instance learning with axis-parallel rectangles that is polynomial in
+size of the bag and dimension of the instance features, then it is possible to
+polynomially PAC-learn DNF-formulars, a problem that is solvable only if
+$\text{RP}=\text{NP}$.
+Recently, \citet{sabato2009homogeneous,DBLP:journals/corr/abs-1107-2021} proved
+bounds on the VC dimension and fat shattering dimension of multiple instance
+learners by the corresponding dimension for a base learner.
 
-All of the results above only consider predicting bag labels. There has been very little theoretical work
-on predicting instance labels. In spite of the large amount of research on multiple instance learning algorithms 
-(for example \citet{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles})
+All of the results above only consider predicting bag labels. There has been
+very little theoretical work on predicting instance labels. In spite of the
+large amount of research on multiple instance learning algorithms (for example
+\citet{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,%
+mangasarian2008multiple,leistner2010miforests,chen2006miles})
 to the authors knowledge, the only quantitative evaluations of predicting
-instance labels are by \citet{gehler2007deterministic} and by \citet{liconvex2010}. While these works show promising results,
-no analysis of instance label prediction is given.
+instance labels are by \citet{gehler2007deterministic} and by
+\citet{liconvex2010}. While these works show promising results, no analysis of
+instance label prediction is given.
 
-Sabato et\,al.\ also studied multiple instance learning as a way
-to reduce labeling complexity~\citep{sabato2010reducing}.
-In the setting of reducing label complexity, the learner creates
-the bags itself. The instances can therefore be assumed to be globally i.i.d.\ and
-the distribution of labels inside the bags is known a priori.
-Under these assumptions, the instance label prediction error is bounded using the bag label error.
-This result is used to evaluate the sample complexity of learning from bags using empirical risk minimization.
-In this work, we generalize some results from \citet{sabato2010reducing} to the setting where the distribution
-of instance labels is unknown and to cases where instances inside a bag may be dependent.
+\citet{sabato2010reducing} also studied multiple instance learning as a way to
+reduce labeling complexity.  In the setting of reducing label complexity, the
+learner creates the bags itself. The instances can therefore be assumed to be
+globally i.i.d.\ and the distribution of labels inside the bags is known a
+priori.
+Under these assumptions, the instance label prediction error is bounded using
+the bag label error.  This result is used to evaluate the sample complexity of
+learning from bags using empirical risk minimization.  In this chapter, we
+generalize some results from \citet{sabato2010reducing} to the setting where
+the distribution of instance labels is unknown and to cases where instances
+inside a bag may be dependent.
 
 \subsection{Preliminaries}
-Throughout this work, $\mathcal{X}$ denotes the space of all instances. Bags are multi sets of instances $X=\{x_1,\dotsc, x_r\}$.
-As usual in supervised learning tasks, we assume our training examples ($X_i,Y_i)$ to be i.i.d.\ from an unknown distribution.
-We denote the label of an instance $x_i$ by $y_i$ and the label of the bag $X_i$ as $Y_i = \max\{y_j | x_j \in  X_i\}$.
-We assume the bag size $r\in \mathbb{N}$ to be constant over all bags in a given problem. This restriction simplifies the treatment
-significantly while our results can in most cases be carried over to settings with varying bag sizes.
-We consider arbitrary classifiers $h_I\colon \mathcal{X} \rightarrow \{0, 1\}$ with associated bag classifier
-$h_B(X) = \max \{h(x_1),\dotsc,h(x_r)\}$.
-The goal is to learn a classifier $h$ which minimizes expected instance misclassification, given training bags and bag labels
-$X_i$ and $Y_i$.
-We do not consider specific hypothesis classes but instead try to bound the expected error of $h$ using the expected error of $h_B$.
-During all of this work, we assume that the true instance labels are never observed.
-We use the following notation: $p(h_I(x)=0):= h_0$, the negative rate of the classifier,
-$p(h_I(x) \neq y) := e_I$, the instance error, and $p(h_B(X) \neq Y) := e_B$, the bag error.
+Throughout this chapter, $\mathcal{X}$ denotes the space of all instances. Bags
+are multi sets of instances $X=\{x_1,\dotsc, x_r\}$.  As usual in supervised
+learning tasks, we assume our training examples ($X_i,Y_i)$ to be i.i.d.\ from
+an unknown distribution.  We denote the label of an instance $x_i$ by $y_i$ and
+the label of the bag $X_i$ as $Y_i = \max\{y_j | x_j \in  X_i\}$.  We assume
+the bag size $r\in \mathbb{N}$ to be constant over all bags in a given problem.
+This restriction simplifies the treatment significantly while our results can
+in most cases be carried over to settings with varying bag sizes.  We consider
+arbitrary classifiers $h_I\colon \mathcal{X} \rightarrow \{0, 1\}$ with
+associated bag classifier $h_B(X) = \max \{h(x_1),\dotsc,h(x_r)\}$.  The goal
+is to learn a classifier $h$ which minimizes expected instance
+misclassification, given training bags and bag labels $X_i$ and $Y_i$.  We do
+not consider specific hypothesis classes but instead try to bound the expected
+error of $h$ using the expected error of $h_B$.  During all of this chapter, we
+assume that the true instance labels are never observed.
 
 \subsection{Necessary Assumptions}\seclabel{assumptions}
-If we allow instances inside bags and their labels to have arbitrary dependencies, we are trying to find
-$y_1,\dotsc,y_r$ maximizing $p(y_1,\dotsc, y_r | x_1, \dotsc, x_r)$. This is a structured 
-prediction problem for which there is no hope solving it, given only bag labels.
-Therefore, as usual in multiple instance learning, we assume there is an underlying concept for instances,
-that is the label of each instance depends only on this instance.
-In other words, we assume $p(y_1,\dotsc, y_r | x_1, \dotsc, x_r)=p(y_1|x_1) \cdot p(y_r|x_r)$.
+If we allow instances inside bags and their labels to have arbitrary
+dependencies, we are trying to find $y_1,\dotsc,y_r$ maximizing $p(y_1,\dotsc,
+y_r | x_1, \dotsc, x_r)$. This is a structured prediction problem for which
+there is no hope solving it, given only bag labels. Therefore, as usual in
+multiple instance learning, we assume there is an underlying concept for
+instances, that is the label of each instance depends only on this instance.
+In other words, we assume $p(y_1,\dotsc, y_r | x_1, \dotsc, x_r)=p(y_1|x_1)
+\cdot p(y_r|x_r)$.
 
-If we allow arbitrary dependencies between the instances $x_i$ in a given bag, there is a simple example 
-\citep{sabato2009homogeneous} that shows that learning instance classification is not necessarily possible.
-Let $\mathcal{X}$ consist of two
-distinct points $x_1$ and $x_2$ with corresponding labels $y_1 = 1$ $y_2 = -1$. If the distribution
-of instances inside bags is such that each bag either contains two copies of $x_1$ or both of $x_1$ and $x_2$, then
-by predicting all bags as positive, one can obtain a perfect bag classifier, while it is impossible to say anything
-about the label of $x_2$.
-Therefore, we need to make additional assumptions about the distribution of instances inside bags to enable
-predicting instance labels.
+If we allow arbitrary dependencies between the instances $x_i$ in a given bag,
+there is a simple example \citep{sabato2009homogeneous} that shows that
+learning instance classification is not necessarily possible. Let
+$\mathcal{X}$ consist of two distinct points $x_1$ and $x_2$ with corresponding
+labels $y_1 = 1$ $y_2 = -1$. If the distribution of instances inside bags is
+such that each bag either contains two copies of $x_1$ or both of $x_1$ and
+$x_2$, then by predicting all bags as positive, one can obtain a perfect bag
+classifier. Even though bag classification is perfect, no statement about the
+label of $x_2$ can sensibly be made.
+Therefore, we need to make additional assumptions about the distribution of
+instances inside bags to make predicting instance labels possible.
 
 \subsection{Independent Case}\seclabel{iid}
-The simplest case in which learning instance prediction is possible is the case when instances are globally i.i.d.,
-that is when instances inside a bag are independent of each other. This is a very strong assumption that might be
-violated in practical applications. On the other hand, strong results can be proved in this setting.
+The simplest case in which learning instance prediction is possible is the case
+when instances are globally i.i.d., that is when instances inside a bag are
+independent of each other. This is a very strong assumption that might be
+violated in practical applications. For example in the above setting of extracting
+segments from an image, the segments within a given image are clearly correlated,
+and often even overlap.
+On the other hand, strong results can be proved using the i.i.d.\ assumption.
 
 \begin{theorem}\label{basicthm}
-Let all instances be independent. With bags $X$, instances $x$, true labels $y,Y$, $h_I$ any classifier
-on instances and $h_B$ the resulting bag classifier, we have
+Let all instances be independent and identically distributed. Let $X$ denote bags, $x$ instances, $y$ true instance labels, and $h_I$ any classifier
+on instances. Let $h_B$ the classifier on bag labels that is induced by $h_I$, that is
+\[
+    h_B(X) = 1 - \prod_{x \in X} (1 - h_I(x)) =
+    \begin{cases}
+        1 \text{\quad if\quad} h_I(x) = 1\ \forall x \in X\\
+        0 \text{\quad else}
+    \end{cases}
+\]
+The we have:
 \begin{align*}
     %\eqlabel{iid}
-   e_I = h_0 + p(y=0) - 2 \left (\frac{1}{2} ( h_0^r + p(y=0)^r - e_B) \right)^\frac{1}{r}.
+   \instanceerror = \hzero + p(y=0) - 2 \left (\frac{1}{2} ( \hzero^r + p(y=0)^r - \bagerror) \right)^\frac{1}{r}.
 \end{align*}
 \end{theorem}
-This is a (slight) generalization of Theorem 1 in \citet{sabato2010reducing} and serves as a basis for all further bounds.
+This is a generalization of Theorem 1 in \citet{sabato2010reducing} and serves as a basis for all further bounds. 
+In particular:
 
-In particular:
 \begin{corollary}\label{perfect}
-If all instances are independent and bag classification is perfect, meaning $e_B=0$, then instance classification is also perfect ($e_I =0$).
+If all instances are independent and bag classification is perfect, meaning $\bagerror=0$, then instance classification is also perfect ($\instanceerror =0$).
 \end{corollary}
@@
 as shown by the example in \Secref{assumptions}.
-
+%
 In practice, $p(y=0)$ is unknown, as we assume a setting where no instance labels are available.
@@
 \begin{theorem}\label{iidbound}
-If $h_0^r  \geq e_B$ and all instances are independent, we have
+If $\hzero^r  \geq \bagerror$ and all instances are independent, we have
 \begin{align}
-    p(h_I(x)\neq y) \leq h_0 - \left ( h_0^r - e_B \right ) ^ \frac{1}{r}
+    p(h_I(x)\neq y) \leq \hzero - \left ( \hzero^r - \bagerror \right ) ^ \frac{1}{r}
 \end{align}
@@
 predictions, without assuming any knowledge of the distribution of instance labels. In this inequality,
-$h_0$ can simply be read off on training and/or test data. The other quantity, $e_B$ can be
+$\hzero$ can simply be read off on training and/or test data. The other quantity, $\bagerror$ can be
 estimated using standard supervised generalization bounds.
@@
 \begin{theorem}\label{equality}
-Let all instances be independent. Given $p(Y)$, the instance error $e_I$ can be calculated by
+Let all instances be independent. Given $p(Y)$, the instance error $\instanceerror$ can be calculated by
 \begin{align}
     %\eqlabel{iid}
-e_I = h_0 + p(Y=0)^\frac{1}{r} - 2 \left (\frac{1}{2} ( h_0^r + p(Y=0) - e_B) \right)^\frac{1}{r}.
+\instanceerror = \hzero + p(Y=0)^\frac{1}{r} - 2 \left (\frac{1}{2} ( \hzero^r + p(Y=0) - \bagerror) \right)^\frac{1}{r}.
 \end{align}
 \end{theorem}
-If bag labels are given, $p(Y=0)$ and $ e_B$ can simply be estimated from validation data.
-This means, in the i.i.d.\ case, instance-level error can simply be calculated from bag label and classifier statistics.
+If bag labels are given, $p(Y=0)$ and $ \bagerror$ can simply be estimated from validation data.
+This means, in the i.i.d.\ case, instance-level error can simply be calculated
+from bag labels and classifier statistics.
 
@@
 \begin{align}
-    p(x_1,\dotsc,x_r)=\int_Z p(x_1)\cdot \dotsc \cdot p(x_r) dZ.
+    p(x_1,\dotsc,x_r)=\int_Z p(x_1)\cdot \dotsc \cdot p(x_r) \, dZ.
 \end{align}
-The intuition behind this assumption is as follows. In the classical case of drug activity prediction~\citep{dietterich1997solving}, a bag corresponds
-to a molecule and each instance corresponds to a different geometric configurations. These are certainly not independent.
-If one knew the ``true'' geometric characteristics of the molecule, each configuration could be derived from that.
+The intuition behind this assumption is as follows.
+For the above example of object class segmentation, one might imagine the class of the present object,
+or even the type of the scene to be this hidden cause. Knowning which class might be present
+in an image disentangles the dependency structure of the segments.
 
-Another popular example of multiple instance learning is image or scene classification~\citep{zhou2007multi,zha2008joint,zhou2009multi}.
-In this case, each image
-is represented as a collection of interest points or segments.
-For the sake of the argument let us assume that interest points or segments correspond to objects.
-Again, objects in a scene are not independent. For example, if we know an image contains a chair and a TV set,
-it is very unlikely to contain a bus.
-On the other hand, if we knew something about the scene, such as ``this is a living room scene'', then the appearance
-of certain objects does not tell us much about other objects.
+Another application with an natural interpretation of the above assumption is
+scene classification~\citep{zhou2007multi,zha2008joint,zhou2009multi}, a very
+popular application of multiple-instance learning.
+In this case, each image is represented as a collection of interest points or
+segments.  For the sake of the argument let us assume that interest points or
+segments correspond to objects.  Again, objects in a scene are not independent.
+For example, if we know an image contains a chair and a TV set, it is very
+unlikely to contain a bus.  On the other hand, if we knew something about the
+scene, such as ``this is a living room scene'', then the appearance of certain
+objects does not tell us much about other objects.
 
-The idea of introducing a hidden cause is similar to the idea of viewing a bag as a manifold and instances as points
-sampled on this manifold, as in \citet{ICML2011Babenko_74}.
+In the classical case of drug activity
+prediction~\citep{dietterich1997solving}, a bag corresponds to a molecule and
+each instance corresponds to a different geometric configurations. These are
+certainly not independent.  If one knew the ``true'' geometric characteristics
+of the molecule, each configuration could be derived from that.
+
+The idea of introducing a hidden cause is similar to the idea of viewing a bag
+as a manifold and instances as points sampled on this manifold, as in
+\citet{ICML2011Babenko_74}.
 
@@
 \begin{corollary}\label{perfect}
-If the bag classification is perfect, meaning $e_B=0$, then instance classification is also perfect.
+If the bag classification is perfect, meaning $\bagerror=0$, then instance classification is also perfect.
 \end{corollary}
@@
     p(h_I(x)\neq y)
-    \leq h_0 - \mathbb{E}_Z \left [\left ( p(h_I(x)=0 | Z)^r - p(h_B(X)\neq Y|Z) \right ) ^ \frac{1}{r} \right ].
+    \leq \hzero - \mathbb{E}_Z \left [\left ( p(h_I(x)=0 | Z)^r - p(h_B(X)\neq Y|Z) \right ) ^ \frac{1}{r} \right ].
 \end{equation}
+Unfortunately, simplifying this expression further requires more assumptions
+about the nature of $Z$ which we will not investigate further here.
 
@@
 \end{figure}
-To get a better handle on the behavior of instance errors, it may also be useful to lower-bound $e_I$.
+To get a better handle on the behavior of instance errors, it may also be useful to lower-bound $\instanceerror$.
 This is possible using Theorem~\ref{basicthm}:
@@
 If the instance labels are independent inside a bag given a hidden cause, the instance
-error $e_I$ has the following lower bound:
+error $\instanceerror$ has the following lower bound:
 \begin{equation}
-e_I \geq h_0 + p(y=0)
-- 2 \left (\frac{1}{2} (p(h_B(X)=0) + p(Y=0) - e_B) \right)^\frac{1}{r}.
+\instanceerror \geq \hzero + p(y=0)
+- 2 \left (\frac{1}{2} (p(h_B(X)=0) + p(Y=0) - \bagerror) \right)^\frac{1}{r}.
 \end{equation}
@@
 As a multiple instance learning algorithm, we choose SVM-SVR from \citet{liconvex2010}, as a state-of-the-art method that
-provides instance predictions. Support vector regression is learned to estimate likelihood ratios of label instances.
+provides instance predictions. In SVM-SVR, support vector regression is learned to estimate likelihood ratios of label instances.
 
@@
 
-\Figref{gehler_iid} compares the true instance error $e_I$, the bag error $e_B$, the estimate of $e_I$ using Theorem~\ref{equality}
+\Figref{gehler_iid} compares the true instance error $\instanceerror$, the bag error $\bagerror$, the estimate of $\instanceerror$ using Theorem~\ref{equality}
 and the bound of Theorem~\ref{iidbound} for datasets with different values of the positive instance rate.
-The $x-axis$ shows different thresholds on estimated likelihood ratios, resulting in varying $h_0$.
+The $x$-axis shows different thresholds on estimated likelihood ratios, resulting in varying $\hzero$.
 It is apparent that the estimate of Theorem~\ref{equality} reflects the true error closely, with only little noise from the restricted
 sample size. Note that the upper bound of Theorem~\ref{iidbound} is not defined everywhere. While it is quite tight when
-$h_I^r \gg e_B$, it is not when $h_I^r \approx e_B$.
+$h_I^r \gg \bagerror$, it is not when $h_I^r \approx \bagerror$.
 
@@
 restricted dependent assumption. \Figref{gehler_soft}
-shows the values of the bound from Theorem~\ref{globalbound} and Theorem~\ref{lowerbound} compared to the
-true instance error $e_I$ for different settings of the positive instance rate.
-Inspecting the lower bound, which uses the same computation as the estimate in the i.i.d.\ case, we see that due to the limited
-dataset size, the estimated lower bound actually lies above the actual error in some cases. On the other hand, for low positive
-rates, the lower bound follows the true error quite closely.
-The upper bound behaves quite different, being quite loose for low instance rates and tighter for high instance rates.
-While the bounds on the true error might not always be very tight, it should be observed that they are clearly
-non-trivial. This is surprising since we made only very general assumptions on the distribution of instances.
+shows the values the lower bound from Theorem~\ref{lowerbound} compared to the
+true instance error $\instanceerror$ for different settings of the positive instance rate.
+Inspecting the lower bound, which uses the same computation as the estimate in
+the i.i.d.\ case, we see that due to the limited dataset size, the estimated
+lower bound actually lies above the actual error in some cases. On the other
+hand, for low positive rates, the lower bound follows the true error quite
+closely.
 
@@
 variable, it is possible to learn instance label prediction from bag labels in
-this setting. We also provided upper and lower bounds on the instance error,
+this setting. We also provided a lower bound on the instance error,
 given known bag labels or an estimage of the bag error.
-Empirical evaluation on a synthetic dataset with known dependen structure and
+Empirical evaluation on a synthetic dataset with known dependency structure and
 known instance labels showed that the bound are non-trivial and yield insight
-into the behavior of instance label errors.  We hope that this work establishes
-a basis for further research on instance label prediction in multiple instance
-learning.
+into the behavior of instance label errors.
+We hope that this analysis establishes a basis for further research on instance
+label prediction in multiple instance learning.

### 2013-08-27 itm/paper.tex
describe use of dimension estimate in itm

@@
 \end{equation}
+We compute the density estimate once, prior to clustering, and then plug the
+estimate $\hat{d}$ into \eqref{graphobjective} in place of $d$.
 We found this estimate to work robustly and give sensible results for all
 datasets we investigated. As we already used a ball-tree data structure to
-build the minimum spanning trees, estimating the dimensionality resulted only in
-little computational overhead.
+build the minimum spanning trees, we can reuse this structure to compute $r_k$.
+Consequently, estimating the dimensionality resulted only in little
+computational overhead.
 
@@
 \end{tabularx}
-\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative clustering and
-ITM on several datasets (higher is better). The best score for
-each dataset is printed in bold.\\$^\dagger$\footnotesize We were unable to make MeanNN scale to 70000 data points, as storing the
-whole pairwise distance matrix seems critical.\label{results}}
+\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative
+    clustering and ITM on several datasets (higher is better). ITM ID refers to
+    ITM using the estimated intrinsic dimensionality. The best score for each
+    dataset is printed in bold.\\$^\dagger$\footnotesize We were unable to make
+    MeanNN scale to 70000 data points, as storing the whole pairwise distance
+    matrix seems critical.\label{results}}
 \end{table}

### 2013-08-28 partial_supervision/paper.tex
Remove MIL theory stuff, start rereading partial supervision part

@@
-Most algorithms for semantic image segmentation and object-class segmentation work with strong supervision: a pixel-wise labeling
-of training images. In this chapter, we will investigate a method that works with annotation
-that is much easier to obtain: whole image labels.
+Most algorithms for semantic image segmentation and object-class segmentation
+work with strong supervision: a pixel-wise labeling of training images. In this
+chapter, we will investigate a method that works with annotation that is much
+easier to obtain: whole image labels.
 
@@
 
-The task of multi-class segmentation is often split up in a segmentation
-and a recognition part. Random forest methods often classify each pixel
-separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
-over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
-computational burden of classifying each pixel separately, but have
-two drawbacks:
-\begin{enumerate}
-    \item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
-    \item Segment boundaries are decided on the lowest level by generating
-        superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
-\end{enumerate}
+%The task of multi-class segmentation is often split up in a segmentation
+%and a recognition part. Random forest methods often classify each pixel
+%separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
+%over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
+%computational burden of classifying each pixel separately, but have
+%two drawbacks:
+%\begin{enumerate}
+    %\item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
+    %\item Segment boundaries are decided on the lowest level by generating
+        %superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
+%\end{enumerate}
 In our approach, we work with a set of candidate segments, generated using constrained parametric min-cuts~\citep{carreira2010constrained}.
@@
 
-We formulate weakly supervised multi-class image segmentation as a multi-instance problem, based upon candidate segments.
+We formulate weakly supervised multi-class image segmentation as a multi-instance problem, based on candidate segments.
 In multi-instance learning~\citep{dietterich1997solving}, each training example is given as a multi-set of instances, called a bag.
@@
 \begin{equation}\eqlabel{multi_instance}
-f_{MI}(X)= \max_{x \in X} f_{I}(x).
+f_{MI}(X)= \max_{x \in X}\ f_{I}(x).
 \end{equation}
@@
 prediction. Part of the reason for this might be that for many of the datasets
-used in multi-instance learning no ground truth exists.
+used in multi-instance learning no ground truth for instances exists.
 
@@
 Since then, many algorithms were proposed to solve the multi-instance learning problem
-using many different approaches \citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
+\citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
 We will discuss only those that are relevant to the present treatment.
@@
 The method of \citet{gaertner2002multi} has a particular appeal in that it transforms a multi-instance problem into a standard
-classification problem by changing the kernel. The downside of this approach is that it does
+classification problem by using an appropriate kernel. The downside of this approach is that it does
 not directly label instances, only bags.
@@
 
-\citet{zhou2009multi} explicitly address non-i.i.d.\  labels, leading to an algorithm that can take
-advantage of correlations inside bags. Computational costs of their algorithm does not scale well with
-the number of instances, although a heuristic algorithm is proposed to overcome this restriction.
-\citet{zhou2009multi} demonstrated only a slight advantage of their algorithm over the MI-kernel of \citet{gaertner2002multi}, so
-we use the MI-kernel for better scalability.
+\citet{zhou2009multi} explicitly address the fact that instances are not
+indpendent within a bag , leading to an algorithm that can take advantage of
+possible correlations. Computational costs of their algorithm does not scale
+well with the number of instances, although a heuristic algorithm is proposed
+to overcome this restriction. 
+%FIXME where does this sentence go?
+\citet{zhou2009multi} demonstrated only a slight
+advantage of their algorithm over the MI-kernel of \citet{gaertner2002multi},
+so we use the MI-kernel for better scalability.
 
-\citet{liconvex2010} compute likelihood ratios for instances, giving a new convex formulation of the multi-instance problem.
-Using these likelihood ratios, classification can be performed directly on the instances, provided an appropriate
-threshold for classifying instances as positive is known.
-We circumvent this problem by applying the same classifier to instances and bags, thereby obtaining hard class
-decisions for each instance.
+\citet{liconvex2010} compute likelihood ratios for instances, giving a new
+convex formulation of the multi-instance problem.  Using these likelihood
+ratios, classification can be performed directly on the instances, provided an
+appropriate threshold for classifying instances as positive is known. We
+circumvent this problem by applying the same classifier to instances and bags,
+thereby obtaining hard class decisions for each instance.
 
@@
 \subsection{Semantic Scene Segmentation via Multi-Instance Learning}
-Recently, several methods have been proposed to obtain semantic segmentations of images using only
-image-level supervision~\citep{vezhnevets2011weakly,verbeek2007region,vezhnevets2010towards}. \citet{vezhnevets2011weakly},
-for example, report impressive results on the MSRC dataset.
+Recently, several methods have been proposed to obtain semantic segmentations
+of images using only image-level
+supervision~\citep{vezhnevets2011weakly,verbeek2007region,vezhnevets2010towards}.
+\citet{vezhnevets2011weakly}, for example, report impressive results on the
+MSRC dataset.
 
-While semantic segmentation is closely related to multi-class image segmentation, there are important
-distinctions: In semantic segmentation, each pixel has a semantic annotation, also containing ``background''
-classes like ``sky'', ``grass'' and ``water''. In multi-class image segmentation, the focus is on
-objects, with possibly large parts of the image being labeled as unspecific ``background''.
-The unspecific background class contains much more clutter than for example ``grass'' and is therefore
-much harder to model.
-This makes disseminating the interesting part in multi-class object recognition challenging, since it
-is not necessary possible to identify non-object regions easily.
+While semantic segmentation is closely related to multi-class image
+segmentation, there are important distinctions: In semantic segmentation, each
+pixel has a semantic annotation, also containing non-object ``stuff'' classes like
+``sky'', ``grass'' and ``water''. In multi-class image segmentation, the focus
+is on objects, with possibly large parts of the image being labeled as
+unspecific ``background''.  The unspecific background class contains much more
+clutter than for example ``grass'' and is therefore much harder to model.
+This makes disseminating the interesting part in multi-class object recognition
+challenging, since it is not necessary possible to identify non-object regions
+easily.
 
@@
 \subsection{Constraint Parametric Min Cuts (CPMC)}
-To generate proposal segments, we use the Constraint Parametric Min Cut (CPCM) method from \citet{carreira2010constrained}. In CPMC, initial
-segments are constructed graph cuts on the pixel grid. The energy function for these cuts uses pixel
-color and the response of the global probability of boundary (gPb) detector~\citep{maire2008using}. As much as ten thousand initial segments are generated from foreground
-and background seeds. A fast rejection based on segment size and ratio cut~\citep{wang2003image} reduced these to about 2000 overlapping
-segments per image. Then, the segments are ranked according to a measure of object-likeness that is based on region and Gestalt properties.
-This ranking is computed using an SVR model~\citep{carreira2010constrained}, which is available online.
-For computing the global probability of boundary (gPb), we used the CUDA implementation
-of \citet{catanzaro2009efficient}, which provides a speedup of two orders of magnitude over the original implementation.
+% Repetition?
+To generate proposal segments, we use the Constraint Parametric Min Cut (CPCM)
+method from \citet{carreira2010constrained}. In CPMC, initial segments are
+constructed using graph cuts on the pixel grid. The energy function for these cuts
+uses pixel color and the response of the global probability of boundary (gPb)
+detector~\citep{maire2008using}. As much as ten thousand initial segments are
+generated from foreground and background seeds. A fast rejection based on
+segment size and ratio cut~\citep{wang2003image} reduces these to about 2000
+overlapping segments per image. Then, the segments are ranked according to a
+measure of object-likeness that is based on region and Gestalt properties.
+This ranking is computed using an SVR model~\citep{carreira2010constrained},
+which is available online. 
+% TODO implementation detail?
+For computing the global probability of boundary
+(gPb), we used the CUDA implementation of \citet{catanzaro2009efficient}, which
+provides a speedup of two orders of magnitude over the original implementation.
 
 \subsection{Multi-Instance Learning using MI-Kernels}
-Since scalability is very important in real-world computer vision
-applications, and natural images might need hundreds of segments to
-account for all possible object boundaries, we use the efficient
-multi-instance kernels~\citep{gaertner2002multi}.
-Multi-instance kernels are a form of set kernels that transform a kernel
-on instance level to a kernel on bag level.
-We reduce the multi-instance multi-class problem to a multi-instance problem by using the one-vs-all approach.
+Since scalability is very important in real-world computer vision applications,
+and natural images might need hundreds of segments to account for all possible
+object boundaries, we use the efficient multi-instance
+kernels~\citep{gaertner2002multi}.  Multi-instance kernels are a form of set
+kernels %TODO add reference to earlier work
+ that transform a kernel on instance level to a kernel on bag level.  We
+reduce the multi-instance multi-class problem to a multi-instance problem by
+using the one-vs-all approach.
 
-With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
+With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the
+corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
 as
@@
 \end{equation}
-where $p \in \mathbb{N}$ is a parameter~\citep{gaertner2002multi}. As we use the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
-are again RBF-kernels, we will not consider $p$ in the following.
+where $p \in \mathbb{N}$ is a free parameter~\citep{gaertner2002multi}. As we use
+the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
+are again RBF-kernels, we will not consider $p$ explicitly in the following.
 
@@
 \end{equation}
-Training an SVM with this kernel produces a bag-level classifier for each class, which we will refer to as MIK.
-This procedure is very efficient since the resulting kernel matrix is of size
-number of bags, which is much smaller than a kernel matrix of size number of
-instances, as is commonly used in the
+Training an SVM with this kernel produces a bag-level classifier for each
+class, which we will refer to as MIK.  This procedure is very efficient since
+the resulting kernel matrix is of size number of bags, which is much smaller
+than a kernel matrix of size number of instances, as is commonly used in the
 literature~\citep{andrews2003support,nguyen2010new,zhang2008m3miml}.  Another
 advantage over other methods is, that it uses a single convex optimization,
-whereas other approaches often use iterative algorithms~\citep{andrews2003support} or need to fit complex
-probabilistic models~\citep{zha2008joint}.
+whereas other approaches often use iterative
+algorithms~\citep{andrews2003support} or need to fit complex probabilistic
+models~\citep{zha2008joint}.
 
@@
 \subsection{Segment Features}
-To describe single segments, we make use of densely computed SIFT~\citep{lowe2004distinctive} and ColorSIFT~\citep{van2009evaluating} features, from which
-we compute bag of visual word histograms. Additionally, we use histograms of oriented gradients~\citep{dalal2005histograms} on the segments.
-We use RBF-kernels for all of the features,
-constructing one MI-kernel per feature. These are then combined using multiple kernel
-learning to produce a single kernel matrix. This kernel matrix can then be used for all classes,
-making classification particularly efficient.
+
+To describe single segments, we make use of densely computed
+SIFT~\citep{lowe2004distinctive} and ColorSIFT~\citep{van2009evaluating}
+features, from which we compute bag of visual word histograms. Additionally, we
+use histograms of oriented gradients~\citep{dalal2005histograms} on the
+segments.  We use RBF-kernels for all of the features, constructing one
+MI-kernel per feature. These are then combined using multiple kernel learning
+to produce a single kernel matrix. This kernel matrix can then be used for all
+classes, making classification particularly efficient.
 
 \subsection{Combining Segments}\seclabel{combiningsegments}
-The framework described above yields an image-level and a segment-level classification.
-In our setup, each segment might be given multiple labels. %TODO explain!
+
+The framework described above yields an image-level and a segment-level
+classification.  In our setup, each segment might be given multiple labels.
+%TODO explain!
 To obtain a pixel-level object-class segmentation, we have to combine these.
-When building the segmentation for a given image, we only consider classes whose presence was predicted on image level.
-Since we do not make use of the ground truth segmentation during training, we cannot learn an optimal combination as 
-in \citet{li2010object} but perform a simple majority vote instead.
-We merge segments into pixel-level class labels by setting
-the label $y_x$ of a pixel $x$ according to
+When building the segmentation for a given image, we only consider classes
+whose presence was predicted on image level.  Since we do not make use of the
+ground truth segmentation during training, we cannot learn an optimal
+combination as in \citet{li2010object} but perform a simple majority vote
+instead.  We merge segments into pixel-level class labels by setting the label
+$y_x$ of a pixel $x$ according to
 \begin{equation}
@@
 \end{equation}
-where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an %FIXME we didn't describe the dataset yet!
+where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an
+%FIXME we didn't describe the dataset yet!
 image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
@@
 \end{figure}
+
 \section{Experiments}
+
 \subsection{Instance-Level Predictions using Multi-Instance Kernels}
-To assess the validity of instance-level predictions using multi-instance kernels, we transform $f_{I}$
-back to an instance-level classifier, using the multi-instance learning
-assumption (\eqref{multi_instance}). We refer to these instance-based MIK predictions
-as MIK-instance. In all experiments, the parameters of the MI-Kernel
-and SVM are adjusted using MIK and then used with both MIK and MIK-instance.
-This facilitates very fast parameter scans since MIK is very efficient to
-compute. Note that we cannot adjust parameters using instance prediction error, as we assume
-no instance labels to be known.
+
+To assess the validity of instance-level predictions using multi-instance
+kernels, we transform $f_{I}$ back to an instance-level classifier, using the
+multi-instance learning assumption (\eqref{multi_instance}). We refer to these
+instance-based MIK predictions as MIK-instance. In all experiments, the
+parameters of the MI-Kernel and SVM are adjusted using MIK and then used with
+both MIK and MIK-instance.  This facilitates very fast parameter scans since
+MIK is very efficient to compute. Note that we cannot adjust parameters using
+instance prediction error, as we assume no instance labels to be known.
 
@@
 methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
-\Tabref{muskacc}. Results were obtained using 10-fold
-cross-validation. While the computational complexity of MIK-instance is
-very low compared to the other methods, it achieves competitive results.
-Using instance-level labels results in a slight loss of accuracy of
-MIK-instances, compared to MIK\@. This small degradation of performance is quite surprising, since the model was
-not trained to provide any instance-level labels.
-i
+\Tabref{muskacc}. Results were obtained using 10-fold cross-validation. While
+the computational complexity of MIK-instance is very low compared to the other
+methods, it achieves competitive results.  Using instance-level labels results
+in a slight loss of accuracy of MIK-instances, compared to MIK\@. This small
+degradation of performance is quite surprising, since the model was not trained
+to provide any instance-level labels.
+
 For multi-class image segmentation, it is beneficial to have a low witness
@@
 miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
-datasets. MIK-instance is able to achieve similar accuracy
-with much less witnesses than the other methods.  Note that Musk1 consists of
-very small bags while Musk2 contains significantly larger bags, more similar to the
+datasets. MIK-instance is able to achieve similar accuracy with much less
+witnesses than the other methods.  Note that Musk1 consists of very small bags
+while Musk2 contains significantly larger bags, more similar to the
 image/segment setup.
+
 \begin{table}
@@
 We evaluate the performance of the proposed algorithm for object-class
-segmentation on the challenging Graz-02 dataset.
-This dataset contains 1096 images of three object classes, bike, car and person.
-Each image may contain multiple instances of the same class.
+segmentation on the challenging Graz-02 dataset.  This dataset contains 1096
+images of three object classes, bike, car and person.  Each image may contain
+multiple instances of the same class.
 
@@
 %feature which are combined using multiple kernel learning.
-We adjusted parameters on a hold-out validation set using bag-level information and used the
-training and test sets as given by the dataset.
-It is straight-forward to extend the binary MIK method to the multi-class
-setting using a one-vs-all strategy.
-We train one multiple kernel learning (MKL) SVM per class using MIK and predict class labels on segment level
-using MIK-instance. If at least one SVM classifies a segment as positive,
-it is associated with the most confident class. Otherwise, it is assigned
-``background'' or no class.
-This yields a classification of each segment into one of four classes: car, bike, person,
-or background. We merge segments into pixel-level class labels as described
-in \Secref{combiningsegments}.
+We adjusted parameters on a hold-out validation set using bag-level information
+and used the training and test sets as given by the dataset.  It is
+straight-forward to extend the binary MIK method to the multi-class setting
+using a one-vs-all strategy.
+We train one multiple kernel learning (MKL) SVM per class using MIK and predict
+class labels on segment level using MIK-instance. If at least one SVM
+classifies a segment as positive, it is associated with the most confident
+class. Otherwise, it is assigned ``background'' or no class.
+This yields a classification of each segment into one of four classes: car,
+bike, person, or background. We merge segments into pixel-level class labels as
+described in \Secref{combiningsegments}.
 
@@
 \end{table}
-Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative results are shown in \figref{results}. The overall
-accuracy on images labels, which is the task that was actually trained, is $90\%$.
-While the performance of our multiple-instance based approach is far
-from current methods that use pixel-level annotations, whose pixel-level accuracy is around $70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level,
-it can serve as a baseline for research on weakly supervised methods for image segmentation.
+
+Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative
+results are shown in \figref{results}. The overall accuracy on images labels,
+which is the task that was actually trained, is $90\%$.  While the performance
+of our multiple-instance based approach is far from current methods that use
+pixel-level annotations, whose pixel-level accuracy is around
+$70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level, it can serve as a
+baseline for research on weakly supervised methods for object class segmentation.
+
+\section{Summary}
+
+We proposed an algorithm for object-class segmentation using only weak
+supervision based on multiple-instance learning. In our approach each image is
+represented as a bag of object-like proposal segments.
+
+We described a way to extent bag level predictions made by the multi-instance
+kernel method to instance level while remaining competitive with the
+state-of-the-art in bag label prediction.
+
+We evaluated the proposed object-class segmentation method on the challenging
+Graz02 dataset. While not reaching the performance of methods using strong
+supervision, our result can work as a baseline for further research into weakly
+supervised object class segmentation.
+
+%In future work, we plan to scale our approach to much larger image datasets. As much
+%more images with weak annotations are available than with pixel level segmentation,
+%we hope that we can improve upon the state-of-the-art in object-class segmentation
+%by making use of larger bodies of training images.

### 2013-08-28 pystruct/paper.tex
serious stab at basics on structured prediction.

@@
 segmentation and visual grouping can be naturally formulated as image labeling tasks.
-
+%
 Arguably the most popular way to approach such labeling problems is via graphical
@@
 modeling spacial dependencies, which are strong in most image-based tasks.
-
+%
 While in earlier approaches, model parameters were set by hand or using
@@
 
-In recent years there has been a wealth of research in methods for learning structured prediction,
-as well as in their application in areas such as natural language processing and computer vision.
-Unfortunately only few implementations are publicly available---many applications are based on
-the non-free implementation of~\citet{joachims2009cutting}.
-
-In this chapter, we will first introduce the concepts and algorithms used in structured prediction,
-in particular in maximum margin methods. We will then accompany this by a description of our
-open source implementation of these algorithms, \pystruct.
-
-\pystruct aims at providing a high-quality implementation with an easy-to-use
-interface, in the high-level Python language.  This allows practitioners to
-efficiently test a range of models, as well as allowing researchers to compare
-to baseline methods much more easily than this is possible with current
-implementations.  \pystruct is BSD-licensed, allowing modification and
-redistribution of the code, as well as use in commercial applications.  By
-embracing paradigms established in the scientific Python community and reusing
-the interface of the widely-used {\sc scikit-learn}
-library~\citep{pedregosa2011scikit}, \pystruct can be used in existing
-projects, replacing standard classifiers.  The online documentation and
-examples help new users understand the somewhat abstract ideas behind
-structured prediction.
-
-%The organization of this chapter is as follows. Section~\ref{sec:api} revisits
-%the basic concepts of structured learning algorithms, and how these map to the
-%\pystruct interface. Section~\ref{sec:examples} exemplifies the use of \pystruct
-%on some standard structured prediction benchmarks and showcasts how \pystruct
-%can interact with existing libraries, foremost scikit-learn and scikit-image.
-%Section~\ref{sec:benchmarks} finally compares \pystruct with SVM$^\text{struct}$,
-%arguably the most widely used implementation of structural SVMs, and
-%SVM$^\text{python}$, a Python interface or the same.
+In recent years there has been a wealth of research in methods for learning
+structured prediction, as well as in their application in areas such as natural
+language processing and computer vision (see \citet{nowozin2011structured} for
+an overview).
+%
+In this chapter, we will first introduce the concepts and algorithms used in
+structured prediction, in particular in maximum margin methods. We will then
+accompany this by a description of our open source implementation of these
+algorithms, \pystruct.
 
@@
 Structured prediction can be defined as making a prediction $f(x)$ by maximizing a
-compati\-bility function between an input $x$ and the possible labels
-$y$~\citep{nowozin2011structured}. Most current approaches use linear functions,
-leading to:
-\begin{equation}\label{eq:main_equation}
-    f(x) = \argmax_{y \in \mathcal{Y}}\  \theta^T \Psi(x, y).
+compatibility function $g$ between an input $x$ and possible labels
+$y$~\citep{nowozin2011structured}:
+\begin{equation}
+    f(x) = \argmax_{y in \mathcal{Y}} g(x, y)
 \end{equation}
-Here, $y$ is a \emph{structured} label, $\Psi$ is a joint feature function of $x$ and $y$, and
-$\theta$ are parameters of the model. \emph{Structured} means that $y$ is more complicated than a single
-output class, for example a label for each word in a sentence or a label for
-each pixel in an image.
+While using arbitrary functions is possible, and was expored
+in some recent works % TODO cite dtf, conditional neural fields
+most current approaches use linear functions, leading to:
+\begin{equation}\eqlabel{main_equation}
+    f(x) = \argmax_{y \in \mathcal{Y}}\  \theta^T \Phi(x, y).
+\end{equation}
+Here, $y$ is a \emph{structured} label, $\Phi$ is a joint feature function of
+$x$ and $y$, and $\theta$ are parameters of the model. \emph{Structured} means
+that $y$ is more complicated than a single output class, for example a label
+for each word in a sentence or a label for each pixel in an image.
 Learning structured prediction means learning the parameters $\theta$ from training data.
+The particular model that is used is completely encoded in $\Phi(x, y)$, which manifests
+the relation between input $x$ and output $y$. As $\mathcal{Y}$ is typically very large,
+it is crucial to exploit the particular form of $\Phi(x, y)$ to solve the maximization problem
+in \Eqref{main_equation}.
 
-% TODO relate to computer vision here
-When learning for structured prediction in the max-margin framework of
-\cite{tsochantaridis2006large}, predictions are made as
-\[ \argmax_y f(x, y, \theta)\]
-where x is the input and $\theta$ are the parameters to be learned.
-Here, $f$ is a compatibility function of $x$ and $y$, that is linear in the parameters
-\[f(x, y, \theta) = \langle \theta, \Psi(x, y) \rangle\]
-where $\Psi(x, y)$ is a joint feature vector of $x$ and $y$. Specifying a specific CRF model
-amounts to specifying $\Psi$.
+While $y$ could be a complicated like a parsing tree or the geometric configuration of a molecule,
+many setting, such as the image segmentation setting we are interested, can be reduced to the case
+where $y$ is a vector of discrete labels $\mathcal{Y} = \{0, \dotsc, k\}^n$. %TODO n varies.
+In the following, we will only discuss this multivatiate case. In general, $n$ is often different
+for different inputs $x$, such as image with different number of (super) pixels.
+We will ignore this in our notation to simplify the presentation.
 
+\subsection{Factor Graphs and the Relation to Graphical Models}
+In the case when $y$ is multivariate, a very general and widely used method to
+specify the structure of a model, and therefore $\Phi$, is using \emph{factor
+graphs}. A factor graph is a bipartite graph $(\mathcal{V}, \mathcal{F},
+\mathcal{E})$, consisting of variable nodes $\mathcal{V}$, factor nodes
+$\mathcal{F}$ and edges $\mathcal{E}$ connecting variables to factors. The
+\emph{scope} $N_F$ of a factor $F \in \mathcal{F}$ is defined as
+\begin{equation}
+    N_F = \{ v \in \mathcal{V} \,|\, (v,F) \in \mathcal{E} \}
+\end{equation}
+The variable nodes of the factor graph correspond to the entries of the variables $y$,
+that is $\mathcal{V} = \{0, \dotsc, n\}$, and each factor node is associated with a \emph{factor} or \emph{potential function}
+$\psi_F$.
+A factor graph represents a function\footnote{Traditionally factor graphs represent \emph{products} of factors.
+To simplify presentation, we work directly in the log-domain of the product representation.}
+\begin{equation}\eqlabel{log_factor_graph}
+    g(x, y) = \sum_{F \in \mathcal{F}} \psi_F(x, y_{N_F})
+\end{equation}
+Here, $y_{N_F}$ denotes the entries of $y$ indexed by $N_F$.
+
+The benefit of using the factor graph representation is that it decomposes the
+function over the variables of interest $y_i$ and makes. This allows us to
+apply efficient optimization procedures for the maximization in
+\Eqref{mainequation} by exploiting the graph structure of the factor graph.
+
+To obtain a linear function as in \Eqref{mainequation} from
+\Eqref{log_factor_graph}, we can simply let each $\psi$ be of the form
+\begin{equation}\eqlabel{general_psi}
+    \psi_F(x, y_{N_F}) = \theta_F^T \Phi_F(x, y_{N_F}).
+\end{equation}
+The by far most common form is 
+\begin{equation}\eqlabel{linear_psi}
+    \psi_F(x, y_{N_F}) = \theta_{F, y_{N_F}}^T \phi_F(x),
+\end{equation}
+where $\phi_F(x)$ is a vector representation of the input $x$, and there are
+different parameter vectors $\theta_{F, y_{N_F}}$ for each possible assignment
+of $y_{N_F}$.
+
+Both, \Eqref{general_psi} and \Eqref{linear_psi} are instantiations of the general
+linear form \Eqref{main_equation}. To see this, for \Eqref{general_psi} we simply
+concatenate the individual components for all $f \in \mathcal{F}$:
+\begin{align}
+    \theta &= \bigoplus_{F \in \mathcal{F}} \theta_F\\
+    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \Phi_F(x, y)
+\end{align}
+%TODO linear psi explicitly
+
+The benefit of using the factor graph representation is that it allows us to
+apply efficient optimization procedures for the maximization in
+\Eqref{mainequation} by exploiting the graph structure of the factor graph.
+
+This approach to structured prediction is closely related to a probabilistic
+approach using \emph{graphical models}.  Probabilistical graphical models are a
+tool to express factorizations of probability distributions.  Similar to
+\Eqref{log_factor_graph}, the joint probability distribution over a
+multi-variate random variable $y$ can be expressed using a factor-graph:
+\begin{equation}\eqlabel{graphicalmodel}
+    p(y | x) = \frac{1}{Z_x}\prod_{F \in \mathcal{F}} \exp\left(\psi_F(x, y_{N_F})\right)
+\end{equation}
+where
+\begin{equation}
+    Z_x = \sum_{y' \in \mathcal{Y}} \prod_{F \in \mathcal{F}} \exp\left(\psi(x, y'_{N_F})\right)
+\end{equation}
+is the normalization constant of the conditional distribution over $y$.
+If $f$ is choosen as in \Eqref{linear_psi}, then the resulting distribution
+belongs to the exponential family.
+
+The most probable prediction $y$ is given as $\argmax_{y \in \mathcal{Y}}p(y|x)$.
+As $Z$ is independent of $y$, and by the monotonicity of the logarithm,
+maximizing $p(y|x)$ is equivalent to maximizing $g(x,y)$ over $y$ in
+\Eqref{log_factor_graph}. Therefore, from a prediction standpoint, the two
+formulations are equivalent.
+
+During learning, the presence of the factor $Z$ in \Eqref{graphicalmodel}
+introduces additional complications. As we only address the problem of making
+predictions, not modeling probabilities, there are no clear benefits from the
+probabilistic approach.  Consequently, we will work with the more direct
+structured prediction approach of \Eqref{mainequation} and
+\Eqref{log_factor_graph} instead.
+
+% TODO relate to computer vision here ??? Pairwise models?
+
+\subsection{The Maximum Margin Approach to Structured Learning}
+% margin BLABLA here?
 For a given loss $\Delta$, the parameters $\theta$ are learned by minimizing
@@
 bound on $\Delta$ empirical risk:
-\[r_i(x_i, y_i, y) = [\max_{y \in \mathcal{Y}} \Delta(y_i, y) + \left<\theta, \psi(x_i, y) - \psi(x_i, y_i) \right >]_+ \]
+\[r_i(x_i, y_i, y) = [\max_{y \in \mathcal{Y}} \Delta(y_i, y) + \left<\theta, \phi(x_i, y) - \phi(x_i, y_i) \right >]_+ \]
 \citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
 \begin{equation}\label{subgradient}
-    g := \psi(x_i, y) - \psi(x_i, y_i) + \theta / C
+    g := \phi(x_i, y) - \phi(x_i, y_i) + \theta / C
 \end{equation}
@@
 \subsection{The Cutting Plane Method}\label{cutting_plane}
-
-When learning for structured prediction in the max-margin framework of
-\citet{tsochantaridis2006large}, predictions are made as
-\[ \argmax_{y \in \mathcal{Y}} f(x, y, \theta),\]
-where $x \in \mathcal{X}$ is the input, $y \in \mathcal{Y}$ the prediction, and $\theta$ are the parameters to be learned.
-We will assume $y$ to be multivariate, $y=(y_1, \dots, y_k)$ with possibly varying $k$.
-
-The function $f$ measures compatibility of $x$ and $y$ and is a linear function of the parameters $\theta$:
-\[f(x, y, \theta) = \langle \theta, \psi(x, y) \rangle.\]
-Here $\psi(x, y)$ is a joint feature vector of $x$ and $y$. Specifying a particular SSVM model
-amounts to specifying $\psi$.
 
@@
 bound on the empirical $\Delta$-risk:
-\[r_i(x^i, y^i, y) = \left [\max_{y \in \mathcal{Y}} \Delta(y^i, y) + \left<\theta, \psi(x^i, y) - \psi(x^i, y^i) \right > \right]_+ \]
+\[r_i(x^i, y^i, y) = \left [\max_{y \in \mathcal{Y}} \Delta(y^i, y) + \left<\theta, \phi(x^i, y) - \phi(x^i, y^i) \right > \right]_+ \]
 
-We solve the following reformulation of Equation~\eqref{mainequation}, known as one-slack QP formulation:
+We solve the following reformulation of Equation~\Eqref{mainequation}, known as one-slack QP formulation:
 \begin{align}\label{eq:oneslack}
@@
     \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
-        &\left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i,
-            \hat{y}^i)] \right \rangle \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
+        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
+            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
             - \xi
@@
 
-The cutting plane method alternates between solving Equation~\eqref{oneslack}
+The cutting plane method alternates between solving Equation~\Eqref{oneslack}
 with a working set $\W$ of constraints, and expanding the working set using the
@@
 \[ 
-    \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y})] \right \rangle
+    \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i, \hat{y})]
 \]
@@
 
-In computer vision applications, or more generally in graph labeling problems,
-$\psi$ is often given as a factor graph over $y$, typically using only unary and pairwise functions:
-\[ \langle \theta, \psi(x, y) \rangle = \sum_{i=0}^k \langle \theta_u,  \psi_u(x, y_k) \rangle + \sum_{(i, j) \in E} \langle \theta_p, \psi_p(x, y_k, y_l) \rangle, \]
-where $E$ a set of pairwise relations. In this form, parameters $\theta_u$ and $\theta_p$ for unary and
-pairwise terms are shared across all entries and pairs.
-The decomposition of $\psi$ into only unary and pairwise interactions allows
-for efficient inference schemes, based on message passing or graph cuts.
-
 \section{Casting Structured Prediction into Software}\label{sec:api}
+Unfortunately only few implementations are publicly
+available---many applications are based on the non-free implementation
+of~\citet{joachims2009cutting}.
+Our implementation, \pystruct, aims at providing a high-quality implementation
+with an easy-to-use interface, in the high-level Python language.  This allows
+practitioners to efficiently test a range of models, as well as allowing
+researchers to compare to baseline methods much more easily than this is
+possible with current implementations.  \pystruct is BSD-licensed, allowing
+modification and redistribution of the code, as well as use in commercial
+applications.  By embracing paradigms established in the scientific Python
+community and reusing the interface of the widely-used {\sc scikit-learn}
+library~\citep{pedregosa2011scikit}, \pystruct can be used in existing
+projects, replacing standard classifiers.  The online documentation and
+examples help new users understand the somewhat abstract ideas behind
+structured prediction.
 
@@
     \item Optimizing the objective with respect to $\theta$.
-    \item Encoding the structure of the problem in a joint feature function $\Psi$.
+    \item Encoding the structure of the problem in a joint feature function $\Phi$.
     \item Solving the maximization problem in Equation~\ref{eq:main_equation}.
@@
 Equation~\ref{eq:main_equation} is usually only feasible by exploiting the
-structure of $\Psi$, while the first one is usually treated as independent.
+structure of $\Phi$, while the first one is usually treated as independent.
 \pystruct takes an object-oriented approach to decouple the task-dependent
@@
 Encoding the structure of the problem is done using \texttt{model} classes, which
-compute $\Psi$ and encode the structure of the problem.
+compute $\Phi$ and encode the structure of the problem.
 \pystruct implements models for many common cases, such as multi-class and
@@
 \paragraph{Modularity} \pystruct separates the algorithms of parameter estimation and
-     inference from the task-dependent formulation of $\Psi$. This allows
+     inference from the task-dependent formulation of $\Phi$. This allows
      practitioners, for example in computer vision or natural language
@@
 %% listing, training time, accuracy, comparision with kraehenbuehl
+%TODO snakes
 

### 2013-08-28 pystruct/paper.tex
more on algorithms for structured learning, but broke the algorithm listings?

@@
 \begin{equation}
-    f(x) = \argmax_{y in \mathcal{Y}} g(x, y)
+    f(x) = \argmax_{y \in \mathcal{Y}} g(x, y)
 \end{equation}
+The computation of $y$ in the above equation is often referred to as inference
+or prediction.
 While using arbitrary functions is possible, and was expored
@@
 the relation between input $x$ and output $y$. As $\mathcal{Y}$ is typically very large,
-it is crucial to exploit the particular form of $\Phi(x, y)$ to solve the maximization problem
-in \Eqref{main_equation}.
+it is crucial to exploit the particular form of $\Phi(x, y)$ to solve the inference problem
+of \Eqref{main_equation}.
 
-While $y$ could be a complicated like a parsing tree or the geometric configuration of a molecule,
-many setting, such as the image segmentation setting we are interested, can be reduced to the case
-where $y$ is a vector of discrete labels $\mathcal{Y} = \{0, \dotsc, k\}^n$. %TODO n varies.
+While $y$ could be a complicated like a parsing tree or the geometric
+configuration of a molecule, many setting, such as the image segmentation
+setting we are interested, can be reduced to the case where $y$ is a vector of
+discrete labels $\mathcal{Y} = \{1, \dotsc, k\}^n$.
 In the following, we will only discuss this multivatiate case. In general, $n$ is often different
@@
 \subsection{Factor Graphs and the Relation to Graphical Models}
+% one factor graph per instance?
 In the case when $y$ is multivariate, a very general and widely used method to
@@
 \end{equation}
-The variable nodes of the factor graph correspond to the entries of the variables $y$,
-that is $\mathcal{V} = \{0, \dotsc, n\}$, and each factor node is associated with a \emph{factor} or \emph{potential function}
-$\psi_F$.
-A factor graph represents a function\footnote{Traditionally factor graphs represent \emph{products} of factors.
-To simplify presentation, we work directly in the log-domain of the product representation.}
+The variable nodes of the factor graph correspond to the entries of the
+variables $y$, that is $\mathcal{V} = \{1, \dotsc, n\}$, and each factor node
+is associated with a \emph{factor} or \emph{potential function} $\psi_F$.
+A factor graph represents a function\footnote{Traditionally factor graphs
+ represent \emph{products} of factors.  To simplify presentation, we work
+directly in the log-domain of the product representation.}
 \begin{equation}\eqlabel{log_factor_graph}
@@
 apply efficient optimization procedures for the maximization in
-\Eqref{mainequation} by exploiting the graph structure of the factor graph.
+\Eqref{main_equation} by exploiting the graph structure of the factor graph.
 
-To obtain a linear function as in \Eqref{mainequation} from
+To obtain a linear function as in \Eqref{main_equation} from
 \Eqref{log_factor_graph}, we can simply let each $\psi$ be of the form
@@
 different parameter vectors $\theta_{F, y_{N_F}}$ for each possible assignment
-of $y_{N_F}$.
+of $y_{N_F} \in \mathcal{Y}_{N_F}$.
 
@@
     \theta &= \bigoplus_{F \in \mathcal{F}} \theta_F\\
-    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \Phi_F(x, y)
+    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \Phi_F(x, y_{N_F})
 \end{align}
-%TODO linear psi explicitly
+Writing down $\Phi$ and $\theta$ for the form \Eqref{linear_psi} is a little less compact:
+\begin{align}
+    \theta &= \bigoplus_{F \in \mathcal{F}}\ \bigoplus_{y_{N_F} \in \mathcal{Y}_{N_F}} \theta_{F, y_{N_F}}\\
+    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \left (\phi_F(x) \otimes e_{y_{N_F}} \right )
+\end{align}
+where $e_{y_{N_F}} \in \mathbb{R}^{|\mathcal{Y}_{N_F}|}$ is the indicator for a
+given variable setting $y_{N_F}$.
+In words, $\Phi_F$ is build simply by creating a vector of
+$|\mathcal{Y}_{N_F}|$ times the size of $\phi_F$, which is zero everywhere,
+except for the place corresponding to $y_{N_F}$. %TODO unclear!
 
@@
 apply efficient optimization procedures for the maximization in
-\Eqref{mainequation} by exploiting the graph structure of the factor graph.
+\Eqref{main_equation} by exploiting the graph structure of the factor graph.
 
@@
 probabilistic approach.  Consequently, we will work with the more direct
-structured prediction approach of \Eqref{mainequation} and
+structured prediction approach of \Eqref{main_equation} and
 \Eqref{log_factor_graph} instead.
@@
 % TODO relate to computer vision here ??? Pairwise models?
-
-\subsection{The Maximum Margin Approach to Structured Learning}
-% margin BLABLA here?
-For a given loss $\Delta$, the parameters $\theta$ are learned by minimizing
-the loss base soft-margin objective
-\begin{equation}\label{mainequation}
-    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  r_i(\theta)
-\end{equation}
-with regularization parameter $C$, where $r_i$ is an hinge-loss like upper
-bound on $\Delta$ empirical risk:
-\[r_i(x_i, y_i, y) = [\max_{y \in \mathcal{Y}} \Delta(y_i, y) + \left<\theta, \phi(x_i, y) - \phi(x_i, y_i) \right >]_+ \]
-\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
-\begin{equation}\label{subgradient}
-    g := \phi(x_i, y) - \phi(x_i, y_i) + \theta / C
-\end{equation}
-leading to a simple iterative optimization.
+% talk about inference?
 
 \section{Learning Algorithms for Linear Max-Margin Structured Prediction}
-% TODO algorithms here
-\subsection{The Cutting Plane Method}\label{cutting_plane}
+Maximum Margin learning is now one of the most popular methods to learn
+structural models in computervision and in other fields.
+There are several reasons for the popularity of linear maximum margin approaches:
+\begin{description}
+    \item[Loss-sensitivity] In contrast to probabilistic approaches, maximum margin learning approaches can directly
+        minimize a user-specified loss.
+    \item[Feasibility] If the loss decomposes over the factor graph that
+        specifies $g$, then learning is feasible as soon as the maximization
+        over $y$ in \Eqref{main_equation} can be caried out.
+    \item[Generalization] The maximum margin principle yields generalization
+        bounds using the effective complexity~\citep{taskar2003max}.
+    \item[Strong Convexity] The resulting optimization problem is strongly
+        convex, leading to efficient optimization and unique solutions.
+\end{description}
 
-For a given loss $\Delta$, the parameters $\theta$ are learned by minimizing
-the loss-based soft-margin objective
-\begin{equation}\label{eq:mainequation}
-\min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  r_i(\theta)
+For learning, a dataset $(x^{(1)}, y^{(1)}),\dotsc,(x^{(k)}, y^{(k)})$, together with a loss
+\begin{equation}
+    \Delta \colon \mathcal{Y} \times \mathcal{Y} \rightarrow \mathbb{R}.
 \end{equation}
-with regularization parameter $C$, where $r_i$ is a hinge-loss-like upper
-bound on the empirical $\Delta$-risk:
-\[r_i(x^i, y^i, y) = \left [\max_{y \in \mathcal{Y}} \Delta(y^i, y) + \left<\theta, \phi(x^i, y) - \phi(x^i, y^i) \right > \right]_+ \]
+The parameters $\theta$ are learned by minimizing the loss based soft-margin
+objective
+\begin{equation}\eqlabel{learning_equation}
+    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x^{(i)}, y^{(i)}, \theta)
+\end{equation}
+with regularization parameter $C$. Here $r_i$ is an hinge-loss like upper bound
+on $\Delta$ empirical risk:
+\begin{equation}\eqlabel{loss_augmentation}
+    \ell(x^{(i)}, y^{(i)}, \theta)= [\max_{y \in \mathcal{Y}} \Delta(y^{(i)}, y) + \theta^T \phi(x^{(i)}, y) - \theta^T \phi(x^{(i)}, y^{(i)}]_+
+\end{equation}
 
-We solve the following reformulation of Equation~\Eqref{mainequation}, known as one-slack QP formulation:
+This is an instance of regularized empirical risk minimization, with a
+piecewise linear, convex upper bound on the loss. 
+Finding the $y$ that corresponds to a maximum in \Eqref{loss_augmentation} is
+a central part of all maximum-margin based learning algorithms, and is referred to
+as loss-augmented prediction. For complex models, such as the ones used for
+image segmentation, this optimization often dominates the learning process in
+terms of computational complexity.
+Therefore, it is often desireable to find learning algorithms that converge with
+as little optimizations of the loss-augmented prediction problem as possible.
+
+There are several popular algorithms to solve \Eqref{learning_equation}. We
+will briefly review three standard algorithms: the one-slack and n-slack
+cutting plane algorithms, a stochastic primal subgradient algorithm, and
+recently proposed stochastic dual coordinate descent method.
+
+\subsection{Stochastic Subgradient Descent}\seclabel{subgradient}
+Arguably the most straight-foward way to approach \Eqref{learning_equation} is using subgradient descent.
+In light of the complexity of solving the loss-augmented prediction problem in \Eqref{loss_augmentation},
+is is natural to work in a stochastic setting (see \citet{ratliff2007online}).
+%\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
+%\begin{equation}\label{subgradient}
+    %g := \phi(x_i, y) - \phi(x_i, y_i) + \theta / C
+%\end{equation}
+%leading to a simple iterative optimization.
+
+\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
+The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
+into a quadratic objective with a combinatorial number of constraints:
+\begin{align}\label{eq:oneslack}
+    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
+    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
+        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
+            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
+            - \xi
+\end{align}
+As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
+is maintained, using the cutting plane method. The algorithm starts with an empty working set,
+and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
+again, with the new set of constraints.
+The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
+which guarantees a suboptimality of at most $\epsilon$.
+
+The complete algorithm is described in Algorithm~\ref{alg_n_slack}.
+
+%\begin{algorithm*}[t]
+    %\caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
+    %%FIXME
+
+    %\begin{algorithmic}[1]
+        %\require training samples $\{ (x^{(1)}, y^{(1)}), \dots, (x^{(n)}, y^{(n)})\}$, regularization parameter $c$, stopping tolerance $\epsilon$.
+        %\ensure parameters $\theta$, slack $\xi$
+        %\state $\w \leftarrow \emptyset$
+        %\repeat
+        %\state $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + c \xi$\label{restricted}\\
+        %\begin{equation*}
+            %\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \in \w:
+            %\left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^i, \hat{y}^{(i)})] \right \rangle \geq \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)})
+                %- \xi
+            %\end{equation*}
+            %\for {i=1, \dots, n}
+            %\state
+            %$\hat{y}^{(i)} \leftarrow i(x^{(i)}, y^{(i)}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{y}} \sum_{i=1}^n \delta(y^{(i)}, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^{(i)}, \hat{y})] \right \rangle$ \label{get_cutting_plane}
+            %\endfor
+            %\state $\w \leftarrow \w \cup \{ (\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \} $
+            %\state $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^{(i)}, \hat{y}^{(i)})] \right \rangle$
+            %\until $\xi' - \xi < \epsilon$ \label{convergence_check}
+        %\end{algorithmic}
+    %\end{algorithm*}
+
+\subsection{The 1-Slack Cutting Plane Method}\seclabel{one_slack}
+
+\begin{algorithm*}[t]
+    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
+    \begin{algorithmic}[1]
+        \require training samples $\{ (x^{(1)}, y^{(1)}), \dots, (x^{(n)}, y^{(n)})\}$, regularization parameter $c$, stopping tolerance $\epsilon$.
+        \ensure parameters $\theta$, slack $\xi$
+        \state $\w \leftarrow \emptyset$
+        \repeat
+            \state $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + c \xi$\label{restricted}\\
+            \begin{equation*}
+                \text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \in \w:
+                \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^i, \hat{y}^{(i)})] \right \rangle \geq \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)})
+                - \xi
+            \end{equation*}
+            \for {i=1, \dots, n}
+                \state
+                $\hat{y}^{(i)} \leftarrow i(x^{(i)}, y^{(i)}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{y}} \sum_{i=1}^n \delta(y^{(i)}, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^{(i)}, \hat{y})] \right \rangle$ \label{get_cutting_plane}
+            \endfor
+            \state $\w \leftarrow \w \cup \{ (\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \} $
+            \state $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^{(i)}, \hat{y}^{(i)})] \right \rangle$
+        \until $\xi' - \xi < \epsilon$ \label{convergence_check}
+    \end{algorithmic}
+\end{algorithm*}
+
+The one-slack cutting plane method~\citep{joachims2009cutting} solves the
+following reformulation of Equation~\Eqref{learning_equation}:
 \begin{align}\label{eq:oneslack}
@@
 \end{align}
-using the cutting plane method described in
-Algorithm~\ref{alg_cutting_plane}~\citep{joachims2009cutting}.
+Informally, the one-slack formulation corresponds to joining all training
+samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
+interactions between variables corresponding to different data points, and
+then applying the n-slack algorithm.
+The complete algorithm is described in Algorithm~\ref{alg_one_slack}.
 
-The cutting plane method alternates between solving Equation~\Eqref{oneslack}
-with a working set $\W$ of constraints, and expanding the working set using the
-current $\theta$ by finding $\mathbf{y}$ corresponding to the most violated constraint,
-using a separation oracle $I$.
-We investigate the construction of $\W$ and the influence of $I$
-on learning.
+By construction only a single constraint is added in each iteration of
+Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
+This has the advantage of making the solution of the QP faster, as it contains
+less variables. The down-side of this is that the loss-augmented prediction
+problem has to be solve much more often until convergence.
+%TODO convergence rate
 
-%TODO unstructured!!
+\citet{joachims2009cutting}, who introduced the method, proposed two enhancements
+to make the algorithm more efficient:
+\begin{description}
+    \item[Constraint Pruning] Members of the working set $\W$ can become inactive during learning.
+        If a constraint has been inactive for a number of itertions, it is removed from $\W$, leading
+        to smaller problem sizes.
+    \item[Inference Caching] In the one-slack algorithm, each constraint is
+        created using a combination of loss-augmented prediction results.
+        Therefore, each of these predictions can be part of multiple
+        constraints during learning.
+        To exploit this, we maintain a set $C^i$ of the last $r$ results of
+        loss-augmented inference for each training example $(x^i, y^i)$
+        (line~\ref{get_cutting_plane} in Algorithm~\ref{alg_one_slack}).
+        For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we
+        find
+        \[ \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
+            \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\phi(x^i, y^i) -
+                \phi(x^i, \hat{y})] \]
+        by enumeration of $C^i$ and continue until
+        line~\ref{convergence_check}.  Only if $\xi' - \xi < \epsilon$, that is
+        the produced constraint is not violated, we return to
+        line~\ref{get_cutting_plane} and actually invoke the loss augmented
+        prediction $I$.
+\end{description}
 
-Intuitively, the one-slack formulation corresponds to joining all training
-samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
-interactions between variables corresponding to different data points.
-Consequently, only a single constraint is added in each iteration of
-Algorithm~\ref{alg_cutting_plane}, leading to very small $\W$. We further use
-pruning of unused constraints, as suggested by \citet{joachims2009cutting},
-resulting in the size of $\W$ being in the order of hundreds for all experiments.
-
-We also use another enhancement of the cutting plane algorithm introduced by
-\citet{joachims2009cutting}, the caching oracle. For each training example $(x^i, y^i)$,
-we maintain a set $C^i$ of the last $r$ results of loss-augmented inference
-(line~\ref{get_cutting_plane} in Algorithm~\ref{alg_cutting_plane}).
-For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we find
-\[ 
-    \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i, \hat{y})]
-\]
-by enumeration of $C^i$ and continue until line~\ref{convergence_check}.
-Only if $\xi' - \xi < \epsilon$, that is the produced constraint is not violated, we
-return to line~\ref{get_cutting_plane} and actually invoke the separation
-oracle $I$.
+\subsection{Stochastic Dual Coordinate Descent}\seclabel{dual_coordinate_descent}
 
@@
 
+\section{Comparing Learning Algorithms}
+%TODO experiments!!
+
 \section{Summary}

### 2013-08-29 partial_supervision/paper.tex
more on algorithms for structured learning, but broke the algorithm listings?
FIX put chapter commands in separate files, so that the text is not on a new page...

@@
+\chapter{Weakly Supervised Object segmentation via Multiple Instance Learning}
+
 Most algorithms for semantic image segmentation and object-class segmentation
@@
 Training an SVM with this kernel produces a bag-level classifier for each
-class, which we will refer to as MIK.  This procedure is very efficient since
+class, which we will refer to as MIK\@.  This procedure is very efficient since
 the resulting kernel matrix is of size number of bags, which is much smaller

### 2013-08-29 pystruct/paper.tex
FIX put chapter commands in separate files, so that the text is not on a new page...
Fixes in the cutting plane algorithms, nicer looks

@@
+\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}\label{ch:structured_pystruct}
+
 Many classical computer vision applications such as stereo, optical flow, semantic
@@
 \section{Learning Algorithms for Linear Max-Margin Structured Prediction}
+
 Maximum Margin learning is now one of the most popular methods to learn
@@
 
-For learning, a dataset $(x^{(1)}, y^{(1)}),\dotsc,(x^{(k)}, y^{(k)})$, together with a loss
+For learning, a dataset $(x\hoch{1}, y\hoch{1}),\dotsc,(x\hoch{k}, y\hoch{k})$ is given, together with a loss
 \begin{equation}
@@
 \begin{equation}\eqlabel{learning_equation}
-    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x^{(i)}, y^{(i)}, \theta)
+    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x\hoch{i}, y\hoch{i}, \theta)
 \end{equation}
@@
 \begin{equation}\eqlabel{loss_augmentation}
-    \ell(x^{(i)}, y^{(i)}, \theta)= [\max_{y \in \mathcal{Y}} \Delta(y^{(i)}, y) + \theta^T \phi(x^{(i)}, y) - \theta^T \phi(x^{(i)}, y^{(i)}]_+
+    \ell(x\hoch{i}, y\hoch{i}, \theta)= [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+
 \end{equation}
@@
 \subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
+\begin{algorithm*}[t]
+    \caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
+    \begin{doublespacing}
+    \begin{algorithmic}[1]
+        \Require training samples $\{ (x\hoch{1}, y\hoch{1}), \dots, (x\hoch{k}, y\hoch{k})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
+        \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
+        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for all } i=1,\dots,k$
+        \Repeat
+            \For {i=1, \dots, k}
+                \State
+                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
+                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
+                    \State $\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
+                    \State
+                    \vspace{-15mm}
+                    \begin{align*}
+                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle &\argmin_{\theta, \xi_1, \dots, \xi_k} \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
+                    &\text{s.t. }\forall i=1,\dots,k \forall \hat{y} \in \W_i:\\
+                    &\theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
+                    \end{align*}
+                \EndIf
+            \EndFor
+            \vspace{-10mm}
+            \Until no $\W_i$ changes any more.
+        \end{algorithmic}
+    \end{doublespacing}
+    \end{algorithm*}
+
 The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
@@
     \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
-    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
-        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
-            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
+    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
+    &\theta^T \sum_{i=1}^n [\phi(x\hoch{i}, y\hoch{i}) - \phi(x\hoch{i},
+        \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
             - \xi
@@
 
-The complete algorithm is described in Algorithm~\ref{alg_n_slack}.
-
-%\begin{algorithm*}[t]
-    %\caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
-    %%FIXME
-
-    %\begin{algorithmic}[1]
-        %\require training samples $\{ (x^{(1)}, y^{(1)}), \dots, (x^{(n)}, y^{(n)})\}$, regularization parameter $c$, stopping tolerance $\epsilon$.
-        %\ensure parameters $\theta$, slack $\xi$
-        %\state $\w \leftarrow \emptyset$
-        %\repeat
-        %\state $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + c \xi$\label{restricted}\\
-        %\begin{equation*}
-            %\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \in \w:
-            %\left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^i, \hat{y}^{(i)})] \right \rangle \geq \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)})
-                %- \xi
-            %\end{equation*}
-            %\for {i=1, \dots, n}
-            %\state
-            %$\hat{y}^{(i)} \leftarrow i(x^{(i)}, y^{(i)}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{y}} \sum_{i=1}^n \delta(y^{(i)}, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^{(i)}, \hat{y})] \right \rangle$ \label{get_cutting_plane}
-            %\endfor
-            %\state $\w \leftarrow \w \cup \{ (\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \} $
-            %\state $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^{(i)}, \hat{y}^{(i)})] \right \rangle$
-            %\until $\xi' - \xi < \epsilon$ \label{convergence_check}
-        %\end{algorithmic}
-    %\end{algorithm*}
+The complete procedure is described in Algorithm~\ref{alg_n_slack}.
 
@@
     \begin{algorithmic}[1]
-        \require training samples $\{ (x^{(1)}, y^{(1)}), \dots, (x^{(n)}, y^{(n)})\}$, regularization parameter $c$, stopping tolerance $\epsilon$.
-        \ensure parameters $\theta$, slack $\xi$
-        \state $\w \leftarrow \emptyset$
-        \repeat
-            \state $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + c \xi$\label{restricted}\\
-            \begin{equation*}
-                \text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \in \w:
-                \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^i, \hat{y}^{(i)})] \right \rangle \geq \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)})
-                - \xi
-            \end{equation*}
-            \for {i=1, \dots, n}
-                \state
-                $\hat{y}^{(i)} \leftarrow i(x^{(i)}, y^{(i)}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{y}} \sum_{i=1}^n \delta(y^{(i)}, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^{(i)}, \hat{y})] \right \rangle$ \label{get_cutting_plane}
-            \endfor
-            \state $\w \leftarrow \w \cup \{ (\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \} $
-            \state $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^{(i)}, \hat{y}^{(i)})] \right \rangle$
-        \until $\xi' - \xi < \epsilon$ \label{convergence_check}
+        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
+        \Ensure parameters $\theta$, slack $\xi$
+        \State $\W \leftarrow \emptyset$
+        \Repeat
+            \State 
+            \vspace{-5mm}
+            \begin{align*}
+            (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi\\
+            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
+            &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
+            \end{align*}
+            \For {i=1, \dots, k}
+                \State
+                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
+            \EndFor
+            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
+            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
+        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
     \end{algorithmic}
@@
 then applying the n-slack algorithm.
-The complete algorithm is described in Algorithm~\ref{alg_one_slack}.
+A detailed description can be found in Algorithm~\ref{alg_one_slack}.
 

### 2013-08-29 itm/paper.tex
FIX put chapter commands in separate files, so that the text is not on a new page...
Shorten some titles because they were ugly

@@
+\chapter{Information Theoretic Clustering}
+
 Before we start our investigation of semantic segmentation and object class segmentation,
@@
 %
-\section{Information Theoretic Clustering Using Nonparametric Entropy-Estimates}
+\section{Clustering Using Nonparametric Entropy-Estimates}
 In general, the goal of clustering can be formulated as follows: 

### 2013-08-29 main.tex
get rid of shorthands in MIL part, some refinement after first reread.
Remove MIL theory stuff, start rereading partial supervision part
more on algorithms for structured learning, but broke the algorithm listings?
FIX put chapter commands in separate files, so that the text is not on a new page...
Shorten some titles because they were ugly

@@
 
+\usepackage{hyperref}
 \usepackage{epsfig}
@@
 \usepackage{xspace}
+\usepackage{setspace}
 
@@
 \newcommand\etal{~et\,al.~}
+
+\newcommand\hzero{\ensuremath{p(h_I(x)=0)}}
+\newcommand\bagerror{\ensuremath{p(h_B(X) \neq Y)}}
+\newcommand\instanceerror{\ensuremath{p(h_I(x) \neq y)}}
+
+%\newcommand{\hoch}[1]{^{(#1)}}
+\newcommand{\hoch}[1]{^{#1}}
 
@@
 
-\chapter{Overview}
 \include{intro}
 
-\chapter{Information Theoretic Clustering}
 \include{itm/paper}
 
-\chapter{Weakly Supervised Object segmentation via Multiple Instance Learning}
 \include{partial_supervision/paper}
-\include{mil/paper}
-
-\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}\label{ch:structured_pystruct}
 
@@
 
-\chapter{Efficient Exact Learning of Loopy CRFs}
-
 \include{exact_learning/paper}
 
-\chapter{Hierarchical Latent Variable Models for Semantic Segmentation}\label{ch:latent}
+\chapter{Latent CRFs for Semantic Segmentation}\label{ch:latent}
 
@@
 \bibliography{main}
-\include{mil/appendix}
 \end{document}

### 2013-08-29 pystruct/paper.tex
Shorten some titles because they were ugly

@@
-\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}\label{ch:structured_pystruct}
+\chapter{Learning Conditional Random Fields}\label{ch:structured_pystruct}
 
@@
 
-\section{Learning Algorithms for Linear Max-Margin Structured Prediction}
+\section{Learning Algorithms for Max-Margin Structured Prediction}
 

### 2013-09-03 pystruct/paper.tex
working on subgradient stuff

@@
 Maximum Margin learning is now one of the most popular methods to learn
-structural models in computervision and in other fields.
+structural models in computervision and text processing.
 There are several reasons for the popularity of linear maximum margin approaches:
@@
 \begin{equation}\eqlabel{loss_augmentation}
-    \ell(x\hoch{i}, y\hoch{i}, \theta)= [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+
+    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+
 \end{equation}
@@
 \subsection{Stochastic Subgradient Descent}\seclabel{subgradient}
-Arguably the most straight-foward way to approach \Eqref{learning_equation} is using subgradient descent.
-In light of the complexity of solving the loss-augmented prediction problem in \Eqref{loss_augmentation},
-is is natural to work in a stochastic setting (see \citet{ratliff2007online}).
+Arguably the most straight-foward way to approach \Eqref{learning_equation} is
+using subgradient descent.  In light of the complexity of solving the
+loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
+to work in a stochastic setting (see \citet{ratliff2007online}).
+Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
+considering a single training example $(x\hoch{i}, y\hoch{i})$
+can be computed simply by solving the loss-augmented prediction problem:
+\begin{align}
+    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
+    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
+\end{align}
+The most commonly used update has the simple form
+\begin{equation}
+    \theta \leftarrow (1 - \eta_t) \theta - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
+\end{equation}
+Where $\eta_t$ is a sequence of learning rates.
+%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
+but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
+Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
+%\begin{equation}
+
+%\end{equation}
+
 %\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
@@
 model = crfs.EdgeFeatureGraphCRF(
-    class_weight=inverse_frequency, symmetric_edge_features=[0, 1],
-    antisymmetric_edge_features=[2], inference_method='qpbo')
+            class_weight=inverse_frequency,
+            symmetric_edge_features=[0, 1],
+            antisymmetric_edge_features=[2],
+            inference_method='qpbo')
 

### 2013-09-03 exact_learning/paper.tex
more on algorithms for structured learning, but broke the algorithm listings?
FIX put chapter commands in separate files, so that the text is not on a new page...
slight improvement of organization and introduction of exact learning chapter

@@
+\chapter{Efficient Exact Learning of Loopy CRFs}
+
 \section{Introduction}
 % Remove duplication with pystruct intro
-As discussed previously, many computer vision algorithms empoy conditional random
-field models on pixel or superpixel graphs. These graphs by nature contain loops,
-making exact prediction and loss-augmented prediction intractable.
-As loss-augmented prediction is a centreal step in the learning algorithms
+As discussed previously, %TODO in the introduction? in pystruct chapter?
+ many computer vision algorithms employ conditional random
+field models on pixel or superpixel graphs. These graphs, by nature, contain loops,
+making exact prediction and loss-augmented prediction in general intractable.
+As loss-augmented prediction is a central step in the learning algorithms
 discussed in Chapter~\ref{ch:structured_pystruct}
-in the inference then lead to approximate learning.
+in the inference lead to only approximate learning.
 
-In this chapter, we investigate the impact of these approximations in
-the one-slack learning algorithms in the context of semantic 
+In this chapter, we investigate the necessity and consequences of these
+approximations in the one-slack learning algorithms in the context of semantic
 image segmentation. We show that despite inference being deemed intractable in
-loopy superpixel models, we are able to learn a structured support vector machine
-for this application exactly.
-learning a model of pairwise
-We will evaluate our approach on the popular MSRC-21 and Pascal VOC datasets.
+loopy superpixel models, we are able to learn a pairwise conditional random
+field model using a structured support vector machine exactly for this task.
+We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. % and nyu?
 
@@
         application, even in the presence of loopy graphs.
+    % also fusion moves is pretty good?
 \end{itemize}
@@
 
-Max margin learning for structured prediction was introduced by
-\citet{taskar2003max}, in the form of maximum-margin Markov models. Later, this
-framework was generalized to structural support vector machines by
-\citet{tsochantaridis2006large}. Both works assume tractable loss-augmented
-inference.
+%Max margin learning for structured prediction was introduced by
+%\citet{taskar2003max}, in the form of maximum-margin Markov models. Later, this
+%framework was generalized to structural support vector machines by
+%\citet{tsochantaridis2006large}. Both works assume tractable loss-augmented
+%inference.
 
-Currently the most widely used method is the one-slack cutting plane formulation
-introduced by \citet{joachims2009cutting}.
-This work also introduced the caching of constraints,
-which serves as a basis for our work. We improve upon their caching scheme, and in
-particular consider how it interacts with approximate inference algorithms.
+%Currently the most widely used method is the one-slack cutting plane formulation
+%introduced by \citet{joachims2009cutting}.
+%This work also introduced the caching of constraints,
+%which serves as a basis for our work. We improve upon their caching scheme, and in
+%particular consider how it interacts with approximate inference algorithms.
 
-% TODO learning for inference?
 Recently, there has been much research in learning structured prediction models
@@
 particular linear programming (LP)---perform best on the considered model.
-They also give a bound on the empirical risk for this case.
-In contrast, we aim at optimizing the non-relaxed objective directly, yielding
-the original, tighter bound.
+They also show that learning parameters with the LP relaxation minimizes a
+bound on the empirical risk when extending the target domain to the relaxed
+solutions.
+In contrast, we aim at optimizing the non-relaxed objective directly, minimizing
+the original empirical risk. This is an important difference, as relaxed solutions
+are usually not acceptable in practice.
 
@@
 not relate their results back to the original SSVM objective they approximate.
+It is unclear how they compare to standard structured prediction approaches
+on real-world applications.
 
-\begin{algorithm*}[t]
-    \caption{Cutting Plane Training of Structural SVMs \label{alg_cutting_plane}}
-
-\begin{algorithmic}[1]
-    \Require Training samples $\{ (x^{1}, y^{1}), \dots, (x^{n}, y^{n})\}$, regularization parameter $C$, stopping tolerance $\epsilon$, separation oracle $I$.
-    \Ensure Parameters $\theta$, slack $\xi$
-    \State $\W \leftarrow \emptyset$
-    \Repeat
-    \State $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi$\label{restricted}\\
-    \begin{equation*}
-        \text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \W:
-        \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i,
-            \hat{y}^i)] \right \rangle \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
-            - \xi
-    \end{equation*}
-    \For {i=1, \dots, n}
-    \State
-    $\hat{y}^i \leftarrow I(x^i, y^i, \theta) \approx \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^n \Delta(y^i, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y})] \right \rangle$ \label{get_cutting_plane}
-    \EndFor
-    \State $\W \leftarrow \W \cup \{ (\hat{y}^1, \dots, \hat{y}^n) \} $
-    \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \Delta(y^i, \hat{y}^i) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^i, \hat{y}^i)] \right \rangle$
-    \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
-\end{algorithmic}
-\end{algorithm*}
-
-\section{Efficient Cutting Plane Training of SSVMs}\label{learning}
-
+\section{Learnint Structure Support Vector Machines with Approximate Inference}
+%TODO ueberleitung zu approximate inference
+% wir machen one-slack und I ist jetzt das und das.
 There are two groups of inference procedures, as identified in
@@
 the properties of under-generating and over-generating inference procedures in
-Section~\ref{bounds}.
+Section~\ref{bounds}. %TODO Filler sentence? wat?
 
@@
 We will now describe how we can further use 1. to both speed up and improve learning.
+
+\section{Efficient Cutting Plane Training of SSVMs}\label{learning}
 
@@
 \section{Experiments}
+
 \subsection{Inference Algorithms}
+
 As a fast under-generating inference algorithm, we used $\alpha$-expansion
@@
 \subsection{Image Segmentation}
-We choose superpixel-based semantic image segmentation as sample application
+We use superpixel-based semantic image segmentation as our application as this is our task of interest. %TODO formulation
 for this work.  CRF based models have a history of success in this application,

### 2013-09-03 main.tex
slight fixes to layout
slight rewrite of acknoledgements

@@
-\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{scrbook}
+\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=calc]{scrbook}
 
@@
 \usepackage{pslatex}
+\usepackage{microtype}
 \usepackage{natbib}
@@
 \usepackage{xspace}
-\usepackage{setspace}
+\onehalfspacing
+\KOMAoptions{DIV=last}
 
@@
 
+\begin{spacing}{1}
 \begin{titlepage}
@@
 \tableofcontents
-
+\end{spacing}
 %\listoffigures
@@
 whose feedback proved to be invaluable during my studies.
-I would like to thank Christoph Lampert, for hosting me as a visiting researcher
+I would like to thank Christoph Lampert for hosting me as a visiting researcher
 at the IST Austria. His guidance and advice helped me in directing my further
@@
 I have been lucky to be a part of the scikit-learn community, whose developers
-have been a constant source of encouragement and feedback to me. My understanding
-of machine learning as well as software engineering has been greatly affected
-by the constant exchange with Olivier Grisel, Lars Buitinck, Gael Varoquax,
-Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and others.
+have been a constant source of encouragement and inspiration to me. My
+understanding of machine learning as well as software engineering has been
+greatly affected by the continuing exchange with Olivier Grisel, Lars Buitinck,
+Gael Varoquax, Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and
+others.
 
-I also thank my parents and my sister for their continuing support.
+I also thank my parents and my sister for their support.
 

### 2013-09-03 pystruct/paper.tex
proof reading of pystruct

@@
 
-\section{Basic Concepts in Structured Prediction}
+\section{Basic Concepts in Structured Prediction}\seclabel{basics_structured_prediction}
 
@@
 
-\subsection{Factor Graphs and the Relation to Graphical Models}
+\subsection{Factor Graphs and the Relation to Graphical Models}\seclabel{factor_graphs}
 % one factor graph per instance?
@@
 \section{Casting Structured Prediction into Software}\label{sec:api}
+
 Unfortunately only few implementations are publicly
@@
 of~\citet{joachims2009cutting}.
-Our implementation, \pystruct, aims at providing a high-quality implementation
-with an easy-to-use interface, in the high-level Python language.  This allows
-practitioners to efficiently test a range of models, as well as allowing
+We provide an implementation, \pystruct, which aims at providing a high-quality
+code with an easy-to-use interface, in the high-level Python language.  This
+allows practitioners to efficiently test a range of models, as well as allowing
 researchers to compare to baseline methods much more easily than this is
@@
 structured prediction.
+We will base the experiments in the rest of this work on the algorithms
+implemented in \pystruct.
 
-Using the above formulation, learning can be broken down into three sub-problems:
+Using the formulation of structured prediction introduced in
+\Secref{basics_structured_prediction}, learning can be broken down into three
+sub-problems:
 \begin{enumerate*}
-    \item Optimizing the objective with respect to $\theta$.
     \item Encoding the structure of the problem in a joint feature function $\Phi$.
-    \item Solving the maximization problem in Equation~\ref{eq:main_equation}.
+    \item Solving the loss-augmented prediction problem in Equation~\eqref{loss_augmentation}.
+    \item Optimizing the objective \Eqref{learning_equation} with respect to $\theta$.
 \end{enumerate*}
-The later two problems are usually tightly coupled, as the maximization in
+The first two problems are usually tightly coupled, as the maximization in
 Equation~\ref{eq:main_equation} is usually only feasible by exploiting the
-structure of $\Phi$, while the first one is usually treated as independent.
+structure of $\Phi$, as described in \Secref{factor_graphs}. The last problem,
+finding $\theta$, on the other hand, is usually treated as independent.
 \pystruct takes an object-oriented approach to decouple the task-dependent
@@
 support cutting plane algorithms for structural support vector
-machines~(SSVMs), subgradient methods for SSVMs, the structured perceptron and
-latent variable SSVMs. The cutting plane implementation uses the {\sc cvxopt}
+machines~(SSVMs), the Block-Coordinate Frank Wolfe (BCFW) algorithm for SSVMs,
+subgradient methods for SSVMs, the structured perceptron and latent variable
+SSVMs. The cutting plane implementation uses the {\sc cvxopt}
 package \citep{dahl2006cvxopt} for quadratic optimization.
@@
 
-Table~\ref{table:comparision} lists algorithms and models that are implemented in \pystruct and
-compares them to other public structured prediction libraries, together with
-the programming language and the project license.
+Table~\ref{table:comparision} lists algorithms and models that are implemented
+in \pystruct and compares them to other public structured prediction libraries,
+together with the programming language and the project license.
 
@@
 \end{tabularx}
-    \caption{\label{table:comparision}Comparison of structured prediction software packages. CP stands
-for cutting plane optimization of SSVMs~\citep{joachims2009cutting}, SG for
-online subgradient optimization of SSVMs~\citep{ratliff2007online}, LV for
-latent variable SSVMs~\citep{yu2009learning}, ML for maximum likelihood
-learning, Chain for chain-structured models
-with pairwise interactions, Graph for arbitrary graphs with pairwise
-interactions, and LDCRF for latent dynamic CRF~\citep{morency2007latent}.
-{\footnotesize $^1$\pystruct itself is BSD licensed, but uses the GPL-licensed package {\sc cvxopt} for cutting-plane learning.}
+    \caption{\label{table:comparision}Comparison of structured prediction
+        software packages. CP stands for cutting plane optimization of
+        SSVMs~\citep{joachims2009cutting}, SG for online subgradient
+        optimization of SSVMs~\citep{ratliff2007online}, LV for latent variable
+        SSVMs~\citep{yu2009learning}, ML for maximum likelihood learning, Chain
+        for chain-structured models with pairwise interactions, Graph for
+        arbitrary graphs with pairwise interactions, and LDCRF for latent
+        dynamic CRF~\citep{morency2007latent}.  {\footnotesize $^1$\pystruct
+        itself is BSD licensed, but uses the GPL-licensed package {\sc cvxopt}
+    for cutting-plane learning.}
 }
@@
 
-\paragraph{Completeness}\pystruct aims at providing full pipelines that can be
-    used in applications. It contains model formulation for many typical
-    applications.  This is in contrast to \svmstruct that provides no
-    models at all, requiring the user to develop significant amounts of code, even
-    for simple tasks.
+\paragraph{Completeness}
+    \pystruct aims at providing complete predictors that can be used directly in
+    applications. It contains model formulation for many typical scenarios.
+    This is in contrast to \svmstruct that provides no models at all, requiring the
+    user to develop significant amounts of code, even for simple tasks.
 
-\paragraph{Modularity} \pystruct separates the algorithms of parameter estimation and
+\paragraph{Modularity}
+    \pystruct separates the algorithms for parameter estimation and
      inference from the task-dependent formulation of $\Phi$. This allows
@@
      implementations is important to allow fast prototyping and scaling to
+     large datasets. \pystruct achieves the same runtime performance
+     as the popular \svmstruct model for cutting plane algorithms, and
+     provides implementations of the BCFW and Subgradient methods that scale to
      large datasets.
@@
 \paragraph{Testing}
-     \pystruct contains a testing-suite with 80\% line-coverage. It also employs continuous integration
-     to ensure stability and seamless user experience.
+     \pystruct contains a testing-suite with 80\% line-coverage. It also
+     employs continuous integration to ensure stability and seamless user
+     experience.
 
 \paragraph{Integration}
-     To improve usability, \pystruct is interoperable with other numeric and scientific Python projects,
-     such as {\sc scikit-learn}~\citep{pedregosa2011scikit},
-     {\sc mahotas}~\citep{coelho:mahotas}, {\sc gensim}~\citep{rehurek_lrec}, and {\sc scikit-image}.
-     This allows users to build powerful applications with little effort. In
-     particular, most of the model-selection methods of {\sc scikit-learn} can be used
-     directly with \pystruct.
+     To improve usability, \pystruct is interoperable with other numeric and
+     scientific Python projects, such as {\sc
+     scikit-learn}~\citep{pedregosa2011scikit},
+     {\sc mahotas}~\citep{coelho:mahotas}, {\sc gensim}~\citep{rehurek_lrec},
+     and {\sc scikit-image}.  This allows users to build powerful applications
+     with little effort. In particular, most of the model-selection methods of
+     {\sc scikit-learn} can be used directly with \pystruct.
 
@@
 
-Conditional random fields are an important tool for semantic image
-segmentation.  We demonstrate how to learn an $n$-slack support
+We demonstrate the use of \pystruct on the task of semantic image segmentation,
+the main focus of this work. The example shows how to learn an $n$-slack support
 vector machine on a superpixel-based CRF on the popular Pascal dataset. We use
@@
 \centering
-\includegraphics[width=.45\textwidth]{times_MNIST}
-\includegraphics[width=.45\textwidth]{accs_MNIST}
+\includegraphics[width=.49\textwidth]{times_MNIST}
+\includegraphics[width=.49\textwidth]{accs_MNIST}
 \caption{Runtime comparison of \pystruct and \svmstruct for multi-class
@@
 \section{Summary}
+Im this chapter, we introduced basic concepts of structured prediction. In particular, we discussed
+Structured Support Vector Machines, a max-margin approach for training linear predictors for
+structured data. %TODO comparison of algorithms?
+We gave a description of several popular learning algorithms, together with their theoretical
+runtime bounds and practical considerations.
 %TODO summarize structured prediction
-This chapter introduced \pystruct, a modular structured learning and prediction library in Python.
-\pystruct is geared towards ease of use, while providing efficient implementations.
+We also introduced \pystruct, our modular structured learning and prediction library in Python.
+\pystruct is geared towards ease of use, while providing efficient implementations and will be the
+basis of our further experiments.
 \pystruct integrates itself into the scientific Python eco-system, making it easy to use with
 existing libraries and applications.
-Currently, \pystruct focuses on max-margin and perceptron-based approaches. In the future,
-we plan to integrate other paradigms, such as sampling-based learning~\citep{wick2011samplerank},
-surrogate objectives (for example pseudo-likelihood), and approaches that allow for a better integration
-of inference and learning~\citep{meshi2010learning}.
+Currently, \pystruct focuses on max-margin and perceptron-based approaches.
+In the future, we plan to integrate other paradigms, such as sampling-based
+learning~\citep{wick2011samplerank}, surrogate objectives (for example
+pseudo-likelihood), and approaches that allow for a better integration of
+inference and learning~\citep{meshi2010learning}.

### 2013-09-03 pystruct/paper.tex
spelling

@@
 or prediction.
-While using arbitrary functions is possible, and was expored
+While using arbitrary functions is possible, and was explored
 in some recent works % TODO cite dtf, conditional neural fields
@@
 discrete labels $\mathcal{Y} = \{1, \dotsc, k\}^n$.
-In the following, we will only discuss this multivatiate case. In general, $n$ is often different
+In the following, we will only discuss this multivariate case. In general, $n$ is often different
 for different inputs $x$, such as image with different number of (super) pixels.
@@
 This approach to structured prediction is closely related to a probabilistic
-approach using \emph{graphical models}.  Probabilistical graphical models are a
+approach using \emph{graphical models}.  Probabilistic graphical models are a
 tool to express factorizations of probability distributions.  Similar to
@@
 is the normalization constant of the conditional distribution over $y$.
-If $f$ is choosen as in \Eqref{linear_psi}, then the resulting distribution
+If $f$ is chosen as in \Eqref{linear_psi}, then the resulting distribution
 belongs to the exponential family.
@@
 Maximum Margin learning is now one of the most popular methods to learn
-structural models in computervision and text processing.
+structural models in computer vision and text processing.
 There are several reasons for the popularity of linear maximum margin approaches:
@@
         specifies $g$, then learning is feasible as soon as the maximization
-        over $y$ in \Eqref{main_equation} can be caried out.
+        over $y$ in \Eqref{main_equation} can be carried out.
     \item[Generalization] The maximum margin principle yields generalization
@@
 terms of computational complexity.
-Therefore, it is often desireable to find learning algorithms that converge with
+Therefore, it is often desirable to find learning algorithms that converge with
 as little optimizations of the loss-augmented prediction problem as possible.
@@
     \item[Constraint Pruning] Members of the working set $\W$ can become inactive during learning.
-        If a constraint has been inactive for a number of itertions, it is removed from $\W$, leading
+        If a constraint has been inactive for a number of iterations, it is removed from $\W$, leading
         to smaller problem sizes.
@@
 \paragraph{Efficiency}
-     While \pystruct focuses on usability, providing efficient and competative
+     While \pystruct focuses on usability, providing efficient and competitive
      implementations is important to allow fast prototyping and scaling to
@@
 important that the implemented learning algorithms run in acceptable time.
-In this section, we compare our implementation of the one-slack cutting plance
+In this section, we compare our implementation of the one-slack cutting plane
 algorithm with the implementation in \svmstruct.
@@
 basis of our further experiments.
-\pystruct integrates itself into the scientific Python eco-system, making it easy to use with
+\pystruct integrates itself into the scientific Python ecosystem, making it easy to use with
 existing libraries and applications.

### 2013-09-04 pystruct/paper.tex
working on pystruct text

@@
 functions lead to learning problems that are convex in the parameters, complex
-constraints complicate their optimization.
+constraints complicate their optimization.  % computation? ;)
 
@@
 structured prediction, as well as in their application in areas such as natural
-language processing and computer vision (see \citet{nowozin2011structured} for
-an overview).
+language processing and computer vision (see \citet{nowozin2011structured} %TODO ms book too!
+for an overview).
 %
-In this chapter, we will first introduce the concepts and algorithms used in
-structured prediction, in particular in maximum margin methods. We will then
+In this chapter, we first introduce the concepts and algorithms used in
+structured prediction, in particular in maximum margin methods. We 
 accompany this by a description of our open source implementation of these
@@
 \end{equation}
-The computation of $y$ in the above equation is often referred to as inference
+Finding $y$ in the above equation is often referred to as inference
 or prediction.
-While using arbitrary functions is possible, and was explored
-in some recent works % TODO cite dtf, conditional neural fields
-most current approaches use linear functions, leading to:
+Most current approaches use linear functions,
+with some notable examples: % TODO cite dtf, conditional neural fields
+
+The linear approach leads to
 \begin{equation}\eqlabel{main_equation}
@@
 for each word in a sentence or a label for each pixel in an image.
-Learning structured prediction means learning the parameters $\theta$ from training data.
-The particular model that is used is completely encoded in $\Phi(x, y)$, which manifests
-the relation between input $x$ and output $y$. As $\mathcal{Y}$ is typically very large,
-it is crucial to exploit the particular form of $\Phi(x, y)$ to solve the inference problem
-of \Eqref{main_equation}.
+Learning structured prediction means learning the parameters $\theta$ from
+training data.  The particular model that is used is completely encoded in
+$\Phi(x, y)$, which manifests the relation between input $x$ and output $y$. As
+$\mathcal{Y}$ is typically very large, it is crucial to exploit the particular
+form of $\Phi(x, y)$ to solve the inference problem of \Eqref{main_equation}.
 
@@
 discrete labels $\mathcal{Y} = \{1, \dotsc, k\}^n$.
-In the following, we will only discuss this multivariate case. In general, $n$ is often different
-for different inputs $x$, such as image with different number of (super) pixels.
-We will ignore this in our notation to simplify the presentation.
+In the following, we will only discuss this multivariate case. In general, $n$
+is often different for different inputs $x$, such as image with different
+number of (super) pixels.  We will ignore this in our notation to simplify the
+presentation.
 
 \subsection{Factor Graphs and the Relation to Graphical Models}\seclabel{factor_graphs}
-% one factor graph per instance?
+% one factor graph per instance? FIXME
 In the case when $y$ is multivariate, a very general and widely used method to

### 2013-09-04 pystruct/paper.tex
splitting pystruct file

@@
-\chapter{Learning Conditional Random Fields}\label{ch:structured_pystruct}
-
-Many classical computer vision applications such as stereo, optical flow, semantic
-segmentation and visual grouping can be naturally formulated as image labeling tasks.
-%
-Arguably the most popular way to approach such labeling problems is via graphical
-models, such as Markov random fields (MRFs) and conditional random fields (CRFs).
-MRFs and CRFs provide a principled way of integrating local evidence and
-modeling spacial dependencies, which are strong in most image-based tasks.
-%
-While in earlier approaches, model parameters were set by hand or using
-cross-validation, more recently parameters are often learned using a max-margin
-approach.
-
-Most models employ linear energy functions of unary and pairwise interactions,
-trained using structural support vector machines (SSVMs). While linear energy
-functions lead to learning problems that are convex in the parameters, complex
-constraints complicate their optimization.  % computation? ;)
-
-In recent years there has been a wealth of research in methods for learning
-structured prediction, as well as in their application in areas such as natural
-language processing and computer vision (see \citet{nowozin2011structured} %TODO ms book too!
-for an overview).
-%
-In this chapter, we first introduce the concepts and algorithms used in
-structured prediction, in particular in maximum margin methods. We 
-accompany this by a description of our open source implementation of these
-algorithms, \pystruct.
-
-\section{Basic Concepts in Structured Prediction}\seclabel{basics_structured_prediction}
-
-Structured prediction can be defined as making a prediction $f(x)$ by maximizing a
-compatibility function $g$ between an input $x$ and possible labels
-$y$~\citep{nowozin2011structured}:
-\begin{equation}
-    f(x) = \argmax_{y \in \mathcal{Y}} g(x, y)
-\end{equation}
-Finding $y$ in the above equation is often referred to as inference
-or prediction.
-Most current approaches use linear functions,
-with some notable examples: % TODO cite dtf, conditional neural fields
-
-The linear approach leads to
-\begin{equation}\eqlabel{main_equation}
-    f(x) = \argmax_{y \in \mathcal{Y}}\  \theta^T \Phi(x, y).
-\end{equation}
-Here, $y$ is a \emph{structured} label, $\Phi$ is a joint feature function of
-$x$ and $y$, and $\theta$ are parameters of the model. \emph{Structured} means
-that $y$ is more complicated than a single output class, for example a label
-for each word in a sentence or a label for each pixel in an image.
-Learning structured prediction means learning the parameters $\theta$ from
-training data.  The particular model that is used is completely encoded in
-$\Phi(x, y)$, which manifests the relation between input $x$ and output $y$. As
-$\mathcal{Y}$ is typically very large, it is crucial to exploit the particular
-form of $\Phi(x, y)$ to solve the inference problem of \Eqref{main_equation}.
-
-While $y$ could be a complicated like a parsing tree or the geometric
-configuration of a molecule, many setting, such as the image segmentation
-setting we are interested, can be reduced to the case where $y$ is a vector of
-discrete labels $\mathcal{Y} = \{1, \dotsc, k\}^n$.
-In the following, we will only discuss this multivariate case. In general, $n$
-is often different for different inputs $x$, such as image with different
-number of (super) pixels.  We will ignore this in our notation to simplify the
-presentation.
-
-\subsection{Factor Graphs and the Relation to Graphical Models}\seclabel{factor_graphs}
-% one factor graph per instance? FIXME
-In the case when $y$ is multivariate, a very general and widely used method to
-specify the structure of a model, and therefore $\Phi$, is using \emph{factor
-graphs}. A factor graph is a bipartite graph $(\mathcal{V}, \mathcal{F},
-\mathcal{E})$, consisting of variable nodes $\mathcal{V}$, factor nodes
-$\mathcal{F}$ and edges $\mathcal{E}$ connecting variables to factors. The
-\emph{scope} $N_F$ of a factor $F \in \mathcal{F}$ is defined as
-\begin{equation}
-    N_F = \{ v \in \mathcal{V} \,|\, (v,F) \in \mathcal{E} \}
-\end{equation}
-The variable nodes of the factor graph correspond to the entries of the
-variables $y$, that is $\mathcal{V} = \{1, \dotsc, n\}$, and each factor node
-is associated with a \emph{factor} or \emph{potential function} $\psi_F$.
-A factor graph represents a function\footnote{Traditionally factor graphs
- represent \emph{products} of factors.  To simplify presentation, we work
-directly in the log-domain of the product representation.}
-\begin{equation}\eqlabel{log_factor_graph}
-    g(x, y) = \sum_{F \in \mathcal{F}} \psi_F(x, y_{N_F})
-\end{equation}
-Here, $y_{N_F}$ denotes the entries of $y$ indexed by $N_F$.
-
-The benefit of using the factor graph representation is that it decomposes the
-function over the variables of interest $y_i$ and makes. This allows us to
-apply efficient optimization procedures for the maximization in
-\Eqref{main_equation} by exploiting the graph structure of the factor graph.
-
-To obtain a linear function as in \Eqref{main_equation} from
-\Eqref{log_factor_graph}, we can simply let each $\psi$ be of the form
-\begin{equation}\eqlabel{general_psi}
-    \psi_F(x, y_{N_F}) = \theta_F^T \Phi_F(x, y_{N_F}).
-\end{equation}
-The by far most common form is 
-\begin{equation}\eqlabel{linear_psi}
-    \psi_F(x, y_{N_F}) = \theta_{F, y_{N_F}}^T \phi_F(x),
-\end{equation}
-where $\phi_F(x)$ is a vector representation of the input $x$, and there are
-different parameter vectors $\theta_{F, y_{N_F}}$ for each possible assignment
-of $y_{N_F} \in \mathcal{Y}_{N_F}$.
-
-Both, \Eqref{general_psi} and \Eqref{linear_psi} are instantiations of the general
-linear form \Eqref{main_equation}. To see this, for \Eqref{general_psi} we simply
-concatenate the individual components for all $f \in \mathcal{F}$:
-\begin{align}
-    \theta &= \bigoplus_{F \in \mathcal{F}} \theta_F\\
-    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \Phi_F(x, y_{N_F})
-\end{align}
-Writing down $\Phi$ and $\theta$ for the form \Eqref{linear_psi} is a little less compact:
-\begin{align}
-    \theta &= \bigoplus_{F \in \mathcal{F}}\ \bigoplus_{y_{N_F} \in \mathcal{Y}_{N_F}} \theta_{F, y_{N_F}}\\
-    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \left (\phi_F(x) \otimes e_{y_{N_F}} \right )
-\end{align}
-where $e_{y_{N_F}} \in \mathbb{R}^{|\mathcal{Y}_{N_F}|}$ is the indicator for a
-given variable setting $y_{N_F}$.
-In words, $\Phi_F$ is build simply by creating a vector of
-$|\mathcal{Y}_{N_F}|$ times the size of $\phi_F$, which is zero everywhere,
-except for the place corresponding to $y_{N_F}$. %TODO unclear!
-
-The benefit of using the factor graph representation is that it allows us to
-apply efficient optimization procedures for the maximization in
-\Eqref{main_equation} by exploiting the graph structure of the factor graph.
-
-This approach to structured prediction is closely related to a probabilistic
-approach using \emph{graphical models}.  Probabilistic graphical models are a
-tool to express factorizations of probability distributions.  Similar to
-\Eqref{log_factor_graph}, the joint probability distribution over a
-multi-variate random variable $y$ can be expressed using a factor-graph:
-\begin{equation}\eqlabel{graphicalmodel}
-    p(y | x) = \frac{1}{Z_x}\prod_{F \in \mathcal{F}} \exp\left(\psi_F(x, y_{N_F})\right)
-\end{equation}
-where
-\begin{equation}
-    Z_x = \sum_{y' \in \mathcal{Y}} \prod_{F \in \mathcal{F}} \exp\left(\psi(x, y'_{N_F})\right)
-\end{equation}
-is the normalization constant of the conditional distribution over $y$.
-If $f$ is chosen as in \Eqref{linear_psi}, then the resulting distribution
-belongs to the exponential family.
-
-The most probable prediction $y$ is given as $\argmax_{y \in \mathcal{Y}}p(y|x)$.
-As $Z$ is independent of $y$, and by the monotonicity of the logarithm,
-maximizing $p(y|x)$ is equivalent to maximizing $g(x,y)$ over $y$ in
-\Eqref{log_factor_graph}. Therefore, from a prediction standpoint, the two
-formulations are equivalent.
-
-During learning, the presence of the factor $Z$ in \Eqref{graphicalmodel}
-introduces additional complications. As we only address the problem of making
-predictions, not modeling probabilities, there are no clear benefits from the
-probabilistic approach.  Consequently, we will work with the more direct
-structured prediction approach of \Eqref{main_equation} and
-\Eqref{log_factor_graph} instead.
-
-% TODO relate to computer vision here ??? Pairwise models?
-% talk about inference?
-
-\section{Learning Algorithms for Max-Margin Structured Prediction}
-
-Maximum Margin learning is now one of the most popular methods to learn
-structural models in computer vision and text processing.
-There are several reasons for the popularity of linear maximum margin approaches:
-\begin{description}
-    \item[Loss-sensitivity] In contrast to probabilistic approaches, maximum margin learning approaches can directly
-        minimize a user-specified loss.
-    \item[Feasibility] If the loss decomposes over the factor graph that
-        specifies $g$, then learning is feasible as soon as the maximization
-        over $y$ in \Eqref{main_equation} can be carried out.
-    \item[Generalization] The maximum margin principle yields generalization
-        bounds using the effective complexity~\citep{taskar2003max}.
-    \item[Strong Convexity] The resulting optimization problem is strongly
-        convex, leading to efficient optimization and unique solutions.
-\end{description}
-
-For learning, a dataset $(x\hoch{1}, y\hoch{1}),\dotsc,(x\hoch{k}, y\hoch{k})$ is given, together with a loss
-\begin{equation}
-    \Delta \colon \mathcal{Y} \times \mathcal{Y} \rightarrow \mathbb{R}.
-\end{equation}
-The parameters $\theta$ are learned by minimizing the loss based soft-margin
-objective
-\begin{equation}\eqlabel{learning_equation}
-    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x\hoch{i}, y\hoch{i}, \theta)
-\end{equation}
-with regularization parameter $C$. Here $r_i$ is an hinge-loss like upper bound
-on $\Delta$ empirical risk:
-\begin{equation}\eqlabel{loss_augmentation}
-    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+
-\end{equation}
-
-This is an instance of regularized empirical risk minimization, with a
-piecewise linear, convex upper bound on the loss. 
-Finding the $y$ that corresponds to a maximum in \Eqref{loss_augmentation} is
-a central part of all maximum-margin based learning algorithms, and is referred to
-as loss-augmented prediction. For complex models, such as the ones used for
-image segmentation, this optimization often dominates the learning process in
-terms of computational complexity.
-Therefore, it is often desirable to find learning algorithms that converge with
-as little optimizations of the loss-augmented prediction problem as possible.
-
-There are several popular algorithms to solve \Eqref{learning_equation}. We
-will briefly review three standard algorithms: the one-slack and n-slack
-cutting plane algorithms, a stochastic primal subgradient algorithm, and
-recently proposed stochastic dual coordinate descent method.
-
-\subsection{Stochastic Subgradient Descent}\seclabel{subgradient}
-Arguably the most straight-foward way to approach \Eqref{learning_equation} is
-using subgradient descent.  In light of the complexity of solving the
-loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
-to work in a stochastic setting (see \citet{ratliff2007online}).
-Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
-considering a single training example $(x\hoch{i}, y\hoch{i})$
-can be computed simply by solving the loss-augmented prediction problem:
-\begin{align}
-    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
-    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
-\end{align}
-The most commonly used update has the simple form
-\begin{equation}
-    \theta \leftarrow (1 - \eta_t) \theta - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
-\end{equation}
-Where $\eta_t$ is a sequence of learning rates.
-%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
-but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
-Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
-%\begin{equation}
-
-%\end{equation}
-
-%\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
-%\begin{equation}\label{subgradient}
-    %g := \phi(x_i, y) - \phi(x_i, y_i) + \theta / C
-%\end{equation}
-%leading to a simple iterative optimization.
-
-\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
-\begin{algorithm*}[t]
-    \caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
-    \begin{doublespacing}
-    \begin{algorithmic}[1]
-        \Require training samples $\{ (x\hoch{1}, y\hoch{1}), \dots, (x\hoch{k}, y\hoch{k})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
-        \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
-        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for all } i=1,\dots,k$
-        \Repeat
-            \For {i=1, \dots, k}
-                \State
-                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
-                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
-                    \State $\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
-                    \State
-                    \vspace{-15mm}
-                    \begin{align*}
-                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle &\argmin_{\theta, \xi_1, \dots, \xi_k} \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
-                    &\text{s.t. }\forall i=1,\dots,k \forall \hat{y} \in \W_i:\\
-                    &\theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
-                    \end{align*}
-                \EndIf
-            \EndFor
-            \vspace{-10mm}
-            \Until no $\W_i$ changes any more.
-        \end{algorithmic}
-    \end{doublespacing}
-    \end{algorithm*}
-
-The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
-into a quadratic objective with a combinatorial number of constraints:
-\begin{align}\label{eq:oneslack}
-    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
-    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
-    &\theta^T \sum_{i=1}^n [\phi(x\hoch{i}, y\hoch{i}) - \phi(x\hoch{i},
-        \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
-            - \xi
-\end{align}
-As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
-is maintained, using the cutting plane method. The algorithm starts with an empty working set,
-and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
-again, with the new set of constraints.
-The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
-which guarantees a suboptimality of at most $\epsilon$.
-
-The complete procedure is described in Algorithm~\ref{alg_n_slack}.
-
-\subsection{The 1-Slack Cutting Plane Method}\seclabel{one_slack}
-
-\begin{algorithm*}[t]
-    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
-    \begin{algorithmic}[1]
-        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
-        \Ensure parameters $\theta$, slack $\xi$
-        \State $\W \leftarrow \emptyset$
-        \Repeat
-            \State 
-            \vspace{-5mm}
-            \begin{align*}
-            (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi\\
-            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
-            &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
-            \end{align*}
-            \For {i=1, \dots, k}
-                \State
-                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
-            \EndFor
-            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
-            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
-        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
-    \end{algorithmic}
-\end{algorithm*}
-
-The one-slack cutting plane method~\citep{joachims2009cutting} solves the
-following reformulation of Equation~\Eqref{learning_equation}:
-\begin{align}\label{eq:oneslack}
-    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
-    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
-        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
-            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
-            - \xi
-\end{align}
-Informally, the one-slack formulation corresponds to joining all training
-samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
-interactions between variables corresponding to different data points, and
-then applying the n-slack algorithm.
-A detailed description can be found in Algorithm~\ref{alg_one_slack}.
-
-By construction only a single constraint is added in each iteration of
-Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
-This has the advantage of making the solution of the QP faster, as it contains
-less variables. The down-side of this is that the loss-augmented prediction
-problem has to be solve much more often until convergence.
-%TODO convergence rate
-
-\citet{joachims2009cutting}, who introduced the method, proposed two enhancements
-to make the algorithm more efficient:
-\begin{description}
-    \item[Constraint Pruning] Members of the working set $\W$ can become inactive during learning.
-        If a constraint has been inactive for a number of iterations, it is removed from $\W$, leading
-        to smaller problem sizes.
-    \item[Inference Caching] In the one-slack algorithm, each constraint is
-        created using a combination of loss-augmented prediction results.
-        Therefore, each of these predictions can be part of multiple
-        constraints during learning.
-        To exploit this, we maintain a set $C^i$ of the last $r$ results of
-        loss-augmented inference for each training example $(x^i, y^i)$
-        (line~\ref{get_cutting_plane} in Algorithm~\ref{alg_one_slack}).
-        For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we
-        find
-        \[ \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
-            \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\phi(x^i, y^i) -
-                \phi(x^i, \hat{y})] \]
-        by enumeration of $C^i$ and continue until
-        line~\ref{convergence_check}.  Only if $\xi' - \xi < \epsilon$, that is
-        the produced constraint is not violated, we return to
-        line~\ref{get_cutting_plane} and actually invoke the loss augmented
-        prediction $I$.
-\end{description}
-
-\subsection{Stochastic Dual Coordinate Descent}\seclabel{dual_coordinate_descent}
-
 \section{Casting Structured Prediction into Software}\label{sec:api}

### 2013-09-04 structured_prediction/paper.tex
splitting pystruct file

\chapter{Learning Conditional Random Fields}\label{ch:structured_pystruct}

Many classical computer vision applications such as stereo, optical flow, semantic
segmentation and visual grouping can be naturally formulated as image labeling tasks.
%
Arguably the most popular way to approach such labeling problems is via graphical
models, such as Markov random fields (MRFs) and conditional random fields (CRFs).
MRFs and CRFs provide a principled way of integrating local evidence and
modeling spacial dependencies, which are strong in most image-based tasks.
%
While in earlier approaches, model parameters were set by hand or using
cross-validation, more recently parameters are often learned using a max-margin
approach.

Most models employ linear energy functions of unary and pairwise interactions,
trained using structural support vector machines (SSVMs). While linear energy
functions lead to learning problems that are convex in the parameters, complex
constraints complicate their optimization.  % computation? ;)

In recent years there has been a wealth of research in methods for learning
structured prediction, as well as in their application in areas such as natural
language processing and computer vision (see \citet{nowozin2011structured} %TODO ms book too!
for an overview).
%
In this chapter, we first introduce the concepts and algorithms used in
structured prediction, in particular in maximum margin methods. We 
accompany this by a description of our open source implementation of these
algorithms, \pystruct.

\section{Basic Concepts in Structured Prediction}\seclabel{basics_structured_prediction}

Structured prediction can be defined as making a prediction $f(x)$ by maximizing a
compatibility function $g$ between an input $x$ and possible labels
$y$~\citep{nowozin2011structured}:
\begin{equation}
    f(x) = \argmax_{y \in \mathcal{Y}} g(x, y)
\end{equation}
Finding $y$ in the above equation is often referred to as inference
or prediction.
Most current approaches use linear functions,
with some notable examples: % TODO cite dtf, conditional neural fields

The linear approach leads to
\begin{equation}\eqlabel{main_equation}
    f(x) = \argmax_{y \in \mathcal{Y}}\  \theta^T \Phi(x, y).
\end{equation}
Here, $y$ is a \emph{structured} label, $\Phi$ is a joint feature function of
$x$ and $y$, and $\theta$ are parameters of the model. \emph{Structured} means
that $y$ is more complicated than a single output class, for example a label
for each word in a sentence or a label for each pixel in an image.
Learning structured prediction means learning the parameters $\theta$ from
training data.  The particular model that is used is completely encoded in
$\Phi(x, y)$, which manifests the relation between input $x$ and output $y$. As
$\mathcal{Y}$ is typically very large, it is crucial to exploit the particular
form of $\Phi(x, y)$ to solve the inference problem of \Eqref{main_equation}.

While $y$ could be a complicated like a parsing tree or the geometric
configuration of a molecule, many setting, such as the image segmentation
setting we are interested, can be reduced to the case where $y$ is a vector of
discrete labels $\mathcal{Y} = \{1, \dotsc, k\}^n$.
In the following, we will only discuss this multivariate case. In general, $n$
is often different for different inputs $x$, such as image with different
number of (super) pixels.  We will ignore this in our notation to simplify the
presentation.

\subsection{Factor Graphs and the Relation to Graphical Models}\seclabel{factor_graphs}
% one factor graph per instance? FIXME
In the case when $y$ is multivariate, a very general and widely used method to
specify the structure of a model, and therefore $\Phi$, is using \emph{factor
graphs}. A factor graph is a bipartite graph $(\mathcal{V}, \mathcal{F},
\mathcal{E})$, consisting of variable nodes $\mathcal{V}$, factor nodes
$\mathcal{F}$ and edges $\mathcal{E}$ connecting variables to factors. The
\emph{scope} $N_F$ of a factor $F \in \mathcal{F}$ is defined as
\begin{equation}
    N_F = \{ v \in \mathcal{V} \,|\, (v,F) \in \mathcal{E} \}
\end{equation}
The variable nodes of the factor graph correspond to the entries of the
variables $y$, that is $\mathcal{V} = \{1, \dotsc, n\}$, and each factor node
is associated with a \emph{factor} or \emph{potential function} $\psi_F$.
A factor graph represents a function\footnote{Traditionally factor graphs
 represent \emph{products} of factors.  To simplify presentation, we work
directly in the log-domain of the product representation.}
\begin{equation}\eqlabel{log_factor_graph}
    g(x, y) = \sum_{F \in \mathcal{F}} \psi_F(x, y_{N_F})
\end{equation}
Here, $y_{N_F}$ denotes the entries of $y$ indexed by $N_F$.

The benefit of using the factor graph representation is that it decomposes the
function over the variables of interest $y_i$ and makes. This allows us to
apply efficient optimization procedures for the maximization in
\Eqref{main_equation} by exploiting the graph structure of the factor graph.

To obtain a linear function as in \Eqref{main_equation} from
\Eqref{log_factor_graph}, we can simply let each $\psi$ be of the form
\begin{equation}\eqlabel{general_psi}
    \psi_F(x, y_{N_F}) = \theta_F^T \Phi_F(x, y_{N_F}).
\end{equation}
The by far most common form is 
\begin{equation}\eqlabel{linear_psi}
    \psi_F(x, y_{N_F}) = \theta_{F, y_{N_F}}^T \phi_F(x),
\end{equation}
where $\phi_F(x)$ is a vector representation of the input $x$, and there are
different parameter vectors $\theta_{F, y_{N_F}}$ for each possible assignment
of $y_{N_F} \in \mathcal{Y}_{N_F}$.

Both, \Eqref{general_psi} and \Eqref{linear_psi} are instantiations of the general
linear form \Eqref{main_equation}. To see this, for \Eqref{general_psi} we simply
concatenate the individual components for all $f \in \mathcal{F}$:
\begin{align}
    \theta &= \bigoplus_{F \in \mathcal{F}} \theta_F\\
    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \Phi_F(x, y_{N_F})
\end{align}
Writing down $\Phi$ and $\theta$ for the form \Eqref{linear_psi} is a little less compact:
\begin{align}
    \theta &= \bigoplus_{F \in \mathcal{F}}\ \bigoplus_{y_{N_F} \in \mathcal{Y}_{N_F}} \theta_{F, y_{N_F}}\\
    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \left (\phi_F(x) \otimes e_{y_{N_F}} \right )
\end{align}
where $e_{y_{N_F}} \in \mathbb{R}^{|\mathcal{Y}_{N_F}|}$ is the indicator for a
given variable setting $y_{N_F}$.
In words, $\Phi_F$ is build simply by creating a vector of
$|\mathcal{Y}_{N_F}|$ times the size of $\phi_F$, which is zero everywhere,
except for the place corresponding to $y_{N_F}$. %TODO unclear!

The benefit of using the factor graph representation is that it allows us to
apply efficient optimization procedures for the maximization in
\Eqref{main_equation} by exploiting the graph structure of the factor graph.

This approach to structured prediction is closely related to a probabilistic
approach using \emph{graphical models}.  Probabilistic graphical models are a
tool to express factorizations of probability distributions.  Similar to
\Eqref{log_factor_graph}, the joint probability distribution over a
multi-variate random variable $y$ can be expressed using a factor-graph:
\begin{equation}\eqlabel{graphicalmodel}
    p(y | x) = \frac{1}{Z_x}\prod_{F \in \mathcal{F}} \exp\left(\psi_F(x, y_{N_F})\right)
\end{equation}
where
\begin{equation}
    Z_x = \sum_{y' \in \mathcal{Y}} \prod_{F \in \mathcal{F}} \exp\left(\psi(x, y'_{N_F})\right)
\end{equation}
is the normalization constant of the conditional distribution over $y$.
If $f$ is chosen as in \Eqref{linear_psi}, then the resulting distribution
belongs to the exponential family.

The most probable prediction $y$ is given as $\argmax_{y \in \mathcal{Y}}p(y|x)$.
As $Z$ is independent of $y$, and by the monotonicity of the logarithm,
maximizing $p(y|x)$ is equivalent to maximizing $g(x,y)$ over $y$ in
\Eqref{log_factor_graph}. Therefore, from a prediction standpoint, the two
formulations are equivalent.

During learning, the presence of the factor $Z$ in \Eqref{graphicalmodel}
introduces additional complications. As we only address the problem of making
predictions, not modeling probabilities, there are no clear benefits from the
probabilistic approach.  Consequently, we will work with the more direct
structured prediction approach of \Eqref{main_equation} and
\Eqref{log_factor_graph} instead.

% TODO relate to computer vision here ??? Pairwise models?
% talk about inference?

\section{Learning Algorithms for Max-Margin Structured Prediction}

Maximum Margin learning is now one of the most popular methods to learn
structural models in computer vision and text processing.
There are several reasons for the popularity of linear maximum margin approaches:
\begin{description}
    \item[Loss-sensitivity] In contrast to probabilistic approaches, maximum margin learning approaches can directly
        minimize a user-specified loss.
    \item[Feasibility] If the loss decomposes over the factor graph that
        specifies $g$, then learning is feasible as soon as the maximization
        over $y$ in \Eqref{main_equation} can be carried out.
    \item[Generalization] The maximum margin principle yields generalization
        bounds using the effective complexity~\citep{taskar2003max}.
    \item[Strong Convexity] The resulting optimization problem is strongly
        convex, leading to efficient optimization and unique solutions.
\end{description}

For learning, a dataset $(x\hoch{1}, y\hoch{1}),\dotsc,(x\hoch{k}, y\hoch{k})$ is given, together with a loss
\begin{equation}
    \Delta \colon \mathcal{Y} \times \mathcal{Y} \rightarrow \mathbb{R}.
\end{equation}
The parameters $\theta$ are learned by minimizing the loss based soft-margin
objective
\begin{equation}\eqlabel{learning_equation}
    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x\hoch{i}, y\hoch{i}, \theta)
\end{equation}
with regularization parameter $C$. Here $r_i$ is an hinge-loss like upper bound
on $\Delta$ empirical risk:
\begin{equation}\eqlabel{loss_augmentation}
    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+
\end{equation}

This is an instance of regularized empirical risk minimization, with a
piecewise linear, convex upper bound on the loss. 
Finding the $y$ that corresponds to a maximum in \Eqref{loss_augmentation} is
a central part of all maximum-margin based learning algorithms, and is referred to
as loss-augmented prediction. For complex models, such as the ones used for
image segmentation, this optimization often dominates the learning process in
terms of computational complexity.
Therefore, it is often desirable to find learning algorithms that converge with
as little optimizations of the loss-augmented prediction problem as possible.

There are several popular algorithms to solve \Eqref{learning_equation}. We
will briefly review three standard algorithms: the one-slack and n-slack
cutting plane algorithms, a stochastic primal subgradient algorithm, and
recently proposed stochastic dual coordinate descent method.

\subsection{Stochastic Subgradient Descent}\seclabel{subgradient}
Arguably the most straight-foward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
to work in a stochastic setting (see \citet{ratliff2007online}).
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align}
The most commonly used update has the simple form
\begin{equation}
    \theta \leftarrow (1 - \eta_t) \theta - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of learning rates.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
%\begin{equation}

%\end{equation}

%\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
%\begin{equation}\label{subgradient}
    %g := \phi(x_i, y) - \phi(x_i, y_i) + \theta / C
%\end{equation}
%leading to a simple iterative optimization.

\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
\begin{algorithm*}[t]
    \caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
    \begin{doublespacing}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{1}, y\hoch{1}), \dots, (x\hoch{k}, y\hoch{k})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for all } i=1,\dots,k$
        \Repeat
            \For {i=1, \dots, k}
                \State
                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
                    \State $\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
                    \State
                    \vspace{-15mm}
                    \begin{align*}
                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle &\argmin_{\theta, \xi_1, \dots, \xi_k} \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
                    &\text{s.t. }\forall i=1,\dots,k \forall \hat{y} \in \W_i:\\
                    &\theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
                    \end{align*}
                \EndIf
            \EndFor
            \vspace{-10mm}
            \Until no $\W_i$ changes any more.
        \end{algorithmic}
    \end{doublespacing}
    \end{algorithm*}

The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
into a quadratic objective with a combinatorial number of constraints:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
    &\theta^T \sum_{i=1}^n [\phi(x\hoch{i}, y\hoch{i}) - \phi(x\hoch{i},
        \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
            - \xi
\end{align}
As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
is maintained, using the cutting plane method. The algorithm starts with an empty working set,
and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
again, with the new set of constraints.
The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
which guarantees a suboptimality of at most $\epsilon$.

The complete procedure is described in Algorithm~\ref{alg_n_slack}.

\subsection{The 1-Slack Cutting Plane Method}\seclabel{one_slack}

\begin{algorithm*}[t]
    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $\xi$
        \State $\W \leftarrow \emptyset$
        \Repeat
            \State 
            \vspace{-5mm}
            \begin{align*}
            (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi\\
            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
            &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
            \end{align*}
            \For {i=1, \dots, k}
                \State
                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
            \EndFor
            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
    \end{algorithmic}
\end{algorithm*}

The one-slack cutting plane method~\citep{joachims2009cutting} solves the
following reformulation of Equation~\Eqref{learning_equation}:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
\end{align}
Informally, the one-slack formulation corresponds to joining all training
samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
interactions between variables corresponding to different data points, and
then applying the n-slack algorithm.
A detailed description can be found in Algorithm~\ref{alg_one_slack}.

By construction only a single constraint is added in each iteration of
Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
This has the advantage of making the solution of the QP faster, as it contains
less variables. The down-side of this is that the loss-augmented prediction
problem has to be solve much more often until convergence.
%TODO convergence rate

\citet{joachims2009cutting}, who introduced the method, proposed two enhancements
to make the algorithm more efficient:
\begin{description}
    \item[Constraint Pruning] Members of the working set $\W$ can become inactive during learning.
        If a constraint has been inactive for a number of iterations, it is removed from $\W$, leading
        to smaller problem sizes.
    \item[Inference Caching] In the one-slack algorithm, each constraint is
        created using a combination of loss-augmented prediction results.
        Therefore, each of these predictions can be part of multiple
        constraints during learning.
        To exploit this, we maintain a set $C^i$ of the last $r$ results of
        loss-augmented inference for each training example $(x^i, y^i)$
        (line~\ref{get_cutting_plane} in Algorithm~\ref{alg_one_slack}).
        For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we
        find
        \[ \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
            \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\phi(x^i, y^i) -
                \phi(x^i, \hat{y})] \]
        by enumeration of $C^i$ and continue until
        line~\ref{convergence_check}.  Only if $\xi' - \xi < \epsilon$, that is
        the produced constraint is not violated, we return to
        line~\ref{get_cutting_plane} and actually invoke the loss augmented
        prediction $I$.
\end{description}

\subsection{Stochastic Dual Coordinate Descent}\seclabel{dual_coordinate_descent}

### 2013-09-04 pystruct/paper.tex
better tables for library comparision

@@
 
+\subsection{Library Structure and Content}
 Using the formulation of structured prediction introduced in
@@
 data-dependent pairwise potentials, and several latent variable models.
+
 The maximization for finding $y$ in Equation~\ref{eq:main_equation} is carried out
@@
 
-Table~\ref{table:comparision} lists algorithms and models that are implemented
-in \pystruct and compares them to other public structured prediction libraries,
-together with the programming language and the project license.
+Table~\ref{table:comparision_algorithms} lists learning algorithms that are implemented
+in \pystruct and compares them to other public structured prediction libraries.
+Table~\ref{table:comparision_models} lists
+
+%TODO what about language / license?
+\begin{table}[t]
+\centering
+\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
+\toprule
+Package       & Cutting Plane& SSGD & BCFW & Latent SSVM & Perceptron & ML\\
+\cmidrule(r){1-1} \cmidrule(r){2-7}
+\pystruct     & \x        & \x   & \x   & \x          & \x         & \o\\
+\svmstruct    & \x        & \o   & \o   & \x          & \o         & \o\\
+\sc{Dlib}     & \x        & \x   & \o   & \o          & \x         & \o\\
+\sc{CRFsuite} & \o        & \o   & \o   & \o          & \x         & \x\\
+
+\bottomrule
+\end{tabularx}
+    \caption{\label{table:comparision_algorithms}Comparison of learning
+        algorithms implemented in popular structured prediction software
+        packages. SSGD stochastic subgradient decent learning of SSVMs, BCFW for
+    Block-Coordinate Frank-Wolfe, Latent SSVM for latent variable SSVMs, ML for
+    maximum likelihood learning}
+\end{table}
 
@@
 \centering
-\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccccccc}
+\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
 \toprule
-Package &     Language &     License&\multicolumn{4}{c}{Algorithms}&\multicolumn{3}{c}{Models} \\
-\cmidrule(r){1-1} \cmidrule(r){2-2} \cmidrule(r){3-3} \cmidrule(r){4-7} \cmidrule{8-10}
-&             &&                     \footnotesize{CP}& \footnotesize{SG}& \footnotesize{LV}& \footnotesize{ML}& \footnotesize{Chain} & \footnotesize{Graph} & \footnotesize{LDCRF}\\
-\pystruct&      Python &       BSD$^1$  & \x$^1$    & \x      & \x   & \o & \x     & \x     & \x \\
-\svmstruct & C++ & non-free         & \x    & \o      & \x   & \o & \o     & \o     & \o \\
-\sc{Dlib}         & C++        & boost            & \x    & \x      & \o   & \o & \x     & \x     &\o\\
-\sc{CRFsuite}     & C++        & BSD              & \o    & \o      & \o   & \x & \x     & \o     &\o\\
+Package        &Multi-Class &  Multi-Label & Chain & Graph & LSVM  &LDCRF \\
+\cmidrule(r){1-1} \cmidrule(r){2-7}
+\pystruct      & \x         & \x           & \x    & \x    & \x    & \x    \\
+\svmstruct     & \x         & \x           & \o    & \o    & \o    & \o    \\
+\sc{Dlib}      & \x         & \o           & \x    & \x    & \o    & \o    \\
+\sc{CRFsuite}  & \o         & \o           & \x    & \o    & \o    & \o    \\
 
@@
 \end{tabularx}
-    \caption{\label{table:comparision}Comparison of structured prediction
-        software packages. CP stands for cutting plane optimization of
-        SSVMs~\citep{joachims2009cutting}, SG for online subgradient
-        optimization of SSVMs~\citep{ratliff2007online}, LV for latent variable
-        SSVMs~\citep{yu2009learning}, ML for maximum likelihood learning, Chain
-        for chain-structured models with pairwise interactions, Graph for
-        arbitrary graphs with pairwise interactions, and LDCRF for latent
-        dynamic CRF~\citep{morency2007latent}.  {\footnotesize $^1$\pystruct
-        itself is BSD licensed, but uses the GPL-licensed package {\sc cvxopt}
-    for cutting-plane learning.}
-}
+    \caption{\label{table:comparision_models}Comparison of models implemented
+    in popular structured prediction software packages. LSVM stands for the latent multi-class SVM, LDCRF for
+    latend dynamic conditional random fields.}
 \end{table}
-\vfill
-\pagebreak
 

### 2013-09-04 structured_prediction/paper.tex
start on stdjsp description

@@
 
-\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
 \begin{algorithm*}[t]
@@
     \end{algorithm*}
+
+\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
 
@@
 
-\subsection{Stochastic Dual Coordinate Descent}\seclabel{dual_coordinate_descent}
+\subsection{The STLJSP algorithm}\seclabel{dual_coordinate_descent}
+Very recently, two groups independently derived a very performant new algorithm.
+\citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
+to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
+and derived a block-coordinate version, where each block corresponds to the constraints associated with a single training example.
+Both groups showed that closed-form line search is possible, and, surprisingly ended up with
+the exact same algorithm and closely related convergence guarantees.
+The algorithm processes a single training example at a time, and can be applied while
+remaining completely in the primal domain, making it applicable to very large datasets.
+We will refer to the algorithm as STDJSP, after the authors of the two papers, to honor
+the independent development of the two views of the method.
+The STDJSP algorithm has properties that are very simliar to stochastic subgradient descent,
+but with two main advantages:
+\begin{description}
+\item[Stopping Criterion] Both views of STLJSP give rise to a dual objective, which can be used as a theoretically sound stopping criterion.
+\item[Learning Rate] TODO %TODO
+\end{description}

### 2013-09-05 structured_prediction/paper.tex
structured prediction: More elaborate description of learning algorithms

@@
 cutting plane algorithms, a stochastic primal subgradient algorithm, and
-recently proposed stochastic dual coordinate descent method.
+recently proposed stochastic dual coordinate descent method, which we
+will now describe in detail.
+
+% discuss online vs batch?
+
+We will also discuss practical implications and implementation.
+One particularly intereresting aspect is parallelization.
+In image segmentation tasks, inference is often costly, making loss-augmented
+prediction the most expensive step in learning. In \emph{batch} algorithms,
+which process all training examples at once, loss-augmented inference is
+embarassingly parallel, leading to a linear speed-up in terms of processes.
 
 \subsection{Stochastic Subgradient Descent}\seclabel{subgradient}
+
 Arguably the most straight-foward way to approach \Eqref{learning_equation} is
@@
 Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
-%\begin{equation}
-
-%\end{equation}
-
-%\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
-%\begin{equation}\label{subgradient}
-    %g := \phi(x_i, y) - \phi(x_i, y_i) + \theta / C
-%\end{equation}
-%leading to a simple iterative optimization.
+\begin{equation}
+    \eta_t = \frac{C}{t},
+\end{equation}
+which has been found to work well in many settings.
+Implementation of the algorithm is straight-forward, but unfortunately
+detecting convergence is often tricky.
+It is possible to use mini-batches instead of processing one sample at a time
+to make use of multiple processors for inference. Unfortunately this negatively
+affects the number of iterations needed, and did not provide a benefit in practice.
 
@@
 which guarantees a suboptimality of at most $\epsilon$.
-
+%
 The complete procedure is described in Algorithm~\ref{alg_n_slack}.
+The n-slack cutting plane can be understood is a sequential algorithm that processes each sample individually.
+While this allows fast process of the optimization with respect to the number of calls to loss-augmented prediction,
+individual iterations become more and more costly. The number of the constraints is usually a multiple of
+the number of training  samples, which leads to very large QP problems, even for medium sized datasets.
+This makes the algorithm often slow in practice.
+%
+Solving the quadratic programm can be accelerated using several techniques. We found that
+agressively removing constraints that are inactive or contribute little to the solution
+often makes the difference between the algorithm being practical or not.
+Another possible technique is to update the quadratic program only every $r$ samples, for some
+small integer $r$ (\citet{joachims2009cutting} suggest 100). While this strategy on its own
+did not provide a large benefits in our experiments, it allows for parallel loss-augmented
+inference on these mini-batches of size $r$.
 
@@
 A detailed description can be found in Algorithm~\ref{alg_one_slack}.
-
+%
 By construction only a single constraint is added in each iteration of
 Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
-This has the advantage of making the solution of the QP faster, as it contains
-less variables. The down-side of this is that the loss-augmented prediction
-problem has to be solve much more often until convergence.
+This has the advantage of making the solution of the QP faster, as it contains far
+less variables than in the $n$-slack algorithm. The down-side of this is that
+the loss-augmented prediction problem has to be solve much more often until
+convergence.
 %TODO convergence rate
@@
 \subsection{The STLJSP algorithm}\seclabel{dual_coordinate_descent}
+
 Very recently, two groups independently derived a very performant new algorithm.
@@
 to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
-and derived a block-coordinate version, where each block corresponds to the constraints associated with a single training example.
-Both groups showed that closed-form line search is possible, and, surprisingly ended up with
-the exact same algorithm and closely related convergence guarantees.
-The algorithm processes a single training example at a time, and can be applied while
-remaining completely in the primal domain, making it applicable to very large datasets.
-We will refer to the algorithm as STDJSP, after the authors of the two papers, to honor
-the independent development of the two views of the method.
-The STDJSP algorithm has properties that are very simliar to stochastic subgradient descent,
-but with two main advantages:
+and derived a block-coordinate version, where each block corresponds to the
+constraints associated with a single training example.  Both groups showed that
+closed-form line search is possible, and, surprisingly ended up with the exact
+same algorithm and closely related convergence guarantees.  The algorithm
+processes a single training example at a time, and can be applied while
+remaining completely in the primal domain, making it applicable to very large
+datasets.  We will refer to the algorithm as STLJSP, after the authors of the
+two papers, to honor the independent development of the two views of the
+method.
+%
+The STDJSP algorithm has properties that are very simliar to
+stochastic subgradient descent, but with two main advantages:
 \begin{description}
-\item[Stopping Criterion] Both views of STLJSP give rise to a dual objective, which can be used as a theoretically sound stopping criterion.
-\item[Learning Rate] TODO %TODO
+\item[Stopping Criterion] Both views of STLJSP give rise to a dual objective,
+    which can be used as a theoretically sound stopping criterion.
+\item[Learning Rate] While there are several theoretical results on choosing
+    the learning rate in stochastic subgradient descent, choosing a concrete
+    schedule is often problematic in practice. By using analytic line-search,
+    STLJSP removes the need for any learning rate parameter, making application
+    of the algorithm in practice much simpler.
 \end{description}
+
+The STLJSP algorithm is usually used with weighted averaging, as this is known
+to improve performance in similar stochastic algorithms, and empirically
+improves convergence. % TODO weighted averaging here? or at SGD?

### 2013-09-05 pystruct/paper.tex
added averaging, added SZLJSP algorithm

@@
 \toprule
-Package       & Cutting Plane& SSGD & BCFW & Latent SSVM & Perceptron & ML\\
+Package       & Cutting Plane& SSGD & STLJSP & Latent SSVM & Perceptron & ML\\
 \cmidrule(r){1-1} \cmidrule(r){2-7}
@@
         algorithms implemented in popular structured prediction software
-        packages. SSGD stochastic subgradient decent learning of SSVMs, BCFW for
-    Block-Coordinate Frank-Wolfe, Latent SSVM for latent variable SSVMs, ML for
-    maximum likelihood learning}
+        packages. SSGD stochastic subgradient decent learning of SSVMs, STLJSP is
+        as described in \secref{dual_coordinate_descent}, Latent SSVM for latent variable SSVMs, ML for
+        maximum likelihood learning.}
 \end{table}

### 2013-09-05 structured_prediction/paper.tex
added averaging, added SZLJSP algorithm

@@
 \begin{equation}
-    \theta \leftarrow (1 - \eta_t) \theta - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
+    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
 \end{equation}
@@
 but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
-Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
+Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
 \begin{equation}
@@
 which has been found to work well in many settings.
-Implementation of the algorithm is straight-forward, but unfortunately
-detecting convergence is often tricky.
+\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
+\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
+a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
+\begin{equation}
+    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
+\end{equation}
+This $t$-weighted averaging can be computed on-the-fly as
+\begin{equation}\eqlabel{averaging}
+    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}
+\end{equation}
+
+Implementation of the stochastic subgradient algorithm (with or without
+averaging) is straight-forward, but unfortunately detecting convergence is
+often tricky.
 It is possible to use mini-batches instead of processing one sample at a time
@@
 
-\subsection{The STLJSP algorithm}\seclabel{dual_coordinate_descent}
+%TODO BMRM?
+
+\subsection{The SZLJSP algorithm}\seclabel{dual_coordinate_descent}
+\begin{algorithm*}[t]
+    \caption{Shalev-Shwartz Zhang Lacoste-Julien Jaggi Schmidt Pletscher\label{alg_szljsp}}
+    \begin{algorithmic}[1]
+        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
+        \Ensure parameters $\theta$
+        \State $\theta_0, \theta_0\hoch{i}, \bar{\theta}_0  \leftarrow \mathbf{0}, \ell_0, \ell_0\hoch{i} \leftarrow 0 $
+        \State $t \leftarrow 0$
+        \Repeat
+            \State $t \leftarrow t + 1$
+            \State Pick $i$ uniformly at random from \{1, \dotsc, k\}.
+            \State $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
+            \State Compute parameter and loss updates based on sample $i$:\\
+                $\theta_s \leftarrow \frac{C}{n} \phi(\hat{y})$
+                $l_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
+            \State Compute optimum step size $\eta$:\\
+                $\eta \leftarrow \frac{(\theta_t\hoch{i} - \theta_s)^T \theta_t + C (\ell_s - \ell_k\hoch{i} )}{\lVert \theta_t\hoch{i} - \theta_s \rVert^2}$ and clip to $[0, 1]$
+            \State Update per-sample parameters and loss estimate:\\
+                $\theta_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \theta_s$\\
+                $\ell_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \ell_s$
+            \State Update global parameters and loss estimate:\\
+                $\theta_{t+1} \leftarrow \theta_{t+1} + \theta_{t}\hoch{i} - \theta_{t+1}\hoch{i}$\\
+                $\ell_{t+1} \leftarrow \ell_{t+1} + \ell_{t}\hoch{i} - \ell_{t+1}\hoch{i}$
+            \State Compute the weighted running average:\\
+            $\bar{\theta}_{t+1} = \frac{k}{k + 2} \bar{\theta}_k + \frac{2}{k + 2}\theta_{k+1}$
+            \vspace{-5mm}
+        \Until
+    \end{algorithmic}
+\end{algorithm*}
 
@@
 remaining completely in the primal domain, making it applicable to very large
-datasets.  We will refer to the algorithm as STLJSP, after the authors of the
+datasets.  We will refer to the algorithm as SZLJSP, after the authors of the
 two papers, to honor the independent development of the two views of the
@@
 %
-The STDJSP algorithm has properties that are very simliar to
+The SZLJSP algorithm has properties that are very simliar to
 stochastic subgradient descent, but with two main advantages:
 \begin{description}
-\item[Stopping Criterion] Both views of STLJSP give rise to a dual objective,
+\item[Stopping Criterion] Both views of SZLJSP give rise to a dual objective,
     which can be used as a theoretically sound stopping criterion.
@@
     schedule is often problematic in practice. By using analytic line-search,
-    STLJSP removes the need for any learning rate parameter, making application
+    SZLJSP removes the need for any learning rate parameter, making application
     of the algorithm in practice much simpler.
@@
 
-The STLJSP algorithm is usually used with weighted averaging, as this is known
-to improve performance in similar stochastic algorithms, and empirically
-improves convergence. % TODO weighted averaging here? or at SGD?
+In practice, the SZLJSP algorithm is usually used with the weighted averaging
+described in \Eqref{averaging}, as this is known to improve performance in the
+closely related stochastic subgradient algorithms, and empirically improves
+convergence.
+
+The detailed procedure as given by \citet{lacoste2012block} is shown in Algorithm~\ref{alg_szljsp}.

### 2013-09-05 structured_prediction/paper.tex
polish SZLJSP algorithm listing

@@
     \caption{Shalev-Shwartz Zhang Lacoste-Julien Jaggi Schmidt Pletscher\label{alg_szljsp}}
+    \begin{doublespacing}
     \begin{algorithmic}[1]
@@
         \Ensure parameters $\theta$
-        \State $\theta_0, \theta_0\hoch{i}, \bar{\theta}_0  \leftarrow \mathbf{0}, \ell_0, \ell_0\hoch{i} \leftarrow 0 $
-        \State $t \leftarrow 0$
+        \State $\theta_0, \theta_0\hoch{i}, \bar{\theta}_0  \leftarrow \mathbf{0},\quad \ell_0, \ell_0\hoch{i}, t \leftarrow 0 $
         \Repeat
             \State $t \leftarrow t + 1$
-            \State Pick $i$ uniformly at random from \{1, \dotsc, k\}.
-            \State $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
-            \State Compute parameter and loss updates based on sample $i$:\\
-                $\theta_s \leftarrow \frac{C}{n} \phi(\hat{y})$
-                $l_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
-            \State Compute optimum step size $\eta$:\\
-                $\eta \leftarrow \frac{(\theta_t\hoch{i} - \theta_s)^T \theta_t + C (\ell_s - \ell_k\hoch{i} )}{\lVert \theta_t\hoch{i} - \theta_s \rVert^2}$ and clip to $[0, 1]$
-            \State Update per-sample parameters and loss estimate:\\
-                $\theta_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \theta_s$\\
-                $\ell_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \ell_s$
-            \State Update global parameters and loss estimate:\\
-                $\theta_{t+1} \leftarrow \theta_{t+1} + \theta_{t}\hoch{i} - \theta_{t+1}\hoch{i}$\\
-                $\ell_{t+1} \leftarrow \ell_{t+1} + \ell_{t}\hoch{i} - \ell_{t+1}\hoch{i}$
-            \State Compute the weighted running average:\\
-            $\bar{\theta}_{t+1} = \frac{k}{k + 2} \bar{\theta}_k + \frac{2}{k + 2}\theta_{k+1}$
-            \vspace{-5mm}
-        \Until
+            \State Pick $i$ uniformly at random from $\{1, \dotsc, k\}$.
+            \State Perform loss-augmented prediction on sample $i$:
+            \Statex[2]   $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \Delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$
+            \State Compute parameter and loss updates based on sample $i$:
+            \Statex[2]     $\theta_s \leftarrow \frac{C}{n} \phi(\hat{y})$
+            \Statex[2]     $l_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
+            \State Compute optimum step size $\eta$:
+            \Statex[2]    $\eta \leftarrow \frac{(\theta_t\hoch{i} - \theta_s)^T \theta_t + C (\ell_s - \ell_k\hoch{i} )}{\lVert \theta_t\hoch{i} - \theta_s \rVert^2}$ and clip to $[0, 1]$
+            \State Update per-sample parameters and loss estimate:
+            \Statex[2]    $\theta_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \theta_s$
+            \Statex[2]    $\ell_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \ell_s$
+            \State Update global parameters and loss estimate:
+            \Statex[2]   $\theta_{t+1} \leftarrow \theta_{t+1} + \theta_{t}\hoch{i} - \theta_{t+1}\hoch{i}$
+            \Statex[2]   $\ell_{t+1} \leftarrow \ell_{t+1} + \ell_{t}\hoch{i} - \ell_{t+1}\hoch{i}$
+            \State Compute the weighted running average:
+            \Statex[2] $\bar{\theta}_{t+1} = \frac{k}{k + 2} \bar{\theta}_k + \frac{2}{k + 2}\theta_{k+1}$
+        \Until $(\theta - \theta_s)^T\theta - \ell + \ell_s \leq \epsilon$
+        \Statex where $\theta_s$ and $\ell_s$ are recomputed over the whole dataset.
     \end{algorithmic}
+    \end{doublespacing}
 \end{algorithm*}

### 2013-09-06 pystruct/paper.tex
some typos, korrektur lesen....

--- 2013-09-04 pystruct/paper.tex
@@
 
-Unfortunately only few implementations are publicly
-available---many applications are based on the non-free implementation
+Unfortunately only few implementations for learning structured prediction are
+publicly available---many applications are based on the non-free implementation
 of~\citet{joachims2009cutting}.
-We provide an implementation, \pystruct, which aims at providing a high-quality
-code with an easy-to-use interface, in the high-level Python language.  This
-allows practitioners to efficiently test a range of models, as well as allowing
-researchers to compare to baseline methods much more easily than this is
-possible with current implementations.  \pystruct is BSD-licensed, allowing
-modification and redistribution of the code, as well as use in commercial
-applications.  By embracing paradigms established in the scientific Python
-community and reusing the interface of the widely-used {\sc scikit-learn}
-library~\citep{pedregosa2011scikit}, \pystruct can be used in existing
-projects, replacing standard classifiers. The online documentation and
+In this section, we introduce our implementation, \pystruct, which aims at
+providing a high-quality code with an easy-to-use interface, in the high-level
+Python language. This allows practitioners to efficiently test a range of
+models, as well as allowing researchers to compare to baseline methods much
+more easily than this is possible with current implementations. \pystruct is
+BSD-licensed, allowing modification and redistribution of the code, as well as
+use in commercial applications.  By embracing paradigms established in the
+scientific Python community and reusing the interface of the widely-used {\sc
+scikit-learn} library~\citep{pedregosa2011scikit}, \pystruct can be used in
+existing projects, replacing standard classifiers. The online documentation and
 examples help new users understand the somewhat abstract ideas behind
 structured prediction.
-We will base the experiments in the rest of this work on the algorithms
-implemented in \pystruct.
+We base the experiments in the rest of this work on the algorithms
+implemented in \pystruct. In the following, we briefly discuss implementation
+and features of the \pystruct library.
 
@@
     \item Solving the loss-augmented prediction problem in Equation~\eqref{loss_augmentation}.
-    \item Optimizing the objective \Eqref{learning_equation} with respect to $\theta$.
+    \item Optimizing the objective in \Eqref{learning_equation} with respect to $\theta$.
 \end{enumerate*}
@@
 support cutting plane algorithms for structural support vector
-machines~(SSVMs), the Block-Coordinate Frank Wolfe (BCFW) algorithm for SSVMs,
+machines~(SSVMs), the SZLJSP algorithm for SSVMs,
 subgradient methods for SSVMs, the structured perceptron and latent variable
-SSVMs. The cutting plane implementation uses the {\sc cvxopt}
-package \citep{dahl2006cvxopt} for quadratic optimization.
+SSVMs. See \Secref{learning_algorithms} for a detailed description of the
+algorithms. The cutting plane implementation uses the {\sc cvxopt} package
+\citep{dahl2006cvxopt} for quadratic optimization.
 
@@
 
-Table~\ref{table:comparision_algorithms} lists learning algorithms that are implemented
-in \pystruct and compares them to other public structured prediction libraries.
-Table~\ref{table:comparision_models} lists
+Table~\ref{table:comparision_algorithms} and
+Table~\ref{table:comparision_models} list learning algorithms and models that
+are implemented in \pystruct and compares them to other publicly available
+structured prediction libraries.
 
@@
 \toprule
-Package       & Cutting Plane& SSGD & BCFW & Latent SSVM & Perceptron & ML\\
+Package       & Cutting Plane& SSGD & SZLJSP & Latent SSVM & Perceptron & ML\\
 \cmidrule(r){1-1} \cmidrule(r){2-7}
@@
         algorithms implemented in popular structured prediction software
-        packages. SSGD stochastic subgradient decent learning of SSVMs, BCFW for
-    Block-Coordinate Frank-Wolfe, Latent SSVM for latent variable SSVMs, ML for
-    maximum likelihood learning}
+        packages. SSGD stochastic subgradient decent learning of SSVMs, SZLJSP is
+        as described in \secref{dual_coordinate_descent}, Latent SSVM for latent variable SSVMs, ML for
+        maximum likelihood learning.}
 \end{table}
@@
      \pystruct contains a testing-suite with 80\% line-coverage. It also
-     employs continuous integration to ensure stability and seamless user
+     employs continuous integration to ensure stability and a seamless user
      experience.
@@
 creates a \texttt{learner} object that will learn the parameters for the given
-model using the n-slack cutting plane method, and line 6 performs the actual
+model using the $n$-slack cutting plane method, and line 6 performs the actual
 learning.  Using this simple setup, we achieve an accuracy of 30.3 on the
@@
 important that the implemented learning algorithms run in acceptable time.
-In this section, we compare our implementation of the one-slack cutting plane
+In this section, we compare our implementation of the $1$-slack cutting plane
 algorithm with the implementation in \svmstruct.
@@
 prediction point of view, this problem is well-suited to benchmark the cutting
-plane solvers with respect to accuracy and speed.
+plane solvers, as loss-augmented prediction is trivial.
 
@@
 
-\section{Comparing Learning Algorithms}
+%\section{Comparing Learning Algorithms}
 %TODO experiments!!

### 2013-09-06 structured_prediction/paper.tex
some typos, korrektur lesen....

@@
 structured prediction, as well as in their application in areas such as natural
-language processing and computer vision (see \citet{nowozin2011structured} %TODO ms book too!
-for an overview).
+language processing and computer vision (see \citet{nowozin2011structured}
+for an introduction and \citet{blake2011markov} for a recent survey).
 %
@@
 algorithms, \pystruct.
+
+\pagebreak %FIXME maybe?
 
@@
 $\mathcal{Y}$ is typically very large, it is crucial to exploit the particular
-form of $\Phi(x, y)$ to solve the inference problem of \Eqref{main_equation}.
+form of $\Phi(x, y)$ to solve the prediction problem of \Eqref{main_equation}.
 
@@
 configuration of a molecule, many setting, such as the image segmentation
-setting we are interested, can be reduced to the case where $y$ is a vector of
+setting we are interested in, can be reduced to the case where $y$ is a vector of
 discrete labels $\mathcal{Y} = \{1, \dotsc, k\}^n$.
 In the following, we will only discuss this multivariate case. In general, $n$
-is often different for different inputs $x$, such as image with different
-number of (super) pixels.  We will ignore this in our notation to simplify the
+is often different for different inputs $x$, such as images with different
+numbers of pixels.  We will ignore this in our notation to simplify the
 presentation.
@@
  represent \emph{products} of factors.  To simplify presentation, we work
-directly in the log-domain of the product representation.}
+directly in the log-domain with respect to the more standard product representation.}
 \begin{equation}\eqlabel{log_factor_graph}
@@
 The benefit of using the factor graph representation is that it decomposes the
-function over the variables of interest $y_i$ and makes. This allows us to
-apply efficient optimization procedures for the maximization in
+function over subsets the variables of interest $y_i$. This allows us to apply
+efficient optimization procedures for the prediction problem in
 \Eqref{main_equation} by exploiting the graph structure of the factor graph.
-
+%
 To obtain a linear function as in \Eqref{main_equation} from
@@
 
-The benefit of using the factor graph representation is that it allows us to
-apply efficient optimization procedures for the maximization in
-\Eqref{main_equation} by exploiting the graph structure of the factor graph.
-
-This approach to structured prediction is closely related to a probabilistic
-approach using \emph{graphical models}.  Probabilistic graphical models are a
+This approach to structured prediction is closely related to 
+approachs using \emph{probabilistic graphical models}.  Probabilistic graphical models are a
 tool to express factorizations of probability distributions.  Similar to
@@
 
-The most probable prediction $y$ is given as $\argmax_{y \in \mathcal{Y}}p(y|x)$.
+The most probable prediction $y$ is given as $\displaystyle \argmax_{y \in \mathcal{Y}}p(y|x)$.
 As $Z$ is independent of $y$, and by the monotonicity of the logarithm,
@@
 
-\section{Learning Algorithms for Max-Margin Structured Prediction}
+\section{Learning Algorithms for Max-Margin Structured Prediction}\seclabel{learning_algorithms}
 
-Maximum Margin learning is now one of the most popular methods to learn
-structural models in computer vision and text processing.
+Maximum margin learning has become one of the most popular methods to learn
+classifiers and structural models in computer vision and text processing.
 There are several reasons for the popularity of linear maximum margin approaches:
 \begin{description}
-    \item[Loss-sensitivity] In contrast to probabilistic approaches, maximum margin learning approaches can directly
+    \item[Loss-sensitivity] In contrast to most probabilistic approaches, maximum margin learning approaches can directly
         minimize a user-specified loss.
@@
 \end{equation}
-with regularization parameter $C$. Here $r_i$ is an hinge-loss like upper bound
-on $\Delta$ empirical risk:
+with regularization parameter $C$. Here $\ell$ is an hinge-loss like upper bound
+on the empirical $\Delta$-risk:
 \begin{equation}\eqlabel{loss_augmentation}
@@
 There are several popular algorithms to solve \Eqref{learning_equation}. We
-will briefly review three standard algorithms: the one-slack and n-slack
-cutting plane algorithms, a stochastic primal subgradient algorithm, and
-recently proposed stochastic dual coordinate descent method, which we
-will now describe in detail.
+will briefly review three standard algorithms, and a very recent one: the
+$1$-slack and $n$-slack cutting plane algorithms, a stochastic primal subgradient
+algorithm, and recently proposed stochastic dual coordinate descent method,
+which we will now describe in detail.
 
@@
 
-We will also discuss practical implications and implementation.
-One particularly intereresting aspect is parallelization.
+We will also discuss practical implications and implementation.  One
+particularly intereresting aspect is the difference between \emph{sequention}
+(or \emph{online}) and \emph{batch} algorithms.  Batch algorithms process the
+whole dataset before adjusting parameters, while sequential algorithms process
+one sample at a time, and adjust parameters incrementally.
 In image segmentation tasks, inference is often costly, making loss-augmented
-prediction the most expensive step in learning. In \emph{batch} algorithms,
-which process all training examples at once, loss-augmented inference is
-embarassingly parallel, leading to a linear speed-up in terms of processes.
+prediction the most expensive step in learning. This often leads to longer
+run-times for batch algorithms. On the other hand, loss-augmented prediction in
+batch algorithms is embarassingly parallel, allowing the use of multiple
+processors with almost linear speed improvements.
 
@@
 It is possible to use mini-batches instead of processing one sample at a time
-to make use of multiple processors for inference. Unfortunately this negatively
-affects the number of iterations needed, and did not provide a benefit in practice.
+to make use of multiple processors for loss-augmented prediction. Unfortunately
+this negatively affects the number of iterations needed, and did not provide a
+benefit in our experiments.
 
@@
 
-The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
+The $n$-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
 into a quadratic objective with a combinatorial number of constraints:
@@
 \end{align}
-As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
+As it is not feasible to deal with all constraints, only a working set $\W$ of active constraints
 is maintained, using the cutting plane method. The algorithm starts with an empty working set,
-and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
+and repeatedly iterates over the training data. For each sample, the most
+violated constraint is added to $\W$, and the quadratic program is solved
 again, with the new set of constraints.
@@
 The complete procedure is described in Algorithm~\ref{alg_n_slack}.
-The n-slack cutting plane can be understood is a sequential algorithm that processes each sample individually.
-While this allows fast process of the optimization with respect to the number of calls to loss-augmented prediction,
-individual iterations become more and more costly. The number of the constraints is usually a multiple of
-the number of training  samples, which leads to very large QP problems, even for medium sized datasets.
-This makes the algorithm often slow in practice.
+The $n$-slack cutting plane is a sequential algorithm that processes each sample
+individually. While this allows fast process of the optimization with respect
+to the number of calls to loss-augmented prediction, individual steps
+become more and more costly. The number of the constraints is usually a
+multiple of the number of training  samples, which leads to very large QP
+problems, even for medium sized datasets. This makes the algorithm often slow
+in practice.
 %
-Solving the quadratic programm can be accelerated using several techniques. We found that
-agressively removing constraints that are inactive or contribute little to the solution
-often makes the difference between the algorithm being practical or not.
-Another possible technique is to update the quadratic program only every $r$ samples, for some
-small integer $r$ (\citet{joachims2009cutting} suggest 100). While this strategy on its own
-did not provide a large benefits in our experiments, it allows for parallel loss-augmented
-inference on these mini-batches of size $r$.
+Solving the quadratic programm can be accelerated using several techniques. We
+found that agressively removing constraints that are inactive or contribute
+little to the solution often makes the difference between the algorithm being
+practical or not.  Another possible technique is to update the quadratic
+program only every $r$ samples, for some small integer $r$
+(\citet{joachims2009cutting} suggest 100). While this strategy on its own did
+not provide a large benefits in our experiments, it allows for parallel
+loss-augmented prediction on these mini-batches of size $r$.
 
@@
 
-The one-slack cutting plane method~\citep{joachims2009cutting} solves the
+The $1$-slack cutting plane method~\citep{joachims2009cutting} solves the
 following reformulation of Equation~\Eqref{learning_equation}:
@@
 \end{align}
-Informally, the one-slack formulation corresponds to joining all training
+Informally, the $1$-slack formulation corresponds to joining all training
 samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
 interactions between variables corresponding to different data points, and
-then applying the n-slack algorithm.
+then applying the $n$-slack algorithm with a single data point.
 A detailed description can be found in Algorithm~\ref{alg_one_slack}.
@@
 \begin{description}
-    \item[Constraint Pruning] Members of the working set $\W$ can become inactive during learning.
-        If a constraint has been inactive for a number of iterations, it is removed from $\W$, leading
-        to smaller problem sizes.
-    \item[Inference Caching] In the one-slack algorithm, each constraint is
+    \item[Constraint Pruning] As in the $n$-slack algorithm, members of the
+        working set $\W$ can become inactive during learning.  If a constraint
+        has been inactive for a number of iterations, it is removed from $\W$,
+        leading to smaller problem sizes.
+    \item[Inference Caching] In the $1$-slack algorithm, each constraint is
         created using a combination of loss-augmented prediction results.
@@
         To exploit this, we maintain a set $C^i$ of the last $r$ results of
-        loss-augmented inference for each training example $(x^i, y^i)$
+        loss-augmented prediction for each training example $(x^i, y^i)$
         (line~\ref{get_cutting_plane} in Algorithm~\ref{alg_one_slack}).
@@
         line~\ref{convergence_check}.  Only if $\xi' - \xi < \epsilon$, that is
-        the produced constraint is not violated, we return to
+        the produced constraint is not violated stronly enough, we return to
         line~\ref{get_cutting_plane} and actually invoke the loss augmented
@@
         \Ensure parameters $\theta$
-        \State $\theta_0, \theta_0\hoch{i}, \bar{\theta}_0  \leftarrow \mathbf{0},\quad \ell_0, \ell_0\hoch{i}, t \leftarrow 0 $
+        \State  $\displaystyle \theta_0^{\vphantom{i}}, \theta_0\hoch{i}, \bar{\theta}_0^{\vphantom{i}} \leftarrow \mathbf{0},\quad \ell_0^{\vphantom{i}}, \ell_0\hoch{i}, t \leftarrow 0 $
         \Repeat
@@
             \Statex[2]     $\theta_s \leftarrow \frac{C}{n} \phi(\hat{y})$
-            \Statex[2]     $l_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
+            \Statex[2]     $\ell_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
             \State Compute optimum step size $\eta$:
@@
 datasets.  We will refer to the algorithm as SZLJSP, after the authors of the
-two papers, to honor the independent development of the two views of the
-method.
+two papers.
 %
@@
     SZLJSP removes the need for any learning rate parameter, making application
-    of the algorithm in practice much simpler.
+    of the algorithm much simpler.
 \end{description}

### 2013-09-06 structured_prediction/paper.tex
convergence rates, citations, mehr todos fertig...

@@
 
-Structured prediction can be defined as making a prediction $f(x)$ by maximizing a
-compatibility function $g$ between an input $x$ and possible labels
+Structured prediction can be defined as making a prediction $f(x)$ by
+maximizing a compatibility function $g$ between an input $x$ and possible
+labels
 $y$~\citep{nowozin2011structured}:
@@
 or prediction.
-Most current approaches use linear functions,
-with some notable examples: % TODO cite dtf, conditional neural fields
+Most current approaches use linear functions, with some notable exceptions:
+\citet{peng2009conditional} introduce a conditional random field for sequence
+prediction where single-node potentials are computed using neural networks,
+while \citet{nowozin2011decision} propose a conditional random field for image
+processing, where all potentials are computed using decision trees.
 
@@
 discrete labels $\mathcal{Y} = \{1, \dotsc, k\}^n$.
-In the following, we will only discuss this multivariate case. In general, $n$
+In the following, we only discuss this multivariate case. In general, $n$
 is often different for different inputs $x$, such as images with different
-numbers of pixels.  We will ignore this in our notation to simplify the
+numbers of pixels.  We ignore this in our notation to simplify the
 presentation.
@@
 predictions, not modeling probabilities, there are no clear benefits from the
-probabilistic approach.  Consequently, we will work with the more direct
+probabilistic approach.  Consequently, we work with the more direct
 structured prediction approach of \Eqref{main_equation} and
@@
 There are several popular algorithms to solve \Eqref{learning_equation}. We
-will briefly review three standard algorithms, and a very recent one: the
+briefly review three standard algorithms, and a very recent one: the
 $1$-slack and $n$-slack cutting plane algorithms, a stochastic primal subgradient
 algorithm, and recently proposed stochastic dual coordinate descent method,
-which we will now describe in detail.
+which we now describe in detail. We also give simplified known
+convergence rates in terms of calls to the loss-augmented prediction.
 
-% discuss online vs batch?
-
-We will also discuss practical implications and implementation.  One
+Additionally, we discuss practical implications and implementation.  One
 particularly intereresting aspect is the difference between \emph{sequention}
@@
 \end{equation}
-Where $\eta_t$ is a sequence of learning rates.
+Where $\eta_t$ is a sequence of step sizes.
 %Theoretical convergence guarantees of \ldots can be given whenever \ldots,
@@
     \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
-    &\theta^T \sum_{i=1}^n [\phi(x\hoch{i}, y\hoch{i}) - \phi(x\hoch{i},
+    &\theta^T \sum_{i=1}^n [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i},
         \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
@@
 %
+\citet{tsochantaridis2006large} showed a convergence rate of $O(\frac{1}{T^2})$
+with respect to calls to loss-augmented prediction. The recent work
+of~\citet{lacoste2012block}, however, suggests a rate of $O(\frac{1}{T})$,
+which empirically seems more plausible. This analysis does not include the cost
+of solving the QP, which depends on the size of the dataset and the number of
+iterations. 
 The complete procedure is described in Algorithm~\ref{alg_n_slack}.
+
 The $n$-slack cutting plane is a sequential algorithm that processes each sample
@@
     \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
-        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
+        &\theta^T \sum_{i=1}^n [\Phi(x^i, y^i) - \Phi(x^i,
             \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
@@
 the loss-augmented prediction problem has to be solve much more often until
-convergence.
-%TODO convergence rate
+convergence. This is reflected in a convergence rate of $O(\frac{1}{kT})$,
+which scales with the inverse of the dataset size.
 
@@
         \[ \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
-            \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\phi(x^i, y^i) -
-                \phi(x^i, \hat{y})] \]
+            \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\Phi(x^i, y^i) -
+                \Phi(x^i, \hat{y})] \]
         by enumeration of $C^i$ and continue until
@@
             \State Compute parameter and loss updates based on sample $i$:
-            \Statex[2]     $\theta_s \leftarrow \frac{C}{n} \phi(\hat{y})$
+            \Statex[2]     $\theta_s \leftarrow \frac{C}{n} \Phi(x, \hat{y})$
             \Statex[2]     $\ell_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
@@
 remaining completely in the primal domain, making it applicable to very large
-datasets.  We will refer to the algorithm as SZLJSP, after the authors of the
+datasets.  We refer to the algorithm as SZLJSP, after the authors of the
 two papers.
 %
-The SZLJSP algorithm has properties that are very simliar to
-stochastic subgradient descent, but with two main advantages:
+The SZLJSP algorithm has the same theoretical convergence rate of $O(\frac{1}{T})$
+as stochastic subgradient decent, but has two distinguishing advantages:
 \begin{description}
@@
 \item[Learning Rate] While there are several theoretical results on choosing
-    the learning rate in stochastic subgradient descent, choosing a concrete
+    the step size in stochastic subgradient descent, choosing a concrete
     schedule is often problematic in practice. By using analytic line-search,
-    SZLJSP removes the need for any learning rate parameter, making application
+    SZLJSP removes the need for any step size parameter, making application
     of the algorithm much simpler.

### 2013-09-07 exact_learning/paper.tex
intermediate

@@
 \section{Introduction}
-% Remove duplication with pystruct intro
 As discussed previously, %TODO in the introduction? in pystruct chapter?
@@
 As loss-augmented prediction is a central step in the learning algorithms
-discussed in Chapter~\ref{ch:structured_pystruct}
-in the inference lead to only approximate learning.
+discussed in Chapter~\ref{ch:structured_pystruct}, approximate inference leads
+to complications in learning.
 
 In this chapter, we investigate the necessity and consequences of these
-approximations in the one-slack learning algorithms in the context of semantic
+approximations in the one-slack cutting-plane learning algorithm in the context of semantic
 image segmentation. We show that despite inference being deemed intractable in
@@
         methods for learning SSVMs using the cutting plane method, relating
-        approximate learning to the exact optimum.
+        approximate learning to the exact optimum. %FIXME really?
     \item We introduce an efficient caching scheme to accelerate cutting plane
@@
 approximate inference alone, certification of optimality allows treating
-learning as a black-box, allowing the researcher to focus attention on
+learning as a black-box, enabling the researcher to focus attention on
 designing the model for the application at hand.
@@
 
-Recently, there has been much research in learning structured prediction models
-where standard exact inference techniques are not applicable.
+Recently, there has been an increase in research in learning structured
+prediction models where standard exact inference techniques are not applicable,
+in particular in the computer vision community.
 The influence of approximate inference on structural support vector machine
@@
 Believe Propagation (LBP)~\citep{lucchi2011spatial}. These works use a single
-inference algorithm for during the whole learning process, and can not provide any
-bounds on the true objective or the empirical risk. In contrast, we combine
-different inference methods that are more appropriate for different stages of
-learning.
+inference algorithm for the whole learning process, and can not provide any
+bounds on the true objective or the empirical risk. In contrast, in this chapter
+we show how to combine different inference methods that are more appropriate
+for different stages of learning.
 
@@
 
-\section{Learnint Structure Support Vector Machines with Approximate Inference}
-%TODO ueberleitung zu approximate inference
-% wir machen one-slack und I ist jetzt das und das.
+\section{Learning Structure Support Vector Machines with Approximate Inference}
+In this chapter, we will use the 1-slack cutting plane algorithm, as we found this
+algorithm to be most suitable for out approach. For reference, the algorithm is outlined
+\Secref{one_slac}, Algorithm~\ref{alg_one_slack}. The inference algorithm $I$ is called
+in line~\ref{get_cutting_plane}. We investigate algorithms $I$ (often called separation
+oracles in the context) that will not yield the exact maximum here.
 There are two groups of inference procedures, as identified in
@@
 of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
-hand, will solve the loss-augmented inference in line~\ref{get_cutting_plane}
+hand, will solve the loss-augmented prediction in line~\ref{get_cutting_plane}
 exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
-that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.  We will elaborate on
-the properties of under-generating and over-generating inference procedures in
-Section~\ref{bounds}. %TODO Filler sentence? wat?
+that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.
+%We will elaborate on
+%the properties of under-generating and over-generating inference procedures in
+%Section~\ref{bounds}. %TODO Filler sentence? wat?
 
@@
 when using inference algorithm $I$, that is $o^I(\theta)$ is the approximation of the primal
-objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.
+objective given by $I$. To simplify exposure, we drop the dependency on $\theta$.
 
@@
 inference.
-We will now describe how we can further use 1. to both speed up and improve learning.
+We now describe how we can further use 1. to both speed up and improve learning.
 
@@
 
-%objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.
-In the following, we will use the notation of Section~\ref{bounds} and indicate
+%objective given by $I$. To simplify exposure, we drop the dependency on $\theta$.
+In the following, we use the notation of Section~\ref{bounds} and indicate
 the choices of oracle $I$ with $Q$ for a chosen inference algorithm and $C$ for
@@
 \subsection{Implementation Details}
-We implemented the cutting plane Algorithm~\ref{alg_cutting_plane} and our pairwise
-model in Python.
-Our cutting plane solver for Equation~\eqref{mainequation} uses
-\texttt{cvxopt}~\citep{dahl2006cvxopt} for solving the QP in the inner loop. Code for
-our structured prediction framework will be released under an open source
-license upon acceptance.
+We implemented the described procedure in \pystruct.
 
@@
 machine learning package~\citep{pedregosa2011scikit}.
+%TODO implementation details here?
 The move-making algorithm using QPBO is implemented with the help of the QPBO-I

### 2013-09-07 itm/paper.tex
intermediate

@@
     MeanNN scale to 70000 data points, as storing the whole pairwise distance
-    matrix seems critical.\label{results}}
+    matrix seems necessary.\label{results}}
 \end{table}

### 2013-09-07 structured_prediction/paper.tex
FIX runtime n-slack: only 1/eps^2 in terms of QP :-/

@@
 \citet{tsochantaridis2006large} showed a convergence rate of $O(\frac{1}{T^2})$
-with respect to calls to loss-augmented prediction. The recent work
+with respect to calls to the QP solver. In the worst-case, only a single new constraint
+could be found in one pass over the dataset, which leads to $O(\frac{1}{kT^2})$ in terms
+of calls to loss-augmented prediction. The recent work
 of~\citet{lacoste2012block}, however, suggests a rate of $O(\frac{1}{T})$,

### 2013-09-07 exact_learning/paper.tex
spelling, slight refinement of exact learning...

@@
 
-These two observations can be used in a number of ways.
-\citet{finley2008training} used 2. to show that learning with an
-over-generating approach minimizes an upper bound on the empirical risk.
-
-We can also use these observations to judge the suboptimality of a given parameter $\theta$,
-that is see how far the current objective is from the true optimum.
-Learning with any under-generating approach, we can use 1. to maintain a lower bound
-on the objective. At any point during learning, in particular if no more constraints
-can be found, we can then use 2., to also find an upper bound.
-This way, we can empirically bound the estimation error, using only approximate
-inference.
-We now describe how we can further use 1. to both speed up and improve learning.
+We can also use these observations to judge the suboptimality of a given
+parameter $\theta$, that is see how far the current objective is from the true
+optimum.  Learning with any under-generating approach, we can use 1. to
+maintain a lower bound on the objective. At any point during learning, in
+particular if no more constraints can be found, we can then use 2., to also
+find an upper bound.  This way, we can empirically bound the estimation error,
+using only approximate inference.  We now describe how we can further use 1. to
+both speed up and improve learning.
 
@@
 The cutting plane method described in Section~\ref{cutting_plane} relies only
-on some separation oracle $I$ that produces violated constraints.
+on some separation oracle $I$ that produces violated constraints when
+performing loss-augmented prediction.
 
@@
     \item Use a fast under-generating algorithm.
-    \item Use exact inference.
+    \item Use a strong but slow algorithm that can certify optimality.
 \end{enumerate*}
@@
 we obtain a certificate of optimality when learning stops.
+% TODO certify optimality != exact...
 
@@
 
-For exact inference, we use the recently developed Alternating Direction Dual
-Decomposition (AD$^3$) method of \citet{martins2011augmented}. AD$^3$ produces
-a solution to the linear programming relaxation, which we use as the basis for
-branch-and-bound.
+For inference with optimality certificate, we use the recently developed
+Alternating Direction Dual Decomposition (AD$^3$) method of
+\citet{martins2011augmented}. AD$^3$ produces a solution to the linear
+programming relaxation, which we use as the basis for branch-and-bound.
 
@@
 
-\subsection{Image Segmentation}
-We use superpixel-based semantic image segmentation as our application as this is our task of interest. %TODO formulation
-for this work.  CRF based models have a history of success in this application,
-with much current work investigating models and
-learning~\citep{gonfaus2010harmony, lucchi2013learning, ladicky2009associative,
-kohli2009robust, krahenbuhl2012efficient}.  We use the MSRC-21 and Pascal VOC 20120 datasets, two
-widely used benchmark in the field.
+\subsection{Semantic Image Segmentation}
+% TODO reference auf davor / intro?
+Our main application is of course semantic segmentation and object class
+segmentation. We evaluate the proposed learning approach on Pascal VOC 2013 and
+MSRC-21
 
@@
 Note that state-of-the-art approaches, some not build on the CRF framework, obtain
-around 40\% Jaccard score (also call VOC score), notably \cite{xia2012segmentation}, who 
+around a Jaccard Index (also call VOC score) of 40\% , notably \cite{xia2012segmentation}, who 
 evaluate on the Pascal VOC 2010 ``test'' set.
@@
 \subsubsection{Results}
-We compare classification results using different inference schemes in
-with results from the literature. As a sanity check, we
-also provide results without pairwise interactions.
+We compare classification results using different inference schemes with
+results from the literature. As a sanity check, we also provide results without
+pairwise interactions.
 
@@
 under-generating approaches can not be used to upper-bound the primal
-objective.
+objective or compute meaningful duality gaps.
 
@@
 \section{Summary}
-In this chapter we demonstrated that it is possible to learn state-of-the-art CRF models
-exactly using structural support vector machines, despite the model containing many loops.
-The key to efficient learning is the combination of different inference mechanisms and
-a novel caching scheme for the one-slack cutting plane method, in combination
-with state-of-the-art inference methods.
+In this chapter we demonstrated that it is possible to learn state-of-the-art
+conditional random field models exactly using structural support vector
+machines, despite the model containing many loops.  The key to efficient
+learning is the combination of different inference mechanisms and a novel
+caching scheme for the one-slack cutting plane method, in combination with
+state-of-the-art inference methods.
 
@@
 computer vision applications.
-
+%
 Even though exact learning does not necessarily lead to a large improvement in
@@
 optimization.
-
-While our model is trained on a relatively small dataset, we expect that scaling to
-datasets of several thousand images, such as Pascal VOC, is unproblematic.
-The time for a single iteration of the cutting plane solver will only grow
-linearly for the inference. The cost of solving the QP is actually independent
-of the number of samples, as only a single constraint will be added in each
-iteration.
 

### 2013-09-07 structured_prediction/paper.tex
spelling, slight refinement of exact learning...

@@
 This approach to structured prediction is closely related to 
-approachs using \emph{probabilistic graphical models}.  Probabilistic graphical models are a
+approaches using \emph{probabilistic graphical models}.  Probabilistic graphical models are a
 tool to express factorizations of probability distributions.  Similar to
@@
 Additionally, we discuss practical implications and implementation.  One
-particularly intereresting aspect is the difference between \emph{sequention}
+particularly interesting aspect is the difference between \emph{sequential}
 (or \emph{online}) and \emph{batch} algorithms.  Batch algorithms process the
@@
 run-times for batch algorithms. On the other hand, loss-augmented prediction in
-batch algorithms is embarassingly parallel, allowing the use of multiple
+batch algorithms is embarrassingly parallel, allowing the use of multiple
 processors with almost linear speed improvements.
@@
 
-Arguably the most straight-foward way to approach \Eqref{learning_equation} is
+Arguably the most straight-forward way to approach \Eqref{learning_equation} is
 using subgradient descent.  In light of the complexity of solving the
@@
 %
-Solving the quadratic programm can be accelerated using several techniques. We
-found that agressively removing constraints that are inactive or contribute
+Solving the quadratic program can be accelerated using several techniques. We
+found that aggressively removing constraints that are inactive or contribute
 little to the solution often makes the difference between the algorithm being
@@
         line~\ref{convergence_check}.  Only if $\xi' - \xi < \epsilon$, that is
-        the produced constraint is not violated stronly enough, we return to
+        the produced constraint is not violated strongly enough, we return to
         line~\ref{get_cutting_plane} and actually invoke the loss augmented

### 2013-09-07 datasets/paper.tex
start going on about stupid datasets

\chapter{Datasets and Evaluation Methodology}
This work uses several public datasets for semantic segmentation
and object class segmentation. This chapter briefly introduces each of the datasets,
and discusses its characteristic. The datasets we use are among the most well-known
and most-used datasets in the current semantic segmentation and object class segmentation
literature.

\section{Graz}
% really?

\section{Microsoft Research Cambridge Segmentation}
% intro + description
The Microsoft Research Cambridge dataset V2 (MSRC-23) was introduced in \citet{shotton2006textonboost}, and
has been widely used since. It is a dataset for semantic segmentation, and contains annotations for
several object categories: car, body, face, dog, boat, plane, bike, cow, cat, sheep,
building, bird, book, chair, horse, and sign. It also contains several ``stuff''
categories: sky, grass, tree, water, road, flower, and mountain.
This division is somewhat arbitrary, as some of the object classes also have a ``stuff'' character:
buildings are often not isolated, but rather whole street-scenes, and books are mostly whole
shelfs.
%
As the dataset contains only very few instances of mountain and horse, we do
not consider these classes, as is common in the literature. The resulting
dataset is known as MSRC-21.
The dataset contains %FIXME
images in total, usually split into a training set of size , a validation set of size ,
and a testing set of size .

% characteristics
The size of the dataset is quite small considering the variety of categories. In particular the
bird class consists mostly of ducks, swans and geese, but also contains two instances of peacocks
and one image showing tits. Few samples combined with large intra-class variety introduces considerable
variance in the performance. Using a fixed training and test set makes comparison with the literature possible,
but does not solve this problem.
%
On the other hand, many of the images obey very similar scene layouts, such as centered cars from the side,
centered planes on the run way, and centered faces. We will keep with the literature in not exploiting
this property of the dataset, as it can be considered an unwanted artifact of the dataset collection.
One striking characteristic of the MSRC dataset is that

Most of the classes in the MSRC-21 dataset can be well distinguished using local texture.
This is particularly true for the stuff classes, but also for the animal classes, buildings, and signs.
Consequently, many approaches use low-level approaches such as texton-boost or random forests as base features,
as these are particularly well-suited for the dataset.

\section{Pascal VOC}

\section{New York University RGB-D Segementation

\section{Evaluation}

### 2013-09-08 datasets/paper.tex
removed dataset blabla, maybe in appendix. added contained papers to the intro.

@@
-\chapter{Datasets and Evaluation Methodology}
+\chapter{Computer Vision Tools, Datasets and Evaluation}
 This work uses several public datasets for semantic segmentation
@@
 literature.
+
+\section{Computer Vision Tools}
+Textonboost
+SIFT, SLIC, CPMC, BOW
 
@@
 
+% Typical methods
 Most of the classes in the MSRC-21 dataset can be well distinguished using local texture.
@@
 as these are particularly well-suited for the dataset.
+%
+Following this line of work, we use the texton-boost implementation provided by \citet{krahenbuhl2012efficient}
+as a basis of all our experiments.
 
 \section{Pascal VOC}
+The Pascal VOC competition has
 
-\section{New York University RGB-D Segementation
+\section{New York University RGB-D Segmentation
 

### 2013-09-08 intro.tex
remove all occurences of i.e. and e.g. minor fixes
FIX put chapter commands in separate files, so that the text is not on a new page...
removed dataset blabla, maybe in appendix. added contained papers to the intro.

--- 2013-08-14 intro.tex
@@
-\section{Overview}
-What does this theses contribute?
-
-Why semantic segmentation?
-Why is it not solved yet?
+\chapter{Introduction}
 
@@
 Datsets? Input?
+
+Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
+Why is it not solved yet?
+
+Structured approaches, latent variable approaches
+
+\section{List of Contributions}
+Contributions:
+- Introduces general clustering algorithm that improves pipeline in serveral places
+- Shows that weakly supervised object class segmentation (not semantic segmentation) is possible
+- Provides general efficient framework for structured prediction
+- Analyze max-margin learning algorithms in different regimes
+- shows exact learning is possible in loopy graphs and improves results (sometimes?)
+- shows importance of edge features
+- Demonstrates how latent variables benefit semantic segmentation
+
+\section{Publications}
+The main material of this thesis has either been published in conference proceedings or has been
+submitted to conferences or journals. We now list the relevant publications.
+\begin{description}
+    \item[Chapter 2] \emph{Information Theoretic Clustering using Minimum Spanning Trees} Andreas C.
+M\"uller, Sebastian Nowozin and Christoph H.
+Lampert.
+Published in the proceedings of the German Conference on Pattern Recognition.
+    \item[Chapter 3] \emph{Multi-Instance Methods for Partially Supervised Image Segmentation} Andreas C.
+M\"uller and Sven Behnke.
+Published in the proceedings of the IARP Worshop on Partially Supervised Learning.
+    \item[Chapter 4] \emph{PyStruct - Structured Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to the Journal of Machine Learning Research, Open Source Software track.
+    \item[Chapter 5] \emph{Learning a Loopy Model for Semantic Segmentation Exactly}.
+Andreas C.
+M\"uller and Sven Behnke.
+Submitted to the International Conference on Computer Vision Theory and Applications.
+    \item[Chapter 6] \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}.
+Andreas C.
+M\"uller and Sven Behnke.
+Submitted to the International Conference on Robotics and Automation.
+\end{description}

### 2013-09-08 exact_learning/paper.tex
added figure to exact learning

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{exact_learning/images/inference_algs}
+\caption{%
+Illustration of the choice of inference algorithm. During the beginning of learning,
+fusion move inference together with caching is used. If no more constraint can be found,
+the fusion move algorithm is replaced by AD$^3$ with branch and bound.
+\figlabel{scheme_inference_algs}}
+\end{figure*}
+
 While using more different oracles is certainly possible, we found
@@
 combination allows us to make fast progress initially and guarantee optimality
-in the end.
+in the end. Our strategy is visualized in \figref{scheme_inference_algs}
 Notably, it is not necessary for an algorithm used as 3) to always produce
@@
 we solve the QP once with a constraint from the cache. If the resulting $o^C$ verifies
-\begin{equation}\label{eq:cache_test}
+\begin{equation}\eqlabel{cache_test}
     o^C - o^Q < \frac{1}{2} (o^Q - o_{\W})

### 2013-09-08 pystruct/paper.tex
convergence rates, citations, mehr todos fertig...
fixed / improved table of implemented methods

@@
 \centering
-\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
+\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lccccccc}
 \toprule
-Package       & Cutting Plane& SSGD & SZLJSP & Latent SSVM & Perceptron & ML\\
-\cmidrule(r){1-1} \cmidrule(r){2-7}
-\pystruct     & \x        & \x   & \x   & \x          & \x         & \o\\
-\svmstruct    & \x        & \o   & \o   & \x          & \o         & \o\\
-\sc{Dlib}     & \x        & \x   & \o   & \o          & \x         & \o\\
-\sc{CRFsuite} & \o        & \o   & \o   & \o          & \x         & \x\\
+Package       & 1-SP & $n$-SP & SSGD & SZLJSP & L-SSVM & Perceptron & ML\\
+\cmidrule(r){1-1} \cmidrule(r){2-8}
+\pystruct     & \x   & \x     & \x   & \x   & \x          & \x         & \o\\
+\svmstruct    & \x   & \x     & \o   & \o   & \x          & \o         & \o\\
+\sc{Dlib}     & \x   & \o     & \o   & \o   & \o          & \x         & \o\\
+\sc{CRFsuite} & \o   & \o     & \o   & \o   & \o          & \x         & \x\\
 
@@
         algorithms implemented in popular structured prediction software
-        packages. SSGD stochastic subgradient decent learning of SSVMs, SZLJSP is
-        as described in \secref{dual_coordinate_descent}, Latent SSVM for latent variable SSVMs, ML for
+        packages. 1-CP stands for 1-slack Cutting Plane, $n$-CP for $n$-slack
+        Cutting plane, SSGD for stochastic subgradient decent learning of
+        SSVMs, SZLJSP is as described in \secref{dual_coordinate_descent},
+        L-SVM stands for latent variable SSVMs, and ML for
         maximum likelihood learning.}
@@
 We also introduced \pystruct, our modular structured learning and prediction library in Python.
-\pystruct is geared towards ease of use, while providing efficient implementations and will be the
+\pystruct is geared towards ease of use, while providing efficient implementations and is be the
 basis of our further experiments.

### 2013-09-08 main.tex
splitting pystruct file
polish SZLJSP algorithm listing
convergence rates, citations, mehr todos fertig...
intermediate
start going on about stupid datasets
removed dataset blabla, maybe in appendix. added contained papers to the intro.
zusammenfassung/abstract.

@@
 \usepackage{multicol}
-\usepackage{pslatex}
+%\usepackage{pslatex}
 \usepackage{microtype}
@@
 \usepackage{minted}
-\usepackage[urw-garamond]{mathdesign}
-\usepackage{garamondx}
+
+%\usepackage{tgpagella}
+%\usepackage[T1]{fontenc}
+
+%\usepackage{garamondx}
+%\usepackage[urw-garamond]{mathdesign}
 \usepackage{xspace}
@@
 
+\makeatletter
+\let\OldStatex\Statex
+\renewcommand{\Statex}[1][3]{%
+\setlength\@tempdima{\algorithmicindent}%
+\OldStatex\hskip\dimexpr#1\@tempdima\relax}
+\makeatother
+
 \definecolor{rulecolor}{rgb}{0.80,0.80,0.80}
@@
 
-\newtheorem{corollary}{Corollary}
-\newtheorem{corollarya}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
-\newtheorem{theorema}{Theorem}
+%\newtheorem{corollary}{Corollary}
+%\newtheorem{corollarya}{Corollary}
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{theorem}{Theorem}
+%\newtheorem{theorema}{Theorem}
 
@@
 \chapter*{Zusammenfassung}
+Automatische Segementierung und Erkennung von semantischen Klassen in
+natuerlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
+Einer der am weitest verbreiteten Ans\"atze zu dieser Aufgabe basiert auf
+Conditional Random Fields und Verfahren der strukturierten Vorhersage. Die
+Vorhersage f\"ur Modelle zur Bild-Markierung mittels Conditional Random Fields
+ist f\"ur gew\"ohnlich nur approiximativ moeglich, bedingt d\"urch das
+Vorhandensein von Kreisen in den zugrunde liegenden Nachbarschaftsgraphen.
+Aus diesem Grund, und trotz der Beliebtheit von Conditional Random Field
+Verfahren, ist das strukturierte Lernen f\"ur solche Anwendungen schlecht
+verstanden.
+%
+Wir untersuchen verschiedene Lernalgorithmen in diesem
+Zusammenhang. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
+Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zu lernen.
+Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
+weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
+durchzuf\"uhren.
+Unsere Studie erlaubt uns eine Charakterisation der betrachteten Algorithmen
+in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
+Objektklassensegmentierung.
+
 \chapter*{Abstract}
+Automatic segmentation and recognition of semantic classes in natural images is an important open problem in computer vision.
+One of the most popular approaches to this task is based on conditional random fields and structured prediction.
+For image labeling tasks, prediction using conditional random fields is usually only approximate, because of the presence
+of loops of the neighborhood graph in the the image. Consequently, learning in
+this setting is still little understood, despite the popularity of these
+methods.
+%
+We investigate several learning algorithms in this setting. We show
+how structured models can be learned exactly in practical settings, even in the
+presence of many loops in the underlying neighborhood graphs.  We introduce a
+software library that allows us to perform extensive empirical comparisons of
+state-of-the-art structured learning approaches. This allows us to characterize
+their practical properties in a range of applications, in particular for semantic segmentation
+and object class segmentation.
 
@@
 
+%\include{datasets/paper}
+
 \include{itm/paper}
@@
 
+\include{structured_prediction/paper}
 \include{pystruct/paper}

### 2013-09-15 nyu/legend.tex
add icra nyu paper

\documentclass{standalone}
\usepackage{color}
\usepackage{tikz}

\definecolor{structure}{RGB}{185, 202, 202}
\definecolor{prop}{RGB}{227, 172, 0}
\definecolor{furniture}{RGB}{255, 94, 28}
\definecolor{ground}{RGB}{9, 68, 83}
%[``structure'', ``prop'', ``furniture'', ``ground'' , ``void'']

\begin{document}
    \begin{tikzpicture}
        \matrix [column sep=.5cm]{%
            \node [rectangle, fill=ground, minimum width=.5cm,minimum height=.5cm] {};&
        \node [rectangle] {Ground};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=structure, minimum width=.5cm,minimum height=.5cm] {};&
        \node [rectangle] {Structure};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=furniture, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Furniture};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=prop, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Props};&
        \node [rectangle, minimum width=.3cm] {};&
        \node [rectangle, fill=black, minimum width=.5cm,minimum height=.5cm] {};&
        \node [draw=none] {Void};\\
    };
    \end{tikzpicture}
\end{document}

### 2013-09-15 nyu/paper.tex
add icra nyu paper

%\begin{abstract}
%    We present a structured learning approach to semantic annotation of RGB-D
%    images. 
%    Our method learns to reason about spatial relations of objects and
%    fuses low-level class predictions to a consistent interpretation of a
%    scene. Our model incorporates color, depth and 3D scene features, on which
%    an energy function is learned to directly optimize object class prediction using
%    the loss-based maximum-margin principle of structural support vector machines.
%    %With the proliferation of cheap % TODO urghs formulation
%    %high-quality RGB-D sensors, it became possible to capture more and more
%    %data not only as color images, but with additional depth information. The
%    %depth channel provides important clues not only for scene structure, but
%    %also for predicting semantic object classes.
%    %We demonstrate how CRF learning, an approach
%    %that has shown great promise for RGB images, can exploit depth information
%    %to improve semantic interpretation of the scene.
%    We evaluate our approach on the NYU V2 dataset of indoor scenes, a challenging
%    dataset covering a wide variety of scene layouts and object classes. We
%    hard-code much less information about the scene layout into our model then previous
%    approaches, and instead learn object relations directly from the data. We
%    find that our conditional random field approach improves upon previous
%    work, setting a new state-of-the-art for the dataset.
%\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Learning Depth-Sensitive Conditional Random Fields for Semantic
Segmentation of RGB-D Images}

\section{Introduction}
For robots to perform varied tasks in unstructured environments, understanding their
surroundings is essential. We formulate the problem of semantic annotation of maps
as a dense labeling of RGB-D images into semantic classes. Dense labeling of objects
and structure classes allows for a detailed reasoning about the scene.

In this work, we propose the use of random forests combined with conditional
random fields (CRFs) to perform robust estimation of structure classes in RGB-D
images. The CRF is learned using a structural support vector machine,
allowing it to integrate the noisy low-level information produced by a pixel-based
random forest to a consistent interpretation of the scene.

We thereby extend the success of learned CRF models for
semantic segmentation in RGB images to the domain of 3D scenes.
Our emphasis lies on exploiting the additional depth and 3D information in all
processing steps, while relying on \emph{learning} to create a model that 
is adjusted to the properties of the sensor input and environment.

Our approach starts with a random forest, providing a noisy local estimate
of semantic classes based on color and depth information. These estimates
are grouped together using a superpixel approach, for which we extend previous
superpixel algorithms from the RGB to the RGB-D domain.
We then build a geometric model of the scene, based on the neighborhood graph
of superpixels.  We use this graph not only to capture spatial relations in the
2D plane of the image, but also to model object distances and surface angles in
3D, using a point cloud generated from the RGB-D image. The process is depicted
in \Figref{teaser}.

We assess the accuracy of our model on the challenging NYU Segmentation V2
dataset \citep{SilbermanECCV12}, where our model outperforms previous
approaches.  Our analysis shows that while our random forest model
already has competitive performance, the superpixel-based grouping and in
particular the loss-based learning are integral ingredients of the success of our method.

\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{nyu/images/teaser}
    \end{center}
    \caption{%
        Overview of the proposed semantic segmentation method.\label{teaser}
    \figlabel{height}}
\end{figure}

\section{Related work}
The task of dense semantic annotation of 3D maps has seen an increased interest
in recent years.  Early work includes \citet{nuchter2008towards}, who combine
6D SLAM, surface annotation, and object recognition to build semantically
annotated maps. They demonstrate their approach on a mobile robot in an indoor
environment.
%
More recently \citet{sengupta2012automatic} introduced a dataset of
semantically annotated street-scenes in a closed track, captured as pairs of
stereo images. They approach the task by jointly reasoning about 3D layout and
semantics of the scenes and produce a dense labeling on image level.  In
\citet{senguptaurban}, they extend their approach to produce a volumetric
reconstruction of the scene, together with a dense semantic labeling of the
volumetric representation. Their approach to image segmentation builds on
the hierarchical CRF approach of
\citet{ladicky2009associative}, which is similar in spirit to our approach,
but uses Potts potentials together with cross-validation to set potentials.

Recent approaches for indoor semantic annotation of maps mostly focused on
RGB-D images, which are now easy to obtain using structured light sensors.
\citet{stuckler2012semantic}, for example, used a Random Forest model to obtain a dense
semantic labeling of images and integrated predictions over multiple views in 3D.
They evaluated their approach on table-top and simple indoors scenes.
%
\citet{silberman2011indoor} introduced the NYU Depth Dataset V1 indoor dataset,
which consisted of a large variety of densely annotated indoor scenes, captured as RGB-D images.
Their work also introduced a baseline method for semantic segmentation of RGB-D
image, which is based on a CRF over superpixels, with unary potentials given by interest point descriptors.
While pairwise potentials for the CRF were carefully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our work, which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
%
\citet{ren2012rgb} evaluated the design of features for semantic labeling of
RGB-D data, and use a hierarchical segmentation to provide context. While they
also define a CRF on superpixels, their model is again not learned, but a
weighted Potts model, using only a probability of boundary map, and not taking
spatial layout into account at all.
%
\citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using in this work.  Their focus is on inferring support relations in indoor
scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} use a complex pipeline, employing significant domain knowledge.
In this work, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

\citet{couprie-iclr-13} approach the task of semantic segmentation of structure classes
in RGB-D using the paradigm of convolutional neural networks, extending
previous work of \citet{farabet-pami-13} and \citet{schulz2012learning}.
Similar to our approach, \citet{couprie-iclr-13} combine the output of a
pixel-based, low-level learning algorithm with an independent unsupervised
segmentation step. In contrast to their work, we improve our results by not
only averaging predictions within superpixels, but also explicitly learning
interactions between neighboring superpixels, favoring a consistent
interpretation of the whole image.

\citet{stueckler2013} extends the approach of \citet{stuckler2012semantic} to
a real-time system for online learning and prediction of semantic classes. Their
use a GPU implementation of random forests, and integrate 3D scene information
in an online fashion. They evaluate their approach on the dataset of
\citet{SilbermanECCV12} with promising results.
We use the implementation of random forest provided by
\citet{stueckler2013}, but instead of integrating predictions over time, we
focus on exploiting the structure within a single frame.

While many of the works mentioned in this chapter make use of a CRF
approach, we are not aware of any prior work on semantic annotation of 3D maps
that fully learns their potentials.
%The task of dense semantic labeling in natural images has been widely studied
%in computer vision.  Our approach of using pixel-based low-level cues is
%similar to the work of \citet{shotton2006textonboost}, who introduced
%TextonBoost for semantic segmentation, and used a hand-coded CRF energy to
%model spatial interactions. More recent work usually employ learned conditional
%random fields, learning object interactions from the
%data to directly maximize a loss function~\citep{lucchi2013learning, krahenbuhl2012efficient},
%which is the approach we are following.

\section{Learning depth-sensitve conditional random fields}\seclabel{learning}
We take a CRF approach, whose nodes represent a labeling of
superpixels.  We use an energy consisting of first and second order factors
(also called unary and pairwise potentials), with learned potential functions.
Let us denote the representation of an input image by $x$ and a labeling of
superpixels into semantic classes by $y$.  Then the general form of the energy is
\begin{equation}\eqlabel{energy}
    g(x, y) = \sum_{v \in V} \psi_v(x, y_v) + \sum_{(v, w) \in E} \psi_{v,w}(x, y_{v}, y_{w}).
\end{equation}
Here $V$ enumerates the superpixels, and $E\subset V \times V$ is a set of
edges, encoding adjacence between superpixels.

We learn the unary and pairwise energy functions $\psi_v$ and $\psi_{v, w}$ from the training data
using a structural support vector machine (SSVM)~\citep{joachims2009cutting}. 
The concept of SSVMs allows for a principled,
maximum-margin based, loss-sensitive training of CRFs. Learning the potential yields much
more complex interactions than the simple Potts potentials that are often used in the literature.

In general, structural SSVMs learn the parameters $\theta$ of a predictor of the form
\begin{equation}\eqlabel{prediction}
f(x) = \argmax_{y \in \mathcal{Y}} \theta^T \Phi(x, y).
\end{equation}
We choose $\psi$ in \Eqref{energy} to be linear in the learnable parameters and
the data-depended features, resulting in a form equivalent to
\Eqref{prediction}. Our features are described in detail below.  We use the
$1$-slack formulation of the structural SVM~\citep{joachims2009cutting} and
solve the maximization in \Eqref{prediction} using a combination of fusion
moves~\citep{lempitsky2010fusion} and the AD$^3$ algorithm of
\citet{martins2011augmented}.  In contrast to graph-cut inference, fusion moves
and AD$^3$ can work with arbitrary potential function, and allow precise
learning using the SSVM approach.

\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{nyu/images/height_success}\\
        \vspace{3mm}
        \includegraphics[width=\linewidth]{nyu/images/height_failure}
    \end{center}
    \caption{%
        Visualization of the height computed using the method described in \Secref{unaries}.
        Input images are shown on the left (depth not shown), the computed height is depicted on the right.
        The top row exemplifies a typical scene, while the bottom row shows a scene without horizontal
        surfaces, where our method fails.
    \figlabel{height}}
\end{figure}

\subsection{Low Level Segmentation}
We take a super-pixel based approach to semantic segmentation.
Our superpixel generation is based on the SLIC algorithm~\citep{achanta2012slic}. We extend the
standard SLIC algorithm, which works on the Lab space, to also include 
depth information. The resulting algorithm is a localized $k$-Means in Lab-D-XY space.
Our implementation is publicly available through the scikit-image library\footnote{\url{http://scikit-image.org}}.
Similar to \citet{SilbermanECCV12}, we found little visual improvement over the
RGB segmentation when using additional depth information. On the other hand,
estimation of per-superpixel features based on the 3D point cloud was more
robust when including depth information into the superpixel procedure.
%TODO visualize?
The resulting superpixels are compact in the 2D image. As the density of the
corresponding point cloud is dependent on the depth, we did not succeed in
creating superpixels that are compact in 3D while maintaining a meaningful
minimum size.

\subsection{Unary Image Features}\seclabel{unaries}
Our method builds on the probability output of a random forest, trained for pixel-wise
classification of the structure classes.
We use the GPU implementation provided by \citet{stueckler2013}\footnote{\url{https://github.com/deeplearningais/curfil}}.
The input for training are the full RGB-D images, transformed to Lab
color space. Each tree in the forest uses training pixels only from a subset of
training images. For each training image, an equal number of pixels for each
occurring class is sampled. Split features are given by difference of regions
on color or depth channels. Region size and offsets are normalized using the
depth at the target pixel.
We accumulate the probabilistic output for all pixels within a superpixel, and
use the resulting distribution as a feature for the unary node potentials in our CRF model.
We augment these prediction with another feature, based on the height of a superpixel in 3D.
This is a very informative feature, in particular to determine the floor.
To compute the height of a (super) pixel, we first find the ``up'' direction.
We use a very simple approach that we found effective: we cluster normal
directions of all pixels into 10 clusters using $k$-means, and use the one that
is most parallel to the $Y$ direction, which roughly corresponds to height in
the dataset.
We then project the 3D point cloud given by the depth along this direction, and
normalize the result between $0$ and $1$.
This procedure works robustly given there is some horizontal surface in the image,
such as the ground or a table. A few scenes contain only walls and furniture,
and the approach fails for these. \Figref{height} illustrates a typical case and
one of the much rarer failure cases.
While we could use a more elaborate scheme,
such as the one from \citet{SilbermanECCV12}, we suspect that the feature is of
little use in scenes without horizontal surfaces.
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{nyu/images/normal_feature}
    \end{center}
    \vspace{-3mm}
    \caption{%
        Visualization of one of the pairwise features, the similarity between superpixel normals.
        The image shows the zoom-in of a bedroom scene, together with the superpixel over-segmentation.
        Lines connect adjacent superpixels, and line-strength gives the magnitude of the orientation similarity.
    \figlabel{normal_feature}}
    \vspace{-9mm}
\end{figure}

\subsection{Pairwise Depth-Sensitive Features}
There are five different features used to build pairwise potentials in our model:

\emph{Constant}
    A constant feature allows to model general neighborhood relations.

\emph{Color Contrast}
    We employ a non-linear color contrast, as is common in the computer
    vision literature, between the superpixel mean colors $c_i$ and $c_j$:
    $\exp\left(-\gamma \lVert c_i - c_j \rVert^2\right)$.

\emph{Vertical Alignment}
    We model the directed angle between superpixel centers in the 2D image
    plane.  This allows the  model to learn that ``structure'' is above
    ``floor'', but not the other way around.

\emph{Depth Difference}
    We include the signed depth difference between superpixels, which
    allows the model to detect depth discontinuities that are not
    represented in the 2D neighborhood graph of the superpixels.

\emph{Normal Orientations}
    Differences in normal vector orientation are a strong clue on
    whether two superpixels belong to the same surface, and therefore the
    same structural class.
    We compute the 3D orientation of normals using the method of \citet{holz_2011_robocup},
    as implemented in the point cloud library (pcl)\footnote{\url{http://pointclouds.org}}.
    All normals within a superpixel are then averaged, to get a single orientation for each superpixel.
    The feature is computed as the difference of $\frac{\pi}{4}$ and the (undirected) angle between the normals belonging
    to two adjacent superpixels.  
    The feature is illustrated in \Figref{normal_feature}. The change
    in normal orientation highlights that pillow and wall are distinct
    objects, even though there is no strong distinction in color or depth.

\section{Experiments}\seclabel{experiments}

\begin{table*}[t]
    \caption{Quantitative comparison of the proposed method with the
    literature.\tablabel{results}}
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
\toprule
                        & ground        &  structure    & furniture     & props         & class average   & pixel average\\
\cmidrule(r){1-7}

RF                              &         90.8  &   81.6        & 67.9          & 19.9          &  65.0        &  68.3 \\
RF + SP                         &         92.5  &   83.3        & \textbf{73.8} & 13.9          &  65.7        &  70.1 \\ 
RF + SP + SVM                   &         94.4  &   79.1        & 64.2          & \textbf{44.0} &  70.4        &  70.3 \\
RF + SP + CRF                   & \textbf{94.9} &   78.9        &          71.1 & 42.7          &\textbf{71.9} &  \textbf{72.3} \\
\cmidrule(r){1-7}
\citet{SilbermanECCV12}         &         68    &   59          & 70           & 42            &  59.6        & 58.6 \\
\citet{couprie-iclr-13}         &         87.3  & \textbf{86.1} & 45.3         & 35.5          &  63.5        & 64.5 \\
\cmidrule(r){1-7}
\citet{stueckler2013}$^\dagger$ & \textbf{95.6} &   83.0        & \textbf{75.1}& 14.2          &  67.0        & 70.9 \\

\bottomrule
\end{tabularx}
\quad\\The best value in each column is printed in bold$^\dagger$. The upper part of
the table shows contributions by different parts of our pipeline. RF stands for random forest prediction, RF + SP for aggregated
random forests prediction within superpixels, RF + SP + SVM for an SVM trained on the unary potentials, and RF + SP + CRF is
our proposed pipeline. We optimized our approach for class average
accuracy.\\
$^\dagger$ \footnotesize Note that the work of \citet{stueckler2013} is not directly
comparable, as they integrated information over multiple frames, and did not
measure accuracy for pixels without valid depth measurement.
\vspace{-5mm}
\end{table*}

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 indoor
RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), which puts more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor.

Our approach is implemented using our \textsc{PyStruct}
library\footnote{\url{http://pystruct.github.io}}.  The code for our
experiments will be released upon acceptance. We use \textsc{scikit-learn}~\citep{pedregosa2011scikit}
for $k$-means clustering and the SVM baseline.

All hyper parameters were adjusted using 5-fold cross-validation. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework of \cite{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using grid-search.
We found 500 superpixels per image to work best, which allow for a maximum
possible performance of 95\% average class accuracy on the validation set.

\subsection{Results}\seclabel{results}
\Tabref{results} compares different components of our approach with the literature.
Please note that we \emph{first} designed our final model, using only the
validation data. We now report accuracies of simpler models for reference,
however these results were not used for model selection. To separate
the influence of loss-based training and the spatial reasoning of the CRF,
we also train a usual support vector machine (SVM) on the unary potentials for comparison.

The random forest prediction, as reported in \citet{stueckler2013} is already quite competitive.
Grouping into superpixels slightly improves performance, by removing
high-frequency noise and snapping to object boundaries. Somewhat surprisingly,
using a standard unstructured SVM with rescaled loss already advances the mean
accuracy decidedly above the previous state-of-the art. We
attribute this mostly to the ability of the SVM to exploit correlation between
classes and uncertainty within the superpixels.
Additionally, the SVM has access to the ``height'' feature, that was not
included in the random forest.
This performance is still improved upon, both in class average and pixel
average performance by the learned CRF approach, yielding
the best published result so far for both measures. The increase over the
standard SVM is $1.5\%$ for class average accuracy and $2.0\%$ for pixel
average accuracy.

A visualization of the impact of each processing step can be found in
\Figref{comparison}, which shows prediction results on the test set.  The four
prediction methods correspond to the rows of \Tabref{results}.  The difference
between the SVM and CRF approaches are clearly visible, with the CRF producing
results that are very close to the ground truth in several complex scenes.

We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the CRF can.
On the other hand, performance of the CRF deteriorates slightly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the CRF has little chance to recover.

\begin{figure}
    \begin{center}
        \includegraphics[width=.45\linewidth]{nyu/images/vertical_alignment}
    \includegraphics[width=.45\linewidth]{nyu/images/depth_difference}
    \end{center}
    \vspace{-3mm}
    \caption{%
        Visualization of some of the learned potentials. The right potential is
        applied to relative depth between superpixels, the second on the
        feature encoding whether one superpixel is above the other in the
        image. See \secref{results} for details.
    \figlabel{weights}}
    \vspace{-3mm}
\end{figure}

\Figref{weights} visualizes some learned potential functions.
Higher values correspond to favored configurations. One can see that the
vertical alignment potential expresses that the floor is much more likely to be below
other classes. It also encodes the fact that props
rest on furniture, but not the other way around.
The potential of the depth feature encodes, for example, that the ground
is usually behind the other classes, while furniture is in front of structures,
such as the wall.
Interestingly, the potential functions are not anti-symmetric, and forcing them to be so
degrades the results. This suggests that the direction of connecting edges, going from
the top left to the bottom right in the image, is also exploited by the potentials.

\begin{figure*}
    \begin{tabu} to \linewidth{@{}XXXXX@{}}
    \includegraphics[width=\linewidth]{nyu/images/00845_image.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00781_image.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01331_image.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00118_image.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01203_image.png}\\
    \multicolumn{5}{@{}c@{}}{Input Image}\vspace{3mm}\\

    \includegraphics[width=\linewidth]{nyu/images/00845_pixel.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00781_pixel.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01331_pixel.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00118_pixel.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01203_pixel.png}\\
    \multicolumn{5}{@{}c@{}}{Random Forest Prediction}\vspace{3mm}\\

    \includegraphics[width=\linewidth]{nyu/images/00845_sp.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00781_sp.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01331_sp.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00118_sp.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01203_sp.png}\\
    \multicolumn{5}{@{}c@{}}{Superpixel Voting}\vspace{3mm}\\

    \includegraphics[width=\linewidth]{nyu/images/00845_svm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00781_svm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01331_svm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00118_svm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01203_svm.png}\\
    \multicolumn{5}{@{}c@{}}{Support Vector Machine on Superpixels}\vspace{3mm}\\

    \includegraphics[width=\linewidth]{nyu/images/00845_ssvm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00781_ssvm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01331_ssvm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00118_ssvm.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01203_ssvm.png}\\
    \multicolumn{5}{@{}c@{}}{Conditional Random Field on Superpixels}\vspace{3mm}\\

    \includegraphics[width=\linewidth]{nyu/images/00845_gt.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00781_gt.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01331_gt.png}&%
    \includegraphics[width=\linewidth]{nyu/images/00118_gt.png}&%
    \includegraphics[width=\linewidth]{nyu/images/01203_gt.png}\\
    \multicolumn{5}{@{}c@{}}{Ground Truth}\vspace{3mm}\\
    \multicolumn{5}{@{}c@{}}{%
        \includegraphics[]{legend.pdf}
    }
    \end{tabu}
\caption{%
Qualitative evaluation of the CRF\@.
The first three images illustrate errors in the original prediction that can be corrected, while
the second two images illustrate failure modes. Pixels marked as void
are excluded from the evaluation. See the \secref{results} for details.
\figlabel{comparison}}
\end{figure*}

\section{Summary and discussion}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the art on the NYU V2
dataset for semantic annotation of structure classes.

While our approach allows modeling of spatial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models or higher
order potentials~\citep{ladicky2009associative}.
Another possible line of future work is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model,
as done by \citet{SilbermanECCV12}. Finally, we could also combine our
single-frame approach with the approach of \citet{stueckler2013}, which fuses
individual views in 3D to exploit temporal coherence.

While this work did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citep{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.

### 2013-09-15 nyu/paper.tex
minor fixes to nyu chapter

@@
 
-\chapter{Learning Depth-Sensitive Conditional Random Fields for Semantic
-Segmentation of RGB-D Images}
+\chapter{Learning Depth-Sensitive Conditional Random Fields}
 
-\section{Introduction}
+%\section{Introduction}
 For robots to perform varied tasks in unstructured environments, understanding their
@@
     \caption{%
-        Overview of the proposed semantic segmentation method.\label{teaser}
+        Overview of the proposed semantic segmentation method.\figlabel{teaser}
     \figlabel{height}}
@@
     \end{center}
-    \vspace{-3mm}
     \caption{%
@@
     \figlabel{normal_feature}}
-    \vspace{-9mm}
 \end{figure}
@@
     vision literature, between the superpixel mean colors $c_i$ and $c_j$:
-    $\exp\left(-\gamma \lVert c_i - c_j \rVert^2\right)$.
+    \[
+        \exp\left(-\gamma \lVert c_i - c_j \rVert^2\right).
+     \]
 
@@
     literature.\tablabel{results}}
-\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
+\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}l@{}cccccc}
 \toprule
-                        & ground        &  structure    & furniture     & props         & class average   & pixel average\\
-\cmidrule(r){1-7}
+                        & \footnotesize{ground}        &  \footnotesize{structure}    & \footnotesize{furniture}     & \footnotesize{props}         & \footnotesize{class avg.}   & \footnotesize{pixel avg.}\\
+\cmidrule{1-7}
 
@@
 RF + SP + CRF                   & \textbf{94.9} &   78.9        &          71.1 & 42.7          &\textbf{71.9} &  \textbf{72.3} \\
-\cmidrule(r){1-7}
+\cmidrule{1-7}
 \citet{SilbermanECCV12}         &         68    &   59          & 70           & 42            &  59.6        & 58.6 \\
 \citet{couprie-iclr-13}         &         87.3  & \textbf{86.1} & 45.3         & 35.5          &  63.5        & 64.5 \\
-\cmidrule(r){1-7}
+\cmidrule{1-7}
 \citet{stueckler2013}$^\dagger$ & \textbf{95.6} &   83.0        & \textbf{75.1}& 14.2          &  67.0        & 70.9 \\
@@
 measure accuracy for pixels without valid depth measurement.
-\vspace{-5mm}
 \end{table*}
@@
     \figlabel{weights}}
-    \vspace{-3mm}
 \end{figure}
@@
     \multicolumn{5}{@{}c@{}}{%
-        \includegraphics[]{legend.pdf}
+        \includegraphics[]{nyu/legend.pdf}
     }

### 2013-09-15 exact_learning/paper.tex
slight fixes to references.

@@
-\chapter{Efficient Exact Learning of Loopy CRFs}
+\chapter{Efficient Exact Learning of Loopy CRFs}\label{ch:exact_learning}
 
@@
 
-\subsection{Bounding the Objective}\label{bounds}
+\subsection{Bounding the Objective}\seclabel{bounds}
 Even using approximate inference procedures, several statements
-about the original exact objective (Equation~\eqref{mainequation}) can be
+about the original exact objective (\eqref{mainequation}) can be
 obtained.
@@
 
-\section{Efficient Cutting Plane Training of SSVMs}\label{learning}
+\section{Efficient Exact Cutting Plane Training of SSVMs}\seclabel{learning}
 
-\subsection{Combining Inference Procedures}\label{combining}
-The cutting plane method described in Section~\ref{cutting_plane} relies only
+\subsection{Combining Inference Procedures}\seclabel{combining}
+The cutting plane method described in \Secref{cutting_plane} relies only
 on some separation oracle $I$ that produces violated constraints when
@@
 
-\subsection{Dynamic Constraint Selection}\label{caching}
-Combining inference algorithm as described in Section~\ref{combining}
+\subsection{Dynamic Constraint Selection}\seclabel{caching}
+Combining inference algorithm as described in \Secref{combining}
 accelerates calls to the separation oracle by using faster, less accurate
@@
 %objective given by $I$. To simplify exposure, we drop the dependency on $\theta$.
-In the following, we use the notation of Section~\ref{bounds} and indicate
+In the following, we use the notation of \Secref{bounds} and indicate
 the choices of oracle $I$ with $Q$ for a chosen inference algorithm and $C$ for
@@
     our exact and approximately learned models and state-of-the-art
-    approaches.\label{msrcacc}}
+    approaches.\tablabel{msrcacc}}
 
@@
 \subsection{Caching}
-First, we compare our caching scheme, as described in Section~\ref{combining}, with the
+First, we compare our caching scheme, as described in \Secref{combining}, with the
 scheme of \citet{joachims2009cutting}, which produces constrains from the cache
@@
 
-Figure~\ref{caching} shows a visual comparison that highlights the differences
+\Figref{caching} shows a visual comparison that highlights the differences
 between the methods. Note that neither $o^Q$ nor $o^C$ provide valid upper bounds on the objective,
@@
 Large dots correspond to $o^Q$, small dots correspond to $o^C$,
-and the line shows $o_\W$. See the text for details.\label{caching}}
+and the line shows $o_\W$. See the text for details.\figlabel{caching}}
 \end{figure*}
@@
     \caption{Accuracies on the Pascal VOC Dataset. We compare our approach
-    against approaches using the same unary potentials.\label{pascalacc}}
+    against approaches using the same unary potentials.\tablabel{pascalacc}}
 
@@
     \caption{Objective function values on the MSRC-21 Dataset}
-    \label{msrc_objective}
+    \tablabel{msrc_objective}
 \end{table}
@@
 
-Results on the MSRC-21 dataset are shown in Table~\ref{msrcacc}.
+Results on the MSRC-21 dataset are shown in \Tabref{msrcacc}.
 We find that our model is on par with state-of-the-art approaches, implying
@@
 
-Results on the Pascal-VOC dataset are shown in Table~\ref{pascalacc}.
+Results on the Pascal-VOC dataset are shown in \Tabref{pascalacc}.
 We compare against several approaches using the same unary potentials.
@@
 up to $\epsilon=10^{-4}$.
-A visual comparison of selected cases is shown in Figure~\ref{visual}.
+A visual comparison of selected cases is shown in \Figref{visual}.
 
 The objective function values using only the under-generating move-making and
-the exact inference are detailed in Table~\ref{msrc_objective} and Table~\ref{pascal_objective}.
+the exact inference are detailed in \Tabref{msrc_objective} and \Tabref{pascal_objective}.
 We see that a significant gap between the cutting plane objective and the primal objective
@@
     \caption{Objective function values on the Pascal Dataset.}
-    \label{pascal_objective}
+    \tablabel{pascal_objective}
 \end{table}
@@
     ground truth.
-\label{visual}}
+\figlabel{visual}}
 \end{figure*}

### 2013-09-15 nyu/paper.tex
slight fixes to references.

@@
 \Eqref{prediction}. Our features are described in detail below.  We use the
-$1$-slack formulation of the structural SVM~\citep{joachims2009cutting} and
-solve the maximization in \Eqref{prediction} using a combination of fusion
-moves~\citep{lempitsky2010fusion} and the AD$^3$ algorithm of
-\citet{martins2011augmented}.  In contrast to graph-cut inference, fusion moves
-and AD$^3$ can work with arbitrary potential function, and allow precise
-learning using the SSVM approach.
+$1$-slack formulation of the structural SVM~\citep{joachims2009cutting} as
+implemented in \pystruct and described in Chapter~\ref{ch:structured_pystruct}. The combination
+of fusion moves and AD$^3$ described in Chapter~\ref{ch:exact_learning} allows us
+to learn the SSVM to optimality \emph{exactly}.
 

### 2013-09-15 intro.tex
zusammenfassung/abstract.
start intro

@@
 \chapter{Introduction}
+This thesis investigates the use of structured learning algorithms to the task of semantic image
+segmentation. Both topics have recieve much attention in the computer vision
+and machine learning communities lately. Unfortunately, learning structured
+prediction in computer vision applications is still little understood.
+We hoe this thesis will further the understanding of structured learning methods,
+give practical guidance and provide implementations to further the use of structured
+prediction methods in computer vision.
 
-What is semantic segmentation?
-Datsets? Input?
+The task of semantic segmentation is defined as predicting a semantic class label
+for each pixel in a natural image. Predicting semantic labels on such a fine scale
+provides ??? an important step towards general scene understanding.
+%
+Applications of scene understanding include automatic image interpretation for retrieval,
+ autonomous driving and mobile robotics.
+%
+related work?? here again?
+Current systems for semantic segmentation emply a wealth of different paradigms, most notably:
+The ranking of figure ground hypotheses
+conditional random field / superpixel? kraehenbuehl?
+detector based
+sparse coding
 
+We will focus on the use of conditional random fields, which have shown very promising.
+Using the paradigm of structural support vector machines, it is possible to learn conditional
+random fields to directly minimize a loss of interest. In particular, conditional random fields allow
+to combine different cues, possibly produced by other approaches, in a principled manner.
+%
+We use four different datasets to evaluate our approach: Graz02, Pascal VOC 2010, MSRC-21
+and NYU V2.
+%
 Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
-Why is it not solved yet?
-
-Structured approaches, latent variable approaches
+Structured approaches
+%
 
@@
 Contributions:
-- Introduces general clustering algorithm that improves pipeline in serveral places
-- Shows that weakly supervised object class segmentation (not semantic segmentation) is possible
-- Provides general efficient framework for structured prediction
-- Analyze max-margin learning algorithms in different regimes
-- shows exact learning is possible in loopy graphs and improves results (sometimes?)
-- shows importance of edge features
-- Demonstrates how latent variables benefit semantic segmentation
+- Introduces general clustering algorithm that improves pipeline in serveral places.
+- Shows that weakly supervised object class segmentation (not semantic segmentation) is possible.
+- Provides general efficient framework for structured prediction.
+- Analyze max-margin learning algorithms in different regimes.
+- shows exact learning is possible in loopy graphs and improves results (sometimes?).
+- shows importance of edge features.
+
+\section{Thesis Outline}
+We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm}, which is used as a building
+block in several places for our semantic segmentation algorithms.
+We will then introducing a mostly unstructured approach to semi-supervised learning for semantic segmentation in Chaper~\ref{ch:semi_supervised},
+investigating the problem of annotating training data for semantic segmentation.
+The central topic of this thesis is introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our software library
+for implementing structured learning and prediction algorithms.
+Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely used structured prediction algorithms, focusing on
+which algorithm is best to use in a given situation. In particular, we investigate learning behavior for semantic segmentation on several datasets.
+
+The problem of learning with approximate inference is investigated in Chapter~\ref{ch:exact_learning}.
+We develop a strategy for efficient caching and a combination of inference algorithms that allows us to
+learn SSVMs for semantic image segmentation exactly, even though the involved factor graphs contain many loops.
+
+Finally, Chapter~\ref{ch:nyu} applies the methods developed in Chapters~\ref{ch:structurd_pystruct} and \ref{ch:exact_learning}
+to the problem of semantic annotation with structure classes in RGB-D data. We demonstrate that we are able to learn
+meaningful spacial relations, and improve upon the state of the art in the NYU V2 datasets that we consider.
 

### 2013-09-16 partial_supervision/paper.tex
Shorten some titles because they were ugly
trying to rescue semi supervised stuff

@@
-\chapter{Weakly Supervised Object segmentation via Multiple Instance Learning}
+\chapter{Weakly Supervised Object Segmentation}
 
@@
 easier to obtain: whole image labels.
+While we do not reach the accuracy of competing fully supervised approaches,
+our efficient, semi-supervised method is potentially able to scale to much
+larger datasets, without the need of time-consuming manual annotation on
+pixel-level.
 
@@
 %\end{enumerate}
-In our approach, we work with a set of candidate segments, generated using constrained parametric min-cuts~\citep{carreira2010constrained}.
-For each image, these segments are a set of overlapping, object-like regions, which serve
-as candidates for object locations.
+In our approach, we work with a set of candidate segments, generated using
+constrained parametric min-cuts~\citep{carreira2010constrained}.  For each
+image, these segments are a set of overlapping, object-like regions, which
+serve as candidates for object locations.
 
-We formulate weakly supervised multi-class image segmentation as a multi-instance problem, based on candidate segments.
-In multi-instance learning~\citep{dietterich1997solving}, each training example is given as a multi-set of instances, called a bag.
-Each instance is represented as a feature vector $x$ and a label $y$.
-A bag is labeled positive if it contains at least one positive example, and negative otherwise.
-During training, only the labels of the training bags, not of the instances inside the bags, are known.
-The goal is to learn a classifier for unseen bags. 
-Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we assume that bags are simply sets, not multi-sets.
-Then a bag is an element of the power set $2^\mathcal{X}$ and the task is
-to learn a function
+We formulate weakly supervised multi-class image segmentation as a
+multi-instance problem, based on these candidate segments.  In multi-instance
+learning~\citep{dietterich1997solving}, each training example is given as a
+multi-set of instances, called a bag.  Each instance is represented as a
+feature vector $x$ and a label $y$.
+
+A bag is labeled positive if it contains at least one positive example, and
+negative otherwise.  During training, only the labels of the training bags, not
+of the instances inside the bags, are known.  The goal is to learn a classifier
+for unseen bags. 
+Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we
+assume that bags are simply sets, not multi-sets.  Then a bag is an element of
+the power set $2^\mathcal{X}$ and the task is to learn a function
 \begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
-Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.
-It is often assumed that the $f_{MI}$ function stems from a so-called underlying concept, given by an (unknown) function
+Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and
+labels $y_i \in \{-1,+1\}$.  It is assumed that the $f_{MI}$ function stems
+from a so-called underlying concept, given by an (unknown) function
 $f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
@@
 
-Sometimes, the goal of finding $f_{MI}$ is extended to finding labels not only
-for new bags, but also for all the separate instances within a bag
-\citep{liconvex2010,zha2008joint}, that is finding $f_{I}$.\\ 
-Even though finding $f_{I}$ is sometimes included in the task statement, there
-has been very little work that actually reported accuracy on instance label
-prediction. Part of the reason for this might be that for many of the datasets
-used in multi-instance learning no ground truth for instances exists.
+Multi-instance learning is a natural formulation for image classification and
+has been successfully applied in this task~\citep{zhou2007multi}. We propose to
+go a step further and apply multi-instance learning to the task of object-class
+segmentation in natural images, by also classifying instances, not only bags.
+In this, we follow the work of \citet{liconvex2010} and \citet{zha2008joint}, who not
+only learn $f_{MI}$, but also $f_{I}$.
 
-We look explicitly at accuracy on instance-level since we are interested in
-actually segmenting images, not just classifying them. For multi-class image
-segmentation, there are some hand-labeled datasets that provide ground truth
-on pixel level. We use this ground truth to evaluate the performance of our
-method. This approach does not exactly correspond to instance-level ground truth --
-since the instances are segments, not pixels -- but relates to it closely.
+In our model, each image forms a bag, while the candidate segments correspond
+to the instances contained in the bag. During learning, only presence of object
+classes is needed as bag level supervision. By learning $f_{I}$, we are then
+able to predict for individual segments whether they contain the object class
+of interest, thereby obtaining a segmentation of the object.
+To measure the performance of our algorithm, we use a dataset that not only contains
+image level annotation, but also pixel level annotation of object. This allows
+us to judge the succes of learning on instance level.
 
-%TODO theoretical part reference?
-% TODO repetition below?
-
-This chapter explores the 
-application of multi-instance learning algorithms to the task of partially supervised image segmentation.
-Multi-instance learning is a natural formulation for image classification and has been
-successfully applied in this task~\citep{zhou2007multi}. We propose to go a step further and apply
-multi-instance learning to the task of object-class segmentation in natural
-images.
+%This chapter explores the 
+%application of multi-instance learning algorithms to the task of partially supervised image segmentation.
+%FIXME not here!
 % FIXME relate to alex here
-To our knowledge, all previous methods in the field use strong
-supervision, meaning manual pixel-wise annotation of training images. This
-approach does not scale to larger datasets, especially if one expects
-consistency and quality in the segmentations.
-
-After our empirical evaluation on image segmentation, we will extend the existing theoretical framework
-on instance classification based on multi-instance prediction.
+%To our knowledge, all previous methods in the field use strong
+%supervision, meaning manual pixel-wise annotation of training images. This
+%approach is problematic, as scaling to larger datasets is limited by the
+%effort to provide annotation, especially if one expects
+%consistency and quality in the segmentations.
 
@@
 \subsection{Proposal Object Segments}\label{related_segments}
-Most work on multi-class segmentation focuses on strong supervision on superpixel level. There is still little work on using candidate segments.
-The method we use for generating candidate segments is Constraint Parametric Min-Cuts (CPMC) from \citet{carreira2010constrained}.
-This method creates a wide variety of overlapping segments. Support vector regression (SVR) is trained on these segments
-to estimate the overlap of segments with ground truth object-class labeling
-from the Pascal VOC dataset~\citep{pascal}. This provides a ranking of candidate segments, according to how ``object-like'' they are, which
-allows for selecting only a limited number of very object-like segments.
-The method performed well on a variety of datasets.
-A similar approach was investigated by \citet{endres2010category}.
+Most work on multi-class segmentation focuses on strong supervision on
+superpixel level. There is still little work on using candidate segments.  The
+method we use for generating candidate segments is Constraint Parametric
+Min-Cuts (CPMC) from \citet{carreira2010constrained}.  This method creates a
+wide variety of overlapping segments. Support vector regression (SVR) is
+trained on these segments to estimate the overlap of segments with ground truth
+object-class labeling from the Pascal VOC dataset~\citep{pascal}. This provides
+a ranking of candidate segments, according to how ``object-like'' they are,
+which allows for selecting only a limited number of very object-like segments.
+The method performed well on a variety of datasets. A similar approach was
+investigated by \citet{endres2010category}.
 %Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
@@
 kernel on bags, defined in terms of a kernel on instances. 
-The basic principle of multi-instance kernel is similar to a soft-max over instances in
-each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. They show that the multi-instance kernel
+The basic principle of the multi-instance kernel is similar to a soft-max over instances in
+each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. \citet{gaertner2002multi} show that the multi-instance kernel
 is able to separate bags if and only if the original kernel on instances is able to separate the underlying concepts.
-The method of \citet{gaertner2002multi} has a particular appeal in that it transforms a multi-instance problem into a standard
+The method of multi-instance kernels has a particular appeal in that it transforms a multi-instance problem into a standard
 classification problem by using an appropriate kernel. The downside of this approach is that it does
@@
 \citet{zhou2009multi} explicitly address the fact that instances are not
-indpendent within a bag , leading to an algorithm that can take advantage of
+independent within a bag , leading to an algorithm that can take advantage of
 possible correlations. Computational costs of their algorithm does not scale
@@
 
-\subsection{Semantic Scene Segmentation via Multi-Instance Learning}
+\subsection{Semantic Scene Segmentation using Weak Annotation }
 Recently, several methods have been proposed to obtain semantic segmentations
@@
 unspecific ``background''.  The unspecific background class contains much more
-clutter than for example ``grass'' and is therefore much harder to model.
-This makes disseminating the interesting part in multi-class object recognition
-challenging, since it is not necessary possible to identify non-object regions
-easily.
+clutter than for example ``grass'' and is therefore much harder to model. Additionally,
+object classes themselves are much harder to capture using low-level textural information only.
+This makes disseminating the distinctive features in multi-class object recognition
+much more challenging.
 
@@
 kernels~\citep{gaertner2002multi}.  Multi-instance kernels are a form of set
-kernels %TODO add reference to earlier work
+kernels %TODO add reference to earlier work -> convolution kernel
  that transform a kernel on instance level to a kernel on bag level.  We
@@
 bag. 
-
-%TODO one classifier per class!
+We extend the MIK SVM to the multi-class case using the one-vs-rest approach,
+learning one SVM for each class.
 
@@
 classification. In our setup, each segment might be given multiple labels.
-%TODO explain!
 To obtain a pixel-level object-class segmentation, we have to combine these.
-When building the segmentation for a given image, we only consider classes
-whose presence was predicted on image level.  Since we do not make use of the
-ground truth segmentation during training, we cannot learn an optimal
+As each image in the training set contains only a single class, we only consider the class
+predicted by the bag-level one-vs-rest classifier.
+Since we do not make use of the ground truth segmentation during training, we cannot learn an optimal
 combination as in \citet{li2010object} but perform a simple majority vote
@@
 \end{equation}
-where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an
-%FIXME we didn't describe the dataset yet!
+where $Y$ is the set of possible object classes, $S_i$ enumerates all segments within an
 image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
@@
 images of three object classes, bike, car and person.  Each image may contain
-multiple instances of the same class.
+multiple instances of the same class, but only one class is present per image.
 
-%From each image, we extract overlapping object-like segments using constraint
-%parametric min cuts (CPMC,\citet{carreira2010constrained}). We use the top 200 ranked segments and discard the
-%rest.
-%Each segment is described using SIFT~\citet{lowe2004distinctive} and
-%ColorSIFT~\citet{van2009evaluating} features, from which we compute bag of
-%visual word histograms as well as histograms of oriented
-%gradients~\citet{dalal2005histograms}. We construct one MI-Kernel per
-%feature which are combined using multiple kernel learning.
 We adjusted parameters on a hold-out validation set using bag-level information
@@
 $70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level, it can serve as a
-baseline for research on weakly supervised methods for object class segmentation.
+baseline for research on weakly supervised methods for object-class segmentation.
 
@@
 supervision, our result can work as a baseline for further research into weakly
-supervised object class segmentation.
-
-%In future work, we plan to scale our approach to much larger image datasets. As much
-%more images with weak annotations are available than with pixel level segmentation,
-%we hope that we can improve upon the state-of-the-art in object-class segmentation
-%by making use of larger bodies of training images.
+supervised object-class segmentation.

### 2013-09-16 partial_supervision/paper.tex
rename MIK-Instance to MIK-I for tables ^^ also some minor fixes.

@@
 
-% TODO contrast with Alex
+Recently, several approaches have been proposed for semi-supervised semantic segmentation.
+While these are close to our work, there are several important distinctions.
+We address the task of object-class segmentation, that concerns object categories, while
+semantic segmentation approaches often focus on "stuff" categories like "sky" and "road" that
+are more easily detected using simple texture features.
+Additionally, in contrast to \citet{vezhnevets2011weakly}, our approach is in
+principle applicable to large datasets, the regime where weak annotation is
+most useful.
 
-%The task of multi-class segmentation is often split up in a segmentation
-%and a recognition part. Random forest methods often classify each pixel
-%separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
-%over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
-%computational burden of classifying each pixel separately, but have
-%two drawbacks:
-%\begin{enumerate}
-    %\item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
-    %\item Segment boundaries are decided on the lowest level by generating
-        %superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
-%\end{enumerate}
 In our approach, we work with a set of candidate segments, generated using
@@
 %They proposed two methods, MIMLBoost and MIMLSvm. MIMLBoost transforms a MIML task into
-%a multi-instance learning task by reformulating the problem of \eqref{miml} as
+%a multi-instance learning task by reformulating the problem of \Eqref{miml} as
 %\begin{equation} f_{MIL} \colon 2^\mathcal{X}\times \mathcal{Y} \rightarrow \{ -1, +1\}.\end{equation}
@@
 classifier. We propose to transform a bag-level classifier $f_{MI}$ as given by
-the SVM and \eqref{mi_kernel} into an instance-level classifier by setting
+the SVM and \Eqref{mi_kernel} into an instance-level classifier by setting
 $f_{I}(x):=f_{MI}(\{x\})$, in other words, by considering each instance as its own
@@
 
-\subsection{Instance-Level Predictions using Multi-Instance Kernels}
+\subsection{Instance-Level Predictions using MI-Kernel}
 
@@
 kernels, we transform $f_{I}$ back to an instance-level classifier, using the
-multi-instance learning assumption (\eqref{multi_instance}). We refer to these
-instance-based MIK predictions as MIK-instance. In all experiments, the
+multi-instance learning assumption (\Eqref{multi_instance}). We refer to these
+instance-based MIK predictions as MIK-I. In all experiments, the
 parameters of the MI-Kernel and SVM are adjusted using MIK and then used with
-both MIK and MIK-instance.  This facilitates very fast parameter scans since
+both MIK and MIK-I.  This facilitates very fast parameter scans since
 MIK is very efficient to compute. Note that we cannot adjust parameters using
@@
     \toprule
-        & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-instance \\
+        & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-I \\
     \cmidrule(rl){2-8}
@@
     \caption{Bag level performance of various MIL algorithms on the standard Musk
-    datasets. All but MIK provide instance-level labeling. \tablabel{muskacc}}
+    datasets. All but MIK provide instance-level labeling.\tablabel{muskacc}}
 
 \end{table}
-We compared the performance of MIK, MIK-instance and state-of-the-art MI
+We compared the performance of MIK, MIK-I and state-of-the-art MI
 methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
 \Tabref{muskacc}. Results were obtained using 10-fold cross-validation. While
-the computational complexity of MIK-instance is very low compared to the other
+the computational complexity of MIK-I is very low compared to the other
 methods, it achieves competitive results.  Using instance-level labels results
-in a slight loss of accuracy of MIK-instances, compared to MIK\@. This small
+in a slight loss of accuracy of MIK-I, compared to MIK\@. This small
 degradation of performance is quite surprising, since the model was not trained
@@
 of segments might correspond to the object.
-\Tabref{musk-witness} compares the witness rates of MIK-instance,
+\Tabref{musk-witness} compares the witness rates of MIK-I,
 miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
-datasets. MIK-instance is able to achieve similar accuracy with much less
+datasets. MIK-I is able to achieve similar accuracy with much less
 witnesses than the other methods.  Note that Musk1 consists of very small bags
@@
     SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
-    MIK-instance& 88.0          & 99\%                &  85.2          & 62.3\%\\
+    MIK-I& 88.0          & 99\%                &  85.2          & 62.3\%\\
     \bottomrule
@@
 We train one multiple kernel learning (MKL) SVM per class using MIK and predict
-class labels on segment level using MIK-instance. If at least one SVM
+class labels on segment level using MIK-I. If at least one SVM
 classifies a segment as positive, it is associated with the most confident
@@
 Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative
-results are shown in \figref{results}. The overall accuracy on images labels,
+results are shown in \Figref{results}. The overall accuracy on images labels,
 which is the task that was actually trained, is $90\%$.  While the performance

### 2013-09-16 exact_learning/paper.tex
remove "will"

@@
 \section{Learning Structure Support Vector Machines with Approximate Inference}
-In this chapter, we will use the 1-slack cutting plane algorithm, as we found this
+In this chapter, we use the 1-slack cutting plane algorithm, as we found this
 algorithm to be most suitable for out approach. For reference, the algorithm is outlined
@@
 in line~\ref{get_cutting_plane}. We investigate algorithms $I$ (often called separation
-oracles in the context) that will not yield the exact maximum here.
+oracles in the context) that do not yield the exact maximum here.
 There are two groups of inference procedures, as identified in
@@
 of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
-hand, will solve the loss-augmented prediction in line~\ref{get_cutting_plane}
+hand, does solve the loss-augmented prediction in line~\ref{get_cutting_plane}
 exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
 that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.
-%We will elaborate on
+%We elaborate on
 %the properties of under-generating and over-generating inference procedures in
@@
 
+%\begin{table}
+    %\begin{center}
+    %\begin{tabularx}{\linewidth}{lccccccccccccccccccccc@{\quad}cc}
+        %\toprule
+                    %& \rotatebox[origin=c]{90}{building}
+                    %& \rotatebox[origin=c]{90}{gras}
+                    %& \rotatebox[origin=c]{90}{tree}
+                    %& \rotatebox[origin=c]{90}{cow}
+                    %& \rotatebox[origin=c]{90}{sheep}
+                    %& \rotatebox[origin=c]{90}{sky}
+                    %& \rotatebox[origin=c]{90}{airplane}
+                    %& \rotatebox[origin=c]{90}{water}
+                    %& \rotatebox[origin=c]{90}{face}
+                    %& \rotatebox[origin=c]{90}{car}
+                    %& \rotatebox[origin=c]{90}{bicycle}
+                    %& \rotatebox[origin=c]{90}{flower}
+                    %& \rotatebox[origin=c]{90}{sign}
+                    %& \rotatebox[origin=c]{90}{bird}
+                    %& \rotatebox[origin=c]{90}{book}
+                    %& \rotatebox[origin=c]{90}{chair}
+                    %& \rotatebox[origin=c]{90}{road}
+                    %& \rotatebox[origin=c]{90}{cat}
+                    %& \rotatebox[origin=c]{90}{dog}
+                    %& \rotatebox[origin=c]{90}{body}
+                    %& \rotatebox[origin=c]{90}{boat}
+                    %& \rotatebox[origin=c]{90}{\textbf{Global}}
+                    %& \rotatebox[origin=c]{90}{\textbf{Average}}
+                    %\\
+        %\cmidrule{1-24}
+    %unary &  27 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11\\
+    %approx.& 30 & 11\\
+    %exact & 30 & 11\\
+        %\cmidrule{1-24}
+        %other\\
+    %\bottomrule
+    %\end{tabularx}
+    %\end{center}
+    %\caption{Accuracies on the MSRC-21 Dataset. We compare our approach
+    %against approaches using the same unary potentials.\label{msrc_detailed}}
+
+%\end{table}
+
 \subsubsection{Pascal VOC 2010}
@@
 \end{table}
+
+%\begin{table}
+    %\begin{center}
+    %\begin{tabularx}{\linewidth}{lccccccccccccccccccccc@{\quad}cc}
+        %\toprule
+                    %& \rotatebox[origin=c]{90}{background}
+                    %& \rotatebox[origin=c]{90}{airplane}
+                    %& \rotatebox[origin=c]{90}{bicycle}
+                    %& \rotatebox[origin=c]{90}{bird}
+                    %& \rotatebox[origin=c]{90}{boat}
+                    %& \rotatebox[origin=c]{90}{bottle}
+                    %& \rotatebox[origin=c]{90}{bus}
+                    %& \rotatebox[origin=c]{90}{car}
+                    %& \rotatebox[origin=c]{90}{cat}
+                    %& \rotatebox[origin=c]{90}{chair}
+                    %& \rotatebox[origin=c]{90}{cow}
+                    %& \rotatebox[origin=c]{90}{dinning table}
+                    %& \rotatebox[origin=c]{90}{dog}
+                    %& \rotatebox[origin=c]{90}{horse}
+                    %& \rotatebox[origin=c]{90}{motorbike}
+                    %& \rotatebox[origin=c]{90}{person}
+                    %& \rotatebox[origin=c]{90}{potted plant}
+                    %& \rotatebox[origin=c]{90}{sheep}
+                    %& \rotatebox[origin=c]{90}{sofa}
+                    %& \rotatebox[origin=c]{90}{train}
+                    %& \rotatebox[origin=c]{90}{TV/monitor}
+                    %& \rotatebox[origin=c]{90}{\textbf{Average}}
+                    %\\
+        %\cmidrule{1-24}
+    %unary &  27 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11 & 11\\
+    %approx.& 30 & 11\\
+    %exact & 30 & 11\\
+        %\cmidrule{1-24}
+        %other \\
+    %\bottomrule
+    %\end{tabularx}
+    %\end{center}
+    %\caption{Accuracies on the Pascal VOC 2010 Dataset. We compare our approach
+    %against approaches using the same unary potentials.\label{pascal_detailed}}
+
+%\end{table}
 

### 2013-09-16 itm/paper.tex
zusammenfassung/abstract.
remove "will"

@@
 Before we start our investigation of semantic segmentation and object class segmentation,
-we will look into a general purpose clustering algorithm.
+we look into a general purpose clustering algorithm.
 %
@@
 components, and rebuilds the spanning tree.
-We will limit our discussion to the most widely used algorithm from~\citep{gower1969minimum}.
+We limit our discussion to the most widely used algorithm from~\citep{gower1969minimum}.
 %
@@
 ourself to optimizing over the set $\mathcal{F}$ of forests over $\mathbf{x}$ with $k$ components,
-as adding edges inside connected components will only decrease the objective.
+as adding edges inside connected components only decrease the objective.
 Thus we can formulate the clustering problem equivalently as
@@
 \toprule
-Algorithm &     Objective &     Deterministic&      Complexity \\
+Algorithm &     Objective &     Det.&      Complexity \\
 \cmidrule{1-4}
@@
 \end{tabularx}
-\caption{Comparing properties of related algorithms.}\label{nowotab}
+\caption{Comparing properties of related algorithms. Det.\ stands for Deterministic}\label{nowotab}
 \end{table}
@@
 %
-%We will also investigate optimizations of the objective 
+%We also investigate optimizations of the objective 
 %\Eqref{graphobjective} that go beyond the proposed method. 

### 2013-09-16 partial_supervision/paper.tex
remove "will"

@@
 work with strong supervision: a pixel-wise labeling of training images. In this
-chapter, we will investigate a method that works with annotation that is much
+chapter, we investigate a method that works with annotation that is much
 easier to obtain: whole image labels.
@@
 
-Recently, several approaches have been proposed for semi-supervised semantic segmentation.
-While these are close to our work, there are several important distinctions.
-We address the task of object-class segmentation, that concerns object categories, while
-semantic segmentation approaches often focus on "stuff" categories like "sky" and "road" that
-are more easily detected using simple texture features.
+Recently, several approaches have been proposed for semi-supervised semantic
+segmentation.  While these are close to our work, there are several important
+distinctions.  We address the task of object-class segmentation, that concerns
+object categories, while semantic segmentation approaches often focus on
+"stuff" categories like "sky" and "road" that are more easily detected using
+simple texture features.
 Additionally, in contrast to \citet{vezhnevets2011weakly}, our approach is in
@@
 
-%This chapter explores the 
-%application of multi-instance learning algorithms to the task of partially supervised image segmentation.
-%FIXME not here!
-% FIXME relate to alex here
-%To our knowledge, all previous methods in the field use strong
-%supervision, meaning manual pixel-wise annotation of training images. This
-%approach is problematic, as scaling to larger datasets is limited by the
-%effort to provide annotation, especially if one expects
-%consistency and quality in the segmentations.
-
 \section{Related Work}
@@
 \citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
-We will discuss only those that are relevant to the present treatment.
+We discuss only those that are relevant to the present treatment.
 
@@
 the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
-are again RBF-kernels, we will not consider $p$ explicitly in the following.
+are again RBF-kernels, we not consider $p$ explicitly in the following.
 
@@
 Training an SVM with this kernel produces a bag-level classifier for each
-class, which we will refer to as MIK\@.  This procedure is very efficient since
+class, which we refer to as MIK\@.  This procedure is very efficient since
 the resulting kernel matrix is of size number of bags, which is much smaller

### 2013-09-17 main.tex
add icra nyu paper
remove "will"
title page, smaller borders

@@
-\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=calc]{scrbook}
+\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11]{scrbook}
+%\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{book}
 
@@
 \usepackage{tabularx}
+\usepackage{tabu}
 \usepackage{graphicx}
@@
 \onehalfspacing
-\KOMAoptions{DIV=last}
+%\KOMAoptions{DIV=last}
 
@@
 
+\setlength{\tabcolsep}{2pt}
+
 \begin{document}
 
-\begin{spacing}{1}
+\begin{spacing}{1.6}
 \begin{titlepage}
+
 \begin{center}
-\vspace*{1in}
-\textbf{{\LARGE On Learning Structured Prediction For Semantic Segmentation}}
-\par
-\vspace{1.5in} {\large Andreas Christian M\"uller}
-\par \vspace{0.5in}
-Abteilung VI, Autonomie Intelligente Systeme \par
-Institut f\"ur Informatik
-\par \vspace{0.5in}
-Friedrich-Willhelm Universit\"at Bonn \par
-\vspace{0.5in} September 2013 \end{center}
-
+\Large\sffamily
+\vspace*{1cm}
+\textbf{{\huge Learning Structured Prediction For Semantic Image Segmentation}}\\[5mm]
+Dissertation\\%[3mm]
+zur\\%[3mm]
+Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\%[3mm]
+der\\%[3mm]
+Mathematisch-Naturwissenschaftlichen Fakult\"at\\%[3mm]
+der\\%[3mm]
+Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\%[3mm]
+vorgelegt von\\%[5mm]
+{\LARGE Andreas Christian M\"uller}\\%[5mm]
+aus\\%[3mm]
+Offenbach am Main\\%[10mm]
+Bonn, September 2013
+\end{center}
 \end{titlepage}
@@
 
-\chapter{Latent CRFs for Semantic Segmentation}\label{ch:latent}
+\include{nyu/paper}
 

### 2013-09-17 intro.tex
second page with advisor / reader, fix page numbering

@@
+\pagenumbering{arabic}
+
 \chapter{Introduction}
-This thesis investigates the use of structured learning algorithms to the task of semantic image
-segmentation. Both topics have recieve much attention in the computer vision
-and machine learning communities lately. Unfortunately, learning structured
-prediction in computer vision applications is still little understood.
-We hoe this thesis will further the understanding of structured learning methods,
-give practical guidance and provide implementations to further the use of structured
-prediction methods in computer vision.
+This thesis investigates the use of structured learning algorithms to the task
+of semantic image segmentation. Both topics have recieve much attention in the
+computer vision and machine learning communities lately. Unfortunately,
+learning structured prediction in computer vision applications is still little
+understood.  We hoe this thesis will further the understanding of structured
+learning methods, give practical guidance and provide implementations to
+further the use of structured prediction methods in computer vision.
 
@@
 for each pixel in a natural image. Predicting semantic labels on such a fine scale
-provides ??? an important step towards general scene understanding.
+constitutes an important step towards general scene understanding.
 %
@@
 
-We will focus on the use of conditional random fields, which have shown very promising.
-Using the paradigm of structural support vector machines, it is possible to learn conditional
-random fields to directly minimize a loss of interest. In particular, conditional random fields allow
-to combine different cues, possibly produced by other approaches, in a principled manner.
+We will focus on the use of conditional random fields, which have shown very
+promising.  Using the paradigm of structural support vector machines, it is
+possible to learn conditional random fields to directly minimize a loss of
+interest. In particular, conditional random fields allow to combine different
+cues, possibly produced by other approaches, in a principled manner.
 %
-We use four different datasets to evaluate our approach: Graz02, Pascal VOC 2010, MSRC-21
-and NYU V2.
+We use four different datasets to evaluate our approach: Graz02, Pascal VOC
+2010, MSRC-21 and NYU V2.
 %
@@
 \section{Thesis Outline}
-We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm}, which is used as a building
-block in several places for our semantic segmentation algorithms.
-We will then introducing a mostly unstructured approach to semi-supervised learning for semantic segmentation in Chaper~\ref{ch:semi_supervised},
-investigating the problem of annotating training data for semantic segmentation.
-The central topic of this thesis is introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our software library
-for implementing structured learning and prediction algorithms.
-Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely used structured prediction algorithms, focusing on
-which algorithm is best to use in a given situation. In particular, we investigate learning behavior for semantic segmentation on several datasets.
+We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm},
+which is used as a building block in several places for our semantic
+segmentation algorithms.  We will then introducing a mostly unstructured
+approach to semi-supervised learning for semantic segmentation in
+Chaper~\ref{ch:semi_supervised}, investigating the problem of annotating
+training data for semantic segmentation.  The central topic of this thesis is
+introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our
+software library for implementing structured learning and prediction
+algorithms.
+Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely
+used structured prediction algorithms, focusing on which algorithm is best to
+use in a given situation. In particular, we investigate learning behavior for
+semantic segmentation on several datasets.
 
-The problem of learning with approximate inference is investigated in Chapter~\ref{ch:exact_learning}.
-We develop a strategy for efficient caching and a combination of inference algorithms that allows us to
-learn SSVMs for semantic image segmentation exactly, even though the involved factor graphs contain many loops.
+The problem of learning with approximate inference is investigated in
+Chapter~\ref{ch:exact_learning}.  We develop a strategy for efficient caching
+and a combination of inference algorithms that allows us to learn SSVMs for
+semantic image segmentation exactly, even though the involved factor graphs
+contain many loops.
 
-Finally, Chapter~\ref{ch:nyu} applies the methods developed in Chapters~\ref{ch:structurd_pystruct} and \ref{ch:exact_learning}
-to the problem of semantic annotation with structure classes in RGB-D data. We demonstrate that we are able to learn
-meaningful spacial relations, and improve upon the state of the art in the NYU V2 datasets that we consider.
+Finally, Chapter~\ref{ch:nyu} applies the methods developed in
+Chapter~\ref{ch:structurd_pystruct} and Chapter~\ref{ch:exact_learning} to the
+problem of semantic annotation with structure classes in RGB-D data. We
+demonstrate that we are able to learn meaningful spacial relations, and improve
+upon the state of the art in the NYU V2 datasets that we consider.
 
 \section{Publications}
-The main material of this thesis has either been published in conference proceedings or has been
-submitted to conferences or journals. We now list the relevant publications.
+The main material of this thesis has either been published in conference
+proceedings or has been submitted to conferences or journals. We now list the
+relevant publications.
 \begin{description}

### 2013-09-17 main.tex
second page with advisor / reader, fix page numbering
working on conclusions

@@
-\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11]{scrbook}
+\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=.5cm]{scrbook}
 %\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{book}
@@
 \end{titlepage}
-
+%
+\thispagestyle{empty}
+\vspace*{5cm}
+\Large\sffamily
+\begin{center}
+Angefertigt mit Genehmigung\\
+der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
+der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
+\end{center}
+1. Gutachter Prof. Dr. Sven Behnke\\
+2. Gutachter Prof. Dr. Reinhard Klein\\
+Tag der Promotion:\\
+Erscheinungsjahr: 2013\\
+\end{spacing}
 \pagenumbering{Roman}
-
 \tableofcontents
-\end{spacing}
 %\listoffigures
@@
 
-\pagenumbering{arabic}
-
 \include{intro}
@@
 
-\chapter{Conclusion}
+\chapter{Conclusion and Future Direction}
+Current progress in computer vision \dots at an incredible speed,
+with new methods and applications emerging constantly.
+Still, there are large differences in the paradigms employed by different researchers.
+While it is generally aknowledged that context is one of the most important cues
+in recognizing objects and interpreting scence, there is little consensus on how
+contextual information should be modelled, or even learned.
+The structured prediction paradigm allows a principled approach to incoporating
+predictions within an image, and to include context information.
+We investigated the potential of structural methods for the application of semantic image segmentation
+and object recognition, appications to which structured methods have been applied successfully before.
+We focus on \emph{learning} of structural models and the interaction of inference and learning
+in the neighborhood models typically employed for semantic segmentation.
+The thesis presents a software implementation of a variety of popular learning algorithms for structural
+support vector machines, together with a thorough evaluation of their properties in the presence
+of approximate inference.
+We demonstated that exact learning is possible, even in the presence of loops in the
+\dots factor graph. We also demonstrate the power of conditioning interactions
+on data, learning spacial interactions in a RGB-D setting.
+
+There are several directions for future research that we think would be interesting to pursue as an extention
+of the presented results: 
+\paragraph[Inference Machines]
+Recently \citet{stoyanov2011empirical} started a new trend in structured
+prediction, which is sometimes called ``inference machines''. The basic
+principle is simple: view the process of prediction with a given inference
+procedure as a prediction method and optimize parameters of this method using
+empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
+loopy belief propagation as their inference algorithm and the optimization is
+carried out simply using gradient descent on the non-convex but differentiable
+loss function.  Other recent work in this direction includes
+\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
+conditional random field and \citet{jancsarylearning}, who used closed form
+inference in a Gaussian CRF\@.
+While these algorithms show great promise, their relation to the traditional
+approach of structured prediction used in this work is mostly unclear. In particular,
+if exact traditional learning is possible in a model, it is unclear how much
+can be gained by direct empirical risk minimization.
+Little empirical comparision was performed, and we are not aware of theoretical
+work in this direction, leaving much room for future investigation of this
+direction.
+\paragraph[Non-Linear Models]
+In this work, we only considered models that are linear in the input features--though
+features are highly non-linear in the original input pixels.
+Kernelization of structural support vector machines is trivial in theory, but
+had only limited success in the context of CRFs for image
+segmentation~\citep{lucchi2012structured}.
+Other
+
+\paragraph[Latent Variable Models]
+\paragraph[Feature Design]
+
+%Vision is challenging
+%context matters
+%structured prediction allows modeling and learning of context and relations
+%thesis gives insight into some relations of approximate inference and learning
+%gives framework for evaluation
+%comparison of methods
+
+which low-level cues?
+superpixels?
+non-linear models?
+latent variables?
+inference engines
 

### 2013-09-17 main.tex
work on conclusions

@@
 segmentation~\citep{lucchi2012structured}.
-Other
+Two major alternatives for non-linear CRFs were proposed in the literature,
+conditional neural fields~\citep{peng2009conditional} based on neural networks,
+and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
+Conditional neural fields have only applied to sequence classification so far,
+and extending them to our setting of semantic image segmentation would be very
+interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
+graphs for image processing, but not for higher level tasks such as semantic
+segmentation. If it is possible to include context in a meaningful way, it
+might be possible to solve even object-centric tasks such as object class
+segmentation with DTFs.
 

### 2013-09-17 structured_prediction/paper.tex
fixes von hannes

@@
 Here, $y$ is a \emph{structured} label, $\Phi$ is a joint feature function of
-$x$ and $y$, and $\theta$ are parameters of the model. \emph{Structured} means
+$x$ and $y$, and $\theta$ contains the parameters of the model. \emph{Structured} means
 that $y$ is more complicated than a single output class, for example a label
@@
 
-While $y$ could be a complicated like a parsing tree or the geometric
-configuration of a molecule, many setting, such as the image segmentation
+While $y$ could be complicated like a parsing tree or the geometric
+configuration of a molecule, many settings, such as the image segmentation
 setting we are interested in, can be reduced to the case where $y$ is a vector of
@@
 The benefit of using the factor graph representation is that it decomposes the
-function over subsets the variables of interest $y_i$. This allows us to apply
+function over subsets of the variables of interest $y_i$. This allows us to apply
 efficient optimization procedures for the prediction problem in
@@
 %
-To obtain a linear function as in \Eqref{main_equation} from
-\Eqref{log_factor_graph}, we can simply let each $\psi$ be of the form
+To obtain a linear function from
+\Eqref{log_factor_graph} as in \Eqref{main_equation}, we can simply let each $\psi$ be of the form
 \begin{equation}\eqlabel{general_psi}
@@
 \end{equation}
-The by far most common form is 
+The most common form by far is 
 \begin{equation}\eqlabel{linear_psi}
@@
     \theta &= \bigoplus_{F \in \mathcal{F}} \theta_F\\
-    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \Phi_F(x, y_{N_F})
+    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \Phi_F(x, y_{N_F}).
 \end{align}
@@
     \theta &= \bigoplus_{F \in \mathcal{F}}\ \bigoplus_{y_{N_F} \in \mathcal{Y}_{N_F}} \theta_{F, y_{N_F}}\\
-    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \left (\phi_F(x) \otimes e_{y_{N_F}} \right )
+    \Phi(x, y) &= \bigoplus_{F \in \mathcal{F}} \left (\phi_F(x) \otimes e_{y_{N_F}} \right ),
 \end{align}
-where $e_{y_{N_F}} \in \mathbb{R}^{|\mathcal{Y}_{N_F}|}$ is the indicator for a
+Here $e_{y_{N_F}} \in \mathbb{R}^{|\mathcal{Y}_{N_F}|}$ is the indicator for a
 given variable setting $y_{N_F}$.
-In words, $\Phi_F$ is build simply by creating a vector of
+In words, $\Phi_F$ is built simply by creating a vector of
 $|\mathcal{Y}_{N_F}|$ times the size of $\phi_F$, which is zero everywhere,
-except for the place corresponding to $y_{N_F}$. %TODO unclear!
+except for the position corresponding to $y_{N_F}$. %TODO unclear!
 
@@
 \begin{equation}\eqlabel{graphicalmodel}
-    p(y | x) = \frac{1}{Z_x}\prod_{F \in \mathcal{F}} \exp\left(\psi_F(x, y_{N_F})\right)
+    p(y | x) = \frac{1}{Z_x}\prod_{F \in \mathcal{F}} \exp\left(\psi_F(x, y_{N_F})\right).
 \end{equation}
-where
+Here
 \begin{equation}
@@
     \item[Generalization] The maximum margin principle yields generalization
-        bounds using the effective complexity~\citep{taskar2003max}.
+        bounds using the effective complexity~\citep{taskar2003max}, that are generally
+        tighter than corresponding VC-theoretical bounds.
     \item[Strong Convexity] The resulting optimization problem is strongly
@@
 \end{equation}
-The parameters $\theta$ are learned by minimizing the loss based soft-margin
+The parameters $\theta$ are learned by minimizing the loss-based soft-margin
 objective
@@
 \end{equation}
-with regularization parameter $C$. Here $\ell$ is an hinge-loss like upper bound
+with regularization parameter $C$. Here, $\ell$ is a hinge-loss like upper bound
 on the empirical $\Delta$-risk:
 \begin{equation}\eqlabel{loss_augmentation}
-    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+
+    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+.
 \end{equation}
-
 This is an instance of regularized empirical risk minimization, with a
@@
 run-times for batch algorithms. On the other hand, loss-augmented prediction in
-batch algorithms is embarrassingly parallel, allowing the use of multiple
+batch algorithms is \emph{embarrassingly parallel}, allowing the use of multiple
 processors with almost linear speed improvements.
@@
 using subgradient descent.  In light of the complexity of solving the
-loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
+loss-augmented prediction problem in \Eqref{loss_augmentation}, it is natural
 to work in a stochastic setting (see \citet{ratliff2007online}).
@@
 %Theoretical convergence guarantees of \ldots can be given whenever \ldots,
-but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
-Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
+In practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
+Many practitioners adopt the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
 \begin{equation}
@@
 \begin{equation}\eqlabel{averaging}
-    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}
+    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}.
 \end{equation}
@@
 It is possible to use mini-batches instead of processing one sample at a time
-to make use of multiple processors for loss-augmented prediction. Unfortunately
+to make use of multiple processors for loss-augmented prediction. Unfortunately,
 this negatively affects the number of iterations needed, and did not provide a
@@
         \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
-        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for all } i=1,\dots,k$
+        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for } i=1,\dots,k$
         \Repeat
@@
             \vspace{-10mm}
-            \Until no $\W_i$ changes any more.
+            \Until no $\W_i$ changes anymore.
         \end{algorithmic}
@@
 again, with the new set of constraints.
-The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
+The algorithm terminates when no constraint can be found that is violated more than $\epsilon$,
 which guarantees a suboptimality of at most $\epsilon$.
@@
 program only every $r$ samples, for some small integer $r$
-(\citet{joachims2009cutting} suggest 100). While this strategy on its own did
+(\citet{joachims2009cutting} suggest $r=100$). While this strategy on its own did
 not provide a large benefits in our experiments, it allows for parallel
@@
 %
-By construction only a single constraint is added in each iteration of
+By construction, only a single constraint is added in each iteration of
 Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
@@
 less variables than in the $n$-slack algorithm. The down-side of this is that
-the loss-augmented prediction problem has to be solve much more often until
+the loss-augmented prediction problem has to be solved much more often until
 convergence. This is reflected in a convergence rate of $O(\frac{1}{kT})$,
@@
         (line~\ref{get_cutting_plane} in Algorithm~\ref{alg_one_slack}).
-        For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we
+        For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$, we
         find

### 2013-09-18 structured_prediction/paper.tex
rest of hannes notes

@@
  represent \emph{products} of factors.  To simplify presentation, we work
-directly in the log-domain with respect to the more standard product representation.}
+directly in the log-domain of the more standard product representation.}
 \begin{equation}\eqlabel{log_factor_graph}
@@
 If $f$ is chosen as in \Eqref{linear_psi}, then the resulting distribution
-belongs to the exponential family.
+belongs to the exponential family, the class of probablility distributions most
+commonly used in the graphical model iterature.
 
@@
 Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
-This has the advantage of making the solution of the QP faster, as it contains far
+This has the advantage of producing a QP that easyer to solve, as it contains far
 less variables than in the $n$-slack algorithm. The down-side of this is that

### 2013-09-18 partial_supervision/paper.tex
slightly better related work for weakly supervised.

@@
 
-\subsection{Semantic Scene Segmentation using Weak Annotation }
-Recently, several methods have been proposed to obtain semantic segmentations
-of images using only image-level
-supervision~\citep{vezhnevets2011weakly,verbeek2007region,vezhnevets2010towards}.
-\citet{vezhnevets2011weakly}, for example, report impressive results on the
-MSRC dataset.
+\subsection{Semantic Scene Segmentation using Weak Annotation}
+Learning semantic segmentation from image-level annotation was first investigated in \citet{verbeek2007region},
+using a semi-supervised conditional random field on patches. \citet{verbeek2007region} evaluated their approach
+on the MSRC-9 datasets.
+More recently, simliar approaches were proposed by
+\citet{vvezhnevets2011weakly} and \citet{vezhnevets2010towards}.
+\citet{vezhnevets2011weakly} independently developed a multiple-instance based
+approach to segmentation, and report impressinve results on the MSRC-21 dataset.
 

### 2013-09-18 partial_supervision/paper.tex
proof reading semi supervised

@@
 simple texture features.
-Additionally, in contrast to \citet{vezhnevets2011weakly}, our approach is in
+Additionally, in contrast to, for example, \citet{vezhnevets2011weakly}, our approach is in
 principle applicable to large datasets, the regime where weak annotation is
@@
 In our approach, we work with a set of candidate segments, generated using
-constrained parametric min-cuts~\citep{carreira2010constrained}.  For each
-image, these segments are a set of overlapping, object-like regions, which
-serve as candidates for object locations.
+constrained parametric min-cuts~\citep{carreira2010constrained}.  The procedure yields
+segments that are overlapping, object-like regions, which serve as candidates
+for object locations.
 
@@
 Multi-instance learning is a natural formulation for image classification and
-has been successfully applied in this task~\citep{zhou2007multi}. We propose to
+has been successfully applied to this task~\citep{zhou2007multi}. We propose to
 go a step further and apply multi-instance learning to the task of object-class
@@
 which allows for selecting only a limited number of very object-like segments.
-The method performed well on a variety of datasets. A similar approach was
-investigated by \citet{endres2010category}.
+The method performed well on a variety of datasets, building the basis of
+a very successful entry to the Pascal VOC segmentation
+challenge~\citep{li2010object}. A similar approach to whole-object segment
+proposals was investigated by \citet{endres2010category}.
 %Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
@@
 More recently, simliar approaches were proposed by
-\citet{vvezhnevets2011weakly} and \citet{vezhnevets2010towards}.
+\citet{vezhnevets2011weakly} and \citet{vezhnevets2010towards}.
 \citet{vezhnevets2011weakly} independently developed a multiple-instance based
@@
 
-While semantic segmentation is closely related to multi-class image
-segmentation, there are important distinctions: In semantic segmentation, each
-pixel has a semantic annotation, also containing non-object ``stuff'' classes like
-``sky'', ``grass'' and ``water''. In multi-class image segmentation, the focus
-is on objects, with possibly large parts of the image being labeled as
-unspecific ``background''.  The unspecific background class contains much more
-clutter than for example ``grass'' and is therefore much harder to model. Additionally,
-object classes themselves are much harder to capture using low-level textural information only.
-This makes disseminating the distinctive features in multi-class object recognition
-much more challenging.
+While semantic segmentation is closely related to the task of multi-class image
+segmentation that we are considering in this chapter, there are important
+distinctions: In semantic segmentation, each pixel has a semantic annotation,
+also containing non-object ``stuff'' classes like ``sky'', ``grass'' and
+``water''. In multi-class image segmentation, the focus is on objects, with
+possibly large parts of the image being labeled as unspecific ``background''.
+The unspecific background class contains much more clutter than for example
+``grass'' and is therefore much harder to model. Additionally, object classes
+themselves are much harder to capture using low-level textural information
+only.  This makes disseminating the distinctive features in multi-class object
+recognition much more challenging, and requires a more holistic apporach to
+recognition than these patch-based or superpixel-based approaches.
 
@@
 
-With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the
+With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, the
 corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
-as
+is defined as
 \begin{equation}\eqlabel{mi_kernel}
@@
 where $p \in \mathbb{N}$ is a free parameter~\citep{gaertner2002multi}. As we use
-the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
+the RBF-kernel $k_{\text{rbf}}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
 are again RBF-kernels, we not consider $p$ explicitly in the following.
@@
 class, which we refer to as MIK\@.  This procedure is very efficient since
-the resulting kernel matrix is of size number of bags, which is much smaller
-than a kernel matrix of size number of instances, as is commonly used in the
+the resulting kernel matrix is of size of the number of bags, which is much smaller
+than a kernel matrix of size of the number of instances, as is commonly used in the
 literature~\citep{andrews2003support,nguyen2010new,zhang2008m3miml}.  Another
-advantage over other methods is, that it uses a single convex optimization,
+advantage over other methods is that it uses a single convex optimization,
 whereas other approaches often use iterative
@@
 bag.
-We extend the MIK SVM to the multi-class case using the one-vs-rest approach,
-learning one SVM for each class.
 
@@
 SIFT~\citep{lowe2004distinctive} and ColorSIFT~\citep{van2009evaluating}
-features, from which we compute bag of visual word histograms. Additionally, we
-use histograms of oriented gradients~\citep{dalal2005histograms} on the
+features on multiple scales, from which we compute bag-of-visual-word histograms. Additionally, we
+use a pyramid of histograms of oriented gradients~\citep{dalal2005histograms} on the
 segments.  We use RBF-kernels for all of the features, constructing one
@@
 The framework described above yields an image-level and a segment-level
-classification. In our setup, each segment might be given multiple labels.
-To obtain a pixel-level object-class segmentation, we have to combine these.
-As each image in the training set contains only a single class, we only consider the class
-predicted by the bag-level one-vs-rest classifier.
-Since we do not make use of the ground truth segmentation during training, we cannot learn an optimal
-combination as in \citet{li2010object} but perform a simple majority vote
-instead.  We merge segments into pixel-level class labels by setting the label
-$y_x$ of a pixel $x$ according to
+classification.  To obtain a pixel-level object-class segmentation, we have to
+combine these.  Since we do not make use of the ground truth segmentation
+during training, we cannot learn an optimal combination as in
+\citet{li2010object} but perform a simple majority vote instead.  We merge
+segments into pixel-level class labels by setting the label $y_x$ of a pixel
+$x$ according to
 \begin{equation}
@@
 parameters of the MI-Kernel and SVM are adjusted using MIK and then used with
-both MIK and MIK-I.  This facilitates very fast parameter scans since
+both MIK and MIK-I.  This facilitates very fast parameter search since
 MIK is very efficient to compute. Note that we cannot adjust parameters using
@@
 We compared the performance of MIK, MIK-I and state-of-the-art MI
-methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
+methods on the Musk benchmark datasets~\citep{dietterich1997solving} in
 \Tabref{muskacc}. Results were obtained using 10-fold cross-validation. While
@@
 methods, it achieves competitive results.  Using instance-level labels results
-in a slight loss of accuracy of MIK-I, compared to MIK\@. This small
-degradation of performance is quite surprising, since the model was not trained
-to provide any instance-level labels.
+in a slight loss of accuracy of MIK-I, compared to MIK\@. Interestingly, even
+though the model was not trained to provide any instance-level labels, the
+performance is still competative.
 
@@
 We adjusted parameters on a hold-out validation set using bag-level information
-and used the training and test sets as given by the dataset.  It is
-straight-forward to extend the binary MIK method to the multi-class setting
-using a one-vs-all strategy.
+and used the training and test sets as specified in the dataset.
 We train one multiple kernel learning (MKL) SVM per class using MIK and predict
-class labels on segment level using MIK-I. If at least one SVM
-classifies a segment as positive, it is associated with the most confident
-class. Otherwise, it is assigned ``background'' or no class.
+class labels on segment level using MIK-I. 
 This yields a classification of each segment into one of four classes: car,
@@
 pixel-level annotations, whose pixel-level accuracy is around
-$70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level, it can serve as a
+$70\%$~\citep{fulkerson2009class,schulz2011}, it can serve as a
 baseline for research on weakly supervised methods for object-class segmentation.

### 2013-09-18 intro.tex
page numbering, layouting...
trying to fix abstract :-/

@@
-\pagenumbering{arabic}
-
 \chapter{Introduction}
@@
 Contributions:
-- Introduces general clustering algorithm that improves pipeline in serveral places.
-- Shows that weakly supervised object class segmentation (not semantic segmentation) is possible.
-- Provides general efficient framework for structured prediction.
-- Analyze max-margin learning algorithms in different regimes.
-- shows exact learning is possible in loopy graphs and improves results (sometimes?).
-- shows importance of edge features.
+\begin{itemize}
+\item Introduces general clustering algorithm that improves pipeline in serveral places.
+\item Shows that weakly supervised object class segmentation (not semantic segmentation) is possible.
+\item Provides general efficient framework for structured prediction.
+\item Analyze max-margin learning algorithms in different regimes.
+\item shows exact learning is possible in loopy graphs and improves results (sometimes?).
+\item shows importance of edge features.
+\end{itemize}
 

### 2013-09-18 main.tex
fixes von hannes
page numbering, layouting...
trying to fix abstract :-/

@@
 \documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=.5cm]{scrbook}
+%\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=calc]{scrbook}
 %\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{book}
@@
 \onehalfspacing
-%\KOMAoptions{DIV=last}
+\KOMAoptions{DIV=last}
 
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
 
@@
 %
-\thispagestyle{empty}
+%\thispagestyle{empty}
+\clearpage
+\pagenumbering{Roman}
+\setcounter{page}{2}
 \vspace*{5cm}
@@
 \end{spacing}
-\pagenumbering{Roman}
 \tableofcontents
@@
 Automatische Segementierung und Erkennung von semantischen Klassen in
-natuerlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
-Einer der am weitest verbreiteten Ans\"atze zu dieser Aufgabe basiert auf
-Conditional Random Fields und Verfahren der strukturierten Vorhersage. Die
-Vorhersage f\"ur Modelle zur Bild-Markierung mittels Conditional Random Fields
-ist f\"ur gew\"ohnlich nur approiximativ moeglich, bedingt d\"urch das
-Vorhandensein von Kreisen in den zugrunde liegenden Nachbarschaftsgraphen.
-Aus diesem Grund, und trotz der Beliebtheit von Conditional Random Field
-Verfahren, ist das strukturierte Lernen f\"ur solche Anwendungen schlecht
-verstanden.
+nat\"urlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
+Diese Arbeit untersucht drei m\"oglichen Ansi\"atzen der Erkennung:
+ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
+von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.
+
+Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
+der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
+der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
+klare Vorteile auf vielen verschiedenen Datens\"atzen hat.
+
+Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
+Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, der
+mit Hilfe von Supervision in Form von Klassen-Pr\"asenz auf Bildern in der Lage
+ist ein Segmentierungsmodell zu Lernen. Nach unserem Wissen ist dies der erste
+Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.
+
+Der dritte und letzte Teil er Arbeit untersucht einen der am weitest
+verbreiteten Ans\"atze zu semantischen Segmentierung und
+Objektklassensegmentierung, n\"amlich Conditional Random Fields verbunden mit Verfahren
+der strukturierten Vorhersage.
+%Die Vorhersage f\"ur Modelle zur Bild-Markierung
+%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
+%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
+%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
+%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
+%solche Anwendungen schlecht verstanden.
 %
-Wir untersuchen verschiedene Lernalgorithmen in diesem
-Zusammenhang. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
-Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zu lernen.
+Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
+im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
+Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
 Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
@@
 \chapter*{Abstract}
-Automatic segmentation and recognition of semantic classes in natural images is an important open problem in computer vision.
-One of the most popular approaches to this task is based on conditional random fields and structured prediction.
-For image labeling tasks, prediction using conditional random fields is usually only approximate, because of the presence
-of loops of the neighborhood graph in the the image. Consequently, learning in
-this setting is still little understood, despite the popularity of these
-methods.
+Automatic segmentation and recognition of semantic classes in natural images is
+an important open problem in computer vision.
+This work investigates three different approaches to recognition: without supervision,
+with supervision on level of images, and with supervision on the level of pixels.
+The thesis comprises three parts.
+
+The first part introduces a clustering algorithm, that optimizes a novel information-theoretic
+objective function. We show that the proposed algorithm has clear advantages over standard algorithm
+from the Literature on a wide array of datasets.
+
+The second part of this work proposes an algorithm for automatic segmentation
+and recognition of object classes in natural images, that learns a segmentation model solely from
+annotation in form of presence and absence of object classes in images. To the best of our knowledge,
+this is the first algorithm of this kind, that explicitly addresses complex object classes.
+
+The third and last part of this work investigates one of the most popular
+approaches to the task of object class segmentation and semantic segmentation,
+based on conditional random fields and structured prediction.
 %
-We investigate several learning algorithms in this setting. We show
-how structured models can be learned exactly in practical settings, even in the
-presence of many loops in the underlying neighborhood graphs.  We introduce a
-software library that allows us to perform extensive empirical comparisons of
+%For image labeling tasks, prediction using conditional random fields is usually
+%only approximate, because of the presence of loops of the neighborhood graph in
+%the the image. Consequently, learning in this setting is still little
+%understood, despite the popularity of these methods.
+%
+We investigate several learning algorithms, in particular in combination with
+approximate inference procedures. We show how structured models for image
+segmentation can be learned exactly in practical settings, even in the presence
+of many loops in the underlying neighborhood graphs.  We introduce a software
+library that allows us to perform extensive empirical comparisons of
 state-of-the-art structured learning approaches. This allows us to characterize
-their practical properties in a range of applications, in particular for semantic segmentation
-and object class segmentation.
+their practical properties in a range of applications, in particular for
+semantic segmentation and object class segmentation.
 
@@
 
-Last but not least I am greatful to Anna M\"uller for her support,
+Last but not least I am greateful to Anna M\"uller for her support,
 encouragement and company. Thank you for giving me the strength needed for this
 endeavour.
-
+\cleardoublepage
+\pagenumbering{arabic}
 \include{intro}
@@
 
-\include{itm/paper}
+%\include{itm/paper}
 
-\include{partial_supervision/paper}
+%\include{partial_supervision/paper}
 
@@
 \chapter{Conclusion and Future Direction}
-Current progress in computer vision \dots at an incredible speed,
-with new methods and applications emerging constantly.
-Still, there are large differences in the paradigms employed by different researchers.
-While it is generally aknowledged that context is one of the most important cues
-in recognizing objects and interpreting scence, there is little consensus on how
+Current research in computer vision progresses at an incredible speed, with new
+methods and applications emerging constantly.  Still, there are large
+differences in the paradigms employed by different researchers.  While it is
+generally acknowledged that context is one of the most important cues in
+recognizing objects and interpreting scenes, there is little consensus on how
 contextual information should be modelled, or even learned.
-The structured prediction paradigm allows a principled approach to incoporating
-predictions within an image, and to include context information.
-We investigated the potential of structural methods for the application of semantic image segmentation
-and object recognition, appications to which structured methods have been applied successfully before.
-We focus on \emph{learning} of structural models and the interaction of inference and learning
-in the neighborhood models typically employed for semantic segmentation.
-The thesis presents a software implementation of a variety of popular learning algorithms for structural
-support vector machines, together with a thorough evaluation of their properties in the presence
-of approximate inference.
-We demonstated that exact learning is possible, even in the presence of loops in the
-\dots factor graph. We also demonstrate the power of conditioning interactions
+The structured prediction paradigm allows a principled approach to
+incorporating predictions within an image, and to include context information.
+We investigated the potential of structural methods for the application of
+semantic image segmentation and object recognition, applications to which
+structured methods have been applied successfully before.  We focus on
+\emph{learning} of structural models and the interaction of inference and
+learning in the neighborhood models typically employed for semantic
+segmentation.
+The thesis presents a software implementation of a variety of popular learning
+algorithms for structural support vector machines, together with a thorough
+evaluation of their properties in the presence of approximate inference.
+We demonstrate that exact learning is possible, even in the presence of loops in the
+underlying factor graph. We also demonstrate the power of conditioning interactions
 on data, learning spacial interactions in a RGB-D setting.
 
-There are several directions for future research that we think would be interesting to pursue as an extention
-of the presented results: 
+There are several directions for future research that we think would be
+interesting to pursue as an extension of the presented results, in particular in
+extending the second and third part:
+\paragraph[Large-Scale Weakly Supervised Object Class Segmentation]
+google
 \paragraph[Inference Machines]
@@
 can be gained by direct empirical risk minimization.
-Little empirical comparision was performed, and we are not aware of theoretical
+Little empirical comparison was performed, and we are not aware of theoretical
 work in this direction, leaving much room for future investigation of this
@@
 \paragraph[Latent Variable Models]
+Something latent
 \paragraph[Feature Design]
-
-%Vision is challenging
-%context matters
-%structured prediction allows modeling and learning of context and relations
-%thesis gives insight into some relations of approximate inference and learning
-%gives framework for evaluation
-%comparison of methods
-
+features are important
 which low-level cues?
 superpixels?
-non-linear models?
-latent variables?
-inference engines
 

### 2013-09-18 intro.tex
state of the art, slight copy-editing....

@@
 \chapter{Introduction}
-This thesis investigates the use of structured learning algorithms to the task
-of semantic image segmentation. Both topics have recieve much attention in the
-computer vision and machine learning communities lately. Unfortunately,
-learning structured prediction in computer vision applications is still little
-understood.  We hoe this thesis will further the understanding of structured
-learning methods, give practical guidance and provide implementations to
-further the use of structured prediction methods in computer vision.
-
 The task of semantic segmentation is defined as predicting a semantic class label
@@
 %
-related work?? here again?
-Current systems for semantic segmentation emply a wealth of different paradigms, most notably:
-The ranking of figure ground hypotheses
-conditional random field / superpixel? kraehenbuehl?
-detector based
-sparse coding
+We will in the following distinguish between the task of semantic segmentation,
+which is usually distinguishing unstructured ``stuff'' classes such as road and
+grass, and object class segmentation, which denotes the segmentation of very
+structured classes, such as cars, planes and people.
+We consider four different datasets in this thesis: the object class segmentation
+datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
+and NYU V2.
+%
+Both tasks have the same ultimate goal of parsing, and therefore understanding, images
+in terms of semantic classes, but usually employ  different mechanisms to represent the input.
+%
+%The main part of this thesis investigates the tasks of semantic segmentation and object class segmentation in natural images.
+The main part of this thesis investigates the use of structured learning
+algorithms to the task of semantic image segmentation. Both topics have receive
+much attention in the computer vision and machine learning communities lately.
+Unfortunately, learning structured prediction in computer vision applications
+is still little understood.  We hoe this thesis will further the understanding
+of structured learning methods, give practical guidance and provide
+implementations to further the use of structured prediction methods in computer
+vision.
+
+%related work?? here again?
+%Current systems for semantic segmentation emply a wealth of different paradigms, most notably:
+%The ranking of figure ground hypotheses
+%conditional random field / superpixel? kraehenbuehl?
+%detector based
+%sparse coding
 
@@
 %
-We use four different datasets to evaluate our approach: Graz02, Pascal VOC
-2010, MSRC-21 and NYU V2.
-%
 Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
@@
 \section{List of Contributions}
-Contributions:
 \begin{itemize}
-\item Introduces general clustering algorithm that improves pipeline in serveral places.
-\item Shows that weakly supervised object class segmentation (not semantic segmentation) is possible.
-\item Provides general efficient framework for structured prediction.
-\item Analyze max-margin learning algorithms in different regimes.
-\item shows exact learning is possible in loopy graphs and improves results (sometimes?).
-\item shows importance of edge features.
+\item Introduction of general clustering algorithm that improves upon widely used
+    approaches from the literature.
+\item Demonstration of an algorithm for  weakly supervised object class segmentation.
+\item Providing a general efficient implementation for structured prediction.
+\item Analysis of max-margin learning algorithms with exact or approximate inference in different applications.
+\item Demonstrating that exact learning for semantic segmentation and object class segmentation is possible in loopy graphs.
+%\item Shows importance of edge features.
 \end{itemize}
@@
 We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm},
-which is used as a building block in several places for our semantic
-segmentation algorithms.  We will then introducing a mostly unstructured
-approach to semi-supervised learning for semantic segmentation in
+which we plan to use as an additional building block for semantic segmentation
+algorithms in future work. Clustering is important both for the initial
+segmentation into superpixels, and in the creation of visual words.
+
+We will then introducing a mostly unstructured
+approach to semi-supervised learning for object class segmentation in
 Chaper~\ref{ch:semi_supervised}, investigating the problem of annotating
@@
 semantic image segmentation exactly, even though the involved factor graphs
-contain many loops.
+contain many loops. We demonstrate our algorithm on the Pascal VOC 2010 and MSRC-21 datasets,
+and provide state-of-the-art results on the MSRC-21 dataset.
 
@@
 demonstrate that we are able to learn meaningful spacial relations, and improve
-upon the state of the art in the NYU V2 datasets that we consider.
+upon the state-of-the-art in the NYU V2 datasets that we consider.
 

### 2013-09-18 main.tex
state of the art, slight copy-editing....

@@
 \chapter*{Zusammenfassung}
-Automatische Segementierung und Erkennung von semantischen Klassen in
+\enlargethispage{10mm}
+Automatische Segmentierung und Erkennung von semantischen Klassen in
 nat\"urlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
@@
 der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
-klare Vorteile auf vielen verschiedenen Datens\"atzen hat.
+klare Vorteile auf vielen verschiedenen Datens\"atzen hat. Clustering ist ein wichtiger
+Baustein in vielen Applikationen des Machinellen Sehens, insbesondere in der
+automatischen Segmentierung.
 
@@
 Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
+Mit den vorgestellten Methoden bringen den Stand der Kunst auf
+auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
+von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.
+
 Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
@@
 objective function. We show that the proposed algorithm has clear advantages over standard algorithm
-from the Literature on a wide array of datasets.
+from the Literature on a wide array of datasets. Clustering algorithms are an important building block
+for higher level computer vision applications, in particular for semantic segmentation.
 
 The second part of this work proposes an algorithm for automatic segmentation
-and recognition of object classes in natural images, that learns a segmentation model solely from
-annotation in form of presence and absence of object classes in images. To the best of our knowledge,
-this is the first algorithm of this kind, that explicitly addresses complex object classes.
+and recognition of object classes in natural images, that learns a segmentation
+model solely from annotation in form of presence and absence of object classes
+in images. To the best of our knowledge, this is the first algorithm of this
+kind, that explicitly addresses complex object classes.
 
@@
 segmentation can be learned exactly in practical settings, even in the presence
-of many loops in the underlying neighborhood graphs.  We introduce a software
-library that allows us to perform extensive empirical comparisons of
-state-of-the-art structured learning approaches. This allows us to characterize
-their practical properties in a range of applications, in particular for
-semantic segmentation and object class segmentation.
+of many loops in the underlying neighborhood graphs.
+The introduced methods provide results advancing the state-of-the art on two complex benchmark datasets
+for semantic segmention, the MSRC-21 Dataset of RGB images and the NYU V2 Dataset or RGB-D
+images of indoor scenes.
+
+We also introduce a software library that allows us to perform extensive empirical
+comparisons of state-of-the-art structured learning approaches. This allows us
+to characterize their practical properties in a range of applications, in
+particular for semantic segmentation and object class segmentation.
 
@@
 greatly affected by the continuing exchange with Olivier Grisel, Lars Buitinck,
-Gael Varoquax, Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and
+Ga\"el Varoquax, Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and
 others.
@@
 
-Last but not least I am greateful to Anna M\"uller for her support,
+Last but not least I am grateful to Anna M\"uller for her support,
 encouragement and company. Thank you for giving me the strength needed for this
@@
 
-%\include{itm/paper}
+\include{itm/paper}
 
-%\include{partial_supervision/paper}
+\include{partial_supervision/paper}
 
@@
 on data, learning spacial interactions in a RGB-D setting.
+%TODO STATE OF THE F*ING ART!
 

### 2013-09-18 exact_learning/paper.tex
more fixes by hannes, working on introduction, adding xkcd image

@@
 
-\section{Learning Structure Support Vector Machines with Approximate Inference}
+\section{Learning SSVMs with Approximate Inference}
 In this chapter, we use the 1-slack cutting plane algorithm, as we found this

### 2013-09-18 intro.tex
more fixes by hannes, working on introduction, adding xkcd image

@@
 \chapter{Introduction}
-The task of semantic segmentation is defined as predicting a semantic class label
-for each pixel in a natural image. Predicting semantic labels on such a fine scale
-constitutes an important step towards general scene understanding.
+\vspace{1cm}
+\hfill%
+\begin{minipage}{.4\linewidth}
+\includegraphics[width=\linewidth]{python_soul}%
+\flushright%
+Randall Munroe, \emph{xkcd.com}
+\end{minipage}%
+\\[2cm]
+In computer vision research, the goal is to automatically extract information from an image.
+In particular discerning semantic information, that is interpreting an image, is a prominent
+research topic.
+While much process has been made in recent years, computer vision systems still lack behind
+humans in most tasks that require semantic information. Many semantic tasks can be formulated
+in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
+Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
+cars are in this image'' or even ``What objects lie on the table?''.
+These questions illustrate a range of possible tasks involving semantic categories,
+from classifying images of single objects, over localization and counting of object classes,
+to parsing a scene fully into objects and object classes, together with their relations.
 %
-Applications of scene understanding include automatic image interpretation for retrieval,
-autonomous driving and mobile robotics.
+While humans can distinguish tens of thousands of object classes, and have little trouble
+in interpreting complex scenes, current methods are often restricted to a much smaller number
+of classes. While some recent approaches are able to classify images into a thousand 
+or even ten thousand classes, results are still somewhat mixed.
+In particular, when refining the task to localization, and allowing multiple objects within
+a scene, few approaches have been evaluated on more than about twenty classes.
+
+In this work, we will investigate the tasks of semantic segmentation and object
+class segmentation.  These ask for a semantic annotation of each pixel with one
+of a set of known classes. We choose this task in particular for several reasons:
+\begin{itemize}
+    \item Pixel-level annotation provides highly detailed information about the scene.
+    \item Joint estimation of multiple classes allows for the use of \emph{context}.
+    \item In contrast to category-agnostic segmentation approaches, object class segmentation
+        has an unambiguous true labeling.
+    \item A variety of annotated datasets is publicly available.
+\end{itemize}
 %
+Applications of semantic segmentation and scene understanding include automatic
+image interpretation for retrieval, autonomous driving and mobile robotics.
+While these are already important applications, we expect that with the
+abundance of camera data, and the proliferation of mobile computing, even
+more applications will gain importance soon.
+
 We will in the following distinguish between the task of semantic segmentation,
@@
 in terms of semantic classes, but usually employ  different mechanisms to represent the input.
-%
-%The main part of this thesis investigates the tasks of semantic segmentation and object class segmentation in natural images.
+
+One of the bottlenecks in learning object class segmentation and semantic
+segmentation is the availability of training data.  While unlabeled image data,
+and even data with semantic
+``tags'' is available in practically unlimited quantities, semantic annotation
+on pixel level is only available through laborious manual annotation. 
+Chapter %FIXME
+introduces an approach to cope with this shortcoming of object class segmentation approaches
+by introducing a method to find segmentations automatically from image level annotations.
+
 The main part of this thesis investigates the use of structured learning
@@
 Unfortunately, learning structured prediction in computer vision applications
-is still little understood.  We hoe this thesis will further the understanding
-of structured learning methods, give practical guidance and provide
-implementations to further the use of structured prediction methods in computer
-vision.
+is still little understood.
+
+We will focus on the use of conditional random fields~(CRFs), which have shown
+great promise for computer vision applications. Using the paradigm of
+structural support vector machines, it is possible to learn conditional random
+fields to directly minimize a loss of interest. In particular, conditional
+random fields allow to combine different cues, possibly produced by other
+approaches, in a principled manner.
+
+There have been several previous studies on learning structural support vector machines,
+and learning for conditional random fields. The impact of approximate inference was first
+investigated by \citet{finley2008training}, applying structural support vector machines
+to multi-label data. Later, different works investigated how to combine approximate inference and
+learning in a single framework. \citet{meshi2010learning, komodakis2011efficient}, and \citet{hazan2010primal} approached
+the problem using duality, and formulate learning and inference as a joint optimization problem.
+\citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and  \citet{krahenbuhlparameter} formulate learning
+structured prediction as optimizing an \emph{prediction engine}, that takes into account all aspects
+of the model, in particular the inference algorithm used.
+In this work, on the other hand, we follow the well-established algorithms for learning structural
+support vector machines, and investigate how we can use the available inference algorithms to
+obtain good results within a reasonable time-frame.
+
+\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
+segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
+number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
+\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
+finding little difference in accuracy.
+We focus our work on the more popular neighborhood graphs, which do not allow for easy inference.
+Therefore, conditional maximum likelihood learning would be much harder in our setting. For our studies,
+we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
+this seemingly limited setup, we can make out large differences in the applicability of
+some of the commonly used methods. For comparison, we also evaluate the learning algorithms
+on chain CRFs, where inference is simple.
+
+%We hoe this thesis will further the understanding
+%of structured learning methods, give practical guidance and provide
+%implementations to further the use of structured prediction methods in computer
+%vision.
 
@@
 
-We will focus on the use of conditional random fields, which have shown very
-promising.  Using the paradigm of structural support vector machines, it is
-possible to learn conditional random fields to directly minimize a loss of
-interest. In particular, conditional random fields allow to combine different
-cues, possibly produced by other approaches, in a principled manner.
 %
-Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
-Structured approaches
+%Structured approaches
 %
@@
 \section{List of Contributions}
+This thesis contains the following contributions:
 \begin{itemize}
@@
 approach to semi-supervised learning for object class segmentation in
-Chaper~\ref{ch:semi_supervised}, investigating the problem of annotating
+Chapter~\ref{ch:semi_supervised}, investigating the problem of annotating
 training data for semantic segmentation. The central topic of this thesis is
@@
 Finally, Chapter~\ref{ch:nyu} applies the methods developed in
-Chapter~\ref{ch:structurd_pystruct} and Chapter~\ref{ch:exact_learning} to the
+Chapter~\ref{ch:structured_pystruct} and Chapter~\ref{ch:exact_learning} to the
 problem of semantic annotation with structure classes in RGB-D data. We
@@
 \begin{description}
-    \item[Chapter 2] \emph{Information Theoretic Clustering using Minimum Spanning Trees} Andreas C.
+    \item[Chapter~\ref{ch:itm}] \emph{Information Theoretic Clustering using Minimum Spanning Trees} Andreas C.
 M\"uller, Sebastian Nowozin and Christoph H.
@@
 Published in the proceedings of the German Conference on Pattern Recognition.
-    \item[Chapter 3] \emph{Multi-Instance Methods for Partially Supervised Image Segmentation} Andreas C.
+    \item[Chapter~\ref{ch:semi_supervised}] \emph{Multi-Instance Methods for Partially Supervised Image Segmentation} Andreas C.
 M\"uller and Sven Behnke.
 Published in the proceedings of the IARP Worshop on Partially Supervised Learning.
-    \item[Chapter 4] \emph{PyStruct - Structured Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to the Journal of Machine Learning Research, Open Source Software track.
-    \item[Chapter 5] \emph{Learning a Loopy Model for Semantic Segmentation Exactly}.
+    \item[Chapter~\ref{ch:structured_pystruct}] \emph{PyStruct - Structured Prediction in Python} Andreas C.
+M\"uller and Sven Behnke.
+Submitted to the Journal of Machine Learning Research, Open Source Software track.
+    \item[Chapter~\ref{ch:exact_learning}] \emph{Learning a Loopy Model for Semantic Segmentation Exactly}.
 Andreas C.
@@
 Submitted to the International Conference on Computer Vision Theory and Applications.
-    \item[Chapter 6] \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}.
+    \item[Chapter~\ref{ch:nyu}] \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}.
 Andreas C.

### 2013-09-18 itm/paper.tex
more fixes by hannes, working on introduction, adding xkcd image

@@
-\chapter{Information Theoretic Clustering}
+\chapter{Information Theoretic Clustering}\label{ch:itm}
 
@@
 %
-\section{Clustering Using Nonparametric Entropy-Estimates}
+\section{Clustering Using Non-Parametric Entropy Estimates}
 In general, the goal of clustering can be formulated as follows: 

### 2013-09-18 main.tex
more fixes by hannes, working on introduction, adding xkcd image

@@
 \usepackage{minted}
+\usepackage[german,english]{babel}
 
@@
 
-%\newtheorem{corollary}{Corollary}
-%\newtheorem{corollarya}{Corollary}
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{theorema}{Theorem}
-
 \newcolumntype{R}{>{\raggedleft\arraybackslash}X}
@@
 \vspace*{1cm}
-\textbf{{\huge Learning Structured Prediction For Semantic Image Segmentation}}\\[5mm]
-Dissertation\\%[3mm]
-zur\\%[3mm]
-Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\%[3mm]
-der\\%[3mm]
-Mathematisch-Naturwissenschaftlichen Fakult\"at\\%[3mm]
-der\\%[3mm]
-Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\%[3mm]
-vorgelegt von\\%[5mm]
-{\LARGE Andreas Christian M\"uller}\\%[5mm]
-aus\\%[3mm]
-Offenbach am Main\\%[10mm]
+\textbf{{\huge Learning Structured Prediction for Semantic Image Segmentation}}\\[5mm]
+Dissertation\\
+zur\\
+Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
+der\\
+Mathematisch-Naturwissenschaftlichen Fakult\"at\\
+der\\
+Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
+vorgelegt von\\
+{\LARGE Andreas Christian M\"uller}\\
+aus\\
+Offenbach am Main\\
 Bonn, September 2013
@@
 \enlargethispage{10mm}
+\selectlanguage{german}
 Automatische Segmentierung und Erkennung von semantischen Klassen in
-nat\"urlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
-Diese Arbeit untersucht drei m\"oglichen Ansi\"atzen der Erkennung:
+nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
+Diese Arbeit untersucht drei m\"oglichen Ans\"atzen der Erkennung:
 ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
@@
 klare Vorteile auf vielen verschiedenen Datens\"atzen hat. Clustering ist ein wichtiger
-Baustein in vielen Applikationen des Machinellen Sehens, insbesondere in der
+Baustein in vielen Applikationen des machinellen Sehens, insbesondere in der
 automatischen Segmentierung.
@@
 Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, der
-mit Hilfe von Supervision in Form von Klassen-Pr\"asenz auf Bildern in der Lage
+mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
 ist ein Segmentierungsmodell zu Lernen. Nach unserem Wissen ist dies der erste
@@
 
-Der dritte und letzte Teil er Arbeit untersucht einen der am weitest
+Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
 verbreiteten Ans\"atze zu semantischen Segmentierung und
-Objektklassensegmentierung, n\"amlich Conditional Random Fields verbunden mit Verfahren
+Objektklassensegmentierung, Conditional Random Fields, verbunden mit Verfahren
 der strukturierten Vorhersage.
@@
 Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
-Mit den vorgestellten Methoden bringen den Stand der Kunst auf
+Mit den vorgestellten Methoden bringen den Stand der Kunst
 auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
@@
 
-Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
+Wir stellen au{\ss}erdem eine Software-Bibliothek vor, die es erlaubt einen
 weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
 durchzuf\"uhren.
-Unsere Studie erlaubt uns eine Charakterisation der betrachteten Algorithmen
+Unsere Studie erlaubt uns eine Charakterisierung der betrachteten Algorithmen
 in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
@@
 
+\selectlanguage{english}
 \chapter*{Abstract}

### 2013-09-18 nyu/paper.tex
more fixes by hannes, working on introduction, adding xkcd image

@@
 
-\chapter{Learning Depth-Sensitive Conditional Random Fields}
+\chapter{Learning Depth-Sensitive Conditional Random Fields}\label{ch:nyu}
 
@@
 
-\section{Related work}
+\section{Related Work}
 The task of dense semantic annotation of 3D maps has seen an increased interest
@@
 
-\section{Learning depth-sensitve conditional random fields}\seclabel{learning}
+\section{Learning Depth-Sensitve Conditional Random Fields}\seclabel{learning}
 We take a CRF approach, whose nodes represent a labeling of
@@
 
-\subsection{Results}\seclabel{results}
 \Tabref{results} compares different components of our approach with the literature.
@@
         feature encoding whether one superpixel is above the other in the
-        image. See \secref{results} for details.
+        image. See \secref{experiments} for details.
     \figlabel{weights}}
@@
 
-\section{Summary and discussion}
+\section{Summary and Discussion}
 We introduce a CRF formulation for semantic segmentation

### 2013-09-18 partial_supervision/paper.tex
more fixes by hannes, working on introduction, adding xkcd image

@@
-\chapter{Weakly Supervised Object Segmentation}
+\chapter{Weakly Supervised Object Segmentation}\label{ch:semi_supervised}
 
@@
 \section{Related Work}
-\subsection{Proposal Object Segments}\label{related_segments}
+\subsection{Object Segment Proposals}\label{related_segments}
 Most work on multi-class segmentation focuses on strong supervision on
@@
 \section{Multi-Instance Kernels for Image Segmentation}
-\subsection{Constraint Parametric Min Cuts (CPMC)}
+\subsection{Constraint Parametric Min-Cuts (CPMC)}
 % Repetition?
-To generate proposal segments, we use the Constraint Parametric Min Cut (CPCM)
+To generate proposal segments, we use the Constraint Parametric Min-Cuts (CPCM)
 method from \citet{carreira2010constrained}. In CPMC, initial segments are

### 2013-09-18 structured_prediction/paper.tex
more fixes by hannes, working on introduction, adding xkcd image

@@
 \begin{algorithm*}[t]
-    \caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
+    \caption{$n$-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
     \begin{doublespacing}
@@
 
-\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
+\subsection{The $n$-Slack Cutting Plane Method}\seclabel{n_slack}
 
@@
 
-\subsection{The SZLJSP algorithm}\seclabel{dual_coordinate_descent}
+\subsection{The SZLJSP Algorithm}\seclabel{dual_coordinate_descent}
 \begin{algorithm*}[t]

### 2013-09-18 itm/paper.tex
fix promises in clustering.

@@
 \begin{itemize}
-    \item Clustering, as a purely unsupervised methods is on one end of the spectrum of algorithms we investigate.
-        When using per-image or per-pixel supervision as in the later chapters, we can use clustering to calibrate our
-        expecation of what semisupervised and supervised algorithms should be able to achieve.
+    \item Clustering, as a purely unsupervised methods is on one end of the
+        spectrum of algorithms we investigate.  When using per-image or
+        per-pixel supervision as in the later chapters, we can use clustering
+        to calibrate our expecation of what semisupervised and supervised
+        algorithms should be able to achieve.
+
+    \item Clustering algorithms build the basis of most superpixel algorithms, and
+        better clustering algorithms can lead to better superpixel algorithms.
+
     \item As many other computer vision algorithms, our segmentation methods
-        with critically depend on bag-of-feature representations of segments or
-        images. These are build using a vocabulary of visual words, that is
-        usually created via clustering.
-    \item When discussing latent variable models in chapter~\ref{ch:latent}, we need to initialize our
-        model to overcome the inherant symmetries. Clustering methods are a
-        straight forward solution to this problem.
+        depends on bag-of-feature representations of segments or images. These
+        are build using a vocabulary of visual words that is usually created
+        via clustering.
 \end{itemize}

### 2013-09-18 main.tex
we vs the thesis

@@
 nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
-Diese Arbeit untersucht drei m\"oglichen Ans\"atzen der Erkennung:
+In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atzen der Erkennung:
 ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
@@
 an important open problem in computer vision.
-This work investigates three different approaches to recognition: without supervision,
+In this work, we investigates three different approaches to recognition: without supervision,
 with supervision on level of images, and with supervision on the level of pixels.
@@
 
-The first part introduces a clustering algorithm, that optimizes a novel information-theoretic
-objective function. We show that the proposed algorithm has clear advantages over standard algorithm
-from the Literature on a wide array of datasets. Clustering algorithms are an important building block
-for higher level computer vision applications, in particular for semantic segmentation.
+The first part introduces a clustering algorithm that optimizes a novel
+information-theoretic objective function. We show that the proposed algorithm
+has clear advantages over standard algorithm from the literature on a wide
+array of datasets. Clustering algorithms are an important building block for
+higher level computer vision applications, in particular for semantic
+segmentation.
 
@@
 model solely from annotation in form of presence and absence of object classes
-in images. To the best of our knowledge, this is the first algorithm of this
-kind, that explicitly addresses complex object classes.
+in images. To the best of our knowledge, this is the first algorithm that is
+also able to addresses complex object classes.
 
@@
 of many loops in the underlying neighborhood graphs.
-The introduced methods provide results advancing the state-of-the art on two complex benchmark datasets
-for semantic segmention, the MSRC-21 Dataset of RGB images and the NYU V2 Dataset or RGB-D
-images of indoor scenes.
+The introduced methods provide results advancing the state-of-the art on two
+complex benchmark datasets for semantic segmention, the MSRC-21 Dataset of RGB
+images and the NYU V2 Dataset or RGB-D images of indoor scenes.
 
-We also introduce a software library that allows us to perform extensive empirical
+Finally, we introduce a software library that allows us to perform extensive empirical
 comparisons of state-of-the-art structured learning approaches. This allows us

### 2013-09-19 intro.tex
much lamer quote in the beginning
much lamer quote

@@
 \begin{minipage}{.4\linewidth}
-\includegraphics[width=\linewidth]{python_soul}%
+    %\[ \argmax_{y \in \mathcal{Y}} f(x, y) \]
+    %\dots I wish I could let it stand on its own like this, but of course there is more to come.
+    %\flushright
+    %\emph{Martial Hebert}
+%Remember that all models are wrong;\\
+%the practical question is how wrong do they have to be to not be useful.
+Essentially, all models are wrong, but some are useful. 
 \flushright%
-Randall Munroe, \emph{xkcd.com}
+\emph{George E. P. Box}
+
+%\includegraphics[width=\linewidth]{python_soul}%
+%\flushright%
+%Randall Munroe, \emph{xkcd.com}
 \end{minipage}%
@@
 class segmentation.  These ask for a semantic annotation of each pixel with one
-of a set of known classes. We choose this task in particular for several reasons:
+of a set of known classes.
+
+We choose this task in particular for several reasons:
 \begin{itemize}
@@
 is still little understood.
-
 We will focus on the use of conditional random fields~(CRFs), which have shown
@@
 random fields allow to combine different cues, possibly produced by other
-approaches, in a principled manner.
+approaches, in a principled manner. One of the main difficulties with CRF
+approaches to computer vision problems is, that context in images is usually
+represented as a neighborhood graph.  These graphs, by nature, contain many
+cycles, making inference intractable in general. Consequently, learning algorithms
+have to rely on approximate inference, with often unknown consequences to learning.
 
@@
 on chain CRFs, where inference is simple.
-
-%We hoe this thesis will further the understanding
-%of structured learning methods, give practical guidance and provide
-%implementations to further the use of structured prediction methods in computer
-%vision.
 

### 2013-09-19 intro.tex
say everything twice. comit and delete lol

@@
 While much process has been made in recent years, computer vision systems still lack behind
-humans in most tasks that require semantic information. Many semantic tasks can be formulated
+humans in most tasks that require semantic information. Often these tasks can be formulated
 in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
 Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
-cars are in this image'' or even ``What objects lie on the table?''.
+cars are in this image?'' or even ``What objects lie on the table?''.
 These questions illustrate a range of possible tasks involving semantic categories,
@@
 in interpreting complex scenes, current methods are often restricted to a much smaller number
-of classes. While some recent approaches are able to classify images into a thousand 
-or even ten thousand classes, results are still somewhat mixed.
-In particular, when refining the task to localization, and allowing multiple objects within
-a scene, few approaches have been evaluated on more than about twenty classes.
-
-In this work, we will investigate the tasks of semantic segmentation and object
-class segmentation.  These ask for a semantic annotation of each pixel with one
-of a set of known classes.
+of classes, and only have limited capabilities to model interactions or relations.
+We believe that context is one of the most important cues when it comes to classifying objects,
+and therefore understanding scenes. Consequently, we target dense labeling of scenes on a fine-grained level.
+The task of densely labeling image pixels into classes is called object class segmentation
+or semantic segmentation.
+\pagebreak
 
@@
         has an unambiguous true labeling.
-    \item A variety of annotated datasets is publicly available.
+    \item A variety of manually annotated datasets is publicly available.
 \end{itemize}
@@
 abundance of camera data, and the proliferation of mobile computing, even
-more applications will gain importance soon.
+more applications will emerge soon.
 
@@
 Both tasks have the same ultimate goal of parsing, and therefore understanding, images
-in terms of semantic classes, but usually employ  different mechanisms to represent the input.
+in terms of semantic classes, but usually employ  different mechanisms to represent and process.
+
+Before we delve into semantic segmentation and object class recognition, we
+will first investigate general grouping mechanisms in Chapter~\ref{ch:itm}.
+Clustering is an important step in most semantic segmentation pipelines, in at
+least two places: bottom-up segmentation and creation of dictionaries for
+feature learning. Chapter~\ref{ch:itm} introduces a novel, information theoretic
+algorithm that compares favourably with algorithms from the literature.
+While we do not apply our algorithm to the task of bottom-up segmentation,
+this is a promising avenue for future research.
 
@@
 on pixel level is only available through laborious manual annotation. 
-Chapter %FIXME
-introduces an approach to cope with this shortcoming of object class segmentation approaches
-by introducing a method to find segmentations automatically from image level annotations.
+Chapter~\ref{ch:semi_supervised} introduces an approach to cope with this
+shortcoming of object class segmentation approaches by introducing a method to
+find segmentations automatically from image level annotations.
 
@@
 on chain CRFs, where inference is simple.
-
-%related work?? here again?
-%Current systems for semantic segmentation emply a wealth of different paradigms, most notably:
-%The ranking of figure ground hypotheses
-%conditional random field / superpixel? kraehenbuehl?
-%detector based
-%sparse coding
-
-%
-%Structured approaches
-%
+More recently, \citet{lucchi2013learning} proposed a novel algorithm for
+efficiently learning structured prediction for semantic segmentation, using
+approximate inference. In Chapter~\ref{ch:exact_learning}, we introduce an
+algorithm that runs in similar time to the one proposed by
+\citet{lucchi2013learning}, but is able to learn a CRF to the exact optimum on
+the same dataset.
 

### 2013-09-19 main.tex
typo
use garamond, pick different sanse, leave math as is.
say everything twice. comit and delete lol

@@
 \usepackage[german,english]{babel}
+%\renewcommand*\sfdefault{avant}
+ %\usepackage{helvet}
+\usepackage[defaultsans]{droidsans}
+%\renewcommand*\sfdefault{lmssq}
 
-%\usepackage{tgpagella}
-%\usepackage[T1]{fontenc}
-
-%\usepackage{garamondx}
+\usepackage{garamondx}
 %\usepackage[urw-garamond]{mathdesign}
@@
 
-The third and last part of this work investigates one of the most popular
+The third and main part of this work investigates one of the most popular
 approaches to the task of object class segmentation and semantic segmentation,
 based on conditional random fields and structured prediction.
-%
-%For image labeling tasks, prediction using conditional random fields is usually
-%only approximate, because of the presence of loops of the neighborhood graph in
-%the the image. Consequently, learning in this setting is still little
-%understood, despite the popularity of these methods.
-%
+
 We investigate several learning algorithms, in particular in combination with
@@
 research. My thanks are also extended to Sebastian Nowozin for his collaboration,
-many helpful discussions, and advise.
+many helpful discussions, and advice.
 
@@
 extending the second and third part:
-\paragraph[Large-Scale Weakly Supervised Object Class Segmentation]
+\paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
 google
-\paragraph[Inference Machines]
+\paragraph{Inference Machines}
 Recently \citet{stoyanov2011empirical} started a new trend in structured
@@
 direction.
-\paragraph[Non-Linear Models]
+\paragraph{Non-Linear Models}
 In this work, we only considered models that are linear in the input features--though
@@
 
-\paragraph[Latent Variable Models]
+\paragraph{Latent Variable Models}
 Something latent
-\paragraph[Feature Design]
+\paragraph{Feature Design}
 features are important

### 2013-09-19 nyu/paper.tex
say everything twice. comit and delete lol

@@
 
-\emph{Constant}
+\paragraph{Constant}
     A constant feature allows to model general neighborhood relations.
 
-\emph{Color Contrast}
+\paragraph{Color Contrast}
     We employ a non-linear color contrast, as is common in the computer
@@
 
-\emph{Vertical Alignment}
+\paragraph{Vertical Alignment}
     We model the directed angle between superpixel centers in the 2D image
@@
 
-\emph{Depth Difference}
+\paragraph{Depth Difference}
     We include the signed depth difference between superpixels, which
@@
 
-\emph{Normal Orientations}
+\paragraph{Normal Orientations}
     Differences in normal vector orientation are a strong clue on

### 2013-09-19 evaluation/paper.tex
Polish intro, add blank chapter on evaluation

\chapter{Empirical Comparison of Learning Algorithms}\label{ch:comparison}

### 2013-09-19 intro.tex
Polish intro, add blank chapter on evaluation

@@
 
-Before we delve into semantic segmentation and object class recognition, we
-will first investigate general grouping mechanisms in Chapter~\ref{ch:itm}.
-Clustering is an important step in most semantic segmentation pipelines, in at
-least two places: bottom-up segmentation and creation of dictionaries for
-feature learning. Chapter~\ref{ch:itm} introduces a novel, information theoretic
-algorithm that compares favourably with algorithms from the literature.
-While we do not apply our algorithm to the task of bottom-up segmentation,
-this is a promising avenue for future research.
-
 One of the bottlenecks in learning object class segmentation and semantic
@@
 \begin{itemize}
-\item Introduction of general clustering algorithm that improves upon widely used
+\item Introducing a general clustering algorithm that improves upon widely used
     approaches from the literature.
-\item Demonstration of an algorithm for  weakly supervised object class segmentation.
-\item Providing a general efficient implementation for structured prediction.
-\item Analysis of max-margin learning algorithms with exact or approximate inference in different applications.
-\item Demonstrating that exact learning for semantic segmentation and object class segmentation is possible in loopy graphs.
-%\item Shows importance of edge features.
+\item Demonstrating a scalable algorithm for weakly supervised object class segmentation.
+\item Providing a general and efficient software framework for structured prediction.
+\item Analysing  max-margin learning algorithms with exact or approximate inference in different applications.
+\item Showing that exact learning for semantic segmentation and object class
+    segmentation is possible in practice, even in loopy graphs.
 \end{itemize}
@@
 \section{Thesis Outline}
-We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm},
-which we plan to use as an additional building block for semantic segmentation
-algorithms in future work. Clustering is important both for the initial
-segmentation into superpixels, and in the creation of visual words.
+Before we delve into semantic segmentation and object class recognition, we
+will first investigate general clustering algorithm in Chapter~\ref{ch:itm}.
+Clustering is an important step in most semantic segmentation pipelines, in at
+least two places: bottom-up segmentation and creation of dictionaries for
+feature learning. We introduce a novel, information theoretic algorithm that
+compares favourably with algorithms from the literature.  While we do not apply
+our algorithm to the task of bottom-up segmentation, this is a promising avenue
+for future research.
 
@@
 Chapter~\ref{ch:semi_supervised}, investigating the problem of annotating
-training data for semantic segmentation. The central topic of this thesis is
-introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our
-software library for implementing structured learning and prediction
-algorithms.
-Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely
+training data for semantic segmentation.
+
+The central topic of this thesis is introduced in
+Chapter~\ref{ch:structured_pystruct}, which also introduces our software
+library for implementing structured learning and prediction algorithms.
+
+Chapter~\ref{ch:comparison} gives a quantitative comparison of the most widely
 used structured prediction algorithms, focusing on which algorithm is best to
@@
 demonstrate that we are able to learn meaningful spacial relations, and improve
-upon the state-of-the-art in the NYU V2 datasets that we consider.
+upon the state-of-the-art on the NYU V2 datasets.
 

### 2013-09-19 intro.tex
polish intro

@@
 research topic.
-While much process has been made in recent years, computer vision systems still lack behind
+While much process has been made in recent years, computer vision systems still lag behind
 humans in most tasks that require semantic information. Often these tasks can be formulated
-in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
+in terms of \emph{semantic classes}, meaning categories of parts, objects or scenes.
 Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
@@
 of classes, and only have limited capabilities to model interactions or relations.
-We believe that context is one of the most important cues when it comes to classifying objects,
-and therefore understanding scenes. Consequently, we target dense labeling of scenes on a fine-grained level.
+We believe that \emph{context} is one of the most important cues when it comes to classifying objects,
+and therefore understanding scenes. Consequently, we target dense labeling of scenes, taking
+object relations into account.
 The task of densely labeling image pixels into classes is called object class segmentation
@@
     \item Pixel-level annotation provides highly detailed information about the scene.
-    \item Joint estimation of multiple classes allows for the use of \emph{context}.
+    \item Joint estimation of multiple classes allows for the use of context.
     \item In contrast to category-agnostic segmentation approaches, object class segmentation
@@
 
-We will in the following distinguish between the task of semantic segmentation,
+In the following we distinguish between the task of semantic segmentation,
 which is usually distinguishing unstructured ``stuff'' classes such as road and
@@
 is still little understood.
-We will focus on the use of conditional random fields~(CRFs), which have shown
+We focus on the use of conditional random fields~(CRFs), which have shown
 great promise for computer vision applications. Using the paradigm of
@@
 finding little difference in accuracy.
-We focus our work on the more popular neighborhood graphs, which do not allow for easy inference.
-Therefore, conditional maximum likelihood learning would be much harder in our setting. For our studies,
+We focus our work on the more popular neighborhood graphs, which do not allow for efficient inference.
+Therefore, conditional maximum likelihood learning is intractable in our setting. For our studies,
 we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
@@
 some of the commonly used methods. For comparison, we also evaluate the learning algorithms
-on chain CRFs, where inference is simple.
+on chain CRFs, where efficient inference is possible.
 More recently, \citet{lucchi2013learning} proposed a novel algorithm for
@@
 \begin{itemize}
-\item Introducing a general clustering algorithm that improves upon widely used
+\item Introducing a clustering algorithm that improves upon widely used
     approaches from the literature.
@@
 Before we delve into semantic segmentation and object class recognition, we
-will first investigate general clustering algorithm in Chapter~\ref{ch:itm}.
+first investigate a general clustering algorithm in Chapter~\ref{ch:itm}.
 Clustering is an important step in most semantic segmentation pipelines, in at
@@
 
-We will then introducing a mostly unstructured
-approach to semi-supervised learning for object class segmentation in
-Chapter~\ref{ch:semi_supervised}, investigating the problem of annotating
-training data for semantic segmentation.
+We introducing an algorithm for  semi-supervised learning of object class
+segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the problem of
+annotating training data for semantic segmentation.
 
@@
 Chapter~\ref{ch:comparison} gives a quantitative comparison of the most widely
-used structured prediction algorithms, focusing on which algorithm is best to
-use in a given situation. In particular, we investigate learning behavior for
+used structured prediction algorithms in diverse settings.
+%, focusing on which algorithm is best to %TODO formulation
+%use in a given situation.
+In particular, we investigate learning behavior for
 semantic segmentation on several datasets.
@@
 \begin{description}
-    \item[Chapter~\ref{ch:itm}] \emph{Information Theoretic Clustering using Minimum Spanning Trees} Andreas C.
-M\"uller, Sebastian Nowozin and Christoph H.
-Lampert.
-Published in the proceedings of the German Conference on Pattern Recognition.
-    \item[Chapter~\ref{ch:semi_supervised}] \emph{Multi-Instance Methods for Partially Supervised Image Segmentation} Andreas C.
-M\"uller and Sven Behnke.
-Published in the proceedings of the IARP Worshop on Partially Supervised Learning.
-    \item[Chapter~\ref{ch:structured_pystruct}] \emph{PyStruct - Structured Prediction in Python} Andreas C.
-M\"uller and Sven Behnke.
-Submitted to the Journal of Machine Learning Research, Open Source Software track.
-    \item[Chapter~\ref{ch:exact_learning}] \emph{Learning a Loopy Model for Semantic Segmentation Exactly}.
-Andreas C.
-M\"uller and Sven Behnke.
-Submitted to the International Conference on Computer Vision Theory and Applications.
-    \item[Chapter~\ref{ch:nyu}] \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}.
-Andreas C.
-M\"uller and Sven Behnke.
-Submitted to the International Conference on Robotics and Automation.
+    \item[Chapter~\ref{ch:itm}] \emph{Information Theoretic Clustering using
+        Minimum Spanning Trees} Andreas C. M\"uller, Sebastian Nowozin and
+        Christoph H. Lampert. Published in the proceedings of the German
+        Conference on Pattern Recognition.
+    \item[Chapter~\ref{ch:semi_supervised}] \emph{Multi-Instance Methods for
+        Partially Supervised Image Segmentation} Andreas C. M\"uller and Sven
+        Behnke. Published in the proceedings of the IARP Worshop on Partially
+        Supervised Learning.
+    \item[Chapter~\ref{ch:structured_pystruct}] \emph{PyStruct - Structured
+        Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to
+        the Journal of Machine Learning Research, Open Source Software track.
+    \item[Chapter~\ref{ch:exact_learning}] \emph{Learning a Loopy Model for
+        Semantic Segmentation Exactly}. Andreas C. M\"uller and Sven Behnke.
+        Submitted to the International Conference on Computer Vision Theory and
+        Applications.
+    \item[Chapter~\ref{ch:nyu}] \emph{Learning Depth-Sensitive Conditional
+        Random Fields for Semantic Segmentation}. Andreas C. M\"uller and Sven
+        Behnke. Submitted to the International Conference on Robotics and
+        Automation.
 \end{description}

### 2013-09-19 exact_learning/paper.tex
starting polishing of exact learning chapter

@@
 
+\begin{algorithm*}[t]
+    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack_again}}
+    \begin{algorithmic}[1]
+        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
+        \Ensure parameters $\theta$, slack $\xi$
+        \State $\W \leftarrow \emptyset$
+        \Repeat
+            \State \label{line_restricted}
+            \vspace{-5mm}
+            \begin{align*}
+                (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi\\
+            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
+            &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
+            \end{align*}
+            \For {i=1, \dots, k}
+                \State
+                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane_again}
+            \EndFor
+            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
+            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
+        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
+    \end{algorithmic}
+\end{algorithm*}
+
 \section{Learning SSVMs with Approximate Inference}
 In this chapter, we use the 1-slack cutting plane algorithm, as we found this
-algorithm to be most suitable for out approach. For reference, the algorithm is outlined
-\Secref{one_slac}, Algorithm~\ref{alg_one_slack}. The inference algorithm $I$ is called
-in line~\ref{get_cutting_plane}. We investigate algorithms $I$ (often called separation
+algorithm to be most suitable for out approach. For reference, the algorithm described in
+\Secref{one_slack}, Algorithm~\ref{alg_one_slack_again} is reproduced as
+Algorithm~\ref{alg_one_slack_again}. The inference algorithm $I$ is called in
+line~\ref{get_cutting_plane_again}. We investigate algorithms $I$ (often called
+separation
 oracles in the context) that do not yield the exact maximum here.
@@
 An under-generating approach satisfies $I(x^i, y^i, \theta) \in
-\mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane}
-of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
-hand, does solve the loss-augmented prediction in line~\ref{get_cutting_plane}
+\mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane_again}
+of Algorithm~\ref{alg_one_slack_again}. An over-generating approach on the other
+hand, does solve the loss-augmented prediction in line~\ref{get_cutting_plane_again}
 exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
@@
 Even using approximate inference procedures, several statements
-about the original exact objective (\eqref{mainequation}) can be
+about the original exact objective (\Eqref{mainequation}) can be
 obtained.
 
-Let $o_{\W}(\theta)$ denote objective \eqref{oneslack} with
+Let $o_{\W}(\theta)$ denote objective \Eqref{oneslack} with
 given parameters $\theta$ restricted to a working set $\W$, as computed in
-line~\ref{restricted} of Algorithm~\ref{alg_cutting_plane} and  let
+line~\ref{line_restricted} of Algorithm~\ref{alg_one_slack_again} and  let
 \[
@@
 \subsection{Combining Inference Procedures}\seclabel{combining}
-The cutting plane method described in \Secref{cutting_plane} relies only
+The cutting plane method described in \Secref{one_slack} relies only
 on some separation oracle $I$ that produces violated constraints when
@@
 combination allows us to make fast progress initially and guarantee optimality
-in the end. Our strategy is visualized in \figref{scheme_inference_algs}
+in the end. Our strategy is visualized in \Figref{scheme_inference_algs}
 Notably, it is not necessary for an algorithm used as 3) to always produce
@@
 
-%More precisely, let $o_{\W}(\theta)$ denote objective \eqref{oneslack} with
+%More precisely, let $o_{\W}(\theta)$ denote objective \Eqref{oneslack} with
 %given parameters $\theta$ restricted to a working set $\W$, as computed in
-%line~\ref{restricted} of Algorithm~\ref{alg_cutting_plane} and  let
+%line~\ref{restricted} of Algorithm~\ref{alg_one_slack} and  let
 %\[
@@
 we continue using the caching oracle. Otherwise we run the inference algorithm again.
-For testing \eqref{cache_test}, the last known value of $o^Q$ is used, as recomputing it would defy
+For testing \Eqref{cache_test}, the last known value of $o^Q$ is used, as recomputing it would defy
 the purpose of the cache.

### 2013-09-19 intro.tex
reformatting intro

@@
 %
-While humans can distinguish tens of thousands of object classes, and have little trouble
-in interpreting complex scenes, current methods are often restricted to a much smaller number
-of classes, and only have limited capabilities to model interactions or relations.
-We believe that \emph{context} is one of the most important cues when it comes to classifying objects,
-and therefore understanding scenes. Consequently, we target dense labeling of scenes, taking
-object relations into account.
-The task of densely labeling image pixels into classes is called object class segmentation
-or semantic segmentation.
+While humans can distinguish tens of thousands of object classes, and have
+little trouble in interpreting complex scenes, current methods are often
+restricted to a much smaller number of classes, and only have limited
+capabilities to model interactions or relations.  We believe that
+\emph{context} is one of the most important cues when it comes to classifying
+objects, and therefore understanding scenes. Consequently, we target dense
+labeling of scenes, taking object relations into account.  The task of densely
+labeling image pixels into classes is called object class segmentation or
+semantic segmentation.
 \pagebreak
@@
 grass, and object class segmentation, which denotes the segmentation of very
-structured classes, such as cars, planes and people.
-We consider four different datasets in this thesis: the object class segmentation
-datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
-and NYU V2.
+structured classes, such as cars, planes and people. We consider four
+different datasets in this thesis: the object class segmentation datasets
+Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21, and
+NYU V2.
 %
-Both tasks have the same ultimate goal of parsing, and therefore understanding, images
-in terms of semantic classes, but usually employ  different mechanisms to represent and process.
+Both tasks have the same ultimate goal of parsing, and therefore understanding,
+images in terms of semantic classes, but usually employ  different mechanisms
+to represent and process.
 
@@
 is still little understood.
-We focus on the use of conditional random fields~(CRFs), which have shown
-great promise for computer vision applications. Using the paradigm of
-structural support vector machines, it is possible to learn conditional random
-fields to directly minimize a loss of interest. In particular, conditional
-random fields allow to combine different cues, possibly produced by other
-approaches, in a principled manner. One of the main difficulties with CRF
-approaches to computer vision problems is, that context in images is usually
-represented as a neighborhood graph.  These graphs, by nature, contain many
-cycles, making inference intractable in general. Consequently, learning algorithms
-have to rely on approximate inference, with often unknown consequences to learning.
+We focus on the use of conditional random fields~(CRFs), which have shown great
+promise for computer vision applications. Using the paradigm of structural
+support vector machines, it is possible to learn conditional random fields to
+directly minimize a loss of interest. In particular, conditional random fields
+allow to combine different cues, possibly produced by other approaches, in a
+principled manner. One of the main difficulties with CRF approaches to computer
+vision problems is, that context in images is usually represented as a
+neighborhood graph.  These graphs, by nature, contain many cycles, making
+inference intractable in general. Consequently, learning algorithms have to
+rely on approximate inference, with often unknown consequences to learning.
 
-There have been several previous studies on learning structural support vector machines,
-and learning for conditional random fields. The impact of approximate inference was first
-investigated by \citet{finley2008training}, applying structural support vector machines
-to multi-label data. Later, different works investigated how to combine approximate inference and
-learning in a single framework. \citet{meshi2010learning, komodakis2011efficient}, and \citet{hazan2010primal} approached
-the problem using duality, and formulate learning and inference as a joint optimization problem.
-\citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and  \citet{krahenbuhlparameter} formulate learning
-structured prediction as optimizing an \emph{prediction engine}, that takes into account all aspects
-of the model, in particular the inference algorithm used.
-In this work, on the other hand, we follow the well-established algorithms for learning structural
-support vector machines, and investigate how we can use the available inference algorithms to
-obtain good results within a reasonable time-frame.
+There have been several previous studies on learning structural support vector
+machines, and learning for conditional random fields. The impact of approximate
+inference was first investigated by \citet{finley2008training}, applying
+structural support vector machines to multi-label data. Later, different works
+investigated how to combine approximate inference and learning in a single
+framework. \citet{meshi2010learning, komodakis2011efficient}, and
+\citet{hazan2010primal} approached the problem using duality, and formulate
+learning and inference as a joint optimization problem.
+\citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and
+\citet{krahenbuhlparameter} formulate learning structured prediction as
+optimizing an \emph{prediction engine}, that takes into account all aspects of
+the model, in particular the inference algorithm used.
+In this work, on the other hand, we follow the well-established algorithms for
+learning structural support vector machines, and investigate how we can use the
+available inference algorithms to obtain good results within a reasonable
+time-frame.
 
-\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
-segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
-number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
-\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
-finding little difference in accuracy.
-We focus our work on the more popular neighborhood graphs, which do not allow for efficient inference.
-Therefore, conditional maximum likelihood learning is intractable in our setting. For our studies,
-we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
-this seemingly limited setup, we can make out large differences in the applicability of
-some of the commonly used methods. For comparison, we also evaluate the learning algorithms
+\citet{nowozin2010parameter} provide a detailed evaluation of different aspects
+of learning object class segmentation, that is somewhat orthogonal to this
+work. Their work considers the choice of features, number superpixels and
+pairwise potentials for conditional maximum likelihood learning of
+tree-structured CRFs.  \citet{nowozin2010parameter} also compare conditional
+maximum likelihood learning with maximum margin learning, finding little
+difference in accuracy.
+We focus our work on the more popular neighborhood graphs, which do not allow
+for efficient inference.  Therefore, conditional maximum likelihood learning is
+intractable in our setting. For our studies, we fix the considered feature set,
+and vary the maximum margin learning algorithm. Despite this seemingly limited
+setup, we can make out large differences in the applicability of some of the
+commonly used methods. For comparison, we also evaluate the learning algorithms
 on chain CRFs, where efficient inference is possible.
@@
 \item Demonstrating a scalable algorithm for weakly supervised object class segmentation.
-\item Providing a general and efficient software framework for structured prediction.
+\item Providing a general and efficient open source software framework for structured prediction.
 \item Analysing max-margin learning algorithms with exact or approximate inference in different applications.

### 2013-09-19 main.tex
Polish intro, add blank chapter on evaluation
new title, polish conclusion

@@
 \vspace*{1cm}
-\textbf{{\huge Learning Structured Prediction for Semantic Image Segmentation}}\\[5mm]
+\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
 Dissertation\\
@@
 Automatic segmentation and recognition of semantic classes in natural images is
-an important open problem in computer vision.
-In this work, we investigates three different approaches to recognition: without supervision,
-with supervision on level of images, and with supervision on the level of pixels.
+an important open problem in computer vision. In this work, we investigates
+three different approaches to recognition: without supervision, with
+supervision on level of images, and with supervision on the level of pixels.
 The thesis comprises three parts.
@@
 array of datasets. Clustering algorithms are an important building block for
-higher level computer vision applications, in particular for semantic
+higher-level computer vision applications, in particular for semantic
 segmentation.
@@
 model solely from annotation in form of presence and absence of object classes
-in images. To the best of our knowledge, this is the first algorithm that is
-also able to addresses complex object classes.
+in images.
 
@@
 based on conditional random fields and structured prediction.
-
+%
 We investigate several learning algorithms, in particular in combination with
@@
 The introduced methods provide results advancing the state-of-the art on two
-complex benchmark datasets for semantic segmention, the MSRC-21 Dataset of RGB
+complex benchmark datasets for semantic segmentation, the MSRC-21 Dataset of RGB
 images and the NYU V2 Dataset or RGB-D images of indoor scenes.
-
+%
 Finally, we introduce a software library that allows us to perform extensive empirical
@@
 
+\include{evaluation/paper}
+
 \include{exact_learning/paper}
@@
 \chapter{Conclusion and Future Direction}
-Current research in computer vision progresses at an incredible speed, with new
-methods and applications emerging constantly.  Still, there are large
-differences in the paradigms employed by different researchers.  While it is
-generally acknowledged that context is one of the most important cues in
-recognizing objects and interpreting scenes, there is little consensus on how
-contextual information should be modelled, or even learned.
-The structured prediction paradigm allows a principled approach to
-incorporating predictions within an image, and to include context information.
-We investigated the potential of structural methods for the application of
-semantic image segmentation and object recognition, applications to which
-structured methods have been applied successfully before.  We focus on
-\emph{learning} of structural models and the interaction of inference and
-learning in the neighborhood models typically employed for semantic
-segmentation.
-The thesis presents a software implementation of a variety of popular learning
-algorithms for structural support vector machines, together with a thorough
-evaluation of their properties in the presence of approximate inference.
-We demonstrate that exact learning is possible, even in the presence of loops in the
-underlying factor graph. We also demonstrate the power of conditioning interactions
-on data, learning spacial interactions in a RGB-D setting.
-%TODO STATE OF THE F*ING ART!
+\section{Summary}
+This thesis explored the use of structured prediction methods for semantic
+segmentation and object class segmentation of natural images, as an important
+step towards general scene understanding.
+We use the paradigm of structured prediction, which allows for a principled
+integration of context and object relations.
+%Current research in computer vision progresses at an incredible speed, with new
+%methods and applications emerging constantly. Still, there are large
+%differences in the paradigms employed by different researchers. While it is
+%generally acknowledged that context is one of the most important cues in
+%recognizing objects and interpreting scenes, there is little consensus on how
+%contextual information should be modelled, or even learned.
+%The structured prediction paradigm allows a principled approach to
+%incorporating predictions within an image, and to include context information.
+%We investigated the potential of structural methods for the application of
+%semantic image segmentation and object recognition, applications to which
+%structured methods have been applied successfully before.
+We focus on \emph{learning} of structural models and the interaction of
+inference and learning in the neighborhood models typically employed for
+semantic segmentation.
+We presents an open source software implementation of a variety of
+popular learning algorithms for structural support vector machines, together
+with a thorough evaluation of their properties in the presence of approximate
+inference. Our software provides a solid basis for future research into learning,
+inference and models for computer vision, by providing extensive examples and
+benchmarks.
 
+We demonstrate that effective use of available inference mechanisms enables
+exact learning, even in the presence of loops in the underlying
+factor graph, leading to state-of-the art results on the MSRC-21 dataset.
+We also demonstrate the power of conditional interactions, learning spacial
+interactions in a RGB-D setting. Our approach improves upon the state-of-the
+art on the NYU V2 benchmark for annotation of semantic structure classes.
+
+We also presented a novel approach for clustering based on information theoretic
+principles. Our algorithm improves upon methods from the literature on a wide
+range of datasets. This indicates that also in the task of extracting
+superpixels, we can hope to achieve better results than the $k$-means based
+SLIC algorithm that we used.
+
+As manual annotation of images for learning semantic segmentation and object
+class recognition is laborious and error-prone, we suggested a method to
+learn object class segmentation for complex object classes from image-level
+annotations alone.
+
+\section{Future Directions}
 There are several directions for future research that we think would be
-interesting to pursue as an extension of the presented results, in particular in
-extending the second and third part:
+interesting to pursue as an extension of the presented results:
 \paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
-google
+We demonstrated a new method for object class segmentation using only weak
+supervision. One of the main advantages of such a method is that it is
+potentially able to exploit the large amount of weakly labeled data that
+is available on the internet. Using additional training data, and evaluating on
+the given, manually annotated data, would be an interesting topic for future
+research.
 \paragraph{Inference Machines}
@@
 In this work, we only considered models that are linear in the input features--though
-features are highly non-linear in the original input pixels.
+features are highly non-linear in the original input pixels. Allowing non-linear
+interactions increases the representational power of a CRF, possibly leading
+to better segmentation results.
 Kernelization of structural support vector machines is trivial in theory, but
@@
 interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
-graphs for image processing, but not for higher level tasks such as semantic
+graphs for image processing, but not for higher-level tasks such as semantic
 segmentation. If it is possible to include context in a meaningful way, it
-might be possible to solve even object-centric tasks such as object class
+might be possible to address even object-centric tasks such as object class
 segmentation with DTFs.
 
-\paragraph{Latent Variable Models}
-Something latent
+\paragraph{Higher Order Potentials and Latent Variable Models}
+While non-linear potentials would allow for more complex interactions between
+inputs and labelings, introducing higher order potentials or latent variables
+allow the model to express more complex interactions within the output
+variables. Possible examples are consistency of larger regions, learning parts
+or learning of scene classes and cooccurcences.
+In principle, higher order potentials and latent variable models are
+equivalent, in that each energy function expressed in either form can be
+transformed into an energy function of the other kind. In practice, learning of
+higher order potentials for semantic segmentation has received little
+attention, while most approaches using latent variables are often limited by
+the non-convexity of learning.  It would be interesting to compare current
+methods using latent variable and higher order approaches, and see how these
+interact with different inference and learning schemes.
+
 \paragraph{Feature Design}
-features are important
-which low-level cues?
-superpixels?
+This work mostly focused on learning methods, and less on the input, with
+the exception of Chapter~\ref{ch:nyu}, that explores the use of 3D features
+for semantic segmentation of indoor scenes. It is clear, however, that
+the input feature play an important role in the performance of any system.
+Using our approach for exact learning of loopy graphs, it seems to be worthwhile to
+revisit the works of \citet{nowozin2010parameter} and \citet{lucchi2011spatial},
+that evaluate the impact of input features and piecewise training, and of the importance
+of global constraints versus global features, respectively.
+In particular the importance of features for pairwise potentials has been
+somewhat overlooked, often being reduced to a single constant or contrast
+sensitive feature.
 

### 2013-09-19 exact_learning/paper.tex
typographic fine-tuning :)

@@
-\chapter{Efficient Exact Learning of Loopy CRFs}\label{ch:exact_learning}
+\chapter{Learning of Loopy CRF Exactly}\label{ch:exact_learning}
 
@@
 In this chapter, we investigate the necessity and consequences of these
-approximations in the one-slack cutting-plane learning algorithm in the context of semantic
+approximations in the $1$-slack cutting-plane learning algorithm in the context of semantic
 image segmentation. We show that despite inference being deemed intractable in
@@
 
-%Currently the most widely used method is the one-slack cutting plane formulation
+%Currently the most widely used method is the $1$-slack cutting plane formulation
 %introduced by \citet{joachims2009cutting}.
@@
 solutions.
-In contrast, we aim at optimizing the non-relaxed objective directly, minimizing
+We argue that extending the target domain in this way is unnatural for many applications,
+and aim at optimizing the non-relaxed objective directly, minimizing
 the original empirical risk. This is an important difference, as relaxed solutions
@@
     \begin{algorithmic}[1]
-        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
+        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$, inference oracle $\hat{I}$.
         \Ensure parameters $\theta$, slack $\xi$
@@
                 \State
-                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane_again}
+                $\hat{y}\hoch{i} \leftarrow \hat{I}(x\hoch{i}, y\hoch{i}, \theta) \approx \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane_again}
             \EndFor
@@
 \section{Learning SSVMs with Approximate Inference}
-In this chapter, we use the 1-slack cutting plane algorithm, as we found this
-algorithm to be most suitable for out approach. For reference, the algorithm described in
-\Secref{one_slack}, Algorithm~\ref{alg_one_slack_again} is reproduced as
-Algorithm~\ref{alg_one_slack_again}. The inference algorithm $I$ is called in
-line~\ref{get_cutting_plane_again}. We investigate algorithms $I$ (often called
+In this chapter, we use the $1$-slack cutting plane algorithm, as we found this
+algorithm to be most suitable for out approach. Recall the optimization problem
+that is solved by the $1$-slack cutting plane algorithm:
+\begin{align}\eqlabel{oneslack_again}
+    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
+    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
+        &\theta^T \sum_{i=1}^n [\Phi(x^i, y^i) - \Phi(x^i,
+            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
+            - \xi
+\end{align}
+For reference, the algorithm described in
+\Secref{one_slack}, Algorithm~\ref{alg_one_slack} is reproduced as
+Algorithm~\ref{alg_one_slack_again}, with the additional input of an inference algorithm $\hat{I}$, which is called in
+line~\ref{get_cutting_plane_again}. We investigate algorithms $\hat{I}$ (often called
 separation
@@
 \citet{finley2008training}: under-generating and over-generating approaches.
-An under-generating approach satisfies $I(x^i, y^i, \theta) \in
+An under-generating approach satisfies $\hat{I}(x^i, y^i, \theta) \in
 \mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane_again}
@@
 exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
-that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.
+that possibly $\hat{I}(x^i, y^i, \theta) \notin \mathcal{Y}$.
 %We elaborate on
@@
 Even using approximate inference procedures, several statements
-about the original exact objective (\Eqref{mainequation}) can be
+about the original exact objective (\Eqref{oneslack_again}) can be
 obtained.
 
-Let $o_{\W}(\theta)$ denote objective \Eqref{oneslack} with
+Let $o_{\W}(\theta)$ denote objective \Eqref{oneslack_again} with
 given parameters $\theta$ restricted to a working set $\W$, as computed in
@@
 \[
-    o^I(\theta) = C\xi' + \frac{||\theta||}{2}^2
+    o^{\hat{I}}(\theta) = C\xi' + \frac{||\theta||}{2}^2
 \]
-when using inference algorithm $I$, that is $o^I(\theta)$ is the approximation of the primal
-objective given by $I$. To simplify exposure, we drop the dependency on $\theta$.
+when using inference algorithm $\hat{I}$, that is $o^{\hat{I}}(\theta)$ is the approximation of the primal
+objective given by $\hat{I}$. To simplify exposure, we drop the dependency on $\theta$.
 
-Depending on the properties of the inference procedure $I$ used, it is easy to see:
+Depending on the properties of the inference procedure $\hat{I}$ used, it is easy to see:
 \begin{enumerate}
@@
 
-    \item If $I$ is an over-generating or exact algorithm, $o^I$ is an upper
+    \item If $\hat{I}$ is an over-generating or exact algorithm, $o^{\hat{I}}$ is an upper
         bound on $o(\theta^*)$.
@@
 The cutting plane method described in \Secref{one_slack} relies only
-on some separation oracle $I$ that produces violated constraints when
+on some separation oracle $\hat{I}$ that produces violated constraints when
 performing loss-augmented prediction.
 
-Using any under-generating oracle $I$, learning can proceed as long as a
+Using any under-generating oracle $\hat{I}$, learning can proceed as long as a
 constraint is found that is violated by more than the stopping tolerance
@@
 In the following, we use the notation of \Secref{bounds} and indicate
-the choices of oracle $I$ with $Q$ for a chosen inference algorithm and $C$ for
+the choices of oracle $\hat{I}$ with $Q$ for a chosen inference algorithm and $C$ for
 using cached constraints.
@@
     Dual Objective $o_\W$ & 65.10 & 67.66  \\
-    Estimated Objective $o^I$ &  67.62& 67.66\\
+    Estimated Objective $o^{\hat{I}}$ &  67.62& 67.66\\
     True Primal Objective $o^E$& 69.92& 67.66\\
@@
 remains when using only under-generating inference.
-Additionally, the estimated primal objective $o^I$ using under-generating inference is
+Additionally, the estimated primal objective $o^{\hat{I}}$ using under-generating inference is
 too optimistic, as can be expected. This underlines the fact that
@@
     Dual Objective $o_\W$ &92.06& 92.24\\
-    Estimated Objective $o^I$ & 92.07 &92.24\\
+    Estimated Objective $o^{\hat{I}}$ & 92.07 &92.24\\
     True Primal Objective $o^E$&92.35& 92.24  \\
@@
 We implemented the described procedure in \pystruct.
-
 We used the SLIC implementation provided by \citet{achanta2012slic} to extract superpixels and
@@
 learning is the combination of different inference mechanisms and a novel
-caching scheme for the one-slack cutting plane method, in combination with
+caching scheme for the $1$-slack cutting plane method, in combination with
 state-of-the-art inference methods.
@@
 optimization.
-
+\enlargethispage{10mm}
 We do not expect learning of pixel-level models, which typically have tens or

### 2013-09-19 intro.tex
typographic fine-tuning :)

@@
 We choose this task in particular for several reasons:
-\begin{itemize}
+\begin{itemize*}
     \item Pixel-level annotation provides highly detailed information about the scene.
@@
     \item A variety of manually annotated datasets is publicly available.
-\end{itemize}
+\end{itemize*}
 %
@@
 
-The main part of this thesis investigates the use of structured learning
+The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
 algorithms to the task of semantic image segmentation. Both topics have receive
-much attention in the computer vision and machine learning communities lately.
+much attention in the computer vision and machine learning communities
+lately~\citep{ladicky2009associative, krahenbuhl2012efficient,
+branson2013efficient, blake2011markov}.
 Unfortunately, learning structured prediction in computer vision applications

### 2013-09-19 itm/paper.tex
typographic fine-tuning :)

@@
 
+\enlargethispage{5mm}
 A comparison of ITM, MeanNN and the baseline methods, $k$-Means and 

### 2013-09-19 main.tex
typographic fine-tuning :)

@@
 \section{Summary}
-This thesis explored the use of structured prediction methods for semantic
-segmentation and object class segmentation of natural images, as an important
+In this thesis, we explored the use of structured prediction methods for semantic
+segmentation and object class segmentation of natural images, an important
 step towards general scene understanding.
@@
 %structured methods have been applied successfully before.
-We focus on \emph{learning} of structural models and the interaction of
+We focused on \emph{learning} of structural models and the interaction of
 inference and learning in the neighborhood models typically employed for
 semantic segmentation.
-We presents an open source software implementation of a variety of
+We presented an open source software implementation of a variety of
 popular learning algorithms for structural support vector machines, together
@@
 
-We demonstrate that effective use of available inference mechanisms enables
+We showed that effective use of available inference mechanisms enables
 exact learning, even in the presence of loops in the underlying
-factor graph, leading to state-of-the art results on the MSRC-21 dataset.
-We also demonstrate the power of conditional interactions, learning spacial
-interactions in a RGB-D setting. Our approach improves upon the state-of-the
+factor graph. Our methods achieve competitive performance with similar
+ methods on the Pascal VOC 2010 dataset, and yield state-of-the art
+ results on the MSRC-21 dataset.
+We demonstrated the power of conditional interactions by learning spacial
+interactions in an RGB-D setting. Here, our approach improves upon the state-of-the
 art on the NYU V2 benchmark for annotation of semantic structure classes.
@@
 principles. Our algorithm improves upon methods from the literature on a wide
-range of datasets. This indicates that also in the task of extracting
-superpixels, we can hope to achieve better results than the $k$-means based
+range of datasets. This indicates that in the task of extracting
+superpixels, we can also hope to achieve better results than the $k$-means based
 SLIC algorithm that we used.
@@
 potentially able to exploit the large amount of weakly labeled data that
-is available on the internet. Using additional training data, and evaluating on
-the given, manually annotated data, would be an interesting topic for future
-research.
+is available on the internet. Using additional, weakly labeled training data,
+and evaluating on the given, manually annotated data, is therefore a promising
+path for improving the presented results.
 \paragraph{Inference Machines}
@@
 prediction, which is sometimes called ``inference machines''. The basic
-principle is simple: view the process of prediction with a given inference
-procedure as a prediction method and optimize parameters of this method using
-empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
-loopy belief propagation as their inference algorithm and the optimization is
-carried out simply using gradient descent on the non-convex but differentiable
-loss function.  Other recent work in this direction includes
-\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
-conditional random field and \citet{jancsarylearning}, who used closed form
-inference in a Gaussian CRF\@.
+principle is simple: the process of prediction \emph{using a given inference
+procedure} is viewed as a feed-forward method for prediction, and parameters of this
+prediction process are optimized directly using empirical risk minimization. The work of
+\citet{stoyanov2011empirical}, used loopy belief propagation as their inference
+algorithm and the optimization is carried out simply using gradient descent on
+the non-convex but differentiable loss function.  Other recent work in this
+direction includes \citet{krahenbuhlparameter}, who used mean-field inference
+in a fully connected conditional random field and \citet{jancsarylearning}, who
+used closed form inference in a Gaussian CRF\@.
 While these algorithms show great promise, their relation to the traditional
 approach of structured prediction used in this work is mostly unclear. In particular,
-if exact traditional learning is possible in a model, it is unclear how much
-can be gained by direct empirical risk minimization.
-Little empirical comparison was performed, and we are not aware of theoretical
-work in this direction, leaving much room for future investigation of this
-direction.
+if exact traditional learning is possible in a model, it is uncertain how much
+in accuracy and efficiency can be gained by direct empirical risk minimization.
+Only limited empirical comparison is available, and we are not aware of theoretical
+work in this direction, leaving much room for future investigation.
 \paragraph{Non-Linear Models}
-In this work, we only considered models that are linear in the input features--though
+In this work, we only considered models that are linear in the input features---though
 features are highly non-linear in the original input pixels. Allowing non-linear
 interactions increases the representational power of a CRF, possibly leading
-to better segmentation results.
-Kernelization of structural support vector machines is trivial in theory, but
+to more accurate prediction results.
+Kernelization of structural support vector machines is straight-forward in theory, but
 had only limited success in the context of CRFs for image
@@
 and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
-Conditional neural fields have only applied to sequence classification so far,
+Conditional neural fields have only been applied to sequence classification so far,
 and extending them to our setting of semantic image segmentation would be very
@@
 allow the model to express more complex interactions within the output
-variables. Possible examples are consistency of larger regions, learning parts
-or learning of scene classes and cooccurcences.
+variables~\citep{dann2012pottics, kohli2009robust}. Possible examples are
+consistency of larger regions, learning parts or learning of scene classes and
+cooccurences.
 In principle, higher order potentials and latent variable models are
@@
 higher order potentials for semantic segmentation has received little
-attention, while most approaches using latent variables are often limited by
+attention, while approaches using latent variables are often limited by
 the non-convexity of learning.  It would be interesting to compare current
@@
 \paragraph{Feature Design}
-This work mostly focused on learning methods, and less on the input, with
-the exception of Chapter~\ref{ch:nyu}, that explores the use of 3D features
+This work mostly focused on learning methods, and less on the input---with
+the exception of Chapter~\ref{ch:nyu}, which explores the use of 3D features
 for semantic segmentation of indoor scenes. It is clear, however, that
-the input feature play an important role in the performance of any system.
+the input features play an important role in the performance of any system.
 Using our approach for exact learning of loopy graphs, it seems to be worthwhile to
@@
 In particular the importance of features for pairwise potentials has been
-somewhat overlooked, often being reduced to a single constant or contrast
-sensitive feature.
+somewhat overlooked in the computer vision literature, often being reduced to a
+single constant or contrast sensitive feature.
 

### 2013-09-19 partial_supervision/paper.tex
typographic fine-tuning :)

@@
 Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we
-assume that bags are simply sets, not multi-sets.  Then a bag is an element of
+assume that bags are simply sets, not multi-sets.
+\pagebreak\\
+Then a bag is an element of
 the power set $2^\mathcal{X}$ and the task is to learn a function
@@
         & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-I \\
-    \cmidrule(rl){2-8}
+    \cmidrule{2-8}
     Musk1 & 87.9 &84.9 &  87.4 &  77.9     & 84.3 & 88.0& 88.0 \\
@@
                 &accuracy & witness-rate & accuracy & witness-rate  \\
-    \cmidrule(rl){2-3}
-    \cmidrule(rl){4-5}
+    \cmidrule{2-3}
+    \cmidrule{4-5}
     mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
@@
                 & car & bike & person \\
-    \cmidrule(l){2-2}
-    \cmidrule(l){3-3}
-    \cmidrule(l){4-4}
+    \cmidrule{2-4}
         This work&   0.30&  0.45&  0.43 \\

### 2013-09-19 structured_prediction/paper.tex
typographic fine-tuning :)

@@
 different parameter vectors $\theta_{F, y_{N_F}}$ for each possible assignment
-of $y_{N_F} \in \mathcal{Y}_{N_F}$.
-
+of $y_{N_F} \in \mathcal{Y}_{N_F}$.\pagebreak\\
+%
 Both, \Eqref{general_psi} and \Eqref{linear_psi} are instantiations of the general
@@
 \end{equation}
-with regularization parameter $C$. Here, $\ell$ is a hinge-loss like upper bound
+with regularization parameter $C$.\pagebreak\\
+Here, $\ell$ is a hinge-loss like upper bound
 on the empirical $\Delta$-risk:
@@
 loss-augmented prediction problem in \Eqref{loss_augmentation}, it is natural
-to work in a stochastic setting (see \citet{ratliff2007online}).
+to work in a stochastic setting (see \citet{ratliff2007online}).\pagebreak\\
 Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
@@
 into a quadratic objective with a combinatorial number of constraints:
-\begin{align}\label{eq:oneslack}
+\begin{align}\label{eq:nslack}
     \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
@@
 \end{align}
+\enlargethispage{10mm}
 Informally, the $1$-slack formulation corresponds to joining all training
@@
 interactions between variables corresponding to different data points, and
-then applying the $n$-slack algorithm with a single data point.
+then applying the $n$-slack algorithm with a single data point.\pagebreak\\
 A detailed description can be found in Algorithm~\ref{alg_one_slack}.
@@
 convergence.
-
+\enlargethispage{10mm}
 The detailed procedure as given by \citet{lacoste2012block} is shown in Algorithm~\ref{alg_szljsp}.

### 2013-09-19 intro.tex
typos julia

@@
 research topic.
-While much process has been made in recent years, computer vision systems still lag behind
+While much progress has been made in recent years, computer vision systems still lag behind
 humans in most tasks that require semantic information. Often these tasks can be formulated
@@
 Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
-cars are in this image?'' or even ``What objects lie on the table?''.
+cars are there in this image?'' or even ``What objects lie on the table?''.
 These questions illustrate a range of possible tasks involving semantic categories,
@@
 The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
-algorithms to the task of semantic image segmentation. Both topics have receive
+algorithms to the task of semantic image segmentation. Both topics have received
 much attention in the computer vision and machine learning communities
@@
 principled manner. One of the main difficulties with CRF approaches to computer
-vision problems is, that context in images is usually represented as a
+vision problems is that context in images is usually represented as a
 neighborhood graph.  These graphs, by nature, contain many cycles, making
@@
 \citet{krahenbuhlparameter} formulate learning structured prediction as
-optimizing an \emph{prediction engine}, that takes into account all aspects of
+optimizing a \emph{prediction engine}, that takes into account all aspects of
 the model, in particular the inference algorithm used.
@@
 least two places: bottom-up segmentation and creation of dictionaries for
-feature learning. We introduce a novel, information theoretic algorithm that
+feature learning. We introduce a novel information theoretic algorithm that
 compares favourably with algorithms from the literature.  While we do not apply
@@
 
-We introducing an algorithm for  semi-supervised learning of object class
+We introduce an algorithm for  semi-supervised learning of object class
 segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the problem of

### 2013-09-19 main.tex
typos julia

@@
 nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
-In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atzen der Erkennung:
+In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atze der Erkennung:
 ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
-von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.
+von Pixeln.
 
+Diese Arbeit setzt sich aus drei Teilen zusammen.
 Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
@@
 Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
-Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, der
+Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, das
 mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
-ist ein Segmentierungsmodell zu Lernen. Nach unserem Wissen ist dies der erste
+ist ein Segmentierungsmodell zu lernen. Nach unserem Wissen ist dies der erste
 Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.
@@
 Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
-verbreiteten Ans\"atze zu semantischen Segmentierung und
+verbreiteten Ans\"atze zur semantischen Segmentierung und
 Objektklassensegmentierung, Conditional Random Fields, verbunden mit Verfahren
@@
 %
-Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
+Wir untersuchen verschiedene Lernalgorithmen des strukturierten Lernens, insbesondere
 im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
@@
 
+\enlargethispage{15mm}
 Wir stellen au{\ss}erdem eine Software-Bibliothek vor, die es erlaubt einen
@@
 Automatic segmentation and recognition of semantic classes in natural images is
-an important open problem in computer vision. In this work, we investigates
+an important open problem in computer vision. In this work, we investigate
 three different approaches to recognition: without supervision, with
@@
 information-theoretic objective function. We show that the proposed algorithm
-has clear advantages over standard algorithm from the literature on a wide
+has clear advantages over standard algorithms from the literature on a wide
 array of datasets. Clustering algorithms are an important building block for

### 2013-09-19 main.tex
typo

@@
 Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
-Mit den vorgestellten Methoden bringen den Stand der Kunst
+Mit den vorgestellten Methoden bringen wir den Stand der Kunst
 auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz

### 2013-09-19 intro.tex
Joe's corrections :)

@@
 of learning object class segmentation, that is somewhat orthogonal to this
-work. Their work considers the choice of features, number superpixels and
+work. Their work considers the choice of features, number of superpixels and
 pairwise potentials for conditional maximum likelihood learning of

### 2013-09-19 itm/paper.tex
Joe's corrections :)

@@
         per-pixel supervision as in the later chapters, we can use clustering
-        to calibrate our expecation of what semisupervised and supervised
+        to calibrate our expectation of what semi-supervised and supervised
         algorithms should be able to achieve.
@@
         depends on bag-of-feature representations of segments or images. These
-        are build using a vocabulary of visual words that is usually created
+        are built using a vocabulary of visual words that is usually created
         via clustering.

### 2013-09-19 nyu/paper.tex
new title, polish conclusion
typographic fine-tuning :)
spelling!

@@
 
-\section{Learning Depth-Sensitve Conditional Random Fields}\seclabel{learning}
+\section{Learning Depth-Sensitive Conditional Random Fields}\seclabel{learning}
 We take a CRF approach, whose nodes represent a labeling of
@@
 crowded scenes, the CRF has little chance to recover.
-
+%
 \begin{figure}
@@
 \end{figure}
-
+\pagebreak\\
+%
 \Figref{weights} visualizes some learned potential functions.
@@
 the second two images illustrate failure modes. Pixels marked as void
-are excluded from the evaluation. See the \secref{results} for details.
+are excluded from the evaluation. See \Secref{experiments} for details.
 \figlabel{comparison}}

### 2013-09-19 partial_supervision/paper.tex
spelling!

@@
 image level annotation, but also pixel level annotation of object. This allows
-us to judge the succes of learning on instance level.
+us to judge the success of learning on instance level.
 
@@
 on the MSRC-9 datasets.
-More recently, simliar approaches were proposed by
+More recently, similar approaches were proposed by
 \citet{vezhnevets2011weakly} and \citet{vezhnevets2010towards}.
 \citet{vezhnevets2011weakly} independently developed a multiple-instance based
-approach to segmentation, and report impressinve results on the MSRC-21 dataset.
+approach to segmentation, and report impressive results on the MSRC-21 dataset.
 
@@
 only.  This makes disseminating the distinctive features in multi-class object
-recognition much more challenging, and requires a more holistic apporach to
+recognition much more challenging, and requires a more holistic approach to
 recognition than these patch-based or superpixel-based approaches.
@@
 though the model was not trained to provide any instance-level labels, the
-performance is still competative.
+performance is still competitive.
 

### 2013-09-19 pystruct/paper.tex
spelling!

@@
     in popular structured prediction software packages. LSVM stands for the latent multi-class SVM, LDCRF for
-    latend dynamic conditional random fields.}
+    latent dynamic conditional random fields.}
 \end{table}
@@
 \section{Summary}
-Im this chapter, we introduced basic concepts of structured prediction. In particular, we discussed
+In this chapter, we introduced basic concepts of structured prediction. In particular, we discussed
 Structured Support Vector Machines, a max-margin approach for training linear predictors for

### 2013-09-19 structured_prediction/paper.tex
typos julia
spelling!

@@
 If $f$ is chosen as in \Eqref{linear_psi}, then the resulting distribution
-belongs to the exponential family, the class of probablility distributions most
-commonly used in the graphical model iterature.
+belongs to the exponential family, the class of probability distributions most
+commonly used in the graphical model literature.
 
@@
 
-\subsection{The 1-Slack Cutting Plane Method}\seclabel{one_slack}
+\subsection{The $1$-Slack Cutting Plane Method}\seclabel{one_slack}
 
 \begin{algorithm*}[t]
-    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
+    \caption{$1$-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
     \begin{algorithmic}[1]
@@
 Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
-This has the advantage of producing a QP that easyer to solve, as it contains far
+This has the advantage of producing a QP that easier to solve, as it contains far
 less variables than in the $n$-slack algorithm. The down-side of this is that

### 2013-09-20 exact_learning/paper.tex
Helvetica, Titel neu setzen, korrekturen einarbeiten.

@@
 approaches, later work employed graph-cuts~\citep{szummer2008learning} or Loopy
-Believe Propagation (LBP)~\citep{lucchi2011spatial}. These works use a single
+Belief Propagation (LBP)~\citep{lucchi2011spatial}. These works use a single
 inference algorithm for the whole learning process, and can not provide any
@@
 combination allows us to make fast progress initially and guarantee optimality
-in the end. Our strategy is visualized in \Figref{scheme_inference_algs}
-Notably, it is not necessary for an algorithm used as 3) to always produce
-exact results. For guaranteeing optimality of the model, it is sufficient that
-we obtain a certificate of optimality when learning stops.
+in the end. Our strategy is visualized in \Figref{scheme_inference_algs}.
+Notably, it is not necessary for an algorithm used as the third strategy to
+always produce exact results. For guaranteeing optimality of the model, it is
+sufficient that we obtain a certificate of optimality when learning stops.
 % TODO certify optimality != exact...

### 2013-09-20 itm/paper.tex
Helvetica, Titel neu setzen, korrekturen einarbeiten.

@@
 %
-\section{Clustering Using Non-Parametric Entropy Estimates}
+\section[Clustering Using Non-Parametric Entropy Estimates]{Clustering Using\linebreak Non-Parametric Entropy Estimates}
 In general, the goal of clustering can be formulated as follows: 

### 2013-09-20 nyu/paper.tex
Helvetica, Titel neu setzen, korrekturen einarbeiten.

@@
 
-\section{Learning Depth-Sensitive Conditional Random Fields}\seclabel{learning}
+\section[Learning Depth-Sensitive Conditional Random Fields]{Learning Depth-Sensitive\linebreak Conditional Random Fields}\seclabel{learning}
 We take a CRF approach, whose nodes represent a labeling of

### 2013-09-20 partial_supervision/paper.tex
Helvetica, Titel neu setzen, korrekturen einarbeiten.

@@
-\chapter{Weakly Supervised Object Segmentation}\label{ch:semi_supervised}
+\chapter[Weakly Supervised Object Segmentation]{Weakly Supervised\linebreak Object Segmentation}\label{ch:semi_supervised}
 
@@
 object categories, while semantic segmentation approaches often focus on
-"stuff" categories like "sky" and "road" that are more easily detected using
+"stuff" categories like "sky" and "road" which are more easily detected using
 simple texture features.
@@
 literature~\citep{andrews2003support,nguyen2010new,zhang2008m3miml}. Another
-advantage over other methods is that it uses a single convex optimization,
+advantage over other methods is the use of a single convex optimization,
 whereas other approaches often use iterative
@@
 	\end{center}
-        \caption{Overview of our method. See text for details.}
+        \caption{Schematic overview. See text for details.}
 	\figlabel{results}
@@
     \cmidrule{2-4}
-        This work&   0.30&  0.45&  0.43 \\
+        MIL-MKL (our approach)&   0.30&  0.45&  0.43 \\
         Best strongly supervised approaches~\citep{fulkerson2009class,schulz2011}&   0.72&  0.72&  0.66 \\

### 2013-09-20 structured_prediction/paper.tex
minor fixes
Helvetica, Titel neu setzen, korrekturen einarbeiten.

@@
-\chapter{Learning Conditional Random Fields}\label{ch:structured_pystruct}
+\chapter[Learning Conditional Random Fields]{Learning\linebreak[4] Conditional Random Fields}\label{ch:structured_pystruct}
 
@@
 
-\pagebreak %FIXME maybe?
+\pagebreak
 
@@
 
-\section{Learning Algorithms for Max-Margin Structured Prediction}\seclabel{learning_algorithms}
+\section{Learning Max-Margin Structured Prediction}\seclabel{learning_algorithms}
 

### 2013-09-20 partial_supervision/paper.tex
minor polish

@@
 % TODO implementation detail?
+\enlargethispage{10mm}
 For computing the global probability of boundary
@@
 while Musk2 contains significantly larger bags, more similar to the
-image/segment setup.
+ setup concerning images and segments.
 

### 2013-09-20 evaluation/paper.tex
starting on datasets for experiments.

@@
 \chapter{Empirical Comparison of Learning Algorithms}\label{ch:comparison}
+In this chapter, we provide an empirical evaluation of the learning algorithms
+described in Chapter~\ref{ch:structured_pystruct}.  We use the open source
+implementations in \pystruct, and publish the evaluation code and datasets with
+the package.
+
+\section{Datasets and Models}
+We consider several qualitatively different datasets, that have been widely used in the literature.
+The problems we consider are multi-class classification, sequence labeling, multi-label prediction,
+and general graph labeling.
+
+\subsection{Multi-Class Classification (MNIST)}
+The simplest task we use is multi-class classification. In this task,
+the inference problem is trivial, as it is assumed the target set is small enough
+to be enumerated efficiently. While this problem could be solved more efficiently
+with specialized algorithms, it nevertheless provides an initial insight into structured
+prediction algoriths.
+
+We choose the classical MNIST dataset of handwritten digits, consisting of
+60000 training images and 10000 test images of the digits zero to nine. Each
+image is a $28 \times 28$ grey-level image, resulting in a $784$-dimensional
+feature vector. We normalize the feature to lie between $0$ and $1$.
+
+The model we use for multi-class classification is the Crammer-Singer
+formulation. We do not include a bias, leading to $784 \cdot 10 = 7840$
+parameters in the model.
+
+\subsection{Sequence Labeling (OCR)}
+A classical application of structured prediction is sequence labeling.
+We choose the ``OCR'' dataset introduced in the seminal work of \citet{taskar2003max}.
+Each input consists of the segmented handwritten letters of a word in lower case. The task is to classify
+all letters in a word, that is assign each segmented input letter one of the
+classes ``a'' to ``z''. As the first letter of each word was capitalized, these
+were removed by \citet{taskar2003max}, leading to somewhat odd-looking
+labels. The words are between \ldots and \ldots letters long, and each letter
+is represented as a binary image of size \ldots.
+We use a simple chain model with a single constant pairwise feature.
+This means the unary potential has \ldots parameters, one for each input
+feature and output class. The pairwise potential consists of a matrix of
+transition potentials with $26 x 26=676$ entries.
+It is well-known that efficient exact inference in chains is possible using
+message passing algorithms. 
+%We use the dynamic programming algorithm
+%implemented in \textsc{OpenGM}.
+
+\subsection{Multi-Label Classification}
+Multi-label classification is a generalization of multi-class classification,
+in which each example can be associated with more than one class. In other
+words, the algorihm must decide for each sample and for each class whether the
+sample belongs to that class or not.
+Multi-label classification was first formulated as a structured prediction problem
+in by \citet{finley2008training}, where it was used to investigate the influence of
+approximate inference on the $n$-slack cutting plane algorithm. In this formulation,
+each class is represented as a binary node in a factor graph---the states representing
+presence of absence of the class. A different factor of pairwise potentials is introduced
+between each pair of classes.
+We also consider a different model, where pairwise potentials are only
+introduced between specific nodes. Build a tree over the binary variables by
+computing the Chow-Liu Tree~\citep{chow1968approximating} of the targets. While
+this results in a less expressive model, a tree-shaped model allows for exact
+inference via message passing.
+We use two of the datasets used in \citet{finley2008training}, the scene
+and yeast datasets. We choose these two as these are real-world datasets for
+which the pairwise approach outlined above actually improves upon the
+baseline~\citep{finley2008training}.
+The scene dataset has six labels, 294 input features, 1211 training samples and
+1196 test samples. This leads to $204 \cdot 6$ parameters for the unary potentials,
+$3 \cdot 5 \cdot 4$ parameters for the full pairwise potentials (four for each edge), and
+$5 \cdot 4$ parameters for the pairwise potentials of the tree-shaped model.
+Using four parameters for each edge is a slight over-parametrization that
+simplifies writing out the model.
+The yeast dataset has 14 labels, 103 features, 1500 training samples and 971
+test samples. The resulting numbers of parameters are $14 \cdot 103$ parameters
+for the unary potentials, $7 \cdot 13 \cdot 4$ parameters for the full pairwise
+potential, and $13 \cdot 4$ parameters for the tree-shaped model.
+
+\subsection{2D Grid CRF (Snakes)}
+The snakes dataset is a synthetic dataset, where samples are labeld 2D grids.
+It was introduced in \citet{nowozin2011decision} to demonstrate the importance
+of learning conditional pairwise potentials. The dataset consists of ``snakes''
+of length ten traversing the 2D grid. Each grid cell that was visited is marked
+as the snake heading out towards the top, bottom, left, or right, while
+unvisited cells are marked as background.  The goal is to predict a labeling of
+the snake from ``head'' to ``tail'', that is assigning numbers from zero to
+nine to the cells that are occupied by the snake.  An illustration is given in
+%TODO
+The task is interesting, as local evidence for the target label is weak except
+around head and tail. Therefor pairwise potentials are necessary to solve the
+task. The dataset is noise-free in the sense that given the above description, a
+human could easily produce the desired labeling without making any mistake.
+The dataset is also interesting as the model of \citet{nowozin2011decision} produced
+notoriously hard to optimize energy functions.
+
+In \citet{nowozin2011decision}, the input was encoded into five RGB colors
+(``up'', ``down'', ``left'', ``right'', ``background'').  To encode the input
+more suitable for our linear methods, we convert this representation to a
+one-hot encoding of the five states.
+
+We use a grid CRF model for this task. Unary for each node are given by the
+input of the 8-neighborhood of the node---using a 4-neighborhood would most
+likely yield better results, but we do not want to encode too much
+task-knowledge into our model.
+Using the one-hot encoding of the input, this leads to $9 \dot 5 = 45$ unary features.
+With 11 output classes, the unary potential has $11 \cdot 9 \cdot 5$ parameters.
+Features for the pairwise potentials are constructed by concatenating the features
+of the two neighboring nodes, taking the direction of the edge into account.
+The pairwise feature therefore has dimensionality $45 \cdot 4$, with the first
+$45$ entries corresponding to the feature of the ``top'' node, the second $45$
+entries to the features ``bottom'' node, followed bye the ``left'' and
+``right'' nodes. As each edge is either horizontal \emph{or} vertical, only two
+of these parts will be non-zero for any given edge.  With $45 \cdot 4$ edge
+features, the pairwise potentials has $45 \cdot 4 \cdot 11^2$ parameters.
+
+\subsection{Superpixel CRFs for Semantic Segmentation}

### 2013-09-20 evaluation/paper.tex
experiment stuff

@@
 \subsection{Superpixel CRFs for Semantic Segmentation}
+Our main attention is devoted to the use of conditional random fields for
+semantic segmentation.  We use the Pascal VOC and MSRC datasets. A detailed
+description of the method will be given in Chapter~\ref{ch:exact_learning}.
+%TODO will it?
+The Pascal VOC has \ldots training and \ldots test images, each having
+around 100 superpixels. There are 20 object categories, plus background. The % TODO check!
+MSRC dataset has \ldots training and \ldots test images, also segmented in
+around 100 superpixels each. We use the MSRC-21 dataset, which has 21 target
+classes.
+Each superpixel is represented as an output variable, with the ground truth obtained
+by majority vote over the pixel belonging to the superpixel.
+Pairwise potentials were introduced for each pair of neighboring superpixels.
+We then removed all superpixels that were labeled as ``void'', together with
+their pairwise potentials.
+As unary potentials, we use the TextonBoost probabilities provided by \citet{krahenbuhl2012efficient}.
+We average probabilities inside each superpixel, yielding 21 input features for both datasets.
+We also used the same pairwise features for the two datasets: a constant feature,
+a color contrast feature, and a feature encoding relative vertical position between superpixels.
+Overall, the models for both datasets had $21 \cdot 21 + 21^2 \cdot 4$ features.
+It is possible to assume symmetric potentials for this dataset, leading to less
+pairwise features.  We found both parametrizations yielded similar results.
+
+\section{Experiments}
+We compare the following algorithms on the above models:
+\begin{itemize}
+    \item Stochastic Subgradient Descent using the Pegasos schedule: %TODO
+    \item Stochastic Subgradient Descent using a momentum heuristic:
+    %\item Stochastic Subgradient Descent using a momentum heuristic: AVERAGING
+    %\item Stochastic Subgradient Descent using a momentum heuristic: AVERAGING
+    \item The $1$-slack cutting plane algorithm without constraint caching.
+    \item The $1$-slack cutting plane algorithm with constraint caching. For
+            each sample, the last $50$ inference results are stored.
+            Details of the caching strategy can be found in Chapter~\ref{ch:exact_learning}.
+            %TODO
+    \item The $n$-slack cutting plane algorithm, where the QP is solved after each training.
+    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 training samples.
+    \item The SZLJSP algorithm.
+\end{itemize}
+All algorithm take a $C$ parameter, which we adjusted on a fully trained model
+on a validation set (experiments not reported here) and held constant for all models.
+We found, however, that the algorithms and models are quite robust to the choice of $C$ within
+one or two orders of magnitude.
+As stopping criterion, we used a duality gap of $0.1$ when possible, and a pre-defined
+number of iterations for the subgradient algorithms.
+
+Our ultimate evaluation criterion is how fast the learning algorithms converge,
+and how robust they are to approximate inference.  To quantify our goals, we
+track primal suboptimality and training set error during learning, as well as
+final error on the test set.
+We report suboptimality as a function of runtime, and as a function of passes over the training set.
+Both are important to practitioners, as they give important insight into the workins of the algorithm.
+The actual runtime is certainly the most important factor, but also highly influenced by implementation details
+and properties of the dataset. We are particularly interested in cases where inference is highly non-trivial.
+In these cases, inference

### 2013-09-20 evaluation/paper.tex
blabla for experiments...

@@
 message passing algorithms.
-%We use the dynamic programming algorithm
-%implemented in \textsc{OpenGM}.
 
@@
 one or two orders of magnitude.
-As stopping criterion, we used a duality gap of $0.1$ when possible, and a pre-defined
-number of iterations for the subgradient algorithms.
+As stopping criterion, we used a duality gap of $0.1$ when possible, and a
+pre-defined number of iterations for the subgradient algorithms. It is worth
+noting that the quadratic programming based algorithms have additional
+hyper-parameters that we do not discuss here in detail, such as the threshold
+for removing a constraint as inactive, how often inactive constraints are
+removed, and parameters of the underlying QP solver. On the other hand, the
+SZLJSP, for example, has no hyper-parameters except for the stopping criterion.
 
@@
 final error on the test set.
-We report suboptimality as a function of runtime, and as a function of passes over the training set.
-Both are important to practitioners, as they give important insight into the workins of the algorithm.
-The actual runtime is certainly the most important factor, but also highly influenced by implementation details
-and properties of the dataset. We are particularly interested in cases where inference is highly non-trivial.
-In these cases, inference
+We report suboptimality as a function of runtime, and as a function of passes
+over the training set.  Both are important to practitioners, as they give
+important insight into the workins of the algorithm.  The actual runtime is
+certainly the most important factor, but also highly influenced by
+implementation details and properties of the dataset. We are particularly
+interested in cases where inference is highly non-trivial, making inference the
+dominating factor of runtime.
+We define one pass over the dataset as calling prediction once for each sample.
+This way, we can get a clear picture how learning times with complexity of the
+inference task.
+
+\subsection{Experiments using Exact Inference}
+In this section, we present results on multi-class classification, sequence
+labeling and multi-label prediction with a tree-structured model. In these
+tasks, exact inference is possible, and we can compute the exact objective of
+the various algorithms easily. We use the dynamic programming algorithm
+implemented in \textsc{OpenGM} for experiments
+
+\subsection{Experiments using Approximate Inference}
+For the remaining tasks of multi-label classification with the full graph, the
+snakes dataset and image segmentation, exact inference is usually not possible.
+
+We use two different approximate inference algorithms, fusion moves, a fast
+local search procedure, and AD$^3$, a dual-subgradient method to solve the LP
+relaxation of the inference problem.
+Evaluation of algorithms where exact inference is not possible is much harder,
+as it is usually not possible to evaluate the exact objective. When using AD$^3$,
+we can still consider the relaxed task, which will be solved exactly in the
+majority of cases (AD$^3$ sometimes fails to find the primal solution to the LP
+relaxation). %TODO is this happening?
+When using fusion moves, there is no obvious interpretation of the produced
+primal objective, other than that it provides a lower bound of the actual
+objective. Nevertheless, we find it informative to analyze the behavior of this lower bound.
+
+For both inference algorithms, we also evaluate the predictive performance on
+the training set.  We explicitly do not consider the relaxed problem
+here---instead we round possible fractional results.
+While the training error does not directly reflect the objective, it provides
+a measure of how effective the ``prediction machine'' given by the learned model
+together with the inference algorithm is.
+
+\section{Discussion}

### 2013-09-20 exact_learning/paper.tex
vladn comments

@@
 As a fast under-generating inference algorithm, we used $\alpha$-expansion
-moves based on non-submodular graph-cuts using Quadratid Pseudo-Boolean
+moves based on non-submodular graph-cuts using Quadratic Pseudo-Boolean
 Optimization (QPBO)~\citep{rother2007optimizing}.  This move-making strategy
@@
 method provided by \citet{rother2007optimizing}.
-We use the excellent implementation of AD$^3$ provided by the authors of
+We use the excellent implementation of AD$^3$ provided by
 \citet{martins2011augmented}.

### 2013-09-20 nyu/paper.tex
vladn comments

@@
 stereo images. They approach the task by jointly reasoning about 3D layout and
-semantics of the scenes and produce a dense labeling on image level.  In
-\citet{senguptaurban}, they extend their approach to produce a volumetric
+semantics of the scenes and produce a dense labeling on image level.
+\citet{senguptaurban} extend the approach to produce a volumetric
 reconstruction of the scene, together with a dense semantic labeling of the
@@
 \citet{ladicky2009associative}, which is similar in spirit to our approach,
-but uses Potts potentials together with cross-validation to set potentials.
+but uses Potts potentials together with cross-validation to adjust parameters.
 

### 2013-09-20 pystruct/paper.tex
minor fixes
vladn comments
make hyperref much prettier!! much!! prettier!!! also fix some stuff...

@@
 and {\sc AD$^3$}~\citep{martins2011augmented}. It also includes an interface to
-a general purpose linear programming solver from \textsc{cvxopt}~\cite{dahl2006cvxopt}.
+a general purpose linear programming solver from \textsc{cvxopt}~\citep{dahl2006cvxopt}.
 
@@
 from \citet{krahenbuhl2012efficient}. The superpixels are generated using SLIC~\citep{achanta2012slic}.%
-\footnote{The preprocessed data can be downloaded at \url{http://www.ais.uni-bonn.de/download/datasets.html}.}
+\footnote{The preprocessed data can be found at \url{http://www.ais.uni-bonn.de/download/datasets.html}.}
 Each sample (corresponding on one entry of the list \texttt{X}) is represented as a

### 2013-09-21 intro.tex
spelling!
more corrections..

@@
 \\[2cm]
-In computer vision research, the goal is to automatically extract information from an image.
-In particular discerning semantic information, that is interpreting an image, is a prominent
-research topic.
+In computer vision research, the goal is to automatically extract information
+from a given image of image sequence.  In particular discerning semantic
+information, that is interpreting an image, is a prominent research topic.
 While much progress has been made in recent years, computer vision systems still lag behind
-humans in most tasks that require semantic information. Often these tasks can be formulated
+human vision in most tasks that require semantic information. These tasks can often be formulated
 in terms of \emph{semantic classes}, meaning categories of parts, objects or scenes.
@@
 These questions illustrate a range of possible tasks involving semantic categories,
-from classifying images of single objects, over localization and counting of object classes,
-to parsing a scene fully into objects and object classes, together with their relations.
+such as classifying images of single objects, localization and counting of object classes,
+and parsing a scene fully into objects and object classes together with their relations.
 %
@@
 little trouble in interpreting complex scenes, current methods are often
-restricted to a much smaller number of classes, and only have limited
+restricted to a much smaller number of classes and only have limited
 capabilities to model interactions or relations.  We believe that
@@
 image interpretation for retrieval, autonomous driving and mobile robotics.
-While these are already important applications, we expect that with the
-abundance of camera data, and the proliferation of mobile computing, even
-more applications will emerge soon.
+%While these are already important applications, we expect that with the
+%abundance of camera data, and the proliferation of mobile computing, even
+%more applications will emerge soon.
+Besides these applications, due to the abundance of camera data and the
+proliferation of mobile computing, we expect semantic annotation of images to
+be a key component in future technologies.
 
 In the following we distinguish between the task of semantic segmentation,
-which is usually distinguishing unstructured ``stuff'' classes such as road and
+which usually distinguishes unstructured ``stuff'' classes such as road and
 grass, and object class segmentation, which denotes the segmentation of very
@@
 Both tasks have the same ultimate goal of parsing, and therefore understanding,
-images in terms of semantic classes, but usually employ  different mechanisms
+images in terms of semantic classes. However they employ different mechanisms
 to represent and process.
@@
 is still little understood.
-We focus on the use of conditional random fields~(CRFs), which have shown great
-promise for computer vision applications. Using the paradigm of structural
+We focus on the use of conditional random fields~(CRFs), which have shown
+promising  results for computer vision applications. Using the paradigm of structural
 support vector machines, it is possible to learn conditional random fields to
-directly minimize a loss of interest. In particular, conditional random fields
+directly minimize a loss of interest. In particular, CRFs
 allow to combine different cues, possibly produced by other approaches, in a
@@
 framework. \citet{meshi2010learning, komodakis2011efficient}, and
-\citet{hazan2010primal} approached the problem using duality, and formulate
+\citet{hazan2010primal} approach the problem using duality, and formulate
 learning and inference as a joint optimization problem.
@@
 on chain CRFs, where efficient inference is possible.
-More recently, \citet{lucchi2013learning} proposed a novel algorithm for
+More recently, \citet{lucchi2013learning} propose a novel algorithm for
 efficiently learning structured prediction for semantic segmentation, using
-approximate inference. In Chapter~\ref{ch:exact_learning}, we introduce an
+approximate inference. In Chapter~\ref{ch:exact_learning}, we develop an
 algorithm that runs in similar time to the one proposed by
-\citet{lucchi2013learning}, but is able to learn a CRF to the exact optimum on
+\citet{lucchi2013learning}, which is able to learn a CRF to the exact optimum on
 the same dataset.
@@
 \item Introducing a clustering algorithm that improves upon widely used
-    approaches from the literature.
-\item Demonstrating a scalable algorithm for weakly supervised object class segmentation.
-\item Providing a general and efficient open source software framework for structured prediction.
-\item Analysing max-margin learning algorithms with exact or approximate inference in different applications.
+    approaches from the literature. Our algorithm yields better clusterings in
+    terms of known classes on a wide range of standard datasets.
+\item Demonstrating a scalable algorithm for weakly supervised object class
+    segmentation.  The proposed method is able to learn to segment complex object
+    classes using image annotation alone.
+\item Providing a general and efficient open source software framework for
+    structured prediction. 
+\item Analysing max-margin learning algorithms with exact or approximate
+    inference in different applications. We give a thorough evaluation of
+    all major SSVM learning algorithms, using a wide array of application.
 \item Showing that exact learning for semantic segmentation and object class
-    segmentation is possible in practice, even in loopy graphs.
+    segmentation is possible in practice, even in loopy graphs. We combine
+    fast inference, caching and inference algorithms which certify optimality
+    to learn a $1$-slack SSVM.
 \end{itemize}
@@
 
-The central topic of this thesis is introduced in
-Chapter~\ref{ch:structured_pystruct}, which also introduces our software
-library for implementing structured learning and prediction algorithms.
+The central topic of this thesis, learning structured prediction for semantic
+segmentation, is introduced in Chapter~\ref{ch:structured_pystruct}. This
+chapter also introduces our software library for implementing structured
+learning and prediction algorithms.
 
@@
 problem of semantic annotation with structure classes in RGB-D data. We
-demonstrate that we are able to learn meaningful spacial relations, and improve
-upon the state-of-the-art on the NYU V2 datasets.
+demonstrate that we are able to learn meaningful spatial relations, and outperform
+state-of-the-art methods on the NYU V2 datasets.
 
@@
         Partially Supervised Image Segmentation} Andreas C. M\"uller and Sven
-        Behnke. Published in the proceedings of the IARP Worshop on Partially
+        Behnke. Published in the proceedings of the IARP Workshop on Partially
         Supervised Learning.

### 2013-09-21 main.tex
Helvetica, Titel neu setzen, korrekturen einarbeiten.
make hyperref much prettier!! much!! prettier!!! also fix some stuff...
more corrections..

@@
 
-\usepackage{hyperref}
-\usepackage{epsfig}
-\usepackage{subfigure}
-\usepackage{calc}
+\usepackage[colorlinks=True, citecolor=blue!50!black, linkcolor=red!50!black, urlcolor=blue!70!black]{hyperref}
 \usepackage{refstyle}
@@
 \usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{multicol}
-%\usepackage{pslatex}
 \usepackage{microtype}
@@
 \usepackage{mdwlist}
-\usepackage{url}
 \usepackage[noend]{algpseudocode}
@@
 %\renewcommand*\sfdefault{avant}
- %\usepackage{helvet}
-\usepackage[defaultsans]{droidsans}
+\usepackage{helvet}
+%\usepackage[defaultsans]{droidsans}
 %\renewcommand*\sfdefault{lmssq}
@@
  results on the MSRC-21 dataset.
-We demonstrated the power of conditional interactions by learning spacial
+We demonstrated the power of conditional interactions by learning spatial
 interactions in an RGB-D setting. Here, our approach improves upon the state-of-the

### 2013-09-21 intro.tex
address more comments...

@@
 different datasets in this thesis: the object class segmentation datasets
-Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21, and
-NYU V2.
+Graz02 and Pascal VOC 2010~\citep{pascal}, and the semantic segmentation datasets
+MSRC-21~\citep{shotton2006textonboost}, and NYU V2~\citep{SilbermanECCV12}.
 %

### 2013-09-21 structured_prediction/paper.tex
more corrections..
address more comments...

@@
 MRFs and CRFs provide a principled way of integrating local evidence and
-modeling spacial dependencies, which are strong in most image-based tasks.
+modeling spatial dependencies, which are strong in most image-based tasks.
 %
@@
 
-While $y$ could be complicated like a parsing tree or the geometric
+While $y$ could be complicated like a parse tree or the geometric
 configuration of a molecule, many settings, such as the image segmentation
@@
                 \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
-                    \State $\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
+                \State \hspace{-3mm}$\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
                     \State
@@
                     \begin{align*}
-                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle &\argmin_{\theta, \xi_1, \dots, \xi_k} \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
-                    &\text{s.t. }\forall i=1,\dots,k \forall \hat{y} \in \W_i:\\
+                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle \argmin_{\theta, \xi_1, \dots, \xi_k}& \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
+                        \text{s.t. }& \text{for } i=1,\dots,k\ \forall \hat{y} \in \W_i:\\
                     &\theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
@@
 \begin{align}\label{eq:nslack}
-    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
-    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
-    &\theta^T \sum_{i=1}^n [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i},
-        \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
-            - \xi
+    \min_{\theta, \xi_1, \dots, \xi_k}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
+    \text{s.t. }&\text{for } i=1,\dots,k\ \forall \hat{y} \in \mathcal{Y}:\\
+    &\theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i},
+        \hat{y})] \geq \Delta(y\hoch{i}, \hat{y})
+            - \xi_i
 \end{align}
@@
             \begin{align*}
-            (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi\\
-            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
+            (\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi}&\frac{||\theta||}{2}^2 + C \xi\\
+            \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
              &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi

### 2013-09-21 structured_prediction/paper.tex
fix some equation numbering.

@@
 considering a single training example $(x\hoch{i}, y\hoch{i})$
-can be computed simply by solving the loss-augmented prediction problem:
-\begin{align}
-    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
+can becomputed simply by solving the loss-augmented prediction problem:
+\begin{align*}
+    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\numberthis\\
     &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
-\end{align}
+\end{align*}
 The most commonly used update has the simple form
@@
 into a quadratic objective with a combinatorial number of constraints:
-\begin{align}\label{eq:nslack}
-    \min_{\theta, \xi_1, \dots, \xi_k}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
+\begin{align*}
+    \min_{\theta, \xi_1, \dots, \xi_k}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i \numberthis \label{eq:nslack}\\
     \text{s.t. }&\text{for } i=1,\dots,k\ \forall \hat{y} \in \mathcal{Y}:\\
@@
             - \xi_i
-\end{align}
+\end{align*}
 As it is not feasible to deal with all constraints, only a working set $\W$ of active constraints
@@
 following reformulation of Equation~\Eqref{learning_equation}:
-\begin{align}\label{eq:oneslack}
-    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
+\begin{align*}
+    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi \numberthis\label{eq:oneslack}\\
     \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
@@
             - \xi
-\end{align}
+\end{align*}
 \enlargethispage{10mm}
@@
         find
-        \[ \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
+        \begin{equation}
+            \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
             \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\Phi(x^i, y^i) -
-                \Phi(x^i, \hat{y})] \]
+                \Phi(x^i, \hat{y})] 
+        \end{equation}
         by enumeration of $C^i$ and continue until

### 2013-09-22 exact_learning/paper.tex
address more comments...
fix some equation numbering.
phi/psi
corrections luisa

@@
-\chapter{Learning of Loopy CRF Exactly}\label{ch:exact_learning}
+\chapter{Learning Loopy CRF Exactly}\label{ch:exact_learning}
 
@@
             &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
-            &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
+            &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
             \end{align*}
@@
                 \State
-                $\hat{y}\hoch{i} \leftarrow \hat{I}(x\hoch{i}, y\hoch{i}, \theta) \approx \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane_again}
+            $\hat{y}\hoch{i} \leftarrow \hat{I}(x\hoch{i}, y\hoch{i}, \theta) \approx \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane_again}
             \EndFor
             \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
-            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
+        \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] $
         \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
@@
 that is solved by the $1$-slack cutting plane algorithm:
-\begin{align}\eqlabel{oneslack_again}
-    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
+\begin{align*}
+    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi \numberthis\label{eq:oneslack_again}\\
     \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
@@
             - \xi
-\end{align}
+\end{align*}
 For reference, the algorithm described in
@@
 \centering
-\includegraphics[width=\linewidth]{exact_learning/images/inference_algs}
+\includegraphics[width=.6\linewidth]{exact_learning/images/inference_algs}
 \caption{%

### 2013-09-22 itm/paper.tex
corrections luisa

@@
 \begin{itemize}
-\item $\displaystyle D_\text{KL}\, = \int_\mathcal{X} p(X) \ln\left(\frac{p(X)}{q(X)}\right)\,dX$ is the Kullback-Leibler divergence,
+    \item $\displaystyle D_\text{KL}(p(X) \parallel q(X))\, = \int_\mathcal{X} p(X) \ln\left(\frac{p(X)}{q(X)}\right)\,dX$ is the Kullback-Leibler divergence,
 \item $\displaystyle H(X)\, = \int_\mathcal{X} p(X) \ln(p(X))\,dX$ is the differential entropy, and
@@
 
+\subsection{Finding Euclidean Minimum Spanning Tree Clusterings}
+The objective given by \Eqref{objective} is a non-linear combinatorial
+optimization problem. It has two properties that make it hard to optimize:
+\begin{enumerate}
+    \item The objective depends in a non-linear way on $L_y$.
+        This makes linear programming techniques, that proved successful for
+        other combinatorial tasks, not directly applicable.
+    \item $L_y$ is defined in terms of minimum spanning trees. This set is hard
+        to characterize, as changing the cluster membership of a single node may change
+        the two minimum spanning trees involved completely.
+\end{enumerate}
+
+\begin{figure}
+\centering
+\includegraphics[width=.32\textwidth]{illustration_left}
+\includegraphics[width=.32\textwidth]{illustration_center}
+\includegraphics[width=.32\textwidth]{illustration_right}
+\caption{Illustration of the optimization algorithm for $k=3$ on synthetic
+dataset. \emph{Left}: Euclidean minimum spanning tree of the data.
+\emph{Center}: The edge that yields the best two-cluster partition in terms of
+\Eqref{objective} was removed, yielding two connected components. \emph{Right}:
+Another edge from the forest was removed, resulting in the desired number of
+three components. Note that the edges that are removed are not the longest edges
+but form a trade-off between edge length and cluster size.}
+\label{fig:illustration}
+\end{figure}
+
 \begin{algorithm}[t]
-\begin{algorithmic}\label{overall}
+\begin{algorithmic}
     \Require Points $\mathbf{x}$, desired number of clusters $k$.
@@
         \For{node $x$}
-        \State $w'_x \leftarrow w_{\text{root}} - w_x$
+        \State $w'_x \leftarrow w_{x_0} - w_x$
         \EndFor
@@
 \end{algorithmic}
-\caption{Information Theoretic MST-based Clustering}
+\caption{Information Theoretic MST-based Clustering\label{overall}}
 
 \end{algorithm}
-
-\subsection{Finding Euclidean Minimum Spanning Tree Clusterings}
-The objective given by \Eqref{objective} is a non-linear combinatorial
-optimization problem. It has two properties that make it hard to optimize:
-\begin{enumerate}
-    \item The objective depends in a non-linear way on $L_y$.
-        This makes linear programming techniques, that proved successful for
-        other combinatorial task, not directly applicable.
-    \item $L_y$ is defined in terms of minimum spanning trees. This set is hard
-        to characterize, as changing the cluster membership of a single node may change
-        the two minimum spanning trees involved completely.
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[width=.32\textwidth]{illustration_left}
-\includegraphics[width=.32\textwidth]{illustration_center}
-\includegraphics[width=.32\textwidth]{illustration_right}
-\caption{Illustration of the optimization algorithm for $k=3$ on synthetic
-dataset. \emph{Left}: Euclidean minimum spanning tree of the data.
-\emph{Center}: The edge that yields the best two-cluster partition in terms of
-\Eqref{objective} was removed, yielding two connected components. \emph{Right}:
-Another edge from the forest was removed, resulting in the desired number of
-three components. Note that the edge that are removed are not the longest edges
-but form a trade-off between edge length and cluster size.}
-\label{fig:illustration}
-\end{figure}
 
@@
 %filler removed Let us first fix some notation.
-Consider a graph $G$ with nodes $\mathbf{x}$ and edge weights given by the
+Consider a graph $G$ with nodes $\mathbf{x}$, an arbitray set of edges, and edge weights given by the
 Euclidean distances between points.  The connected components of $G$ induce a
@@
 \end{equation}
-
 Optimization over forests remains hard, and we further restrict ourself to
@@
 improve upon this, we use a dynamic programming approach as described in
-Algorithm~1, in function SplitCluster, which has only linear
+function \textsc{SplitCluster} of Algorithm~\ref{overall} , which has only linear
 complexity. Using this algorithm, run time in the two cluster case is dominated
@@
 components equals the number of desired clusters. The overall procedure is
-summarized in Algorithm~1, an illustration can be found in
-\Figref{illustration}. We refer to Algorithm~1 as \emph{Information
+summarized in Algorithm~\ref{overall}, an illustration can be found in
+\Figref{illustration}. We refer to Algorithm~\ref{overall} as \emph{Information
 Theoretic MST-based (ITM) clustering}.
@@
 $\alpha$ is the inverse of the Ackerman function. The dynamic programming solution of
-Algorithm~1 has a run time of $O(n)$ per removed edge, leading to an overall
+Algorithm~\ref{overall} has a run time of $O(n)$ per removed edge, leading to an overall
 run time of $O(n \log(n) \alpha(n)+ nk)$. The $O(nk)$ comes from a worst case
@@
 As single link agglomerative clustering is sensitive to outliers, we set
-a hard limit on the minimum number of samples per cluster of five for
+a hard limit of five on the minimum number of samples per cluster for
 the quantitative analysis.
@@
 We evaluated the experiments using the \emph{adjusted Rand index 
-(ARI}~\citep{hubert1985comparing}  popular measures 
+(ARI}~\citep{hubert1985comparing} a popular measure
 of cluster quality~\citep{gomes2010discriminative,kamvar2003spectral}.
@@
 \cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
-digits      &  1797 &   64 &  10 &     0.62 &      0.67            &0.10          &\B{0.85}       &0.73\\ 
-faces       &   400 & 4096 &  40 &     0.41 &      0.49            &0.08          &0.02           &\B{0.54}\\ 
-iris        &   150 &    4 &   3 &     0.72 &      0.75            &0.55          &\B{0.88}       &0.88\\
-usps        &  9298 &  256 &  10 &     0.52 &      0.54            &0.00          &   0.44        &\B{0.64}\\ 
-vehicle     &   846 &   18 &   4 &     0.10 &      0.09            &0.00          &\B{0.10}       &0.10\\
-vowel       &   990 &   10 &  11 &     0.17 &      0.19            &0.00          &\B{0.20}       &0.19\\
-waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
-mnist       & 70000 &  784 &  10 &     0.37 &      N/A$^\dagger$         &0.00          &   0.50        &0.77\\
+digits      &$ 1797$&$  64$&$ 10$& $    0.62 $&     $0.67$           &$0.10$         &$\mathbf{0.85}$       &$0.73$\\ 
+faces       &$  400$&$4096$&$ 40$& $    0.41 $&     $0.49$           &$0.08$         &$0.02$           &$\B{0.54}$\\ 
+iris        &$  150$&$   4$&$  3$& $    0.72 $&     $0.75$           &$0.55$         &$\B{0.88}$       &$0.88$\\
+usps        &$ 9298$&$ 256$&$ 10$& $    0.52 $&     $0.54$           &$0.00$         &$   0.44$        &$\B{0.64}$\\ 
+vehicle     &$  846$&$  18$&$  4$& $    0.10 $&     $0.09$           &$0.00$         &$\B{0.10}$       &$0.10$\\
+vowel       &$  990$&$  10$&$ 11$& $    0.17 $&     $0.19$           &$0.00$         &$\B{0.20}$       &$0.19$\\
+waveform    &$ 5000$&$  21$&$  2$& $\B{0.37} $&     $0.30$           &$0.00$         &$   0.23 $       &$0.23$\\
+mnist       &$70000$&$ 784$&$ 10$& $    0.37 $&     N/A$^\dagger$    &$0.00$         &$   0.50$        &$\B{0.77}$\\
 \bottomrule

### 2013-09-22 evaluation/paper.tex
visualizations for experiment datasets.

@@
 \subsection{Multi-Class Classification (MNIST)}
+\begin{figure}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_0}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_1}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_2}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_3}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_4}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_5}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_6}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_7}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_8}
+    \includegraphics[width=.092\linewidth, frame]{evaluation/images/mnist_digit_9}
+    \caption{%
+        Visualization of samples from the ten classes in the MNIST dataset.\figlabel{mnist_illustration}
+    }
+\end{figure}
 The simplest task we use is multi-class classification. In this task,
@@
 prediction algoriths.
-
 We choose the classical MNIST dataset of handwritten digits, consisting of
-60000 training images and 10000 test images of the digits zero to nine. Each
-image is a $28 \times 28$ grey-level image, resulting in a $784$-dimensional
-feature vector. We normalize the feature to lie between $0$ and $1$.
+60000 training images and 10000 test images of the digits zero to nine.
+\Figref{mnist_illustration} shows some examples. Each image is a $28 \times 28$
+grey-level image, resulting in a $784$-dimensional feature vector. We normalize
+the feature to lie between $0$ and $1$.
 
@@
 
+\begin{figure}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_00}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_01}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_02}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_03}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_04}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_05}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_06}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_07}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_08}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_09}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_10}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_11}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/justifications_12}\\[3mm]
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_00}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_01}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_02}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_03}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/skiing_04}\\[3mm]
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_00}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_01}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_02}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_03}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_04}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_05}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_06}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_07}
+    \includegraphics[width=.068\linewidth, frame]{evaluation/images/unexpected_08}\\
+    \caption{%
+        Visualization of some words from the OCR dataset. The first letters were removed by \citet{taskar2003max} as 
+        they were capitals.
+        The words are: (j)ustifications, (s)kiing and (u)nexpected. The letters in parentheses are not part of the dataset.
+        \figlabel{ocr_illustration}
+    }
+\end{figure}
 \subsection{Sequence Labeling (OCR)}
@@
 We choose the ``OCR'' dataset introduced in the seminal work of \citet{taskar2003max}.
-Each input consists of the segmented handwritten letters of a word in lower case. The task is to classify
-all letters in a word, that is assign each segmented input letter one of the
-classes ``a'' to ``z''. As the first letter of each word was capitalized, these
-were removed by \citet{taskar2003max}, leading to somewhat odd-looking
-labels. The words are between \ldots and \ldots letters long, and each letter
-is represented as a binary image of size \ldots.
+Each input consists of the segmented handwritten letters of a word in lower
+case. The task is to classify all letters in a word, that is assign each
+segmented input letter one of the classes ``a'' to ``z''. As the first letter
+of each word was capitalized, these were removed by \citet{taskar2003max},
+leading to somewhat odd-looking labels. The words are between three and fourteen
+letters long, and each letter is represented as a binary image of size $16
+\time 8$, with a toal of $6877$ words.
+The dataset is devided into ten folds. We consider two setups: learning
+on one fold, and testing on the remaining nine folds, following \citet{taskar2003max},
+and learning on nine folds and testing on the remaining fold, following \citet{lacoste}.
+We refer to the learning on one fold as OCR-small, and learning on nine folds as OCR-big.
 We use a simple chain model with a single constant pairwise feature.
-This means the unary potential has \ldots parameters, one for each input
+This means the unary potential has $16 \cdot 8 \cdot 26$ parameters, one for each input
 feature and output class. The pairwise potential consists of a matrix of
-transition potentials with $26 x 26=676$ entries.
+transition potentials with $26 \cdot 26=676$ entries.
 It is well-known that efficient exact inference in chains is possible using
@@
 
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XXXXXXX@{}}
+    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0001}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0002}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0003}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0004}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0005}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0006}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_input_0007}\\
+    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0001}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0002}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0003}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0004}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0005}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0006}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snake_label_0007}
+    \end{tabu}
+\caption{%
+    Visualization of the snakes dataset. The top row shows input patterns, the bottom row the corresponding
+    labels. The colors in the image showing the labels correspond to dark blue for background, and encoding
+    the length of the snake from head (red) to tail (blue).
+\figlabel{snakes_illustration}}
+\end{figure}
+
 \subsection{2D Grid CRF (Snakes)}
@@
 the snake from ``head'' to ``tail'', that is assigning numbers from zero to
-nine to the cells that are occupied by the snake.  An illustration is given in
-%TODO
-The task is interesting, as local evidence for the target label is weak except
-around head and tail. Therefor pairwise potentials are necessary to solve the
+nine to the cells that are occupied by the snake.  \Figref{snakes_illustration} illustrates the
+principle.
+Local evidence for the target label is weak except
+around head and tail, making this a challenging task. Pairwise potentials are necessary to solve the
 task. The dataset is noise-free in the sense that given the above description, a
@@
 This way, we can get a clear picture how learning times with complexity of the
-inference task.
+inference task. For the subgradient and SZLJSP algorithms, learning time is linear
+in the number of passes ofver the dataset, while chaching and solving of a QP makes
+the learning time depend on the number of iterations in hard-to-predict ways.
 
@@
 the various algorithms easily. We use the dynamic programming algorithm
-implemented in \textsc{OpenGM} for experiments
+implemented in \textsc{OpenGM} for experiments.
+
+We start our experiments with the MNIST dataset.
+The convergence of the primal objective and training set loss in terms of
+passes over the training set is shown in \figref{mnist_iterations}.
+Several trends can be observed: the $n$-slack cutting plane algorithms are converge fastest, with
+the version that recomputes the QP at every step leading the race. The algorithms are
+closely followed by the caching $1$-slack cutting plane algorithm.
+They are followed by the significantly slower SZLJSP algorithm, and finally
+the non-caching $1$-slack cutting plane algorithm and subgradient descent.
+These results are intiutive, as they reflect ``how much work'' each algorithm does
+for each loss-augmented prediction step. More work towards the objective leads to faster convergence.
+This ``more work'' is quantified in \figref{mnist_times}.
+Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow convergence.
+Also, caching does not speed up learing learning on this dataset, as loss-augmented predction
+is as fast as querying the cache.
+
+We now consider datasets with non-trivial inference.
+The results for OCR-small are shown in \figref{ocr_small_iterations} and \figref{ocr_small_time}.
+The trends seen in the figures are similar to what we saw for MNIST above.
+Again, even though inference is non-trivial, the non-caching $1$-slack
+algorithm is slower than the caching one. Again, the plots against passes over the trainingset
+do not reflect the runtime complexity of the $n$-slack approach.
+
+There is one interesting phenomenon to be seen here: The cutting-plane algorithms stop immediately
+when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
+on learning much longer---even longer than it took to get to the desired suboptimality. This is caused
+by a much looser bound given by the dual. Remember that the stopping criterion for both cutting-plane 
+and SZLJSP are given by the duality gap. It seems that while the primal objective converges much
+faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
+This means that in for this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
+in guaranteeing the desired suboptimality.
+\figref{ocr_small_duals} illustrates this point by ploting primal and dual
+objectives for the $1$-slack cutting plane and the SZLJSP algorithms.
 
@@
 \section{Discussion}
+From the experiments above, it is clear there is currently \emph{no single best algorithm}.
+We can say, however, that the stochastic subgradient descent algorithm with
+pegasos step-size always converged slower, both in terms of learning time and
+in passes over the dataset, as the SZLJSP algorithm.
+Together with the fact that stochastic subgradient descent has no good stopping
+criterion, there seems to be no reason to use the stochastic gradient descent.

### 2013-09-22 intro.tex
citations for graz

--- 2013-09-21 intro.tex
@@
 different datasets in this thesis: the object class segmentation datasets
-Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21, and
-NYU V2.
+Graz-02~\citep{marszatek2007accurate} and Pascal VOC 2010~\citep{pascal}, and the semantic segmentation datasets
+MSRC-21~\citep{shotton2006textonboost}, and NYU V2~\citep{SilbermanECCV12}.
 %

### 2013-09-22 partial_supervision/paper.tex
citations for graz

@@
 
-\subsection{Partially Supervised Image Segmentation on Graz 02}
+\subsection{Partially Supervised Image Segmentation on Graz-02}
 
 We evaluate the performance of the proposed algorithm for object-class
-segmentation on the challenging Graz-02 dataset.  This dataset contains 1096
+segmentation on the challenging Graz-02 dataset~\citep{marszatek2007accurate}.  This dataset contains 1096
 images of three object classes, bike, car and person.  Each image may contain
@@
 We evaluated the proposed object-class segmentation method on the challenging
-Graz02 dataset. While not reaching the performance of methods using strong
+Graz-02 dataset. While not reaching the performance of methods using strong
 supervision, our result can work as a baseline for further research into weakly

### 2013-09-22 structured_prediction/paper.tex
phi/psi
fix psi/Psi/phi/Phi mess :D

@@
 \begin{equation}\eqlabel{loss_augmentation}
-    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+.
+    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y) - \theta^T \Phi(x\hoch{i}, y\hoch{i}]_+.
 \end{equation}
@@
 to work in a stochastic setting (see \citet{ratliff2007online}).\pagebreak\\
-Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
+Given a model through $\Phi$ and a set of parameters $\theta$, a subgradient
 considering a single training example $(x\hoch{i}, y\hoch{i})$
@@
 \begin{align*}
-    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\numberthis\\
-    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
+    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\numberthis\\
+    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y)
 \end{align*}
@@
 \begin{equation}
-    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
+    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i})\right]
 \end{equation}
@@
                 \State
-                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
-                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
+                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})] $
+                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
                 \State \hspace{-3mm}$\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
@@
                         \text{s.t. }& \text{for } i=1,\dots,k\ \forall \hat{y} \in \W_i:\\
-                    &\theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
+                    &\theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
                     \end{align*}
@@
             \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
-             &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
+             &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
             \end{align*}
@@
                 \State
-                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
+                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
             \EndFor
             \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
-            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
+            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] $
         \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
@@
             \State Perform loss-augmented prediction on sample $i$:
-            \Statex[2]   $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \Delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$
+            \Statex[2]   $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \Delta(y\hoch{i}, \hat{y}) - \theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})]$
             \State Compute parameter and loss updates based on sample $i$:

### 2013-09-22 exact_learning/paper.tex
call fusion moves fusiom moves everywhere.

@@
 
-As a fast under-generating inference algorithm, we used $\alpha$-expansion
-moves based on non-submodular graph-cuts using Quadratic Pseudo-Boolean
-Optimization (QPBO)~\citep{rother2007optimizing}.  This move-making strategy
-can be seen as a simple instantiation of the more general framework of fusion
-moves, as introduced by \citet{lempitsky2010fusion}
+As a fast under-generating inference algorithm, we used fusion
+moves~\citep{rother2007optimizing, lempitsky2010fusion}. Fusion moves are a local
+search procedure, using moves that are generated using an auxiliar binary problem,
+which is solved using QPBO~\cite{rother2007optimizing}.
 
@@
 %TODO implementation details here?
-The move-making algorithm using QPBO is implemented with the help of the QPBO-I
+The we implement fusion moves with the help of the QPBO-I
 method provided by \citet{rother2007optimizing}.

### 2013-09-22 partial_supervision/paper.tex
call fusion moves fusiom moves everywhere.

@@
 larger datasets, without the need of time-consuming manual annotation on
-pixel-level.
+pixel level.
 
@@
 to the instances contained in the bag. During learning, only presence of object
-classes is needed as bag level supervision. By learning $f_{I}$, we are then
+classes is needed as bag-level supervision. By learning $f_{I}$, we are then
 able to predict for individual segments whether they contain the object class
@@
 To measure the performance of our algorithm, we use a dataset that not only contains
-image level annotation, but also pixel level annotation of object. This allows
+image-level annotation, but also pixel-level annotation of object. This allows
 us to judge the success of learning on instance level.
@@
     \end{tabularx}
-    \caption{Bag level performance of various MIL algorithms on the standard Musk
+    \caption{Bag-level performance of various MIL algorithms on the standard Musk
     datasets. All but MIK provide instance-level labeling.\tablabel{muskacc}}
@@
 
-We described a way to extent bag level predictions made by the multi-instance
+We described a way to extent bag-level predictions made by the multi-instance
 kernel method to instance level while remaining competitive with the

### 2013-09-22 pystruct/paper.tex
call fusion moves fusiom moves everywhere.

@@
 includes support for using {\sc OpenGM}~\citep{kappes2013comparative}, {\sc
-LibDAI}~\citep{Mooij_libDAI_10}, QPBO fusion moves~\citep{rother2007optimizing},
+LibDAI}~\citep{Mooij_libDAI_10}, fusion moves~\citep{rother2007optimizing, lempitsky2010fusion},
 and {\sc AD$^3$}~\citep{martins2011augmented}. It also includes an interface to
@@
 are declared to be symmetric with respect to the direction of an edge, the last
-is antisymmetric. The inference method used is QPBO-fusion moves.  Line 5
+is antisymmetric. We use fusion moves for inference.  Line 5
 creates a \texttt{learner} object that will learn the parameters for the given

### 2013-09-22 exact_learning/paper.tex
use past tense when discussing literature.

@@
 learning was first analyzed by \citet{finley2008training}.
-\citet{finley2008training} show convergence results for under-generating and
+\citet{finley2008training} showed convergence results for under-generating and
 over-generating inference procedures, meaning methods that find suboptimal, but
@@
 respectively.
-\citet{finley2008training} demonstrate that over-generating approaches---in
+\citet{finley2008training} demonstrated that over-generating approaches---in
 particular linear programming (LP)---perform best on the considered model.
@@
 learning using duality.
-In particular, \citet{hazan2010primal} demonstrate the performance of their
+In particular, \citet{hazan2010primal} demonstrated the performance of their
 model on an image denoising task, where it is possible to learn a large number

### 2013-09-22 intro.tex
use past tense when discussing literature.

@@
 framework. \citet{meshi2010learning, komodakis2011efficient}, and
-\citet{hazan2010primal} approach the problem using duality, and formulate
+\citet{hazan2010primal} approached the problem using duality, and formulate
 learning and inference as a joint optimization problem.
 \citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and
-\citet{krahenbuhlparameter} formulate learning structured prediction as
+\citet{krahenbuhlparameter} formulated learning structured prediction as
 optimizing a \emph{prediction engine}, that takes into account all aspects of
@@
 
-\citet{nowozin2010parameter} provide a detailed evaluation of different aspects
+\citet{nowozin2010parameter} provided a detailed evaluation of different aspects
 of learning object class segmentation, that is somewhat orthogonal to this
@@
 pairwise potentials for conditional maximum likelihood learning of
-tree-structured CRFs.  \citet{nowozin2010parameter} also compare conditional
+tree-structured CRFs.  \citet{nowozin2010parameter} also compared conditional
 maximum likelihood learning with maximum margin learning, finding little
@@
 on chain CRFs, where efficient inference is possible.
-More recently, \citet{lucchi2013learning} propose a novel algorithm for
+More recently, \citet{lucchi2013learning} proposed a novel algorithm for
 efficiently learning structured prediction for semantic segmentation, using

### 2013-09-22 itm/paper.tex
use past tense when discussing literature.

@@
 %
-\citet{barber2006kernelized} construct a soft clustering by using a parametric model of $p(Y\given X)$.
+\citet{barber2006kernelized} constructed a soft clustering by using a parametric model of $p(Y\given X)$.
 %
@@
 tree-based approach in which the largest edge is removed until the desired
-number of components is reached.  \citet{zahn1971graph} refines this criterion
+number of components is reached.  \citet{zahn1971graph} refined this criterion
 by cutting edges that are longer than other edges in the vicinity. This

### 2013-09-22 main.tex
fix some equation numbering.
corrections luisa
visualizations for experiment datasets.
use past tense when discussing literature.

@@
-\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=.5cm]{scrbook}
+\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=1cm]{scrbook}
 %\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=calc]{scrbook}
@@
 \usepackage[german,english]{babel}
+\usepackage[export]{adjustbox}
+
 %\renewcommand*\sfdefault{avant}
@@
 
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
+
 \newcommand{\pystruct}{\textsc{PyStruct}\xspace}
@@
 \newcommand{\given}{\, | \,}
-\newcommand{\B}[1]{\textbf{#1}}
+\newcommand{\B}[1]{\mathbf{#1}}
 
@@
 prediction process are optimized directly using empirical risk minimization. The work of
-\citet{stoyanov2011empirical}, used loopy belief propagation as their inference
+\citet{stoyanov2011empirical} used loopy belief propagation as their inference
 algorithm and the optimization is carried out simply using gradient descent on

### 2013-09-22 nyu/paper.tex
use past tense when discussing literature.

@@
 The task of dense semantic annotation of 3D maps has seen an increased interest
-in recent years.  Early work includes \citet{nuchter2008towards}, who combine
+in recent years.  Early work includes \citet{nuchter2008towards}, who combined
 6D SLAM, surface annotation, and object recognition to build semantically
@@
 semantics of the scenes and produce a dense labeling on image level.
-\citet{senguptaurban} extend the approach to produce a volumetric
+\citet{senguptaurban} extended the approach to produce a volumetric
 reconstruction of the scene, together with a dense semantic labeling of the
@@
 \citet{ladicky2009associative}, which is similar in spirit to our approach,
-but uses Potts potentials together with cross-validation to adjust parameters.
+but used Potts potentials together with cross-validation to adjust parameters.
 
@@
 optimized with support relations and structure classes.
-\citet{SilbermanECCV12} use a complex pipeline, employing significant domain knowledge.
+\citet{SilbermanECCV12} used a complex pipeline, employing significant domain knowledge.
 In this work, on the other hand, we try to learn all relevant domain specific features
@@
 
-\citet{couprie-iclr-13} approach the task of semantic segmentation of structure classes
+\citet{couprie-iclr-13} approached the task of semantic segmentation of structure classes
 in RGB-D using the paradigm of convolutional neural networks, extending
 previous work of \citet{farabet-pami-13} and \citet{schulz2012learning}.
-Similar to our approach, \citet{couprie-iclr-13} combine the output of a
+Similar to our approach, \citet{couprie-iclr-13} combined the output of a
 pixel-based, low-level learning algorithm with an independent unsupervised
@@
 
-\citet{stueckler2013} extends the approach of \citet{stuckler2012semantic} to
-a real-time system for online learning and prediction of semantic classes. Their
-use a GPU implementation of random forests, and integrate 3D scene information
-in an online fashion. They evaluate their approach on the dataset of
-\citet{SilbermanECCV12} with promising results.
-We use the implementation of random forest provided by
-\citet{stueckler2013}, but instead of integrating predictions over time, we
-focus on exploiting the structure within a single frame.
+\citet{stueckler2013} extended the approach of \citet{stuckler2012semantic} to
+a real-time system for online learning and prediction of semantic classes.
+Their method use a GPU implementation of random forests, and integrate 3D scene
+information in an online fashion. They evaluated their approach on the dataset
+of \citet{SilbermanECCV12} with promising results.  We use the implementation
+of random forest provided by \citet{stueckler2013}, but instead of integrating
+predictions over time, we focus on exploiting the structure within a single
+frame.
 

### 2013-09-22 partial_supervision/paper.tex
use past tense when discussing literature.

@@
 In this, we follow the work of \citet{liconvex2010} and \citet{zha2008joint}, who not
-only learn $f_{MI}$, but also $f_{I}$.
+only learned $f_{MI}$, but also $f_{I}$.
 
@@
 The basic principle of the multi-instance kernel is similar to a soft-max over instances in
-each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. \citet{gaertner2002multi} show that the multi-instance kernel
+each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. \citet{gaertner2002multi} showed that the multi-instance kernel
 is able to separate bags if and only if the original kernel on instances is able to separate the underlying concepts.
@@
 
-\citet{zhou2009multi} explicitly address the fact that instances are not
+\citet{zhou2009multi} explicitly addressed the fact that instances are not
 independent within a bag , leading to an algorithm that can take advantage of
@@
 
-\citet{liconvex2010} compute likelihood ratios for instances, giving a new
+\citet{liconvex2010} computed likelihood ratios for instances, giving a new
 convex formulation of the multi-instance problem. Using these likelihood

### 2013-09-22 pystruct/paper.tex
use past tense when discussing literature.

@@
 publicly available---many applications are based on the non-free implementation
-of~\citet{joachims2009cutting}.
+of \citet{joachims2009cutting}.
 In this section, we introduce our implementation, \pystruct, which aims at

### 2013-09-22 structured_prediction/paper.tex
hack around spacing in align in algorihtms
use past tense when discussing literature.

@@
 Most current approaches use linear functions, with some notable exceptions:
-\citet{peng2009conditional} introduce a conditional random field for sequence
+\citet{peng2009conditional} introduced a conditional random field for sequence
 prediction where single-node potentials are computed using neural networks,
-while \citet{nowozin2011decision} propose a conditional random field for image
+while \citet{nowozin2011decision} proposed a conditional random field for image
 processing, where all potentials are computed using decision trees.
@@
                     \vspace{-15mm}
-                    \begin{align*}
-                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle \argmin_{\theta, \xi_1, \dots, \xi_k}& \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
-                        \text{s.t. }& \text{for } i=1,\dots,k\ \forall \hat{y} \in \W_i:\\
-                    &\theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
-                    \end{align*}
+                    \begin{flalign*}
+                        \qquad\qquad(\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle \argmin_{\theta, \xi_1, \dots, \xi_k}& \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i&\\
+                        \text{s.t. }& \text{for } i=1,\dots,k\ \forall \hat{y} \in \W_i:&\\
+                                    &\theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i&
+                    \end{flalign*}
                 \EndIf
@@
             \vspace{-5mm}
-            \begin{align*}
-            (\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi}&\frac{||\theta||}{2}^2 + C \xi\\
-            \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
-             &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
-            \end{align*}
+            \begin{flalign*}
+                \quad\,\,(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi}&\frac{||\theta||}{2}^2 + C \xi&\\
+                \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:&\\
+                            &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi&
+            \end{flalign*}
             \For {i=1, \dots, k}

### 2013-09-23 evaluation/paper.tex
fixes von riccarda, ein paar plots fuer auswertung

@@
 the feature to lie between $0$ and $1$.
-
+%
+\enlargethispage{15mm}
 The model we use for multi-class classification is the Crammer-Singer
@@
 The convergence of the primal objective and training set loss in terms of
-passes over the training set is shown in \figref{mnist_iterations}.
+passes over the training set is shown in \Figref{mnist_iterations}.
 Several trends can be observed: the $n$-slack cutting plane algorithms are converge fastest, with
@@
 for each loss-augmented prediction step. More work towards the objective leads to faster convergence.
-This ``more work'' is quantified in \figref{mnist_times}.
+This ``more work'' is quantified in \Figref{mnist_times}.
 Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow convergence.
@@
 
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \includegraphics[width=\linewidth]{evaluation/images/letters_small}&%
+    \includegraphics[width=\linewidth]{evaluation/images/letters_small_time}
+    \end{tabu}
+\caption{%
+   Convergence of the primal objective on OCR-small. 
+\figlabel{ocr_small_results}}
+\end{figure}
+
 We now consider datasets with non-trivial inference.
-The results for OCR-small are shown in \figref{ocr_small_iterations} and \figref{ocr_small_time}.
+The results for OCR-small are shown in \Figref{ocr_small_results}.
 The trends seen in the figures are similar to what we saw for MNIST above.
@@
 do not reflect the runtime complexity of the $n$-slack approach.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=.5\linewidth]{evaluation/images/letters_small_dual}
+\end{center}
+\caption{%
+   Convergence of the primal objective on OCR-small. 
+\figlabel{ocr_small_dual}}
+\end{figure}
 
@@
 in guaranteeing the desired suboptimality.
-\figref{ocr_small_duals} illustrates this point by ploting primal and dual
+\Figref{ocr_small_dual} illustrates this point by ploting primal and dual
 objectives for the $1$-slack cutting plane and the SZLJSP algorithms.
+
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \includegraphics[width=\linewidth]{evaluation/images/letters_big}&%
+    \includegraphics[width=\linewidth]{evaluation/images/letters_big_time}
+    \end{tabu}
+\caption{%
+   Convergence of the primal objective on OCR-large. 
+\figlabel{ocr_large_results}}
+\end{figure}
+
+The results for the OCR-large dataset are shown in \figref{ocr_large_results}
+The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
+in terms of passes through the dataset, but $n$-slack cutting plane is impractical slow
+with respect to runtime. The SZLJSP algorithm converges fast intially, but the $1$-slack cutting plane
+is faster at high precision and faster in certifying the desired duality gap.
 

### 2013-09-23 partial_supervision/paper.tex
fixes von riccarda, ein paar plots fuer auswertung

@@
 work with strong supervision: a pixel-wise labeling of training images. In this
-chapter, we investigate a method that works with annotation that is much
+chapter we investigate a method that works with annotation which is much
 easier to obtain: whole image labels.
@@
 segmentation.  While these are close to our work, there are several important
-distinctions.  We address the task of object-class segmentation, that concerns
+distinctions.  We address the task of object-class segmentation which concerns
 object categories, while semantic segmentation approaches often focus on
@@
 simple texture features.
-Additionally, in contrast to, for example, \citet{vezhnevets2011weakly}, our approach is in
+In contrast to, for example, \citet{vezhnevets2011weakly}, who build a joint model over all
+segmentation decisions, our approach is in
 principle applicable to large datasets, the regime where weak annotation is
@@
 
-In our approach, we work with a set of candidate segments, generated using
+In our approach we work with a set of candidate segments, generated using
 constrained parametric min-cuts~\citep{carreira2010constrained}.  The procedure yields
-segments that are overlapping, object-like regions, which serve as candidates
+segments that are overlapping, object-like regions which serve as candidates
 for object locations.
@@
 We formulate weakly supervised multi-class image segmentation as a
-multi-instance problem, based on these candidate segments.  In multi-instance
-learning~\citep{dietterich1997solving}, each training example is given as a
+multi-instance problem based on these candidate segments.  In multi-instance
+learning~\citep{dietterich1997solving} each training example is given as a
 multi-set of instances, called a bag.  Each instance is represented as a
@@
 
-A bag is labeled positive if it contains at least one positive example, and
-negative otherwise.  During training, only the labels of the training bags, not
+A bag is labeled positive if it contains at least one positive example and
+negative otherwise.  During training only the labels of the training bags, not
 of the instances inside the bags, are known.  The goal is to learn a classifier
 for unseen bags. 
-Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we
+Formally, let $\mathcal{X}$ be the set of instances. To simplify notation we
 assume that bags are simply sets, not multi-sets.
@@
 go a step further and apply multi-instance learning to the task of object-class
-segmentation in natural images, by also classifying instances, not only bags.
-In this, we follow the work of \citet{liconvex2010} and \citet{zha2008joint}, who not
+segmentation in natural images by also classifying instances, not only bags.
+In this we follow the work of \citet{liconvex2010} and \citet{zha2008joint}, who not
 only learned $f_{MI}$, but also $f_{I}$.
 
-In our model, each image forms a bag, while the candidate segments correspond
-to the instances contained in the bag. During learning, only presence of object
+In our model each image forms a bag, while the candidate segments correspond
+to the instances contained in the bag. During learning only presence of object
 classes is needed as bag-level supervision. By learning $f_{I}$, we are then
@@
 of interest, thereby obtaining a segmentation of the object.
-To measure the performance of our algorithm, we use a dataset that not only contains
+To measure the performance of our algorithm we use a dataset that not only contains
 image-level annotation, but also pixel-level annotation of object. This allows
@@
 object-class labeling from the Pascal VOC dataset~\citep{pascal}. This provides
-a ranking of candidate segments, according to how ``object-like'' they are,
+a ranking of candidate segments according to how ``object-like'' they are,
 which allows for selecting only a limited number of very object-like segments.
@@
 challenge~\citep{li2010object}. A similar approach to whole-object segment
-proposals was investigated by \citet{endres2010category}.
+proposals was investigated by \citet{endres2010category}, but they did not compare
+their results with the state-of-the-art approach of \citep{carreira2010constrained}.
 %Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
@@
 object boundaries, we use the efficient multi-instance
-kernels~\citep{gaertner2002multi}.  Multi-instance kernels are a form of set
+kernel~\citep{gaertner2002multi}.  Multi-instance kernels are a form of set
 kernels %TODO add reference to earlier work -> convolution kernel
@@
 \begin{equation}\eqlabel{mi_kernel}
-k_{MI}(X,X') := \sum_{x \in X, x' \in X'} k^p(x,x'),
+k_{MI}(X,X') := \sum_{x \in X, x' \in X'} k_I^p(x,x'),
 \end{equation}
@@
 the RBF-kernel $k_{\text{rbf}}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
-are again RBF-kernels, we not consider $p$ explicitly in the following.
+are again RBF-kernels, we do not consider $p$ explicitly in the following.
 
@@
 \begin{equation}
-    y_x = \argmax_{y \in Y} \#\{S_i | p \in S_i \land y_{S_i}=y \},
+    y_x = \argmax_{y \in Y} \#\{S_i | x \in S_i \land y_{S_i}=y \},
 \end{equation}
@@
     \end{tabularx}
-    \caption{Bag-level performance of various MIL algorithms on the standard Musk
+    \caption{Bag-level accuracy (in percent) of various MIL algorithms on the standard Musk
     datasets. All but MIK provide instance-level labeling.\tablabel{muskacc}}

### 2013-09-23 evaluation/paper.tex
more figures in evaluation.

@@
 This ``more work'' is quantified in \Figref{mnist_times}.
-Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow convergence.
-Also, caching does not speed up learing learning on this dataset, as loss-augmented predction
-is as fast as querying the cache.
+Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow
+convergence.  Also, caching does not speed up learing learning on this dataset,
+as loss-augmented predction is as fast as querying the cache.
 
@@
 
-The results for the OCR-large dataset are shown in \figref{ocr_large_results}
+The results for the OCR-large dataset are shown in \Figref{ocr_large_results}
 The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
@@
 is faster at high precision and faster in certifying the desired duality gap.
+%TODO initially 1-slack is beaten by ssgd
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree}&%
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_time}
+    \end{tabu}
+\caption{%
+   Convergence of the primal objective on the multi-label yeast dataset with a tree-shaped model.
+\figlabel{yeast_tree}}
+\end{figure}
+
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \includegraphics[width=\linewidth]{evaluation/images/scene_tree}&%
+    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_time}
+    \end{tabu}
+\caption{%
+   Convergence of the primal objective on the multi-label scene dataset with a tree-shaped model.
+   \figlabel{scene_tree}}
+\end{figure}
+
+Next, we consider the multi-label task, starting with the tree-shaped models.
+The results for the yeast and scene datasets are shown in \Figref{yeast_tree} and \Figref{scene_tree},
+respectively. The caching $1$-slack algorithm is much faster than the non-caching one on both datasets.
+The caching $1$-slack algorithm is the fastest algorithm to achieve the desired primal suboptimality.
+On the scene dataset, the other cutting-plane base algorithms are second fastest, while they
+lag behind the stoachastic algorithms on the larger yeast dataset. Again, initial convergence is
+faster for the stochastic algorithms.
+
+The results when using the full model are shown in \Figref{yeast_full} and \Figref{scene_full}.
+Here, inference is much more complex than in the previous tasks.
 

### 2013-09-23 intro.tex
feedback from a real expert....

@@
 In computer vision research, the goal is to automatically extract information
-from a given image of image sequence.  In particular discerning semantic
+from a given image or image sequence.  In particular discerning semantic
 information, that is interpreting an image, is a prominent research topic.
@@
 \emph{context} is one of the most important cues when it comes to classifying
-objects, and therefore understanding scenes. Consequently, we target dense
+objects, and therefore understanding scenes. Therefore, we target dense
 labeling of scenes, taking object relations into account.  The task of densely
@@
 
-We choose this task in particular for several reasons:
+We choose this task in particular for following reasons:
 \begin{itemize*}
@@
 ``tags'' is available in practically unlimited quantities, semantic annotation
-on pixel level is only available through laborious manual annotation. 
+on pixel level is scarce and only available through laborious manual annotation. 
 Chapter~\ref{ch:semi_supervised} introduces an approach to cope with this
 shortcoming of object class segmentation approaches by introducing a method to
-find segmentations automatically from image level annotations.
+learn segmentation automatically from image level annotations.
 
@@
 directly minimize a loss of interest. In particular, CRFs
-allow to combine different cues, possibly produced by other approaches, in a
+allow to combine different cues, possibly produced using different paradgims in a
 principled manner. One of the main difficulties with CRF approaches to computer
 vision problems is that context in images is usually represented as a
-neighborhood graph.  These graphs, by nature, contain many cycles, making
+neighborhood graph of pizels or superpixels.  These graphs, by nature, contain many cycles, making
 inference intractable in general. Consequently, learning algorithms have to
@@
     to learn a $1$-slack SSVM.
+\item Learning 3D relations of semantic structure categories for indoor scenes.
+    We extend the CRF approach to learning spacial relations from RGB-D data and
+    improve upon the state-of-the-art in semantic annotation on the NYU V2 dataset
+    of indoor scenes.
 \end{itemize}
@@
 We introduce an algorithm for semi-supervised learning of object class
-segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the problem of
-annotating training data for semantic segmentation.
+segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the difficulty of
+obtraining annotated training data for semantic segmentation.
 

### 2013-09-23 itm/paper.tex
feedback from a real expert....

@@
 So not only is clustering one of the fundamental problems in machine learning, it is also
-an important building block for the methods in this work.
+an important building block for other methods in this work.
 %
@@
 over. We overcome this by proposing an efficient approximate optimization
-based on the Euclidean minimum spanning tree algorithm. 
+based on Euclidean minimum spanning trees.
 %
@@
 which makes it very easy to use in practice.
+%
+The non-parametric entropy estimate we
+are using is applicable only if the data distribution is absolute continuous and
+therefore can not be applied if the data lies on a proper submanifold.
+We show how to overcome this obstacle in practice by using an estimate of the
+intrinsic dimensionality of the data.
 %
@@
     \item Showing the practicality on a number of synthetic and real datasets.
-    \item Extending the clustering to data on submanifolds by estimating implicit dimensionality.
+    \item Extending the clustering to data on submanifolds by estimating intrinsic dimensionality.
 \end{itemize}

### 2013-09-23 main.tex
fixes von riccarda, ein paar plots fuer auswertung
feedback from a real expert....

@@
 Tag der Promotion:\\
-Erscheinungsjahr: 2013\\
+Erscheinungsjahr:\\
 \end{spacing}
@@
 mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
-ist ein Segmentierungsmodell zu lernen. Nach unserem Wissen ist dies der erste
-Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.
+ist ein Segmentierungsmodell zu lernen.
 
@@
 %
-Wir untersuchen verschiedene Lernalgorithmen des strukturierten Lernens, insbesondere
-im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
-Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
-Mit den vorgestellten Methoden bringen wir den Stand der Kunst
-auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
-von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.
+Wir untersuchen verschiedene Lernalgorithmen des strukturierten Lernens,
+insbesondere im Zusammenhang mit approximativer Vorhersage. Wir zeigen, dass es
+m\"oglich ist trotz des Vorhandenseins von Kreisen in den betrachteten Nachbarschaftsgraphen
+exakte strukturierte Modelle zur Bildsegmentierung zu lernen.  Mit den
+vorgestellten Methoden bringen wir den Stand der Kunst auf zwei komplexen
+Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz von
+RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.
 

### 2013-09-23 nyu/paper.tex
feedback from a real expert....

@@
-%\begin{abstract}
-%    We present a structured learning approach to semantic annotation of RGB-D
-%    images. 
-%    Our method learns to reason about spatial relations of objects and
-%    fuses low-level class predictions to a consistent interpretation of a
-%    scene. Our model incorporates color, depth and 3D scene features, on which
-%    an energy function is learned to directly optimize object class prediction using
-%    the loss-based maximum-margin principle of structural support vector machines.
-%    %With the proliferation of cheap % TODO urghs formulation
-%    %high-quality RGB-D sensors, it became possible to capture more and more
-%    %data not only as color images, but with additional depth information. The
-%    %depth channel provides important clues not only for scene structure, but
-%    %also for predicting semantic object classes.
-%    %We demonstrate how CRF learning, an approach
-%    %that has shown great promise for RGB images, can exploit depth information
-%    %to improve semantic interpretation of the scene.
-%    We evaluate our approach on the NYU V2 dataset of indoor scenes, a challenging
-%    dataset covering a wide variety of scene layouts and object classes. We
-%    hard-code much less information about the scene layout into our model then previous
-%    approaches, and instead learn object relations directly from the data. We
-%    find that our conditional random field approach improves upon previous
-%    work, setting a new state-of-the-art for the dataset.
-%\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \chapter{Learning Depth-Sensitive Conditional Random Fields}\label{ch:nyu}
@@
 For robots to perform varied tasks in unstructured environments, understanding their
-surroundings is essential. We formulate the problem of semantic annotation of maps
-as a dense labeling of RGB-D images into semantic classes. Dense labeling of objects
+surroundings is essential. In this chapter, we look at the semantic annotation of maps
+as a dense labeling of \mbox{RGB-D} images into semantic classes. We formulate the problem
+as learning a CRF over a superpixel segmentation of the \mbox{RGB-D} image, producing a labeling
+that takes 3D layout into account. Dense labeling of objects
 and structure classes allows for a detailed reasoning about the scene.
 
-In this work, we propose the use of random forests combined with conditional
-random fields (CRFs) to perform robust estimation of structure classes in RGB-D
-images. The CRF is learned using a structural support vector machine,
-allowing it to integrate the noisy low-level information produced by a pixel-based
-random forest to a consistent interpretation of the scene.
+%We use of random forests to obtain a noisy low-level estimate of structure class on pixel level.
+%These estimates serve as the unary input to the learned CRF model which integrates
+%these low-level cues to a consistent interpretation of the scene.
 
-We thereby extend the success of learned CRF models for
-semantic segmentation in RGB images to the domain of 3D scenes.
+We thereby extend the success of learned CRF models for semantic segmentation
+in RGB images as considered in Chapter~\ref{exact_learning} to the domain of 3D scenes.
 Our emphasis lies on exploiting the additional depth and 3D information in all
@@
 are grouped together using a superpixel approach, for which we extend previous
-superpixel algorithms from the RGB to the RGB-D domain.
+superpixel algorithms from the RGB to the \mbox{RGB-D} domain.
 We then build a geometric model of the scene, based on the neighborhood graph
@@
 2D plane of the image, but also to model object distances and surface angles in
-3D, using a point cloud generated from the RGB-D image. The process is depicted
+3D, using a point cloud generated from the \mbox{RGB-D} image. The process is depicted
 in \Figref{teaser}.
@@
 Recent approaches for indoor semantic annotation of maps mostly focused on
-RGB-D images, which are now easy to obtain using structured light sensors.
+\mbox{RGB-D} images, which are now easy to obtain using structured light sensors.
 \citet{stuckler2012semantic}, for example, used a Random Forest model to obtain a dense
@@
 \citet{silberman2011indoor} introduced the NYU Depth Dataset V1 indoor dataset,
-which consisted of a large variety of densely annotated indoor scenes, captured as RGB-D images.
-Their work also introduced a baseline method for semantic segmentation of RGB-D
+which consisted of a large variety of densely annotated indoor scenes, captured as \mbox{RGB-D} images.
+Their work also introduced a baseline method for semantic segmentation of \mbox{RGB-D}
 image, which is based on a CRF over superpixels, with unary potentials given by interest point descriptors.
@@
 designed for the dataset, potentials were either directly set by hand or estimated using empirical
-frequencies. This is in contrast to our work, which applies structured prediction techniques
+frequencies. This is in contrast to our approach which applies structured prediction techniques
 to learn potentials automatically to optimize predictive performance.
@@
 \citet{ren2012rgb} evaluated the design of features for semantic labeling of
-RGB-D data, and use a hierarchical segmentation to provide context. While they
+\mbox{RGB-D} data, and use a hierarchical segmentation to provide context. While they
 also define a CRF on superpixels, their model is again not learned, but a
@@
 \citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
-using in this work.  Their focus is on inferring support relations in indoor
+using here.  Their focus is on inferring support relations in indoor
 scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
@@
 \citet{SilbermanECCV12} used a complex pipeline, employing significant domain knowledge.
-In this work, on the other hand, we try to learn all relevant domain specific features
+In our approach, on the other hand, we try to learn all relevant domain specific features
 directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
@@
 \citet{couprie-iclr-13} approached the task of semantic segmentation of structure classes
-in RGB-D using the paradigm of convolutional neural networks, extending
+in \mbox{RGB-D} using the paradigm of convolutional neural networks, extending
 previous work of \citet{farabet-pami-13} and \citet{schulz2012learning}.
@@
 We use the GPU implementation provided by \citet{stueckler2013}\footnote{\url{https://github.com/deeplearningais/curfil}}.
-The input for training are the full RGB-D images, transformed to Lab
+The input for training are the full \mbox{RGB-D} images, transformed to Lab
 color space. Each tree in the forest uses training pixels only from a subset of
@@
 indoor scenes.  The dataset comes with a detailed annotation of 1449 indoor
-RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
+\mbox{RGB-D} images belonging to a wide variety of indoor scenes, categorized into 26
 scene classes.  The annotation contains four semantic structural classes:
@@
 We introduce a CRF formulation for semantic segmentation
-of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
+of structure classes in \mbox{RGB-D} images. We base our model in the output of an efficient GPU implementation
 of random forest, and model spatial neighborhood using a superpixel-based approach.
@@
 order potentials~\citep{ladicky2009associative}.
-Another possible line of future work is to combine our approach with a more
+Another possible line of future investigation is to combine our approach with a more
 task-specific one, directly including support plane assumptions into the model,
@@
 
-While this work did not explicitly address real time application, the random
+We did not explicitly address real time application, the random
 forest implementation that we build upon allows for real-time

### 2013-09-23 partial_supervision/paper.tex
feedback from a real expert....

@@
 While we do not reach the accuracy of competing fully supervised approaches,
-our efficient, semi-supervised method is potentially able to scale to much
+our efficient, weakly supervised method is potentially able to scale to much
 larger datasets, without the need of time-consuming manual annotation on
@@
 
-Recently, several approaches have been proposed for semi-supervised semantic
+Recently, several approaches have been proposed for weakly supervised semantic
 segmentation.  While these are close to our work, there are several important

### 2013-09-23 exact_learning/paper.tex
feedback from a real expert....
copyedit nyu, be clear about improving upon the state of the art on msrc.

@@
 field model using a structured support vector machine exactly for this task.
-We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. % and nyu?
+We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. Our
+approach improves upon the state-of-the art on the MSRC-21 dataset, and is
+competative with comparable approaches on the Pascal VOC dataset.% and nyu?
 
-The contribution of this chapter is threefold:
+The contribution of this chapter are:
 \begin{itemize}
@@
         application, even in the presence of loopy graphs.
+    \item We show that even using a strong approximate inference procedure can
+        improve upon the state-of-the art on the MSRC-21 dataset.
     % also fusion moves is pretty good?
@@
 
-\section{Efficient Exact Cutting Plane Training of SSVMs}\seclabel{learning}
+\section[Efficient Exact Cutting Plane Training of SSVMs]{Efficient Exact\linebreak Cutting Plane Training of SSVMs}\seclabel{learning}
 
@@
 Results on the MSRC-21 dataset are shown in \Tabref{msrcacc}.
-We find that our model is on par with state-of-the-art approaches, implying
-that our model is realistic for this task. In particular, our results are comparable to those of
-\citet{lucchi2013learning}, who use a stochastic subgradient method with working sets.
+We find that our model is improves upon state-of-the-art approaches. In
+particular, our results improve upon to those of \citet{lucchi2013learning},
+who use a stochastic subgradient method with working sets.
 Their best model takes 583s for training, while training our model exactly takes 1814s.

### 2013-09-23 intro.tex
copyedit nyu, be clear about improving upon the state of the art on msrc.

@@
 semantic image segmentation exactly, even though the involved factor graphs
-contain many loops. We demonstrate our algorithm on the Pascal VOC 2010 and MSRC-21 datasets,
-and provide state-of-the-art results on the MSRC-21 dataset.
+contain many loops. We demonstrate our algorithm on the Pascal VOC 2010, where we are competative
+with comparable approaches, and MSRC-21 datasets where we improve upon the state-of-the-art.
 

### 2013-09-23 main.tex
copyedit nyu, be clear about improving upon the state of the art on msrc.

@@
 
-Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
+Der dritte Teil der Arbeit untersucht einen der am weitesten
 verbreiteten Ans\"atze zur semantischen Segmentierung und
@@
 popular learning algorithms for structural support vector machines, together
-with a thorough evaluation of their properties in the presence of approximate
-inference. Our software provides a solid basis for future research into learning,
-inference and models for computer vision, by providing extensive examples and
+with a thorough evaluation of their properties, in particulare when using approximate
+inference. Our software provides a foundation for future research into learning,
+inference and models for computer vision by providing extensive examples and
 benchmarks.
@@
 factor graph. Our methods achieve competitive performance with similar
- methods on the Pascal VOC 2010 dataset, and yield state-of-the art
+ methods on the Pascal VOC 2010 dataset, and improve upon state-of-the art
  results on the MSRC-21 dataset.
@@
 
-We also presented a novel approach for clustering based on information theoretic
-principles. Our algorithm improves upon methods from the literature on a wide
-range of datasets. This indicates that in the task of extracting
-superpixels, we can also hope to achieve better results than the $k$-means based
-SLIC algorithm that we used.
+We also presented a novel approach for clustering based on information
+theoretic principles. Our algorithm improves upon methods from the literature
+in finding pre-defined classes on a wide range of datasets. This indicates that
+in the task of extracting superpixels, we can also hope to achieve better
+results than the $k$-means based SLIC algorithm that we used.
 
@@
 learn object class segmentation for complex object classes from image-level
-annotations alone.
+annotations alone. Our approach is formulated using multiple instance learning
+over a set of candidate segments. We demonstrated the feasibility and effectiveness
+of our approach on the challenging Graz-02 dataset of street scenes.
 

### 2013-09-23 nyu/paper.tex
copyedit nyu, be clear about improving upon the state of the art on msrc.

@@
 We take a super-pixel based approach to semantic segmentation.
-Our superpixel generation is based on the SLIC algorithm~\citep{achanta2012slic}. We extend the
-standard SLIC algorithm, which works on the Lab space, to also include 
-depth information. The resulting algorithm is a localized $k$-Means in Lab-D-XY space.
+Our superpixel generation is based on the SLIC
+algorithm~\citep{achanta2012slic}, as in Chapter~\ref{exact_learning}. We
+extend the standard SLIC algorithm, which works on the Lab space, to also
+include depth information. The resulting algorithm is a localized $k$-Means in
+Lab-D-XY space.
 Our implementation is publicly available through the scikit-image library\footnote{\url{http://scikit-image.org}}.
@@
 scene classes.  The annotation contains four semantic structural classes:
-structure, floor, furniture and prop. There is an additional ``void'' class,
-which is used for object boundaries and hard-to-annotate regions. We follow the
+structure, floor, furniture and prop. As in the MSRC-21 and Pascal VOC
+datasets, there is an additional ``void'' class, which is used for object
+boundaries and hard-to-annotate regions. We follow the
 literature in excluding these pixels completely from the evaluation.
@@
 Our approach is implemented using our \textsc{PyStruct}
-library\footnote{\url{http://pystruct.github.io}}.  The code for our
-experiments will be released upon acceptance. We use \textsc{scikit-learn}~\citep{pedregosa2011scikit}
-for $k$-means clustering and the SVM baseline.
-
+library introduced in \Secref{pystruct}.
+%
 All hyper parameters were adjusted using 5-fold cross-validation. The
@@
 possible performance of 95\% average class accuracy on the validation set.
+
+We observed that the linear programming relaxation
+often found an integer solution, without the need for a branch-and-bound procedure. We also found
+that using fusion moves alone produced inferior results.
 

### 2013-09-23 pystruct/paper.tex
copyedit nyu, be clear about improving upon the state of the art on msrc.
starting blabla on superpixels and CRFS

@@
-\section{Casting Structured Prediction into Software}\label{sec:api}
+\section{Casting Structured Prediction into Software}\label{sec:pystruct}
 
@@
 the main focus of this work. The example shows how to learn an $n$-slack support
-vector machine on a superpixel-based CRF on the popular Pascal dataset. We use
-unary potentials generated using TextonBoost
-from \citet{krahenbuhl2012efficient}. The superpixels are generated using SLIC~\citep{achanta2012slic}.%
-\footnote{The preprocessed data can be found at \url{http://www.ais.uni-bonn.de/download/datasets.html}.}
+vector machine on a superpixel-based CRF on the popular Pascal dataset. Details of the experiment
+can be found in Chapter~\ref{exact_learning}.
 Each sample (corresponding on one entry of the list \texttt{X}) is represented as a

### 2013-09-23 structured_prediction/paper.tex
starting blabla on superpixels and CRFS

@@
 The detailed procedure as given by \citet{lacoste2012block} is shown in Algorithm~\ref{alg_szljsp}.
+
+\section{CRFs for Semantic Segmentation}
+Conditional Random Fields have been established as an important tool in many
+areas of computer vision. Applications include dense stereo, optical flow, 
+inpainting, de-noising, image editing, low-level segmentation and semantic segmentation.
+
+There are broadly two types of models that are used: models in which each pixel
+in an image is represented as a node, and models in which pixels are first
+grouped together into superpixels and each superpixel is represented as a node.
+In pixel-based CRFs edges are usually introduced between adjacent pixels, that is using
+either 4-neighborhoods or 8-neighborhoods. Superpixel-based approaches introduce
+edges between adjacent superpixels, that is those that share a boundary.
+%TODO images of superpixels!
+%
+The motivation for using superpixels is that it is often easy to group nearby
+pixels that belong to the same object just based on local color or texture cues.
+While it is hard to even define what makes a sensible segmentation of an image,
+it is much easier to define, and also to compute an \emph{over-segmentation}.
+The only requirement for an over-segmentation is that each segment only
+contains pixels belonging to the same entity.
+
+Using superpixels instead of pixels has large computational advantages.
+A model that uses superpixels contains much fewer variables than one that works on pixel-level.
+Typical sizes for superpixels are in the hundereds of pixels, leading to a hundred-fold decrease
+in complexity for prediction. There are also semantic advantages. Using superpixels
+can constitute a form of regularization, including prior knowledge about the structure
+of images. Modeling interactions between superpixels also allows more distant parts of an image
+to interact, instead of modeling long-term relations only indirectly over neighbors.
+Finally, modeling decisions on a superpixel-level allows for more context for a local decision,
+while semantic decisions on pixel-level are not meaningful near object boundaries.
+The main disadvantage of superpixel-based approaches is that the initial oversegmentation can
+not be influenced by later reasoning. This can be a problem if fine structures need to be
+segmented for which local evidence does not suffice.
+
+% about algorithms, slic?
+% hierarchical\dots has superpixels and pixels?
+%  pro/con superpixels?
+
+% submodular/non-submodular? higher order?
+
+In early models, parameters were usually set by hand. Recent works mostly use a
+maximum margin based formulation based on maximum margin Markov
+models~\cite{taskar2003max} or structural ssvms~\cite{tsochantaridis2006large}
+for parameter learning~\cite{lucchi2013learning, lucchi2011spatial}.
+
+Most CRF-based image segmentation approaches use loopy neighborhood graphs,
+leading to intractable inference problems during learning -- see
+\citep{nowozin2010parameter} for a tree based approach. There has been much
+research in learning structured prediction models using approximate inference.
+The influence of approximate inference on structured support vector machine
+training was first analyzed in  \citep{finley2008training}.
+
+More recently, \cite{meshi2010learning}, \cite{hazan2010primal} and
+\cite{komodakis2011efficient} introduced formulations for joint inference and
+learning using duality.
+
+\citep{finley2008training} shows that using certain Linear Programming (LP)
+relaxations, generalization guarantees can be preserved. Unfortunately,
+standard LP solvers are too slow for many applications, in particular semantic
+segmentation as formulated in this work.
+
+Most related to our approach is \citep{szummer2008learning}, which utilize graph cut
+based optimization for inference.
+Instead of using the classical approach of \cite{boykov2001fast}, which is restricted
+to submodular potentials, we use a $\alpha$-expansion move-making algorithm based on binary QPBO
+\cite{rother2007optimizing}. While relying on local search loses the guarantees of binary QPBO,
+this method is fast and we find it to perform well in practice.

### 2013-09-23 structured_prediction/paper.tex
more crf blabla

--- 2013-09-22 structured_prediction/paper.tex
@@
 The detailed procedure as given by \citet{lacoste2012block} is shown in Algorithm~\ref{alg_szljsp}.
+
+\section{CRFs for Semantic Segmentation}
+Structured models based on factor graphs, as described in \Secref{factor_graphs} are often
+referred to as Conditional Random Fields (CRFs), pointing to their
+probabilistic interpretation.
+Conditional Random Fields have been established as an important tool in many
+areas of computer vision. Applications include dense stereo, optical flow, 
+inpainting, de-noising, image editing, low-level segmentation and semantic segmentation.
+
+Most conditional random fields only apply unary and pairwise potential functions. In this case, FIXME equation 4.4
+becomes
+\begin{equation}\eqlabel{energy_crf}
+    g(x, y) = \sum_{v \in V} \psi_v(x, y_v) + \sum_{(v, w) \in E} \psi_{v,w}(x, y_{v}, y_{w}).
+\end{equation}
+Here $V$ enumerates variables, and $E\subset V \times V$ is a set of edges,
+representing pairwise factors.
+
+There are broadly two types of models that are used: models in which each pixel
+in an image is represented as a variable, and models in which pixels are first
+grouped together into superpixels and each superpixel is represented as a variable.
+In pixel-based CRFs edges are usually introduced between adjacent pixels, that is using
+either 4-neighborhoods or 8-neighborhoods. Superpixel-based approaches introduce
+edges between adjacent superpixels, that is those that share a boundary.
+%TODO images of superpixels!
+%TODO looooppsss!!!!
+
+The motivation for using superpixels is that it is often easy to group nearby
+pixels that belong to the same object just based on local color or texture cues.
+While it is hard to even define what makes a sensible segmentation of an image,
+it is much easier to define, and also to compute an \emph{over-segmentation}.
+The only requirement for an over-segmentation is that each segment only
+contains pixels belonging to the same entity.
+
+Using superpixels instead of pixels has large computational advantages.  A
+model that uses superpixels contains much fewer variables than one that works
+on pixel-level.  Typical sizes for superpixels are in the hundereds of pixels,
+leading to a hundred-fold decrease in the number of variables. There are also
+semantic advantages. Using superpixels can constitute a form of regularization,
+including prior knowledge about the structure of images. Modeling interactions
+between superpixels also allows more distant parts of an image to interact,
+instead of modeling long-term relations only indirectly over neighbors.
+Finally, modeling decisions on a superpixel-level allows for more context for a
+local decision, while semantic decisions on pixel-level are not meaningful near
+object boundaries.  The main disadvantage of superpixel-based approaches is
+that the initial oversegmentation can not be influenced by later reasoning.
+This can be a problem if fine structures need to be segmented for which local
+evidence does not suffice.
+
+% about algorithms, slic?
+% hierarchical\dots has superpixels and pixels?
+
+% submodular/non-submodular? higher order?
+
+Conditional random fields were first used in the context of semantic
+segmentation by \citet{he2004multiscale}.  Who used contrastive divergence for
+learning parameters. Many other early models, such as the one of
+\citet{shotton2006textonboost}, did not learn parameters at all, but used
+\emph{contrast-sensitive Potts potential}. These potentials have a smoothing
+effect over labels by encouraging neighboring variables to take the same value.
+The penalty for taking different values is dependent on color contrast between
+the pixels.
+
+This smoothing approach was improved upon by \citet{kohli2009robust}, who
+introduces higher order potentials to enforce label consistency within larger
+regions. Later, \citet{ladicky2009associative} proposed a hierarchical model,
+including the higher order potentials of \citet{kohli2009robust}, and
+additional lateral and hierarchical connections. A similar approach, using
+superpixels as the finest resolution, and additionally modelling object class
+coocurrences was suggested by \citet{gonfaus2010harmony}.  However, all these
+models did not learn the potentials of the CRF model in a principled manner.
+The above approaches all learned unary potentials using a non-structured
+approach, for example SVMs or boosting, and then set pairwise and higher order
+potentials in the model by hand.
+
+\citet{szummer2008learning} on the other hand used a structured support vector
+machine approach to learn unary and pairwise parameters. They use the classical
+graph cut approach of \citet{boykov2001fast} for inference.  However, graph-cut
+inference is only applicable to submular energies, and therefore severly
+restricts the expressivity of the resulting model.
+
+\citet{lucchi2011spatial} investigated the importance of global constraints,
+using an approach similar to \citet{szummer2008learning}, but also learning
+global interactions---however, these did not improve performance, compared to
+simply including global descriptors into local classifiers.
+
+The problem of approximate inference was addressed elegantly by
+\citet{yao2012describing}, who learn a joint model for scene classification,
+object localization and semantic segmentation.  Their work is based on
+\citet{hazan2010primal}, who integrate learning and inference in a joint
+optimization problem.
+
+%TODO where does that leave us now?
+In Chapter~\ref{exact_learning} and Chapter~\ref{nyu}, we investigate
+superpixel-based conditional random fields learned with conditional random
+fields. Our emphasis is on understanding the learning, and how to cope with the
+complexity of inference in loopy models.

### 2013-09-24 evaluation/paper.tex
add new plots...

@@
     \begin{tabu} to \linewidth{@{}XX@{}}
-    \includegraphics[width=\linewidth]{evaluation/images/letters_small}&%
-    \includegraphics[width=\linewidth]{evaluation/images/letters_small_time}
+    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log}&%
+    \includegraphics[width=\linewidth]{evaluation/images/letters_small_log_time}
     \end{tabu}
@@
     \begin{tabu} to \linewidth{@{}XX@{}}
-    \includegraphics[width=\linewidth]{evaluation/images/letters_big}&%
-    \includegraphics[width=\linewidth]{evaluation/images/letters_big_time}
+    \includegraphics[width=\linewidth]{evaluation/images/letter_big_log}&%
+    \includegraphics[width=\linewidth]{evaluation/images/letter_big_log_time}
     \end{tabu}
@@
 %TODO initially 1-slack is beaten by ssgd
-\begin{figure}
-    \begin{tabu} to \linewidth{@{}XX@{}}
-    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree}&%
-    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_time}
-    \end{tabu}
-\caption{%
-   Convergence of the primal objective on the multi-label yeast dataset with a tree-shaped model.
-\figlabel{yeast_tree}}
-\end{figure}
 
@@
     \begin{tabu} to \linewidth{@{}XX@{}}
-    \includegraphics[width=\linewidth]{evaluation/images/scene_tree}&%
-    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_time}
+    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log}&%
+    \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log}&%
+    \includegraphics[width=\linewidth]{evaluation/images/scene_full_log_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
     \end{tabu}
 \caption{%
-   Convergence of the primal objective on the multi-label scene dataset with a tree-shaped model.
-   \figlabel{scene_tree}}
+   Convergence of the primal objective on the multi-label datasets. From top to bottom: the scene dataset using a tree model,
+   the scene dataset using a full pairwise model, the yeast dataset using a tree model and the yeast dataset using a full pairwise model.
+\figlabel{yeast_tree}}
 \end{figure}

### 2013-09-24 structured_prediction/paper.tex
add new plots...

@@
 %TODO images of superpixels!
-%TODO looooppsss!!!!
+By nature, both approaches create graphs with a large number of loops. This
+complicates prediction and learning, and is the central topic of
+Chapter~\ref{ch:exact_learning}.
 
@@
 evidence does not suffice.
+
+We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
+our experiments.  It has been shown to provide competative results with a
+minimum of computational complexity.  Algorithmically, SLIC simply computes a
+$k$-means clustering over pixels. Each pixels is represented as 5D point, using
+three color channels and the $x$ and $y$ coordinates in the image.  In a
+post-processing step, small segments are removed. To make clustering of so many
+points feasible, the search for the nearest cluster in $k$-means is restricted
+to a local neighborhood in the image.
 
@@
 restricts the expressivity of the resulting model.
-
 \citet{lucchi2011spatial} investigated the importance of global constraints,

### 2013-09-24 evaluation/paper.tex
add table about datasets

@@
 
+\begin{table}
+    \caption{Summary of Datasets.\tablabel{datasets}}
+    \begin{tabu} to \linewidth{@{}XXXXXX@{}}
+    %\multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\
+    \toprule
+    Dataset    & samples & nodes & edges & dim w & labels\\
+    \cmidrule{1-6}
+    MNIST      & 60000   & 1     & none  &       & 10\\
+    OCR-small  & 704     & 3--14 & chain &       & 26\\
+    OCR-large  & 6173    & 3--14 & chain &       & 26\\
+    scene-tree & 1211    & 6     & tree  &       & 2\\
+    scene-full & 1211    & 6     & loopy &       & 2\\
+    yeast-tree & 1500    & 14    & tree  &       & 2\\
+    yeast-full & 1500    & 14    & loopy &       & 2\\
+    snake      & 200     &84--168& loopy &       & 11\\
+    MSRC21     & 276     & 7--113& loopy &       & 21\\
+    Pascal VOC & 964     & 8--112& loopy &       & 21\\
+    \bottomrule
+
+    \end{tabu}
+\end{table}
+
 \section{Experiments}
@@
     \item Stochastic Subgradient Descent using the Pegasos schedule: %TODO
-    \item Stochastic Subgradient Descent using a momentum heuristic:
-    %\item Stochastic Subgradient Descent using a momentum heuristic: AVERAGING
-    %\item Stochastic Subgradient Descent using a momentum heuristic: AVERAGING
+    %\item Stochastic Subgradient Descent using the Pegasos schedule: AVERAGING
     \item The $1$-slack cutting plane algorithm without constraint caching.
@@
     \begin{tabu} to \linewidth{@{}XX@{}}
+    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
     \includegraphics[width=\linewidth]{evaluation/images/letters_small_log}&%
@@
     \begin{tabu} to \linewidth{@{}XX@{}}
-    \includegraphics[width=\linewidth]{evaluation/images/letter_big_log}&%
-    \includegraphics[width=\linewidth]{evaluation/images/letter_big_log_time}
+    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
+    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log}&%
+    \includegraphics[width=\linewidth]{evaluation/images/letters_big_log_time}
     \end{tabu}
@@
     \begin{tabu} to \linewidth{@{}XX@{}}
+    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
     \includegraphics[width=\linewidth]{evaluation/images/scene_tree_log}&%
@@
     \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
-    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}
     \end{tabu}
 \caption{%
-   Convergence of the primal objective on the multi-label datasets. From top to bottom: the scene dataset using a tree model,
-   the scene dataset using a full pairwise model, the yeast dataset using a tree model and the yeast dataset using a full pairwise model.
-\figlabel{yeast_tree}}
+   Convergence of the primal objective on the multi-label datasets. From top to
+   bottom: the scene dataset using a tree model, the scene dataset using a full
+   model, the yeast dataset using a tree model, and the yeast dataset using a
+   full model.
+\figlabel{multi-label-results}}
 \end{figure}

### 2013-09-24 evaluation/paper.tex
compute feature numbers for all models ^^

@@
 We use a simple chain model with a single constant pairwise feature.
-This means the unary potential has $16 \cdot 8 \cdot 26$ parameters, one for each input
+This means the unary potential has $16 \cdot 8 \cdot 26 = 768$ parameters, one for each input
 feature and output class. The pairwise potential consists of a matrix of
@@
 The scene dataset has six labels, 294 input features, 1211 training samples and
-1196 test samples. This leads to $204 \cdot 6$ parameters for the unary potentials,
-$3 \cdot 5 \cdot 4$ parameters for the full pairwise potentials (four for each edge), and
-$5 \cdot 4$ parameters for the pairwise potentials of the tree-shaped model.
+1196 test samples. This leads to $204 \cdot 6 = 1224$ parameters for the unary potentials,
+$3 \cdot 5 \cdot 4 = 60$ parameters for the full pairwise potentials (four for each edge), and
+$5 \cdot 4 = 20$ parameters for the pairwise potentials of the tree-shaped model.
 Using four parameters for each edge is a slight over-parametrization that
@@
 The yeast dataset has 14 labels, 103 features, 1500 training samples and 971
-test samples. The resulting numbers of parameters are $14 \cdot 103$ parameters
-for the unary potentials, $7 \cdot 13 \cdot 4$ parameters for the full pairwise
-potential, and $13 \cdot 4$ parameters for the tree-shaped model.
+test samples. The resulting numbers of parameters are $14 \cdot 103=1442$ parameters
+for the unary potentials, $7 \cdot 13 \cdot 4=364$ parameters for the full pairwise
+potential, and $13 \cdot 4=52$ parameters for the tree-shaped model.
 
@@
 Using the one-hot encoding of the input, this leads to $9 \dot 5 = 45$ unary features.
-With 11 output classes, the unary potential has $11 \cdot 9 \cdot 5$ parameters.
+With 11 output classes, the unary potential has $11 \cdot 9 \cdot 5=495$ parameters.
 Features for the pairwise potentials are constructed by concatenating the features
@@
 of these parts will be non-zero for any given edge.  With $45 \cdot 4$ edge
-features, the pairwise potentials has $45 \cdot 4 \cdot 11^2$ parameters.
+features, the pairwise potentials have $45 \cdot 4 \cdot 11^2 =21780$ parameters.
 
@@
 %TODO will it?
-The Pascal VOC has \ldots training and \ldots test images, each having
-around 100 superpixels. There are 20 object categories, plus background. The % TODO check!
+The Pascal VOC has \ldots training and \ldots validation images, each having
+around 100 superpixels. There are 20 object categories, plus background. The % TODO check! refer to other part
 MSRC dataset has \ldots training and \ldots test images, also segmented in
@@
 Each superpixel is represented as an output variable, with the ground truth obtained
-by majority vote over the pixel belonging to the superpixel.
+by majority vote over the pixel belonging to the superpixel. We removed all
+superpixels which were mostly ``void'', leading to some samples haven much
+fewer than 100 variables.
 Pairwise potentials were introduced for each pair of neighboring superpixels.
@@
 their pairwise potentials.
-As unary potentials, we use the TextonBoost probabilities provided by \citet{krahenbuhl2012efficient}.
-We average probabilities inside each superpixel, yielding 21 input features for both datasets.
-We also used the same pairwise features for the two datasets: a constant feature,
-a color contrast feature, and a feature encoding relative vertical position between superpixels.
-Overall, the models for both datasets had $21 \cdot 21 + 21^2 \cdot 4$ features.
-It is possible to assume symmetric potentials for this dataset, leading to less
-pairwise features. We found both parametrizations yielded similar results.
+As unary potentials, we use the TextonBoost probabilities provided by  % FIXME this !!
+\citet{krahenbuhl2012efficient}.  We average probabilities inside each
+superpixel, yielding 21 input features for both datasets.  We also used the
+same pairwise features for the two datasets: a constant feature, a color
+contrast feature, and a feature encoding relative vertical position between
+superpixels.  Overall, the models for both datasets had $21 \cdot 21 + 21^2
+\cdot 4$ features.  It is possible to assume symmetric potentials for this
+dataset, leading to less pairwise features. We found both parametrizations
+yielded similar results.
 
@@
     \cmidrule{1-6}
-    MNIST      & 60000   & 1     & none  &       & 10\\
-    OCR-small  & 704     & 3--14 & chain &       & 26\\
-    OCR-large  & 6173    & 3--14 & chain &       & 26\\
-    scene-tree & 1211    & 6     & tree  &       & 2\\
-    scene-full & 1211    & 6     & loopy &       & 2\\
-    yeast-tree & 1500    & 14    & tree  &       & 2\\
-    yeast-full & 1500    & 14    & loopy &       & 2\\
-    snake      & 200     &84--168& loopy &       & 11\\
+    MNIST      & 60000   & 1     & none  & 7840  & 10\\
+    OCR-small  & 704     & 3--14 & chain & 1444  & 26\\
+    OCR-large  & 6173    & 3--14 & chain & 1444  & 26\\
+    scene-tree & 1211    & 6     & tree  & 1244  & 2\\
+    scene-full & 1211    & 6     & loopy & 1284  & 2\\
+    yeast-tree & 1500    & 14    & tree  & 1494  & 2\\
+    yeast-full & 1500    & 14    & loopy & 1806  & 2\\
+    snake      & 200     &84--168& loopy & 22275 & 11\\
     MSRC21     & 276     & 7--113& loopy &       & 21\\

### 2013-09-24 evaluation/paper.tex
freakingtable !!!!
finish off table

@@
 their pairwise potentials.
-As unary potentials, we use the TextonBoost probabilities provided by  % FIXME this !!
-\citet{krahenbuhl2012efficient}.  We average probabilities inside each
-superpixel, yielding 21 input features for both datasets.  We also used the
-same pairwise features for the two datasets: a constant feature, a color
-contrast feature, and a feature encoding relative vertical position between
-superpixels.  Overall, the models for both datasets had $21 \cdot 21 + 21^2
-\cdot 4$ features.  It is possible to assume symmetric potentials for this
-dataset, leading to less pairwise features. We found both parametrizations
-yielded similar results.
+We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
+for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC21 dataset.
+We use the same three pairwise features for both dataset: a constant, a color
+feature and a feature encoding relative vertical position.
+Overall, this results in $21 \cdot  21 + 3 \cdot 21 \cdot 21 = 1764$ parameters
+for the potentials on Pascal VOC and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
+2646$ parameters for the potentials on the MSRC21 dataset.
+
+%For Pascal VOC we use the probabilities provided by
+%\citet{krahenbuhl2012efficient}, which combine low-level Texton-Boost
+%potentials with bounding-box object detectors.  We average probabilities inside
+%each superpixel, yielding 21 input features for both datasets.  We also used
+%the
+%same pairwise features for the two datasets: a constant feature, a color
+%contrast feature, and a feature encoding relative vertical position between
+%superpixels.  Overall, the models for both datasets had $21 \cdot 21 + 21^2
+%\cdot 4$ features.  It is possible to assume symmetric potentials for this
+%dataset, leading to less pairwise features. We found both parametrizations
+%yielded similar results.
 
@@
     \caption{Summary of Datasets.\tablabel{datasets}}
-    \begin{tabu} to \linewidth{@{}XXXXXX@{}}
+    \begin{tabu} to \linewidth{@{}X[1,l]YX[1,c]X[1,c]YY@{}}
     %\multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\
     \toprule
-    Dataset    & samples & nodes & edges & dim w & labels\\
+    Dataset    & {samples} & nodes & edges & {dim w} & {labels}\\
     \cmidrule{1-6}
@@
     snake      & 200     &84--168& loopy & 22275 & 11\\
-    MSRC21     & 276     & 7--113& loopy &       & 21\\
-    Pascal VOC & 964     & 8--112& loopy &       & 21\\
+    MSRC21     & 276     & 7--113& loopy &  1764 & 21\\
+    Pascal VOC & 964     & 8--112& loopy &  2646 & 21\\
     \bottomrule

### 2013-09-24 nyu/paper.tex
caption below tables

@@
 \begin{table*}[t]
-    \caption{Quantitative comparison of the proposed method with the
-    literature.\tablabel{results}}
 \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}l@{}cccccc}
@@
 \end{tabularx}
-\quad\\The best value in each column is printed in bold$^\dagger$. The upper part of
+    \caption{Quantitative comparison of the proposed method with the
+        literature.
+The best value in each column is printed in bold$^\dagger$. The upper part of
 the table shows contributions by different parts of our pipeline. RF stands for random forest prediction, RF + SP for aggregated
 random forests prediction within superpixels, RF + SP + SVM for an SVM trained on the unary potentials, and RF + SP + CRF is
-our proposed pipeline. We optimized our approach for class average
+our proposed pipeline. We optimized our approach for class average\tablabel{results}
 accuracy.\\
@@
 comparable, as they integrated information over multiple frames, and did not
-measure accuracy for pixels without valid depth measurement.
+measure accuracy for pixels without valid depth measurement.}
 \end{table*}

### 2013-09-24 structured_prediction/paper.tex
caption below tables

@@
 
-\section{CRFs for Semantic Segmentation}
+\section{CRFs for Semantic Segmentation}\seclabel{semantic_segmentation}
 Structured models based on factor graphs, as described in \Secref{factor_graphs} are often

### 2013-09-24 evaluation/paper.tex
caption below tables
snakes and msrc plots

@@
 \begin{table}
-    \caption{Summary of Datasets.\tablabel{datasets}}
     \begin{tabu} to \linewidth{@{}X[1,l]YX[1,c]X[1,c]YY@{}}
@@
     \toprule
-    Dataset    & {samples} & nodes & edges & {dim w} & {labels}\\
+    Dataset    & {samples} & variables & graph & {dim $\theta$} & {labels}\\
     \cmidrule{1-6}
@@
     \bottomrule
-
     \end{tabu}
+    \caption{Summary of datasets used in the evaluation. \tablabel{datasets}}
 \end{table}
@@
     \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
     \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}
@@
 
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_loss}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time_loss}\\
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_loss}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time_loss}
+    \end{tabu}
+\caption{%
+    Snakes
+\figlabel{snakes_results}}
+\end{figure}
+
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
+    \includegraphics[width=\linewidth]{evaluation/images/msrc}&%
+    \includegraphics[width=\linewidth]{evaluation/images/msrc_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/msrc_loss}&%
+    \includegraphics[width=\linewidth]{evaluation/images/msrc_time_loss}\\
+    \includegraphics[width=\linewidth]{evaluation/images/msrc}&%
+    \includegraphics[width=\linewidth]{evaluation/images/msrc_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/msrc_loss}&%
+    \includegraphics[width=\linewidth]{evaluation/images/msrc_time_loss}
+    \end{tabu}
+\caption{%
+    MSRC, Pascal VOC
+\figlabel{msrc_pascal}}
+\end{figure}
+
 \section{Discussion}

### 2013-09-24 pystruct/paper.tex
snakes and msrc plots

@@
 the main focus of this work. The example shows how to learn an $n$-slack support
-vector machine on a superpixel-based CRF on the popular Pascal dataset. Details of the experiment
-can be found in Chapter~\ref{exact_learning}.
+vector machine on a superpixel-based CRF on the Pascal dataset. Details of the experiment
+can be found in Section~\secref{semantic_segmentation}.
 Each sample (corresponding on one entry of the list \texttt{X}) is represented as a

### 2013-09-24 evaluation/paper.tex
vlad first mail

@@
 Our main attention is devoted to the use of conditional random fields for
-semantic segmentation.  We use the Pascal VOC and MSRC datasets. A detailed
-description of the method will be given in Chapter~\ref{ch:exact_learning}.
-%TODO will it?
-The Pascal VOC has \ldots training and \ldots validation images, each having
+semantic segmentation.  We use the Pascal VOC and MSRC datasets.
+The Pascal VOC has 964 training, each having
 around 100 superpixels. There are 20 object categories, plus background. The % TODO check! refer to other part
-MSRC dataset has \ldots training and \ldots test images, also segmented in
-around 100 superpixels each. We use the MSRC-21 dataset, which has 21 target
-classes.
+MSRC-21 dataset has 276 training images, also segmented in around 100 superpixels
+each. There are 21 classes in the MSRC-21 dataset.
 Each superpixel is represented as an output variable, with the ground truth obtained
@@
 Pairwise potentials were introduced for each pair of neighboring superpixels.
-We then removed all superpixels that were labeled as ``void'', together with
-their pairwise potentials.
 We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
@@
 \begin{itemize}
-    \item Stochastic Subgradient Descent using the Pegasos schedule: %TODO
+    \item Stochastic Subgradient Descent (SSGD) using the Pegasos schedule for step-sizes.
     %\item Stochastic Subgradient Descent using the Pegasos schedule: AVERAGING
@@
 \end{itemize}
+
 All algorithm take a $C$ parameter, which we adjusted on a fully trained model
-on a validation set (experiments not reported here) and held constant for all models.
-We found, however, that the algorithms and models are quite robust to the choice of $C$ within
-one or two orders of magnitude.
+on a validation set (experiments not reported here) and held constant for all
+models.  We found, however, that the algorithms and models are quite robust to
+the choice of $C$ within one or two orders of magnitude.
 As stopping criterion, we used a duality gap of $0.1$ when possible, and a
@@
 In this section, we present results on multi-class classification, sequence
-labeling and multi-label prediction with a tree-structured model. In these
+labeling and multi-label prediction. In these
 tasks, exact inference is possible, and we can compute the exact objective of
 the various algorithms easily. We use the dynamic programming algorithm
-implemented in \textsc{OpenGM} for experiments.
+implemented in \textsc{OpenGM} for experiments for the tree and chain models.
+For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
+We found that using the linear programming relaxation on the multi-label
+dataset is often tight, and there was no need to resort to approximate
+inference.
 
@@
 passes over the training set is shown in \Figref{mnist_iterations}.
-Several trends can be observed: the $n$-slack cutting plane algorithms are converge fastest, with
-the version that recomputes the QP at every step leading the race. The algorithms are
-closely followed by the caching $1$-slack cutting plane algorithm.
-They are followed by the significantly slower SZLJSP algorithm, and finally
-the non-caching $1$-slack cutting plane algorithm and subgradient descent.
-These results are intiutive, as they reflect ``how much work'' each algorithm does
-for each loss-augmented prediction step. More work towards the objective leads to faster convergence.
-This ``more work'' is quantified in \Figref{mnist_times}.
-Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow
-convergence.  Also, caching does not speed up learing learning on this dataset,
-as loss-augmented predction is as fast as querying the cache.
+Several trends can be observed: the $n$-slack cutting plane algorithms are
+converge fastest, with the version that recomputes the QP at every step leading
+the race. The algorithms are closely followed by the caching $1$-slack cutting
+plane algorithm.  They are followed by the significantly slower SZLJSP
+algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
+subgradient descent.  These results are intiutive, as they reflect ``how much
+work'' each algorithm does for each loss-augmented prediction step. More work
+towards the objective leads to faster convergence.  This ``more work'' is
+quantified in \Figref{mnist_times}.  Clearly the $1$-slack algorithms to ``to
+much'' work, leading to very slow convergence.  Also, caching does not speed up
+learing learning on this dataset, as loss-augmented predction is as fast as
+querying the cache.
 
@@
 Next, we consider the multi-label task, starting with the tree-shaped models.
-The results for the yeast and scene datasets are shown in \Figref{yeast_tree} and \Figref{scene_tree},
+The results for the yeast and scene datasets are shown in \Figref{multi-label-results},
 respectively. The caching $1$-slack algorithm is much faster than the non-caching one on both datasets.
+
 The caching $1$-slack algorithm is the fastest algorithm to achieve the desired primal suboptimality.
@@
 lag behind the stoachastic algorithms on the larger yeast dataset. Again, initial convergence is
-faster for the stochastic algorithms.
+faster for the stochastic algorithms
 
-The results when using the full model are shown in \Figref{yeast_full} and \Figref{scene_full}.
-Here, inference is much more complex than in the previous tasks.
+%, inference is much more complex than in the previous tasks.
 

### 2013-09-24 exact_learning/paper.tex
vlad first mail

@@
 We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. Our
-approach improves upon the state-of-the art on the MSRC-21 dataset, and is
+approach improves upon the state-of-the-art on the MSRC-21 dataset, and is
 competative with comparable approaches on the Pascal VOC dataset.% and nyu?
@@
     \item We show that even using a strong approximate inference procedure can
-        improve upon the state-of-the art on the MSRC-21 dataset.
+        improve upon the state-of-the-art on the MSRC-21 dataset.
     % also fusion moves is pretty good?
@@
 We used the SLIC implementation provided by \citet{achanta2012slic} to extract superpixels and
-the SIFT implementation in the \texttt{vlfeat} package~\citep{vedaldi08vlfeat}.
+the SIFT implementation in the \textsc{vlfeat} package~\citep{vedaldi08vlfeat}.
 For clustering visual words, piecewise training of unary potentials and the
-approximation to the $\chi^2$-kernel, we made use of the \texttt{scikit-learn}
+approximation to the $\chi^2$-kernel, we made use of the \textsc{scikit-learn}
 machine learning package~\citep{pedregosa2011scikit}.

### 2013-09-24 intro.tex
vlad first mail

@@
 Graz-02~\citep{marszatek2007accurate} and Pascal VOC 2010~\citep{pascal}, and the semantic segmentation datasets
-MSRC-21~\citep{shotton2006textonboost}, and NYU V2~\citep{SilbermanECCV12}.
+MSRC-21~\citep{shotton2006textonboost} and NYU V2~\citep{SilbermanECCV12}.
 %
@@
 promising  results for computer vision applications. Using the paradigm of structural
-support vector machines, it is possible to learn conditional random fields to
+support vector machines~(SSVMs), it is possible to learn conditional random fields to
 directly minimize a loss of interest. In particular, CRFs
@@
 state-of-the-art methods on the NYU V2 datasets.
+\pagebreak
 
@@
 \begin{description}
-    \item[Chapter~\ref{ch:itm}] \emph{Information Theoretic Clustering using
-        Minimum Spanning Trees} Andreas C. M\"uller, Sebastian Nowozin and
-        Christoph H. Lampert. Published in the proceedings of the German
-        Conference on Pattern Recognition.
-    \item[Chapter~\ref{ch:semi_supervised}] \emph{Multi-Instance Methods for
-        Partially Supervised Image Segmentation} Andreas C. M\"uller and Sven
-        Behnke. Published in the proceedings of the IARP Workshop on Partially
-        Supervised Learning.
-    \item[Chapter~\ref{ch:structured_pystruct}] \emph{PyStruct - Structured
-        Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to
-        the Journal of Machine Learning Research, Open Source Software track.
-    \item[Chapter~\ref{ch:exact_learning}] \emph{Learning a Loopy Model for
-        Semantic Segmentation Exactly}. Andreas C. M\"uller and Sven Behnke.
-        Submitted to the International Conference on Computer Vision Theory and
-        Applications.
-    \item[Chapter~\ref{ch:nyu}] \emph{Learning Depth-Sensitive Conditional
-        Random Fields for Semantic Segmentation}. Andreas C. M\"uller and Sven
-        Behnke. Submitted to the International Conference on Robotics and
-        Automation.
+    \item[Chapter~\ref{ch:itm}]%
+        \emph{Information Theoretic Clustering using Minimum Spanning Trees}\\
+        Andreas C. M\"uller, Sebastian Nowozin and Christoph H. Lampert.
+        Published in the proceedings of the German Conference on Pattern
+        Recognition.
+    \item[Chapter~\ref{ch:semi_supervised}]%
+        \emph{Multi-Instance Methods for Partially Supervised Image
+        Segmentation}\\
+        Andreas C. M\"uller and Sven Behnke. Published in the proceedings of
+        the IARP Workshop on Partially Supervised Learning.
+    \item[Chapter~\ref{ch:structured_pystruct}]%
+        \emph{PyStruct - Structured Prediction in Python}\\
+        Andreas C. M\"uller and Sven Behnke. Submitted to the Journal of
+        Machine Learning Research, Open Source Software track.
+    \item[Chapter~\ref{ch:exact_learning}]%
+        \emph{Learning a Loopy Model for Semantic Segmentation Exactly}\\
+        Andreas C. M\"uller and Sven Behnke.  Submitted to the International
+        Conference on Computer Vision Theory and Applications.
+    \item[Chapter~\ref{ch:nyu}]%
+        \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}\\
+        Andreas C. M\"uller and Sven Behnke. Submitted to the International
+        Conference on Robotics and Automation.
 \end{description}

### 2013-09-24 itm/paper.tex
vlad first mail

@@
 \begin{itemize}
-    \item Clustering, as a purely unsupervised methods is on one end of the
+    \item Clustering, as a purely unsupervised method, is on one end of the
         spectrum of algorithms we investigate.  When using per-image or
@@
 Clusters are given by the Voronoi tessellation of the cluster means and
-therefore are always convex.  Another draw-back is the non-determinism of the
+therefore are always convex.  Another drawback is the non-determinism of the
 procedure, caused by the dependence on random initialization.
@@
          I(X, Y)
-        &= D_{\text{KL}}(p(X, Y) \parallel p(X) p(Y)) = H(X) - \sum_{y=1}^k p(Y{=}y) H(X \given Y {=} y)
+        &= D_{\text{KL}}\bigl(p(X, Y) \parallel p(X) p(Y)\bigr) = H(X) - \sum_{y=1}^k p(Y{=}y) H(X \given Y {=} y)
 \end{align}
@@
 \begin{itemize}
-    \item $\displaystyle D_\text{KL}(p(X) \parallel q(X))\, = \int_\mathcal{X} p(X) \ln\left(\frac{p(X)}{q(X)}\right)\,dX$ is the Kullback-Leibler divergence,
-\item $\displaystyle H(X)\, = \int_\mathcal{X} p(X) \ln(p(X))\,dX$ is the differential entropy, and
-\item $\displaystyle H(X \given Y{=}y)\, = \int_\mathcal{X} p(X \given Y {=}y) \ln(p(X \given Y {=} y))\,dX$ is the conditional differential entropy.
+    \item $\displaystyle D_\text{KL}\bigl(p(X) \parallel q(X)\bigr)\, = \int_\mathcal{X} p(X) \ln\left(\frac{p(X)}{q(X)}\right)\,dX$ is the Kullback-Leibler divergence,
+    \item $\displaystyle H(X)\, = \int_\mathcal{X} p(X) \ln\bigl(p(X)\bigr)\,dX$ is the differential entropy, and
+    \item $\displaystyle H(X \given Y{=}y)\, = \int_\mathcal{X} p(X \given Y {=}y) \ln\bigl(p(X \given Y {=} y)\bigr)\,dX$ is the conditional differential entropy.
     \end{itemize}
@@
 \[
-    p(Y\,{=}\,y)= \frac{n_y}{n} \text{,\quad with\quad} n_y = |\{i \mid y_i = y \}|.
+    p(Y\,{=}\,y)= \frac{n_y}{n} \text{,\quad with\quad} n_y = \bigl|\{i \mid y_i = y \}\bigr|.
 \]
@@
 %
-Then $\hat{I}(G) \geq \hat{I}(\mathbf{y}(G))$, by the definition of
+Then $\hat{I}(G) \geq \hat{I}\bigl(\mathbf{y}(G)\bigr)$, by the definition of
 the minimum spanning tree, and equality holds if and only if $G_y$ is the
@@
 exactly and efficiently by searching over all of $\mathcal{G}$.
-This amounts to searching for the edge $e$ that maximizes $\hat{I}(T(\mathbf{x})
-\setminus e)$. The naive algorithm that computes the objective for each edge
+This amounts to searching for the edge $e$ that maximizes $\hat{I}\bigl(T(\mathbf{x})
+\setminus e\bigr)$. The naive algorithm that computes the objective for each edge
 separately has run time that is quadratic in the number of data points.  To
@@
 components equals the number of desired clusters. The overall procedure is
-summarized in Algorithm~\ref{overall}, an illustration can be found in
-\Figref{illustration}. We refer to Algorithm~\ref{overall} as \emph{Information
-Theoretic MST-based (ITM) clustering}.
+summarized in Algorithm~\ref{overall}, which we call \emph{Information
+Theoretic MST-based (ITM) clustering}.  An illustration can be found in
+\Figref{illustration}
 
@@
 in practice than specialized algorithms for euclidean minimum spanning trees, 
-which achieve a better theoretical runtime of $O(n \log(n) \alpha(n))$ \citep{dtb2010}. Here
+which achieve a better theoretical runtime of $O\bigl(n \log(n) \alpha(n)\bigr)$ \citep{dtb2010}. Here
 $\alpha$ is the inverse of the Ackerman function. The dynamic programming solution of
 Algorithm~\ref{overall} has a run time of $O(n)$ per removed edge, leading to an overall
-run time of $O(n \log(n) \alpha(n)+ nk)$. The $O(nk)$ comes from a worst case
+run time of $O\bigl(n \log(n) \alpha(n)+ nk\bigr)$. The $O(nk)$ comes from a worst case
 scenario, in which each step in the hierarchical clustering procedure only
@@
 that the individual clusters are much smaller than the original dataset. In
-this case, the $O(nk)$ factor would improve to $O(n\log(k))$.
+this case, the $O(nk)$ factor would improve to $O\bigl(n\log(k)\bigr)$.
 
@@
 We found this estimate to work robustly and give sensible results for all
-datasets we investigated. As we already used a ball-tree data structure to
+datasets we investigated. As we already used a ball tree data structure to
 build the minimum spanning trees, we can reuse this structure to compute $r_k$.
@@
 We implemented the ITM clustering procedure as well as MeanNN in Python.  We
-used the $k$-Means implementation available in the scikit-learn
+used the $k$-Means implementation available in the \textsc{scikit-learn}
 library~\citep{pedregosa2011scikit}. The source
@@
 Single Link &   -- &    Yes    &    $O(n\log n)$\\
-ITM & $\displaystyle \sum_{y=0}^k d p(y) \log(\bar{L}_y  ) + p(y) \log{p(y)}$ & Yes & $O(\alpha(n) n \log n + nk)$\\
+ITM & $\displaystyle \sum_{y=0}^k d p(y) \log(\bar{L}_y  ) + p(y) \log{p(y)}$ & Yes & $O\bigl(\alpha(n) n \log n + nk\bigr)$\\
 \bottomrule
@@
 \Figref{qualitative} shows qualitative results on three synthetic datasets.
-For well separated, convex clusters, all four algorithms produce the same
-clustering (see top row).  If the structure of the data is more complex, the
+For well separated, convex clusters, all four algorithms produce equivalent
+clusterings (see top row).  If the structure of the data is more complex, the
 advantage of the proposed method is apparent.  Note that there was no need to

### 2013-09-24 main.tex
freakingtable !!!!
vlad first mail

@@
 \usepackage{tabularx}
+\usepackage{siunitx}
 \usepackage{tabu}
@@
 \setlength{\tabcolsep}{2pt}
+
+%\newcolumntype Y{S[group-four-digits=true, per-mode=symbol, round-precision=0, table-number-alignment=right]}
+\newcolumntype Y{X[c,1]{S[group-four-digits=true, table-format=2.2, table-number-alignment=right]}}
+\tabucolumn Y
 
@@
 and recognition of object classes in natural images, that learns a segmentation
-model solely from annotation in form of presence and absence of object classes
+model solely from annotation in the form of presence and absence of object classes
 in images.
@@
 of many loops in the underlying neighborhood graphs.
-The introduced methods provide results advancing the state-of-the art on two
+The introduced methods provide results advancing the state-of-the-art on two
 complex benchmark datasets for semantic segmentation, the MSRC-21 Dataset of RGB
@@
 
-I have been lucky to be a part of the scikit-learn community, whose developers
+I have been lucky to be a part of the \textsc{scikit-learn} community, whose developers
 have been a constant source of encouragement and inspiration to me. My
@@
 
-\chapter{Conclusion and Future Direction}
-\section{Summary}
+\chapter{Conclusion}
+%\section{Summary}
 In this thesis, we explored the use of structured prediction methods for semantic
@@
 factor graph. Our methods achieve competitive performance with similar
- methods on the Pascal VOC 2010 dataset, and improve upon state-of-the art
+ methods on the Pascal VOC 2010 dataset, and improve upon state-of-the-art
  results on the MSRC-21 dataset.
@@
 results than the $k$-means based SLIC algorithm that we used.
-
+\enlargethispage{1cm}
 As manual annotation of images for learning semantic segmentation and object

### 2013-09-24 nyu/paper.tex
vlad first mail

@@
 All hyper parameters were adjusted using 5-fold cross-validation. The
-hyper parameters of the random forests were found using the \texttt{hyperopt}
+hyper parameters of the random forests were found using the \textsc{hyperopt}
 framework of \cite{bergstra2011algorithms}. For the CRF model, the only
@@
 using a standard unstructured SVM with rescaled loss already advances the mean
-accuracy decidedly above the previous state-of-the art. We
+accuracy decidedly above the previous state-of-the-art. We
 attribute this mostly to the ability of the SVM to exploit correlation between
@@
 that is learned using the SSVM approach. By explicitly modeling 3D relations in
-a fully learned framework, we improve the state-of-the art on the NYU V2
+a fully learned framework, we improve the state-of-the-art on the NYU V2
 dataset for semantic annotation of structure classes.

### 2013-09-24 partial_supervision/paper.tex
vlad first mail

@@
 
+\enlargethispage{10mm}
 We formulate weakly supervised multi-class image segmentation as a
@@
 feature vector $x$ and a label $y$.
-
 A bag is labeled positive if it contains at least one positive example and
@@
 only learned $f_{MI}$, but also $f_{I}$.
-
 In our model each image forms a bag, while the candidate segments correspond
@@
 proposals was investigated by \citet{endres2010category}, but they did not compare
-their results with the state-of-the-art approach of \citep{carreira2010constrained}.
+their results with the state-of-the-art approach of \citet{carreira2010constrained}.
 %Brendel and Todorovic~\citet{NIPS2010_0122} also use a set of candidate segments is used, though all of these are kept throughout the whole 
@@
 \subsection{Multi-Instance Methods}
-Multi-instance learning was formally introduced in \citet{dietterich1997solving}.
+Multi-instance learning was formally introduced by \citet{dietterich1997solving}.
 Since then, many algorithms were proposed to solve the multi-instance learning problem
@@
 combine these.  Since we do not make use of the ground truth segmentation
-during training, we cannot learn an optimal combination as in
-\citet{li2010object} but perform a simple majority vote instead.  We merge
+during training, we cannot learn an optimal combination as
+\citet{li2010object} did, but perform a simple majority vote instead.  We merge
 segments into pixel-level class labels by setting the label $y_x$ of a pixel
-$x$ according to
+$x$ according to:
 \begin{equation}
-    y_x = \argmax_{y \in Y} \#\{S_i | x \in S_i \land y_{S_i}=y \},
+    y_x = \argmax_{y \in Y} |\{S_i | x \in S_i \land y_{S_i}=y \}|,
 \end{equation}
-where $Y$ is the set of possible object classes, $S_i$ enumerates all segments within an
-image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
+Here $Y$ is the set of possible object classes, $S_i$ enumerates all segments within an
+image and $y_{S_i}$ is the label of segment $S_i$. In other words each pixel is
 assigned the class with the highest number of class segments containing it.

### 2013-09-24 evaluation/paper.tex
more corrections by vladn :)

@@
 with specialized algorithms, it nevertheless provides an initial insight into structured
-prediction algoriths.
+prediction algorithms.
 We choose the classical MNIST dataset of handwritten digits, consisting of
@@
 letters long, and each letter is represented as a binary image of size $16
-\time 8$, with a toal of $6877$ words.
-The dataset is devided into ten folds. We consider two setups: learning
+\time 8$, with a total of $6877$ words.
+The dataset is divided into ten folds. We consider two setups: learning
 on one fold, and testing on the remaining nine folds, following \citet{taskar2003max},
@@
 in which each example can be associated with more than one class. In other
-words, the algorihm must decide for each sample and for each class whether the
+words, the algorithm must decide for each sample and for each class whether the
 sample belongs to that class or not.
@@
 We also consider a different model, where pairwise potentials are only
-introduced between specific nodes. Build a tree over the binary variables by
+introduced between specific nodes. We build a tree over the binary variables by
 computing the Chow-Liu Tree~\citep{chow1968approximating} of the targets. While
@@
 \subsection{2D Grid CRF (Snakes)}
-The snakes dataset is a synthetic dataset, where samples are labeld 2D grids.
+The snakes dataset is a synthetic dataset, where samples are labeled 2D grids.
 It was introduced in \citet{nowozin2011decision} to demonstrate the importance
@@
 over the training set.  Both are important to practitioners, as they give
-important insight into the workins of the algorithm.  The actual runtime is
+important insight into the working of the algorithm.  The actual runtime is
 certainly the most important factor, but also highly influenced by
@@
 inference task. For the subgradient and SZLJSP algorithms, learning time is linear
-in the number of passes ofver the dataset, while chaching and solving of a QP makes
+in the number of passes over the dataset, while caching and solving of a QP makes
 the learning time depend on the number of iterations in hard-to-predict ways.
@@
 passes over the training set is shown in \Figref{mnist_iterations}.
-Several trends can be observed: the $n$-slack cutting plane algorithms are
+Several trends can be observed: the $n$-slack cutting plane algorithms
 converge fastest, with the version that recomputes the QP at every step leading
@@
 algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
-subgradient descent.  These results are intiutive, as they reflect ``how much
+subgradient descent.  These results are intuitive, as they reflect ``how much
 work'' each algorithm does for each loss-augmented prediction step. More work
@@
 much'' work, leading to very slow convergence.  Also, caching does not speed up
-learing learning on this dataset, as loss-augmented predction is as fast as
+learning learning on this dataset, as loss-augmented prediction is as fast as
 querying the cache.
@@
 Again, even though inference is non-trivial, the non-caching $1$-slack
-algorithm is slower than the caching one. Again, the plots against passes over the trainingset
+algorithm is slower than the caching one. Again, the plots against passes over the training set
 do not reflect the runtime complexity of the $n$-slack approach.
@@
 On the scene dataset, the other cutting-plane base algorithms are second fastest, while they
-lag behind the stoachastic algorithms on the larger yeast dataset. Again, initial convergence is
+lag behind the stochastic algorithms on the larger yeast dataset. Again, initial convergence is
 faster for the stochastic algorithms

### 2013-09-24 pystruct/paper.tex
more corrections by vladn :)

@@
 The source code is shown in Listing 1.
-Lines 1-3 declare a model using parametric edge potentials for arbitrary graphs.
-Here \texttt{class\_weight} re-weights the hamming loss according to inverse class
+Lines 1--5 declare a model using parametric edge potentials for arbitrary graphs.
+Here \texttt{class\_weight} re-weights the Hamming loss according to inverse class
 frequencies. The parametric pairwise interactions have three features: a

### 2013-09-24 structured_prediction/paper.tex
more corrections by vladn :)

@@
 with regularization parameter $C$.\pagebreak\\
-Here, $\ell$ is a hinge-loss like upper bound
+Here, $\ell$ is a hinge-loss-like upper bound
 on the empirical $\Delta$-risk:
 \begin{equation}\eqlabel{loss_augmentation}
-    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y) - \theta^T \Phi(x\hoch{i}, y\hoch{i}]_+.
+    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y) - \theta^T \Phi(x\hoch{i}, y\hoch{i})]_+.
 \end{equation}
@@
 considering a single training example $(x\hoch{i}, y\hoch{i})$
-can becomputed simply by solving the loss-augmented prediction problem:
+can be computed simply by solving the loss-augmented prediction problem:
 \begin{align*}
@@
         \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
-        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for } i=1,\dots,k$
+        \State $\W_i \leftarrow \varnothing, \xi_i \leftarrow 0 \text{ for } i=1,\dots,k$
         \Repeat

### 2013-09-25 evaluation/paper.tex
more results, discussion for experiments

@@
 the learning time depend on the number of iterations in hard-to-predict ways.
+All algorithms were run on a single core i7 processor.
 
@@
 
-The results for the OCR-large dataset are shown in \Figref{ocr_large_results}
+The results for the OCR-large dataset are shown in \Figref{ocr_large_results}. It was not feasible to run the $n$-slack
+algorithm when solving the QP at every step here.
 The trends are the same as for the smaller dataset: $n$-slack and caching $1$-slack cutting plane are very fast
@@
 is faster at high precision and faster in certifying the desired duality gap.
-%TODO initially 1-slack is beaten by ssgd
 
@@
 The results for the yeast and scene datasets are shown in \Figref{multi-label-results},
-respectively. The caching $1$-slack algorithm is much faster than the non-caching one on both datasets.
+respectively. The caching $1$-slack algorithm is much faster than the
+non-caching one on both datasets. While the two variants of the $n$-slack algorithm are equally
+fast with respect to passes over the training data, solving the QP every 100 steps is among the
+fastest algorithms, while solving the QP every step is among the slowest.
 
@@
 lag behind the stochastic algorithms on the larger yeast dataset. Again, initial convergence is
-faster for the stochastic algorithms
+faster for the stochastic algorithms.
 
@@
 we can still consider the relaxed task, which will be solved exactly in the
-majority of cases (AD$^3$ sometimes fails to find the primal solution to the LP
-relaxation). %TODO is this happening?
+majority of cases (AD$^3$ can fail to find the primal solution to the LP
+relaxation).
 When using fusion moves, there is no obvious interpretation of the produced
 primal objective, other than that it provides a lower bound of the actual
-objective. Nevertheless, we find it informative to analyze the behavior of this lower bound.
+objective. Nevertheless, we find it informative to analyze the behavior of this
+lower bound.
 
@@
 the training set.  We explicitly do not consider the relaxed problem
-here---instead we round possible fractional results.
-While the training error does not directly reflect the objective, it provides
-a measure of how effective the ``prediction machine'' given by the learned model
-together with the inference algorithm is.
+here---instead we round possible fractional results. While the training error
+does not directly reflect the objective, it provides a measure of how effective
+the ``prediction machine'' given by the learned model together with the
+inference algorithm is.
 
@@
     \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_loss}&%
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time_loss}\\
     \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3}&%
@@
     \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_loss}&%
-    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time_loss}\\
-    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo}&%
-    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time}\\
-    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_loss}&%
-    \includegraphics[width=\linewidth]{evaluation/images/snakes_qpbo_time_loss}
+    \includegraphics[width=\linewidth]{evaluation/images/snakes_ad3_time_loss}
     \end{tabu}
 \caption{%
-    Snakes
+    Results on the snakes dataset. The top four plots show results using fusion
+    move inference, the bottom four plots using AD$^3$. The training set error shows
+    that none of the algorithms is able to solve the task using fusion moves.
+    The objective values for fusion move inference are only approximate, giving
+    an upper bound on the true primal objective, while the objective for AD$^3$
+    is the true objective for the relaxed task.
 \figlabel{snakes_results}}
@@
     \includegraphics[width=\linewidth]{evaluation/images/msrc_time_loss}\\
-    \includegraphics[width=\linewidth]{evaluation/images/msrc}&%
-    \includegraphics[width=\linewidth]{evaluation/images/msrc_time}\\
-    \includegraphics[width=\linewidth]{evaluation/images/msrc_loss}&%
-    \includegraphics[width=\linewidth]{evaluation/images/msrc_time_loss}
+    \includegraphics[width=\linewidth]{evaluation/images/pascal}&%
+    \includegraphics[width=\linewidth]{evaluation/images/pascal_time}\\
+    \includegraphics[width=\linewidth]{evaluation/images/pascal_loss}&%
+    \includegraphics[width=\linewidth]{evaluation/images/pascal_time_loss}
     \end{tabu}
 \caption{%
-    MSRC, Pascal VOC
+    Objective and traning set loss on the segmentation task. The top four plots
+    show results for the MSRC-21 dataset, the bottom four plots for the Pascal
+    VOC dataset. As approximate inference was used, the shown objective values
+    are only lower bounds on the actual primal objective.
 \figlabel{msrc_pascal}}
@@
 
-\section{Discussion}
+\section{Summary}
 From the experiments above, it is clear there is currently \emph{no single best algorithm}.
-We can say, however, that the stochastic subgradient descent algorithm with
-pegasos step-size always converged slower, both in terms of learning time and
-in passes over the dataset, as the SZLJSP algorithm.
-Together with the fact that stochastic subgradient descent has no good stopping
-criterion, there seems to be no reason to use the stochastic gradient descent.
+However, there are some clear trends:
+\begin{itemize}
+    \item SZJLSP always converges faster than SSGD, with each iteration having
+        the same time complexity.  This indicates that SZJLSP should be
+        preferred over SSGD, apart from settings with strong memory
+        constraints.
+    \item In nearly all experiments, caching significantly improved performance
+        of the 1-slack cutting plane algorithm.
+    \item Initially SZJLSP and even SSGD converge faster than the 1-slack
+        cutting plane algorithm, while the 1-slack cutting plane algorithm was
+        faster in achieving high precision solutions and guaranteeing low
+        duality gap.
+    \item The n-slack cutting plane algorithm was often fastest to converge in
+        terms of passes over the training set, but often slow in terms of
+        runtime. The caching $1$-slack algorithm converged faster in terms of
+        cpu-time on almost all datasets.
+\end{itemize}
+There are however some caveats with our analysis, in particular with respect to
+the cutting plane solvers. The performance of the cutting plane algorithms
+depends not only on the performance of the QP solver used (in our case
+quadratic cone programming), but also on heuristics for pruning variables from
+the QP, caching constraints (for the $1$-slack algorithm) and deciding how
+often to solve the QP (for the $n$-slack algorithm). Better heuristics could
+have a beneficial effect on these algorithms. On the other hand, it is
+remarkable how competative the SZLJSP algorithm is, without the need for any
+such implementation tricks. This makes the SZJLSP algorithm very attractive for
+the practitioner who does not want to spend much time tuning the
+implementation.
+
+We do not consider SSGD using averaging here, but found it so produce similar
+results in tentative experiments. In particular, it was never faster than SZLJSP.
+
+With respect to approximate learning, we found that the inference algorithm has
+a stronger impact on the performance than the learning algorithm.
+While the SSGD and SZJLSP algorithms have somewhat lower training error on the snakes than
+the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.
+Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
+task near perfect.
+For the segmentation datasets, the approximate inference seem to be less of a
+problem, as most algorithms achieve the desired approximate duality gap (which
+is not the actual duality gap, see Chapter~\ref{ch:exact_learning} for a
+discussion).
+The large label space made using the linear programming relaxation intractable,
+though, and we can not draw a final conclusion here. In the next chapter,
+Chapter~\ref{ch:exact_learning}, we will demonstrate how we can learn exactly,
+even in this setting.

### 2013-09-25 main.tex
more results, discussion for experiments

@@
 \begin{document}
-
+%
 \begin{spacing}{1.6}
 \begin{titlepage}
-
+\vspace*{1mm}
 \begin{center}
 \Large\sffamily
-\vspace*{1cm}
-\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
+\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[4mm]
 Dissertation\\
@@
 Offenbach am Main\\
-Bonn, September 2013
+Bonn, September 2013\\[5mm]
+
+%\vfill
+\includegraphics[height=1.5cm]{ais_logo_164px}
 \end{center}

### 2013-09-25 structured_prediction/paper.tex
more results, discussion for experiments

@@
 different parameter vectors $\theta_{F, y_{N_F}}$ for each possible assignment
-of $y_{N_F} \in \mathcal{Y}_{N_F}$.\pagebreak\\
+of $y_{N_F} \in \mathcal{Y}_{N_F}$.
 %

### 2013-09-25 main.tex
add logos to title page

--- 2013-09-24 main.tex
@@
 \begin{document}
-
-\begin{spacing}{1.6}
+%
+\begin{spacing}{1.5}
 \begin{titlepage}
-
+\vspace*{1mm}
 \begin{center}
 \Large\sffamily
-\vspace*{1cm}
 \textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
@@
 Offenbach am Main\\
-Bonn, September 2013
+Bonn, September 2013\\[10mm]
+
 \end{center}
+\hspace{2.4cm}\includegraphics[height=2.0cm]{ais_uni_logo}
 \end{titlepage}

### 2013-09-25 evaluation/paper.tex
work on evaluation blub

@@
 For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
-We found that using the linear programming relaxation on the multi-label
-dataset is often tight, and there was no need to resort to approximate
-inference.
+We found that using the linear programming relaxation on the scene dataset is
+often tight, and there was no need to resort to approximate inference. Inference
+on the yeast dataset was more complex, and on the verge of being non-practical.
+We use this dataset as an example for very cosly inference.
 
@@
 
-Next, we consider the multi-label task, starting with the tree-shaped models.
-The results for the yeast and scene datasets are shown in \Figref{multi-label-results},
-respectively. The caching $1$-slack algorithm is much faster than the
-non-caching one on both datasets. While the two variants of the $n$-slack algorithm are equally
-fast with respect to passes over the training data, solving the QP every 100 steps is among the
+Next, we consider the multi-label task.  The results for the yeast and scene
+datasets are shown in \Figref{multi-label-results}, respectively. The caching
+$1$-slack algorithm is much faster than the non-caching one on both datasets.
+While the two variants of the $n$-slack algorithm are equally fast with respect
+to passes over the training data, solving the QP every 100 steps is among the
 fastest algorithms, while solving the QP every step is among the slowest.
 
-The caching $1$-slack algorithm is the fastest algorithm to achieve the desired primal suboptimality.
-On the scene dataset, the other cutting-plane base algorithms are second fastest, while they
-lag behind the stochastic algorithms on the larger yeast dataset. Again, initial convergence is
-faster for the stochastic algorithms.
+The caching $1$-slack algorithm is the fastest algorithm to achieve the desired
+primal suboptimality.  On the scene dataset, the other cutting-plane base
+algorithms are second fastest, while they lag behind the stochastic algorithms
+on the larger yeast dataset. Again, initial convergence is faster for the
+stochastic algorithms.
 
@@
 \subsection{Experiments using Approximate Inference}
-For the remaining tasks of multi-label classification with the full graph, the
-snakes dataset and image segmentation, exact inference is usually not possible.
+For the remaining tasks of the snakes dataset and image segmentation, exact
+inference is intractable for learning.
 
 We use two different approximate inference algorithms, fusion moves, a fast
-local search procedure, and AD$^3$, a dual-subgradient method to solve the LP
-relaxation of the inference problem.
+local search procedure, and the linear programming relaxation provided by AD$^3$,
+not using branch-and-bound.
+
 Evaluation of algorithms where exact inference is not possible is much harder,
@@
 inference algorithm is.
+
+\Figref{snakes_results} shows the results on the snakes dataset.
+The top row shows the objective when learning with fusion-moves, the next row the training set loss.
+First, we notice that even the approximate
 

### 2013-09-25 main.tex
align logo slightly differently, add stuff to future work.

@@
 \end{center}
-\hspace{2.4cm}\includegraphics[height=2.0cm]{ais_uni_logo}
+\hspace{3.4cm}\includegraphics[height=2.0cm]{ais_uni_logo}
 \end{titlepage}
@@
 interesting to pursue as an extension of the presented results:
+%
 \paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
@@
 path for improving the presented results.
+%
+\paragraph{Cached Inference for SZLJSP}
+We saw in Chapter~\ref{ch:comparison} that the $1$-slack cutting plane algorithm
+benefits immensely from caching inference results during training. Therefore, investigating
+the influence of caching for SZLJSP seems a promising topic for future research.
+%
+\paragraph{Theoretical Analysis of $n$-slack Algorithm}
+As we have seen in Chapter~\ref{ch:comparison}, the $n$-slack algorithm often
+converges very fast in terms of passes over the training data This is in stark
+contrast to the known theoretical convergence guarantee, which is the slowest
+of all the algorithms we considered with
+$O(\frac{1}{\epsilon^2})$\footnote{This is in terms of calls to the QP\@. We are
+not aware of any analysis in terms of inference calls or passes over the
+training set.}. It seem as if the approach of \citet{lacoste2012block} can
+yield a better convergence guarantee, but it is also worth investigating the
+direction pursued by \citet{shalev2012proximal}.
+%
 \paragraph{Inference Machines}
@@
 work in this direction, leaving much room for future investigation.
+%
 \paragraph{Non-Linear Models}
@@
 segmentation with DTFs.
-
+%
 \paragraph{Higher Order Potentials and Latent Variable Models}
@@
 interact with different inference and learning schemes.
-
+%
 \paragraph{Feature Design}

### 2013-09-25 evaluation/paper.tex
bring experiments section into some sort of shape... at least for exact inference.

@@
 In this section, we present results on multi-class classification, sequence
-labeling and multi-label prediction. In these
-tasks, exact inference is possible, and we can compute the exact objective of
-the various algorithms easily. We use the dynamic programming algorithm
-implemented in \textsc{OpenGM} for experiments for the tree and chain models.
+labeling and multi-label prediction. In these tasks, exact inference is
+possible, and we can compute the exact objective of the various algorithms
+easily. We use the dynamic programming algorithm implemented in \textsc{OpenGM}
+for experiments for the tree and chain models.
 For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
@@
 on the yeast dataset was more complex, and on the verge of being non-practical.
-We use this dataset as an example for very cosly inference.
+We use this dataset as an example for very costly inference.
+
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
+    \includegraphics[width=\linewidth]{evaluation/images/mnist}&%
+    \includegraphics[width=\linewidth]{evaluation/images/mnist_loss}
+    \end{tabu}
+\caption{%
+   Convergence of the primal suboptimality and training set loss on MNIST. 
+\figlabel{mnist_curves}}
+\end{figure}
 
 We start our experiments with the MNIST dataset.
-The convergence of the primal objective and training set loss in terms of
-passes over the training set is shown in \Figref{mnist_iterations}.
-Several trends can be observed: the $n$-slack cutting plane algorithms
-converge fastest, with the version that recomputes the QP at every step leading
-the race. The algorithms are closely followed by the caching $1$-slack cutting
-plane algorithm.  They are followed by the significantly slower SZLJSP
-algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
-subgradient descent.  These results are intuitive, as they reflect ``how much
-work'' each algorithm does for each loss-augmented prediction step. More work
-towards the objective leads to faster convergence.  This ``more work'' is
-quantified in \Figref{mnist_times}.  Clearly the $1$-slack algorithms to ``to
-much'' work, leading to very slow convergence.  Also, caching does not speed up
-learning learning on this dataset, as loss-augmented prediction is as fast as
-querying the cache.
+The convergence of the primal suboptimality and training set loss in terms of
+passes over the training set is shown in \Figref{mnist_curves}. While inference
+is trivial, this is by far the largest dataset, making the $n$-slack algorithms
+not feasible. We see that the stochastic algorithms are much faster than the
+cutting plane algorithm, in particular initially, even in terms of iterations,
+with the SZLJSP is clearly leading the way. Interestingly, the $1$-slack algorithm
+catches up near the desired suboptimality. Looking at the training set loss,
+it is clear that the high precision we demanded was not necessary, and we could have
+terminated the stochastic algorithms much earlier.
 
@@
 \caption{%
-   Convergence of the primal objective on OCR-small. 
+   Convergence of the primal suboptimality on OCR-small. 
 \figlabel{ocr_small_results}}
@@
 The results for OCR-small are shown in \Figref{ocr_small_results}.
-The trends seen in the figures are similar to what we saw for MNIST above.
-Again, even though inference is non-trivial, the non-caching $1$-slack
-algorithm is slower than the caching one. Again, the plots against passes over the training set
-do not reflect the runtime complexity of the $n$-slack approach.
+Considering the plot against passes over the dataset on the left, several
+trends can be observed: the $n$-slack cutting plane algorithms converge
+fastest, with the version that recomputes the QP at every step leading the
+race. The algorithms are closely followed by the caching $1$-slack cutting
+plane algorithm.  They are followed by the significantly slower SZLJSP
+algorithm, and finally the non-caching $1$-slack cutting plane algorithm and
+subgradient descent.  These results are intuitive, as they reflect ``how much
+work'' each algorithm does for each loss-augmented prediction step. More work
+towards the objective leads to faster convergence.  This ``more work'' is
+quantified on the right hand side, where the suboptimality is plotted against time.
+Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow
+convergence.  Also, caching does not speed up learning learning on this
+dataset.
+
+%The trends seen in the figures are similar to what we saw for MNIST above.
+%Again, even though inference is non-trivial, the non-caching $1$-slack
+%algorithm is slower than the caching one. Again, the plots against passes over the training set
+%do not reflect the runtime complexity of the $n$-slack approach.
 
@@
 \caption{%
-   Convergence of the primal objective on OCR-small. 
+    Primal and dual objectives for SZLJSP and $1$-slack cutting plane on OCR-small.
 \figlabel{ocr_small_dual}}
@@
 
-There is one interesting phenomenon to be seen here: The cutting-plane algorithms stop immediately
+We want to highlight an interesting phenomenon here: The cutting-plane algorithms stop immediately
 when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
@@
 faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
-This means that in for this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
+This means that in this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
 in guaranteeing the desired suboptimality.
-\Figref{ocr_small_dual} illustrates this point by ploting primal and dual
+\Figref{ocr_small_dual} illustrates this point by plotting primal and dual
 objectives for the $1$-slack cutting plane and the SZLJSP algorithms.
@@
 \caption{%
-   Convergence of the primal objective on OCR-large. 
+   Convergence of the primal suboptimality on OCR-large. 
 \figlabel{ocr_large_results}}
@@
 in terms of passes through the dataset, but $n$-slack cutting plane is impractical slow
-with respect to runtime. The SZLJSP algorithm converges fast intially, but the $1$-slack cutting plane
+with respect to runtime. The SZLJSP algorithm converges fast initially, but the $1$-slack cutting plane
 is faster at high precision and faster in certifying the desired duality gap.
@@
     \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}\\
-    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log}&%
-    \includegraphics[width=\linewidth]{evaluation/images/yeast_tree_log_time}
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_full}&%
+    \includegraphics[width=\linewidth]{evaluation/images/yeast_full_time}
     \end{tabu}
@@
 Next, we consider the multi-label task.  The results for the yeast and scene
-datasets are shown in \Figref{multi-label-results}, respectively. The caching
-$1$-slack algorithm is much faster than the non-caching one on both datasets.
+datasets are shown in \Figref{multi-label-results}. The caching $1$-slack
+algorithm is much faster than the non-caching one on both datasets and both
+graph structures.
 While the two variants of the $n$-slack algorithm are equally fast with respect
@@
 fastest algorithms, while solving the QP every step is among the slowest.
+We applied this method only to the smaller scene dataset, as applying it on the
+yeast dataset was not practicable.
 
 The caching $1$-slack algorithm is the fastest algorithm to achieve the desired
-primal suboptimality.  On the scene dataset, the other cutting-plane base
-algorithms are second fastest, while they lag behind the stochastic algorithms
-on the larger yeast dataset. Again, initial convergence is faster for the
-stochastic algorithms.
+primal suboptimality for tree-structured graphs, while it is out-performed by
+the $n$-slack cutting plane algorithm for the full graphs. This is intuitive,
+as the additional work the $n$-slack algorithm does becomes more valuable, the
+longer the inference takes.
 
@@
 \begin{itemize}
-    \item SZJLSP always converges faster than SSGD, with each iteration having
+    \item SZLJSP always converges faster than SSGD, with each iteration having
         the same time complexity.  This indicates that SZJLSP should be
@@
         duality gap.
-    \item The n-slack cutting plane algorithm was often fastest to converge in
+    \item The $n$-slack cutting plane algorithm was often fastest to converge in
         terms of passes over the training set, but often slow in terms of
-        runtime. The caching $1$-slack algorithm converged faster in terms of
-        cpu-time on almost all datasets.
+        runtime. For problems with very slow inference, the $n$-slack algorithm
+        might be a better choice than the others---if the interval between re-solving
+        the QP is chosen right.
 \end{itemize}
@@
 have a beneficial effect on these algorithms. On the other hand, it is
-remarkable how competative the SZLJSP algorithm is, without the need for any
-such implementation tricks. This makes the SZJLSP algorithm very attractive for
+remarkable how competitive the SZLJSP algorithm is, without the need for any
+such implementation tricks. This makes the SZLJSP algorithm very attractive for
 the practitioner who does not want to spend much time tuning the
@@
 
-We do not consider SSGD using averaging here, but found it so produce similar
-results in tentative experiments. In particular, it was never faster than SZLJSP.
+We do not consider SSGD with averaging~\cite{lacoste2012simpler} here, but
+found it so produce similar results in tentative experiments. In particular, it
+never outperformed SZLJSP.
 
@@
 a stronger impact on the performance than the learning algorithm.
-While the SSGD and SZJLSP algorithms have somewhat lower training error on the snakes than
+While the SSGD and SZLJSP algorithms have somewhat lower training error on the snakes than
 the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.

### 2013-09-25 intro.tex
add archive preprint to exact learning paper reference
align logo slightly differently, add stuff to future work.
some graphics, minor fixes.

@@
 labeling image pixels into classes is called object class segmentation or
-semantic segmentation.
-\pagebreak
+semantic segmentation.\pagebreak\\
 
@@
 different datasets in this thesis: the object class segmentation datasets
-Graz-02~\citep{marszatek2007accurate} and Pascal VOC 2010~\citep{pascal}, and the semantic segmentation datasets
-MSRC-21~\citep{shotton2006textonboost} and NYU V2~\citep{SilbermanECCV12}.
+Graz-02~\citep{marszatek2007accurate} and Pascal VOC 2010~\citep{pascal}, and
+the semantic segmentation datasets MSRC-21~\citep{shotton2006textonboost} and
+NYU V2~\citep{SilbermanECCV12}. Examples can be found in \Figref{pascal_msrc}.
 %
@@
 learn segmentation automatically from image level annotations.
+
+\begin{figure}[t]
+    \begin{center}
+    \begin{tabu} to 0.8\linewidth{@{}XX@{}}
+    \includegraphics[width=\linewidth]{msrc_scene}&%
+    \includegraphics[width=\linewidth]{msrc_scene_GT}\\
+    \includegraphics[width=\linewidth]{pascal}&%
+    \includegraphics[width=\linewidth]{pascal_gt}
+    \end{tabu}
+    \end{center}
+\caption{%
+    Examples from the MSRC-21 (top) and Pascal VOC (bottom) datasets with
+    ground-truth annotation. MSRC-21 contains mostly texture classes, such as
+    tree, building, street and sky, but also objects, like cars in this example.
+    Pascal VOC contains only object classes, such as person, cat, table and
+    bottle, and an additional background class (black).
+\figlabel{pascal_msrc}}
+\end{figure}
 
@@
 state-of-the-art methods on the NYU V2 datasets.
-\pagebreak
 
@@
         \emph{Learning a Loopy Model for Semantic Segmentation Exactly}\\
-        Andreas C. M\"uller and Sven Behnke.  Submitted to the International
+        Andreas C. M\"uller and Sven Behnke.\ arXiv preprint 1309.4061, Submitted to the International
         Conference on Computer Vision Theory and Applications.

### 2013-09-25 structured_prediction/paper.tex
some graphics, minor fixes.

@@
 \section{CRFs for Semantic Segmentation}\seclabel{semantic_segmentation}
+%
 Structured models based on factor graphs, as described in \Secref{factor_graphs} are often
@@
 evidence does not suffice.
+%
+\begin{figure}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_org}&%
+    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp}\\
+    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_graph}&%
+    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_gt}
+\end{tabu}
+\caption{%
+    Schematic of CRFs over superpixels. From left to right: input image, superpixel segmentation
+    using SLIC into about 100 superpixels, pairwise potential in the CRF, and desired labeling.
+\figlabel{superpixel_illustration}}
+\end{figure}
 
@@
 %TODO where does that leave us now?
-In Chapter~\ref{exact_learning} and Chapter~\ref{nyu}, we investigate
-superpixel-based conditional random fields learned with conditional random
-fields. Our emphasis is on understanding the learning, and how to cope with the
-complexity of inference in loopy models.
+In Chapter~\ref{ch:exact_learning} and Chapter~\ref{ch:nyu}, we investigate
+superpixel-based conditional random fields learned using structural support
+vector machines. The emphasis of Chapter~\ref{ch:exact_learning} is on
+understanding the learning, and how to cope with the complexity of inference in
+loopy models, while Chapter~\ref{ch:nyu} extends the approach to RGB-D images
+and shows how we can learn spatial relations using parametric pairwise
+potentials.

### 2013-09-25 evaluation/paper.tex
working on approximate inference result discussion

@@
 \subsection{Experiments using Approximate Inference}
-For the remaining tasks of the snakes dataset and image segmentation, exact
+For the remaining tasks, the snakes dataset and image segmentation, exact
 inference is intractable for learning.
-
+%
 We use two different approximate inference algorithms, fusion moves, a fast
 local search procedure, and the linear programming relaxation provided by AD$^3$,
-not using branch-and-bound.
+not using branch-and-bound, in contrast to the multi-label setup above.
 
@@
 relaxation).
-When using fusion moves, there is no obvious interpretation of the produced
+When using fusion moves, there is no obvious interpretation of the approximate
 primal objective, other than that it provides a lower bound of the actual
 objective. Nevertheless, we find it informative to analyze the behavior of this
-lower bound.
+lower bound. In contrast to the previous section, we show primal and dual objective
+values instead of the primal suboptimality when using approximate inference.
+The dual values do provide lower bounds
+to the exact objective, and are therefore somewhat more informative here, see
+Chapter~\ref{ch:exact_learning} for a discussion.
 
@@
 \Figref{snakes_results} shows the results on the snakes dataset.
-The top row shows the objective when learning with fusion-moves, the next row the training set loss.
-First, we notice that even the approximate
+The top row shows the objective when learning with fusion-moves, the next row
+the training set loss.  First, we observe that using fusion-moves even the
+approximate objective does not reach the desired optimality gap for the
+$1$-slack and SZLJSP algorithms. This is caused by the algorithms not finding
+any further constraints. The $n$-slack algorithm fairs a bit better and
+achieves a higher dual value. Looking at the loss, it is clear that the stochastic
+algorithms have an advantage, possibly caused by not breaking to early.
+Still all algorithms ultimately fail to solve the task.
+
+The situation is very different when using the linear programming relaxation.
+Looking at the four plots at the bottom of \Figref{snakes_results}, we see that all
+algorithms are able to solve the task perfectly. %FIXME more details!
+Surprisingly, SSGD was very successful on this dataset, even faster then the non-caching $n$-slack
+algorithm. Caching provides nearly an order of magnitude speed-up on the dataset.
+
+Using branch-and-bound to obtain exact results was not feasible on this dataset.
 
@@
     Results on the snakes dataset. The top four plots show results using fusion
-    move inference, the bottom four plots using AD$^3$. The training set error shows
-    that none of the algorithms is able to solve the task using fusion moves.
-    The objective values for fusion move inference are only approximate, giving
-    an upper bound on the true primal objective, while the objective for AD$^3$
-    is the true objective for the relaxed task.
+    move inference, the bottom four plots using AD$^3$. Dashed lines indicate the dual objective.
 \figlabel{snakes_results}}
 \end{figure}
+
+The last set of experiments is on the segmentation dataset MSRC-21 and Pascal
+VOC\@.  Because of the large number of labels and samples, it is only feasible
+to use the fusion move inference algorithms. The results are shown in
+figure~\cite{msrc_pascal_curves}, with the top four plots showing results on
+MSRC-21, and the bottom four plots showing results on Pascal VOC.
+
+In both datasets, the $1$-slack algorithm is somewhat slow in minimizing the
+training set error, and the $n$-slack and stochastic algorithms are much
+faster. All algorithms achieve the desired suboptimality, at least in the
+approximate objective. This suggests that inference is somewhat easier for this
+task.
 
@@
     show results for the MSRC-21 dataset, the bottom four plots for the Pascal
-    VOC dataset. As approximate inference was used, the shown objective values
-    are only lower bounds on the actual primal objective.
-\figlabel{msrc_pascal}}
+    VOC dataset.
+\figlabel{msrc_pascal_curves}}
 \end{figure}
@@
 discussion).
-The large label space made using the linear programming relaxation intractable,
-though, and we can not draw a final conclusion here. In the next chapter,
-Chapter~\ref{ch:exact_learning}, we will demonstrate how we can learn exactly,
-even in this setting.
+In the next chapter, Chapter~\ref{ch:exact_learning}, we will demonstrate how
+we can learn exactly, even in this setting.

### 2013-09-26 evaluation/paper.tex
new plots for msrc and pascal

@@
 figure~\cite{msrc_pascal_curves}, with the top four plots showing results on
-MSRC-21, and the bottom four plots showing results on Pascal VOC.
-
+MSRC-21, and the bottom four plots showing results on Pascal VOC\@.
+First, we notice that all algorithms achieve low (approximate) duality gap, and none
+of the algorithms terminates prematurely. Also, all algorithms achieve a similar dual value.
+All algorithms achieve similar training errors after convergence.
 In both datasets, the $1$-slack algorithm is somewhat slow in minimizing the
 training set error, and the $n$-slack and stochastic algorithms are much
-faster. All algorithms achieve the desired suboptimality, at least in the
-approximate objective. This suggests that inference is somewhat easier for this
-task.
+faster. Solving the QP only every 100 steps in the $n$-slack algorithm significantly slows
+down learning here. This is in contrast to the yeast datasets, where the converse was true,
+even though both datasets have a similar number of training examples.
+Given the outcome described above, we suspect that learning was not affected by approximate
+inference too much. We leave a more detailed analysis for Chapter~\ref{ch:exact_learning}.
 
@@
 We do not consider SSGD with averaging~\cite{lacoste2012simpler} here, but
-found it so produce similar results in tentative experiments. In particular, it
-never outperformed SZLJSP.
+found it so produce results similar to plain SSGD in tentative experiments. In
+particular, it never outperformed SZLJSP.
 

### 2013-09-26 evaluation/paper.tex
proof reading of experiments

--- 2013-09-25 evaluation/paper.tex
@@
 on one fold, and testing on the remaining nine folds, following \citet{taskar2003max},
-and learning on nine folds and testing on the remaining fold, following \citet{lacoste}.
+and learning on nine folds and testing on the remaining fold, following \citet{lacoste2012block}.
 We refer to the learning on one fold as OCR-small, and learning on nine folds as OCR-big.
@@
 Multi-label classification was first formulated as a structured prediction problem
-in by \citet{finley2008training}, where it was used to investigate the influence of
-approximate inference on the $n$-slack cutting plane algorithm. In this formulation,
+in by \citet{finley2008training}, who used to investigate the influence of
+approximate inference on the $n$-slack cutting plane algorithm. In their formulation
 each class is represented as a binary node in a factor graph---the states representing
@@
 \subsection{2D Grid CRF (Snakes)}
-The snakes dataset is a synthetic dataset, where samples are labeled 2D grids.
+The snakes dataset is a synthetic dataset where samples are labeled 2D grids.
 It was introduced in \citet{nowozin2011decision} to demonstrate the importance
@@
 the snake from ``head'' to ``tail'', that is assigning numbers from zero to
-nine to the cells that are occupied by the snake.  \Figref{snakes_illustration} illustrates the
-principle.
+nine to the cells that are occupied by the snake.  \Figref{snakes_illustration}
+illustrates the principle.
 Local evidence for the target label is weak except
-around head and tail, making this a challenging task. Pairwise potentials are necessary to solve the
-task. The dataset is noise-free in the sense that given the above description, a
-human could easily produce the desired labeling without making any mistake.
-The dataset is also interesting as the model of \citet{nowozin2011decision} produced
-notoriously hard to optimize energy functions.
+around head and tail, making this a challenging task, requiring strong pairwise
+potentials. The dataset is noise-free in the sense that given the above
+description, a human could easily produce the desired labeling without making
+any mistake. The dataset is also interesting as the model of
+\citet{nowozin2011decision} produced notoriously hard to optimize energy
+functions.
 
@@
 (``up'', ``down'', ``left'', ``right'', ``background'').  To encode the input
-more suitable for our linear methods, we convert this representation to a
+more suitably for our linear methods, we convert this representation to a
 one-hot encoding of the five states.
 
-We use a grid CRF model for this task. Unary for each node are given by the
+We use a grid CRF model for this task. Unary potentials for each node are given by the
 input of the 8-neighborhood of the node---using a 4-neighborhood would most
@@
 $45$ entries corresponding to the feature of the ``top'' node, the second $45$
-entries to the features ``bottom'' node, followed bye the ``left'' and
+entries to the features ``bottom'' node, followed by the ``left'' and
 ``right'' nodes. As each edge is either horizontal \emph{or} vertical, only two
@@
 semantic segmentation.  We use the Pascal VOC and MSRC datasets.
-The Pascal VOC has 964 training, each having
+The Pascal VOC has 964 training images, each having
 around 100 superpixels. There are 20 object categories, plus background. The % TODO check! refer to other part
@@
 each. There are 21 classes in the MSRC-21 dataset.
-Each superpixel is represented as an output variable, with the ground truth obtained
-by majority vote over the pixel belonging to the superpixel. We removed all
-superpixels which were mostly ``void'', leading to some samples haven much
-fewer than 100 variables.
-Pairwise potentials were introduced for each pair of neighboring superpixels.
+Each superpixel is represented as an output variable, with the ground truth
+obtained by majority vote over the pixel belonging to the superpixel. We
+removed all superpixels which were mostly ``void'', leading to some samples
+having much fewer than 100 variables.
+Pairwise potentials are introduced for each pair of neighboring superpixels.
 We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
-for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC21 dataset.
-We use the same three pairwise features for both dataset: a constant, a color
+for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC-21 dataset.
+We use the same three pairwise features for both dataset: a constant feature, a color
 feature and a feature encoding relative vertical position.
@@
 for the potentials on Pascal VOC and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
-2646$ parameters for the potentials on the MSRC21 dataset.
+2646$ parameters for the potentials on the MSRC-21 dataset.
 
@@
     snake      & 200     &84--168& loopy & 22275 & 11\\
-    MSRC21     & 276     & 7--113& loopy &  1764 & 21\\
+    MSRC-21     & 276     & 7--113& loopy &  1764 & 21\\
     Pascal VOC & 964     & 8--112& loopy &  2646 & 21\\
@@
 \section{Experiments}
-We compare the following algorithms on the above models:
+We compare the following algorithms on the above models (see \Secref{learning_algorithms} for details):
 \begin{itemize}
@@
             %TODO
-    \item The $n$-slack cutting plane algorithm, where the QP is solved after each training.
-    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 training samples.
-    \item The SZLJSP algorithm.
+    \item The $n$-slack cutting plane algorithm, where the QP is solved after each sample
+    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 samples.
+    \item The SZLJSP algorithm with weighted averaging.
 \end{itemize}
@@
 and how robust they are to approximate inference.  To quantify our goals, we
-track primal suboptimality and training set error during learning, as well as
-final error on the test set.
-We report suboptimality as a function of runtime, and as a function of passes
-over the training set.  Both are important to practitioners, as they give
+track primal suboptimality and training set error during learning.
+%, as well as final error on the test set. FIXME?
+We report primal suboptimality as a function of runtime, and as a function of passes
+over the training set.  Both are informative to practitioners, as they give
 important insight into the working of the algorithm.  The actual runtime is
-certainly the most important factor, but also highly influenced by
+arguably the most relevant factor, but also highly influenced by
 implementation details and properties of the dataset. We are particularly
-interested in cases where inference is highly non-trivial, making inference the
-dominating factor of runtime.
+interested in cases where inference is highly non-trivial, making it the
+dominating factor with respect to runtime.
 We define one pass over the dataset as calling prediction once for each sample.
-This way, we can get a clear picture how learning times with complexity of the
+This way, we can get a clear picture of how learning times scale with complexity of the
 inference task. For the subgradient and SZLJSP algorithms, learning time is linear
@@
 for experiments for the tree and chain models.
-For the full pairwise multi-label model, we use branch-and-bound together with the %FIXME AD3 solver description?
+For the full pairwise multi-label model, we use branch-and-bound together with
+the AD$^3$ algorithm.%FIXME AD3 solver description?
 We found that using the linear programming relaxation on the scene dataset is
@@
 cutting plane algorithm, in particular initially, even in terms of iterations,
-with the SZLJSP is clearly leading the way. Interestingly, the $1$-slack algorithm
-catches up near the desired suboptimality. Looking at the training set loss,
-it is clear that the high precision we demanded was not necessary, and we could have
-terminated the stochastic algorithms much earlier.
+with the SZLJSP clearly leading the way.
+%Interestingly, the $1$-slack algorithm
+%catches up near the desired suboptimality and is faster in achieving the desired duality gap--even though
+%the primal objective is always above SZLJSP\@. A zoom-in can be found in
+%\Figref{ocr_small_duality}, right hand side.
+
+We want to highlight an interesting phenomenon here: The cutting-plane algorithms stop immediately
+when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
+on learning much longer---even though the primal of the $1$-slack algorithms is
+always above the primal of the SZLJSP\@. This is caused by a looser bound
+given by the dual. Remember that the stopping criterion for both cutting-plane 
+and SZLJSP are given by the duality gap. It seems that while the primal objective converges much
+faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
+This means that in this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
+in guaranteeing the desired suboptimality. \Figref{ocr_small_dual} illustrates this point by plotting primal and dual
+objectives for the $1$-slack cutting plane and the SZLJSP algorithms.
+ Looking at the training set loss in \Figref{mnist_curves} suggests that the
+ high precision we demanded was not necessary, and we could have terminated the
+ stochastic algorithms much earlier.
 
@@
 towards the objective leads to faster convergence.  This ``more work'' is
-quantified on the right hand side, where the suboptimality is plotted against time.
-Clearly the $1$-slack algorithms to ``to much'' work, leading to very slow
+quantified on the right hand side of \Figref{ocr_small_results}, where the suboptimality is plotted against time.
+Clearly the $n$-slack algorithms do ``too much'' work, leading to very slow
 convergence.  Also, caching does not speed up learning learning on this
 dataset.
+%
+As on MNIST, the $1$-slack algorithm stops before the SZLJSP (in terms of
+iterations) because of a tighter dual.  A detailed view is shown in
+\Figref{ocr_small_dual} on the right.
 
@@
 \begin{figure}
-    \begin{center}
-    \includegraphics[width=.5\linewidth]{evaluation/images/letters_small_dual}
-\end{center}
+    \begin{tabu} to \linewidth{@{}XX@{}}
+    \multicolumn{2}{@{}c@{}}{\includegraphics[width=\linewidth]{evaluation/legend.pdf}}\\[-3mm]
+    \includegraphics[width=\linewidth]{evaluation/images/mnist_duality_gap}&%
+    \includegraphics[width=\linewidth]{evaluation/images/letters_small_duality_gap}
+    \end{tabu}
 \caption{%
-    Primal and dual objectives for SZLJSP and $1$-slack cutting plane on OCR-small.
+    Primal objective and dual objective (dashed lines) for SZLJSP and $1$-slack cutting plane on MNIST (left) and OCR-small (right).
 \figlabel{ocr_small_dual}}
 \end{figure}
-
-We want to highlight an interesting phenomenon here: The cutting-plane algorithms stop immediately
-when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
-on learning much longer---even longer than it took to get to the desired suboptimality. This is caused
-by a much looser bound given by the dual. Remember that the stopping criterion for both cutting-plane 
-and SZLJSP are given by the duality gap. It seems that while the primal objective converges much
-faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
-This means that in this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
-in guaranteeing the desired suboptimality.
-\Figref{ocr_small_dual} illustrates this point by plotting primal and dual
-objectives for the $1$-slack cutting plane and the SZLJSP algorithms.
 
@@
 is faster at high precision and faster in certifying the desired duality gap.
-
 \begin{figure}
@@
 \caption{%
-   Convergence of the primal objective on the multi-label datasets. From top to
+   Convergence of the primal suboptimality on the multi-label datasets. From top to
    bottom: the scene dataset using a tree model, the scene dataset using a full
@@
 to passes over the training data, solving the QP every 100 steps is among the
-fastest algorithms, while solving the QP every step is among the slowest.
+fastest algorithms, while solving the QP at every step is among the slowest.
 We applied this method only to the smaller scene dataset, as applying it on the
@@
 We use two different approximate inference algorithms, fusion moves, a fast
-local search procedure, and the linear programming relaxation provided by AD$^3$,
+local search procedure, and the linear programming relaxation provided by AD$^3$
 not using branch-and-bound, in contrast to the multi-label setup above.
@@
 \Figref{snakes_results} shows the results on the snakes dataset.
-The top row shows the objective when learning with fusion-moves, the next row
-the training set loss.  First, we observe that using fusion-moves even the
+The top row shows the objective when learning with fusion moves, the next row
+the training set loss.  First, we observe that using fusion moves even the
 approximate objective does not reach the desired optimality gap for the
 $1$-slack and SZLJSP algorithms. This is caused by the algorithms not finding
-any further constraints. The $n$-slack algorithm fairs a bit better and
-achieves a higher dual value. Looking at the loss, it is clear that the stochastic
-algorithms have an advantage, possibly caused by not breaking to early.
+any further constraints. The $n$-slack algorithm fares a bit better and
+achieves a higher dual value. Looking at the training set loss, it is clear that the stochastic
+algorithms have an advantage, possibly caused by not stopping too early.
 Still all algorithms ultimately fail to solve the task.
@@
 Looking at the four plots at the bottom of \Figref{snakes_results}, we see that all
-algorithms are able to solve the task perfectly. %FIXME more details!
+algorithms are able to solve the task perfectly. Evaluating on the test-set, all algorithms
+achieve an accuracy of around 99.5\%, confirming the models actually learned the task.
 Surprisingly, SSGD was very successful on this dataset, even faster then the non-caching $n$-slack
 algorithm. Caching provides nearly an order of magnitude speed-up on the dataset.
-
+%
 Using branch-and-bound to obtain exact results was not feasible on this dataset.
@@
 VOC\@.  Because of the large number of labels and samples, it is only feasible
-to use the fusion move inference algorithms. The results are shown in
-figure~\cite{msrc_pascal_curves}, with the top four plots showing results on
-MSRC-21, and the bottom four plots showing results on Pascal VOC.
-
+to use the fusion move inference algorithm. The results are shown in
+\Figref{msrc_pascal_curves}, with the top four plots showing results on
+MSRC-21, and the bottom four plots showing results on Pascal VOC\@.
+First, we notice that all algorithms achieve low (approximate) duality gap, and none
+of the algorithms terminates prematurely. Also, all algorithms achieve a
+similar dual objectives and similar training errors after convergence.
 In both datasets, the $1$-slack algorithm is somewhat slow in minimizing the
 training set error, and the $n$-slack and stochastic algorithms are much
-faster. All algorithms achieve the desired suboptimality, at least in the
-approximate objective. This suggests that inference is somewhat easier for this
-task.
+faster. Solving the QP only every 100 steps in the $n$-slack algorithm significantly slows
+down learning on both datasets. This is in contrast to the yeast datasets, where the converse was true,
+even though yeast has a similar number of training examples to Pascal VOC.
+
+Given the outcome described above, we suspect that learning was not affected by approximate
+inference too much. We leave a more detailed analysis for Chapter~\ref{ch:exact_learning}.\pagebreak\\
 
@@
     \item SZLJSP always converges faster than SSGD, with each iteration having
-        the same time complexity.  This indicates that SZJLSP should be
+        the same time complexity.  This indicates that SZLJSP should be
         preferred over SSGD, apart from settings with strong memory
@@
         of the 1-slack cutting plane algorithm.
-    \item Initially SZJLSP and even SSGD converge faster than the 1-slack
+    \item Initially SZLJSP and even SSGD converge faster than the 1-slack
         cutting plane algorithm, while the 1-slack cutting plane algorithm was
@@
     \item The $n$-slack cutting plane algorithm was often fastest to converge in
-        terms of passes over the training set, but often slow in terms of
+        terms of passes over the training set, but slow in terms of
         runtime. For problems with very slow inference, the $n$-slack algorithm
@@
 
-We do not consider SSGD with averaging~\cite{lacoste2012simpler} here, but
-found it so produce similar results in tentative experiments. In particular, it
-never outperformed SZLJSP.
+We do not consider SSGD with averaging~\citep{lacoste2012simpler} here, but
+found it so produce results similar to plain SSGD in tentative experiments. In
+particular, it never outperformed SZLJSP.
 
@@
 a stronger impact on the performance than the learning algorithm.
-While the SSGD and SZLJSP algorithms have somewhat lower training error on the snakes than
+While the SSGD and SZLJSP algorithms have somewhat lower training error on the snakes dataset than
 the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.
 Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
-task near perfect.
-For the segmentation datasets, the approximate inference seem to be less of a
-problem, as most algorithms achieve the desired approximate duality gap (which
-is not the actual duality gap, see Chapter~\ref{ch:exact_learning} for a
-discussion).
-In the next chapter, Chapter~\ref{ch:exact_learning}, we will demonstrate how
+task near perfectly.
+For the segmentation datasets, it is harder to judge the outcome, though there seems to be little difference
+between the learning algorithms in terms of the final result.
+In the next chapter, we will demonstrate how
 we can learn exactly, even in this setting.

### 2013-09-26 structured_prediction/paper.tex
trying to work on crf part.

@@
 \subsection{The SZLJSP Algorithm}\seclabel{dual_coordinate_descent}
-\begin{algorithm*}[t]
+\begin{algorithm}[tp]
     \caption{Shalev-Shwartz Zhang Lacoste-Julien Jaggi Schmidt Pletscher\label{alg_szljsp}}
@@
     \end{doublespacing}
-\end{algorithm*}
+\end{algorithm}
 
@@
 
-\section{CRFs for Semantic Segmentation}\seclabel{semantic_segmentation}
+\section{Conditional Random Fields for Semantic Segmentation}\seclabel{semantic_segmentation}
 %
+\subsection{Fundamentals of Conditional Random Fields}
 Structured models based on factor graphs, as described in \Secref{factor_graphs} are often
@@
 probabilistic interpretation.
-Conditional Random Fields have been established as an important tool in many
-areas of computer vision. Applications include dense stereo, optical flow, 
-inpainting, de-noising, image editing, low-level segmentation and semantic segmentation.
+CRFs have been established as an important tool in many areas of computer
+vision. Applications include dense stereo, optical flow, inpainting,
+denoising, image editing, low-level segmentation and semantic segmentation.
 
-Most conditional random fields only apply unary and pairwise potential functions. In this case, FIXME equation 4.4
-becomes
+Most conditional random fields only apply unary and pairwise potential
+functions. In this case, \Eqref{log_factor_graph} becomes
 \begin{equation}\eqlabel{energy_crf}
@@
 \end{equation}
-Here $V$ enumerates variables, and $E\subset V \times V$ is a set of edges,
-representing pairwise factors.
+Here $V$ enumerates variables and $E\subset V \times V$ is a set of edges which
+represent pairwise factors.
 
@@
 in an image is represented as a variable, and models in which pixels are first
-grouped together into superpixels and each superpixel is represented as a variable.
-In pixel-based CRFs edges are usually introduced between adjacent pixels, that is using
-either 4-neighborhoods or 8-neighborhoods. Superpixel-based approaches introduce
-edges between adjacent superpixels, that is those that share a boundary.
-%TODO images of superpixels!
-By nature, both approaches create graphs with a large number of loops. This
-complicates prediction and learning, and is the central topic of
-Chapter~\ref{ch:exact_learning}.
+grouped together into superpixels and each superpixel is represented as a
+variable.  In pixel-based CRFs edges are usually introduced between adjacent
+pixels, that is using either 4-neighborhoods or 8-neighborhoods.
+Superpixel-based approaches introduce edges between adjacent superpixels, that
+is those that share a boundary. \Figref{superpixel_illustration} shows a
+sample image together with extracted superpixels and the neighborhood graph.
+
+By nature, models over pixels and models over superpixels create graphs with a
+large number of loops. This complicates prediction and learning, and is the
+central topic of Chapter~\ref{ch:exact_learning}.
 
@@
 While it is hard to even define what makes a sensible segmentation of an image,
-it is much easier to define, and also to compute an \emph{over-segmentation}.
+it is much easier to define an \emph{over-segmentation}.
 The only requirement for an over-segmentation is that each segment only
@@
 
-We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
-our experiments.  It has been shown to provide competative results with a
-minimum of computational complexity.  Algorithmically, SLIC simply computes a
-$k$-means clustering over pixels. Each pixels is represented as 5D point, using
-three color channels and the $x$ and $y$ coordinates in the image.  In a
-post-processing step, small segments are removed. To make clustering of so many
-points feasible, the search for the nearest cluster in $k$-means is restricted
-to a local neighborhood in the image.
-
-% about algorithms, slic?
 % hierarchical\dots has superpixels and pixels?
@@
 % submodular/non-submodular? higher order?
-
+\subsection{Previous Work}
 Conditional random fields were first used in the context of semantic
@@
 introduces higher order potentials to enforce label consistency within larger
-regions. Later, \citet{ladicky2009associative} proposed a hierarchical model,
-including the higher order potentials of \citet{kohli2009robust}, and
-additional lateral and hierarchical connections. A similar approach, using
-superpixels as the finest resolution, and additionally modelling object class
-coocurrences was suggested by \citet{gonfaus2010harmony}.  However, all these
-models did not learn the potentials of the CRF model in a principled manner.
-The above approaches all learned unary potentials using a non-structured
-approach, for example SVMs or boosting, and then set pairwise and higher order
-potentials in the model by hand.
+regions. Later, \citet{ladicky2009associative} proposed a hierarchical model
+over pixels and superpixels, including the higher order potentials of
+\citet{kohli2009robust} and additional lateral and hierarchical connections. A
+similar approach, using superpixels as the finest resolution, and additionally
+modelling object class coocurrences was suggested by
+\citet{gonfaus2010harmony}.  However, all these models did not learn the
+potentials of the CRF model in a principled manner.  The above approaches all
+learned unary potentials using a non-structured approach, for example SVMs or
+boosting, and then set pairwise and higher order potentials in the model by
+hand.
 
@@
 
-%TODO where does that leave us now?
-In Chapter~\ref{ch:exact_learning} and Chapter~\ref{ch:nyu}, we investigate
-superpixel-based conditional random fields learned using structural support
-vector machines. The emphasis of Chapter~\ref{ch:exact_learning} is on
-understanding the learning, and how to cope with the complexity of inference in
-loopy models, while Chapter~\ref{ch:nyu} extends the approach to RGB-D images
-and shows how we can learn spatial relations using parametric pairwise
-potentials.
+\subsection{Data, Features and Superpixels}
+When evaluating learning algorithms in Chapter~\ref{ch:evaluation} and
+Chapter~\ref{ch:exact_learning} we use the well-established Pascal VOC 2010 and
+MSRC-21 datasets, shown in the introduction in \Figref{pascal_msrc}.
+As we focus on the learning algorithms in these chapters, we use features for
+unary potentials from the literature.
+
+For both datasets we use the TextonBoost class probablities provided by
+\citet{krahenbuhl2012efficient}. For the Pascal VOC dataset, the potentials
+provided by \citet{krahenbuhl2012efficient} also include the responses of
+object detectors~\citet{}, to better capture the complex object classes.
+We average the potentials inside superpixels and use the resulting feature
+as input to our unary potentials.
+For the MSRC-21 dataset, we compute TextonBoost on two different scales, as
+suggested by \citet{mottaghianalyzing}. We additionally extract SIFT and color descriptors and create
+bag-of-word descriptors for each superpixel. Following the approach of
+\citet{lucchi2011spatial} we augment these with a global bag-of-word descriptor
+for each image. We train a linear SVM using an approximation to the additive
+$\chi^2$ kernel~\citet{vedaldi2010efficient} and use the response as an
+additional input to our CRF\@. This \emph{piecewise training} simplifies
+learning and was found to have little effect on
+accuracy~\citet{nowozin2010parameter}. In total, there are $63$ features
+for the MSRC-21 dataset, $21$ for each scale of TextonBoost and an additional
+$21$ for the bag-of-word model using an SVM.
+
+We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
+our experiments.  It has been shown to provide competative results with a
+minimum of computational complexity.  Algorithmically, SLIC simply computes a
+$k$-means clustering over pixels. Each pixels is represented as 5D point, using
+three color channels and the $x$ and $y$ coordinates in the image.  In a
+post-processing step, small segments are removed. To make clustering of so many
+points feasible, the search for the nearest cluster in $k$-means is restricted
+to a local neighborhood in the image. The superpixels shown in
+\Figref{superpixels_illustration} are computed using SLIC.

### 2013-09-26 evaluation/paper.tex
trying to work on crf part.
corrections hannes

@@
 \section{Datasets and Models}
-We consider several qualitatively different datasets, that have been widely used in the literature.
+We consider several qualitatively different datasets that have been widely used in the literature.
 The problems we consider are multi-class classification, sequence labeling, multi-label prediction,
@@
 We choose the classical MNIST dataset of handwritten digits, consisting of
-60000 training images and 10000 test images of the digits zero to nine.
-\Figref{mnist_illustration} shows some examples. Each image is a $28 \times 28$
+$60000$ training images and $10000$ test images of the digits zero to nine.
+\Figref{mnist_illustration} shows some examples. Each image is a $28{\times}28$
 grey-level image, resulting in a $784$-dimensional feature vector. We normalize
-the feature to lie between $0$ and $1$.
+the features between $0$ and $1$.
 %
@@
 Each input consists of the segmented handwritten letters of a word in lower
-case. The task is to classify all letters in a word, that is assign each
-segmented input letter one of the classes ``a'' to ``z''. As the first letter
+case. The task is to classify all letters in a word, that is, assign each
+segmented input letter to one of the classes ``a'' to ``z''. As the first letter
 of each word was capitalized, these were removed by \citet{taskar2003max},
@@
 We use a simple chain model with a single constant pairwise feature.
-This means the unary potential has $16 \cdot 8 \cdot 26 = 768$ parameters, one for each input
+This means that the unary potential has $16 \cdot 8 \cdot 26 = 768$ parameters, one for each input
 feature and output class. The pairwise potential consists of a matrix of
@@
 each class is represented as a binary node in a factor graph---the states representing
-presence of absence of the class. A different factor of pairwise potentials is introduced
+presence or absence of the class. A different factor of pairwise potentials is introduced
 between each pair of classes.
@@
 unvisited cells are marked as background.  The goal is to predict a labeling of
-the snake from ``head'' to ``tail'', that is assigning numbers from zero to
+the snake from ``head'' to ``tail'', that is, assigning numbers from zero to
 nine to the cells that are occupied by the snake.  \Figref{snakes_illustration}
@@
 any mistake. The dataset is also interesting as the model of
-\citet{nowozin2011decision} produced notoriously hard to optimize energy
+\citet{nowozin2011decision} produced notoriously hard-to-optimize energy
 functions.
@@
 task-knowledge into our model.
-Using the one-hot encoding of the input, this leads to $9 \dot 5 = 45$ unary features.
+Using the one-hot encoding of the input, this leads to $9 \cdot 5 = 45$ unary features.
 With 11 output classes, the unary potential has $11 \cdot 9 \cdot 5=495$ parameters.
@@
 semantic segmentation.  We use the Pascal VOC and MSRC datasets.
-The Pascal VOC has 964 training images, each having
-around 100 superpixels. There are 20 object categories, plus background. The % TODO check! refer to other part
+The Pascal VOC has 964 training images, each divided into
+around 100 superpixels. There are 20 object categories, and an additional background class. The % TODO check! refer to other part
 MSRC-21 dataset has 276 training images, also segmented in around 100 superpixels
@@
 Each superpixel is represented as an output variable, with the ground truth
-obtained by majority vote over the pixel belonging to the superpixel. We
-removed all superpixels which were mostly ``void'', leading to some samples
+obtained by majority vote over the pixels belonging to the superpixel. We
+removed all superpixels in which the majority of pixels is labeled ``void'', leading to some samples
 having much fewer than 100 variables.
@@
 for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC-21 dataset.
-We use the same three pairwise features for both dataset: a constant feature, a color
+We use the same three pairwise features for both datasets: a constant feature, a color
 feature and a feature encoding relative vertical position.
@@
 \begin{itemize}
-    \item Stochastic Subgradient Descent (SSGD) using the Pegasos schedule for step-sizes.
+    \item Stochastic Subgradient Descent (SSGD) using the Pegasos schedule for step-sizes
     %\item Stochastic Subgradient Descent using the Pegasos schedule: AVERAGING
-    \item The $1$-slack cutting plane algorithm without constraint caching.
-    \item The $1$-slack cutting plane algorithm with constraint caching. For
-            each sample, the last $50$ inference results are stored.
-            Details of the caching strategy can be found in Chapter~\ref{ch:exact_learning}.
-            %TODO
+    \item The $1$-slack cutting plane algorithm without inference caching
+    \item The $1$-slack cutting plane algorithm with caching the last 50
+        inference results for each sample (see Chapter~\ref{ch:exact_learning}
+        for details)
     \item The $n$-slack cutting plane algorithm, where the QP is solved after each sample
-    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 samples.
-    \item The SZLJSP algorithm with weighted averaging.
+    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 samples
+    \item The SZLJSP algorithm with weighted averaging
 \end{itemize}
 
-All algorithm take a $C$ parameter, which we adjusted on a fully trained model
+All algorithms take a $C$ parameter, which we adjusted on a fully trained model
 on a validation set (experiments not reported here) and held constant for all
@@
 removed, and parameters of the underlying QP solver. On the other hand, the
-SZLJSP, for example, has no hyper-parameters except for the stopping criterion.
+SZLJSP algorithm, has no hyper-parameters except for the stopping criterion.
 
@@
 For the full pairwise multi-label model, we use branch-and-bound together with
-the AD$^3$ algorithm.%FIXME AD3 solver description?
+the AD$^3$ algorithm. %FIXME AD3 solver description?
 We found that using the linear programming relaxation on the scene dataset is
@@
 The top row shows the objective when learning with fusion moves, the next row
-the training set loss.  First, we observe that using fusion moves even the
+the training set loss.  First, we observe that using fusion moves, even the
 approximate objective does not reach the desired optimality gap for the
@@
 algorithms have an advantage, possibly caused by not stopping too early.
-Still all algorithms ultimately fail to solve the task.
+Still, all algorithms ultimately fail to solve the task.
 
@@
 MSRC-21, and the bottom four plots showing results on Pascal VOC\@.
-First, we notice that all algorithms achieve low (approximate) duality gap, and none
+First, we notice that all algorithms achieve low (approximate) duality gaps, and none
 of the algorithms terminates prematurely. Also, all algorithms achieve a
@@
         might be a better choice than the others---if the interval between re-solving
-        the QP is chosen right.
+        the QP is chosen appropriately.
 \end{itemize}
-There are however some caveats with our analysis, in particular with respect to
+There are however some caveats to our analysis, in particular with respect to
 the cutting plane solvers. The performance of the cutting plane algorithms
@@
 
-We do not consider SSGD with averaging~\citep{lacoste2012simpler} here, but
-found it so produce results similar to plain SSGD in tentative experiments. In
+We do not present results for SSGD with averaging~\citep{lacoste2012simpler} here, but
+found it to produce results similar to plain SSGD in tentative experiments. In
 particular, it never outperformed SZLJSP.
@@
 Using the linear programming relaxation, on the other hand, all algorithms are able to solve the
-task near perfectly.
-For the segmentation datasets, it is harder to judge the outcome, though there seems to be little difference
-between the learning algorithms in terms of the final result.
-In the next chapter, we will demonstrate how
-we can learn exactly, even in this setting.
+task nearly perfectly.
+For the segmentation datasets, it is harder to judge the outcome, though there
+seems to be little difference between the learning algorithms in terms of the
+final result.  In the next chapter, we will demonstrate how we can learn
+exactly, even with the complex models used for semantic segmentation.

### 2013-09-26 itm/paper.tex
more corrections by vladn :)
corrections hannes

@@
 compute, because clustering $n$ points requires computing the eigenvalues and
--vectors of an $n\times n$ matrix.
+-vectors of an $n{\times}n$ matrix.
 
@@
 \Figref{qualitative} shows qualitative results on three synthetic datasets.
-For well separated, convex clusters, all four algorithms produce equivalent
-clusterings (see top row).  If the structure of the data is more complex, the
+For well separated, convex clusters, the four algorithms produce very similar
+clusters (see top row).  If the structure of the data is more complex, the
 advantage of the proposed method is apparent.  Note that there was no need to

### 2013-09-26 nyu/paper.tex
some graphics, minor fixes.
corrections hannes

@@
 We thereby extend the success of learned CRF models for semantic segmentation
-in RGB images as considered in Chapter~\ref{exact_learning} to the domain of 3D scenes.
+in RGB images as considered in Chapter~\ref{ch:exact_learning} to the domain of 3D scenes.
 Our emphasis lies on exploiting the additional depth and 3D information in all
@@
 Our superpixel generation is based on the SLIC
-algorithm~\citep{achanta2012slic}, as in Chapter~\ref{exact_learning}. We
+algorithm~\citep{achanta2012slic} described in \Secref{semantic_segmentation}. We
 extend the standard SLIC algorithm, which works on the Lab space, to also
@@
 \end{figure}
-\pagebreak\\
 %

### 2013-09-26 structured_prediction/paper.tex
corrections hannes

@@
 \subsection{Data, Features and Superpixels}
-When evaluating learning algorithms in Chapter~\ref{ch:evaluation} and
+When evaluating learning algorithms in Chapter~\ref{ch:comparison} and
 Chapter~\ref{ch:exact_learning} we use the well-established Pascal VOC 2010 and
@@
 points feasible, the search for the nearest cluster in $k$-means is restricted
-to a local neighborhood in the image. The superpixels shown in
-\Figref{superpixels_illustration} are computed using SLIC.
+to a local neighborhood in the image. An example of superpixels computed with the SLIC algorithm can be found in
+\Figref{superpixel_illustration}.
+
+The features and superpixel algorithm used for the RGB-D dataset NYU used in Chapter~\ref{ch:nyu} will be discussed there.

### 2013-09-26 evaluation/paper.tex
smallcaps for datasets, some more corrections by vlad

@@
 and learning on nine folds and testing on the remaining fold, following \citet{lacoste2012block}.
-We refer to the learning on one fold as OCR-small, and learning on nine folds as OCR-big.
+We refer to the learning on one fold as \textsc{OCR-small}, and learning on nine folds as OCR-big.
 We use a simple chain model with a single constant pairwise feature.
@@
 Multi-label classification was first formulated as a structured prediction problem
-in by \citet{finley2008training}, who used to investigate the influence of
+by \citet{finley2008training}, who used to investigate the influence of
 approximate inference on the $n$-slack cutting plane algorithm. In their formulation
@@
 inference via message passing.
-We use two of the datasets used in \citet{finley2008training}, the scene
-and yeast datasets. We choose these two as these are real-world datasets for
+We use two of the datasets used in \citet{finley2008training}, the \textsc{scene}
+and \textsc{yeast} datasets. We choose these two as these are real-world datasets for
 which the pairwise approach outlined above actually improves upon the
 baseline~\citep{finley2008training}.
-The scene dataset has six labels, 294 input features, 1211 training samples and
+The \textsc{scene} dataset has six labels, 294 input features, 1211 training samples and
 1196 test samples. This leads to $204 \cdot 6 = 1224$ parameters for the unary potentials,
@@
 simplifies writing out the model.
-The yeast dataset has 14 labels, 103 features, 1500 training samples and 971
+The \textsc{yeast} dataset has 14 labels, 103 features, 1500 training samples and 971
 test samples. The resulting numbers of parameters are $14 \cdot 103=1442$ parameters
@@
 \caption{%
-    Visualization of the snakes dataset. The top row shows input patterns, the bottom row the corresponding
+    Visualization of the \textsc{snakes} dataset. The top row shows input patterns, the bottom row the corresponding
     labels. The colors in the image showing the labels correspond to dark blue for background, and encoding
@@
 \subsection{2D Grid CRF (Snakes)}
-The snakes dataset is a synthetic dataset where samples are labeled 2D grids.
-It was introduced in \citet{nowozin2011decision} to demonstrate the importance
+The \textsc{snakes} dataset is a synthetic dataset where samples are labeled 2D grids.
+It was introduced by \citet{nowozin2011decision} to demonstrate the importance
 of learning conditional pairwise potentials. The dataset consists of ``snakes''
@@
 description, a human could easily produce the desired labeling without making
-any mistake. The dataset is also interesting as the model of
+any mistake. The dataset is also interesting as the model proposed by
 \citet{nowozin2011decision} produced notoriously hard-to-optimize energy
@@
 
-In \citet{nowozin2011decision}, the input was encoded into five RGB colors
+Originally, the input is encoded into five RGB colors
 (``up'', ``down'', ``left'', ``right'', ``background'').  To encode the input
@@
 Our main attention is devoted to the use of conditional random fields for
-semantic segmentation.  We use the Pascal VOC and MSRC datasets.
-The Pascal VOC has 964 training images, each divided into
+semantic segmentation.  We use the \textsc{Pascal VOC} and MSRC datasets.
+The \textsc{Pascal VOC} has 964 training images, each divided into
 around 100 superpixels. There are 20 object categories, and an additional background class. The % TODO check! refer to other part
@@
 We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
-for the Pascal VOC dataset 21 unary potentials and 63 for the MSRC-21 dataset.
+21 unary features for the \textsc{Pascal VOC} dataset and 63 unary features for the MSRC-21 dataset.
 We use the same three pairwise features for both datasets: a constant feature, a color
@@
 Overall, this results in $21 \cdot  21 + 3 \cdot 21 \cdot 21 = 1764$ parameters
-for the potentials on Pascal VOC and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
+for the potentials on \textsc{Pascal VOC} and $63 \cdot  21 + 3 \cdot 21 \cdot 21 =
 2646$ parameters for the potentials on the MSRC-21 dataset.
@@
     MNIST               & 60000   & 1     & none  & 7840  & 10\\
-    OCR-small  & 704     & 3--14 & chain & 1444  & 26\\
-    OCR-large  & 6173    & 3--14 & chain & 1444  & 26\\
-    scene-tree & 1211    & 6     & tree  & 1244  & 2\\
-    scene-full & 1211    & 6     & loopy & 1284  & 2\\
-    yeast-tree & 1500    & 14    & tree  & 1494  & 2\\
-    yeast-full & 1500    & 14    & loopy & 1806  & 2\\
-    snake      & 200     &84--168& loopy & 22275 & 11\\
+    \textsc{OCR-small}  & 704     & 3--14 & chain & 1444  & 26\\
+    \textsc{OCR-large}  & 6173    & 3--14 & chain & 1444  & 26\\
+    \textsc{scene-tree} & 1211    & 6     & tree  & 1244  & 2\\
+    \textsc{scene-full} & 1211    & 6     & loopy & 1284  & 2\\
+    \textsc{yeast-tree} & 1500    & 14    & tree  & 1494  & 2\\
+    \textsc{yeast-full} & 1500    & 14    & loopy & 1806  & 2\\
+    \textsc{snakes}     & 200     &84--168& loopy & 22275 & 11\\
     MSRC-21             & 276     & 7--113& loopy &  1764 & 21\\
-    Pascal VOC & 964     & 8--112& loopy &  2646 & 21\\
+    \textsc{Pascal VOC} & 964     & 8--112& loopy &  2646 & 21\\
     \bottomrule
@@
 the AD$^3$ algorithm. %FIXME AD3 solver description?
-We found that using the linear programming relaxation on the scene dataset is
+We found that using the linear programming relaxation on the \textsc{scene} dataset is
 often tight, and there was no need to resort to approximate inference. Inference
-on the yeast dataset was more complex, and on the verge of being non-practical.
+on the \textsc{yeast} dataset was more complex, and on the verge of being non-practical.
 We use this dataset as an example for very costly inference.
@@
 We now consider datasets with non-trivial inference.
-The results for OCR-small are shown in \Figref{ocr_small_results}.
+The results for \textsc{OCR-small} are shown in \Figref{ocr_small_results}.
 Considering the plot against passes over the dataset on the left, several
@@
 \caption{%
-    Primal objective and dual objective (dashed lines) for SZLJSP and $1$-slack cutting plane on MNIST (left) and OCR-small (right).
+    Primal objective and dual objective (dashed lines) for SZLJSP and $1$-slack cutting plane on MNIST (left) and \textsc{OCR-small} (right).
 \figlabel{ocr_small_dual}}
@@
    Convergence of the primal suboptimality on the multi-label datasets. From top to
-   bottom: the scene dataset using a tree model, the scene dataset using a full
-   model, the yeast dataset using a tree model, and the yeast dataset using a
+   bottom: the \textsc{scene} dataset using a tree model, the \textsc{scene} dataset using a full
+   model, the \textsc{yeast} dataset using a tree model, and the \textsc{yeast} dataset using a
    full model.
@@
 
-Next, we consider the multi-label task.  The results for the yeast and scene
+Next, we consider the multi-label task.  The results for the \textsc{yeast} and \textsc{scene}
 datasets are shown in \Figref{multi-label-results}. The caching $1$-slack
@@
 fastest algorithms, while solving the QP at every step is among the slowest.
-We applied this method only to the smaller scene dataset, as applying it on the
-yeast dataset was not practicable.
+We applied this method only to the smaller \textsc{scene} dataset, as applying it on the
+\textsc{yeast} dataset was not practicable.
 
@@
 \subsection{Experiments using Approximate Inference}
-For the remaining tasks, the snakes dataset and image segmentation, exact
+For the remaining tasks, the \textsc{snakes} dataset and image segmentation, exact
 inference is intractable for learning.
@@
 
-\Figref{snakes_results} shows the results on the snakes dataset.
+\Figref{snakes_results} shows the results on the \textsc{snakes} dataset.
 The top row shows the objective when learning with fusion moves, the next row
@@
 \caption{%
-    Results on the snakes dataset. The top four plots show results using fusion
+    Results on the \textsc{snakes} dataset. The top four plots show results using fusion
     move inference, the bottom four plots using AD$^3$. Dashed lines indicate the dual objective.
@@
 
-The last set of experiments is on the segmentation dataset MSRC-21 and Pascal
-VOC\@.  Because of the large number of labels and samples, it is only feasible
+The last set of experiments is on the segmentation dataset MSRC-21 and \textsc{Pascal
+VOC}.  Because of the large number of labels and samples, it is only feasible
 to use the fusion move inference algorithm. The results are shown in
 \Figref{msrc_pascal_curves}, with the top four plots showing results on
-MSRC-21, and the bottom four plots showing results on Pascal VOC\@.
+MSRC-21, and the bottom four plots showing results on \textsc{Pascal VOC}.
 First, we notice that all algorithms achieve low (approximate) duality gaps, and none
@@
 faster. Solving the QP only every 100 steps in the $n$-slack algorithm significantly slows
-down learning on both datasets. This is in contrast to the yeast datasets, where the converse was true,
-even though yeast has a similar number of training examples to Pascal VOC.
+down learning on both datasets. This is in contrast to the \textsc{yeast} datasets, where the converse was true,
+even though \textsc{yeast} has a similar number of training examples to \textsc{Pascal VOC}.
 
@@
     Objective and traning set loss on the segmentation task. The top four plots
-    show results for the MSRC-21 dataset, the bottom four plots for the Pascal
-    VOC dataset.
+    show results for the MSRC-21 dataset, the bottom four plots for the \textsc{Pascal
+    VOC} dataset.
 \figlabel{msrc_pascal_curves}}
@@
 a stronger impact on the performance than the learning algorithm.
-While the SSGD and SZLJSP algorithms have somewhat lower training error on the snakes dataset than
+While the SSGD and SZLJSP algorithms have somewhat lower training error on the \textsc{snakes} dataset than
 the cutting-plane algorithm when using fusion-move inference, they, too, fail at solving the task.

### 2013-09-26 exact_learning/paper.tex
smallcaps for datasets, some more corrections by vlad

@@
 field model using a structured support vector machine exactly for this task.
-We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. Our
+We evaluate our approach on the popular MSRC-21 and \textsc{Pascal VOC} datasets. Our
 approach improves upon the state-of-the-art on the MSRC-21 dataset, and is
-competative with comparable approaches on the Pascal VOC dataset.% and nyu?
+competative with comparable approaches on the \textsc{Pascal VOC} dataset.% and nyu?
 
@@
 Our main application is of course semantic segmentation and object class
-segmentation. We evaluate the proposed learning approach on Pascal VOC 2013 and
+segmentation. We evaluate the proposed learning approach on \textsc{Pascal} VOC 2013 and
 MSRC-21
@@
     \end{center}
-    \caption{Accuracies on the Pascal VOC Dataset. We compare our approach
+    \caption{Accuracies on the \textsc{Pascal} VOC Dataset. We compare our approach
     against approaches using the same unary potentials.\tablabel{pascalacc}}
@@
 \subsubsection{Pascal VOC 2010}
-For the Pascal VOC 2010 dataset, we follow the procedure from \citet{krahenbuhl2012efficient}
+For the \textsc{Pascal} VOC 2010 dataset, we follow the procedure from \citet{krahenbuhl2012efficient}
 in using the official ``validation'' set as our evaluation set, and splitting the training set again.
@@
 around a Jaccard Index (also call VOC score) of 40\% , notably \cite{xia2012segmentation}, who 
-evaluate on the Pascal VOC 2010 ``test'' set.
+evaluate on the \textsc{Pascal} VOC 2010 ``test'' set.
 
@@
 
-Results on the Pascal-VOC dataset are shown in \Tabref{pascalacc}.
+Results on the \textsc{Pascal} VOC dataset are shown in \Tabref{pascalacc}.
 We compare against several approaches using the same unary potentials.
@@
     \end{center}
-    \caption{Objective function values on the Pascal Dataset.}
+    \caption{Objective function values on the \textsc{Pascal} VOC Dataset.}
     \tablabel{pascal_objective}

### 2013-09-26 intro.tex
smallcaps for datasets, some more corrections by vlad

@@
 different datasets in this thesis: the object class segmentation datasets
-Graz-02~\citep{marszatek2007accurate} and Pascal VOC 2010~\citep{pascal}, and
+\textsc{Graz}-02~\citep{marszatek2007accurate} and \textsc{Pascal} VOC 2010~\citep{pascal}, and
 the semantic segmentation datasets MSRC-21~\citep{shotton2006textonboost} and
@@
 \caption{%
-    Examples from the MSRC-21 (top) and Pascal VOC (bottom) datasets with
+    Examples from the MSRC-21 (top) and \textsc{Pascal} VOC (bottom) datasets with
     ground-truth annotation. MSRC-21 contains mostly texture classes, such as
     tree, building, street and sky, but also objects, like cars in this example.
-    Pascal VOC contains only object classes, such as person, cat, table and
+    \textsc{Pascal} VOC contains only object classes, such as person, cat, table and
     bottle, and an additional background class (black).
@@
 semantic image segmentation exactly, even though the involved factor graphs
-contain many loops. We demonstrate our algorithm on the Pascal VOC 2010, where we are competative
+contain many loops. We demonstrate our algorithm on the \textsc{Pascal} VOC 2010, where we are competative
 with comparable approaches, and MSRC-21 datasets where we improve upon the state-of-the-art.

### 2013-09-26 itm/paper.tex
smallcaps for datasets, some more corrections by vlad

@@
 We see that ITM does well when the intrinsic dimensionality of the data matches
-the feature dimension, but degraded otherwise (see ``faces'' and ``usps'').
+the feature dimension, but degraded otherwise (see \textsc{faces} and \textsc{usps}).
 Estimating the intrinsic dimensionality of the data overcomes this weakness,
@@
 \cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
-digits      &$ 1797$&$  64$&$ 10$& $    0.62 $&     $0.67$           &$0.10$         &$\mathbf{0.85}$       &$0.73$\\ 
-faces       &$  400$&$4096$&$ 40$& $    0.41 $&     $0.49$           &$0.08$         &$0.02$           &$\B{0.54}$\\ 
-iris        &$  150$&$   4$&$  3$& $    0.72 $&     $0.75$           &$0.55$         &$\B{0.88}$       &$0.88$\\
-usps        &$ 9298$&$ 256$&$ 10$& $    0.52 $&     $0.54$           &$0.00$         &$   0.44$        &$\B{0.64}$\\ 
-vehicle     &$  846$&$  18$&$  4$& $    0.10 $&     $0.09$           &$0.00$         &$\B{0.10}$       &$0.10$\\
-vowel       &$  990$&$  10$&$ 11$& $    0.17 $&     $0.19$           &$0.00$         &$\B{0.20}$       &$0.19$\\
-waveform    &$ 5000$&$  21$&$  2$& $\B{0.37} $&     $0.30$           &$0.00$         &$   0.23 $       &$0.23$\\
-mnist       &$70000$&$ 784$&$ 10$& $    0.37 $&     N/A$^\dagger$    &$0.00$         &$   0.50$        &$\B{0.77}$\\
+\textsc{digits}      &$ 1797$&$  64$&$ 10$& $    0.62 $&     $0.67$           &$0.10$         &$\mathbf{0.85}$       &$0.73$\\ 
+\textsc{faces}       &$  400$&$4096$&$ 40$& $    0.41 $&     $0.49$           &$0.08$         &$0.02$           &$\B{0.54}$\\ 
+\textsc{iris}        &$  150$&$   4$&$  3$& $    0.72 $&     $0.75$           &$0.55$         &$\B{0.88}$       &$0.88$\\
+\textsc{usps}        &$ 9298$&$ 256$&$ 10$& $    0.52 $&     $0.54$           &$0.00$         &$   0.44$        &$\B{0.64}$\\ 
+\textsc{vehicle}     &$  846$&$  18$&$  4$& $    0.10 $&     $0.09$           &$0.00$         &$\B{0.10}$       &$0.10$\\
+\textsc{vowel}       &$  990$&$  10$&$ 11$& $    0.17 $&     $0.19$           &$0.00$         &$\B{0.20}$       &$0.19$\\
+\textsc{waveform}    &$ 5000$&$  21$&$  2$& $\B{0.37} $&     $0.30$           &$0.00$         &$   0.23 $       &$0.23$\\
+\textsc{MNIST}       &$70000$&$ 784$&$ 10$& $    0.37 $&     N/A$^\dagger$    &$0.00$         &$   0.50$        &$\B{0.77}$\\
 \bottomrule

### 2013-09-26 main.tex
some graphics, minor fixes.
smallcaps for datasets, some more corrections by vlad

@@
 factor graph. Our methods achieve competitive performance with similar
- methods on the Pascal VOC 2010 dataset, and improve upon state-of-the-art
+methods on the \textsc{Pascal} VOC 2010 dataset, and improve upon state-of-the-art
  results on the MSRC-21 dataset.
@@
 over a set of candidate segments. We demonstrated the feasibility and effectiveness
-of our approach on the challenging Graz-02 dataset of street scenes.
+of our approach on the challenging \textsc{Graz}-02 dataset of street scenes.
 
@@
 %
-\paragraph{Theoretical Analysis of $n$-slack Algorithm}
+\paragraph{Theoretical Analysis of the $n$-Slack Algorithm}
 As we have seen in Chapter~\ref{ch:comparison}, the $n$-slack algorithm often

### 2013-09-26 nyu/paper.tex
smallcaps for datasets, some more corrections by vlad

@@
 scene classes.  The annotation contains four semantic structural classes:
-structure, floor, furniture and prop. As in the MSRC-21 and Pascal VOC
+structure, floor, furniture and prop. As in the MSRC-21 and \textsc{Pascal} VOC
 datasets, there is an additional ``void'' class, which is used for object

### 2013-09-26 partial_supervision/paper.tex
smallcaps for datasets, some more corrections by vlad

@@
 trained on these segments to estimate the overlap of segments with ground truth
-object-class labeling from the Pascal VOC dataset~\citep{pascal}. This provides
+object-class labeling from the \textsc{Pascal} VOC dataset~\citep{pascal}. This provides
 a ranking of candidate segments according to how ``object-like'' they are,
@@
 The method performed well on a variety of datasets, building the basis of
-a very successful entry to the Pascal VOC segmentation
+a very successful entry to the \textsc{Pascal} VOC segmentation
 challenge~\citep{li2010object}. A similar approach to whole-object segment
@@
 We evaluate the performance of the proposed algorithm for object-class
-segmentation on the challenging Graz-02 dataset~\citep{marszatek2007accurate}.  This dataset contains 1096
+segmentation on the challenging \textsc{Graz}-02 dataset~\citep{marszatek2007accurate}.  This dataset contains 1096
 images of three object classes, bike, car and person.  Each image may contain
@@
 	\end{center}
-        \caption{Qualitative results on on the Graz-02 dataset. Top: Results on
+    \caption{Qualitative results on on the \textsc{Graz}-02 dataset. Top: Results on
         category ``car''. Bottom: Results on category ``person''. From left to
@@
     \end{tabularx}
-    \caption{Pixel-level accuracy on the Graz-02 dataset.
+    \caption{Pixel-level accuracy on the \textsc{Graz}-02 dataset.
         \tablabel{tab:graz}}
@@
 We evaluated the proposed object-class segmentation method on the challenging
-Graz-02 dataset. While not reaching the performance of methods using strong
+\textsc{Graz}-02 dataset. While not reaching the performance of methods using strong
 supervision, our result can work as a baseline for further research into weakly

### 2013-09-26 pystruct/paper.tex
smallcaps for datasets, some more corrections by vlad

@@
      as the popular \svmstruct model for cutting plane algorithms, and
-     provides implementations of the BCFW and Subgradient methods that scale to
+     provides implementations of the SZLJSP and Subgradient methods that scale to
      large datasets.
@@
 the main focus of this work. The example shows how to learn an $n$-slack support
-vector machine on a superpixel-based CRF on the Pascal dataset. Details of the experiment
+vector machine on a superpixel-based CRF on the \textsc{Pascal VOC} dataset. Details of the experiment
 can be found in Section~\secref{semantic_segmentation}.

### 2013-09-26 structured_prediction/paper.tex
smallcaps for datasets, some more corrections by vlad

@@
 \Eqref{log_factor_graph} instead.
-
-% TODO relate to computer vision here ??? Pairwise models?
-% talk about inference?
 
@@
 
-%TODO BMRM?
-
 \subsection{The SZLJSP Algorithm}\seclabel{dual_coordinate_descent}
@@
 model that uses superpixels contains much fewer variables than one that works
-on pixel-level.  Typical sizes for superpixels are in the hundereds of pixels,
+on pixel-level. Typical sizes for superpixels are in the hundreds of pixels,
 leading to a hundred-fold decrease in the number of variables. There are also
@@
 object boundaries.  The main disadvantage of superpixel-based approaches is
-that the initial oversegmentation can not be influenced by later reasoning.
+that the initial over-segmentation can not be influenced by later reasoning.
 This can be a problem if fine structures need to be segmented for which local
@@
 graph cut approach of \citet{boykov2001fast} for inference.  However, graph-cut
-inference is only applicable to submular energies, and therefore severly
-restricts the expressivity of the resulting model.
+inference is only applicable to submodular energies, and therefore severely
+restricts the expressiveness of the resulting model.
 \citet{lucchi2011spatial} investigated the importance of global constraints,
@@
 When evaluating learning algorithms in Chapter~\ref{ch:comparison} and
-Chapter~\ref{ch:exact_learning} we use the well-established Pascal VOC 2010 and
+Chapter~\ref{ch:exact_learning} we use the well-established \textsc{Pascal} VOC 2010 and
 MSRC-21 datasets, shown in the introduction in \Figref{pascal_msrc}.
@@
 
-For both datasets we use the TextonBoost class probablities provided by
-\citet{krahenbuhl2012efficient}. For the Pascal VOC dataset, the potentials
+For both datasets we use the TextonBoost class probabilities provided by
+\citet{krahenbuhl2012efficient}. For the \textsc{Pascal} VOC dataset, the potentials
 provided by \citet{krahenbuhl2012efficient} also include the responses of
@@
 We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
-our experiments.  It has been shown to provide competative results with a
+our experiments.  It has been shown to provide competitive results with a
 minimum of computational complexity.  Algorithmically, SLIC simply computes a
@@
 points feasible, the search for the nearest cluster in $k$-means is restricted
-to a local neighborhood in the image. An example of superpixels computed with the SLIC algorithm can be found in
-\Figref{superpixel_illustration}.
+to a local neighborhood in the image. An example of superpixels computed with
+the SLIC algorithm can be found in \Figref{superpixel_illustration}.
 
-The features and superpixel algorithm used for the RGB-D dataset NYU used in Chapter~\ref{ch:nyu} will be discussed there.
+The features and superpixel algorithm used for the RGB-D dataset NYU used in
+Chapter~\ref{ch:nyu} will be discussed there.

### 2013-09-26 evaluation/paper.tex
move interesting phenomenon back to letters.

@@
 with the SZLJSP clearly leading the way.
-%Interestingly, the $1$-slack algorithm
-%catches up near the desired suboptimality and is faster in achieving the desired duality gap--even though
-%the primal objective is always above SZLJSP\@. A zoom-in can be found in
-%\Figref{ocr_small_duality}, right hand side.
-
-We want to highlight an interesting phenomenon here: The cutting-plane algorithms stop immediately
-when reaching the desired suboptimality of $0.1$. The SZLJSP algorithm, on the other hand, keeps
-on learning much longer---even though the primal of the $1$-slack algorithms is
-always above the primal of the SZLJSP\@. This is caused by a looser bound
-given by the dual. Remember that the stopping criterion for both cutting-plane 
-and SZLJSP are given by the duality gap. It seems that while the primal objective converges much
-faster in the SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does not.
-This means that in this practical experiment, the non-caching $1$-slack algorithm was \emph{faster}
-in guaranteeing the desired suboptimality. \Figref{ocr_small_dual} illustrates this point by plotting primal and dual
-objectives for the $1$-slack cutting plane and the SZLJSP algorithms.
- Looking at the training set loss in \Figref{mnist_curves} suggests that the
- high precision we demanded was not necessary, and we could have terminated the
- stochastic algorithms much earlier.
+Interestingly, the $1$-slack algorithm catches up near the desired
+suboptimality.  Looking at the training set loss in \Figref{mnist_curves}
+suggests that the high precision we demanded was not necessary, and we could
+have terminated the stochastic algorithms much earlier.
 
@@
 dataset.
-%
-As on MNIST, the $1$-slack algorithm stops before the SZLJSP (in terms of
-iterations) because of a tighter dual.  A detailed view is shown in
-\Figref{ocr_small_dual} on the right.
+
+We want to highlight an interesting phenomenon here: The cutting-plane
+algorithms stop immediately when reaching the desired suboptimality of $0.1$.
+The SZLJSP algorithm, on the other hand, keeps on learning much longer---even
+though the primal of the $1$-slack algorithms is always above the primal of the
+SZLJSP\@. This is caused by a looser bound given by the dual. Remember that the
+stopping criterion for both cutting-plane and SZLJSP are given by the duality
+gap. It seems that while the primal objective converges much faster in the
+SZLJSP algorithm than in the non-caching $1$-slack algorithm, the dual does
+not. This means that in this practical experiment, the non-caching $1$-slack
+algorithm was \emph{faster} in guaranteeing the desired suboptimality, and
+therefore in terminating.  \Figref{ocr_small_dual} illustrates this point by
+plotting primal and dual objectives for the $1$-slack cutting plane and the
+SZLJSP algorithms. Taking a closer look, we find that the same behavior
+occurred for MNIST, which is also shown in \Figref{ocr_small_dual}.
 

### 2013-09-26 evaluation/paper.tex
typo
try to refer to the rows in the plots in a less confusing way

@@
 We applied this method only to the smaller \textsc{scene} dataset, as applying it on the
-\textsc{yeast} dataset was not practicable.
+\textsc{yeast} dataset was not practical.
 
@@
 The situation is very different when using the linear programming relaxation.
-Looking at the four plots at the bottom of \Figref{snakes_results}, we see that all
+Looking at last two rows of \Figref{snakes_results}, we see that all
 algorithms are able to solve the task perfectly. Evaluating on the test-set, all algorithms
@@
 \caption{%
-    Results on the \textsc{snakes} dataset. The top four plots show results using fusion
-    move inference, the bottom four plots using AD$^3$. Dashed lines indicate the dual objective.
+    Results on the \textsc{snakes} dataset. The first two rows show results using fusion
+    move inference, the two rows below using AD$^3$. Dashed lines indicate the dual objective.
 \figlabel{snakes_results}}
@@
 to use the fusion move inference algorithm. The results are shown in
-\Figref{msrc_pascal_curves}, with the top four plots showing results on
-MSRC-21, and the bottom four plots showing results on \textsc{Pascal VOC}.
+\Figref{msrc_pascal_curves}, with the plots in the two upper rows showing results on
+MSRC-21, and the tow rows below showing results on \textsc{Pascal VOC}.
 First, we notice that all algorithms achieve low (approximate) duality gaps, and none
@@
 \caption{%
-    Objective and traning set loss on the segmentation task. The top four plots
-    show results for the MSRC-21 dataset, the bottom four plots for the \textsc{Pascal
+    Objective and traning set loss on the segmentation task. The first two rows
+    show results for the MSRC-21 dataset, the rows below for the \textsc{Pascal
     VOC} dataset.

### 2013-09-26 exact_learning/paper.tex
more fixes via vlad :)

@@
 approach improves upon the state-of-the-art on the MSRC-21 dataset, and is
-competative with comparable approaches on the \textsc{Pascal VOC} dataset.% and nyu?
+competative with comparable approaches on the \textsc{Pascal VOC} dataset.
 
@@
 learning as a black-box, enabling the researcher to focus attention on
-designing the model for the application at hand.
+designing the model for the application at hand. It also makes research more
+reproducible, as the particular optimization methods that are used become less
+relevant to the result.
 
@@
             \vspace{-5mm}
-            \begin{align*}
-                (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi\\
-            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
-            &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
-            \end{align*}
+            \begin{flalign*}
+            \quad\quad    (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi&\\
+            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:&\\
+            &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi&
+            \end{flalign*}
             \For {i=1, \dots, k}
@@
 oracles in the context) that do not yield the exact maximum here.
-There are two groups of inference procedures, as identified in
+There are two groups of inference procedures, as identified by
 \citet{finley2008training}: under-generating and over-generating approaches.
@@
 
-Let $o_{\W}(\theta)$ denote objective \Eqref{oneslack_again} with
+Let $o_{\W}(\theta)$ denote the objective in \Eqref{oneslack_again} with
 given parameters $\theta$ restricted to a working set $\W$, as computed in
@@
 search procedure, using moves that are generated using an auxiliar binary problem,
-which is solved using QPBO~\cite{rother2007optimizing}.
+which is solved using QPBO~\citep{rother2007optimizing}.
 
@@
 between the methods. Note that neither $o^Q$ nor $o^C$ provide valid upper bounds on the objective,
-which is particularly visible for $o^C$ using the method of \cite{joachims2009cutting}.
+which is particularly visible for $o^C$ using the method of \citet{joachims2009cutting}.
 Using no cache leads to a relatively smooth, but slow convergence, as inference is run often.
@@
 Our heuristic instead probes the cache, and only proceeds using cached constraints if the resulting
-$o^C$ is not to far from $o^Q$.
+$o^C$ is not too far from $o^Q$.
 
@@
 Note that state-of-the-art approaches, some not build on the CRF framework, obtain
-around a Jaccard Index (also call VOC score) of 40\% , notably \cite{xia2012segmentation}, who 
+around a Jaccard Index (also call VOC score) of 40\% , notably \citet{xia2012segmentation}, who 
 evaluate on the \textsc{Pascal} VOC 2010 ``test'' set.

### 2013-09-26 nyu/paper.tex
more fixes via vlad :)

@@
 of \citet{SilbermanECCV12} with promising results.  We use the implementation
-of random forest provided by \citet{stueckler2013}, but instead of integrating
+of random forests provided by \citet{stueckler2013}, but instead of integrating
 predictions over time, we focus on exploiting the structure within a single
@@
 hyper parameters of the random forests were found using the \textsc{hyperopt}
-framework of \cite{bergstra2011algorithms}. For the CRF model, the only
+framework of \citet{bergstra2011algorithms}. For the CRF model, the only
 hyper-parameters are related to the superpixel segmentation, and the single
@@
 
-We did not explicitly address real time application, the random
+We did not explicitly address real time application; the random
 forest implementation that we build upon allows for real-time

### 2013-09-26 evaluation/paper.tex
remove typos, thank vlad some more :)

@@
 easily. We use the dynamic programming algorithm implemented in \textsc{OpenGM}
-for experiments for the tree and chain models.
+for experiments using the tree and chain models.
 For the full pairwise multi-label model, we use branch-and-bound together with
-the AD$^3$ algorithm. %FIXME AD3 solver description?
+the AD$^3$ algorithm.
 We found that using the linear programming relaxation on the \textsc{scene} dataset is
@@
 %
-We use two different approximate inference algorithms, fusion moves, a fast
-local search procedure, and the linear programming relaxation provided by AD$^3$
+We use two different approximate inference algorithms, fusion
+moves~\citep{lempitsky2010fusion}, a fast local search procedure, and the
+linear programming relaxation provided by AD$^3$~\citep{martins2011augmented}
 not using branch-and-bound, in contrast to the multi-label setup above.

### 2013-09-26 main.tex
more fixes via vlad :)
remove typos, thank vlad some more :)

@@
 many helpful discussions, and advice.
-
+%
 I would also like to thank Carsten Rother for allowing me to work with him
@@
 
-I have been lucky to be a part of the \textsc{scikit-learn} community, whose developers
-have been a constant source of encouragement and inspiration to me. My
-understanding of machine learning as well as software engineering has been
+My research has been partially funded by the B-IT research school that I want to thank for my scholarship.
+
+I have been lucky to be a part of the \textsc{scikit-learn} community, whose
+developers have been a constant source of encouragement and inspiration to me.
+My understanding of machine learning as well as software engineering has been
 greatly affected by the continuing exchange with Olivier Grisel, Lars Buitinck,
-Ga\"el Varoquax, Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and
-others.
+Ga\"el Varoquax, Mathieu Blondel, Gilles Louppe, Arnaud Joly and others. In
+particular I want to thank Vlad Niculae for his help during the formation of
+this work.
 
@@
 benefits immensely from caching inference results during training. Therefore, investigating
-the influence of caching for SZLJSP seems a promising topic for future research.
+the influence of caching for SZLJSP~(see \Secref{dual_coordinate_descent})
+seems a promising topic for future research.
 %

### 2013-09-27 main.tex
Move table of content, fix second reader

@@
 1. Gutachter Prof. Dr. Sven Behnke\\
-2. Gutachter Prof. Dr. Reinhard Klein\\
+2. Gutachter Prof. Dr. J\"urgen Gall\\
 Tag der Promotion:\\
@@
 \end{spacing}
-\tableofcontents
 %\listoffigures
@@
 
-My research has been partially funded by the B-IT research school that I want to thank for my scholarship.
+The B-IT research school kindly provided funding for parts of my research,
+which I am also thankful for.
 
@@
 endeavour.
+
+\tableofcontents
+
 \cleardoublepage

### 2013-09-27 evaluation/paper.tex
k-means not capitalizes, minor fixes

@@
 semantic segmentation.  We use the \textsc{Pascal VOC} and MSRC datasets.
-The \textsc{Pascal VOC} has 964 training images, each divided into
-around 100 superpixels. There are 20 object categories, and an additional background class. The % TODO check! refer to other part
-MSRC-21 dataset has 276 training images, also segmented in around 100 superpixels
-each. There are 21 classes in the MSRC-21 dataset.
+The \textsc{Pascal VOC} has $964$ training images, each divided into
+around $100$ superpixels. There are $20$ object categories, and an additional background class. The % TODO check! refer to other part
+MSRC-21 dataset has $276$ training images, also segmented in around $100$ superpixels
+each. There are $21$ semantic classes in the MSRC-21 dataset.
 Each superpixel is represented as an output variable, with the ground truth
@@
 removed all superpixels in which the majority of pixels is labeled ``void'', leading to some samples
-having much fewer than 100 variables.
+having much fewer than $100$ variables.
 Pairwise potentials are introduced for each pair of neighboring superpixels.
 We use the unary and pairwise potentials described in \Secref{semantic_segmentation}:
-21 unary features for the \textsc{Pascal VOC} dataset and 63 unary features for the MSRC-21 dataset.
+$21$ unary features for the \textsc{Pascal VOC} dataset and $63$ unary features for the MSRC-21 dataset.
 We use the same three pairwise features for both datasets: a constant feature, a color
@@
     \cmidrule{1-6}
-    MNIST               & 60000   & 1     & none  & 7840  & 10\\
-    \textsc{OCR-small}  & 704     & 3--14 & chain & 1444  & 26\\
-    \textsc{OCR-large}  & 6173    & 3--14 & chain & 1444  & 26\\
-    \textsc{scene-tree} & 1211    & 6     & tree  & 1244  & 2\\
-    \textsc{scene-full} & 1211    & 6     & loopy & 1284  & 2\\
-    \textsc{yeast-tree} & 1500    & 14    & tree  & 1494  & 2\\
-    \textsc{yeast-full} & 1500    & 14    & loopy & 1806  & 2\\
-    \textsc{snakes}     & 200     &84--168& loopy & 22275 & 11\\
-    MSRC-21             & 276     & 7--113& loopy &  1764 & 21\\
-    \textsc{Pascal VOC} & 964     & 8--112& loopy &  2646 & 21\\
+    MNIST               &$60000$  &$ 1     $ & none  &$7840 $& $10$ \\
+    \textsc{OCR-small}  &$704  $  &$ 3--14 $ & chain &$1444 $& $26$ \\
+    \textsc{OCR-large}  &$6173 $  &$ 3--14 $ & chain &$1444 $& $26$ \\
+    \textsc{scene-tree} &$1211 $  &$ 6     $ & tree  &$1244 $& $ 2$ \\
+    \textsc{scene-full} &$1211 $  &$ 6     $ & loopy &$1284 $& $ 2$ \\
+    \textsc{yeast-tree} &$1500 $  &$ 14    $ & tree  &$1494 $& $ 2$ \\
+    \textsc{yeast-full} &$1500 $  &$ 14    $ & loopy &$1806 $& $ 2$ \\
+    \textsc{snakes}     &$200  $  &$84--168$ & loopy &$22275$& $11$ \\
+    MSRC-21             &$276  $  &$ 7--113$ & loopy &$ 1764$& $21$ \\
+    \textsc{Pascal VOC} &$964  $  &$ 8--112$ & loopy &$ 2646$& $21$ \\
     \bottomrule
@@
     \item The $n$-slack cutting plane algorithm, where the QP is solved after each sample
-    \item The $n$-slack cutting plane algorithm, where the QP is solved every 100 samples
+    \item The $n$-slack cutting plane algorithm, where the QP is solved every $100$ samples
     \item The SZLJSP algorithm with weighted averaging
@@
 While the two variants of the $n$-slack algorithm are equally fast with respect
-to passes over the training data, solving the QP every 100 steps is among the
+to passes over the training data, solving the QP every $100$ steps is among the
 fastest algorithms, while solving the QP at every step is among the slowest.
@@
 algorithms are able to solve the task perfectly. Evaluating on the test-set, all algorithms
-achieve an accuracy of around 99.5\%, confirming the models actually learned the task.
+achieve an accuracy of around $99.5\%$, confirming the models actually learned the task.
 Surprisingly, SSGD was very successful on this dataset, even faster then the non-caching $n$-slack
@@
 training set error, and the $n$-slack and stochastic algorithms are much
-faster. Solving the QP only every 100 steps in the $n$-slack algorithm significantly slows
+faster. Solving the QP only every $100$ steps in the $n$-slack algorithm significantly slows
 down learning on both datasets. This is in contrast to the \textsc{yeast} datasets, where the converse was true,
@@
     \item In nearly all experiments, caching significantly improved performance
-        of the 1-slack cutting plane algorithm.
-    \item Initially SZLJSP and even SSGD converge faster than the 1-slack
-        cutting plane algorithm, while the 1-slack cutting plane algorithm was
+        of the $1$-slack cutting plane algorithm.
+    \item Initially SZLJSP and even SSGD converge faster than the $1$-slack
+        cutting plane algorithm, while the $1$-slack cutting plane algorithm was
         faster in achieving high precision solutions and guaranteeing low

### 2013-09-27 exact_learning/paper.tex
some moving around of figures and tables
k-means not capitalizes, minor fixes

@@
 
-\begin{algorithm*}[t]
-    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack_again}}
-    \begin{algorithmic}[1]
-        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$, inference oracle $\hat{I}$.
-        \Ensure parameters $\theta$, slack $\xi$
-        \State $\W \leftarrow \emptyset$
-        \Repeat
-            \State \label{line_restricted}
-            \vspace{-5mm}
-            \begin{flalign*}
-            \quad\quad    (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi&\\
-            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:&\\
-            &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi&
-            \end{flalign*}
-            \For {i=1, \dots, k}
-                \State
-            $\hat{y}\hoch{i} \leftarrow \hat{I}(x\hoch{i}, y\hoch{i}, \theta) \approx \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane_again}
-            \EndFor
-            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
-        \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] $
-        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
-    \end{algorithmic}
-\end{algorithm*}
-
 \section{Learning SSVMs with Approximate Inference}
@@
 %Section~\ref{bounds}. %TODO Filler sentence? wat?
+\begin{algorithm*}[t]
+    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack_again}}
+    \begin{algorithmic}[1]
+        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$, inference oracle $\hat{I}$.
+        \Ensure parameters $\theta$, slack $\xi$
+        \State $\W \leftarrow \emptyset$
+        \Repeat
+            \State \label{line_restricted}
+            \vspace{-5mm}
+            \begin{flalign*}
+            \quad\quad    (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi&\\
+            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:&\\
+            &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi&
+            \end{flalign*}
+            \For {i=1, \dots, k}
+                \State
+            $\hat{y}\hoch{i} \leftarrow \hat{I}(x\hoch{i}, y\hoch{i}, \theta) \approx \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane_again}
+            \EndFor
+            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
+        \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] $
+        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
+    \end{algorithmic}
+\end{algorithm*}
 
@@
 It is easy to see that our heuristic runs inference only $O(\log(o^Q -
-o_{\W}))$ times more often than the strategy from \citet{joachims2009cutting} in the
+o_{\W}))$ times more often than the strategy of \citet{joachims2009cutting} in the
 worst case.
@@
 \subsubsection{Pascal VOC 2010}
-For the \textsc{Pascal} VOC 2010 dataset, we follow the procedure from \citet{krahenbuhl2012efficient}
+For the \textsc{Pascal} VOC 2010 dataset, we follow the procedure of \citet{krahenbuhl2012efficient}
 in using the official ``validation'' set as our evaluation set, and splitting the training set again.
@@
 
-\begin{figure*}
+\begin{figure*}[tb]
 \centering

### 2013-09-27 intro.tex
k-means not capitalizes, minor fixes

@@
     structured prediction. 
-\item Analysing max-margin learning algorithms with exact or approximate
+\item Analysing max-margin learning algorithms with exact and approximate
     inference in different applications. We give a thorough evaluation of
-    all major SSVM learning algorithms, using a wide array of application.
+    all major SSVM learning algorithms in a wide array of application.
 \item Showing that exact learning for semantic segmentation and object class

### 2013-09-27 itm/paper.tex
k-means not capitalizes, minor fixes

@@
 \section{Related Work}
-The most commonly used clustering algorithm is the $k$-Means algorithm, originally
+The most commonly used clustering algorithm is the $k$-means algorithm, originally
 investigated by \citet{macqueen1967some} and most commonly implemented using
 Lloyd's algorithm \citep{macqueen1967some,lloyd1982least}. While
-$k$-Means often works well in practice, one of its main drawbacks is the
+$k$-means often works well in practice, one of its main drawbacks is the
 restriction in cluster shape. 
@@
 graph constructed from the data. While spectral clustering is much more
-flexible than $k$-Means it is quite sensitive to the particular choice of graph
+flexible than $k$-means it is quite sensitive to the particular choice of graph
 construction and similarity measure. It is also computationally expensive to
@@
 %clusters, in practice the result is strongly biased towards compact, convex
-%regions. This can be explained by a close connections of MeanNN to $k$-Means,
+%regions. This can be explained by a close connections of MeanNN to $k$-means,
 %see Section~\ref{discussion} for an in-depth discussion.
@@
 %
-\section[Clustering Using Non-Parametric Entropy Estimates]{Clustering Using\linebreak Non-Parametric Entropy Estimates}
+\section[Clustering using Non-Parametric Entropy Estimates]{Clustering Using\linebreak Non-Parametric Entropy Estimates}
 In general, the goal of clustering can be formulated as follows: 
@@
 Algorithm~\ref{overall} has a run time of $O(n)$ per removed edge, leading to an overall
-run time of $O\bigl(n \log(n) \alpha(n)+ nk\bigr)$. The $O(nk)$ comes from a worst case
+run time of $O\bigl(n \log(n) \alpha(n)+ nk\bigr)$. The $O(nk)$ term comes from a worst case
 scenario, in which each step in the hierarchical clustering procedure only
@@
 that the individual clusters are much smaller than the original dataset. In
-this case, the $O(nk)$ factor would improve to $O\bigl(n\log(k)\bigr)$.
+this case, the $O(nk)$ term would improve to $O\bigl(n\log(k)\bigr)$.
 
@@
 accurately reflect the intrinsic dimensionality of the data.
-We use the estimate analyzed in \citet{massoud2007manifold}.
+We use the estimate analyzed by \citet{massoud2007manifold}.
 In their method, for each data point $x$, a local estimate $\hat{d}(x)$ of the
@@
 \includegraphics[width=.24\textwidth]{qualitative_twomoons_mst} \\ 
-\caption{Comparison of $k$-Means (left), MeanNN (center left), single link
+\caption{Comparison of $k$-means (left), MeanNN (center left), single link
 (center right) and ITM (right) on four synthetic datasets.  Without the need to
@@
 recover non-convex clusters (third row) but often produces similar results to
-$k$-Means (second and last row). Single link clustering is very sensitive
+$k$-means (second and last row). Single link clustering is very sensitive
 to noise, as it does not take cluster size into account.}
@@
 \section{Experiments}
-We compared ITM to the popular $k$-Means algorithm 
+We compared ITM to the popular $k$-means algorithm 
 \citep{macqueen1967some,lloyd1982least}, to the MeanNN algorithm of
@@
 \enlargethispage{5mm}
-A comparison of ITM, MeanNN and the baseline methods, $k$-Means and 
+A comparison of ITM, MeanNN and the baseline methods, $k$-means and 
 single link agglomerative clustering, in terms of their objective, 
@@
 We implemented the ITM clustering procedure as well as MeanNN in Python.  We
-used the $k$-Means implementation available in the \textsc{scikit-learn}
+used the $k$-means implementation available in the \textsc{scikit-learn}
 library~\citep{pedregosa2011scikit}. The source
@@
 \subsection{Experimental Setup}
-For both $k$-Means and MeanNN, we restart the algorithm ten times using
+For both $k$-means and MeanNN, we restart the algorithm ten times using
 different random initializations, keeping the result with the best objective
@@
 \cmidrule{1-4}
-$k$-Means &     $\displaystyle \sum_y \sum_{i, y_i = y} \| x_i - \mu_y \|^2$ & No  & $O(nk)$ per iteration%
+$k$-means &     $\displaystyle \sum_y \sum_{i, y_i = y} \| x_i - \mu_y \|^2$ & No  & $O(nk)$ per iteration%
 %\footnote{The worst-time complexity of the Lloyd-Algorithm is exponential~\cite{arthur2006slow}, though online algorithm are observed to converge in linear time~\cite{bottou1995convergence}.}
@@
 results.  It is also noteworthy that the results of MeanNN are very close to
-those produces by $k$-Means in most cases. This similarity can be explained
+those produces by $k$-means in most cases. This similarity can be explained
 by the close relation of the objective functions, listed in Table~\ref{nowotab}.
@@
 of cluster quality~\citep{gomes2010discriminative,kamvar2003spectral}.
-The Rand index~\citep{rand1971objective} between two clusterings counts on how
+The Rand index~\citep{rand1971objective} between two clusterings counts how
 many pairs of points two clusterings agree. The adjusted Rand index contains a
@@
 \cmidrule{5-9}
-Description &     n &    d &   k &  $k$-Means      & MeanNN        & SL            &   ITM        & ITM ID \\
+Description &     n &    d &   k &  $k$-means      & MeanNN        & SL            &   ITM        & ITM ID \\
 \cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
@@
 \end{tabularx}
-\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative
+\caption{Adjusted Rand Index of $k$-means, MeanNN, single link agglomerative
     clustering and ITM on several datasets (higher is better). ITM ID refers to
@@
 algorithms and the non-parametric entropy based clustering of
-\citep{faivishevsky2010nonparametric} on multiple benchmark datasets. We
+\citet{faivishevsky2010nonparametric} on multiple benchmark datasets. We
 demonstrated that ITM is able to detect non-convex clusters,

### 2013-09-27 nyu/paper.tex
remove typos, thank vlad some more :)
Fix figure allignment in nyu
some moving around of figures and tables
k-means not capitalizes, minor fixes

@@
 
-\begin{figure}
+\begin{figure}[t]
     \begin{center}
@@
 extend the standard SLIC algorithm, which works on the Lab space, to also
-include depth information. The resulting algorithm is a localized $k$-Means in
+include depth information. The resulting algorithm is a localized $k$-means in
 Lab-D-XY space.
@@
 We use a very simple approach that we found effective: we cluster normal
-directions of all pixels into 10 clusters using $k$-means, and use the one that
+directions of all pixels into $10$ clusters using $k$-means, and use the one that
 is most parallel to the $Y$ direction, which roughly corresponds to height in
@@
 little use in scenes without horizontal surfaces.
-\begin{figure}
+\begin{figure}[tb]
     \begin{center}
@@
 \toprule
-                        & \footnotesize{ground}        &  \footnotesize{structure}    & \footnotesize{furniture}     & \footnotesize{props}         & \footnotesize{class avg.}   & \footnotesize{pixel avg.}\\
+                        & \footnotesize{ground}        &  \footnotesize{structure}    & \footnotesize{furniture}     & \footnotesize{prop}         & \footnotesize{class avg.}   & \footnotesize{pixel avg.}\\
 \cmidrule{1-7}
@@
 diagonal of the confusion matrix), which puts more emphasis on the harder
-classes of props and furniture, which have smaller area than structure and
+classes of prop and furniture, which have smaller area than structure and
 floor.
@@
 %
-All hyper parameters were adjusted using 5-fold cross-validation. The
+All hyper-parameters were adjusted using 5-fold cross-validation. The
 hyper parameters of the random forests were found using the \textsc{hyperopt}
@@
 \Tabref{results} compares different components of our approach with the literature.
-Please note that we \emph{first} designed our final model, using only the
+Note that we \emph{first} designed our final model, using only the
 validation data. We now report accuracies of simpler models for reference,
@@
 which is not surprising. We see that evidence from the random forest is often very noisy, and biased
-away from the ``props'' class. While the unstructured SVM can correct somewhat
+away from the ``prop'' class. While the unstructured SVM can correct somewhat
 for the class imbalance, it has no way to make larger areas consistent, which
@@
     \begin{center}
-        \includegraphics[width=.45\linewidth]{nyu/images/vertical_alignment}
-    \includegraphics[width=.45\linewidth]{nyu/images/depth_difference}
+        \includegraphics[height=.40\linewidth]{nyu/images/vertical_alignment}
+    \includegraphics[height=.40\linewidth]{nyu/images/depth_difference}
     \end{center}
@@
 vertical alignment potential expresses that the floor is much more likely to be below
-other classes. It also encodes the fact that props
+other classes. It also encodes the fact that prop
 rest on furniture, but not the other way around.

### 2013-09-27 partial_supervision/paper.tex
k-means not capitalizes, minor fixes

@@
 method we use for generating candidate segments is Constraint Parametric
-Min-Cuts (CPMC) from \citet{carreira2010constrained}.  This method creates a
+Min-Cuts (CPMC) of \citet{carreira2010constrained}.  This method creates a
 wide variety of overlapping segments. Support vector regression (SVR) is
@@
 To generate proposal segments, we use the Constraint Parametric Min-Cuts (CPCM)
-method from \citet{carreira2010constrained}. In CPMC, initial segments are
+method of \citet{carreira2010constrained}. In CPMC, initial segments are
 constructed using graph cuts on the pixel grid. The energy function for these cuts

### 2013-09-27 pystruct/paper.tex
move interesting phenomenon back to letters.
k-means not capitalizes, minor fixes

@@
         algorithms implemented in popular structured prediction software
-        packages. 1-CP stands for 1-slack Cutting Plane, $n$-CP for $n$-slack
+        packages. 1-CP stands for $1$-slack Cutting Plane, $n$-CP for $n$-slack
         Cutting plane, SSGD for stochastic subgradient decent learning of
@@
     \caption{\label{table:comparision_models}Comparison of models implemented
-    in popular structured prediction software packages. LSVM stands for the latent multi-class SVM, LDCRF for
-    latent dynamic conditional random fields.}
+    in popular structured prediction software packages. LSVM stands for the latent
+    multi-class SVM, LDCRF for latent dynamic conditional random fields.}
 \end{table}
@@
 vector machine on a superpixel-based CRF on the \textsc{Pascal VOC} dataset. Details of the experiment
-can be found in Section~\secref{semantic_segmentation}.
+can be found in \Secref{semantic_segmentation}.
 Each sample (corresponding on one entry of the list \texttt{X}) is represented as a
@@
 model using the $n$-slack cutting plane method, and line 6 performs the actual
-learning.  Using this simple setup, we achieve an accuracy of 30.3 on the
+learning.  Using this simple setup, we achieve an accuracy (Jaccard index) of 30.3 on the
 validation set following the protocol of \citet{krahenbuhl2012efficient}, who
@@
 Results are shown in \Figref{timings}. We report learning times and accuracy for
-varying regularization parameter $C$. The MNIST dataset has 60\,000 training
-examples, 784 features and 10 classes.%
-\footnote{Details about the experiment and code for the experiments can be found on the project website.}
+varying regularization parameter $C$. The MNIST dataset has $60\,000$ training
+examples, $784$ features and $10$ classes.%
+The setup of the experiment is the same as in Chapter~\ref{ch:comparison}.
 The figure indicates that \pystruct has competitive performance, while using

### 2013-09-27 structured_prediction/paper.tex
some moving around of figures and tables
k-means not capitalizes, minor fixes

@@
 In this chapter, we first introduce the concepts and algorithms used in
-structured prediction, in particular in maximum margin methods. We 
-accompany this by a description of our open source implementation of these
+structured prediction, in particular in maximum margin methods. Then, we review
+the use of CRFs in computer vision, and introduce our methods. Finally we give
+a description of our open source implementation of structured learning
 algorithms, \pystruct.
@@
 \end{equation}
-Finding $y$ in the above equation is often referred to as inference
-or prediction.
-Most current approaches use linear functions, with some notable exceptions:
-\citet{peng2009conditional} introduced a conditional random field for sequence
-prediction where single-node potentials are computed using neural networks,
-while \citet{nowozin2011decision} proposed a conditional random field for image
-processing, where all potentials are computed using decision trees.
-
-The linear approach leads to
+Finding $y$ in the above equation is often referred to as inference or
+prediction.  We will use the most common approach of using a linear parametrization of $g$,
+which leads to
 \begin{equation}\eqlabel{main_equation}
@@
 \end{equation}
-Where $\eta_t$ is a sequence of step sizes.
+Here $\eta_t$ is a sequence of step sizes.
 %Theoretical convergence guarantees of \ldots can be given whenever \ldots,
@@
 \citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
-to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
-and derived a block-coordinate version, where each block corresponds to the
-constraints associated with a single training example.  Both groups showed that
-closed-form line search is possible, and, surprisingly ended up with the exact
-same algorithm and closely related convergence guarantees.  The algorithm
-processes a single training example at a time, and can be applied while
-remaining completely in the primal domain, making it applicable to very large
-datasets.  We refer to the algorithm as SZLJSP, after the authors of the
-two papers.
+to structured support vector machines. \citet{lacoste2012block} started from
+the Frank-Wolfe algorithm and derived a block-coordinate version, where each
+block corresponds to the constraints associated with a single training example.
+Both groups showed that closed-form line search is possible, and, surprisingly
+ended up with the exact same algorithm and closely related convergence
+guarantees.  The algorithm processes a single training example at a time, and
+can be applied while remaining completely in the primal domain, making it
+applicable to very large datasets.  We refer to the algorithm as SZLJSP, after
+the authors of the two papers.
 %
@@
 \end{description}
+The only disadvantage of SZLJSP compared to SSGD is the larger memory requirement.
+While SSGD only needs to store a single copy of the parameters $\theta$, SZLJSP
+needs to store a separate (though possibly sparse) copy for each training sample.
 
-In practice, the SZLJSP algorithm is usually used with the weighted averaging
+In practice, the SZLJSP algorithm is commonly used with the weighted averaging
 described in \Eqref{averaging}, as this is known to improve performance in the
@@
 
-\section{Conditional Random Fields for Semantic Segmentation}\seclabel{semantic_segmentation}
+\section[Conditional Random Fields for Semantic Segmentation]{Conditional Random Fields\linebreak for Semantic Segmentation}\seclabel{semantic_segmentation}
 %
@@
     Schematic of CRFs over superpixels. From left to right: input image, superpixel segmentation
-    using SLIC into about 100 superpixels, pairwise potential in the CRF, and desired labeling.
+    using SLIC into about 100 superpixels, pairwise potential in the CRF, and desired labeling of superpixels.
 \figlabel{superpixel_illustration}}
@@
 % submodular/non-submodular? higher order?
+\subsection{Data, Features and Superpixels}
+When evaluating learning algorithms in Chapter~\ref{ch:comparison} and
+Chapter~\ref{ch:exact_learning} we use the well-established \textsc{Pascal} VOC 2010 and
+MSRC-21 datasets, shown in the introduction in \Figref{pascal_msrc}.
+As we focus on the learning algorithms in these chapters, we use features for
+unary potentials from the literature.
+
+For both datasets we use the TextonBoost class probabilities provided by
+\citet{krahenbuhl2012efficient}. For the \textsc{Pascal} VOC dataset, the potentials
+provided by \citet{krahenbuhl2012efficient} also include the responses of
+object detectors~\citet{}, to better capture the complex object classes.
+We average the potentials inside superpixels and use the resulting feature
+as input to our unary potentials.
+For the MSRC-21 dataset, we compute TextonBoost on two different scales, as
+suggested by \citet{mottaghianalyzing}. We additionally extract SIFT and color descriptors and create
+bag-of-word descriptors for each superpixel. Following the approach of
+\citet{lucchi2011spatial} we augment these with a global bag-of-word descriptor
+for each image. We train a linear SVM using an approximation to the additive
+$\chi^2$ kernel~\citet{vedaldi2010efficient} and use the response as an
+additional input to our CRF\@. This \emph{piecewise training} simplifies
+learning and was found to have little effect on
+accuracy~\citet{nowozin2010parameter}. In total, there are $63$ features
+for the MSRC-21 dataset, $21$ for each scale of TextonBoost and an additional
+$21$ for the bag-of-word model using an SVM.
+
+We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
+our experiments.  It has been shown to provide competitive results with a
+minimum of computational complexity.  Algorithmically, SLIC simply computes a
+$k$-means clustering over pixels. Each pixels is represented as 5D point, using
+three color channels and the $x$ and $y$ coordinates in the image.  In a
+post-processing step, small segments are removed. To make clustering of so many
+points feasible, the search for the nearest cluster in $k$-means is restricted
+to a local neighborhood in the image. An example of superpixels computed with
+the SLIC algorithm can be found in \Figref{superpixel_illustration}.
+%
+The features and superpixel algorithm used for the RGB-D dataset NYU used in
+Chapter~\ref{ch:nyu} will be discussed there.
+
 \subsection{Previous Work}
@@
 optimization problem.
-
-\subsection{Data, Features and Superpixels}
-When evaluating learning algorithms in Chapter~\ref{ch:comparison} and
-Chapter~\ref{ch:exact_learning} we use the well-established \textsc{Pascal} VOC 2010 and
-MSRC-21 datasets, shown in the introduction in \Figref{pascal_msrc}.
-As we focus on the learning algorithms in these chapters, we use features for
-unary potentials from the literature.
-
-For both datasets we use the TextonBoost class probabilities provided by
-\citet{krahenbuhl2012efficient}. For the \textsc{Pascal} VOC dataset, the potentials
-provided by \citet{krahenbuhl2012efficient} also include the responses of
-object detectors~\citet{}, to better capture the complex object classes.
-We average the potentials inside superpixels and use the resulting feature
-as input to our unary potentials.
-For the MSRC-21 dataset, we compute TextonBoost on two different scales, as
-suggested by \citet{mottaghianalyzing}. We additionally extract SIFT and color descriptors and create
-bag-of-word descriptors for each superpixel. Following the approach of
-\citet{lucchi2011spatial} we augment these with a global bag-of-word descriptor
-for each image. We train a linear SVM using an approximation to the additive
-$\chi^2$ kernel~\citet{vedaldi2010efficient} and use the response as an
-additional input to our CRF\@. This \emph{piecewise training} simplifies
-learning and was found to have little effect on
-accuracy~\citet{nowozin2010parameter}. In total, there are $63$ features
-for the MSRC-21 dataset, $21$ for each scale of TextonBoost and an additional
-$21$ for the bag-of-word model using an SVM.
-
-We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
-our experiments.  It has been shown to provide competitive results with a
-minimum of computational complexity.  Algorithmically, SLIC simply computes a
-$k$-means clustering over pixels. Each pixels is represented as 5D point, using
-three color channels and the $x$ and $y$ coordinates in the image.  In a
-post-processing step, small segments are removed. To make clustering of so many
-points feasible, the search for the nearest cluster in $k$-means is restricted
-to a local neighborhood in the image. An example of superpixels computed with
-the SLIC algorithm can be found in \Figref{superpixel_illustration}.
-
-The features and superpixel algorithm used for the RGB-D dataset NYU used in
-Chapter~\ref{ch:nyu} will be discussed there.

### 2013-09-27 main.tex
set binding correction to 2cm (as this is the total correction).
more minor fixes

@@
-\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=1cm]{scrbook}
+\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=2cm]{scrbook}
 %\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=calc]{scrbook}
@@
 prediction, which is sometimes called ``inference machines''. The basic
-principle is simple: the process of prediction \emph{using a given inference
+principle is simple: the process of \emph{prediction using a given inference
 procedure} is viewed as a feed-forward method for prediction, and parameters of this
@@
 conditional neural fields~\citep{peng2009conditional} based on neural networks,
-and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
+and decision tree fields~(DTFs)~\citep{nowozin2011decision}, based on decision trees.
 Conditional neural fields have only been applied to sequence classification so far,
 and extending them to our setting of semantic image segmentation would be very
-interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
+interesting. DTFs on the other hand have been applied to loopy
 graphs for image processing, but not for higher-level tasks such as semantic
@@
 While non-linear potentials would allow for more complex interactions between
-inputs and labelings, introducing higher order potentials or latent variables
-allow the model to express more complex interactions within the output
-variables~\citep{dann2012pottics, kohli2009robust}. Possible examples are
-consistency of larger regions, learning parts or learning of scene classes and
-cooccurences.
+inputs and labelings, introducing higher order
+potentials~\citep{kohli2009robust, ladicky2009associative} or latent
+variables~\citep{dann2012pottics} allow the model to express more complex
+interactions within the output variables. Possible examples are consistency of
+larger regions, learning parts or learning of scene classes and cooccurences.
 In principle, higher order potentials and latent variable models are
@@
 higher order potentials for semantic segmentation has received little
-attention, while approaches using latent variables are often limited by
-the non-convexity of learning.  It would be interesting to compare current
-methods using latent variable and higher order approaches, and see how these
-interact with different inference and learning schemes.
+attention, while approaches using latent variables are often limited by the
+non-convexity of learning.  It would be interesting to compare current methods
+using latent variable and higher order approaches, and see how these interact
+with different inference and learning schemes.
 %
